### 2019-11-19 data_cleaning_initial.Rmd
initial commit

---
title: "KidsAIR initial data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
```

```{r}

```

### 2019-11-21 data_cleaning_initial.Rmd
pm data cleaning

@@
 ```{r}
+# Load Dusttrack data
+## List sheets within the excel file
+list_sheets_nm <- excel_sheets("Input/nm_kids_dusttrack_20191121.xlsx")
+## Apply the read_xlsx function to each item in the sheet list
+## Bind the items from each sheet into a single dataframe
+pm_initial_nm <- data.frame(do.call("rbind", lapply(list_sheets_nm, read_xlsx, 
+                                      path = "Input/nm_kids_dusttrack_20191121.xlsx")))
 
+list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
+pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
+                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx")))
+
+# Bind data together from different study areas
+pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
+
+# Rename and create new variables
+kids_pm <- pm_initial %>% 
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  mutate(id_home = as.factor(homeid),
+         date = ymd(date),
+         time = hms(time),
+         pm25 = as.numeric(aerosol)) %>% 
+  unite(datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime = ymd_hms(datetime)) %>% 
+  select(id_home, date, time, datetime, pm25) %>% 
+  arrange(id_home, date) %>% 
+  group_by(id_home) %>% 
+  mutate(obs = n(),
+         total = sum(pm25),
+         min_datetime = as.character(datetime),
+         min_datetime = min(min_datetime)),
+         max_datetime = as.character(datetime),
+         max_datetime = max(max_datetime)),
+         diff_datetime = ymd_hms(max_datetime) - ymd_hms(min_datetime))
+
+test <- kids_pm %>% 
+  group_by(id_home) %>% 
+  summarize(n())
 ```

### 2019-11-22 data_cleaning_initial.Rmd
data cleaning updates

--- 2019-11-19 data_cleaning_initial.Rmd
@@
 library(lubridate)
+library(zoo)
 ```
 
+# Load, format, and save KidsAIR Dusttrack data
 ```{r}
+# Load Dusttrack data
+## List sheets within the excel file
+list_sheets_nm <- excel_sheets("Input/nm_kids_dusttrack_20191121.xlsx")
+## Apply the read_xlsx function to each item in the sheet list
+## Bind the items from each sheet into a single dataframe
+pm_initial_nm <- data.frame(do.call("rbind", lapply(list_sheets_nm, read_xlsx, 
+                                      path = "Input/nm_kids_dusttrack_20191121.xlsx"))) %>% 
+  mutate(area = "NM")
 
+list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
+pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
+                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
+  mutate(area = "WMT")
+
+# Bind data together from different study areas
+pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
+
+# Rename and create new variables
+kids_pm <- pm_initial %>% 
+  # separate time from incorrect/associated date
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  # format/rename variables
+  mutate(id_home = as.factor(homeid),
+         pm_date = ymd(date),
+         pm_time = hms(time),
+         pm = as.numeric(aerosol)) %>% 
+  # new datetime variable
+  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
+  select(id_home, area, pm_date, pm_time, pm_datetime, pm) %>% 
+  arrange(id_home, pm_date) %>% 
+  group_by(id_home) %>% 
+  # new variables summarizing pm2.5 within each id_home
+  mutate(pm_total_observations = n(),
+         pm_total_summed = sum(pm),
+         min_datetime = first(pm_datetime),
+         max_datetime = last(pm_datetime),
+         min_datetime = ymd_hms(min_datetime),
+         max_datetime = ymd_hms(max_datetime),
+         diff_datetime = interval(min_datetime, max_datetime),
+         pm_sampling_days = diff_datetime/86400,
+         pm_24hr_mean = pm_total_summed/pm_sampling_days,
+         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right")) %>% 
+  select(-min_datetime, -max_datetime, -diff_datetime) %>% 
+  ungroup()
+
+# Save pm dataset as RDS and CSV
+write_rds(kids_pm, "Output/kids_pm.rds")
+write_csv(kids_pm, "Output/kids_pm.csv")
 ```
+
+# Load, format, and save fuel moisture content data
+```{r}
+moisture_initial_wmt <- read_xlsx("Input/wmt_kids_moisture_20191122.xlsx")
+
+kids_moisture <- moisture_initial_wmt %>% 
+  mutate(id = as.factor(ID), # ask about what these IDs are 
+         id_home = as.factor(HomeWinterID),
+         moisture_date = ymd(CompletedDate),
+         moisture_1 = as.numeric(Moisture1),
+         moisture_2 = as.numeric(Moisture2),
+         moisture_3 = as.numeric(Moisture3),
+         moisture_ave = as.numeric(MoistureAve),
+         moisture_missing = as.character(WhatMissing),
+         moisture_incomplete = as.factor(Incomplete),
+         moisture_ignore_missing = as.factor(IgnoreMissingData),
+         moisture_ignore_reason = as.character(IgnoreReason),
+         moisture_tech = as.character(Technician),
+         moisture_split = as.factor(Split),
+         area = "WMT") %>% 
+  select(id, id_home, area, moisture_date, moisture_1, moisture_2, moisture_3, 
+         moisture_ave, moisture_missing, moisture_incomplete, moisture_ignore_missing,
+         moisture_ignore_reason, moisture_tech, moisture_split)
+
+# Save pm dataset as RDS and CSV
+write_rds(kids_moisture, "Output/kids_moisture.rds")
+write_csv(kids_moisture, "Output/kids_moisture.csv")
+```

### 2019-11-22 moisture_summary.Rmd
data cleaning updates

---
title: "Kids Moisture Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

```{r}
# Load data
kids_moisture <- read_rds("Output/kids_moisture.rds")
```

# Summary stats
```{r}
summary_kids_moisture <- kids_moisture %>% 
  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
            min(pm_24hr_mean), max(pm_24hr_mean),
            min(pm), max(pm),
            min(pm_sampling_days), max(pm_sampling_days),
            min(pm_total_observations), max(pm_total_observations)) %>% 
  t()
summary_kids_pm
```

# Boxplots
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- kids_pm %>% 
  ggplot(aes(area, pm_24hr_mean)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(0, 400) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

### 2019-11-22 pm_summary.Rmd
data cleaning updates

--- 2019-11-22 moisture_summary.Rmd
@@
 ---
-title: "Kids Moisture Summary"
+title: "Kids PM Summary"
 author: "Ethan Walker"
@@
 # Load data
-kids_moisture <- read_rds("Output/kids_moisture.rds")
+kids_pm <- read_rds("Output/kids_pm.rds")
 ```
@@
 ```{r}
-summary_kids_moisture <- kids_moisture %>% 
+summary_kids_pm <- kids_pm %>% 
   summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),

### 2019-11-25 data_cleaning_initial.Rmd
data cleaning and summary updates

--- 2019-11-19 data_cleaning_initial.Rmd
@@
 library(lubridate)
+library(zoo)
 ```
 
+# Load, format, and save KidsAIR Dusttrack data
 ```{r}
+# Load Dusttrack data
+## List sheets within the excel file
+list_sheets_nm <- excel_sheets("Input/nm_kids_dusttrack_20191121.xlsx")
+## Apply the read_xlsx function to each item in the sheet list
+## Bind the items from each sheet into a single dataframe
+pm_initial_nm <- data.frame(do.call("rbind", lapply(list_sheets_nm, read_xlsx, 
+                                      path = "Input/nm_kids_dusttrack_20191121.xlsx"))) %>% 
+  mutate(area = "NM")
 
+list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
+pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
+                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
+  mutate(area = "WMT")
+
+# Bind data together from different study areas
+pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
+
+# Rename and create new variables
+kids_pm <- pm_initial %>% 
+  # separate time from incorrect/associated date
+  separate(time, into = c("trash", "time"), sep = " ") %>% 
+  # format/rename variables
+  mutate(id_home = as.factor(homeid),
+         pm_date = ymd(date),
+         pm_time = hms(time),
+         pm = as.numeric(aerosol)) %>% 
+  # new datetime variable
+  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
+  select(id_home, area, pm_date, pm_time, pm_datetime, pm) %>% 
+  arrange(id_home, pm_date) %>% 
+  group_by(id_home) %>% 
+  # new variables summarizing pm2.5 within each id_home
+  mutate(pm_total_observations = n(),
+         pm_total_summed = sum(pm),
+         min_datetime = first(pm_datetime),
+         max_datetime = last(pm_datetime),
+         min_datetime = ymd_hms(min_datetime),
+         max_datetime = ymd_hms(max_datetime),
+         diff_datetime = interval(min_datetime, max_datetime),
+         pm_total_days = diff_datetime/86400,
+         pm_24hr_mean = pm_total_summed/pm_total_days,
+         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
+         obs = row_number(),
+         sampling_day = if_else(obs < 1441, 1, 0),
+         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
+         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
+         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
+         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
+         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
+         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
+         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
+  group_by(id_home, sampling_day) %>% 
+  mutate(pm_24hr_daily = sum(pm),
+         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
+         pm_sampling_day = sampling_day) %>% 
+  ungroup() %>% 
+  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
+
+# Save pm dataset as RDS and CSV
+write_rds(kids_pm, "Output/kids_pm.rds")
+write_csv(kids_pm, "Output/kids_pm.csv")
 ```
+
+# Load, format, and save fuel moisture content data
+```{r}
+moisture_initial_wmt <- read_xlsx("Input/wmt_kids_moisture_20191122.xlsx")
+
+kids_moisture <- moisture_initial_wmt %>% 
+  mutate(id = as.factor(ID), # ask about what these IDs are 
+         id_home = as.factor(HomeWinterID),
+         moisture_date = ymd(CompletedDate),
+         moisture_1 = as.numeric(Moisture1),
+         moisture_2 = as.numeric(Moisture2),
+         moisture_3 = as.numeric(Moisture3),
+         moisture_ave = as.numeric(MoistureAve),
+         moisture_missing = as.character(WhatMissing),
+         moisture_incomplete = as.factor(Incomplete),
+         moisture_ignore_missing = as.factor(IgnoreMissingData),
+         moisture_ignore_reason = as.character(IgnoreReason),
+         moisture_tech = as.character(Technician),
+         moisture_split = as.factor(Split),
+         area = "WMT") %>% 
+  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
+                                 moisture_3 = 0, moisture_ave = 0)) %>% 
+  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
+                                 moisture_3 = -999)) %>% 
+  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
+  group_by(id_home, moisture_date) %>% 
+  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  spread(sample, value) %>% 
+  select(id, id_home, area, moisture_date, moisture_1, moisture_2, moisture_3, 
+         moisture_ave, moisture_missing, moisture_incomplete, moisture_ignore_missing,
+         moisture_ignore_reason, moisture_tech, moisture_split)
+
+# Save pm dataset as RDS and CSV
+write_rds(kids_moisture, "Output/kids_moisture.rds")
+write_csv(kids_moisture, "Output/kids_moisture.csv")
+```

### 2019-11-25 moisture_summary.Rmd
data cleaning and summary updates

@@
 summary_kids_moisture <- kids_moisture %>% 
-  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
-            min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm),
-            min(pm_sampling_days), max(pm_sampling_days),
-            min(pm_total_observations), max(pm_total_observations)) %>% 
+  summarize(mean(moisture_ave, na.rm = TRUE), sd(moisture_ave, na.rm = TRUE),
+            median(moisture_ave, na.rm = TRUE),
+            min(moisture_ave, na.rm = TRUE), max(moisture_ave, na.rm = TRUE)) %>% 
   t()
-summary_kids_pm
+summary_kids_moisture
 ```
@@
 ```{r, fig.height=6, fig.width=5}
-boxplots_kids_pm <- kids_pm %>% 
-  ggplot(aes(area, pm_24hr_mean)) +
+boxplots_kids_moisture <- kids_moisture %>% 
+  select(area, moisture_ave) %>% 
+  na.exclude() %>% 
+  ggplot(aes(area, moisture_ave)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
     theme_minimal() +
-    ylim(0, 400) +
+    ylim(0, 50) +
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +
-    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    labs(y = "Mean moisture content (%)") +
     #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
@@
           axis.ticks = element_blank())
-boxplots_kids_pm
+boxplots_kids_moisture
 ```

### 2019-11-25 pm_summary.Rmd
data cleaning and summary updates

@@
             min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm),
-            min(pm_sampling_days), max(pm_sampling_days),
-            min(pm_total_observations), max(pm_total_observations)) %>% 
+            min(pm), max(pm), mean(pm_total_days),
+            min(pm_total_days), max(pm_total_days),
+            min(pm_total_observations), max(pm_total_observations),
+            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
   t()

### 2019-11-26 data_cleaning_initial.Rmd
further data processing/cleaning

--- 2019-11-22 data_cleaning_initial.Rmd
@@
   # format/rename variables
-  mutate(id_home = as.factor(homeid),
+  mutate(home_winter_id = as.factor(homeid),
          pm_date = ymd(date),
@@
   unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(id_home, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(id_home, pm_date) %>% 
-  group_by(id_home) %>% 
-  # new variables summarizing pm2.5 within each id_home
+  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
+  arrange(home_winter_id, pm_date) %>% 
+  group_by(home_winter_id) %>% 
+  # new variables summarizing pm2.5 within each home_winter_id
   mutate(pm_total_observations = n(),
@@
          diff_datetime = interval(min_datetime, max_datetime),
-         pm_sampling_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_sampling_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right")) %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime) %>% 
-  ungroup()
+         pm_total_days = diff_datetime/86400,
+         pm_24hr_mean = pm_total_summed/pm_total_days,
+         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
+         obs = row_number(),
+         sampling_day = if_else(obs < 1441, 1, 0),
+         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
+         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
+         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
+         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
+         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
+         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
+         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
+  group_by(home_winter_id, sampling_day) %>% 
+  mutate(pm_24hr_daily = sum(pm),
+         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
+         pm_sampling_day = sampling_day) %>% 
+  ungroup() %>% 
+  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
 
@@
 ```{r}
+# Load data
 moisture_initial_wmt <- read_xlsx("Input/wmt_kids_moisture_20191122.xlsx")
@@
 kids_moisture <- moisture_initial_wmt %>% 
-  mutate(id = as.factor(ID), # ask about what these IDs are 
-         id_home = as.factor(HomeWinterID),
+  # rename variables
+  mutate(moisture_id = as.factor(ID),  
+         home_winter_id = as.factor(HomeWinterID),
          moisture_date = ymd(CompletedDate),
@@
          area = "WMT") %>% 
-  select(id, id_home, area, moisture_date, moisture_1, moisture_2, moisture_3, 
+  # replace 0 and -999 values with NA
+  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
+                                 moisture_3 = 0, moisture_ave = 0)) %>% 
+  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
+                                 moisture_3 = -999)) %>% 
+  # gather 3 moisture samples to calculate average
+  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
+  group_by(home_winter_id, moisture_date) %>% 
+  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  # spread 3 moisture samples back to wide format
+  spread(sample, value) %>% 
+  select(moisture_id, home_winter_id, area, moisture_date, moisture_1, moisture_2, moisture_3, 
          moisture_ave, moisture_missing, moisture_incomplete, moisture_ignore_missing,
@@
 
-# Save pm dataset as RDS and CSV
+# Save as RDS and CSV
 write_rds(kids_moisture, "Output/kids_moisture.rds")
@@
 ```
+
+# Load, format, and save home and child overview data
+## These files will link the home ID and child ID with the intervention status
+```{r}
+# Load files
+wmt_child_initial <- read_xlsx("Input/wmt_child_20191126.xlsx")
+wmt_home_initial <- read_xlsx("Input/wmt_home_20191126.xlsx")
+wmt_homewinter_initial <- read_xlsx("Input/wmt_homewinter_20191126.xlsx")
+wmt_parent_demographics_initial <- read_xlsx("Input/wmt_parent_demographics_20191126.xlsx")
+
+# join all files from above
+joined1 <- wmt_child_initial %>% 
+  mutate(hispanic_child = as.factor(Hispanic),
+         race_child = as.factor(Race),
+         comments_child = Comments,
+         what_missing_child = WhatMissing,
+         incomplete_child = Incomplete,
+         ignore_missing_child = IgnoreMissingData,
+         ignore_reason_child = IgnoreReason) %>% 
+  select(-ID, -Hispanic, -Race, -Comments, -WhatMissing, -Incomplete,
+         -IgnoreMissingData, -IgnoreReason) %>% 
+  left_join(wmt_homewinter_initial, by = "HomeID") %>% 
+  left_join(wmt_home_initial, by = "HomeID") %>% 
+  left_join(wmt_parent_demographics_initial, by = "HomeWinterID") %>%
+  mutate(gender_parent = as.factor(Gender),
+         hispanic_parent = as.factor(Hispanic),
+         race_parent = as.factor(Race),
+         what_missing_parent = WhatMissing,
+         incomplete_parent = Incomplete,
+         ignore_missing_parent = IgnoreMissingData,
+         ignore_reason_parent = IgnoreReason) %>%
+  select(-ID, -Hispanic, -Race, -WhatMissing, -Incomplete,
+         -IgnoreMissingData, -IgnoreReason, -Gender, -EnterDate) %>%
+  rename_all(tolower)
+
+# create dataset of id's and treatment assignment
+wmt_ids_linked <- joined1 %>% 
+  select(childid, homeid, home, homewinterid, winterid, txcondition) %>% 
+  mutate(child_id = as.character(childid),
+         home_id = as.character(homeid),
+         home_winter_id = as.character(homewinterid),
+         winter_id = as.character(winterid),
+         treatment = as.factor(txcondition)) %>% 
+  select(-childid, -homeid, -homewinterid, -winterid, -txcondition)
+
+# create dataset of demographic data
+wmt_demographic_data <- joined1 %>% 
+  mutate(child_id = as.character(childid),
+         home_id = as.character(homeid),
+         age_nov1 = as.factor(agenov1),
+         child_assent_date = ymd(childassentdate),
+         parent_consent_date = ymd(parentconsentdate),
+         home_winter_id = as.character(homewinterid),
+         total_residents = as.numeric(totalres),
+         total_residents_under5 = as.numeric(underfive),
+         ignore_blanks_parent = ignoreblanks,
+         income = as.factor(income),
+         education = as.factor(education)) %>% 
+  select(child_id, home_id, hispanic_child, age_nov1, race_child, child_assent_date,
+         parent_consent_date, comments_child, what_missing_child, incomplete_child,
+         enrolled, ignore_missing_child, ignore_reason_child,
+         home_winter_id, gender_parent, hispanic_parent, race_parent, income,
+         education, total_residents, total_residents_under5, incomplete_parent,
+         ignore_blanks_parent, what_missing_parent, ignore_missing_parent,
+         ignore_reason_parent) %>% 
+  distinct(child_id, .keep_all = TRUE)
+
+# create dataset of data on home details
+## lower on priority list - see which variables we will use and clean later
+wmt_home_details_uncleaned <- wmt_homewinter_initial %>%
+  left_join(wmt_home_initial, by = "HomeID") %>% 
+  rename_all(tolower)
+
+# Save as RDS and CSV
+write_rds(wmt_ids_linked, "Output/wmt_ids_linked.rds")
+write_rds(wmt_demographic_data, "Output/wmt_demographic_data.rds")
+write_csv(wmt_ids_linked, "Output/wmt_ids_linked.csv")
+write_csv(wmt_demographic_data, "Output/wmt_demographic_data.csv")
+```

### 2019-12-04 data_cleaning_initial.Rmd
additional dusttrack data cleaning

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019"
+date: "Started 19 Nov 2019, Updated 4 Dec 2019"
 output: pdf_document
@@
 ```{r}
-# Load Dusttrack data
+##### Read in dusttrack .txt files (from WMT location) #####
+
+# List files
+list_txt_files <- list.files("Input/WMT/dusttrack_txt_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/dusttrack_txt_files")
+initial_data = tibble(file_list = list_txt_files) %>%
+  extract(file_list, "home_winter_id", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 28)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other WMT/PM data below
+wmt_pm_txt <- initial_data %>% 
+  mutate(date = mdy(`MM/dd/yyyy`),
+         time = hms(`hh:mm:ss`),
+         pm = as.numeric(`mg/m^3`),
+         area = "WMT") %>% 
+  select(home_winter_id, date, time, pm, area)
+
+##### Next, load other files from all locations #####
+
+##### Load combined dusttrack files from Emily W #####
+##### Compare these with the raw data pulled in above #####
+
 ## List sheets within the excel file
-list_sheets_nm <- excel_sheets("Input/nm_kids_dusttrack_20191121.xlsx")
+list_sheets_nn <- excel_sheets("Input/nn_kids_dusttrack_20191121.xlsx")
 ## Apply the read_xlsx function to each item in the sheet list
 ## Bind the items from each sheet into a single dataframe
-pm_initial_nm <- data.frame(do.call("rbind", lapply(list_sheets_nm, read_xlsx, 
-                                      path = "Input/nm_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "NM")
+pm_initial_nn <- data.frame(do.call("rbind", lapply(list_sheets_nn, read_xlsx, 
+                                      path = "Input/nn_kids_dusttrack_20191121.xlsx"))) %>% 
+  mutate(area = "NN")
 
@@
 
-# Load, format, and save home and child overview data
-## These files will link the home ID and child ID with the intervention status
+# Load, format, and save home and child overview/demographic data
 ```{r}
@@
 ```
+
+```{r}
+
+```

### 2019-12-04 pm_summary.Rmd
additional dusttrack data cleaning

--- 2019-11-22 pm_summary.Rmd
@@
 summary_kids_pm <- kids_pm %>% 
+  #filter(pm < 0)
   summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
             min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm),
-            min(pm_sampling_days), max(pm_sampling_days),
-            min(pm_total_observations), max(pm_total_observations)) %>% 
+            min(pm), max(pm), mean(pm_total_days),
+            min(pm_total_days), max(pm_total_days),
+            min(pm_total_observations), max(pm_total_observations),
+            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
   t()

### 2019-12-05 data_cleaning_initial.Rmd
PM data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 4 Dec 2019"
+date: "Started 19 Nov 2019, Updated 5 Dec 2019"
 output: pdf_document
@@
 ```{r}
-##### Read in dusttrack .txt files (from WMT location) #####
+##### Read in dusttrack .txt files (from WMT) #####
 
@@
 
-# Format combined file from above; combine with other WMT/PM data below
+# Format combined file from above; combine with other PM data below
 wmt_pm_txt <- initial_data %>% 
   mutate(date = mdy(`MM/dd/yyyy`),
-         time = hms(`hh:mm:ss`),
+         time = `hh:mm:ss`,
          pm = as.numeric(`mg/m^3`),
          area = "WMT") %>% 
-  select(home_winter_id, date, time, pm, area)
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  select(home_winter_id, datetime, date, time, pm, area)
 
-##### Next, load other files from all locations #####
+##### Read in dusttrack .csv files (all areas) #####
 
+### WMT ###
+# List files
+list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/dusttrack_csv_files")
+initial_data = tibble(file_list = list_csv_files) %>%
+  extract(file_list, "home_winter_id", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_pm_csv <- initial_data %>% 
+  mutate(date = mdy(Date),
+         time = Time,
+         pm = as.numeric(AEROSOL),
+         area = "WMT") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  select(home_winter_id, datetime, date, time, pm, area) %>% 
+  filter(!is.na(pm))
+
+### NN ###
+# List files
+list_csv_files <- list.files("Input/NN/dusttrack_csv_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/NN/dusttrack_csv_files")
+initial_data = tibble(file_list = list_csv_files) %>%
+  extract(file_list, "home_winter_id", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+nn_pm_csv <- initial_data %>% 
+  mutate(date = mdy(Date),
+         time = Time,
+         pm = as.numeric(AEROSOL),
+         area = "NN") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  select(home_winter_id, datetime, date, time, pm, area) %>% 
+  filter(!is.na(pm))
+
+### AK ###
+# List files
+list_csv_files <- list.files("Input/AK/dusttrack_csv_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/AK/dusttrack_csv_files")
+initial_data = tibble(file_list = list_csv_files) %>%
+  extract(file_list, "home_winter_id", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+ak_pm_csv <- initial_data %>% 
+  mutate(date = mdy(Date),
+         time = Time,
+         pm = as.numeric(AEROSOL),
+         area = "AK") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  select(home_winter_id, datetime, date, time, pm, area) %>% 
+  filter(!is.na(pm))
+
+##### Read in other dusttrack .csv files (WMT and NN) #####
+
+### WMT ### 
+# List files
+list_other_files <- list.files("Input/WMT/dusttrack_other_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/dusttrack_other_files")
+initial_data = tibble(file_list = list_other_files) %>%
+  extract(file_list, "home_winter_id", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_pm_other <- initial_data %>% 
+  mutate(trash1 = `Instrument Name`,
+         trash2 = `DustTrak II`) %>% 
+  select(home_winter_id, trash1, trash2) %>% 
+  group_by(home_winter_id) %>% 
+  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
+                              trash2, "NA"),
+         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
+                              trash1, "NA"),
+         trash1 = as.numeric(trash1),
+         trash1 = as.character(trash1),
+         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
+  filter(!is.na(trash5)) %>% 
+  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
+  spread(trash4, trash3) %>% 
+  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
+  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
+  mutate(datetime = lead(datetime)) %>% 
+  replace_with_na(replace = list(datetime = "NA NA")) %>% 
+  mutate(datetime2 = first(datetime),
+         datetime = mdy_hms(datetime2),
+         trash1 = as.numeric(trash1)) %>% 
+  filter(!is.na(trash1)) %>% 
+  mutate(datetime2 = datetime + seconds(trash1),
+         datetime2 = ymd_hms(datetime2)) %>% 
+  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(pm = as.numeric(trash2),
+         datetime = datetime2,
+         area = "WMT") %>% 
+  select(home_winter_id, datetime, date, time, pm, area) %>% 
+  ungroup()
+
+### NN ### 
+# List files
+list_other_files <- list.files("Input/NN/dusttrack_other_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/NN/dusttrack_other_files")
+initial_data = tibble(file_list = list_other_files) %>%
+  extract(file_list, "home_winter_id", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+nn_pm_other <- initial_data %>% 
+  mutate(trash1 = `Instrument Name`,
+         trash2 = `DustTrak II`) %>% 
+  select(home_winter_id, trash1, trash2) %>% 
+  filter(!is.na(trash1)) %>% 
+  group_by(home_winter_id) %>% 
+  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
+                              trash1, "NA"),
+         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
+                              trash2, "NA"),
+         trash1 = as.numeric(trash1),
+         trash1 = as.character(trash1),
+         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
+  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
+  filter(!is.na(trash5)) %>% 
+  spread(trash3, trash4) %>% 
+  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
+  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
+  mutate(datetime = lead(datetime)) %>% 
+  replace_with_na(replace = list(datetime = "NA NA")) %>% 
+  mutate(datetime2 = first(datetime),
+         datetime = mdy_hms(datetime2),
+         trash1 = as.numeric(trash1)) %>% 
+  filter(!is.na(trash1)) %>% 
+  mutate(datetime2 = datetime + seconds(trash1),
+         datetime2 = ymd_hms(datetime2)) %>% 
+  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(pm = as.numeric(trash2),
+         datetime = datetime2,
+         area = "NN") %>% 
+  select(home_winter_id, datetime, date, time, pm, area) %>% 
+  ungroup() %>% 
+  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
+                                  "22", home_winter_id),
+         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
+                                  "55", home_winter_id),
+         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
+                                  "81", home_winter_id))
+
+##### Bind all files from above #####
+kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
+                nn_pm_csv, nn_pm_other,
+                ak_pm_csv)
+
+# Save pm dataset as RDS and CSV
+write_rds(kids_pm, "Output/kids_pm_full.rds")
+#write_csv(kids_pm, "Output/kids_pm_full.csv")
+
+# Load data and clean up
+kids_pm <- read_rds("Output/kids_pm_full.rds")
+
+kids_pm_check <- kids_pm %>% 
+  group_by(home_winter_id) %>% 
+  summarise(obs = n()) %>% 
+  ungroup()
+
+ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
+nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
+wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")
+
+dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
+  rename_all(tolower) %>% 
+  filter(!is.na(homewinterid)) %>% 
+  mutate(home_winter_id = as.character(homewinterid)) %>% 
+  left_join(kids_pm_check, by = "home_winter_id")
+
+##### Add to new variables to Kids PM data #####
+kids_pm_new <- kids_pm %>% 
+  mutate(home_winter_id = as.character(home_winter_id),
+         area = as.factor(area),
+         pm_date = date,
+         pm_time = time,
+         pm_datetime = datetime,
+         pm = as.numeric(pm)) %>% 
+  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
+  arrange(home_winter_id, pm_date) %>% 
+  group_by(home_winter_id) %>% 
+  # new variables summarizing pm2.5 within each home_winter_id
+  mutate(pm_total_observations = n(),
+         pm_total_summed = sum(pm),
+         min_datetime = first(pm_datetime),
+         max_datetime = last(pm_datetime),
+         min_datetime = ymd_hms(min_datetime),
+         max_datetime = ymd_hms(max_datetime),
+         diff_datetime = interval(min_datetime, max_datetime),
+         pm_total_days = diff_datetime/86400,
+         pm_24hr_mean = pm_total_summed/pm_total_days,
+         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
+         obs = row_number(),
+         sampling_day = if_else(obs < 1441, 1, 0),
+         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
+         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
+         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
+         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
+         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
+         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
+         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
+  group_by(home_winter_id, sampling_day) %>% 
+  mutate(pm_24hr_daily = sum(pm),
+         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
+         pm_sampling_day = sampling_day) %>% 
+  ungroup() %>% 
+  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
+
+# Save pm dataset as RDS and CSV
+write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+#write_csv(kids_pm_new, "Output/kids_pm_new.csv")
+```
+
+# Original code/files from Emily - save for future reference
+```{r}
 ##### Load combined dusttrack files from Emily W #####

### 2019-12-05 pm_summary.Rmd
PM data cleaning updates

--- 2019-11-22 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019"
+date: "Started 22 Nov 2019, Updated 5 Dec 2019"
 output: pdf_document
@@
 # Load data
-kids_pm <- read_rds("Output/kids_pm.rds")
+kids_pm <- read_rds("Output/kids_pm_new.rds")
 ```
@@
 summary_kids_pm <- kids_pm %>% 
+  na.exclude() %>% 
+  #filter(pm < 0)
   summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
             min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm),
-            min(pm_sampling_days), max(pm_sampling_days),
-            min(pm_total_observations), max(pm_total_observations)) %>% 
+            min(pm), max(pm), mean(pm_total_days),
+            min(pm_total_days), max(pm_total_days),
+            min(pm_total_observations), max(pm_total_observations),
+            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
   t()

### 2019-12-06 data_cleaning_initial.Rmd
dusttrak data cleaning

@@
 
-# Load, format, and save KidsAIR Dusttrack data
+# Load, format, and save KidsAIR Dusttrak data
 ```{r}
@@
 #write_csv(kids_pm, "Output/kids_pm_full.csv")
+```
 
-# Load data and clean up
-kids_pm <- read_rds("Output/kids_pm_full.rds")
+# Load Dusttrak data and clean up
+```{r}
+kids_pm_full <- read_rds("Output/kids_pm_full.rds")
 
-kids_pm_check <- kids_pm %>% 
-  group_by(home_winter_id) %>% 
+kids_pm_check <- kids_pm_full %>% 
+  group_by(area, home_winter_id) %>% 
   summarise(obs = n()) %>% 
@@
 
-ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
-nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
-wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")
+ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
+  mutate(area = "AK")
+nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
+  mutate(area = "NN")
+wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
+  mutate(area = "WMT")
 
@@
   mutate(home_winter_id = as.character(homewinterid)) %>% 
-  left_join(kids_pm_check, by = "home_winter_id")
+  left_join(kids_pm_check, by = c("area", "home_winter_id"))
+
+%>% 
+  group_by(home_winter_id) %>% 
+  mutate(n_hwid = n()) %>% 
+  ungroup() %>% 
+  filter(n_hwid == 2)
 
 ##### Add to new variables to Kids PM data #####
-kids_pm_new <- kids_pm %>% 
+kids_pm_new <- kids_pm_full %>%  
+  #filter(area == "AK") %>% 
+  # apply correction factor to pm
+  mutate(pm = pm/1.65) %>% 
   mutate(home_winter_id = as.character(home_winter_id),
@@
          pm_time = time,
-         pm_datetime = datetime,
-         pm = as.numeric(pm)) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
+         pm = as.numeric(pm),
+         day_of_week = weekdays(date)) %>% 
+  unite("pm_datetime", date:time, sep = " ") %>% 
+  select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
+  arrange(area, home_winter_id, pm_date) %>% 
+  group_by(area, home_winter_id) %>% 
   # new variables summarizing pm2.5 within each home_winter_id
@@
          pm_total_summed = sum(pm),
+         # new datetime values to calculate sampling interval
          min_datetime = first(pm_datetime),
@@
          diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
+         # total sampling interval, based on first and last datetimes
+         pm_sample_interval = diff_datetime/86400,
+         # expected number of observations with the sampling interval
+         expected_obs = pm_sample_interval*24*60,
+         # difference between actual and expected observations
+         actual_vs_expected_diff = pm_total_observations - expected_obs,
+         # indicator variable if actual vs expected obs > 10 or greater
+         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
+                                      actual_vs_expected_diff > -10, 1, 0),
+         # mean pm based on 24hr periods in the sampling interval
+         pm_24hr_mean_interval = pm_total_summed/pm_sample_interval,
+         # mean pm based on 24hr periods of total observations
+         pm_24hr_mean_obs = pm_total_summed/(pm_total_observations/1440),
+         # rolling 24hr mean pm
          pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
+         # breaking sampling period into days
          obs = row_number(),
@@
          sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
+  # calculate pm means for each sampling day
   group_by(home_winter_id, sampling_day) %>% 

### 2019-12-06 pm_summary.Rmd
dusttrak data cleaning

@@
 ```{r}
+summary <- kids_pm %>% 
+  filter(pm < 0)
+
 summary_kids_pm <- kids_pm %>%
-  na.exclude() %>% 
   #filter(pm < 0)
-  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
-            min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm), mean(pm_total_days),
-            min(pm_total_days), max(pm_total_days),
-            min(pm_total_observations), max(pm_total_observations),
-            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
+  summarize(mean(pm_24hr_mean_interval, na.rm = TRUE), 
+            sd(pm_24hr_mean_interval, na.rm = TRUE),
+            min(pm_24hr_mean_interval, na.rm = TRUE), 
+            max(pm_24hr_mean_interval, na.rm = TRUE),
+            mean(pm_24hr_mean_obs, na.rm = TRUE), 
+            sd(pm_24hr_mean_obs, na.rm = TRUE),
+            min(pm_24hr_mean_obs, na.rm = TRUE), 
+            max(pm_24hr_mean_obs, na.rm = TRUE),
+            min(pm, na.rm = TRUE), 
+            max(pm, na.rm = TRUE), 
+            mean(pm_sample_interval, na.rm = TRUE),
+            min(pm_sample_interval, na.rm = TRUE), 
+            max(pm_sample_interval, na.rm = TRUE),
+            min(pm_total_observations, na.rm = TRUE), 
+            max(pm_total_observations, na.rm = TRUE),
+            mean(pm_24hr_daily, na.rm = TRUE), 
+            mean(pm_24hr_daily_weighted, na.rm = TRUE)) %>% 
   t()
-summary_kids_pm
+kable(summary_kids_pm, digits = 2)
 ```

### 2019-12-09 data_cleaning_initial.Rmd
dusttrak cleaning and summary stats

--- 2019-12-05 data_cleaning_initial.Rmd
@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 5 Dec 2019"
+date: "Started 19 Nov 2019, Updated 9 Dec 2019"
 output: pdf_document
@@
 # Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm_full.rds")
+write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
 #write_csv(kids_pm, "Output/kids_pm_full.csv")
+```
 
-# Load data and clean up
-kids_pm <- read_rds("Output/kids_pm_full.rds")
+# Load Dusttrak data and clean up
+```{r}
+# Load full, raw dataset compiled in chunk above
+kids_pm_full <- read_rds("Output/kids_pm_full_raw.rds")
 
-kids_pm_check <- kids_pm %>% 
-  group_by(home_winter_id) %>% 
-  summarise(obs = n()) %>% 
-  ungroup()
+# Load and combine data sheets with QAQC comments, etc
+ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
+  mutate(area = "AK")
+ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
+  mutate(area = "AK") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, dtid, area)
+nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
+  mutate(area = "NN")
+nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
+  mutate(area = "NN") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, dtid, area)
+wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
+  mutate(area = "WMT")
+wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
+  mutate(area = "WMT") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, dtid, area)
 
-ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
-nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
-wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")
+dt <- rbind(ak_dt, nn_dt, wmt_dt) %>% 
+  mutate(home_winter_id = as.character(homewinterid)) %>% 
+  select(-homewinterid)
 
@@
   mutate(home_winter_id = as.character(homewinterid)) %>% 
-  left_join(kids_pm_check, by = "home_winter_id")
+  select(-homewinterid) %>% 
+  left_join(dt, by = c("area", "home_winter_id"))
+
+# Join dt_home with full dataset
+kids_pm_joined <- kids_pm_full %>% 
+  left_join(dt_home, by = c("area", "home_winter_id")) %>% 
+  rename(treatment = condition)
 
 ##### Add to new variables to Kids PM data #####
-kids_pm_new <- kids_pm %>% 
+kids_pm_new <- kids_pm_joined %>%  
+  #filter(area == "AK") %>% 
+  # apply correction factor to pm and change to ug
+  mutate(pm = pm/1.65,
+         pm = pm*1000) %>% 
   mutate(home_winter_id = as.character(home_winter_id),
@@
          pm_time = time,
-         pm_datetime = datetime,
-         pm = as.numeric(pm)) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
+         pm = as.numeric(pm),
+         day_of_week = weekdays(date)) %>% 
+  unite("pm_datetime", date:time, sep = " ") %>% 
+  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
+  arrange(area, home_winter_id, pm_date) %>% 
+  group_by(area, home_winter_id) %>% 
   # new variables summarizing pm2.5 within each home_winter_id
   mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
+         # new datetime values to calculate sampling interval
          min_datetime = first(pm_datetime),
@@
          diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
+         # total sampling interval, based on first and last datetimes
+         pm_sample_interval = diff_datetime/86400,
+         # expected number of observations with the sampling interval
+         expected_obs = pm_sample_interval*24*60,
+         # difference between actual and expected observations
+         actual_vs_expected_diff = pm_total_observations - expected_obs,
+         # indicator variable if actual vs expected obs > 10 or greater
+         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
+                                      actual_vs_expected_diff > -10, 1, 0),
+         pm_mean = mean(pm),
+         # breaking sampling period into days
          obs = row_number(),
@@
          sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
+  # calculate pm means for each sampling day
+  group_by(area, home_winter_id, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm),
          pm_sampling_day = sampling_day) %>% 
   ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
+  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, -datetime)
 

### 2019-12-09 pm_summary.Rmd
dusttrak cleaning and summary stats

--- 2019-12-05 pm_summary.Rmd
@@
 ```{r}
-summary_kids_pm <- kids_pm %>% 
-  na.exclude() %>% 
-  #filter(pm < 0)
-  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
-            min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm), mean(pm_total_days),
-            min(pm_total_days), max(pm_total_days),
-            min(pm_total_observations), max(pm_total_observations),
-            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
+# summary by area
+summary_area <- kids_pm %>%
+  #filter(actual_vs_expected == 1) %>% 
+  #filter(pm <0) %>% 
+  group_by(home_winter_id) %>% 
+  summarize("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE),
+            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
+            "Min Sample Days" = min(pm_sample_interval, na.rm = TRUE), 
+            "Max Sample Days" = max(pm_sample_interval, na.rm = TRUE),
+            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
+            "Min Sample Obs" = min(pm_total_observations, na.rm = TRUE), 
+            "Max Sample Obs" = max(pm_total_observations, na.rm = TRUE)) %>% 
   t()
-summary_kids_pm
+summary <- as.data.frame(summary_area)
+my_names <- summary_area[1, ]
+colnames(summary) <- my_names
+summary_area <- summary %>% 
+  rownames_to_column() %>% 
+  rename(Statistic = rowname) %>% 
+  filter(Statistic != "area") 
+#summary_area
+
+# summary by home_winter_id
+summary_id <- kids_pm %>% 
+  group_by(area, home_winter_id) %>% 
+  summarise("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE),
+            "Comments" = first(qaqccomments))
+
+# summary by dusttrak id
+summary_dtid <- kids_pm %>% 
+  group_by(area, dtid) %>% 
+  summarise("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE))
 ```
@@
 boxplots_kids_pm <- kids_pm %>% 
-  ggplot(aes(area, pm_24hr_mean)) +
+  filter(pm < 500) %>% 
+  ggplot(aes(area, pm)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
     theme_minimal() +
-    ylim(0, 400) +
+    ylim(-100, 500) +
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +

### 2019-12-12 pm_summary.Rmd
pm cleaning

@@
 library(knitr)
+library(lubridate)
+jvPalette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 ```
@@
             "Max PM" = max(pm, na.rm = TRUE),
+            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
+            "Act vs Exp" = first(actual_vs_expected),
             "Comments" = first(qaqccomments))
+write_csv(summary_id, "output/kids_summary_id.csv")
 
@@
             "Max PM" = max(pm, na.rm = TRUE))
+
+histogram_id <- kids_pm %>% 
+  filter(home_winter_id == "10" & area == "AK") %>% 
+  ggplot() + 
+    geom_histogram(aes(pm), bins = 50) +
+    theme_classic()
+histogram_id
+
+##### 2 Functions to look at PM data across time #####
+
+# Plot PM by datetime to look at trends, gaps, and peaks
+pm_time_plot_function <- function(id, location) {
+pm_time_plot <- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  #filter(pm < 30000) %>% 
+  arrange(pm_datetime) %>% 
+  mutate(pm_sampling_day = as.character(pm_sampling_day),
+         day_of_week = factor(day_of_week,
+                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
+                                         "Thursday", "Friday", "Saturday"))) %>%
+  ggplot() + 
+    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
+    theme_classic() +
+    labs(title = id,
+         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
+         x = "Datetime",
+         color = "Day of week") +
+    theme(axis.title.y = element_text(size = 12,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_text(size = 12),
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 12, color = "black"),
+          axis.line.x = element_line(colour = "black", size = 1), 
+          axis.line.y = element_line(colour = "black", size = 1), 
+          axis.ticks = element_blank()) +
+    scale_color_manual(values = jvPalette)
+
+gaps <<- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  mutate(pm_datetime = ymd_hms(pm_datetime),
+         datetime_lead = ymd_hms(lead(pm_datetime)),
+         datetime_1 = seconds(pm_datetime),
+         datetime_2 = seconds(datetime_lead),
+         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
+         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
+  arrange(desc(datetime_diff_minutes)) %>% 
+  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
+         pm_datetime, datetime_lead)
+
+pm_desc <<- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  arrange(desc(pm)) %>% 
+  select(pm_datetime, pm, pm_sampling_day)
+
+pm_time_plot
+}
+
+# Change function input to look at specific home
+# Run all of these lines together
+pm_time_plot_function(id = "228", location = "NN")
+head(gaps, 25)
+head(pm_desc, 25)
+tail(pm_desc, 25)
+
+# filter data for a single home
+check <- kids_pm %>% 
+  filter(area == "NN" & home_winter_id == "224") %>% 
+  filter(pm < 0)
 ```

### 2019-12-13 dustrak_data_cleaning.Rmd
further cleaning for PM data

--- 2019-12-09 data_cleaning_initial.Rmd
@@
 ---
-title: "KidsAIR initial data work"
+title: "KidsAIR initial PM data work"
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 9 Dec 2019"
+date: "Started 19 Nov 2019, Updated 13 Dec 2019"
 output: pdf_document
@@
 ```
-
-# Load, format, and save fuel moisture content data
-```{r}
-# Load data
-moisture_initial_wmt <- read_xlsx("Input/wmt_kids_moisture_20191122.xlsx")
-
-kids_moisture <- moisture_initial_wmt %>% 
-  # rename variables
-  mutate(moisture_id = as.factor(ID),  
-         home_winter_id = as.factor(HomeWinterID),
-         moisture_date = ymd(CompletedDate),
-         moisture_1 = as.numeric(Moisture1),
-         moisture_2 = as.numeric(Moisture2),
-         moisture_3 = as.numeric(Moisture3),
-         moisture_ave = as.numeric(MoistureAve),
-         moisture_missing = as.character(WhatMissing),
-         moisture_incomplete = as.factor(Incomplete),
-         moisture_ignore_missing = as.factor(IgnoreMissingData),
-         moisture_ignore_reason = as.character(IgnoreReason),
-         moisture_tech = as.character(Technician),
-         moisture_split = as.factor(Split),
-         area = "WMT") %>% 
-  # replace 0 and -999 values with NA
-  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
-                                 moisture_3 = 0, moisture_ave = 0)) %>% 
-  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
-                                 moisture_3 = -999)) %>% 
-  # gather 3 moisture samples to calculate average
-  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
-  group_by(home_winter_id, moisture_date) %>% 
-  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
-  ungroup() %>% 
-  # spread 3 moisture samples back to wide format
-  spread(sample, value) %>% 
-  select(moisture_id, home_winter_id, area, moisture_date, moisture_1, moisture_2, moisture_3, 
-         moisture_ave, moisture_missing, moisture_incomplete, moisture_ignore_missing,
-         moisture_ignore_reason, moisture_tech, moisture_split)
-
-# Save as RDS and CSV
-write_rds(kids_moisture, "Output/kids_moisture.rds")
-write_csv(kids_moisture, "Output/kids_moisture.csv")
-```
-
-# Load, format, and save home and child overview/demographic data
-```{r}
-# Load files
-wmt_child_initial <- read_xlsx("Input/wmt_child_20191126.xlsx")
-wmt_home_initial <- read_xlsx("Input/wmt_home_20191126.xlsx")
-wmt_homewinter_initial <- read_xlsx("Input/wmt_homewinter_20191126.xlsx")
-wmt_parent_demographics_initial <- read_xlsx("Input/wmt_parent_demographics_20191126.xlsx")
-
-# join all files from above
-joined1 <- wmt_child_initial %>% 
-  mutate(hispanic_child = as.factor(Hispanic),
-         race_child = as.factor(Race),
-         comments_child = Comments,
-         what_missing_child = WhatMissing,
-         incomplete_child = Incomplete,
-         ignore_missing_child = IgnoreMissingData,
-         ignore_reason_child = IgnoreReason) %>% 
-  select(-ID, -Hispanic, -Race, -Comments, -WhatMissing, -Incomplete,
-         -IgnoreMissingData, -IgnoreReason) %>% 
-  left_join(wmt_homewinter_initial, by = "HomeID") %>% 
-  left_join(wmt_home_initial, by = "HomeID") %>% 
-  left_join(wmt_parent_demographics_initial, by = "HomeWinterID") %>%
-  mutate(gender_parent = as.factor(Gender),
-         hispanic_parent = as.factor(Hispanic),
-         race_parent = as.factor(Race),
-         what_missing_parent = WhatMissing,
-         incomplete_parent = Incomplete,
-         ignore_missing_parent = IgnoreMissingData,
-         ignore_reason_parent = IgnoreReason) %>%
-  select(-ID, -Hispanic, -Race, -WhatMissing, -Incomplete,
-         -IgnoreMissingData, -IgnoreReason, -Gender, -EnterDate) %>%
-  rename_all(tolower)
-
-# create dataset of id's and treatment assignment
-wmt_ids_linked <- joined1 %>% 
-  select(childid, homeid, home, homewinterid, winterid, txcondition) %>% 
-  mutate(child_id = as.character(childid),
-         home_id = as.character(homeid),
-         home_winter_id = as.character(homewinterid),
-         winter_id = as.character(winterid),
-         treatment = as.factor(txcondition)) %>% 
-  select(-childid, -homeid, -homewinterid, -winterid, -txcondition)
-
-# create dataset of demographic data
-wmt_demographic_data <- joined1 %>% 
-  mutate(child_id = as.character(childid),
-         home_id = as.character(homeid),
-         age_nov1 = as.factor(agenov1),
-         child_assent_date = ymd(childassentdate),
-         parent_consent_date = ymd(parentconsentdate),
-         home_winter_id = as.character(homewinterid),
-         total_residents = as.numeric(totalres),
-         total_residents_under5 = as.numeric(underfive),
-         ignore_blanks_parent = ignoreblanks,
-         income = as.factor(income),
-         education = as.factor(education)) %>% 
-  select(child_id, home_id, hispanic_child, age_nov1, race_child, child_assent_date,
-         parent_consent_date, comments_child, what_missing_child, incomplete_child,
-         enrolled, ignore_missing_child, ignore_reason_child,
-         home_winter_id, gender_parent, hispanic_parent, race_parent, income,
-         education, total_residents, total_residents_under5, incomplete_parent,
-         ignore_blanks_parent, what_missing_parent, ignore_missing_parent,
-         ignore_reason_parent) %>% 
-  distinct(child_id, .keep_all = TRUE)
-
-# create dataset of data on home details
-## lower on priority list - see which variables we will use and clean later
-wmt_home_details_uncleaned <- wmt_homewinter_initial %>%
-  left_join(wmt_home_initial, by = "HomeID") %>% 
-  rename_all(tolower)
-
-# Save as RDS and CSV
-write_rds(wmt_ids_linked, "Output/wmt_ids_linked.rds")
-write_rds(wmt_demographic_data, "Output/wmt_demographic_data.rds")
-write_csv(wmt_ids_linked, "Output/wmt_ids_linked.csv")
-write_csv(wmt_demographic_data, "Output/wmt_demographic_data.csv")
-```
-
-```{r}
-
-```

### 2019-12-13 home_child_data_cleaning.Rmd
further cleaning for PM data

--- 2019-12-05 data_cleaning_initial.Rmd
@@
 ---
-title: "KidsAIR initial data work"
+title: "KidsAIR initial home/child data work"
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 5 Dec 2019"
+date: "Started 19 Nov 2019, Updated 13 Dec 2019"
 output: pdf_document
@@
 library(zoo)
-```
-
-# Load, format, and save KidsAIR Dusttrack data
-```{r}
-##### Read in dusttrack .txt files (from WMT) #####
-
-# List files
-list_txt_files <- list.files("Input/WMT/dusttrack_txt_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_txt_files")
-initial_data = tibble(file_list = list_txt_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv, skip = 28)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_txt <- initial_data %>% 
-  mutate(date = mdy(`MM/dd/yyyy`),
-         time = `hh:mm:ss`,
-         pm = as.numeric(`mg/m^3`),
-         area = "WMT") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area)
-
-##### Read in dusttrack .csv files (all areas) #####
-
-### WMT ###
-# List files
-list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "WMT") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-### NN ###
-# List files
-list_csv_files <- list.files("Input/NN/dusttrack_csv_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/NN/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-nn_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "NN") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-### AK ###
-# List files
-list_csv_files <- list.files("Input/AK/dusttrack_csv_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/AK/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-ak_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "AK") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-##### Read in other dusttrack .csv files (WMT and NN) #####
-
-### WMT ### 
-# List files
-list_other_files <- list.files("Input/WMT/dusttrack_other_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_other_files")
-initial_data = tibble(file_list = list_other_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_other <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(home_winter_id, trash1, trash2) %>% 
-  group_by(home_winter_id) %>% 
-  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash2, "NA"),
-         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash1, "NA"),
-         trash1 = as.numeric(trash1),
-         trash1 = as.character(trash1),
-         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
-  filter(!is.na(trash5)) %>% 
-  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
-  spread(trash4, trash3) %>% 
-  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
-  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
-  mutate(datetime = lead(datetime)) %>% 
-  replace_with_na(replace = list(datetime = "NA NA")) %>% 
-  mutate(datetime2 = first(datetime),
-         datetime = mdy_hms(datetime2),
-         trash1 = as.numeric(trash1)) %>% 
-  filter(!is.na(trash1)) %>% 
-  mutate(datetime2 = datetime + seconds(trash1),
-         datetime2 = ymd_hms(datetime2)) %>% 
-  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = datetime2,
-         area = "WMT") %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  ungroup()
-
-### NN ### 
-# List files
-list_other_files <- list.files("Input/NN/dusttrack_other_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/NN/dusttrack_other_files")
-initial_data = tibble(file_list = list_other_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-nn_pm_other <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(home_winter_id, trash1, trash2) %>% 
-  filter(!is.na(trash1)) %>% 
-  group_by(home_winter_id) %>% 
-  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash1, "NA"),
-         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash2, "NA"),
-         trash1 = as.numeric(trash1),
-         trash1 = as.character(trash1),
-         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
-  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
-  filter(!is.na(trash5)) %>% 
-  spread(trash3, trash4) %>% 
-  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
-  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
-  mutate(datetime = lead(datetime)) %>% 
-  replace_with_na(replace = list(datetime = "NA NA")) %>% 
-  mutate(datetime2 = first(datetime),
-         datetime = mdy_hms(datetime2),
-         trash1 = as.numeric(trash1)) %>% 
-  filter(!is.na(trash1)) %>% 
-  mutate(datetime2 = datetime + seconds(trash1),
-         datetime2 = ymd_hms(datetime2)) %>% 
-  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = datetime2,
-         area = "NN") %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  ungroup() %>% 
-  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
-                                  "22", home_winter_id),
-         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
-                                  "55", home_winter_id),
-         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
-                                  "81", home_winter_id))
-
-##### Bind all files from above #####
-kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
-                nn_pm_csv, nn_pm_other,
-                ak_pm_csv)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm_full.rds")
-#write_csv(kids_pm, "Output/kids_pm_full.csv")
-
-# Load data and clean up
-kids_pm <- read_rds("Output/kids_pm_full.rds")
-
-kids_pm_check <- kids_pm %>% 
-  group_by(home_winter_id) %>% 
-  summarise(obs = n()) %>% 
-  ungroup()
-
-ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
-nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
-wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")
-
-dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
-  rename_all(tolower) %>% 
-  filter(!is.na(homewinterid)) %>% 
-  mutate(home_winter_id = as.character(homewinterid)) %>% 
-  left_join(kids_pm_check, by = "home_winter_id")
-
-##### Add to new variables to Kids PM data #####
-kids_pm_new <- kids_pm %>% 
-  mutate(home_winter_id = as.character(home_winter_id),
-         area = as.factor(area),
-         pm_date = date,
-         pm_time = time,
-         pm_datetime = datetime,
-         pm = as.numeric(pm)) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
-         pm_sampling_day = sampling_day) %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm_new, "Output/kids_pm_new.rds")
-#write_csv(kids_pm_new, "Output/kids_pm_new.csv")
-```
-
-# Original code/files from Emily - save for future reference
-```{r}
-##### Load combined dusttrack files from Emily W #####
-##### Compare these with the raw data pulled in above #####
-
-## List sheets within the excel file
-list_sheets_nn <- excel_sheets("Input/nn_kids_dusttrack_20191121.xlsx")
-## Apply the read_xlsx function to each item in the sheet list
-## Bind the items from each sheet into a single dataframe
-pm_initial_nn <- data.frame(do.call("rbind", lapply(list_sheets_nn, read_xlsx, 
-                                      path = "Input/nn_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "NN")
-
-list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
-pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
-                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "WMT")
-
-# Bind data together from different study areas
-pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
-
-# Rename and create new variables
-kids_pm <- pm_initial %>% 
-  # separate time from incorrect/associated date
-  separate(time, into = c("trash", "time"), sep = " ") %>% 
-  # format/rename variables
-  mutate(home_winter_id = as.factor(homeid),
-         pm_date = ymd(date),
-         pm_time = hms(time),
-         pm = as.numeric(aerosol)) %>% 
-  # new datetime variable
-  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
-         pm_sampling_day = sampling_day) %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm.rds")
-write_csv(kids_pm, "Output/kids_pm.csv")
-```
-
-# Load, format, and save fuel moisture content data
-```{r}
-# Load data
-moisture_initial_wmt <- read_xlsx("Input/wmt_kids_moisture_20191122.xlsx")
-
-kids_moisture <- moisture_initial_wmt %>% 
-  # rename variables
-  mutate(moisture_id = as.factor(ID),  
-         home_winter_id = as.factor(HomeWinterID),
-         moisture_date = ymd(CompletedDate),
-         moisture_1 = as.numeric(Moisture1),
-         moisture_2 = as.numeric(Moisture2),
-         moisture_3 = as.numeric(Moisture3),
-         moisture_ave = as.numeric(MoistureAve),
-         moisture_missing = as.character(WhatMissing),
-         moisture_incomplete = as.factor(Incomplete),
-         moisture_ignore_missing = as.factor(IgnoreMissingData),
-         moisture_ignore_reason = as.character(IgnoreReason),
-         moisture_tech = as.character(Technician),
-         moisture_split = as.factor(Split),
-         area = "WMT") %>% 
-  # replace 0 and -999 values with NA
-  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
-                                 moisture_3 = 0, moisture_ave = 0)) %>% 
-  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
-                                 moisture_3 = -999)) %>% 
-  # gather 3 moisture samples to calculate average
-  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
-  group_by(home_winter_id, moisture_date) %>% 
-  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
-  ungroup() %>% 
-  # spread 3 moisture samples back to wide format
-  spread(sample, value) %>% 
-  select(moisture_id, home_winter_id, area, moisture_date, moisture_1, moisture_2, moisture_3, 
-         moisture_ave, moisture_missing, moisture_incomplete, moisture_ignore_missing,
-         moisture_ignore_reason, moisture_tech, moisture_split)
-
-# Save as RDS and CSV
-write_rds(kids_moisture, "Output/kids_moisture.rds")
-write_csv(kids_moisture, "Output/kids_moisture.csv")
 ```

### 2019-12-13 moisture_data_cleaning.Rmd
further cleaning for PM data

--- 2019-12-05 data_cleaning_initial.Rmd
@@
 ---
-title: "KidsAIR initial data work"
+title: "KidsAIR initial moisture content data work"
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 5 Dec 2019"
+date: "Started 19 Nov 2019, Updated 13 Dec 2019"
 output: pdf_document
@@
 library(zoo)
-```
-
-# Load, format, and save KidsAIR Dusttrack data
-```{r}
-##### Read in dusttrack .txt files (from WMT) #####
-
-# List files
-list_txt_files <- list.files("Input/WMT/dusttrack_txt_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_txt_files")
-initial_data = tibble(file_list = list_txt_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv, skip = 28)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_txt <- initial_data %>% 
-  mutate(date = mdy(`MM/dd/yyyy`),
-         time = `hh:mm:ss`,
-         pm = as.numeric(`mg/m^3`),
-         area = "WMT") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area)
-
-##### Read in dusttrack .csv files (all areas) #####
-
-### WMT ###
-# List files
-list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "WMT") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-### NN ###
-# List files
-list_csv_files <- list.files("Input/NN/dusttrack_csv_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/NN/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-nn_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "NN") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-### AK ###
-# List files
-list_csv_files <- list.files("Input/AK/dusttrack_csv_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/AK/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-ak_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "AK") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-##### Read in other dusttrack .csv files (WMT and NN) #####
-
-### WMT ### 
-# List files
-list_other_files <- list.files("Input/WMT/dusttrack_other_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_other_files")
-initial_data = tibble(file_list = list_other_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_other <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(home_winter_id, trash1, trash2) %>% 
-  group_by(home_winter_id) %>% 
-  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash2, "NA"),
-         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash1, "NA"),
-         trash1 = as.numeric(trash1),
-         trash1 = as.character(trash1),
-         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
-  filter(!is.na(trash5)) %>% 
-  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
-  spread(trash4, trash3) %>% 
-  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
-  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
-  mutate(datetime = lead(datetime)) %>% 
-  replace_with_na(replace = list(datetime = "NA NA")) %>% 
-  mutate(datetime2 = first(datetime),
-         datetime = mdy_hms(datetime2),
-         trash1 = as.numeric(trash1)) %>% 
-  filter(!is.na(trash1)) %>% 
-  mutate(datetime2 = datetime + seconds(trash1),
-         datetime2 = ymd_hms(datetime2)) %>% 
-  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = datetime2,
-         area = "WMT") %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  ungroup()
-
-### NN ### 
-# List files
-list_other_files <- list.files("Input/NN/dusttrack_other_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/NN/dusttrack_other_files")
-initial_data = tibble(file_list = list_other_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-nn_pm_other <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(home_winter_id, trash1, trash2) %>% 
-  filter(!is.na(trash1)) %>% 
-  group_by(home_winter_id) %>% 
-  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash1, "NA"),
-         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash2, "NA"),
-         trash1 = as.numeric(trash1),
-         trash1 = as.character(trash1),
-         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
-  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
-  filter(!is.na(trash5)) %>% 
-  spread(trash3, trash4) %>% 
-  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
-  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
-  mutate(datetime = lead(datetime)) %>% 
-  replace_with_na(replace = list(datetime = "NA NA")) %>% 
-  mutate(datetime2 = first(datetime),
-         datetime = mdy_hms(datetime2),
-         trash1 = as.numeric(trash1)) %>% 
-  filter(!is.na(trash1)) %>% 
-  mutate(datetime2 = datetime + seconds(trash1),
-         datetime2 = ymd_hms(datetime2)) %>% 
-  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = datetime2,
-         area = "NN") %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  ungroup() %>% 
-  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
-                                  "22", home_winter_id),
-         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
-                                  "55", home_winter_id),
-         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
-                                  "81", home_winter_id))
-
-##### Bind all files from above #####
-kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
-                nn_pm_csv, nn_pm_other,
-                ak_pm_csv)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm_full.rds")
-#write_csv(kids_pm, "Output/kids_pm_full.csv")
-
-# Load data and clean up
-kids_pm <- read_rds("Output/kids_pm_full.rds")
-
-kids_pm_check <- kids_pm %>% 
-  group_by(home_winter_id) %>% 
-  summarise(obs = n()) %>% 
-  ungroup()
-
-ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
-nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
-wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")
-
-dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
-  rename_all(tolower) %>% 
-  filter(!is.na(homewinterid)) %>% 
-  mutate(home_winter_id = as.character(homewinterid)) %>% 
-  left_join(kids_pm_check, by = "home_winter_id")
-
-##### Add to new variables to Kids PM data #####
-kids_pm_new <- kids_pm %>% 
-  mutate(home_winter_id = as.character(home_winter_id),
-         area = as.factor(area),
-         pm_date = date,
-         pm_time = time,
-         pm_datetime = datetime,
-         pm = as.numeric(pm)) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
-         pm_sampling_day = sampling_day) %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm_new, "Output/kids_pm_new.rds")
-#write_csv(kids_pm_new, "Output/kids_pm_new.csv")
-```
-
-# Original code/files from Emily - save for future reference
-```{r}
-##### Load combined dusttrack files from Emily W #####
-##### Compare these with the raw data pulled in above #####
-
-## List sheets within the excel file
-list_sheets_nn <- excel_sheets("Input/nn_kids_dusttrack_20191121.xlsx")
-## Apply the read_xlsx function to each item in the sheet list
-## Bind the items from each sheet into a single dataframe
-pm_initial_nn <- data.frame(do.call("rbind", lapply(list_sheets_nn, read_xlsx, 
-                                      path = "Input/nn_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "NN")
-
-list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
-pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
-                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "WMT")
-
-# Bind data together from different study areas
-pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
-
-# Rename and create new variables
-kids_pm <- pm_initial %>% 
-  # separate time from incorrect/associated date
-  separate(time, into = c("trash", "time"), sep = " ") %>% 
-  # format/rename variables
-  mutate(home_winter_id = as.factor(homeid),
-         pm_date = ymd(date),
-         pm_time = hms(time),
-         pm = as.numeric(aerosol)) %>% 
-  # new datetime variable
-  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
-         pm_sampling_day = sampling_day) %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm.rds")
-write_csv(kids_pm, "Output/kids_pm.csv")
 ```
@@
 ```
-
-# Load, format, and save home and child overview/demographic data
-```{r}
-# Load files
-wmt_child_initial <- read_xlsx("Input/wmt_child_20191126.xlsx")
-wmt_home_initial <- read_xlsx("Input/wmt_home_20191126.xlsx")
-wmt_homewinter_initial <- read_xlsx("Input/wmt_homewinter_20191126.xlsx")
-wmt_parent_demographics_initial <- read_xlsx("Input/wmt_parent_demographics_20191126.xlsx")
-
-# join all files from above
-joined1 <- wmt_child_initial %>% 
-  mutate(hispanic_child = as.factor(Hispanic),
-         race_child = as.factor(Race),
-         comments_child = Comments,
-         what_missing_child = WhatMissing,
-         incomplete_child = Incomplete,
-         ignore_missing_child = IgnoreMissingData,
-         ignore_reason_child = IgnoreReason) %>% 
-  select(-ID, -Hispanic, -Race, -Comments, -WhatMissing, -Incomplete,
-         -IgnoreMissingData, -IgnoreReason) %>% 
-  left_join(wmt_homewinter_initial, by = "HomeID") %>% 
-  left_join(wmt_home_initial, by = "HomeID") %>% 
-  left_join(wmt_parent_demographics_initial, by = "HomeWinterID") %>%
-  mutate(gender_parent = as.factor(Gender),
-         hispanic_parent = as.factor(Hispanic),
-         race_parent = as.factor(Race),
-         what_missing_parent = WhatMissing,
-         incomplete_parent = Incomplete,
-         ignore_missing_parent = IgnoreMissingData,
-         ignore_reason_parent = IgnoreReason) %>%
-  select(-ID, -Hispanic, -Race, -WhatMissing, -Incomplete,
-         -IgnoreMissingData, -IgnoreReason, -Gender, -EnterDate) %>%
-  rename_all(tolower)
-
-# create dataset of id's and treatment assignment
-wmt_ids_linked <- joined1 %>% 
-  select(childid, homeid, home, homewinterid, winterid, txcondition) %>% 
-  mutate(child_id = as.character(childid),
-         home_id = as.character(homeid),
-         home_winter_id = as.character(homewinterid),
-         winter_id = as.character(winterid),
-         treatment = as.factor(txcondition)) %>% 
-  select(-childid, -homeid, -homewinterid, -winterid, -txcondition)
-
-# create dataset of demographic data
-wmt_demographic_data <- joined1 %>% 
-  mutate(child_id = as.character(childid),
-         home_id = as.character(homeid),
-         age_nov1 = as.factor(agenov1),
-         child_assent_date = ymd(childassentdate),
-         parent_consent_date = ymd(parentconsentdate),
-         home_winter_id = as.character(homewinterid),
-         total_residents = as.numeric(totalres),
-         total_residents_under5 = as.numeric(underfive),
-         ignore_blanks_parent = ignoreblanks,
-         income = as.factor(income),
-         education = as.factor(education)) %>% 
-  select(child_id, home_id, hispanic_child, age_nov1, race_child, child_assent_date,
-         parent_consent_date, comments_child, what_missing_child, incomplete_child,
-         enrolled, ignore_missing_child, ignore_reason_child,
-         home_winter_id, gender_parent, hispanic_parent, race_parent, income,
-         education, total_residents, total_residents_under5, incomplete_parent,
-         ignore_blanks_parent, what_missing_parent, ignore_missing_parent,
-         ignore_reason_parent) %>% 
-  distinct(child_id, .keep_all = TRUE)
-
-# create dataset of data on home details
-## lower on priority list - see which variables we will use and clean later
-wmt_home_details_uncleaned <- wmt_homewinter_initial %>%
-  left_join(wmt_home_initial, by = "HomeID") %>% 
-  rename_all(tolower)
-
-# Save as RDS and CSV
-write_rds(wmt_ids_linked, "Output/wmt_ids_linked.rds")
-write_rds(wmt_demographic_data, "Output/wmt_demographic_data.rds")
-write_csv(wmt_ids_linked, "Output/wmt_ids_linked.csv")
-write_csv(wmt_demographic_data, "Output/wmt_demographic_data.csv")
-```
-
-```{r}
-
-```

### 2019-12-13 pm_summary.Rmd
further cleaning for PM data

--- 2019-12-06 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 5 Dec 2019"
+date: "Started 22 Nov 2019, Updated 13 Dec 2019"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = FALSE)
 ```
@@
 # Summary stats
+## Updated as data is cleaned
 ```{r}
-summary <- kids_pm %>% 
-  filter(pm < 0)
-
-summary_kids_pm <- kids_pm %>%
-  #filter(pm < 0)
-  summarize(mean(pm_24hr_mean_interval, na.rm = TRUE), 
-            sd(pm_24hr_mean_interval, na.rm = TRUE),
-            min(pm_24hr_mean_interval, na.rm = TRUE), 
-            max(pm_24hr_mean_interval, na.rm = TRUE),
-            mean(pm_24hr_mean_obs, na.rm = TRUE), 
-            sd(pm_24hr_mean_obs, na.rm = TRUE),
-            min(pm_24hr_mean_obs, na.rm = TRUE), 
-            max(pm_24hr_mean_obs, na.rm = TRUE),
-            min(pm, na.rm = TRUE), 
-            max(pm, na.rm = TRUE), 
-            mean(pm_sample_interval, na.rm = TRUE),
-            min(pm_sample_interval, na.rm = TRUE), 
-            max(pm_sample_interval, na.rm = TRUE),
-            min(pm_total_observations, na.rm = TRUE), 
-            max(pm_total_observations, na.rm = TRUE),
-            mean(pm_24hr_daily, na.rm = TRUE), 
-            mean(pm_24hr_daily_weighted, na.rm = TRUE)) %>% 
+# summary by area
+summary_area <- kids_pm %>%
+  # filtering out specific houses during data cleaning
+  ## decide for sure if they should be exluded going forward
+  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
+         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
+         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
+         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
+         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
+         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
+         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
+         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
+         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
+                              1, do_not_use)) %>% 
+  filter(do_not_use != 1) %>% 
+  # filter out high PM values
+  filter(pm < 5000) %>% 
+  #filter(actual_vs_expected == 1) %>% 
+  #filter(pm <0) %>% 
+  group_by(area) %>% 
+  # Run summary stats
+  summarize("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE),
+            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
+            "Min Sample Days" = min(pm_sample_interval, na.rm = TRUE), 
+            "Max Sample Days" = max(pm_sample_interval, na.rm = TRUE),
+            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
+            "Min Sample Obs" = min(pm_total_observations, na.rm = TRUE), 
+            "Max Sample Obs" = max(pm_total_observations, na.rm = TRUE)) %>% 
+  # transpose data for ease of reading
   t()
-kable(summary_kids_pm, digits = 2)
+# create column names and make it look nicer
+summary <- as.data.frame(summary_area)
+my_names <- summary_area[1, ]
+colnames(summary) <- my_names
+summary_area <- summary %>% 
+  rownames_to_column() %>% 
+  rename(Statistic = rowname) %>% 
+  filter(Statistic != "area") 
+kable(summary_area, digits = 2)
 ```
 
-# Boxplots
+# Summarize by ID and Dusttrak ID
+```{r}
+# summary by home_winter_id
+summary_id <- kids_pm %>% 
+  group_by(area, home_winter_id) %>% 
+  summarise("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE),
+            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
+            "Act vs Exp" = first(actual_vs_expected),
+            "Comments" = first(qaqccomments))
+# write_csv(summary_id, "output/kids_summary_id.csv")
+
+# summary by dusttrak id
+summary_dtid <- kids_pm %>% 
+  group_by(dtid) %>% 
+  summarise("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE))
+```
+
+# Histogram for distribution of PM by individual ID
+```{r}
+histogram_id <- kids_pm %>% 
+  filter(home_winter_id == "10" & area == "AK") %>% 
+  ggplot() + 
+    geom_histogram(aes(pm), bins = 50) +
+    theme_classic()
+histogram_id
+```
+
+# Function to look at PM data across time
+```{r}
+# Plot PM by datetime to look at trends, gaps, and peaks
+pm_time_plot_function <- function(id, location) {
+pm_time_plot <- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  #filter(pm < 30000) %>% 
+  arrange(pm_datetime) %>% 
+  mutate(pm_sampling_day = as.character(pm_sampling_day),
+         day_of_week = factor(day_of_week,
+                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
+                                         "Thursday", "Friday", "Saturday"))) %>%
+  ggplot() + 
+    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
+    theme_classic() +
+    labs(title = id,
+         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
+         x = "Datetime",
+         color = "Day of week") +
+    theme(axis.title.y = element_text(size = 12,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_text(size = 12),
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 12, color = "black"),
+          axis.line.x = element_line(colour = "black", size = 1), 
+          axis.line.y = element_line(colour = "black", size = 1), 
+          axis.ticks = element_blank()) +
+    scale_color_manual(values = jvPalette)
+
+gaps <<- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  mutate(pm_datetime = ymd_hms(pm_datetime),
+         datetime_lead = ymd_hms(lead(pm_datetime)),
+         datetime_1 = seconds(pm_datetime),
+         datetime_2 = seconds(datetime_lead),
+         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
+         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
+  arrange(desc(datetime_diff_minutes)) %>% 
+  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
+         pm_datetime, datetime_lead)
+
+pm_desc <<- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  arrange(desc(pm)) %>% 
+  select(pm_datetime, pm, pm_sampling_day)
+
+pm_time_plot
+}
+```
+
+# Run function from above to look at results
+```{r}
+# Change function input to look at specific home
+# Run all of these lines together
+pm_time_plot_function(id = "85", location = "AK")
+head(gaps, 25)
+head(pm_desc, 25)
+tail(pm_desc, 25)
+
+# filter data for a single home
+check_data <- kids_pm %>% 
+  filter(area == "WMT" & home_winter_id == "513") %>% 
+  filter(pm < 0)
+```
+
+# Boxplots for PM by study area
 ```{r, fig.height=6, fig.width=5}
 boxplots_kids_pm <- kids_pm %>% 
-  ggplot(aes(area, pm_24hr_mean)) +
+  filter(pm < 500) %>% 
+  ggplot(aes(area, pm)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 

### 2019-12-17 dustrak_data_cleaning.Rmd
general data cleaning

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 13 Dec 2019"
+date: "Started 19 Nov 2019, Updated 16 Dec 2019"
 output: pdf_document
@@
 
-# Load, format, and save KidsAIR Dusttrak data
+# Initial load, format, and save KidsAIR Dusttrak data
 ```{r}
@@
   rename_all(tolower) %>% 
-  select(homewinterid, dtid, area)
+  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
 nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
@@
   rename_all(tolower) %>% 
-  select(homewinterid, dtid, area)
+  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
 wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
@@
   rename_all(tolower) %>% 
-  select(homewinterid, dtid, area)
+  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
 
@@
          pm_sampling_day = sampling_day) %>% 
+  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
+  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
   ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, -datetime)
+  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
+         -datetime, -trash1, -trash2)
 
@@
 #write_csv(kids_pm_new, "Output/kids_pm_new.csv")
+```
+
+# Summarize by ID and Dusttrak ID to look for trends
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+# summary by home_winter_id
+summary_id <- kids_pm %>% 
+  group_by(area, home_winter_id) %>% 
+  summarise("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE),
+            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
+            "Act vs Exp" = first(actual_vs_expected),
+            "Comments" = first(qaqccomments))
+ write_csv(summary_id, "output/kids_summary_id.csv")
+
+# summary by dusttrak id
+summary_dtid <- kids_pm %>% 
+  group_by(dtid) %>% 
+  summarise("n" = n(),
+            "Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "Median PM" = median(pm, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE))
+```
+
+# Histogram for distribution of PM by individual ID
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+histogram_id <- kids_pm %>% 
+  filter(home_winter_id == "10" & area == "AK") %>% 
+  ggplot() + 
+    geom_histogram(aes(pm), bins = 50) +
+    theme_classic()
+histogram_id
+```
+
+# Function to look at PM data across time
+```{r}
+# Plot PM by datetime to look at trends, gaps, and peaks
+pm_time_plot_function <- function(id, location) {
+pm_time_plot <- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  #filter(pm < 30000) %>% 
+  arrange(pm_datetime) %>% 
+  mutate(pm_sampling_day = as.character(pm_sampling_day),
+         day_of_week = factor(day_of_week,
+                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
+                                         "Thursday", "Friday", "Saturday"))) %>%
+  ggplot() + 
+    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
+    theme_classic() +
+    labs(title = id,
+         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
+         x = "Datetime",
+         color = "Day of week") +
+    theme(axis.title.y = element_text(size = 12,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_text(size = 12),
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 12, color = "black"),
+          axis.line.x = element_line(colour = "black", size = 1), 
+          axis.line.y = element_line(colour = "black", size = 1), 
+          axis.ticks = element_blank()) +
+    scale_color_manual(values = jvPalette)
+
+gaps <<- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  mutate(pm_datetime = ymd_hms(pm_datetime),
+         datetime_lead = ymd_hms(lead(pm_datetime)),
+         datetime_1 = seconds(pm_datetime),
+         datetime_2 = seconds(datetime_lead),
+         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
+         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
+  arrange(desc(datetime_diff_minutes)) %>% 
+  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
+         pm_datetime, datetime_lead)
+
+pm_desc <<- kids_pm %>% 
+  filter(home_winter_id == id & area == location) %>% 
+  arrange(desc(pm)) %>% 
+  select(pm_datetime, pm, pm_sampling_day)
+
+pm_time_plot
+}
+```
+
+# Run function from above to look at results
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+# Change function input to look at specific home
+# Run all of these lines together
+pm_time_plot_function(id = "512", location = "WMT")
+head(gaps, 25)
+head(pm_desc, 25)
+tail(pm_desc, 25)
+
+# filter data for a single home
+check_data <- kids_pm %>% 
+  filter(area == "AK" & home_winter_id == "10")
+
+%>% 
+  filter(pm < 0)
+```
+
+# Check sample start times
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+# Checking difference between dustrak start datetime and field data log datetime
+sample_times <- kids_pm %>% 
+  #filter(area == "AK") %>% 
+  group_by(area, home_winter_id) %>% 
+  arrange(pm_datetime) %>% 
+  unite(start_datetime, c("startdate", "starttime"), 
+        sep = " ", remove = FALSE) %>% 
+  mutate(first_datetime = first(pm_datetime),
+         first_datetime = ymd_hms(first_datetime),
+         start_datetime = ymd_hms(start_datetime),
+         time_diff = first_datetime - start_datetime) %>% 
+  distinct(time_diff, .keep_all = TRUE) %>% 
+  select(area, home_winter_id, pm_datetime, first_datetime,
+         start_datetime, time_diff, startdate, starttime) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id)
+#write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
+```
+
+# Loading and combining notes from time and PM logs
+```{r}
+time_notes <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
+  select(area, home_winter_id, first_datetime, start_datetime,
+         time_diff, next_steps_time, check_log_time)
+
+pm_notes <- read_xlsx("Input/kids_summary_id.xlsx") %>% 
+  select(area, home_winter_id, n, further_comments_pm, next_steps_pm, check_log_pm)
+
+kids_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+  mutate(home_winter_id = as.numeric(home_winter_id))
+
+notes_joined <- pm_notes %>% 
+  left_join(time_notes, by = c("area", "home_winter_id")) %>% 
+  left_join(kids_ids, by = c("area", "home_winter_id"))
+#write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
 ```

### 2019-12-17 home_child_data_cleaning.Rmd
general data cleaning

@@
 
-# Load, format, and save home and child overview/demographic data
+# Load, format, and save home and child overview/demographic data for WMT
 ```{r}
 # Load files
-wmt_child_initial <- read_xlsx("Input/wmt_child_20191126.xlsx")
-wmt_home_initial <- read_xlsx("Input/wmt_home_20191126.xlsx")
-wmt_homewinter_initial <- read_xlsx("Input/wmt_homewinter_20191126.xlsx")
-wmt_parent_demographics_initial <- read_xlsx("Input/wmt_parent_demographics_20191126.xlsx")
+wmt_child_initial <- read_xlsx("Input/WMT/wmt_child_20191126.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
+  mutate(child_id_num = as.character(childid),
+         home_id = as.character(homeid),
+         child_id_char = as.character(id),
+         hispanic_child = as.factor(hispanic),
+         race_child = as.factor(race),
+         age_child_nov = as.character(agenov1)) %>% 
+  select(child_id_num, home_id, child_id_char, hispanic_child,
+         race_child, age_child_nov, enrolled)
+
+wmt_home_initial <- read_xlsx("Input/WMT/wmt_home_20191126.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homeid, home, txcondition) %>% 
+  mutate(home_id = as.character(homeid),
+         home = as.character(home),
+         treatment = as.factor(txcondition)) %>% 
+  select(home_id, home, treatment)
+
+wmt_homewinter_initial <- read_xlsx("Input/WMT/wmt_homewinter_20191126.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         home_id = as.character(homeid),
+         winter_id = as.character(winterid),
+         intervention_date = interventiondate,
+         intervention_time = interventiontime) %>% 
+  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
+
+wmt_parent_demographics_initial <-
+  read_xlsx("Input/WMT/wmt_parent_demographics_20191126.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, gender, hispanic, race, income, education,
+         totalres, underfive) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         gender_parent = as.factor(gender),
+         hispanic_parent = as.factor(hispanic),
+         race_parent = as.factor(race),
+         income = as.factor(income),
+         education = as.factor(education),
+         total_residents = as.character(totalres),
+         residents_under_five = as.character(underfive)) %>% 
+  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
+         income, education, total_residents, residents_under_five) %>% 
+  filter(!is.na(gender_parent))
 
 # join all files from above
-joined1 <- wmt_child_initial %>% 
-  mutate(hispanic_child = as.factor(Hispanic),
-         race_child = as.factor(Race),
-         comments_child = Comments,
-         what_missing_child = WhatMissing,
-         incomplete_child = Incomplete,
-         ignore_missing_child = IgnoreMissingData,
-         ignore_reason_child = IgnoreReason) %>% 
-  select(-ID, -Hispanic, -Race, -Comments, -WhatMissing, -Incomplete,
-         -IgnoreMissingData, -IgnoreReason) %>% 
-  left_join(wmt_homewinter_initial, by = "HomeID") %>% 
-  left_join(wmt_home_initial, by = "HomeID") %>% 
-  left_join(wmt_parent_demographics_initial, by = "HomeWinterID") %>%
-  mutate(gender_parent = as.factor(Gender),
-         hispanic_parent = as.factor(Hispanic),
-         race_parent = as.factor(Race),
-         what_missing_parent = WhatMissing,
-         incomplete_parent = Incomplete,
-         ignore_missing_parent = IgnoreMissingData,
-         ignore_reason_parent = IgnoreReason) %>%
-  select(-ID, -Hispanic, -Race, -WhatMissing, -Incomplete,
-         -IgnoreMissingData, -IgnoreReason, -Gender, -EnterDate) %>%
-  rename_all(tolower)
+joined1 <- wmt_homewinter_initial %>% 
+  left_join(wmt_parent_demographics_initial, by = "home_winter_id") %>% 
+  left_join(wmt_home_initial, by = "home_id") %>% 
+  left_join(wmt_child_initial, by = "home_id") %>% 
+  mutate(area = "WMT")
 
 # create dataset of id's and treatment assignment
-wmt_ids_linked <- joined1 %>% 
-  select(childid, homeid, home, homewinterid, winterid, txcondition) %>% 
-  mutate(child_id = as.character(childid),
-         home_id = as.character(homeid),
-         home_winter_id = as.character(homewinterid),
-         winter_id = as.character(winterid),
-         treatment = as.factor(txcondition)) %>% 
-  select(-childid, -homeid, -homewinterid, -winterid, -txcondition)
+wmt_home_ids_linked <- joined1 %>% 
+  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
+  distinct(home_winter_id, .keep_all = TRUE) %>% 
+  arrange(home_id, winter_id)
 
-# create dataset of demographic data
-wmt_demographic_data <- joined1 %>% 
-  mutate(child_id = as.character(childid),
-         home_id = as.character(homeid),
-         age_nov1 = as.factor(agenov1),
-         child_assent_date = ymd(childassentdate),
-         parent_consent_date = ymd(parentconsentdate),
-         home_winter_id = as.character(homewinterid),
-         total_residents = as.numeric(totalres),
-         total_residents_under5 = as.numeric(underfive),
-         ignore_blanks_parent = ignoreblanks,
-         income = as.factor(income),
-         education = as.factor(education)) %>% 
-  select(child_id, home_id, hispanic_child, age_nov1, race_child, child_assent_date,
-         parent_consent_date, comments_child, what_missing_child, incomplete_child,
-         enrolled, ignore_missing_child, ignore_reason_child,
-         home_winter_id, gender_parent, hispanic_parent, race_parent, income,
-         education, total_residents, total_residents_under5, incomplete_parent,
-         ignore_blanks_parent, what_missing_parent, ignore_missing_parent,
-         ignore_reason_parent) %>% 
-  distinct(child_id, .keep_all = TRUE)
-
-# create dataset of data on home details
-## lower on priority list - see which variables we will use and clean later
-wmt_home_details_uncleaned <- wmt_homewinter_initial %>%
-  left_join(wmt_home_initial, by = "HomeID") %>% 
-  rename_all(tolower)
+# create dataset of child demographic data
+wmt_child_demographics <- joined1 %>% 
+  distinct(child_id_num, .keep_all = TRUE) %>% 
+  arrange(home_id, child_id_num, winter_id) %>% 
+  filter(enrolled == 1)
 
 # Save as RDS and CSV
-write_rds(wmt_ids_linked, "Output/wmt_ids_linked.rds")
-write_rds(wmt_demographic_data, "Output/wmt_demographic_data.rds")
-write_csv(wmt_ids_linked, "Output/wmt_ids_linked.csv")
-write_csv(wmt_demographic_data, "Output/wmt_demographic_data.csv")
+write_rds(wmt_home_ids_linked, "Output/wmt_ids_linked.rds")
+write_rds(wmt_child_demographics, "Output/wmt_demographic_data.rds")
+write_csv(wmt_home_ids_linked, "Output/wmt_ids_linked.csv")
+write_csv(wmt_child_demographics, "Output/wmt_demographic_data.csv")
 ```
 
+# Load, format, and save home and child overview/demographic data for NN
 ```{r}
+# Load files
+nn_child_initial <- read_xlsx("Input/NN/nn_child_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
+  mutate(child_id_num = as.character(childid),
+         home_id = as.character(homeid),
+         child_id_char = as.character(id),
+         hispanic_child = as.factor(hispanic),
+         race_child = as.factor(race),
+         age_child_nov = as.character(agenov1)) %>% 
+  select(child_id_num, home_id, child_id_char, hispanic_child,
+         race_child, age_child_nov, enrolled)
 
+nn_home_initial <- read_xlsx("Input/NN/nn_home_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homeid, home, txcondition) %>% 
+  mutate(home_id = as.character(homeid),
+         home = as.character(home),
+         treatment = as.factor(txcondition)) %>% 
+  select(home_id, home, treatment)
+
+nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         home_id = as.character(homeid),
+         winter_id = as.character(winterid),
+         intervention_date = interventiondate,
+         intervention_time = interventiontime) %>% 
+  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
+
+nn_parent_demographics_initial <-
+  read_xlsx("Input/NN/nn_parent_demographics_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, gender, hispanic, race, income, education,
+         totalres, underfive) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         gender_parent = as.factor(gender),
+         hispanic_parent = as.factor(hispanic),
+         race_parent = as.factor(race),
+         income = as.factor(income),
+         education = as.factor(education),
+         total_residents = as.character(totalres),
+         residents_under_five = as.character(underfive)) %>% 
+  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
+         income, education, total_residents, residents_under_five) %>% 
+  filter(!is.na(gender_parent))
+
+# join all files from above
+joined1 <- nn_homewinter_initial %>% 
+  left_join(nn_parent_demographics_initial, by = "home_winter_id") %>% 
+  left_join(nn_home_initial, by = "home_id") %>% 
+  left_join(nn_child_initial, by = "home_id") %>% 
+  mutate(area = "NN")
+
+# create dataset of id's and treatment assignment
+nn_home_ids_linked <- joined1 %>% 
+  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
+  distinct(home_winter_id, .keep_all = TRUE) %>% 
+  arrange(home_id, winter_id)
+
+# create dataset of child demographic data
+nn_child_demographics <- joined1 %>% 
+  distinct(child_id_num, .keep_all = TRUE) %>% 
+  arrange(home_id, child_id_num, winter_id) %>% 
+  filter(enrolled == 1)
+
+# Save as RDS and CSV
+write_rds(nn_home_ids_linked, "Output/nn_ids_linked.rds")
+write_rds(nn_child_demographics, "Output/nn_demographic_data.rds")
+write_csv(nn_home_ids_linked, "Output/nn_ids_linked.csv")
+write_csv(nn_child_demographics, "Output/nn_demographic_data.csv")
 ```
+
+# Load, format, and save home and child overview/demographic data for AK
+```{r}
+# Load files
+ak_child_initial <- read_xlsx("Input/AK/ak_child_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
+  mutate(child_id_num = as.character(childid),
+         home_id = as.character(homeid),
+         child_id_char = as.character(id),
+         hispanic_child = as.factor(hispanic),
+         race_child = as.factor(race),
+         age_child_nov = as.character(agenov1)) %>% 
+  select(child_id_num, home_id, child_id_char, hispanic_child,
+         race_child, age_child_nov, enrolled)
+
+ak_home_initial <- read_xlsx("Input/AK/ak_home_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homeid, home, txcondition) %>% 
+  mutate(home_id = as.character(homeid),
+         home = as.character(home),
+         treatment = as.factor(txcondition)) %>% 
+  select(home_id, home, treatment)
+
+ak_homewinter_initial <- read_xlsx("Input/AK/ak_homewinter_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         home_id = as.character(homeid),
+         winter_id = as.character(winterid),
+         intervention_date = interventiondate,
+         intervention_time = interventiontime) %>% 
+  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
+
+ak_parent_demographics_initial <-
+  read_xlsx("Input/AK/ak_parent_demographics_20191216.xlsx") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, gender, hispanic, race, income, education,
+         totalres, underfive) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         gender_parent = as.factor(gender),
+         hispanic_parent = as.factor(hispanic),
+         race_parent = as.factor(race),
+         income = as.factor(income),
+         education = as.factor(education),
+         total_residents = as.character(totalres),
+         residents_under_five = as.character(underfive)) %>% 
+  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
+         income, education, total_residents, residents_under_five) %>% 
+  filter(!is.na(gender_parent))
+
+# join all files from above
+joined1 <- ak_homewinter_initial %>% 
+  left_join(ak_parent_demographics_initial, by = "home_winter_id") %>% 
+  left_join(ak_home_initial, by = "home_id") %>% 
+  left_join(ak_child_initial, by = "home_id") %>% 
+  mutate(area = "AK")
+
+# create dataset of id's and treatment assignment
+ak_home_ids_linked <- joined1 %>% 
+  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
+  distinct(home_winter_id, .keep_all = TRUE) %>% 
+  arrange(home_id, winter_id)
+
+# create dataset of child demographic data
+ak_child_demographics <- joined1 %>% 
+  distinct(child_id_num, .keep_all = TRUE) %>% 
+  arrange(home_id, child_id_num, winter_id) %>% 
+  filter(enrolled == 1)
+
+# Save as RDS and CSV
+write_rds(ak_home_ids_linked, "Output/ak_ids_linked.rds")
+write_rds(ak_child_demographics, "Output/ak_demographic_data.rds")
+write_csv(ak_home_ids_linked, "Output/ak_ids_linked.csv")
+write_csv(ak_child_demographics, "Output/ak_demographic_data.csv")
+```
+
+# Combine linked id data for WMT, NN, AK
+```{r}
+ak_ids <- read_rds("Output/ak_ids_linked.rds")
+nn_ids <- read_rds("Output/nn_ids_linked.rds")
+wmt_ids <- read_rds("Output/wmt_ids_linked.rds")
+
+kids_ids <- rbind(ak_ids, nn_ids, wmt_ids)
+
+#write_rds(kids_ids, "Output/kids_linked_ids.rds")
+```
+
+# Combine demographic data for WMT, NN, AK
+```{r}
+ak_demographics <- read_rds("Output/ak_demographic_data.rds")
+nn_demographics <- read_rds("Output/nn_demographic_data.rds")
+wmt_demographics <- read_rds("Output/wmt_demographic_data.rds")
+
+kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics)
+
+#write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
+```

### 2019-12-17 pm_summary.Rmd
general data cleaning

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 13 Dec 2019"
+date: "Started 22 Nov 2019, Updated 16 Dec 2019"
 output: pdf_document
@@
 library(readr)
+library(readxl)
 library(knitr)
@@
 
-# Summarize by ID and Dusttrak ID
-```{r}
-# summary by home_winter_id
-summary_id <- kids_pm %>% 
-  group_by(area, home_winter_id) %>% 
-  summarise("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "Median PM" = median(pm, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
-            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
-            "Act vs Exp" = first(actual_vs_expected),
-            "Comments" = first(qaqccomments))
-# write_csv(summary_id, "output/kids_summary_id.csv")
-
-# summary by dusttrak id
-summary_dtid <- kids_pm %>% 
-  group_by(dtid) %>% 
-  summarise("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "Median PM" = median(pm, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE))
-```
-
-# Histogram for distribution of PM by individual ID
-```{r}
-histogram_id <- kids_pm %>% 
-  filter(home_winter_id == "10" & area == "AK") %>% 
-  ggplot() + 
-    geom_histogram(aes(pm), bins = 50) +
-    theme_classic()
-histogram_id
-```
-
-# Function to look at PM data across time
-```{r}
-# Plot PM by datetime to look at trends, gaps, and peaks
-pm_time_plot_function <- function(id, location) {
-pm_time_plot <- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  #filter(pm < 30000) %>% 
-  arrange(pm_datetime) %>% 
-  mutate(pm_sampling_day = as.character(pm_sampling_day),
-         day_of_week = factor(day_of_week,
-                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
-                                         "Thursday", "Friday", "Saturday"))) %>%
-  ggplot() + 
-    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
-    theme_classic() +
-    labs(title = id,
-         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
-         x = "Datetime",
-         color = "Day of week") +
-    theme(axis.title.y = element_text(size = 12,
-                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
-          axis.title.x = element_text(size = 12),
-          axis.text.x = element_blank(),
-          axis.text.y = element_text(size = 12, color = "black"),
-          axis.line.x = element_line(colour = "black", size = 1), 
-          axis.line.y = element_line(colour = "black", size = 1), 
-          axis.ticks = element_blank()) +
-    scale_color_manual(values = jvPalette)
-
-gaps <<- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  mutate(pm_datetime = ymd_hms(pm_datetime),
-         datetime_lead = ymd_hms(lead(pm_datetime)),
-         datetime_1 = seconds(pm_datetime),
-         datetime_2 = seconds(datetime_lead),
-         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
-         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
-  arrange(desc(datetime_diff_minutes)) %>% 
-  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
-         pm_datetime, datetime_lead)
-
-pm_desc <<- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  arrange(desc(pm)) %>% 
-  select(pm_datetime, pm, pm_sampling_day)
-
-pm_time_plot
-}
-```
-
-# Run function from above to look at results
-```{r}
-# Change function input to look at specific home
-# Run all of these lines together
-pm_time_plot_function(id = "85", location = "AK")
-head(gaps, 25)
-head(pm_desc, 25)
-tail(pm_desc, 25)
-
-# filter data for a single home
-check_data <- kids_pm %>% 
-  filter(area == "WMT" & home_winter_id == "513") %>% 
-  filter(pm < 0)
-```
-
 # Boxplots for PM by study area

### 2019-12-18 dustrak_data_cleaning.Rmd
PM data cleaning

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 16 Dec 2019"
+date: "Started 19 Nov 2019, Updated 18 Dec 2019"
 output: pdf_document
@@
 library(zoo)
+jvPalette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 ```
@@
 # Run all of these lines together
-pm_time_plot_function(id = "512", location = "WMT")
+pm_time_plot_function(id = "140", location = "NN")
 head(gaps, 25)
@@
 check_data <- kids_pm %>% 
-  filter(area == "AK" & home_winter_id == "10")
+  filter(area == "NN" & home_winter_id == "140")
 
@@
 #write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
+```
+
+# Fix dusttrak sample start times
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+# Checking difference between dustrak start datetime and field data log datetime
+sample_times <- kids_pm %>% 
+  #filter(area == "NN") %>% 
+  group_by(area, home_winter_id) %>% 
+  arrange(pm_datetime) %>% 
+  unite(start_datetime, c("startdate", "starttime"), 
+        sep = " ", remove = FALSE) %>% 
+  mutate(first_datetime = first(pm_datetime),
+         first_datetime = ymd_hms(first_datetime),
+         start_datetime = ymd_hms(start_datetime),
+         time_diff = first_datetime - start_datetime) %>% 
+  #distinct(time_diff, .keep_all = TRUE) %>% 
+  # make general time changes based on ranges
+  mutate(pm_datetime = ymd_hms(pm_datetime),
+         pm_datetime_new = if_else(time_diff > 1800 & time_diff < 5400, 
+                                   pm_datetime - 3600, pm_datetime),
+         pm_datetime_new = if_else(time_diff < -1800 & time_diff > -5400, 
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff > 5400 & time_diff < 9000, 
+                                   pm_datetime_new - 7200, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff < -5400 & time_diff > -9000, 
+                                   pm_datetime_new + 7200, pm_datetime_new)) %>% 
+  # make specific time changes
+  mutate(pm_datetime_new = if_else(time_diff == 10903, # WMA457
+                                   pm_datetime_new - 10800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == (-82721), # WMA440
+                                   pm_datetime_new + 82800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 10069, # WMA459
+                                   pm_datetime_new - 9900, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 9817, # WMA314
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff < -31535900, # YKC06-103 
+                                   pm_datetime_new + 31539519, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff < -31532200, # WMA305
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 9120, # WMA228
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == (-21497), # WMA130
+                                   pm_datetime_new + 21497, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 684, # WMA237
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 149793, # WMA238
+                                   pm_datetime_new - 149793, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -46227, # CH320
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 13514, # TC360 
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -10696, # TC357 
+                                   pm_datetime_new + 10800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -43427, # TC355 
+                                   pm_datetime_new + 43200, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -11388, # TC353 
+                                   pm_datetime_new + 10800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -221, # TC351 
+                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
+  mutate(first_datetime_new = first(pm_datetime_new),
+         first_datetime_new = ymd_hms(first_datetime_new),
+         start_datetime = ymd_hms(start_datetime),
+         time_diff_new = first_datetime_new - start_datetime) %>% 
+  distinct(time_diff, .keep_all = TRUE)
+
 ```

### 2019-12-18 home_child_data_cleaning.Rmd
PM data cleaning

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 16 Dec 2019"
+date: "Started 19 Nov 2019, Updated 17 Dec 2019"
 output: pdf_document
@@
 # Load files
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > WMT > child.xlsx
 wmt_child_initial <- read_xlsx("Input/WMT/wmt_child_20191126.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > WMT > home.xlsx
 wmt_home_initial <- read_xlsx("Input/WMT/wmt_home_20191126.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > WMT > homewinter.xlsx
 wmt_homewinter_initial <- read_xlsx("Input/WMT/wmt_homewinter_20191126.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > WMT > Home Data > parentdemographics.xlsx
 wmt_parent_demographics_initial <-
@@
 # Load files
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > Navajo > child.xlsx
 nn_child_initial <- read_xlsx("Input/NN/nn_child_20191216.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > Navajo > home.xlsx
 nn_home_initial <- read_xlsx("Input/NN/nn_home_20191216.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > Navajo > homewinter.xlsx
 nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20191216.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > Navajo > Home Data > parentdemographics.xlsx
 nn_parent_demographics_initial <-
@@
 # Load files
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > AK > child.xlsx
 ak_child_initial <- read_xlsx("Input/AK/ak_child_20191216.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > AK > home.xlsx
 ak_home_initial <- read_xlsx("Input/AK/ak_home_20191216.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > AK > homewinter.xlsx
 ak_homewinter_initial <- read_xlsx("Input/AK/ak_homewinter_20191216.xlsx") %>% 
@@
 
+## Box location: All Files > NoonGroupData > KidsAIR > 
+## 2019-November-Data > AK > Home Data > parentdemographics.xlsx
 ak_parent_demographics_initial <-

### 2019-12-20 dustrak_data_cleaning.Rmd
PM cleaning

--- 2019-12-17 dustrak_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 16 Dec 2019"
+date: "Started 19 Nov 2019, Updated 19 Dec 2019"
 output: pdf_document
@@
 library(zoo)
+library(knitr)
+jv_palette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 ```
@@
 
-# Load Dusttrak data and clean up
+# Load Dusttrak data and do initial clean-up
 ```{r}
@@
 # Save pm dataset as RDS and CSV
-write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+#write_rds(kids_pm_new, "Output/kids_pm_new.rds")
 #write_csv(kids_pm_new, "Output/kids_pm_new.csv")
@@
           axis.ticks = element_blank()) +
-    scale_color_manual(values = jvPalette)
+    scale_color_manual(values = jv_palette)
 
@@
 # Run all of these lines together
-pm_time_plot_function(id = "512", location = "WMT")
+pm_time_plot_function(id = "137", location = "NN")
 head(gaps, 25)
@@
 #write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
+```
+
+# Fix dusttrak sample start times
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+# Checking difference between dustrak start datetime and field data log datetime
+kids_pm_new2 <- kids_pm_new %>% 
+  #filter(area == "NN") %>% 
+  group_by(area, home_winter_id) %>% 
+  arrange(pm_datetime) %>% 
+  unite(start_datetime, c("startdate", "starttime"), 
+        sep = " ", remove = FALSE) %>% 
+  mutate(first_datetime = first(pm_datetime),
+         first_datetime = ymd_hms(first_datetime),
+         start_datetime = ymd_hms(start_datetime),
+         time_diff = first_datetime - start_datetime) %>% 
+  #distinct(time_diff, .keep_all = TRUE) %>% 
+  # make general time changes based on ranges
+  mutate(pm_datetime = ymd_hms(pm_datetime),
+         pm_datetime_new = if_else(time_diff > 1800 & time_diff < 5400, 
+                                   pm_datetime - 3600, pm_datetime),
+         pm_datetime_new = if_else(time_diff < -1800 & time_diff > -5400, 
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff > 5400 & time_diff < 9000, 
+                                   pm_datetime_new - 7200, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff < -5400 & time_diff > -9000, 
+                                   pm_datetime_new + 7200, pm_datetime_new)) %>% 
+  # make specific time changes
+  mutate(pm_datetime_new = if_else(time_diff == 10903, # WMA457
+                                   pm_datetime_new - 10800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == (-82721), # WMA440
+                                   pm_datetime_new + 82800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 10069, # WMA459
+                                   pm_datetime_new - 9900, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 9817, # WMA314
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff < -31535900, # YKC06-103 
+                                   pm_datetime_new + 31539519, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff < -31532200, # WMA305
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 9120, # WMA228
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == (-21497), # WMA130
+                                   pm_datetime_new + 21497, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 684, # WMA237
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 149793, # WMA238
+                                   pm_datetime_new - 149793, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -46227, # CH320
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 13514, # TC360 
+                                   pm_datetime_new - 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -10696, # TC357 
+                                   pm_datetime_new + 10800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -43427, # TC355 
+                                   pm_datetime_new + 43200, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -11388, # TC353 
+                                   pm_datetime_new + 10800, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == -221, # TC351 
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 4999, # CH214 
+                                   pm_datetime_new + 3600, pm_datetime_new),
+         pm_datetime_new = if_else(time_diff == 4061, # WMA203 
+                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
+  mutate(first_datetime_new = first(pm_datetime_new),
+         first_datetime_new = ymd_hms(first_datetime_new),
+         start_datetime = ymd_hms(start_datetime),
+         time_diff_new = first_datetime_new - start_datetime) %>%  
+  distinct(time_diff, .keep_all = TRUE)
+
+kids_pm_new2 <- kids_pm_new %>%
+  ungroup() %>% 
+  arrange(area, home_winter_id, pm_datetime_new)
+
+# write_rds(kids_pm_new2, "Output/kids_pm_new.rds")
+```
+
+# Further cleaning:
+## Remove specific homes from PM dataset
+## Remove extremely high values
+## Change negative values
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+
+kids_pm_new <- kids_pm %>%
+  # filtering out specific homes where PM data is not suitable for analysis
+  # removes 119,520 obs
+         # YKC06_105: only 12 hours of data
+  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
+         # YKC06_313: sample hourly instead of every minute
+         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
+         # CH_302: sample hourly instead of every minute
+         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
+         # CH_304: Home was resampled
+         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
+         # CH_306: sample hourly instead of every minute
+         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
+         # CH_307: only 8 datapoints
+         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
+         # TC_357: Home was resampled
+         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
+         # CH_322: sample hourly instead of every minute
+         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
+         # CH_325: 6300 obs are negative, down to -12.12. Median is negative.
+         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
+         # CH_400: Home was resampled
+         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
+         # CH_403: Home was resampled
+         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
+         # CH_108: Only 300 datapoints and no other files found
+         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
+         # CH_207: Only 400 datapoints; no other files found
+         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
+         # TC_230: Home was resampled
+         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
+         # CH_223: Home was resampled
+         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
+         # TC_242: 1 day worth of data with lots of gaps
+         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
+         # WMA277: Home was resampled
+         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
+         # WMA305: Home was resampled
+         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
+         # WMA306: Home was resampled
+         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
+         # WMA358: Home was resampled
+         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
+         # WMA448: Home was resampled
+         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
+         # WMA200: 4700 negative values, down to -4.85. Median is negative
+         do_not_use = if_else(area == "WMT" & home_winter_id == "152", 1, do_not_use),
+         # WMA201: 5000 negative values, down to -10. Median negative. 17 hour gap 
+         do_not_use = if_else(area == "WMT" & home_winter_id == "153", 1, do_not_use),
+         # WMA603: Home was resampled
+         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
+                              1, do_not_use)) %>% 
+  filter(do_not_use != 1) %>% 
+  # replace PM > 10,000 with NA (33 obs)
+  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
+  replace_with_na(replace = list(pm = 999999)) %>% 
+  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
+  # 291,490 obs less than 1; 56,684 obs less than 0 
+  mutate(pm = if_else(pm < 1, 0.5, pm))
+
+# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+```
+
+# New variables to use for analysis
+```{r}
+# Load data
+kids_pm <- read_rds("Output/kids_pm_new.rds")
+cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
+  select(area, home_winter_id, final_cleaning_pm, sensitivity_analysis_pm)
+
+kids_pm_new <- kids_pm %>% 
+  mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
+  left_join(cleaning_summary, by = c("area", "home_winter_id")) %>% 
+  select(-samplingdate, -samplingtime, -pickupdate, -pickuptime,
+         -dtmin, -dtmax, -dtaverage, -qaqccomments, -data_comments,
+         -pm_sampling_day, -pm_mean_daily) %>% 
+  # create sampling days based on datetime differences
+  group_by(area, home_winter_id) %>% 
+  mutate(pm_datetime_new = if_else(is.na(pm_datetime_new), 
+                                   pm_datetime, pm_datetime_new),
+         first_datetime_new = first(pm_datetime_new),
+         first_datetime_new = ymd_hms(first_datetime_new)) %>% 
+  mutate(datetime_diff = pm_datetime_new - first_datetime_new,
+         sampling_day = if_else(datetime_diff <= 86400, 1, 0),
+         sampling_day = if_else(datetime_diff > 86400 & datetime_diff <= 172800, 
+                                2, sampling_day),
+         sampling_day = if_else(datetime_diff > 172800 & datetime_diff <= 259200, 
+                                3, sampling_day),
+         sampling_day = if_else(datetime_diff > 259200 & datetime_diff <= 345600, 
+                                4, sampling_day),
+         sampling_day = if_else(datetime_diff > 345600 & datetime_diff <= 432000, 
+                                5, sampling_day),
+         sampling_day = if_else(datetime_diff > 432000 & datetime_diff <= 518400, 
+                                6, sampling_day),
+         sampling_day = if_else(datetime_diff > 518400 & datetime_diff <= 604800, 
+                                7, sampling_day),
+         sampling_day = if_else(datetime_diff > 604800 & datetime_diff <= 691200, 
+                                8, sampling_day),
+         sampling_day = if_else(datetime_diff > 691200 & datetime_diff <= 777600, 
+                                9, sampling_day),
+         sampling_day = if_else(datetime_diff > 777600 & datetime_diff <= 864000, 
+                                10, sampling_day),
+         sampling_day = if_else(datetime_diff > 864000, 
+                                11, sampling_day)) %>% 
+  # new vars for daily means and % expected observations for each day
+  group_by(area, home_winter_id, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm),
+         percent_daily_obs_pm = n()/1440) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, pm_datetime_new) %>% 
+  select(-first_datetime, -time_diff, -do_not_use)
+
+test <- kids_pm_new %>% 
+  filter(percent_daily_obs_pm >= 0.9) %>% 
+  summarise(n()/2241695)
+# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
 ```

### 2019-12-20 pm_summary.Rmd
PM cleaning

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 16 Dec 2019"
+date: "Started 22 Nov 2019, Updated 19 Dec 2019"
 output: pdf_document
@@
 library(lubridate)
-jvPalette <- c("#330099","#CC0066","#FF6633", 
+jv_palette <- c("#330099","#CC0066","#FF6633", 
                  "#0099CC", "#FF9900","#CC6633",
@@
 # Summary stats
-## Updated as data is cleaned
 ```{r}
-# summary by area
-summary_area <- kids_pm %>%
-  # filtering out specific houses during data cleaning
-  ## decide for sure if they should be exluded going forward
-  mutate(do_not_use = if_else(area == "AK" & home_winter_id == "39", 1, 0),
-         do_not_use = if_else(area == "AK" & home_winter_id == "89", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "127", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "1330", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "134", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "135", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "1510", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "161", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "165", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "2150", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "2210", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "29", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "58", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "620", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "840", 1, do_not_use),
-         do_not_use = if_else(area == "NN" & home_winter_id == "86", 1, do_not_use),
-         do_not_use = if_else(area == "WMT" & home_winter_id == "1910", 1, do_not_use),
-         do_not_use = if_else(area == "WMT" & home_winter_id == "2500", 1, do_not_use),
-         do_not_use = if_else(area == "WMT" & home_winter_id == "2510", 1, do_not_use),
-         do_not_use = if_else(area == "WMT" & home_winter_id == "2860", 1, do_not_use),
-         do_not_use = if_else(area == "WMT" & home_winter_id == "3750", 1, do_not_use),
-         do_not_use = if_else(area == "WMT" & home_winter_id == "5100", 
-                              1, do_not_use)) %>% 
-  filter(do_not_use != 1) %>% 
-  # filter out high PM values
-  filter(pm < 5000) %>% 
-  #filter(actual_vs_expected == 1) %>% 
-  #filter(pm <0) %>% 
-  group_by(area) %>% 
-  # Run summary stats
+pm_summary <- kids_pm %>% 
+  # group by: area, treatment, dtid, day_of_week, Sampling_day
+  group_by(day_of_week) %>% 
   summarize("n" = n(),
@@
             "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
-            "Min Sample Days" = min(pm_sample_interval, na.rm = TRUE), 
-            "Max Sample Days" = max(pm_sample_interval, na.rm = TRUE),
             "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
-            "Min Sample Obs" = min(pm_total_observations, na.rm = TRUE), 
-            "Max Sample Obs" = max(pm_total_observations, na.rm = TRUE)) %>% 
-  # transpose data for ease of reading
-  t()
-# create column names and make it look nicer
-summary <- as.data.frame(summary_area)
-my_names <- summary_area[1, ]
-colnames(summary) <- my_names
-summary_area <- summary %>% 
-  rownames_to_column() %>% 
-  rename(Statistic = rowname) %>% 
-  filter(Statistic != "area") 
-kable(summary_area, digits = 2)
+            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE))
 ```
@@
 boxplots_kids_pm <- kids_pm %>% 
-  filter(pm < 500) %>% 
-  ggplot(aes(area, pm)) +
+  #filter(pm < 500) %>% 
+  ggplot(aes(area, log(pm))) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
     theme_minimal() +
-    ylim(-100, 500) +
+    #ylim(0, 500) +
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +

### 2019-12-20 dustrak_data_cleaning.Rmd
PM cleaning

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 19 Dec 2019"
+date: "Started 19 Nov 2019, Updated 20 Dec 2019"
 output: pdf_document
@@
 # Initial load, format, and save KidsAIR Dusttrak data
+## Files here are in 3 formats: csv (2 different formats) and txt
+## I had to do some manual cleaning of the files to get them named
+## consistently and in reasonably similar formats to be able to load
+## into R in batches, as seen below. The files I upload in the code below
+## will also be uploaded to Box in a folder with my name.
+## Box folder directory: 
 ```{r}
@@
 # Run all of these lines together
-pm_time_plot_function(id = "137", location = "NN")
+pm_time_plot_function(id = "129", location = "WMT")
 head(gaps, 25)
@@
 check_data <- kids_pm %>% 
-  filter(area == "NN" & home_winter_id == "140")
-
-%>% 
+  filter(area == "WMT" & home_winter_id == "129") %>% 
   filter(pm < 0)
@@
 # Checking difference between dustrak start datetime and field data log datetime
-kids_pm_new2 <- kids_pm_new %>% 
+kids_pm_new <- kids_pm %>% 
   #filter(area == "NN") %>% 
@@
          time_diff_new = first_datetime_new - start_datetime) %>%  
-  distinct(time_diff, .keep_all = TRUE)
-
-kids_pm_new2 <- kids_pm_new %>%
+  #distinct(time_diff, .keep_all = TRUE) %>%
   ungroup() %>% 
@@
 
-# write_rds(kids_pm_new2, "Output/kids_pm_new.rds")
+# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
 ```
@@
 ```
-
-# Original code/files from Emily - save for future reference
-```{r}
-##### Load combined dusttrack files from Emily W #####
-##### Compare these with the raw data pulled in above #####
-
-## List sheets within the excel file
-list_sheets_nn <- excel_sheets("Input/nn_kids_dusttrack_20191121.xlsx")
-## Apply the read_xlsx function to each item in the sheet list
-## Bind the items from each sheet into a single dataframe
-pm_initial_nn <- data.frame(do.call("rbind", lapply(list_sheets_nn, read_xlsx, 
-                                      path = "Input/nn_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "NN")
-
-list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
-pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
-                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "WMT")
-
-# Bind data together from different study areas
-pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
-
-# Rename and create new variables
-kids_pm <- pm_initial %>% 
-  # separate time from incorrect/associated date
-  separate(time, into = c("trash", "time"), sep = " ") %>% 
-  # format/rename variables
-  mutate(home_winter_id = as.factor(homeid),
-         pm_date = ymd(date),
-         pm_time = hms(time),
-         pm = as.numeric(aerosol)) %>% 
-  # new datetime variable
-  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
-         pm_sampling_day = sampling_day) %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm.rds")
-write_csv(kids_pm, "Output/kids_pm.csv")
-```

### 2019-12-20 pm_summary.Rmd
PM cleaning

@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(day_of_week) %>% 
+  group_by(sampling_day) %>% 
   summarize("n" = n(),
@@
   #filter(pm < 500) %>% 
-  ggplot(aes(area, log(pm))) +
+  ggplot(aes(area, pm)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
     theme_minimal() +
-    #ylim(0, 500) +
+    ylim(0, 100) +
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +

### 2019-12-20 home_child_data_cleaning.Rmd
PM cleaning
Cleaning updates

@@
 #write_rds(kids_ids, "Output/kids_linked_ids.rds")
+
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 ```
@@
 #write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
+
+kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
 ```

### 2019-12-20 pm_summary.Rmd
Cleaning updates

@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(sampling_day) %>% 
+  group_by(area) %>% 
   summarize("n" = n(),

### 2019-12-23 sums_data_cleaning.Rmd
New SUMs data cleaning RMD

--- 2019-12-18 dustrak_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 18 Dec 2019"
+date: "Started 23 Dec 2019, Updated 23 Dec 2019"
 output: pdf_document
@@
 library(zoo)
-jvPalette <- c("#330099","#CC0066","#FF6633", 
+library(knitr)
+jv_palette <- c("#330099","#CC0066","#FF6633", 
                  "#0099CC", "#FF9900","#CC6633",
@@
 
+```{r}
+wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         date = ChangeDate,
+         time = ChangeTime,
+         alarms = Alarms,
+         file_name = ibutFile,
+         notes = `REPLACE(REPLACE(Notes`) %>% 
+  select(home_winter_id, date, time, alarms, file_name, notes)
+
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+  filter(area == "WMT") %>% 
+  select(home_winter_id, home)
+
+wmt_sums_log <- wmt_sums_data_log %>% 
+  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, date, time, alarms, file_name, notes)
+
+write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")
+```
+
 # Initial load, format, and save KidsAIR Dusttrak data
+## Files here are in 3 formats: csv (2 different formats) and txt
+## I had to do some manual cleaning of the files to get them named
+## consistently and in reasonably similar formats to be able to load
+## into R in batches, as seen below. The files I upload in the code below
+## will also be uploaded to Box in a folder with my name.
+## Box folder directory: 
 ```{r}
-##### Read in dusttrack .txt files (from WMT) #####
+##### Read in sums csv files (from WMT) #####
 
 # List files
-list_txt_files <- list.files("Input/WMT/dusttrack_txt_files")
+list_csv_files <- list.files("Input/WMT/sums_csv_new")
 
 # Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_txt_files")
-initial_data = tibble(file_list = list_txt_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv, skip = 28)) %>%
+## run next 12 lines together
+setwd("Input/WMT/sums_csv_new")
+initial_data = tibble(file_list = list_csv_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 1,
+                            col_names = c("line", "date", "time", "temp_c", "trash"),
+                            col_types = cols(line = col_double(),
+                                             date = col_date(format = ""),
+                                             time = col_time(format = ""),
+                                             temp_c = col_character(),
+                                             trash = col_character()))) %>%
   unnest(save_data) %>%
@@
 # Format combined file from above; combine with other PM data below
-wmt_pm_txt <- initial_data %>% 
-  mutate(date = mdy(`MM/dd/yyyy`),
-         time = `hh:mm:ss`,
-         pm = as.numeric(`mg/m^3`),
-         area = "WMT") %>% 
+wmt_sums_csv <- initial_data %>% 
+  mutate(area = "WMT",
+         home2 = "WMA") %>% 
   unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area)
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = datetime,
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums)
 
+# List files
+list_csv_files_2 <- list.files("Input/WMT/sums_csv_old/format_1")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 7 lines together
+setwd("Input/WMT/sums_csv_old/format_1")
+initial_data = tibble(file_list = list_csv_files_2) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 22,
+                            col_names = c("datetime_sums", "temp_c_sums"))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_csv_2 <- initial_data %>% 
+  mutate(area = "WMT",
+         home2 = "WMA",
+         datetime_sums = ymd_hms(datetime_sums)) %>% 
+  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(date_sums = ymd(date_sums)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  na.exclude()
+
+# List files
+list_csv_files_3 <- list.files("Input/WMT/sums_csv_old/format_2")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 12 lines together
+setwd("Input/WMT/sums_csv_old/format_2")
+initial_data = tibble(file_list = list_csv_files_3) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 2,
+                            col_names = c("line", "date", "time", "temp_f", "trash"),
+                            col_types = cols(line = col_double(),
+                                             date = col_character(),
+                                             time = col_time(format = ""),
+                                             temp_f = col_double(),
+                                             trash = col_character()))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_csv_3 <- initial_data %>% 
+  mutate(area = "WMT",
+         home2 = "WMA") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = mdy_hms(datetime),
+         date_sums = mdy(date),
+         time_sums = time,
+         temp_c_sums = (temp_f-32)/1.8) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums)
+
+##### Read in sums csv files (from WMT) #####
+
+# List files
+list_csv_files <- list.files("Input/WMT/sums_xlsx_files")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 12 lines together
+setwd("Input/WMT/sums_xlsx_files")
+initial_data = tibble(file_list = list_csv_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 1,
+                            col_names = c("line", "date", "time", "temp_c", "trash"),
+                            col_types = cols(line = col_double(),
+                                             date = col_date(format = ""),
+                                             time = col_time(format = ""),
+                                             temp_c = col_character(),
+                                             trash = col_character()))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_csv <- initial_data %>% 
+  mutate(area = "WMT",
+         home2 = "WMA") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = datetime,
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums)
+```
+
+# Code from PM cleaning
+```{r}
 ##### Read in dusttrack .csv files (all areas) #####
@@
 ```
-
-# Load Dusttrak data and clean up
-```{r}
-# Load full, raw dataset compiled in chunk above
-kids_pm_full <- read_rds("Output/kids_pm_full_raw.rds")
-
-# Load and combine data sheets with QAQC comments, etc
-ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
-  mutate(area = "AK")
-ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
-  mutate(area = "AK") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
-nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
-  mutate(area = "NN")
-nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
-  mutate(area = "NN") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
-wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
-  mutate(area = "WMT")
-wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
-  mutate(area = "WMT") %>% 
-  rename_all(tolower) %>% 
-  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
-
-dt <- rbind(ak_dt, nn_dt, wmt_dt) %>% 
-  mutate(home_winter_id = as.character(homewinterid)) %>% 
-  select(-homewinterid)
-
-dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
-  rename_all(tolower) %>% 
-  filter(!is.na(homewinterid)) %>% 
-  mutate(home_winter_id = as.character(homewinterid)) %>% 
-  select(-homewinterid) %>% 
-  left_join(dt, by = c("area", "home_winter_id"))
-
-# Join dt_home with full dataset
-kids_pm_joined <- kids_pm_full %>% 
-  left_join(dt_home, by = c("area", "home_winter_id")) %>% 
-  rename(treatment = condition)
-
-##### Add to new variables to Kids PM data #####
-kids_pm_new <- kids_pm_joined %>%  
-  #filter(area == "AK") %>% 
-  # apply correction factor to pm and change to ug
-  mutate(pm = pm/1.65,
-         pm = pm*1000) %>% 
-  mutate(home_winter_id = as.character(home_winter_id),
-         area = as.factor(area),
-         pm_date = date,
-         pm_time = time,
-         pm = as.numeric(pm),
-         day_of_week = weekdays(date)) %>% 
-  unite("pm_datetime", date:time, sep = " ") %>% 
-  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
-  arrange(area, home_winter_id, pm_date) %>% 
-  group_by(area, home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         # new datetime values to calculate sampling interval
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         # total sampling interval, based on first and last datetimes
-         pm_sample_interval = diff_datetime/86400,
-         # expected number of observations with the sampling interval
-         expected_obs = pm_sample_interval*24*60,
-         # difference between actual and expected observations
-         actual_vs_expected_diff = pm_total_observations - expected_obs,
-         # indicator variable if actual vs expected obs > 10 or greater
-         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
-                                      actual_vs_expected_diff > -10, 1, 0),
-         pm_mean = mean(pm),
-         # breaking sampling period into days
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  # calculate pm means for each sampling day
-  group_by(area, home_winter_id, sampling_day) %>% 
-  mutate(pm_mean_daily = mean(pm),
-         pm_sampling_day = sampling_day) %>% 
-  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
-  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
-         -datetime, -trash1, -trash2)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm_new, "Output/kids_pm_new.rds")
-#write_csv(kids_pm_new, "Output/kids_pm_new.csv")
-```
-
-# Summarize by ID and Dusttrak ID to look for trends
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-
-# summary by home_winter_id
-summary_id <- kids_pm %>% 
-  group_by(area, home_winter_id) %>% 
-  summarise("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "Median PM" = median(pm, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
-            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
-            "Act vs Exp" = first(actual_vs_expected),
-            "Comments" = first(qaqccomments))
- write_csv(summary_id, "output/kids_summary_id.csv")
-
-# summary by dusttrak id
-summary_dtid <- kids_pm %>% 
-  group_by(dtid) %>% 
-  summarise("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "Median PM" = median(pm, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE))
-```
-
-# Histogram for distribution of PM by individual ID
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-
-histogram_id <- kids_pm %>% 
-  filter(home_winter_id == "10" & area == "AK") %>% 
-  ggplot() + 
-    geom_histogram(aes(pm), bins = 50) +
-    theme_classic()
-histogram_id
-```
-
-# Function to look at PM data across time
-```{r}
-# Plot PM by datetime to look at trends, gaps, and peaks
-pm_time_plot_function <- function(id, location) {
-pm_time_plot <- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  #filter(pm < 30000) %>% 
-  arrange(pm_datetime) %>% 
-  mutate(pm_sampling_day = as.character(pm_sampling_day),
-         day_of_week = factor(day_of_week,
-                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
-                                         "Thursday", "Friday", "Saturday"))) %>%
-  ggplot() + 
-    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
-    theme_classic() +
-    labs(title = id,
-         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
-         x = "Datetime",
-         color = "Day of week") +
-    theme(axis.title.y = element_text(size = 12,
-                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
-          axis.title.x = element_text(size = 12),
-          axis.text.x = element_blank(),
-          axis.text.y = element_text(size = 12, color = "black"),
-          axis.line.x = element_line(colour = "black", size = 1), 
-          axis.line.y = element_line(colour = "black", size = 1), 
-          axis.ticks = element_blank()) +
-    scale_color_manual(values = jvPalette)
-
-gaps <<- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  mutate(pm_datetime = ymd_hms(pm_datetime),
-         datetime_lead = ymd_hms(lead(pm_datetime)),
-         datetime_1 = seconds(pm_datetime),
-         datetime_2 = seconds(datetime_lead),
-         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
-         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
-  arrange(desc(datetime_diff_minutes)) %>% 
-  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
-         pm_datetime, datetime_lead)
-
-pm_desc <<- kids_pm %>% 
-  filter(home_winter_id == id & area == location) %>% 
-  arrange(desc(pm)) %>% 
-  select(pm_datetime, pm, pm_sampling_day)
-
-pm_time_plot
-}
-```
-
-# Run function from above to look at results
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-
-# Change function input to look at specific home
-# Run all of these lines together
-pm_time_plot_function(id = "140", location = "NN")
-head(gaps, 25)
-head(pm_desc, 25)
-tail(pm_desc, 25)
-
-# filter data for a single home
-check_data <- kids_pm %>% 
-  filter(area == "NN" & home_winter_id == "140")
-
-%>% 
-  filter(pm < 0)
-```
-
-# Check sample start times
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-
-# Checking difference between dustrak start datetime and field data log datetime
-sample_times <- kids_pm %>% 
-  #filter(area == "AK") %>% 
-  group_by(area, home_winter_id) %>% 
-  arrange(pm_datetime) %>% 
-  unite(start_datetime, c("startdate", "starttime"), 
-        sep = " ", remove = FALSE) %>% 
-  mutate(first_datetime = first(pm_datetime),
-         first_datetime = ymd_hms(first_datetime),
-         start_datetime = ymd_hms(start_datetime),
-         time_diff = first_datetime - start_datetime) %>% 
-  distinct(time_diff, .keep_all = TRUE) %>% 
-  select(area, home_winter_id, pm_datetime, first_datetime,
-         start_datetime, time_diff, startdate, starttime) %>% 
-  ungroup() %>% 
-  arrange(area, home_winter_id)
-#write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
-```
-
-# Loading and combining notes from time and PM logs
-```{r}
-time_notes <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
-  select(area, home_winter_id, first_datetime, start_datetime,
-         time_diff, next_steps_time, check_log_time)
-
-pm_notes <- read_xlsx("Input/kids_summary_id.xlsx") %>% 
-  select(area, home_winter_id, n, further_comments_pm, next_steps_pm, check_log_pm)
-
-kids_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
-  mutate(home_winter_id = as.numeric(home_winter_id))
-
-notes_joined <- pm_notes %>% 
-  left_join(time_notes, by = c("area", "home_winter_id")) %>% 
-  left_join(kids_ids, by = c("area", "home_winter_id"))
-#write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
-```
-
-# Fix dusttrak sample start times
-```{r}
-# Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
-
-# Checking difference between dustrak start datetime and field data log datetime
-sample_times <- kids_pm %>% 
-  #filter(area == "NN") %>% 
-  group_by(area, home_winter_id) %>% 
-  arrange(pm_datetime) %>% 
-  unite(start_datetime, c("startdate", "starttime"), 
-        sep = " ", remove = FALSE) %>% 
-  mutate(first_datetime = first(pm_datetime),
-         first_datetime = ymd_hms(first_datetime),
-         start_datetime = ymd_hms(start_datetime),
-         time_diff = first_datetime - start_datetime) %>% 
-  #distinct(time_diff, .keep_all = TRUE) %>% 
-  # make general time changes based on ranges
-  mutate(pm_datetime = ymd_hms(pm_datetime),
-         pm_datetime_new = if_else(time_diff > 1800 & time_diff < 5400, 
-                                   pm_datetime - 3600, pm_datetime),
-         pm_datetime_new = if_else(time_diff < -1800 & time_diff > -5400, 
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff > 5400 & time_diff < 9000, 
-                                   pm_datetime_new - 7200, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff < -5400 & time_diff > -9000, 
-                                   pm_datetime_new + 7200, pm_datetime_new)) %>% 
-  # make specific time changes
-  mutate(pm_datetime_new = if_else(time_diff == 10903, # WMA457
-                                   pm_datetime_new - 10800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == (-82721), # WMA440
-                                   pm_datetime_new + 82800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 10069, # WMA459
-                                   pm_datetime_new - 9900, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 9817, # WMA314
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff < -31535900, # YKC06-103 
-                                   pm_datetime_new + 31539519, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff < -31532200, # WMA305
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 9120, # WMA228
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == (-21497), # WMA130
-                                   pm_datetime_new + 21497, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 684, # WMA237
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 149793, # WMA238
-                                   pm_datetime_new - 149793, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -46227, # CH320
-                                   pm_datetime_new + 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == 13514, # TC360 
-                                   pm_datetime_new - 3600, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -10696, # TC357 
-                                   pm_datetime_new + 10800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -43427, # TC355 
-                                   pm_datetime_new + 43200, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -11388, # TC353 
-                                   pm_datetime_new + 10800, pm_datetime_new),
-         pm_datetime_new = if_else(time_diff == -221, # TC351 
-                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
-  mutate(first_datetime_new = first(pm_datetime_new),
-         first_datetime_new = ymd_hms(first_datetime_new),
-         start_datetime = ymd_hms(start_datetime),
-         time_diff_new = first_datetime_new - start_datetime) %>% 
-  distinct(time_diff, .keep_all = TRUE)
-
-```
-
-# Original code/files from Emily - save for future reference
-```{r}
-##### Load combined dusttrack files from Emily W #####
-##### Compare these with the raw data pulled in above #####
-
-## List sheets within the excel file
-list_sheets_nn <- excel_sheets("Input/nn_kids_dusttrack_20191121.xlsx")
-## Apply the read_xlsx function to each item in the sheet list
-## Bind the items from each sheet into a single dataframe
-pm_initial_nn <- data.frame(do.call("rbind", lapply(list_sheets_nn, read_xlsx, 
-                                      path = "Input/nn_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "NN")
-
-list_sheets_wmt <- excel_sheets("Input/wmt_kids_dusttrack_20191121.xlsx")
-pm_initial_wmt <- data.frame(do.call("rbind", lapply(list_sheets_wmt, read_xlsx, 
-                                      path = "Input/wmt_kids_dusttrack_20191121.xlsx"))) %>% 
-  mutate(area = "WMT")
-
-# Bind data together from different study areas
-pm_initial <- rbind(pm_initial_nm, pm_initial_wmt)
-
-# Rename and create new variables
-kids_pm <- pm_initial %>% 
-  # separate time from incorrect/associated date
-  separate(time, into = c("trash", "time"), sep = " ") %>% 
-  # format/rename variables
-  mutate(home_winter_id = as.factor(homeid),
-         pm_date = ymd(date),
-         pm_time = hms(time),
-         pm = as.numeric(aerosol)) %>% 
-  # new datetime variable
-  unite(pm_datetime, c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
-  arrange(home_winter_id, pm_date) %>% 
-  group_by(home_winter_id) %>% 
-  # new variables summarizing pm2.5 within each home_winter_id
-  mutate(pm_total_observations = n(),
-         pm_total_summed = sum(pm),
-         min_datetime = first(pm_datetime),
-         max_datetime = last(pm_datetime),
-         min_datetime = ymd_hms(min_datetime),
-         max_datetime = ymd_hms(max_datetime),
-         diff_datetime = interval(min_datetime, max_datetime),
-         pm_total_days = diff_datetime/86400,
-         pm_24hr_mean = pm_total_summed/pm_total_days,
-         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
-         obs = row_number(),
-         sampling_day = if_else(obs < 1441, 1, 0),
-         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
-         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
-         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
-         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
-         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
-         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
-         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
-  group_by(home_winter_id, sampling_day) %>% 
-  mutate(pm_24hr_daily = sum(pm),
-         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
-         pm_sampling_day = sampling_day) %>% 
-  ungroup() %>% 
-  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm.rds")
-write_csv(kids_pm, "Output/kids_pm.csv")
-```

### 2019-12-24 sums_data_cleaning.Rmd
SUMs cleaning

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 23 Dec 2019"
+date: "Started 23 Dec 2019, Updated 24 Dec 2019"
 output: pdf_document
@@
 ```{r}
-wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>% 
+wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
   filter(area == "WMT") %>% 
-  select(home_winter_id, home)
+  select(home_winter_id, home, home_id)
 
@@
   left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, date, time, alarms, file_name, notes)
+  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes)
 

### 2019-12-28 sums_data_cleaning.Rmd
SUMs cleaning

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 24 Dec 2019"
+date: "Started 23 Dec 2019, Updated 27 Dec 2019"
 output: pdf_document
@@
 
+# Add linked IDs to WMT SUMs data logs
 ```{r}
-wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
+wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         ibid = iButtonID,
+         date = InstallDate,
+         time = InstallTime) %>% 
+  select(home_winter_id, ibid, date, time)
+
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+  filter(area == "WMT") %>% 
+  select(home_winter_id, home, home_id)
+
+wmt_sums_log <- wmt_sums_data_log %>% 
+  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, ibid) %>% 
+  arrange(home, date)
+
+write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")
+
+wmt_sums_data_log_updated <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
+wmt_sums_log <- wmt_sums_data_log_updated %>% 
+  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
+  arrange(home, date)
+
+write_csv(wmt_sums_log, "Output/wmt_sums_data_log_updated.csv")
+```
+
+# Add linked IDs to NN SUMs data logs
+```{r}
+nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         ibid = iButtonID,
+         date = InstallDate,
+         time = InstallTime) %>% 
+  select(home_winter_id, ibid, date, time)
+
 kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
-  filter(area == "WMT") %>% 
+  filter(area == "NN") %>% 
   select(home_winter_id, home, home_id)
 
-wmt_sums_log <- wmt_sums_data_log %>% 
+nn_sums_log <- nn_sums_data_log %>% 
   left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes)
+  select(home_winter_id, home, home_id, date, time, ibid) %>% 
+  arrange(home, date)
 
-write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")
+write_csv(nn_sums_log, "Output/nn_sums_data_log.csv")
+
+nn_sums_data_log_updated <- read_xlsx("Input/NN/nn_sums_data_log_updated.xlsx") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         date = ChangeDate,
+         time = ChangeTime,
+         alarms = Alarms,
+         file_name = ibutFile,
+         notes = IgnoreReason) %>% 
+  select(home_winter_id, date, time, alarms, file_name, notes)
+
+nn_sums_log <- nn_sums_data_log_updated %>% 
+  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
+  arrange(home, date)
+
+write_csv(nn_sums_log, "Output/nn_sums_data_log_updated.csv")
 ```
 
-# Initial load, format, and save KidsAIR Dusttrak data
-## Files here are in 3 formats: csv (2 different formats) and txt
+# Add linked IDs to AK SUMs data logs
+```{r}
+ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         ibid = iButtonID,
+         date = InstallDate,
+         time = InstallTime) %>% 
+  select(home_winter_id, ibid, date, time)
+
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+  filter(area == "AK") %>% 
+  select(home_winter_id, home, home_id)
+
+ak_sums_log <- ak_sums_data_log %>% 
+  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, ibid) %>% 
+  arrange(home, date)
+
+write_csv(ak_sums_log, "Output/ak_sums_data_log.csv")
+
+ak_sums_data_log_updated <- read_xlsx("Input/AK/ak_sums_data_log_updated.xlsx") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         date = ChangeDate,
+         time = ChangeTime,
+         alarms = Alarms,
+         file_name = ibutFile,
+         notes = IgnoreReason) %>% 
+  select(home_winter_id, date, time, alarms, file_name, notes)
+
+ak_sums_log <- ak_sums_data_log_updated %>% 
+  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
+  arrange(home, date)
+
+write_csv(ak_sums_log, "Output/ak_sums_data_log_updated.csv")
+```
+
+# Initial load, format, and save KidsAIR WMT SUMs data
+## Files here are in several formats: csv (3 different formats) and txt (5 formats)
 ## I had to do some manual cleaning of the files to get them named
@@
          temp_c_sums = temp_c) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums)
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
 
@@
   select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
-  na.exclude()
+  arrange(home, datetime_sums) %>% 
+  distinct()
 
@@
          temp_c_sums = (temp_f-32)/1.8) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums)
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
 
-##### Read in sums csv files (from WMT) #####
+##### Read in sums xlsx files (from WMT) #####
 
 # List files
-list_csv_files <- list.files("Input/WMT/sums_xlsx_files")
+list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip16")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/sums_xlsx_files/header_skip16")
+initial_data = tibble(file_list = list_xlsx_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx, skip = 16)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_xlsx <- initial_data %>% 
+  mutate(datetime = ymd_hms(`Log data: date/time`),
+         temp_c = as.numeric(`Temperature °C`),
+         area = "WMT",
+         home2 = "WMA") %>% 
+  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = datetime,
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c,
+         home = as.factor(home)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip21")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/sums_xlsx_files/header_skip21")
+initial_data = tibble(file_list = list_xlsx_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx, skip = 21)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_xlsx2 <- initial_data %>% 
+  mutate(datetime = ymd_hms(`Log data: date/time`),
+         temp_c = as.numeric(`Temperature °C`),
+         area = "WMT",
+         home2 = "WMA") %>% 
+  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = datetime,
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c,
+         home = as.factor(home)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_2col_c")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/sums_xlsx_files/no_header_2col_c")
+initial_data = tibble(file_list = list_xlsx_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_xlsx3 <- initial_data %>% 
+  mutate(datetime = ymd_hms(datetime),
+         temp_c = as.numeric(temp_c),
+         area = "WMT",
+         home2 = "WMA") %>% 
+  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = datetime,
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c,
+         home = as.factor(home)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_c")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/sums_xlsx_files/no_header_4col_c")
+initial_data = tibble(file_list = list_xlsx_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_xlsx4 <- initial_data %>% 
+  mutate(temp_c = as.numeric(temp_c),
+         area = "WMT",
+         home2 = "WMA") %>% 
+  separate("time", c("trash", "time"), sep = " ") %>%
+  unite("time", c("time", "am_pm"), sep = " ") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = ymd_hms(datetime),
+         date_sums = ymd(date),
+         time_sums = time,
+         temp_c_sums = temp_c,
+         home = as.factor(home)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_f")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/WMT/sums_xlsx_files/no_header_4col_f")
+initial_data = tibble(file_list = list_xlsx_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+wmt_sums_xlsx5 <- initial_data %>% 
+  mutate(temp_f = as.numeric(temp_f),
+         area = "WMT",
+         home2 = "WMA") %>% 
+  separate("time", c("trash", "time"), sep = " ") %>%
+  unite("time", c("time", "am_pm"), sep = " ") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  unite("home", c("home2", "home"), sep = "") %>% 
+  mutate(datetime_sums = ymd_hms(datetime),
+         date_sums = ymd(date),
+         time_sums = time,
+         temp_c_sums = (temp_f-32)/1.8,
+         home = as.factor(home)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# Bind all WMT SUMs data together
+wmt_sums_raw <- rbind(wmt_sums_csv, wmt_sums_csv_2, wmt_sums_csv_3,
+                      wmt_sums_xlsx, wmt_sums_xlsx2, wmt_sums_xlsx3,
+                      wmt_sums_xlsx4, wmt_sums_xlsx5) %>% 
+  select(-datetime_sums) %>% 
+  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  filter(!is.na(date_sums)) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+write_rds(wmt_sums_raw, "Output/wmt_sums_raw.rds")
+```
+
+# Initial load, format, and save KidsAIR NN SUMs data
+```{r}
+##### Read in sums csv files (from NN) #####
+
+# List files
+list_csv_files <- list.files("Input/NN/sums_csv_new")
 
@@
 ## run next 12 lines together
-setwd("Input/WMT/sums_xlsx_files")
+setwd("Input/NN/sums_csv_new")
 initial_data = tibble(file_list = list_csv_files) %>%
   extract(file_list, "home", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv, skip = 1,
-                            col_names = c("line", "date", "time", "temp_c", "trash"),
+  mutate(save_data = lapply(file_list, read_csv,
+                            col_names = c("line", "date", "time", "temp_c"),
                             col_types = cols(line = col_double(),
@@
                                              time = col_time(format = ""),
-                                             temp_c = col_character(),
+                                             temp_c = col_character()))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+nn_sums_csv <- initial_data %>% 
+  mutate(area = "NN") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = datetime,
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_csv_files_2 <- list.files("Input/NN/sums_csv_old/2col-c")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 7 lines together
+setwd("Input/NN/sums_csv_old/2col-c")
+initial_data = tibble(file_list = list_csv_files_2) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv)) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+nn_sums_csv_2 <- initial_data %>% 
+  mutate(area = "NN",
+         datetime_sums = mdy_hm(datetime)) %>% 
+  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>%
+  mutate(date_sums = ymd(date_sums),
+         temp_c_sums = as.numeric(temp_c)) %>%
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 12 lines together
+setwd("Input/NN/sums_csv_old/3col-f")
+initial_data = tibble(file_list = list_csv_files_3) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 1,
+                            col_names = c("date", "time", "temp_f"),
+                            col_types = cols(date = col_character(),
+                                             time = col_time(format = ""),
+                                             temp_f = col_double()))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+nn_sums_csv_3 <- initial_data %>% 
+  mutate(area = "NN") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = mdy_hms(datetime),
+         date_sums = mdy(date),
+         time_sums = time,
+         temp_c_sums = (temp_f-32)/1.8) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# Bind all WMT SUMs data together
+nn_sums_raw <- rbind(nn_sums_csv, nn_sums_csv_2, nn_sums_csv_3) %>% 
+  select(-datetime_sums) %>% 
+  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  filter(!is.na(date_sums)) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+write_rds(nn_sums_raw, "Output/nn_sums_raw.rds")
+```
+
+# Initial load, format, and save KidsAIR AK SUMs data
+```{r}
+##### Read in sums csv files (from AK) #####
+
+# List files
+list_csv_files <- list.files("Input/AK/sums_csv_new")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 12 lines together
+setwd("Input/AK/sums_csv_new")
+initial_data = tibble(file_list = list_csv_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv,
+                            col_names = c("line", "date", "time", "temp_c"),
+                            col_types = cols(line = col_double(),
+                                             date = col_date(format = ""),
+                                             time = col_time(format = ""),
+                                             temp_c = col_double()))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+ak_sums_csv <- initial_data %>% 
+  mutate(area = "AK") %>% 
+  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>%
+  mutate(datetime_sums = ymd_hms(datetime),
+         date_sums = date,
+         time_sums = time,
+         temp_c_sums = temp_c) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_csv_files_2 <- list.files("Input/AK/sums_csv_old")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 7 lines together
+setwd("Input/AK/sums_csv_old")
+initial_data = tibble(file_list = list_csv_files_2) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_csv, skip = 1,
+                            col_names = c("line", "date", "time", "temp_f", "trash"),
+                            col_types = cols(line = col_double(),
+                                             date = col_character(),
+                                             time = col_time(format = ""),
+                                             temp_f = col_double(),
                                              trash = col_character())))  %>%
@@
 # Format combined file from above; combine with other PM data below
-wmt_sums_csv <- initial_data %>% 
-  mutate(area = "WMT",
-         home2 = "WMA") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  unite("home", c("home2", "home"), sep = "") %>% 
-  mutate(datetime_sums = datetime,
-         date_sums = date,
+ak_sums_csv_2 <- initial_data %>% 
+  mutate(area = "AK") %>% 
+  mutate(date_sums = mdy(date),
          time_sums = time,
-         temp_c_sums = temp_c) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums)
+         temp_c_sums = (temp_f-32)/1.8) %>% 
+  unite("datetime", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = ymd_hms(datetime)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+##### Read in sums xlsx files (from AK) #####
+
+# List files
+list_xlsx_files <- list.files("Input/AK/sums_xlsx_temp_c")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/AK/sums_xlsx_temp_c")
+initial_data = tibble(file_list = list_xlsx_files) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx,
+                            col_names = c("date", "time", "am_pm", 
+                                          "trash", "temp_c"))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+ak_sums_xlsx <- initial_data %>% 
+  mutate(area = "AK") %>% 
+  separate("time", c("trash", "time"), sep = " ", remove = FALSE) %>% 
+  unite("time", c("time", "am_pm"), sep = " ") %>% 
+  mutate(date_sums = ymd(date),
+         time_sums = time,
+         temp_c_sums = as.numeric(temp_c),
+         home = as.factor(home)) %>% 
+  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# List files
+list_xlsx_files2 <- list.files("Input/AK/sums_xlsx_temp_f")
+
+# Set working directory and load files in list; extract file name and add as column
+## run next 6 lines together
+setwd("Input/AK/sums_xlsx_temp_f")
+initial_data = tibble(file_list = list_xlsx_files2) %>%
+  extract(file_list, "home", remove = FALSE) %>%
+  mutate(save_data = lapply(file_list, read_xlsx,
+                            col_names = c("date", "time", "am_pm", 
+                                          "trash", "temp_f"))) %>%
+  unnest(save_data) %>%
+  select(-file_list)
+
+# Format combined file from above; combine with other PM data below
+ak_sums_xlsx2 <- initial_data %>% 
+  mutate(area = "AK") %>% 
+  separate("time", c("trash", "time"), sep = " ", remove = FALSE) %>% 
+  unite("time", c("time", "am_pm"), sep = " ") %>% 
+  mutate(date_sums = ymd(date),
+         time_sums = time,
+         temp_c_sums = as.numeric((temp_f-32)/1.8),
+         home = as.factor(home)) %>% 
+  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+# Bind all WMT SUMs data together
+ak_sums_raw <- rbind(ak_sums_csv, ak_sums_csv_2, 
+                     ak_sums_xlsx, ak_sums_xlsx2) %>% 
+  select(-datetime_sums) %>% 
+  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  filter(!is.na(date_sums)) %>% 
+  arrange(home, datetime_sums) %>% 
+  distinct()
+
+write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
 ```

### 2020-01-03 sums_data_cleaning.Rmd
Clean SUMs data

@@
 ---
-title: "KidsAIR initial PM data work"
+title: "KidsAIR initial SUMs data work"
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 27 Dec 2019"
+date: "Started 23 Dec 2019, Updated 3 Jan 2020"
 output: pdf_document
@@
          ibid = iButtonID,
-         date = InstallDate,
-         time = InstallTime) %>% 
-  select(home_winter_id, ibid, date, time)
+         date_log = InstallDate,
+         time_log = InstallTime) %>% 
+  select(home_winter_id, ibid, date_log, time_log)
 
@@
   filter(area == "WMT") %>% 
-  select(home_winter_id, home, home_id)
+  select(home_winter_id, winter_id, home, home_id)
 
@@
   left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, ibid) %>% 
-  arrange(home, date)
+  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
+  arrange(home, date_log) %>% 
+  mutate(date_log = mdy(date_log))
 
+write_rds(wmt_sums_log, "Output/wmt_sums_data_log.rds")
 write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")
@@
   filter(area == "NN") %>% 
-  select(home_winter_id, home, home_id)
+  select(home_winter_id, winter_id, home, home_id)
 
@@
   left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, ibid) %>% 
-  arrange(home, date)
+  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
+  arrange(home, date_log) %>% 
+  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
+  select(-trash)
 
+write_rds(nn_sums_log, "Output/nn_sums_data_log.rds")
 write_csv(nn_sums_log, "Output/nn_sums_data_log.csv")
@@
          ibid = iButtonID,
-         date = InstallDate,
-         time = InstallTime) %>% 
-  select(home_winter_id, ibid, date, time)
+         date_log = InstallDate,
+         time_log = InstallTime) %>% 
+  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
+  select(home_winter_id, ibid, date_log, time_log)
 
@@
   filter(area == "AK") %>% 
-  select(home_winter_id, home, home_id)
+  select(home_winter_id, winter_id, home, home_id)
 
@@
   left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, ibid) %>% 
-  arrange(home, date)
+  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
+  arrange(home, date_log)
 
 write_csv(ak_sums_log, "Output/ak_sums_data_log.csv")
+write_rds(ak_sums_log, "Output/ak_sums_data_log.rds")
 
@@
 
-# Bind all WMT SUMs data together
+# Bind all AK SUMs data together
 ak_sums_raw <- rbind(ak_sums_csv, ak_sums_csv_2, 
@@
 
-write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
+# write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
 ```
 
-# Code from PM cleaning
+# Bind all SUMs files together
 ```{r}
-##### Read in dusttrack .csv files (all areas) #####
+wmt_sums_raw <- read_rds("Output/wmt_sums_raw.rds")
 
-### WMT ###
-# List files
-list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")
+nn_sums_raw <- read_rds("Output/nn_sums_raw.rds")
 
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
+ak_sums_raw <- read_rds("Output/ak_sums_raw.rds")
 
-# Format combined file from above; combine with other PM data below
-wmt_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "WMT") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
+all_sums_raw <- rbind(wmt_sums_raw, nn_sums_raw, ak_sums_raw)
 
-### NN ###
-# List files
-list_csv_files <- list.files("Input/NN/dusttrack_csv_files")
+# write_rds(all_sums_raw, "Output/all_sums_raw.rds")
+```
 
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/NN/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
+# Check sample start times vs log start times
+```{r}
+# Load all SUMs data
+all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_sums)
 
-# Format combined file from above; combine with other PM data below
-nn_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "NN") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
+# Load SUMs data logs
+wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
+nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
+ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
 
-### AK ###
-# List files
-list_csv_files <- list.files("Input/AK/dusttrack_csv_files")
+# Combine SUMs logs from all areas
+all_sums_log <- rbind(wmt_sums_log, nn_sums_log, ak_sums_log) %>% 
+  mutate(winter_id = as.character(winter_id))
 
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/AK/dusttrack_csv_files")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
+# Combine SUMs logs and data
+## Get ready to compare start times from logs and data files
+sums_time_comparison <- all_sums_raw %>%
+  group_by(area, home) %>% 
+  mutate(datetime_initial = first(datetime_sums)) %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums),
+         datetime_initial = ymd_hms(datetime_initial),
+         winter_id = if_else((datetime_sums - datetime_initial) > 20736000,
+                             2, 1), 
+         winter_id = as.character(winter_id)) %>% 
+  ungroup() %>% 
+  left_join(all_sums_log, by = c("home", "winter_id")) %>% 
+  group_by(home, winter_id) %>% 
+  unite(datetime_log, c("date_log", "time_log"), 
+        sep = " ", remove = FALSE) %>% 
+  mutate(datetime_log = ymd_hms(datetime_log),
+         datetime_initial = first(datetime_sums),
+         time_diff_sums = datetime_initial - datetime_log) %>% 
+  distinct(time_diff_sums, .keep_all = TRUE) %>%
+  ungroup() %>% 
+  arrange(home, winter_id)
 
-# Format combined file from above; combine with other PM data below
-ak_pm_csv <- initial_data %>% 
-  mutate(date = mdy(Date),
-         time = Time,
-         pm = as.numeric(AEROSOL),
-         area = "AK") %>% 
-  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  filter(!is.na(pm))
-
-##### Read in other dusttrack .csv files (WMT and NN) #####
-
-### WMT ### 
-# List files
-list_other_files <- list.files("Input/WMT/dusttrack_other_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/WMT/dusttrack_other_files")
-initial_data = tibble(file_list = list_other_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-wmt_pm_other <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(home_winter_id, trash1, trash2) %>% 
-  group_by(home_winter_id) %>% 
-  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash2, "NA"),
-         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash1, "NA"),
-         trash1 = as.numeric(trash1),
-         trash1 = as.character(trash1),
-         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
-  filter(!is.na(trash5)) %>% 
-  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
-  spread(trash4, trash3) %>% 
-  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
-  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
-  mutate(datetime = lead(datetime)) %>% 
-  replace_with_na(replace = list(datetime = "NA NA")) %>% 
-  mutate(datetime2 = first(datetime),
-         datetime = mdy_hms(datetime2),
-         trash1 = as.numeric(trash1)) %>% 
-  filter(!is.na(trash1)) %>% 
-  mutate(datetime2 = datetime + seconds(trash1),
-         datetime2 = ymd_hms(datetime2)) %>% 
-  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = datetime2,
-         area = "WMT") %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  ungroup()
-
-### NN ### 
-# List files
-list_other_files <- list.files("Input/NN/dusttrack_other_files")
-
-# Set working directory and load files in list; extract file name and add as column
-## run next 6 lines together
-setwd("Input/NN/dusttrack_other_files")
-initial_data = tibble(file_list = list_other_files) %>%
-  extract(file_list, "home_winter_id", remove = FALSE) %>%
-  mutate(save_data = lapply(file_list, read_csv)) %>%
-  unnest(save_data) %>%
-  select(-file_list)
-
-# Format combined file from above; combine with other PM data below
-nn_pm_other <- initial_data %>% 
-  mutate(trash1 = `Instrument Name`,
-         trash2 = `DustTrak II`) %>% 
-  select(home_winter_id, trash1, trash2) %>% 
-  filter(!is.na(trash1)) %>% 
-  group_by(home_winter_id) %>% 
-  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash1, "NA"),
-         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
-                              trash2, "NA"),
-         trash1 = as.numeric(trash1),
-         trash1 = as.character(trash1),
-         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
-  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
-  filter(!is.na(trash5)) %>% 
-  spread(trash3, trash4) %>% 
-  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
-  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
-  mutate(datetime = lead(datetime)) %>% 
-  replace_with_na(replace = list(datetime = "NA NA")) %>% 
-  mutate(datetime2 = first(datetime),
-         datetime = mdy_hms(datetime2),
-         trash1 = as.numeric(trash1)) %>% 
-  filter(!is.na(trash1)) %>% 
-  mutate(datetime2 = datetime + seconds(trash1),
-         datetime2 = ymd_hms(datetime2)) %>% 
-  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
-  mutate(pm = as.numeric(trash2),
-         datetime = datetime2,
-         area = "NN") %>% 
-  select(home_winter_id, datetime, date, time, pm, area) %>% 
-  ungroup() %>% 
-  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
-                                  "22", home_winter_id),
-         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
-                                  "55", home_winter_id),
-         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
-                                  "81", home_winter_id))
-
-##### Bind all files from above #####
-kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
-                nn_pm_csv, nn_pm_other,
-                ak_pm_csv)
-
-# Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
-#write_csv(kids_pm, "Output/kids_pm_full.csv")
+write_csv(sums_time_comparison, "Output/sums_time_comparison.csv")
+write_rds(sums_time_comparison, "Output/sums_time_comparison.rds")
 ```

### 2020-01-07 sums_data_cleaning.Rmd
SUMs cleaning

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 3 Jan 2020"
+date: "Started 23 Dec 2019, Updated 6 Jan 2020"
 output: pdf_document
@@
 ```
+
+# Compile start and stop times for SUMs based on field logs
+```{r}
+# Load all SUMs data
+all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_sums)
+
+# Load and format SUMs data logs for WMT
+wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
+
+wmt_sums_stop <- read_csv("Input/WMT/wmt_medicalvisit_20200107.csv") %>% 
+  mutate(home_winter_id = HomeWinterID,
+         final_visit_date = if_else(VisitDate == "11/4/2025", "11/4/2015", VisitDate),
+         final_visit_date = mdy(final_visit_date)) %>% 
+  group_by(home_winter_id) %>% 
+  arrange(final_visit_date) %>% 
+  mutate(date_last = last(final_visit_date),
+         date_filter = if_else(date_last == final_visit_date, 1, 0)) %>% 
+  ungroup() %>% 
+  filter(date_filter == 1) %>% 
+  select(home_winter_id, final_visit_date) %>% 
+  distinct(.keep_all = TRUE) %>% 
+  arrange(home_winter_id) %>% 
+  mutate(time = "01:00:00") %>% 
+  unite(datetime_stop, c("final_visit_date", "time"), sep = " ") %>% 
+  mutate(datetime_stop = ymd_hms(datetime_stop))
+
+wmt_sums_start <- read_xlsx("Input/WMT/wmt_sums_start_times_20200107.xlsx") %>% 
+  mutate(datetime_start = ymd_hms(datetime_start))
+
+wmt_sums_start_stop <- wmt_sums_start %>% 
+  left_join(wmt_sums_stop, by = "home_winter_id")
+
+wmt_sums_raw <- all_sums_raw %>% 
+  filter(area == "WMT") %>% 
+  left_join(wmt_sums_start_stop, by = c("area", "home")) %>% 
+  mutate(filter_var = 
+         if_else(datetime_sums > datetime_start & datetime_sums < datetime_stop,
+                 1, 0))
+
+wmt_sums_filtered <- wmt_sums_raw %>% 
+  filter(filter_var == 1) %>% 
+  arrange(area, home, winter_id, datetime_sums)
+
+# Start here next:
+## Make sure datetime_sums lines up with start/stop range
+## Work on cleaning up temp_c_sums
+## Ask about NN/AK finish dates and move on to repeating steps for those homes
+write_rds(wmt_sums_filtered, "Output/wmt_sums_filtered.rds")
+
+nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
+# Need to get the NN medicalvisit file once data collection is finished
+nn_sums_stop <- read_csv("Input/NN/nn_medicalvisit_20200107.csv") %>% 
+  mutate(home_winter_id = HomeWinterID,
+         final_visit_date = if_else(VisitDate == "11/4/2025", "11/4/2015", VisitDate),
+         final_visit_date = mdy(final_visit_date)) %>% 
+  group_by(home_winter_id) %>% 
+  arrange(final_visit_date) %>% 
+  mutate(date_last = last(final_visit_date),
+         date_filter = if_else(date_last == final_visit_date, 1, 0)) %>% 
+  ungroup() %>% 
+  filter(date_filter == 1) %>% 
+  select(home_winter_id, final_visit_date) %>% 
+  distinct(.keep_all = TRUE) %>% 
+  arrange(home_winter_id)
+
+ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
+# Need to get the AK medicalvisit file once data collection is finished
+ak_sums_stop <- read_csv("Input/AK/ak_medicalvisit_20200107.csv") %>% 
+  mutate(home_winter_id = HomeWinterID,
+         final_visit_date = if_else(VisitDate == "11/4/2025", "11/4/2015", VisitDate),
+         final_visit_date = mdy(final_visit_date)) %>% 
+  group_by(home_winter_id) %>% 
+  arrange(final_visit_date) %>% 
+  mutate(date_last = last(final_visit_date),
+         date_filter = if_else(date_last == final_visit_date, 1, 0)) %>% 
+  ungroup() %>% 
+  filter(date_filter == 1) %>% 
+  select(home_winter_id, final_visit_date) %>% 
+  distinct(.keep_all = TRUE) %>% 
+  arrange(home_winter_id)
+```

### 2020-01-09 sums_data_cleaning.Rmd
SUMs cleaning

--- 2020-01-03 sums_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 3 Jan 2020"
+date: "Started 23 Dec 2019, Updated 8 Jan 2020"
 output: pdf_document
@@
 ```
+
+# Compile start and stop times for SUMs based on field logs
+```{r}
+# Load all SUMs data
+all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_sums)
+
+# Load and format SUMs data logs for WMT
+wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
+
+wmt_sums_stop <- read_csv("Input/WMT/wmt_medicalvisit_20200107.csv") %>% 
+  mutate(home_winter_id = HomeWinterID,
+         final_visit_date = if_else(VisitDate == "11/4/2025", "11/4/2015", VisitDate),
+         final_visit_date = mdy(final_visit_date)) %>% 
+  group_by(home_winter_id) %>% 
+  arrange(final_visit_date) %>% 
+  mutate(date_last = last(final_visit_date),
+         date_filter = if_else(date_last == final_visit_date, 1, 0)) %>% 
+  ungroup() %>% 
+  filter(date_filter == 1) %>% 
+  select(home_winter_id, final_visit_date) %>% 
+  distinct(.keep_all = TRUE) %>% 
+  arrange(home_winter_id) %>% 
+  mutate(time = "01:00:00") %>% 
+  unite(datetime_stop, c("final_visit_date", "time"), sep = " ") %>% 
+  mutate(datetime_stop = ymd_hms(datetime_stop))
+
+wmt_sums_start <- read_xlsx("Input/WMT/wmt_sums_start_times_20200107.xlsx") %>% 
+  mutate(datetime_start = ymd_hms(datetime_start))
+
+wmt_sums_start_stop <- wmt_sums_start %>% 
+  left_join(wmt_sums_stop, by = "home_winter_id")
+
+wmt_sums_raw <- all_sums_raw %>% 
+  filter(area == "WMT") %>% 
+  left_join(wmt_sums_start_stop, by = c("area", "home")) %>% 
+  mutate(filter_var = 
+         if_else(datetime_sums > datetime_start & datetime_sums < datetime_stop,
+                 1, 0))
+
+wmt_sums_filtered <- wmt_sums_raw %>% 
+  filter(filter_var == 1) %>% 
+  arrange(area, home, winter_id, datetime_sums)
+
+# Clean up this file and combine with other areas
+# write_rds(wmt_sums_filtered, "Output/wmt_sums_filtered.rds")
+
+#####
+nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
+
+nn_sums_stop <- read_xlsx("Input/NN/nn_medicalvisit_20200108.xlsx") %>% 
+  mutate(home_winter_id = HomeWinterID,
+         final_visit_date = ymd(VisitDate)) %>% 
+  group_by(home_winter_id) %>% 
+  arrange(final_visit_date) %>% 
+  mutate(date_last = last(final_visit_date),
+         date_filter = if_else(date_last == final_visit_date, 1, 0)) %>% 
+  ungroup() %>% 
+  filter(date_filter == 1) %>% 
+  select(home_winter_id, final_visit_date) %>% 
+  distinct(.keep_all = TRUE) %>% 
+  arrange(home_winter_id) %>% 
+  mutate(time = "01:00:00") %>% 
+  unite(datetime_stop, c("final_visit_date", "time"), sep = " ") %>% 
+  mutate(datetime_stop = ymd_hms(datetime_stop))
+
+nn_sums_start <- read_xlsx("Input/NN/nn_sums_start_times_20200108.xlsx") %>% 
+  mutate(datetime_start = ymd_hms(datetime_start))
+
+nn_sums_start_stop <- nn_sums_start %>% 
+  left_join(nn_sums_stop, by = "home_winter_id")
+
+nn_sums_raw <- all_sums_raw %>% 
+  filter(area == "NN") %>% 
+  left_join(nn_sums_start_stop, by = c("area", "home")) %>% 
+  mutate(filter_var = 
+         if_else(datetime_sums > datetime_start & datetime_sums < datetime_stop,
+                 1, 0))
+
+nn_sums_filtered <- nn_sums_raw %>% 
+  filter(filter_var == 1) %>% 
+  arrange(area, home, winter_id, datetime_sums)
+
+# Clean up this file and combine with other areas
+# write_rds(nn_sums_filtered, "Output/nn_sums_filtered.rds")
+
+#####
+ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
+
+ak_sums_stop <- read_xlsx("Input/AK/ak_medicalvisit_20200108.xlsx") %>% 
+  mutate(home_winter_id = HomeWinterID,
+         final_visit_date = ymd(VisitDate)) %>% 
+  group_by(home_winter_id) %>% 
+  arrange(final_visit_date) %>% 
+  mutate(date_last = last(final_visit_date),
+         date_filter = if_else(date_last == final_visit_date, 1, 0)) %>% 
+  ungroup() %>% 
+  filter(date_filter == 1) %>% 
+  select(home_winter_id, final_visit_date) %>% 
+  distinct(.keep_all = TRUE) %>% 
+  arrange(home_winter_id) %>% 
+  mutate(time = "01:00:00") %>% 
+  unite(datetime_stop, c("final_visit_date", "time"), sep = " ") %>% 
+  mutate(datetime_stop = ymd_hms(datetime_stop))
+
+ak_sums_start <- read_xlsx("Input/AK/ak_sums_start_times_20200108.xlsx") %>% 
+  mutate(datetime_start = ymd_hms(datetime_start))
+
+ak_sums_start_stop <- ak_sums_start %>% 
+  left_join(ak_sums_stop, by = "home_winter_id")
+
+ak_sums_raw <- all_sums_raw %>% 
+  filter(area == "AK") %>% 
+  left_join(ak_sums_start_stop, by = c("area", "home")) %>% 
+  mutate(filter_var = 
+         if_else(datetime_sums > datetime_start & datetime_sums < datetime_stop,
+                 1, 0))
+
+ak_sums_filtered <- ak_sums_raw %>% 
+  filter(filter_var == 1) %>% 
+  arrange(area, home, winter_id, datetime_sums)
+
+# Clean up this file and combine with other areas
+# fix doubled observations
+# write_rds(ak_sums_filtered, "Output/wmt_sums_filtered.rds")
+```

### 2020-01-09 sums_data_cleaning.Rmd
SUMs cleaning and summary stats

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 8 Jan 2020"
+date: "Started 23 Dec 2019, Updated 9 Jan 2020"
 output: pdf_document
@@
          time_sums = time,
-         temp_c_sums = temp_c) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+         temp_c_sums = temp_c,
+         file_type = "logtag") %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
-  mutate(date_sums = ymd(date_sums)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  mutate(date_sums = ymd(date_sums),
+         file_type = "temp_c") %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
          time_sums = time,
-         temp_c_sums = (temp_f-32)/1.8) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+         temp_c_sums = (temp_f-32)/1.8,
+         file_type = "temp_f") %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
          temp_c_sums = temp_c,
-         home = as.factor(home)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+         home = as.factor(home),
+         file_type = "xlsx_c") %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
          temp_c_sums = (temp_f-32)/1.8,
-         home = as.factor(home)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+         home = as.factor(home),
+         file_type = "xlsx_f") %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
+                                                  "temp_f", "xlsx_c",
+                                                  "xlsx_f")))
 
-write_rds(wmt_sums_raw, "Output/wmt_sums_raw.rds")
+# This filtering will keep only the first instance when there are multiple
+# observations for a specific datetime_sums
+wmt_sums_filter <- wmt_sums_raw %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(home, datetime_sums)
+
+write_rds(wmt_sums_filter, "Output/wmt_sums_raw.rds")
 ```
@@
   mutate(date_sums = ymd(date_sums),
-         temp_c_sums = as.numeric(temp_c)) %>%
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+         temp_c_sums = as.numeric(temp_c),
+         file_type = "temp_c") %>%
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
+                                                  "temp_f")))
 
-write_rds(nn_sums_raw, "Output/nn_sums_raw.rds")
+# This filtering will keep only the first instance when there are multiple
+# observations for a specific datetime_sums
+nn_sums_filter <- nn_sums_raw %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(home, datetime_sums)
+
+write_rds(nn_sums_filter, "Output/nn_sums_raw.rds")
 ```
@@
          time_sums = time,
-         temp_c_sums = (temp_f-32)/1.8) %>% 
+         temp_c_sums = (temp_f-32)/1.8,
+         file_type = "temp_f") %>% 
   unite("datetime", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = ymd_hms(datetime)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
          temp_c_sums = as.numeric(temp_c),
-         home = as.factor(home)) %>% 
+         home = as.factor(home),
+         file_type = "xlsx_c") %>% 
   unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
          temp_c_sums = as.numeric((temp_f-32)/1.8),
-         home = as.factor(home)) %>% 
+         home = as.factor(home),
+         file_type = "xlsx_f") %>% 
   unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
   arrange(home, datetime_sums) %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  mutate(file_type = factor(file_type, levels = c("logtag", "temp_f",
+                                                  "xlsx_c", "xlsx_f")))
 
-# write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
+# This filtering will keep only the first instance when there are multiple
+# observations for a specific datetime_sums
+ak_sums_filter <- ak_sums_raw %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(home, datetime_sums)
+
+# write_rds(ak_sums_filter, "Output/ak_sums_raw.rds")
 ```
@@
 # Load and format SUMs data logs for WMT
-wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
+#wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
 
@@
 
-# Clean up this file and combine with other areas
 # write_rds(wmt_sums_filtered, "Output/wmt_sums_filtered.rds")
@@
 #####
-nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
+#nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
 
@@
 
-# Clean up this file and combine with other areas
 # write_rds(nn_sums_filtered, "Output/nn_sums_filtered.rds")
@@
 #####
-ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
+#ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
 
@@
 
-# Clean up this file and combine with other areas
-# fix doubled observations
-# write_rds(ak_sums_filtered, "Output/wmt_sums_filtered.rds")
+# write_rds(ak_sums_filtered, "Output/ak_sums_filtered.rds")
 ```
+
+# Checking to make sure multiple observations per datetime_sums were removed
+# Number of obs in each dataset below should match # obs in check_dates
+```{r}
+wmt_sums_filtered <- read_rds("Output/wmt_sums_filtered.rds")
+nn_sums_filtered <- read_rds("Output/nn_sums_filtered.rds")
+ak_sums_filtered <- read_rds("Output/ak_sums_filtered.rds")
+
+check_dates <- nn_sums_filtered %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE)
+```
+
+# Combine filtered data from each area
+```{r}
+wmt_sums_filtered <- read_rds("Output/wmt_sums_filtered.rds")
+nn_sums_filtered <- read_rds("Output/nn_sums_filtered.rds")
+ak_sums_filtered <- read_rds("Output/ak_sums_filtered.rds")
+
+all_sums_filtered <- rbind(wmt_sums_filtered, nn_sums_filtered, ak_sums_filtered)
+
+# write_rds(all_sums_filtered, "Output/all_sums_filtered.rds")
+```
+
+# Continue cleaning up SUMs data
+```{r}
+all_sums_filtered <- read_rds("Output/all_sums_filtered.rds")
+
+sums_clean <- all_sums_filtered %>% 
+   mutate(temp_c_sums = as.numeric(temp_c_sums),
+          winter_id = as.factor(winter_id),
+          home_winter_id = as.factor(home_winter_id),
+          home_id = as.factor(home_id),
+          day_of_week = weekdays(date_sums),
+          month_sums = as.factor(month(date_sums, label = TRUE, abbr = TRUE))) %>% 
+  select(-filter_var) %>% 
+  arrange(area, home, winter_id, datetime_sums)
+
+# write_rds(sums_clean, "Output/sums_clean.rds")
+```
+
+# Plot temp over time for individual home to check for trends visually
+```{r}
+sums_clean <- read_rds("Output/sums_clean.rds")
+
+sums_time_trends_plot <- sums_clean %>% 
+  filter(home == "WMA102") %>% 
+  ggplot() + 
+    geom_point(aes(datetime_sums, temp_c_sums, color = month_sums), size = 1.5) +
+    theme_classic() +
+    labs(y = "Temperature (degrees C)",
+         x = "Datetime",
+         color = "Month of Year") +
+    theme(axis.title.y = element_text(size = 12,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_text(size = 12),
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 12, color = "black"),
+          axis.line.x = element_line(colour = "black", size = 1), 
+          axis.line.y = element_line(colour = "black", size = 1), 
+          axis.ticks = element_blank()) +
+    scale_color_manual(values = jv_palette)
+sums_time_trends_plot
+```

### 2020-01-09 sums_summary.Rmd
SUMs cleaning and summary stats

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 9 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
sums_clean <- read_rds("Output/sums_clean.rds")
```

# Summary stats
```{r}
sums_summary <- sums_clean %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean Temp" = mean(temp_c_sums, na.rm = TRUE), 
            "SD Temp" = sd(temp_c_sums, na.rm = TRUE),
            "Min Temp" = min(temp_c_sums, na.rm = TRUE), 
            "Median Temp" = median(temp_c_sums, na.rm = TRUE),
            "Max Temp" = max(temp_c_sums, na.rm = TRUE))
```

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=5}
boxplots_kids_sums <- sums_clean %>%
  ggplot(aes(area, temp_c_sums)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(-50, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = "Temperature (degrees C)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_sums
```

### 2020-01-10 dustrak_data_cleaning.Rmd
SUMs cleaning updates. Add extensive notes to PM and SUMs code.

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 20 Dec 2019"
+date: "Started 19 Nov 2019, Updated 10 Jan 2020"
 output: pdf_document
@@
 
-# Initial load, format, and save KidsAIR Dusttrak data
-## Files here are in 3 formats: csv (2 different formats) and txt
-## I had to do some manual cleaning of the files to get them named
-## consistently and in reasonably similar formats to be able to load
-## into R in batches, as seen below. The files I upload in the code below
-## will also be uploaded to Box in a folder with my name.
-## Box folder directory: 
+# Initial load, format, and save KidsAIR DustTrak (PM) data
+## Files were downloaded from Box: KidsAIR > 2019-Nov-Data > WMT > uploads > Dusttrack
+## Files were not named consistently, so I renamed them by their home ID number.
+## Files also had several different types and formats.
+## Manual cleaning of some files had to be done to get them in 
+## consistent formats and file types to work with in R.
+## Unfortunately, this process had to be manual due to the large number of files
+## and inconsistent QC process across the various files.
+## As a result, this process will be difficult to replicate if needed;
+## The code below will highlight how the files were formatted and grouped together.
+## Once files were in consistent formats, I load them to R in batches, as seen below.
+## The following code cleans and combines all PM data
 ```{r}
@@
 # List files
+## Numerous files had this format with a 28 row header
+## 3 columns: date, time, PM
 list_txt_files <- list.files("Input/WMT/dusttrack_txt_files")
@@
 # List files
+## Numerous files had this format with no header
+## 3 columns: date, time, PM
 list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")
@@
 # List files
+## Numerous files had this format with no header
+## 3 columns: date, time, PM
 list_csv_files <- list.files("Input/NN/dusttrack_csv_files")
@@
 # List files
+## Numerous files had this format with no header
+## 3 columns: date, time, PM
 list_csv_files <- list.files("Input/AK/dusttrack_csv_files")
@@
 # List files
+## 4 files had this format with a 19 row header
+## 4 columns: elapsed time (start time extracted from header), PM, notes, notes
 list_other_files <- list.files("Input/WMT/dusttrack_other_files")
@@
 # List files
+## Numerous files had this format with a 19 row header
+## 4 columns: elapsed time (start time extracted from header), PM, notes, notes
 list_other_files <- list.files("Input/NN/dusttrack_other_files")
@@
 # Save pm dataset as RDS and CSV
-write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
-#write_csv(kids_pm, "Output/kids_pm_full.csv")
+# write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
+# write_csv(kids_pm, "Output/kids_pm_full.csv")
 ```
@@
 
-##### Add to new variables to Kids PM data #####
+##### Add new variables to Kids PM data #####
 kids_pm_new <- kids_pm_joined %>%  
@@
 # Save pm dataset as RDS and CSV
-#write_rds(kids_pm_new, "Output/kids_pm_new.rds")
-#write_csv(kids_pm_new, "Output/kids_pm_new.csv")
+# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
 ```
@@
 # Run function from above to look at results
+## Function run for each household to check for odd trends/gaps in the data
+## Notes from this process on uploaded to Box:
+### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
 ```{r}
@@
   arrange(area, home_winter_id)
-#write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
+# write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
 ```
@@
 # Loading and combining notes from time and PM logs
+## Notes from the process of comparing actual vs expect start times 
+## on uploaded to Box:
+### KidsAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
 ```{r}
@@
   left_join(kids_ids, by = c("area", "home_winter_id"))
-#write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
+
+# write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
 ```
@@
 # Fix dusttrak sample start times
+## Notes from above were used to make the following changes in the data
 ```{r}
@@
   summarise(n()/2241695)
+
 # write_rds(kids_pm_new, "Output/kids_pm_new.rds")

### 2020-01-10 sums_data_cleaning.Rmd
SUMs cleaning updates. Add extensive notes to PM and SUMs code.

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 9 Jan 2020"
+date: "Started 23 Dec 2019, Updated 10 Jan 2020"
 output: pdf_document
@@
 # Add linked IDs to WMT SUMs data logs
+## Use these throughout the data cleaning process to check start times
+## Files and notes can be found on Box: 
+### NoonanGroupData > KidsAIR > Ethan - data cleaning and analysis
 ```{r}
+# Input file from Box: ibutton.xlsx
 wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>%
@@
 
+# Input file from Box: ibuttonupd.xlsx
 wmt_sums_data_log_updated <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
@@
 # Add linked IDs to NN SUMs data logs
+## Use these throughout the data cleaning process to check start times
+## Files and notes can be found on Box: 
+### NoonanGroupData > KidsAIR > Ethan - data cleaning and analysis
 ```{r}
+# Input file from Box: ibutton.xlsx
 nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
@@
 
+# Input file from Box: ibuttonupd.xlsx
 nn_sums_data_log_updated <- read_xlsx("Input/NN/nn_sums_data_log_updated.xlsx") %>% 
@@
 # Add linked IDs to AK SUMs data logs
+## Use these throughout the data cleaning process to check start times
+## Files and notes can be found on Box: 
+### NoonanGroupData > KidsAIR > Ethan - data cleaning and analysis
 ```{r}
+# Input file from Box: ibutton.xlsx
 ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
@@
 
+# Input file from Box: ibutton.xlsx
 ak_sums_data_log_updated <- read_xlsx("Input/AK/ak_sums_data_log_updated.xlsx") %>% 
@@
 # Initial load, format, and save KidsAIR WMT SUMs data
-## Files here are in several formats: csv (3 different formats) and txt (5 formats)
-## I had to do some manual cleaning of the files to get them named
-## consistently and in reasonably similar formats to be able to load
-## into R in batches, as seen below. The files I upload in the code below
-## will also be uploaded to Box in a folder with my name.
-## Box folder directory: 
+## Files were downloaded from Box: KidsAIR > 2019-Nov-Data > WMT > uploads > IButton
+## Files were not named consistently, so I renamed them by their home ID number.
+## Files also had several different types and formats.
+## Manual cleaning of some files had to be done to get them in 
+## consistent formats and file types to work with in R.
+## Unfortunately, this process had to be manual due to the large number of files
+## and inconsistent QC process across the various files.
+## As a result, this process will be difficult to replicate if needed;
+## The code below will highlight how the files were formatted and grouped together.
+## Once files were in consistent formats, I load them to R in batches, as seen below.
+## The following code cleans and combines all SUMs temperature data
+## Notes and cleaned data files are uploaded to Box:
+### KidsAIR > Ethan - data cleaning and analysis
 ```{r}
@@
 # List files
+## These files were directly exported by me from the Logtag software.
+## I am confident in their formatting and content since I pulled the data
+## from the original file type of the field instrument.
+## The files have 5 columns: line, date, time, temp (C), notes
 list_csv_files <- list.files("Input/WMT/sums_csv_new")
@@
 # Set working directory and load files in list; extract file name and add as column
-## run next 12 lines together
-setwd("Input/WMT/sums_csv_new")
-initial_data = tibble(file_list = list_csv_files) %>%
-  extract(file_list, "home", remove = FALSE) %>%
+## run next 13 lines together
+setwd("Input/WMT/sums_csv_new") # set new working directory
+initial_data = tibble(file_list = list_csv_files) %>% # use list of files from above
+  extract(file_list, "home", remove = FALSE) %>% # extract file name, name "home"
+  # apply read_csv function; name and format columns
   mutate(save_data = lapply(file_list, read_csv, skip = 1, 
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_csv <- initial_data %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
+  filter(temp_c != "Paused") %>% 
+  # separate temp to remove instances of ">" in the temp data
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = datetime,
@@
          time_sums = time,
-         temp_c_sums = temp_c,
+         temp_c = if_else(is.na(temp_c), trash2, temp_c),
+         temp_c_sums = as.numeric(temp_c),
          file_type = "logtag") %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Only 1 file had this format: 
+## a 21 row header, then columns of datetime and temp (C)
 list_csv_files_2 <- list.files("Input/WMT/sums_csv_old/format_1")
@@
 # Set working directory and load files in list; extract file name and add as column
-## run next 7 lines together
+## run next 9 lines together
 setwd("Input/WMT/sums_csv_old/format_1")
@@
   mutate(save_data = lapply(file_list, read_csv, skip = 22,
-                            col_names = c("datetime_sums", "temp_c_sums"))) %>%
+                            col_names = c("datetime_sums", "temp_c"),
+                            col_types = cols(datetime_sums = col_character(),
+                                             temp_c = col_character()))) %>%
   unnest(save_data) %>%
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_csv_2 <- initial_data %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(date_sums = ymd(date_sums),
+         temp_c = if_else(is.na(temp_c), trash2, temp_c),
+         temp_c_sums = as.numeric(temp_c),
          file_type = "temp_c") %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format: 
+## 5 columns: line, date, time, temp (F), notes
 list_csv_files_3 <- list.files("Input/WMT/sums_csv_old/format_2")
@@
                                              time = col_time(format = ""),
-                                             temp_f = col_double(),
+                                             temp_f = col_character(),
                                              trash = col_character()))) %>%
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_csv_3 <- initial_data %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
+  filter(temp_f != "Paused") %>% 
+  separate(temp_f, c("trash2", "temp_f"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = mdy_hms(datetime),
@@
          time_sums = time,
+         temp_f = if_else(is.na(temp_f), trash2, temp_f),
+         temp_f = as.numeric(temp_f),
          temp_c_sums = (temp_f-32)/1.8,
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
@@
 # List files
+## Numerous files had this format with a 16 row header
+## 2 columns: datetime, temp (C)
 list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip16")
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_xlsx <- initial_data %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = datetime,
@@
          time_sums = time,
-         temp_c_sums = temp_c,
+         temp_c = if_else(is.na(temp_c), trash2, temp_c),
+         temp_c_sums = as.numeric(temp_c),
          home = as.factor(home),
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## 6 files had this format with a 21 row header
+## 2 columns: datetime, temp (C)
 list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip21")
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_xlsx2 <- initial_data %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>%
   mutate(datetime_sums = datetime,
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format with no header
+## 2 columns: datetime, temp (C)
 list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_2col_c")
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_xlsx3 <- initial_data %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format with no header
+## 4 columns: date, time, am/pm, temp (C)
 list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_c")
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_xlsx4 <- initial_data %>% 
@@
   unite("home", c("home2", "home"), sep = "") %>% 
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>%
   mutate(datetime_sums = ymd_hms(datetime),
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format with no header
+## 4 columns: date, time, am/pm, temp (F)
 list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_f")
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 wmt_sums_xlsx5 <- initial_data %>% 
@@
   arrange(home, datetime_sums) %>% 
-  distinct() %>% 
+  distinct() %>% # remove duplicate lines that have same info from different files
   mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
@@
 
-write_rds(wmt_sums_filter, "Output/wmt_sums_raw.rds")
+# write_rds(wmt_sums_filter, "Output/wmt_sums_raw.rds")
 ```
@@
 # Initial load, format, and save KidsAIR NN SUMs data
+## This chunk performs the same process that was done previously for WMT files
 ```{r}
@@
 # List files
+## These files were directly exported by me from the Logtag software.
+## I am confident in their formatting and content since I pulled the data
+## from the original file type of the field instrument.
+## The files have 5 columns: line, date, time, temp (C), notes
 list_csv_files <- list.files("Input/NN/sums_csv_new")
@@
   mutate(save_data = lapply(file_list, read_csv,
-                            col_names = c("line", "date", "time", "temp_c"),
-                            col_types = cols(line = col_double(),
+                            col_names = c("line", "date", "time", "temp_c", "trash"),
+                            col_types = cols(line = col_character(),
                                              date = col_date(format = ""),
                                              time = col_time(format = ""),
-                                             temp_c = col_character()))) %>%
+                                             temp_c = col_character(),
+                                             trash = col_character()))) %>%
   unnest(save_data) %>%
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 nn_sums_csv <- initial_data %>% 
@@
   unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = datetime,
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+# 1 file had this format: No header, columns of datetime and temp (C)
 list_csv_files_2 <- list.files("Input/NN/sums_csv_old/2col-c")
@@
   separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>%
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(date_sums = ymd(date_sums),
+         temp_c = if_else(is.na(temp_c), trash2, temp_c),
          temp_c_sums = as.numeric(temp_c),
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format
+## No header, 3 columns: date, time, temp (F)
 list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")
@@
                                              time = col_time(format = ""),
-                                             temp_f = col_double()))) %>%
+                                             temp_f = col_character()))) %>%
   unnest(save_data) %>%
@@
   unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
+  filter(temp_f != "Paused") %>% 
+  separate(temp_f, c("trash2", "temp_f"), sep = " ", remove = FALSE) %>%
   mutate(datetime_sums = mdy_hms(datetime),
@@
 
-write_rds(nn_sums_filter, "Output/nn_sums_raw.rds")
+# write_rds(nn_sums_filter, "Output/nn_sums_raw.rds")
 ```
@@
 # List files
+## These files were directly exported by me from the Logtag software.
+## I am confident in their formatting and content since I pulled the data
+## from the original file type of the field instrument.
+## The files have 5 columns: line, date, time, temp (C), notes
 list_csv_files <- list.files("Input/AK/sums_csv_new")
@@
   mutate(save_data = lapply(file_list, read_csv,
-                            col_names = c("line", "date", "time", "temp_c"),
-                            col_types = cols(line = col_double(),
+                            col_names = c("line", "date", "time", "temp_c", "trash"),
+                            col_types = cols(line = col_character(),
                                              date = col_date(format = ""),
                                              time = col_time(format = ""),
-                                             temp_c = col_double()))) %>%
+                                             temp_c = col_character(),
+                                             trash = col_character()))) %>%
   unnest(save_data) %>%
@@
   unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>%
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(datetime_sums = ymd_hms(datetime),
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format with no header
+## 5 columns: line, date, time, temp (F), notes
 list_csv_files_2 <- list.files("Input/AK/sums_csv_old")
@@
                             col_names = c("line", "date", "time", "temp_f", "trash"),
-                            col_types = cols(line = col_double(),
+                            col_types = cols(line = col_character(),
                                              date = col_character(),
                                              time = col_time(format = ""),
-                                             temp_f = col_double(),
+                                             temp_f = col_character(),
                                              trash = col_character())))  %>%
@@
   mutate(area = "AK") %>% 
+  filter(temp_f != "Paused") %>% 
+  separate(temp_f, c("trash2", "temp_f"), sep = " ", remove = FALSE) %>%
   mutate(date_sums = mdy(date),
          time_sums = time,
+         temp_f = if_else(is.na(temp_f), trash2, temp_f),
+         temp_f = as.numeric(temp_f),
          temp_c_sums = (temp_f-32)/1.8,
@@
 # List files
+## 3 files had this format with no header
+## 5 columns: date, time, am or pm, C, temp
 list_xlsx_files <- list.files("Input/AK/sums_xlsx_temp_c")
@@
   unite("time", c("time", "am_pm"), sep = " ") %>% 
+  filter(temp_c != "Paused") %>% 
+  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
   mutate(date_sums = ymd(date),
          time_sums = time,
+         temp_c = if_else(is.na(temp_c), trash2, temp_c),
          temp_c_sums = as.numeric(temp_c),
@@
   arrange(home, datetime_sums) %>% 
-  distinct()
+  distinct() %>% 
+  filter(!is.na(temp_c_sums))
 
 # List files
+## Numerous files had this format with no header
+## 5 columns: date, time, am or pm, F, temp
 list_xlsx_files2 <- list.files("Input/AK/sums_xlsx_temp_f")
@@
 
-# Bind all SUMs files together
+# Bind all SUMs files together after initial upload
 ```{r}
@@
 # Check sample start times vs log start times
+## This chunk of code produced a log of sample start times for each home
+## that we have SUMs/Logtag/IButton data for.
+## I used this log to go through individual homes and assess start times
+## vs expected start times, find missing start times, etc.
+## This log, with additional notes, are found on Box:
+### KidsAIR > Ethan - data cleaning and analysis > sums_time_comparison
 ```{r}
@@
 # Compile start and stop times for SUMs based on field logs
+## Start times were compiled based on the beginning of data collection
+## recorded in the paper logs. A list of start times is uploaded to Box:
+### KidsAIR > Ethan - data cleaning and analysis > sums_time_comparison
+## Stope times were based on the date of the final visit to a home each winter
+## This timestamp was extracted from the "medicalvisit" files found on Box
 ```{r}
@@
 # Load and format SUMs data logs for WMT
-#wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
 
@@
 #####
-#nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
 
@@
 #####
-#ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
 
@@
 
-check_dates <- nn_sums_filtered %>% 
+check_dates <- ak_sums_filtered %>% 
   group_by(home) %>% 

### 2020-01-13 home_char_act_cleaning.Rmd
data cleaning updates

---
title: "KidsAIR home activity and characteristics cleaning"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 13 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save home char data
```{r}
##### WMT #####
# Load data
home_char_initial_wmt <- read_csv("Input/WMT/wmt_home_char_20200113.csv")

wmt_home_char <- home_char_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         residents_total = as.factor(Residents),
         residents_under5 = as.factor(Under5),
         residents_under18 = as.factor(Under18),
         residents_smoke = as.factor(Smoke),
         residents_smoke_inside = as.factor(SmokeInside),
         home_type = as.factor(HomeType),
         home_year_built = as.factor(YearHomeBuilt),
         home_sqft = as.numeric(SqFootage),
         home_floors = as.factor(NumFloors),
         home_bedrooms = as.factor(NumBedrooms),
         home_windows = as.factor(NumWindows),
         home_pets = as.factor(NumPets),
         home_furry = as.factor(NumFurryPets),
         primary_heat = as.factor(PrimaryHeat),
         secondary_heat = as.factor(SecondaryHeat),
         stove_model = as.character(WoodstoveModel),
         stove_age = as.factor(WoodstoveAge),
         stove_cert = as.factor(WoodstoveEPACert),
         chimney_clean = as.factor(ChimneyCleaned),
         home_damp = as.factor(Damp),
         home_mold = as.factor(Mold),
         home_fireplace = as.factor(ExistingFireplace),
         home_furnace = as.factor(ExistingFurnace),
         home_woodstove = as.factor(ExistingWoodstove),
         home_insert = as.factor(ExistingInsert),
         area = "WMT") %>% 
  select(home_winter_id, area, residents_total, residents_under5,
         residents_under18, residents_smoke, residents_smoke_inside,
         home_type, home_year_built, home_sqft, home_floors, home_bedrooms,
         home_windows, home_pets, home_furry, primary_heat,
         secondary_heat, stove_model, stove_age, stove_cert, chimney_clean,
         home_damp, home_mold, home_fireplace, home_furnace,
         home_woodstove, home_insert) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(wmt_home_char, "Output/wmt_home_char.rds")

##### NN #####
# Load data
home_char_initial_nn <- read_xlsx("Input/NN/nn_home_char_20200113.xlsx")

nn_home_char <- home_char_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         residents_total = as.factor(Residents),
         residents_under5 = as.factor(Under5),
         residents_under18 = as.factor(Under18),
         residents_smoke = as.factor(Smoke),
         residents_smoke_inside = as.factor(SmokeInside),
         home_type = as.factor(HomeType),
         home_year_built = as.factor(YearHomeBuilt),
         home_sqft = as.numeric(SqFootage),
         home_floors = as.factor(NumFloors),
         home_bedrooms = as.factor(NumBedrooms),
         home_windows = as.factor(NumWindows),
         home_pets = as.factor(NumPets),
         home_furry = as.factor(NumFurryPets),
         primary_heat = as.factor(PrimaryHeat),
         secondary_heat = as.factor(SecondaryHeat),
         stove_model = as.character(WoodstoveModel),
         stove_age = as.factor(WoodstoveAge),
         stove_cert = as.factor(WoodstoveEPACert),
         chimney_clean = as.factor(ChimneyCleaned),
         home_damp = as.factor(Damp),
         home_mold = as.factor(Mold),
         home_fireplace = as.factor(ExistingFireplace),
         home_furnace = as.factor(ExistingFurnace),
         home_woodstove = as.factor(ExistingWoodstove),
         home_insert = as.factor(ExistingInsert),
         area = "NN") %>% 
  select(home_winter_id, area, residents_total, residents_under5,
         residents_under18, residents_smoke, residents_smoke_inside,
         home_type, home_year_built, home_sqft, home_floors, home_bedrooms,
         home_windows, home_pets, home_furry, primary_heat,
         secondary_heat, stove_model, stove_age, stove_cert, chimney_clean,
         home_damp, home_mold, home_fireplace, home_furnace,
         home_woodstove, home_insert) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(nn_home_char, "Output/nn_home_char.rds")

##### AK #####
# Load data
home_char_initial_ak <- read_xlsx("Input/AK/ak_home_char_20200113.xlsx")

ak_home_char <- home_char_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         residents_total = as.factor(Residents),
         residents_under5 = as.factor(Under5),
         residents_under18 = as.factor(Under18),
         residents_smoke = as.factor(Smoke),
         residents_smoke_inside = as.factor(SmokeInside),
         home_type = as.factor(HomeType),
         home_year_built = as.factor(YearHomeBuilt),
         home_sqft = as.numeric(SqFootage),
         home_floors = as.factor(NumFloors),
         home_bedrooms = as.factor(NumBedrooms),
         home_windows = as.factor(NumWindows),
         home_pets = as.factor(NumPets),
         home_furry = as.factor(NumFurryPets),
         primary_heat = as.factor(PrimaryHeat),
         secondary_heat = as.factor(SecondaryHeat),
         stove_model = as.character(WoodstoveModel),
         stove_age = as.factor(WoodstoveAge),
         stove_cert = as.factor(WoodstoveEPACert),
         chimney_clean = as.factor(ChimneyCleaned),
         home_damp = as.factor(Damp),
         home_mold = as.factor(Mold),
         home_fireplace = as.factor(ExistingFireplace),
         home_furnace = as.factor(ExistingFurnace),
         home_woodstove = as.factor(ExistingWoodstove),
         home_insert = as.factor(ExistingInsert),
         area = "AK") %>% 
  select(home_winter_id, area, residents_total, residents_under5,
         residents_under18, residents_smoke, residents_smoke_inside,
         home_type, home_year_built, home_sqft, home_floors, home_bedrooms,
         home_windows, home_pets, home_furry, primary_heat,
         secondary_heat, stove_model, stove_age, stove_cert, chimney_clean,
         home_damp, home_mold, home_fireplace, home_furnace,
         home_woodstove, home_insert) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(ak_home_char, "Output/ak_home_char.rds")
```

# Join home characteristic data from different areas
```{r}
wmt_home_char <- read_rds("Output/wmt_home_char.rds")
nn_home_char <- read_rds("Output/nn_home_char.rds")
ak_home_char <- read_rds("Output/ak_home_char.rds")

kids_home_char <- rbind(wmt_home_char, nn_home_char, ak_home_char)

# write_rds(kids_home_char, "Output/kids_home_char.rds")
```

# Clean up all home characteristics data
```{r}
kids_home_char <- read_rds("Output/kids_home_char.rds")

kids_home_char_clean <- kids_home_char

# write_rds(kids_home_char_clean, "Output/kids_home_char_clean.rds")
```

###############################################

# Load, format, and save home activity data
```{r}
##### WMT #####
# Load data
home_act_initial_wmt <- read_csv("Input/WMT/wmt_home_activity_20200113.csv")

wmt_home_act <- home_act_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         home_act_day = as.factor(Day),
         home_act_gas = as.factor(Gas),
         home_act_elect = as.factor(Electrical),
         home_act_propane = as.factor(Propane),
         home_act_oil = as.factor(Oil),
         home_act_other = as.factor(Other),
         home_act_smoking = as.factor(Smoking),
         home_act_incense = as.factor(Incense),
         home_act_candle = as.factor(Candle),
         home_act_kerosene = as.factor(Kerosene),
         home_act_windows = as.factor(OpenWindows),
         home_act_door = as.factor(OpenDoor),
         home_act_sweep = as.factor(Sweeping),
         home_act_comments = as.character(Comments),
         area = "WMT") %>% 
  select(home_winter_id, area, home_act_day,
         home_act_gas, home_act_elect, home_act_propane,
         home_act_oil, home_act_other, home_act_smoking,
         home_act_incense, home_act_candle, home_act_kerosene,
         home_act_windows, home_act_door, home_act_sweep, home_act_comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id, home_act_day)

# Save as RDS
# write_rds(wmt_home_act, "Output/wmt_home_act.rds")

##### NN #####
# Load data
home_act_initial_nn <- read_xlsx("Input/NN/nn_home_activity_20200113.xlsx")

nn_home_act <- home_act_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         home_act_day = as.factor(Day),
         home_act_gas = as.factor(Gas),
         home_act_elect = as.factor(Electrical),
         home_act_propane = as.factor(Propane),
         home_act_oil = as.factor(Oil),
         home_act_other = as.factor(Other),
         home_act_smoking = as.factor(Smoking),
         home_act_incense = as.factor(Incense),
         home_act_candle = as.factor(Candle),
         home_act_kerosene = as.factor(Kerosene),
         home_act_windows = as.factor(OpenWindows),
         home_act_door = as.factor(OpenDoor),
         home_act_sweep = as.factor(Sweeping),
         home_act_comments = as.character(IgnoreReason),
         area = "NN") %>% 
  select(home_winter_id, area, home_act_day,
         home_act_gas, home_act_elect, home_act_propane,
         home_act_oil, home_act_other, home_act_smoking,
         home_act_incense, home_act_candle, home_act_kerosene,
         home_act_windows, home_act_door, home_act_sweep, home_act_comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id, home_act_day)

# Save as RDS
# write_rds(nn_home_act, "Output/nn_home_act.rds")

##### AK #####
# Load data
home_act_initial_ak <- read_xlsx("Input/AK/ak_home_activity_20200113.xlsx")

ak_home_act <- home_act_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         home_act_day = as.factor(Day),
         home_act_gas = as.factor(Gas),
         home_act_elect = as.factor(Electrical),
         home_act_propane = as.factor(Propane),
         home_act_oil = as.factor(Oil),
         home_act_other = as.factor(Other),
         home_act_smoking = as.factor(Smoking),
         home_act_incense = as.factor(Incense),
         home_act_candle = as.factor(Candle),
         home_act_kerosene = as.factor(Kerosene),
         home_act_windows = as.factor(OpenWindows),
         home_act_door = as.factor(OpenDoor),
         home_act_sweep = as.factor(Sweeping),
         home_act_comments = as.character(IgnoreReason),
         area = "AK") %>% 
  select(home_winter_id, area, home_act_day,
         home_act_gas, home_act_elect, home_act_propane,
         home_act_oil, home_act_other, home_act_smoking,
         home_act_incense, home_act_candle, home_act_kerosene,
         home_act_windows, home_act_door, home_act_sweep, home_act_comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id, home_act_day)

# Save as RDS
# write_rds(ak_home_act, "Output/ak_home_act.rds")
```

# Join home activity data from different areas
```{r}
wmt_home_act <- read_rds("Output/wmt_home_act.rds")
nn_home_act <- read_rds("Output/nn_home_act.rds")
ak_home_act <- read_rds("Output/ak_home_act.rds")

kids_home_act <- rbind(wmt_home_act, nn_home_act, ak_home_act)

# write_rds(kids_home_act, "Output/kids_home_act.rds")
```

# Clean up all home activity data
```{r}
kids_home_act <- read_rds("Output/kids_home_act.rds")

kids_home_act_clean <- kids_home_act

# write_rds(kids_home_act_clean, "Output/kids_home_act_clean.rds")
```

### 2020-01-13 indoor_temp_rh_cleaning.Rmd
data cleaning updates

---
title: "KidsAIR indoor temp and humidity cleaning"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 13 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save indoor temp/rh data
```{r}
##### WMT #####
# Load data
temp_rh_initial_wmt <- read_csv("Input/WMT/wmt_indoor_temp_rh_20200113.csv")

wmt_temp_rh <- temp_rh_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(`REPLACE(REPLACE(Comments`),
         area = "WMT") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(wmt_temp_rh, "Output/wmt_temp_rh.rds")

##### NN #####
# Load data
temp_rh_initial_nn <- read_xlsx("Input/NN/nn_indoor_temp_rh_20200113.xlsx")

nn_temp_rh <- temp_rh_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(IgnoreReason),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(nn_temp_rh, "Output/nn_temp_rh.rds")

##### AK #####
# Load data
temp_rh_initial_ak <- read_xlsx("Input/AK/ak_indoor_temp_rh_20200113.xlsx")

ak_temp_rh <- temp_rh_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         rh_indoor_current = as.numeric(RelHumCur),
         rh_indoor_max = as.numeric(RelHumMax),
         rh_indoor_min = as.numeric(RelHumMin),
         temp_indoor_current = as.numeric(TempCCur),
         temp_indoor_max = as.numeric(TempCMax),
         temp_indoor_min = as.numeric(TempCMin),
         rh_temp_comments = as.character(IgnoreReason),
         area = "AK") %>% 
  # replace -999 values with NA
  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
                                 rh_indoor_min = -999, temp_indoor_current = -999,
                                 temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
                                 rh_indoor_min = 0, temp_indoor_current = 0,
                                 temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
  select(home_winter_id, area, rh_indoor_current, 
         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
  arrange(home_winter_id)

# Save as RDS 
# write_rds(ak_temp_rh, "Output/ak_temp_rh.rds")
```

# Join temp/rh data from different areas
```{r}
wmt_temp_rh <- read_rds("Output/wmt_temp_rh.rds")
nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
ak_temp_rh <- read_rds("Output/ak_temp_rh.rds")

kids_indoor_temp_rh <- rbind(wmt_temp_rh, nn_temp_rh, ak_temp_rh)

# write_rds(kids_indoor_temp_rh, "Output/kids_indoor_temp_rh.rds")
```

# Clean up all temp/rh data
```{r}
kids_indoor_temp_rh <- read_rds("Output/kids_indoor_temp_rh.rds")

kids_indoor_temp_rh_clean <- kids_indoor_temp_rh

# write_rds(kids_indoor_temp_rh_clean, "Output/kids_indoor_temp_rh_clean.rds")
```

### 2020-01-13 moisture_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 13 Dec 2019"
+date: "Started 19 Nov 2019, Updated 13 Jan 2020"
 output: pdf_document
@@
 ```{r}
+##### WMT #####
 # Load data
-moisture_initial_wmt <- read_xlsx("Input/wmt_kids_moisture_20191122.xlsx")
+moisture_initial_wmt <- read_xlsx("Input/WMT/wmt_kids_moisture_20191122.xlsx")
 
-kids_moisture <- moisture_initial_wmt %>% 
+wmt_moisture <- moisture_initial_wmt %>% 
   # rename variables
-  mutate(moisture_id = as.factor(ID),  
-         home_winter_id = as.factor(HomeWinterID),
+  mutate(home_winter_id = as.factor(HomeWinterID),
          moisture_date = ymd(CompletedDate),
@@
          moisture_ave = as.numeric(MoistureAve),
-         moisture_missing = as.character(WhatMissing),
-         moisture_incomplete = as.factor(Incomplete),
-         moisture_ignore_missing = as.factor(IgnoreMissingData),
-         moisture_ignore_reason = as.character(IgnoreReason),
-         moisture_tech = as.character(Technician),
          moisture_split = as.factor(Split),
@@
   spread(sample, value) %>% 
-  select(moisture_id, home_winter_id, area, moisture_date, moisture_1, moisture_2, moisture_3, 
-         moisture_ave, moisture_missing, moisture_incomplete, moisture_ignore_missing,
-         moisture_ignore_reason, moisture_tech, moisture_split)
+  select(home_winter_id, area, moisture_date, 
+         moisture_1, moisture_2, moisture_3, 
+         moisture_ave, moisture_split) %>% 
+  arrange(home_winter_id, moisture_date)
 
-# Save as RDS and CSV
-write_rds(kids_moisture, "Output/kids_moisture.rds")
-write_csv(kids_moisture, "Output/kids_moisture.csv")
+# Save as RDS 
+# write_rds(wmt_moisture, "Output/wmt_moisture.rds")
+
+##### NN #####
+# Load data
+moisture_initial_nn <- read_xlsx("Input/NN/nn_moisture_20200113.xlsx")
+
+nn_moisture <- moisture_initial_nn %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         moisture_date = ymd(CompletedDate),
+         moisture_1 = as.numeric(Moisture1),
+         moisture_2 = as.numeric(Moisture2),
+         moisture_3 = as.numeric(Moisture3),
+         moisture_ave = as.numeric(MoistureAve),
+         moisture_split = as.factor(Split),
+         area = "NN") %>% 
+  # replace 0 and -999 values with NA
+  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
+                                 moisture_3 = 0, moisture_ave = 0)) %>% 
+  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
+                                 moisture_3 = -999)) %>% 
+  # gather 3 moisture samples to calculate average
+  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
+  group_by(home_winter_id, moisture_date) %>% 
+  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  # spread 3 moisture samples back to wide format
+  spread(sample, value) %>% 
+  select(home_winter_id, area, moisture_date, 
+         moisture_1, moisture_2, moisture_3, 
+         moisture_ave, moisture_split) %>% 
+  arrange(home_winter_id, moisture_date)
+
+# Save as RDS 
+# write_rds(nn_moisture, "Output/nn_moisture.rds")
+
+##### AK #####
+# Load data
+moisture_initial_ak <- read_xlsx("Input/AK/ak_moisture_20200113.xlsx")
+
+ak_moisture <- moisture_initial_ak %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         moisture_date = ymd(CompletedDate),
+         moisture_1 = as.numeric(Moisture1),
+         moisture_2 = as.numeric(Moisture2),
+         moisture_3 = as.numeric(Moisture3),
+         moisture_ave = as.numeric(MoistureAve),
+         moisture_split = as.factor(Split),
+         area = "AK") %>% 
+  # replace 0 and -999 values with NA
+  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
+                                 moisture_3 = 0, moisture_ave = 0)) %>% 
+  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
+                                 moisture_3 = -999)) %>% 
+  # gather 3 moisture samples to calculate average
+  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
+  group_by(home_winter_id, moisture_date) %>% 
+  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  # spread 3 moisture samples back to wide format
+  spread(sample, value) %>% 
+  select(home_winter_id, area, moisture_date, 
+         moisture_1, moisture_2, moisture_3, 
+         moisture_ave, moisture_split) %>% 
+  arrange(home_winter_id, moisture_date)
+
+# Save as RDS 
+# write_rds(ak_moisture, "Output/ak_moisture.rds")
 ```
+
+# Join moisture data from different areas
+```{r}
+wmt_moisture <- read_rds("Output/wmt_moisture.rds")
+nn_moisture <- read_rds("Output/nn_moisture.rds")
+ak_moisture <- read_rds("Output/ak_moisture.rds")
+
+kids_moisture <- rbind(wmt_moisture, nn_moisture, ak_moisture)
+
+# write_rds(kids_moisture, "Output/kids_moisture.rds")
+```
+
+# Clean up all moisture data
+```{r}
+kids_moisture <- read_rds("Output/kids_moisture.rds")
+
+kids_moisture_clean <- kids_moisture
+
+# write_rds(kids_moisture_clean, "Output/kids_moisture_clean.rds")
+```

### 2020-01-13 pm_summary.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 19 Dec 2019"
+date: "Started 22 Nov 2019, Updated 13 Jan 2020"
 output: pdf_document
@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(area) %>% 
+  group_by(treatment) %>% 
   summarize("n" = n(),

### 2020-01-13 stove_grade_usage_cleaning.Rmd
data cleaning updates

---
title: "KidsAIR wood stove grades and usage"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 13 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load, format, and save wood stove grade data
```{r}
##### WMT #####
# Load data
stove_grades_initial_wmt <- read_csv("Input/WMT/wmt_stove_grades_20200113.csv")

wmt_stove_grades <- stove_grades_initial_wmt %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         comments = as.character(`REPLACE(REPLACE(Notes`),
         area = "WMT") %>% 
  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(wmt_stove_grades, "Output/wmt_stove_grades.rds")

##### NN #####
# Load data
stove_grades_initial_nn <- read_xlsx("Input/NN/nn_stove_grades_20200113.xlsx")

nn_stove_grades <- stove_grades_initial_nn %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         comments = as.character(IgnoreReason),
         area = "NN") %>% 
  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(nn_stove_grades, "Output/nn_stove_grades.rds")

##### AK #####
# Load data
stove_grades_initial_ak <- read_xlsx("Input/AK/ak_stove_grades_20200113.xlsx")

ak_stove_grades <- stove_grades_initial_ak %>% 
  # rename variables
  mutate(home_id = as.factor(HomeID),
         home_winter_id = as.factor(HomeWinterID),
         stove_grade = as.factor(Score1),
         comments = as.character(IgnoreReason),
         area = "AK") %>% 
  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(ak_stove_grades, "Output/ak_stove_grades.rds")
```

# Join stove grade data from different areas
```{r}
wmt_stove_grades <- read_rds("Output/wmt_stove_grades.rds")
nn_stove_grades <- read_rds("Output/nn_stove_grades.rds")
ak_stove_grades <- read_rds("Output/ak_stove_grades.rds")

kids_stove_grades <- rbind(wmt_stove_grades, nn_stove_grades, ak_stove_grades)

# write_rds(kids_stove_grades, "Output/kids_stove_grades.rds")
```

# Clean up all stove grade data
```{r}
kids_stove_grades <- read_rds("Output/kids_stove_grades.rds")

kids_stove_grades_clean <- kids_stove_grades

# write_rds(kids_stove_grades_clean, "Output/kids_stove_grades_clean.rds")
```

###############################################

# Load, format, and save wood stove use data
```{r}
##### WMT #####
# Load data
stove_use_initial_wmt <- read_csv("Input/WMT/wmt_stove_use_20200113.csv")

wmt_stove_use <- stove_use_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(HowGet),
         wood_cords = as.numeric(Cords),
         area = "WMT") %>% 
  select(home_winter_id, area, burn_level, wood_type, wood_type_other,
         wood_collect, wood_collect_method, wood_cords) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(wmt_stove_use, "Output/wmt_stove_use.rds")

##### NN #####
# Load data
stove_use_initial_nn <- read_xlsx("Input/NN/nn_stove_use_20200113.xlsx")

nn_stove_use <- stove_use_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(HowGet),
         wood_cords = as.numeric(Cords),
         area = "NN") %>% 
  select(home_winter_id, area, burn_level, wood_type, wood_type_other,
         wood_collect, wood_collect_method, wood_cords) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(nn_stove_use, "Output/nn_stove_use.rds")

##### AK #####
# Load data
stove_use_initial_ak <- read_xlsx("Input/AK/ak_stove_use_20200113.xlsx")

ak_stove_use <- stove_use_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         burn_level = as.factor(Burning),
         wood_type = as.factor(Wood),
         wood_type_other = as.character(WhOtherWood),
         wood_collect = as.factor(WhenGet),
         wood_collect_method = as.factor(HowGet),
         wood_cords = as.numeric(Cords),
         area = "AK") %>% 
  select(home_winter_id, area, burn_level, wood_type, wood_type_other,
         wood_collect, wood_collect_method, wood_cords) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

# Save as RDS
# write_rds(ak_stove_use, "Output/ak_stove_use.rds")
```

# Join stove use data from different areas
```{r}
wmt_stove_use <- read_rds("Output/wmt_stove_use.rds")
nn_stove_use <- read_rds("Output/nn_stove_use.rds")
ak_stove_use <- read_rds("Output/ak_stove_use.rds")

kids_stove_use <- rbind(wmt_stove_use, nn_stove_use, ak_stove_use)

# write_rds(kids_stove_use, "Output/kids_stove_use.rds")
```

# Clean up all stove use data
```{r}
kids_stove_use <- read_rds("Output/kids_stove_use.rds")

kids_stove_use_clean <- kids_stove_use %>% 
  replace_with_na(replace = list(wood_cords = -999, wood_cords = 0))

# write_rds(kids_stove_use_clean, "Output/kids_stove_use_clean.rds")
```

### 2020-01-14 home_char_act_cleaning.Rmd
cleaning joining datasets begin summarizing data

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 13 Jan 2020"
+date: "Started 13 Jan 2020, Updated 14 Jan 2020"
 output: pdf_document
@@
 kids_home_char <- read_rds("Output/kids_home_char.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_home_char_clean <- kids_home_char
+home_char_clean <- kids_home_char %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_home_char_clean, "Output/kids_home_char_clean.rds")
+# write_rds(home_char_clean, "Output/home_char_clean.rds")
+```
+
+# Summary data for home characteristics
+```{r}
+# Load dataset
+home_char_clean <- read_rds("Output/home_char_clean.rds")
+
+# Summarize numeric vars
+home_char_summary <- home_char_clean %>% 
+  group_by(area) %>% 
+  summarize(mean(home_sqft, na.rm = TRUE), 
+            min(home_sqft, na.rm = TRUE), 
+            median(home_sqft, na.rm = TRUE), 
+            max(home_sqft, na.rm = TRUE))
+home_char_summary
+
+# Summarize factor vars
+## Use this as a quick look at all vars
+table_list <- home_char_clean %>% 
+  #filter(area == "WMT") %>% 
+  select(-home_winter_id, -home_sqft, -stove_model) 
+lapply(table_list, table)
+
+## Select specific var to assess individually/group by area
+table(table_list$residents_total, table_list$area, useNA = "ifany")
 ```
@@
 kids_home_act <- read_rds("Output/kids_home_act.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_home_act_clean <- kids_home_act
+home_act_clean <- kids_home_act %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_home_act_clean, "Output/kids_home_act_clean.rds")
+# write_rds(home_act_clean, "Output/home_act_clean.rds")
 ```
+
+# Summary data for home characteristics
+```{r}
+# Load dataset
+home_act_clean <- read_rds("Output/home_act_clean.rds")
+
+# Summarize factor vars
+## Use this as a quick look at all vars
+table_list <- home_act_clean %>% 
+  #filter(area == "WMT") %>% 
+  select(-home_winter_id, -home_act_comments) 
+lapply(table_list, table)
+
+## Select specific var to assess individually/group by area
+table(table_list$home_act_sweep, table_list$area, useNA = "ifany")
+```

### 2020-01-14 indoor_temp_rh_cleaning.Rmd
cleaning joining datasets begin summarizing data

@@
 kids_indoor_temp_rh <- read_rds("Output/kids_indoor_temp_rh.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_indoor_temp_rh_clean <- kids_indoor_temp_rh
+indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_indoor_temp_rh_clean, "Output/kids_indoor_temp_rh_clean.rds")
+# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
 ```
+
+# Summarize temp/rh data
+```{r}
+indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
+
+temp_rh_summary <- indoor_temp_rh_clean %>%
+  group_by(area) %>% 
+  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
+                                med = median, max = max), na.rm = TRUE) %>% 
+  t()
+temp_rh_summary
+```

### 2020-01-14 join_format_kids_data.Rmd
cleaning joining datasets begin summarizing data

---
title: "KidsAIR: Join data and format full dataset"
author: "Ethan Walker"
date: "Started 14 Jan 2020, Updated 14 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```
# Use this Markdown to join datasets for analysis

# Load individual datasets
```{r}
# These files are millions of observations long and take time to load/edit
# Make a separate way to join these data and only use when necessary
# Also consider how to join/analyze these data since they were collected
# at different frequencies (20 or 25 min for SUMs; 1 min for PM)
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

# These other files have fewer observations and are easier to work with
# Make a full dataset by joining these datasets together
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Join small datasets
```{r}

```

# Format small dataset
```{r}

```

# Join large datasets
```{r}

```

### 2020-01-14 moisture_data_cleaning.Rmd
cleaning joining datasets begin summarizing data

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 13 Jan 2020"
+date: "Started 19 Nov 2019, Updated 14 Jan 2020"
 output: pdf_document
@@
 kids_moisture <- read_rds("Output/kids_moisture.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_moisture_clean <- kids_moisture
+moisture_clean <- kids_moisture %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_moisture_clean, "Output/kids_moisture_clean.rds")
+# write_rds(moisture_clean, "Output/moisture_clean.rds")
 ```

### 2020-01-14 moisture_summary.Rmd
cleaning joining datasets begin summarizing data

--- 2019-11-22 moisture_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019"
+date: "Started 22 Nov 2019, Updated 14 Jan 2020"
 output: pdf_document
@@
 # Load data
-kids_moisture <- read_rds("Output/kids_moisture.rds")
+moisture_clean <- read_rds("Output/moisture_clean.rds")
 ```
@@
 ```{r}
-summary_kids_moisture <- kids_moisture %>% 
-  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
-            min(pm_24hr_mean), max(pm_24hr_mean),
-            min(pm), max(pm),
-            min(pm_sampling_days), max(pm_sampling_days),
-            min(pm_total_observations), max(pm_total_observations)) %>% 
+summary_moisture <- moisture_clean %>% 
+  group_by(area) %>% 
+  summarize(mean(moisture_ave, na.rm = TRUE), sd(moisture_ave, na.rm = TRUE),
+            median(moisture_ave, na.rm = TRUE),
+            min(moisture_ave, na.rm = TRUE), max(moisture_ave, na.rm = TRUE)) %>% 
   t()
-summary_kids_pm
+summary_moisture
 ```
@@
 ```{r, fig.height=6, fig.width=5}
-boxplots_kids_pm <- kids_pm %>% 
-  ggplot(aes(area, pm_24hr_mean)) +
+boxplots_moisture <- moisture_clean %>% 
+  select(area, moisture_ave) %>% 
+  na.exclude() %>% 
+  ggplot(aes(area, moisture_ave)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
           axis.ticks = element_blank())
-boxplots_kids_pm
+boxplots_moisture
 ```

### 2020-01-14 pm_summary.Rmd
cleaning joining datasets begin summarizing data

--- 2019-12-20 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 19 Dec 2019"
+date: "Started 22 Nov 2019, Updated 14 Jan 2020"
 output: pdf_document
@@
 # Load data
-kids_pm <- read_rds("Output/kids_pm_new.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
 ```{r}
-pm_summary <- kids_pm %>% 
+pm_summary <- pm_clean %>% 
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(area) %>% 
+  group_by(treatment) %>% 
   summarize("n" = n(),
@@
 ```{r, fig.height=6, fig.width=5}
-boxplots_kids_pm <- kids_pm %>% 
+boxplots_kids_pm <- pm_clean %>% 
   #filter(pm < 500) %>% 

### 2020-01-14 stove_grade_usage_cleaning.Rmd
cleaning joining datasets begin summarizing data

@@
 kids_stove_grades <- read_rds("Output/kids_stove_grades.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_stove_grades_clean <- kids_stove_grades
+stove_grades_clean <- kids_stove_grades %>% 
+  filter(home_id != 0) %>% 
+  rename(stove_grade_comments = comments) %>% 
+  select(-home_id) %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_stove_grades_clean, "Output/kids_stove_grades_clean.rds")
+# write_rds(stove_grades_clean, "Output/stove_grades_clean.rds")
+
+# Dataset showing home with no stove grade
+stove_grades_na <- stove_grades_clean %>% 
+  filter(is.na(stove_grade)) %>% 
+  distinct(home_id) 
+```
+
+# Summary data for stove grades
+```{r}
+# Load dataset
+stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
+
+stove_grade_summary <- stove_grades_clean %>% 
+  group_by(area, stove_grade) %>% 
+  summarize(n())
+stove_grade_summary
 ```
@@
 kids_stove_use <- read_rds("Output/kids_stove_use.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_stove_use_clean <- kids_stove_use %>% 
-  replace_with_na(replace = list(wood_cords = -999, wood_cords = 0))
+stove_use_clean <- kids_stove_use %>% 
+  replace_with_na(replace = list(wood_cords = -999, wood_cords = 0)) %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_stove_use_clean, "Output/kids_stove_use_clean.rds")
+# write_rds(stove_use_clean, "Output/stove_use_clean.rds")
 ```
+
+# Summary data for stove use
+```{r}
+# Load dataset
+stove_use_clean <- read_rds("Output/stove_use_clean.rds")
+
+# Summarize numeric vars
+stove_use_summary <- stove_use_clean %>% 
+  group_by(area) %>% 
+  summarize(mean(wood_cords, na.rm = TRUE), 
+            min(wood_cords, na.rm = TRUE), 
+            median(wood_cords, na.rm = TRUE), 
+            max(wood_cords, na.rm = TRUE))
+stove_use_summary
+
+# Summarize factor vars
+## Use this as a quick look at all vars
+table_list <- stove_use_clean %>% 
+  #filter(area == "WMT") %>% 
+  select(-home_winter_id, -wood_cords) 
+lapply(table_list, table)
+
+## Select specific var to assess individually/group by area
+table(stove_use_clean$burn_level, stove_use_clean$area, useNA = "ifany")
+```

### 2020-01-14 sums_data_cleaning.Rmd
cleaning joining datasets begin summarizing data

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 10 Jan 2020"
+date: "Started 23 Dec 2019, Updated 14 Jan 2020"
 output: pdf_document
@@
 all_sums_filtered <- read_rds("Output/all_sums_filtered.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+  select(home_winter_id, area, treatment)
 
@@
           winter_id = as.factor(winter_id),
-          home_winter_id = as.factor(home_winter_id),
+          home_winter_id = as.character(home_winter_id),
           home_id = as.factor(home_id),
@@
   select(-filter_var) %>% 
-  arrange(area, home, winter_id, datetime_sums)
+  arrange(area, home, winter_id, datetime_sums) %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 

### 2020-01-16 home_char_act_cleaning.Rmd
cleaning and summary stat updates

--- 2020-01-13 home_char_act_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 13 Jan 2020"
+date: "Started 13 Jan 2020, Updated 15 Jan 2020"
 output: pdf_document
@@
 kids_home_char <- read_rds("Output/kids_home_char.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_home_char_clean <- kids_home_char
+home_char_clean <- kids_home_char %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_home_char_clean, "Output/kids_home_char_clean.rds")
+# write_rds(home_char_clean, "Output/home_char_clean.rds")
 ```
@@
 kids_home_act <- read_rds("Output/kids_home_act.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 
-kids_home_act_clean <- kids_home_act
+home_act_clean <- kids_home_act %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
 
-# write_rds(kids_home_act_clean, "Output/kids_home_act_clean.rds")
+# write_rds(home_act_clean, "Output/home_act_clean.rds")
 ```

### 2020-01-16 home_char_act_summary.Rmd
cleaning and summary stat updates

---
title: "Kids home characteristics/activity summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 15 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(readxl)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load dataset
home_char_clean <- read_rds("Output/home_char_clean.rds")

home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Summary data for home characteristics
```{r}
# Summarize numeric vars
home_char_summary <- home_char_clean %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(home_sqft != 0) %>% 
  group_by(area, winter_id) %>% 
  summarize("N" = n(),
            "Mean Sqft" = mean(home_sqft, na.rm = TRUE), 
            "SD Sqft" = sd(home_sqft, na.rm = TRUE), 
            "Min Sqft" = min(home_sqft, na.rm = TRUE), 
            "Median Sqft" = median(home_sqft, na.rm = TRUE), 
            "Max Sqft" = max(home_sqft, na.rm = TRUE))
kable(home_char_summary, digits = 1, align = "c")

# Summarize factor vars
## Use this as a quick look at all vars
table_list <- home_char_clean %>%  
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(residents_total != 0) %>% 
  #filter(area == "WMT") %>% 
  select(-home_winter_id, -home_sqft, -stove_model, -home_id, -home, -treatment) 
#lapply(table_list, table)

## Select specific var to assess individually/group by area
table(table_list$winter_id, table_list$area, table_list$home_type, useNA = "ifany")
```

# Summary data for home activity
```{r}
# Summarize factor vars
## Use this as a quick look at all vars
table_list <- home_act_clean %>% 
  filter(winter_id == 1) %>% 
  #filter(area == "WMT") %>% 
  select(-home_winter_id, -home_act_comments, -home_id, -home, -treatment) 
#lapply(table_list, table)

## Select specific var to assess individually/group by area
#table(table_list$home_act_sweep, table_list$area, useNA = "ifany")
```

# Home characteristics summary plots
```{r}
home_char_function <- function(data, variable) {

home_plot <- data %>% 
  filter(winter_id == 1) %>% 
  filter(home_sqft != 0) %>%
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
home_plot

}

home_char_function(home_char_clean, "residents_total")
home_char_function(home_char_clean, "residents_under5")
home_char_function(home_char_clean, "residents_under18")
home_char_function(home_char_clean, "residents_smoke")
home_char_function(home_char_clean, "residents_smoke_inside")
home_char_function(home_char_clean, "home_type")
home_char_function(home_char_clean, "home_floors")
home_char_function(home_char_clean, "home_windows")
home_char_function(home_char_clean, "home_bedrooms")
home_char_function(home_char_clean, "home_pets")
home_char_function(home_char_clean, "home_furry")
home_char_function(home_char_clean, "primary_heat")
home_char_function(home_char_clean, "secondary_heat")
home_char_function(home_char_clean, "stove_age")
home_char_function(home_char_clean, "stove_cert")
home_char_function(home_char_clean, "chimney_clean")
home_char_function(home_char_clean, "home_mold")
home_char_function(home_char_clean, "home_damp")
home_char_function(home_char_clean, "home_fireplace")
home_char_function(home_char_clean, "home_furnace")
home_char_function(home_char_clean, "home_woodstove")
home_char_function(home_char_clean, "home_insert")
```

# Home activity summary plots
```{r}
home_act_function <- function(data, variable) {

home_plot <- data %>% 
  filter(winter_id == 1) %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
home_plot

}

home_act_function(home_act_clean, "home_act_gas")
home_act_function(home_act_clean, "home_act_elect")
home_act_function(home_act_clean, "home_act_propane")
home_act_function(home_act_clean, "home_act_oil")
home_act_function(home_act_clean, "home_act_smoking")
home_act_function(home_act_clean, "home_act_incense")
home_act_function(home_act_clean, "home_act_candle")
home_act_function(home_act_clean, "home_act_kerosene")
home_act_function(home_act_clean, "home_act_windows")
home_act_function(home_act_clean, "home_act_door")
home_act_function(home_act_clean, "home_act_sweep")
```

### 2020-01-16 moisture_summary.Rmd
cleaning and summary stat updates

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 14 Jan 2020"
+date: "Started 22 Nov 2019, Updated 15 Jan 2020"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
+                      include = TRUE, warning = FALSE)
 ```
@@
 
-# Summary stats
+# Moisture content summary stats
+## Average of 3 measurements
 ```{r}
@@
   group_by(area) %>% 
-  summarize(mean(moisture_ave, na.rm = TRUE), sd(moisture_ave, na.rm = TRUE),
-            median(moisture_ave, na.rm = TRUE),
-            min(moisture_ave, na.rm = TRUE), max(moisture_ave, na.rm = TRUE)) %>% 
+  summarize(mean = mean(moisture_ave, na.rm = TRUE), 
+            SD = sd(moisture_ave, na.rm = TRUE),
+            median = median(moisture_ave, na.rm = TRUE),
+            min = min(moisture_ave, na.rm = TRUE), 
+            max = max(moisture_ave, na.rm = TRUE)) %>%
+  select(-area) %>% 
   t() 
-summary_moisture
+kable(summary_moisture, col.names = c("Alaska", "Navajo", "Montana"), 
+      digits = 1, align = "c")
+```
+
+# Was wood split?
+```{r}
+wood_split <- moisture_clean %>% 
+  filter(!is.na(moisture_split)) %>% 
+  group_by(area, moisture_split) %>% 
+  summarise(n = n())
+kable(wood_split, align = "c")
 ```

### 2020-01-16 stove_grade_usage_summary.Rmd
cleaning and summary stat updates

---
title: "Kids stove grades and usage summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 15 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
```

```{r, eval=FALSE}
# Save this for potential future use

# Summarize factor vars
## Use this as a quick look at all vars
table_list <- stove_use_clean %>% 
  #filter(area == "WMT") %>% 
  filter(winter_id == 1) %>% 
  select(-home_winter_id, -wood_cords, -wood_type_other,
         -home_id, -home, -treatment) 
lapply(table_list, table)
```

# Summary of stove grades
```{r}
## Look further into NA values? ##
stove_grade_summary <- stove_grades_clean %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  select(area, home, stove_grade) %>% 
  filter(!is.na(stove_grade)) %>% 
  group_by(area, stove_grade) %>% 
  distinct(home) %>% 
  summarize(n())
kable(stove_grade_summary, align = "c")
```

## 18 homes with grades from 2 winters
## 4 of the homes were assigned different grades in each winter
### Was there an upgrade/downgrade to the stove? Check for comments/notes.
```{r}
multiple_grades <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(!is.na(stove_grade)) %>%
  group_by(home) %>% 
  filter(n() > 1) %>% 
  arrange(home, winter_id)
kable(multiple_grades)
```

# Summary of stove use
```{r}
# Summarize numeric vars
stove_use_summary <- stove_use_clean %>% 
  group_by(area) %>% 
  summarize("N" = n(),
            "Mean Cords" = mean(wood_cords, na.rm = TRUE),
            "SD Cords" = sd(wood_cords, na.rm = TRUE),
            "Min Cords" = min(wood_cords, na.rm = TRUE), 
            "Median Cords" = median(wood_cords, na.rm = TRUE), 
            "Max Cords" = max(wood_cords, na.rm = TRUE))
kable(stove_use_summary, digits = 1, align = "c")

## Select specific var to assess individually/group by area
table(stove_use_clean$burn_level, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_type, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_collect, stove_use_clean$area, useNA = "ifany")
table(stove_use_clean$wood_collect_method, stove_use_clean$area, useNA = "ifany")
```

# Stove use summary plots
```{r}
stove_use_plot_function <- function(data, variable) {

stove_use_plot <- data %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
stove_use_plot

}

stove_use_plot_function(stove_use_clean, "burn_level")
stove_use_plot_function(stove_use_clean, "wood_type")
stove_use_plot_function(stove_use_clean, "wood_collect")
stove_use_plot_function(stove_use_clean, "wood_collect_method")
```

### 2020-01-16 dustrak_data_cleaning.Rmd
commit from new computer

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 10 Jan 2020"
+date: "Started 19 Nov 2019, Updated 14 Jan 2020"
 output: pdf_document
@@
 ```
+
+# Save cleaned dataset
+```{r}
+kids_pm_new <- read_rds("Output/kids_pm_new.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+
+pm_clean <- kids_pm_new %>% 
+  mutate(home_winter_id = as.character(home_winter_id)) %>% 
+  select(-home, -treatment) %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+
+# write_rds(pm_clean, "Output/pm_clean.rds")
+```

### 2020-01-16 home_char_act_summary.Rmd
commit from new computer

@@
 date: "Started 15 Jan 2020, Updated 15 Jan 2020"
-output: pdf_document
+output:
+  word_document: default
+  pdf_document: default
 ---

### 2020-01-16 home_child_data_cleaning.Rmd
SUMs cleaning
commit from new computer

@@
 
-#write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
+# write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
 
@@
 ```
+
+# Clean demographics data
+```{r}
+kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
+
+demographics_clean <- kids_demographics %>% 
+  separate(intervention_time, c("trash", "intervention_time"), sep = " ") %>% 
+  unite(intervention_datetime, c("intervention_date", "intervention_time"),
+        sep = " ", remove = FALSE) %>% 
+  mutate(intervention_datetime = ymd_hms(intervention_datetime)) %>% 
+  select(-trash)
+
+# write_rds(demographics_clean, "Output/demographics_clean.rds")
+```

### 2020-01-16 indoor_temp_rh_cleaning.Rmd
commit from new computer

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 14 Jan 2020"
+date: "Started 13 Jan 2020, Updated 15 Jan 2020"
 output: pdf_document
@@
 ```
-
-# Summarize temp/rh data
-```{r}
-indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
-
-temp_rh_summary <- indoor_temp_rh_clean %>%
-  group_by(area) %>% 
-  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
-                                med = median, max = max), na.rm = TRUE) %>% 
-  t()
-temp_rh_summary
-```

### 2020-01-16 pm_summary.tex
commit from new computer

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Kids PM Summary},
  pdfauthor={Ethan Walker},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Kids PM Summary}
\author{Ethan Walker}
\date{Started 22 Nov 2019, Updated 5 Dec 2019}

\begin{document}
\maketitle

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(readr)}
\KeywordTok{library}\NormalTok{(knitr)}
\KeywordTok{library}\NormalTok{(lubridate)}
\NormalTok{jvPalette <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"#330099"}\NormalTok{,}\StringTok{"#CC0066"}\NormalTok{,}\StringTok{"#FF6633"}\NormalTok{, }
                 \StringTok{"#0099CC"}\NormalTok{, }\StringTok{"#FF9900"}\NormalTok{,}\StringTok{"#CC6633"}\NormalTok{,}
                  \StringTok{"#FF3366"}\NormalTok{, }\StringTok{"#33CC99"}\NormalTok{, }\StringTok{"#33999"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Load data}
\NormalTok{kids_pm <-}\StringTok{ }\KeywordTok{read_rds}\NormalTok{(}\StringTok{"Output/kids_pm_new.rds"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{summary-stats}{%
\section{Summary stats}\label{summary-stats}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# summary by area}
\NormalTok{summary_area <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}
\StringTok{  }\CommentTok{#filter(actual_vs_expected == 1) %>% }
\StringTok{  }\CommentTok{#filter(pm <0) %>% }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(home_winter_id) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\StringTok{"n"}\NormalTok{ =}\StringTok{ }\KeywordTok{n}\NormalTok{(),}
            \StringTok{"Mean PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{mean}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"SD PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{sd}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Min PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{min}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"Median PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{median}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Max PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{max}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Mean Sample Days"}\NormalTok{ =}\StringTok{ }\KeywordTok{mean}\NormalTok{(pm_sample_interval, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Min Sample Days"}\NormalTok{ =}\StringTok{ }\KeywordTok{min}\NormalTok{(pm_sample_interval, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"Max Sample Days"}\NormalTok{ =}\StringTok{ }\KeywordTok{max}\NormalTok{(pm_sample_interval, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Mean Sample Obs"}\NormalTok{ =}\StringTok{ }\KeywordTok{mean}\NormalTok{(pm_total_observations, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Min Sample Obs"}\NormalTok{ =}\StringTok{ }\KeywordTok{min}\NormalTok{(pm_total_observations, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"Max Sample Obs"}\NormalTok{ =}\StringTok{ }\KeywordTok{max}\NormalTok{(pm_total_observations, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{t}\NormalTok{()}
\NormalTok{summary <-}\StringTok{ }\KeywordTok{as.data.frame}\NormalTok{(summary_area)}
\NormalTok{my_names <-}\StringTok{ }\NormalTok{summary_area[}\DecValTok{1}\NormalTok{, ]}
\KeywordTok{colnames}\NormalTok{(summary) <-}\StringTok{ }\NormalTok{my_names}
\NormalTok{summary_area <-}\StringTok{ }\NormalTok{summary }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rownames_to_column}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{Statistic =}\NormalTok{ rowname) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(Statistic }\OperatorTok{!=}\StringTok{ "area"}\NormalTok{) }
\CommentTok{#summary_area}

\CommentTok{# summary by home_winter_id}
\NormalTok{summary_id <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(area, home_winter_id) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\StringTok{"n"}\NormalTok{ =}\StringTok{ }\KeywordTok{n}\NormalTok{(),}
            \StringTok{"Mean PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{mean}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"SD PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{sd}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Min PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{min}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"Median PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{median}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Max PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{max}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Act vs Exp Diff"}\NormalTok{ =}\StringTok{ }\KeywordTok{first}\NormalTok{(}\KeywordTok{as.numeric}\NormalTok{(actual_vs_expected_diff)),}
            \StringTok{"Act vs Exp"}\NormalTok{ =}\StringTok{ }\KeywordTok{first}\NormalTok{(actual_vs_expected),}
            \StringTok{"Comments"}\NormalTok{ =}\StringTok{ }\KeywordTok{first}\NormalTok{(qaqccomments))}
\KeywordTok{write_csv}\NormalTok{(summary_id, }\StringTok{"output/kids_summary_id.csv"}\NormalTok{)}

\CommentTok{# summary by dusttrak id}
\NormalTok{summary_dtid <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(area, dtid) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\StringTok{"n"}\NormalTok{ =}\StringTok{ }\KeywordTok{n}\NormalTok{(),}
            \StringTok{"Mean PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{mean}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"SD PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{sd}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Min PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{min}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{), }
            \StringTok{"Median PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{median}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{),}
            \StringTok{"Max PM"}\NormalTok{ =}\StringTok{ }\KeywordTok{max}\NormalTok{(pm, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{))}

\NormalTok{histogram_id <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(home_winter_id }\OperatorTok{==}\StringTok{ "10"} \OperatorTok{&}\StringTok{ }\NormalTok{area }\OperatorTok{==}\StringTok{ "AK"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }
\StringTok{    }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(pm), }\DataTypeTok{bins =} \DecValTok{50}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{()}
\NormalTok{histogram_id}
\end{Highlighting}
\end{Shaded}

\includegraphics{pm_summary_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{##### 2 Functions to look at PM data across time #####}

\CommentTok{# Plot PM by datetime to look at trends, gaps, and peaks}
\NormalTok{pm_time_plot_function <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(id, location) \{}
\NormalTok{pm_time_plot <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(home_winter_id }\OperatorTok{==}\StringTok{ }\NormalTok{id }\OperatorTok{&}\StringTok{ }\NormalTok{area }\OperatorTok{==}\StringTok{ }\NormalTok{location) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\CommentTok{#filter(pm < 30000) %>% }
\StringTok{  }\KeywordTok{arrange}\NormalTok{(pm_datetime) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{pm_sampling_day =} \KeywordTok{as.character}\NormalTok{(pm_sampling_day),}
         \DataTypeTok{day_of_week =} \KeywordTok{factor}\NormalTok{(day_of_week,}
                              \DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Sunday"}\NormalTok{, }\StringTok{"Monday"}\NormalTok{, }\StringTok{"Tuesday"}\NormalTok{, }\StringTok{"Wednesday"}\NormalTok{, }
                                         \StringTok{"Thursday"}\NormalTok{, }\StringTok{"Friday"}\NormalTok{, }\StringTok{"Saturday"}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }
\StringTok{    }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(pm_datetime, pm, }\DataTypeTok{color =}\NormalTok{ day_of_week), }\DataTypeTok{size =} \FloatTok{1.5}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_classic}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =}\NormalTok{ id,}
         \DataTypeTok{y =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"PM"}\NormalTok{[}\FloatTok{2.5}\NormalTok{], }\StringTok{" ("}\NormalTok{, mu, g}\OperatorTok{/}\NormalTok{m}\OperatorTok{^}\DecValTok{3}\NormalTok{, }\StringTok{")"}\NormalTok{)),}
         \DataTypeTok{x =} \StringTok{"Datetime"}\NormalTok{,}
         \DataTypeTok{color =} \StringTok{"Day of week"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.title.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{12}\NormalTok{,}
                                      \DataTypeTok{margin =} \KeywordTok{margin}\NormalTok{(}\DataTypeTok{t =} \DecValTok{0}\NormalTok{, }\DataTypeTok{r =} \DecValTok{20}\NormalTok{, }\DataTypeTok{b =} \DecValTok{0}\NormalTok{, }\DataTypeTok{l =} \DecValTok{0}\NormalTok{)),}
          \DataTypeTok{axis.title.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{12}\NormalTok{),}
          \DataTypeTok{axis.text.x =} \KeywordTok{element_blank}\NormalTok{(),}
          \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{12}\NormalTok{, }\DataTypeTok{color =} \StringTok{"black"}\NormalTok{),}
          \DataTypeTok{axis.line.x =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{), }
          \DataTypeTok{axis.line.y =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{), }
          \DataTypeTok{axis.ticks =} \KeywordTok{element_blank}\NormalTok{()) }\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values =}\NormalTok{ jvPalette)}

\NormalTok{gaps <<-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(home_winter_id }\OperatorTok{==}\StringTok{ }\NormalTok{id }\OperatorTok{&}\StringTok{ }\NormalTok{area }\OperatorTok{==}\StringTok{ }\NormalTok{location) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{pm_datetime =} \KeywordTok{ymd_hms}\NormalTok{(pm_datetime),}
         \DataTypeTok{datetime_lead =} \KeywordTok{ymd_hms}\NormalTok{(}\KeywordTok{lead}\NormalTok{(pm_datetime)),}
         \DataTypeTok{datetime_1 =} \KeywordTok{seconds}\NormalTok{(pm_datetime),}
         \DataTypeTok{datetime_2 =} \KeywordTok{seconds}\NormalTok{(datetime_lead),}
         \DataTypeTok{datetime_diff_minutes =} \KeywordTok{as.numeric}\NormalTok{(datetime_}\DecValTok{2} \OperatorTok{-}\StringTok{ }\NormalTok{datetime_}\DecValTok{1}\NormalTok{)}\OperatorTok{/}\DecValTok{60}\NormalTok{,}
         \DataTypeTok{datetime_diff_hours =} \KeywordTok{as.numeric}\NormalTok{(datetime_diff_minutes}\OperatorTok{/}\DecValTok{60}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{arrange}\NormalTok{(}\KeywordTok{desc}\NormalTok{(datetime_diff_minutes)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, }
\NormalTok{         pm_datetime, datetime_lead)}

\NormalTok{pm_desc <<-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(home_winter_id }\OperatorTok{==}\StringTok{ }\NormalTok{id }\OperatorTok{&}\StringTok{ }\NormalTok{area }\OperatorTok{==}\StringTok{ }\NormalTok{location) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{arrange}\NormalTok{(}\KeywordTok{desc}\NormalTok{(pm)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(pm_datetime, pm, pm_sampling_day)}

\NormalTok{pm_time_plot}
\NormalTok{\}}

\CommentTok{# Change function input to look at specific home}
\CommentTok{# Run all of these lines together}
\KeywordTok{pm_time_plot_function}\NormalTok{(}\DataTypeTok{id =} \StringTok{"228"}\NormalTok{, }\DataTypeTok{location =} \StringTok{"NN"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Note: method with signature 'Period#ANY' chosen for function '-',
##  target signature 'Period#Period'.
##  "ANY#Period" would also be valid
\end{verbatim}

\includegraphics{pm_summary_files/figure-latex/unnamed-chunk-3-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(gaps, }\DecValTok{25}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 25 x 6
##    day_of_week pm_sampling_day datetime_diff_m~ datetime_diff_h~
##    <chr>                 <dbl>            <dbl>            <dbl>
##  1 Wednesday                 1                1           0.0167
##  2 Wednesday                 1                1           0.0167
##  3 Wednesday                 1                1           0.0167
##  4 Wednesday                 1                1           0.0167
##  5 Wednesday                 1                1           0.0167
##  6 Wednesday                 1                1           0.0167
##  7 Wednesday                 1                1           0.0167
##  8 Wednesday                 1                1           0.0167
##  9 Wednesday                 1                1           0.0167
## 10 Wednesday                 1                1           0.0167
## # ... with 15 more rows, and 2 more variables: pm_datetime <dttm>,
## #   datetime_lead <dttm>
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(pm_desc, }\DecValTok{25}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 25 x 3
##    pm_datetime            pm pm_sampling_day
##    <chr>               <dbl>           <dbl>
##  1 2019-03-07 19:38:49 5485.               2
##  2 2019-03-07 00:32:49 3394.               1
##  3 2019-03-06 21:12:49 2582.               1
##  4 2019-03-07 21:02:49 2515.               2
##  5 2019-03-07 09:54:49 2061.               1
##  6 2019-03-09 21:05:49 1982.               4
##  7 2019-03-09 13:08:49 1952.               3
##  8 2019-03-07 11:01:49 1939.               1
##  9 2019-03-08 18:51:49 1903.               3
## 10 2019-03-07 19:39:49 1879.               2
## # ... with 15 more rows
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{tail}\NormalTok{(pm_desc, }\DecValTok{25}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 25 x 3
##    pm_datetime            pm pm_sampling_day
##    <chr>               <dbl>           <dbl>
##  1 2019-03-08 09:14:49  2.42               2
##  2 2019-03-08 09:15:49  2.42               2
##  3 2019-03-08 09:16:49  2.42               2
##  4 2019-03-08 09:17:49  2.42               2
##  5 2019-03-08 09:18:49  2.42               2
##  6 2019-03-08 09:19:49  2.42               2
##  7 2019-03-08 09:20:49  2.42               2
##  8 2019-03-08 09:21:49  2.42               2
##  9 2019-03-08 09:22:49  2.42               2
## 10 2019-03-08 09:23:49  2.42               2
## # ... with 15 more rows
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# filter data for a single home}
\NormalTok{check <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(area }\OperatorTok{==}\StringTok{ "NN"} \OperatorTok{&}\StringTok{ }\NormalTok{home_winter_id }\OperatorTok{==}\StringTok{ "224"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(pm }\OperatorTok{<}\StringTok{ }\DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{boxplots}{%
\section{Boxplots}\label{boxplots}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{boxplots_kids_pm <-}\StringTok{ }\NormalTok{kids_pm }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(pm }\OperatorTok{<}\StringTok{ }\DecValTok{500}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(area, pm)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_boxplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =} \DecValTok{1}\NormalTok{), }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{, }\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }
               \DataTypeTok{fatten =} \DecValTok{1}\NormalTok{, }\DataTypeTok{outlier.size =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{width =} \FloatTok{0.35}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{stat_summary}\NormalTok{(}\DataTypeTok{fun.y=}\NormalTok{mean, }\DataTypeTok{geom=}\StringTok{"point"}\NormalTok{, }\DataTypeTok{shape=}\DecValTok{17}\NormalTok{, }\DataTypeTok{size=}\DecValTok{3}\NormalTok{, }\DataTypeTok{color=}\StringTok{"red"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_minimal}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{ylim}\NormalTok{(}\OperatorTok{-}\DecValTok{100}\NormalTok{, }\DecValTok{500}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\CommentTok{#geom_hline(yintercept = 50, color = "red", size = 1.2) +}
\StringTok{    }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"PM"}\NormalTok{[}\FloatTok{2.5}\NormalTok{], }\StringTok{" ("}\NormalTok{, mu, g}\OperatorTok{/}\NormalTok{m}\OperatorTok{^}\DecValTok{3}\NormalTok{, }\StringTok{")"}\NormalTok{))) }\OperatorTok{+}
\StringTok{    }\CommentTok{#labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{16}\NormalTok{, }\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{),}
          \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{16}\NormalTok{, }\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{),}
          \DataTypeTok{axis.title.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \DecValTok{16}\NormalTok{,}
                                      \DataTypeTok{margin =} \KeywordTok{margin}\NormalTok{(}\DataTypeTok{t =} \DecValTok{0}\NormalTok{, }\DataTypeTok{r =} \DecValTok{20}\NormalTok{, }\DataTypeTok{b =} \DecValTok{0}\NormalTok{, }\DataTypeTok{l =} \DecValTok{0}\NormalTok{)),}
          \DataTypeTok{axis.title.x =} \KeywordTok{element_blank}\NormalTok{(),}
          \DataTypeTok{axis.line.x =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{size =} \FloatTok{1.2}\NormalTok{), }
          \DataTypeTok{axis.line.y =} \KeywordTok{element_line}\NormalTok{(}\DataTypeTok{colour =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{size =} \FloatTok{1.2}\NormalTok{), }
          \DataTypeTok{axis.ticks =} \KeywordTok{element_blank}\NormalTok{())}
\NormalTok{boxplots_kids_pm}
\end{Highlighting}
\end{Shaded}

\includegraphics{pm_summary_files/figure-latex/unnamed-chunk-4-1.pdf}

\end{document}

### 2020-01-16 stove_grade_usage_cleaning.Rmd
commit from new computer

@@
 
-# Summary data for stove grades
-```{r}
-# Load dataset
-stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
-
-stove_grade_summary <- stove_grades_clean %>% 
-  group_by(area, stove_grade) %>% 
-  summarize(n())
-stove_grade_summary
-```
-
 ###############################################
@@
   replace_with_na(replace = list(wood_cords = -999, wood_cords = 0)) %>% 
+  mutate(wood_collect = factor(wood_collect,
+                               levels = c("Less than 1 week",
+                                          "Between 1 week and 1 month",
+                                          "Between 1 month and 3 months",
+                                          "Between 3 months and 6 months",
+                                          "Between 6 months and 1 year",
+                                          "More than 1 year"),
+                               labels = c("< 1 week",
+                                          "1 week to 1 mo",
+                                          "1 mo to 3 mo",
+                                          "3 mo to 6 mo",
+                                          "6 mo to 1 year",
+                                          "> 1 yr"))) %>% 
   left_join(kids_linked_ids, by = c("area", "home_winter_id"))
@@
 ```
-
-# Summary data for stove use
-```{r}
-# Load dataset
-stove_use_clean <- read_rds("Output/stove_use_clean.rds")
-
-# Summarize numeric vars
-stove_use_summary <- stove_use_clean %>% 
-  group_by(area) %>% 
-  summarize(mean(wood_cords, na.rm = TRUE), 
-            min(wood_cords, na.rm = TRUE), 
-            median(wood_cords, na.rm = TRUE), 
-            max(wood_cords, na.rm = TRUE))
-stove_use_summary
-
-# Summarize factor vars
-## Use this as a quick look at all vars
-table_list <- stove_use_clean %>% 
-  #filter(area == "WMT") %>% 
-  select(-home_winter_id, -wood_cords) 
-lapply(table_list, table)
-
-## Select specific var to assess individually/group by area
-table(stove_use_clean$burn_level, stove_use_clean$area, useNA = "ifany")
-```

### 2020-01-16 sums_summary.Rmd
commit from new computer

@@
 author: "Ethan Walker"
-date: "Started 9 Jan 2020, Updated 9 Jan 2020"
+date: "Started 9 Jan 2020, Updated 15 Jan 2020"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
+                      include = TRUE, warning = FALSE)
 ```
@@
 
-# Summary stats
+# Summary stats by area (winter)
 ```{r}
@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(area) %>% 
+  group_by(area, winter_id) %>% 
   summarize("n" = n(),
@@
             "Median Temp" = median(temp_c_sums, na.rm = TRUE),
-            "Max Temp" = max(temp_c_sums, na.rm = TRUE))
+            "Max Temp" = max(temp_c_sums, na.rm = TRUE)) %>% 
+  select(-area, -winter_id) %>% 
+  t()
+
+kable(sums_summary, col.names = c("Alaska (1)", "Alaska (2)",
+                                  "Navajo (1)", "Navajo (2)",
+                                  "Montana (1)", "Montana (2)"), 
+      digits = 1, align = "c")
 ```
 
-# Boxplots for PM by study area
+# Boxplots for SUMs temp (C) by study area - Winter 1
 ```{r, fig.height=6, fig.width=5}
 boxplots_kids_sums <- sums_clean %>%
+  filter(winter_id == 1) %>% 
   ggplot(aes(area, temp_c_sums)) +
@@
 ```
+
+# Boxplots for SUMs temp (C) by study area - Winter 2
+```{r, fig.height=6, fig.width=5}
+boxplots_kids_sums <- sums_clean %>%
+  filter(winter_id == 2) %>% 
+  ggplot(aes(area, temp_c_sums)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    theme_minimal() +
+    ylim(-50, 100) +
+    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
+    labs(y = "Temperature (degrees C)") +
+    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
+    theme(axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_kids_sums
+```

### 2020-01-16 indoor_temp_rh_summary.Rmd
minor edit

---
title: "Kids indoor temp/RH summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 16 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

```{r}
# Load data
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
```

# Summary stats
```{r}
temp_rh_summary <- indoor_temp_rh_clean %>%
  group_by(area) %>% 
  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
                                med = median, max = max), na.rm = TRUE) %>% 
  select(-area) %>% 
  t() 
kable(temp_rh_summary, col.names = c("Alaska", "Navajo", "Montana"), 
      digits = 1, align = "c")
```

```{r, fig.height=6, fig.width=5}
# Boxplot function
boxplots_temp_rh_function <- function(data, variable){

boxplots_temp_rh <- data %>% 
  # change input to plot different temp/RH variable
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  na.exclude() %>% 
  ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    #ylim(0, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    #labs(y = "Max indoor RH (%)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_temp_rh

}
```

## Current indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current")
```

## Max indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max")
```

## Min indoor humidity (%)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min")
```

## Current indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current")
```

## Max indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max")
```

## Min indoor temperature (C)
```{r}
boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min")
```

### 2020-01-16 pm_summary.Rmd

@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(treatment) %>% 
+  group_by(area) %>% 
   summarize("n" = n(),
@@
             "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
-            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE))
+            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
+  select(-area) %>% 
+  t()
+
+kable(pm_summary, col.names = c("Alaska", "Navajo", "Montana"), 
+      digits = 1, align = "c")
 ```

### 2020-01-17 home_char_act_summary.Rmd
data cleaning

--- 2020-01-16 home_char_act_summary.Rmd
@@
 date: "Started 15 Jan 2020, Updated 15 Jan 2020"
-output: pdf_document
+output:
+  pdf_document: default
+  word_document: default
 ---

### 2020-01-17 join_format_kids_data.Rmd
data cleaning

@@
 
-# Join large datasets
+# Join SUMs and DustTrak datasets
+## Fix vars/dates in data dictionary once finalized
 ```{r}
+pm_join <- read_rds("Output/pm_clean.rds") %>% 
+  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
+  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  group_by(area, home) %>% 
+  arrange(sample_datetime) %>% 
+  mutate(pm_last_datetime = last(sample_datetime),
+         pm_first_datetime = first_datetime_new,
+         sample_datetime = round_date(sample_datetime, unit = "minute"),
+         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime)
 
+sums_join <- read_rds("Output/sums_clean.rds") %>% 
+  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
+  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime)
+
+pm_sums_joined <- sums_join %>% 
+  left_join(pm_join, by = c("area", "home", "sample_datetime")) %>% 
+  filter(pm_rolling_20 < 1000)
+
+cor.test(pm_sums_joined$temp_c_sums, pm_sums_joined$pm_rolling_20, method = "spearman")
+pm_sums_model <- lm(log(pm_sums_joined$pm_rolling_20) ~ pm_sums_joined$temp_c_sums)
+
+summary(pm_sums_model)
+exp((0.0235293)-1)*100
+plot(pm_sums_model)
+plot(pm_sums_joined$pm_rolling_20, pm_sums_joined$temp_c_sums)
 ```

### 2020-01-17 pm_summary.Rmd
data cleaning

@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
+  filter(pm < 500) %>% 
   group_by(area) %>% 
@@
 boxplots_kids_pm <- pm_clean %>% 
-  #filter(pm < 500) %>% 
+  filter(pm < 500) %>% 
   ggplot(aes(area, pm)) +
@@
     theme_minimal() +
-    ylim(0, 100) +
+    ylim(0, 500) +
     #geom_hline(yintercept = 50, color = "red", size = 1.2) +

### 2020-01-21 dustrak_data_cleaning.Rmd
data cleaning updates

--- 2020-01-10 dustrak_data_cleaning.Rmd
@@
 
-# Save pm dataset as RDS and CSV
+# Save pm dataset as RDS 
 # write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
-# write_csv(kids_pm, "Output/kids_pm_full.csv")
 ```
@@
 ##### Add new variables to Kids PM data #####
-kids_pm_new <- kids_pm_joined %>%  
+kids_pm_new1 <- kids_pm_joined %>%  
   #filter(area == "AK") %>% 
@@
 
-# Save pm dataset as RDS and CSV
-# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
-# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
+# Save pm dataset as RDS 
+# write_rds(kids_pm_new1, "Output/kids_pm_new1.rds")
 ```
@@
 # Load data 
-kids_pm <- read_rds("Output/kids_pm_new.rds")
+kids_pm <- read_rds("Output/kids_pm_new1.rds")
 
 # Checking difference between dustrak start datetime and field data log datetime
-kids_pm_new <- kids_pm %>% 
+kids_pm_new2 <- kids_pm %>% 
   #filter(area == "NN") %>% 
@@
 
-# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+# write_rds(kids_pm_new2, "Output/kids_pm_new2.rds")
 ```
@@
 # Load data 
-kids_pm <- read_rds("Output/kids_pm_new.rds")
+kids_pm <- read_rds("Output/kids_pm_new2.rds")
 
-kids_pm_new <- kids_pm %>%
+kids_pm_new3 <- kids_pm %>%
   # filtering out specific homes where PM data is not suitable for analysis
@@
 
-# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
+# write_rds(kids_pm_new3, "Output/kids_pm_new3.rds")
 ```
@@
 # Load data 
-kids_pm <- read_rds("Output/kids_pm_new.rds")
+kids_pm <- read_rds("Output/kids_pm_new3.rds")
 cleaning_summary <- read_xlsx("Input/cleaning_summary.xlsx") %>% 
@@
 
-kids_pm_new <- kids_pm %>% 
+kids_pm_new4 <- kids_pm %>% 
   mutate(home_winter_id = as.numeric(home_winter_id)) %>% 
@@
 
-# write_rds(kids_pm_new, "Output/kids_pm_new.rds")
-# write_csv(kids_pm_new, "Output/kids_pm_new.csv")
+# write_rds(kids_pm_new4, "Output/kids_pm_new4.rds")
 ```
+
+# Save cleaned dataset
+```{r}
+kids_pm <- read_rds("Output/kids_pm_new4.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+
+kids_pm_new5 <- kids_pm %>% 
+  mutate(home_winter_id = as.character(home_winter_id)) %>% 
+  select(-home, -treatment) %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+
+# write_rds(kids_pm_new5, "Output/pm_clean.rds")
+```

### 2020-01-21 home_child_data_cleaning.Rmd
data cleaning updates

@@
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+  mutate(winter_id = as.numeric(winter_id),
+         winter_id = if_else(home == "CH218", winter_id + 1, winter_id)) %>% 
+  filter(winter_id > 0)
+
+#write_rds(kids_linked_ids, "Output/kids_linked_ids.rds")
 ```

### 2020-01-21 placebo_exposure_data_cleaning.Rmd
data cleaning updates

---
title: "KidsAIR: placebo home exposure paper data cleaning"
author: "Ethan Walker"
date: "Started 21 Jan 2020, Updated 21 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Filter winter 1 and placebo homes from PM dataset
## These homes will be the focus of the initial exposure abstract/paper
```{r}
pm_homes <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Join selected homes from above with other datasets and filter homes of interest

# PM data
## 93 placebo homes from Winter 1 have PM data
```{r}
pm_filtered <- pm_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# SUMs data
## 78 placebo homes from Winter 1 have SUMs data
#### Look for missing files/data? ####
```{r}
sums_filtered <- sums_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)

sums_homes <- sums_clean %>% 
  mutate(sums_home = "true") %>% 
  left_join(pm_homes, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(sums_pm_match = "true") %>% 
  select(home, area, home_winter_id, winter_id, home_id, treatment, 
         sums_home, sums_pm_match)

sums_pm_check <- pm_homes %>% 
  left_join(sums_homes, by = c("area", "home", "home_winter_id")) %>% 
  filter(is.na(sums_pm_match))

# Should have SUMs data for YKC06113
# Should have SUMs data for CH107, CH218
# No files for CH102, CH103 (file but no content), CH106, TC357 
# Should have SUMs data for WMA102, WMA132, WMA141
# No files for WMA507, WMA673
# Download files for WMA114, WMA121, WMA125

check <- sums_clean %>% 
  filter(home == "YKC06113")
```

# Demographics data
## 91 placebo homes from Winter 1 have demographic data
#### Look for missing files/data? ####
```{r}
demographics_filtered <- demographics_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# Moisture data
## 60 placebo homes from Winter 1 have moisture data
#### Look for missing files/data? ####
```{r}
moisture_filtered <- moisture_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# Indoor temp/rh data
## 93 placebo homes from Winter 1 have indoor temp/rh data
#### Check for missing data ####
```{r}
indoor_temp_rh_filtered <- indoor_temp_rh_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# Stove use data
## 93 placebo homes from Winter 1 have stove use data
#### Check for missing data ####
```{r}
stove_use_filtered <- stove_use_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# Stove grades data
## 93 placebo homes from Winter 1 have stove grades data
#### Check for missing grades ####
```{r}
stove_grades_filtered <- stove_grades_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home, stove_grade) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# Home characteristics data
## 93 placebo homes from Winter 1 have home characteristics data
#### Continue cleaning up data ####
```{r}
home_char_filtered <- home_char_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

# Home activity data
## 93 placebo homes from Winter 1 have home activity data
#### Continue cleaning up data; make sure there are 6 obs per home ####
```{r}
home_act_filtered <- home_act_clean %>% 
  left_join(pm_homes, by = c("area", "home")) %>% 
  filter(select_home == "select") %>% 
  filter(winter_id == 1) %>% 
  #group_by(area, home) %>% 
  #summarize(n())
  ungroup() %>% 
  arrange(area, home)
```

### 2020-01-21 sums_data_cleaning.Rmd
data cleaning updates

@@
 ```
+
+# Check to see if there are missing files to search for
+```{r}
+# Load SUMs data
+sums_clean <- read_rds("Output/sums_clean.rds")
+# Load SUMs data logs
+wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds") %>% 
+  mutate(area = "WMT")
+nn_sums_log <- read_rds("Output/nn_sums_data_log.rds") %>% 
+  mutate(area = "NN")
+ak_sums_log <- read_rds("Output/ak_sums_data_log.rds") %>% 
+  mutate(area = "AK")
+
+# Combine SUMs logs from all areas
+all_sums_log <- rbind(wmt_sums_log, nn_sums_log, ak_sums_log) %>% 
+  mutate(winter_id = as.character(winter_id),
+         sums_log = "true")
+
+sums_check <- sums_clean %>% 
+  group_by(area, home, winter_id) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  full_join(all_sums_log, by = c("area", "home", "winter_id")) %>% 
+  ungroup() %>% 
+  arrange(area, home, winter_id)
+
+write_csv(sums_check, "Output/sums_check.csv")
+```

### 2020-01-22 home_child_data_cleaning.Rmd
data cleaning updates

@@
 wmt_child_demographics <- joined1 %>% 
-  distinct(child_id_num, .keep_all = TRUE) %>% 
+  #distinct(child_id_num, .keep_all = TRUE) %>% 
   arrange(home_id, child_id_num, winter_id) %>% 
-  filter(enrolled == 1)
+  filter(winter_id == 1 | winter_id == 2)
 
@@
 write_rds(wmt_child_demographics, "Output/wmt_demographic_data.rds")
-write_csv(wmt_home_ids_linked, "Output/wmt_ids_linked.csv")
-write_csv(wmt_child_demographics, "Output/wmt_demographic_data.csv")
 ```
@@
 nn_child_demographics <- joined1 %>% 
-  distinct(child_id_num, .keep_all = TRUE) %>% 
+  #distinct(child_id_num, .keep_all = TRUE) %>% 
   arrange(home_id, child_id_num, winter_id) %>% 
-  filter(enrolled == 1)
+  filter(winter_id == 1 | winter_id == 2)
 
@@
 write_rds(nn_child_demographics, "Output/nn_demographic_data.rds")
-write_csv(nn_home_ids_linked, "Output/nn_ids_linked.csv")
-write_csv(nn_child_demographics, "Output/nn_demographic_data.csv")
 ```
@@
 ak_child_demographics <- joined1 %>% 
-  distinct(child_id_num, .keep_all = TRUE) %>% 
+  #distinct(child_id_num, .keep_all = TRUE) %>% 
   arrange(home_id, child_id_num, winter_id) %>% 
-  filter(enrolled == 1)
+  filter(winter_id == 1 | winter_id == 2)
 
@@
 write_rds(ak_child_demographics, "Output/ak_demographic_data.rds")
-write_csv(ak_home_ids_linked, "Output/ak_ids_linked.csv")
-write_csv(ak_child_demographics, "Output/ak_demographic_data.csv")
 ```

### 2020-01-22 placebo_exposure_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 21 Jan 2020, Updated 21 Jan 2020"
+date: "Started 21 Jan 2020, Updated 22 Jan 2020"
 output: pdf_document
@@
 ```{r}
-pm_homes <- pm_clean %>% 
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
   filter(winter_id == 1) %>% 
@@
   arrange(area, home)
+
+# 88 Winter 1 education homes have PM data
+pm_homes_education <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Education") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+# 86 Winter 1 filter homes have PM data
+pm_homes_filter <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Filter") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
 ```
 
-# Join selected homes from above with other datasets and filter homes of interest
+#### Join selected homes from above with other datasets and filter homes of interest ####
 
@@
 ## 93 placebo homes from Winter 1 have PM data
+## 88 education homes from Winter 1 have PM data
+## 86 filter homes from Winter 1 have PM data
+## All data is accounted for; missing data is truly missing from data collection issues
 ```{r}
 pm_filtered <- pm_clean %>% 
-  left_join(pm_homes, by = c("area", "home")) %>% 
-  filter(select_home == "select") %>% 
+  left_join(pm_homes_placebo, by = c("area", "home")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+pm_obs <- pm_clean %>% 
+  left_join(pm_homes_placebo, by = c("area", "home")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, home) %>% 
+  summarize(n())
+```
+
+# SUMs data
+## 80/93 placebo homes from Winter 1 have SUMs data
+## 77/88 education homes from Winter 1 have SUMs data
+## 70/86 filter homes from Winter 1 have SUMs data
+## All data is accounted for; missing data is truly missing from data collection issues
+```{r}
+sums_filtered <- sums_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_filter, by = c("area", "home")) %>% 
+  filter(pm_home == "true") %>% 
   filter(winter_id == 1) %>% 
@@
 
-# SUMs data
-## 78 placebo homes from Winter 1 have SUMs data
-#### Look for missing files/data? ####
+# Demographics data
+## 93/93 placebo homes from Winter 1 have demographic data
+## 88/88 education homes from Winter 1 have demographic data
+## 86/86 filter homes from Winter 1 have demographic data
 ```{r}
-sums_filtered <- sums_clean %>% 
-  left_join(pm_homes, by = c("area", "home")) %>% 
-  filter(select_home == "select") %>% 
-  filter(winter_id == 1) %>% 
-  #group_by(area, home) %>% 
-  #summarize(n())
-  ungroup() %>% 
-  arrange(area, home)
-
-sums_homes <- sums_clean %>% 
-  mutate(sums_home = "true") %>% 
-  left_join(pm_homes, by = c("area", "home", "home_winter_id")) %>% 
+demographics_check1 <- demographics_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_filter, by = c("area", "home")) %>% 
   filter(pm_home == "true") %>% 
   filter(winter_id == 1) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  mutate(sums_pm_match = "true") %>% 
-  select(home, area, home_winter_id, winter_id, home_id, treatment, 
-         sums_home, sums_pm_match)
+  group_by(area, home) %>% 
+  summarize(n())
 
-sums_pm_check <- pm_homes %>% 
-  left_join(sums_homes, by = c("area", "home", "home_winter_id")) %>% 
-  filter(is.na(sums_pm_match))
+demographics_check2 <- demographics_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_filter, by = c("area", "home")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, home) %>% 
+  summarize(n()) %>% 
+  mutate(dem_filtered = "true") %>% 
+  # change to correct joining data of interest
+  full_join(pm_homes_filter, by = c("area", "home"))
 
-# Should have SUMs data for YKC06113
-# Should have SUMs data for CH107, CH218
-# No files for CH102, CH103 (file but no content), CH106, TC357 
-# Should have SUMs data for WMA102, WMA132, WMA141
-# No files for WMA507, WMA673
-# Download files for WMA114, WMA121, WMA125
-
-check <- sums_clean %>% 
-  filter(home == "YKC06113")
-```
-
-# Demographics data
-## 91 placebo homes from Winter 1 have demographic data
-#### Look for missing files/data? ####
-```{r}
 demographics_filtered <- demographics_clean %>% 
-  left_join(pm_homes, by = c("area", "home")) %>% 
-  filter(select_home == "select") %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_filter, by = c("area", "home")) %>% 
+  filter(pm_home == "true") %>% 
   filter(winter_id == 1) %>% 
-  #group_by(area, home) %>% 
-  #summarize(n())
   ungroup() %>% 

### 2020-01-22 sums_data_cleaning.Rmd
data cleaning updates

--- 2020-01-14 sums_data_cleaning.Rmd
@@
 
-check_dates <- ak_sums_filtered %>% 
+check_dates <- wmt_sums_filtered %>% 
   group_by(home) %>% 
@@
 ```
+
+# Check to see if there are missing files to search for
+## Through 22 Jan 2020, all files are accounted for that are not documented as missing
+## Update once data collection is finished in April 2020
+```{r}
+# Load SUMs data
+sums_clean <- read_rds("Output/sums_clean.rds")
+
+check <- sums_clean %>% 
+  filter(home_winter_id == 82 & area == "AK")
+
+# Load SUMs data logs
+wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds") %>% 
+  mutate(area = "WMT")
+nn_sums_log <- read_rds("Output/nn_sums_data_log.rds") %>% 
+  mutate(area = "NN")
+ak_sums_log <- read_rds("Output/ak_sums_data_log.rds") %>% 
+  mutate(area = "AK")
+
+# Combine SUMs logs from all areas
+all_sums_log <- rbind(wmt_sums_log, nn_sums_log, ak_sums_log) %>% 
+  mutate(winter_id = as.character(winter_id),
+         sums_log = "true")
+
+sums_check <- sums_clean %>% 
+  group_by(area, home, winter_id) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  full_join(all_sums_log, by = c("area", "home", "winter_id")) %>% 
+  ungroup() %>% 
+  arrange(area, home, winter_id)
+
+#write_csv(sums_check, "Output/sums_check.csv")
+```

### 2020-01-23 filter_treatment_obs_check.Rmd
data cleaning

---
title: "KidsAIR: treatment filtering and missing data check"
author: "Ethan Walker"
date: "Started 21 Jan 2020, Updated 23 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Filter winter 1 and specific treatment homes from PM dataset
## These homes will be the focus of the initial exposure abstract/paper
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

# 88 Winter 1 education homes have PM data
pm_homes_education <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Education") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

# 86 Winter 1 filter homes have PM data
pm_homes_filter <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Filter") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)
```

#### Join selected homes from above with other datasets and filter homes of interest ####

# PM data
## 93 placebo homes from Winter 1 have PM data
## 88 education homes from Winter 1 have PM data
## 86 filter homes from Winter 1 have PM data
## All data is accounted for; missing data is truly missing from data collection issues
```{r}
pm_filtered <- pm_clean %>% 
  left_join(pm_homes_placebo, by = c("area", "home")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  ungroup() %>% 
  arrange(area, home)

pm_obs <- pm_clean %>% 
  left_join(pm_homes_placebo, by = c("area", "home")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n())
```

# SUMs data
## 80/93 placebo homes from Winter 1 have SUMs data
## 77/88 education homes from Winter 1 have SUMs data
## 70/86 filter homes from Winter 1 have SUMs data
## All data is accounted for; missing data is truly missing from data collection issues
```{r}
sums_check <- sums_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n())

sums_filtered <- sums_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Demographics data
## 93/93 placebo homes from Winter 1 have demographic data
## 88/88 education homes from Winter 1 have demographic data
## 86/86 filter homes from Winter 1 have demographic data
#### Some missing data within individual homes/rows; will be assessed in further data cleaning ####
#### Check for different answers across Winter 1 and Winter 2 ####
```{r}
demographics_check1 <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n())

demographics_check2 <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  summarize(n()) %>% 
  mutate(dem_filtered = "true") %>% 
  # change to correct joining data of interest
  full_join(pm_homes_placebo, by = c("area", "home"))

demographics_filtered <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Moisture data
## 91/93 placebo homes from Winter 1 have moisture data
## 83/88 education homes from Winter 1 have moisture data
## 83/86 filter homes from Winter 1 have moisture data
```{r}
moisture_check <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(moisture_ave)) %>% 
  group_by(area, home) %>% 
  summarize(n())

moisture_filtered <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(moisture_ave)) %>% 
  ungroup() %>% 
  arrange(area, home)

# Quick data check for odd values
min(moisture_filtered$moisture_ave)
max(moisture_filtered$moisture_ave)
```

# Indoor temp/rh data
## 75/93 placebo homes from Winter 1 have temp/rh data
## 70/88 education homes from Winter 1 have temp/rh data
## 76/86 filter homes from Winter 1 have temp/rh data
#### Some further missing data within individual homes/rows ####
```{r}
indoor_temp_rh_check <- indoor_temp_rh_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(rh_indoor_current) & is.na(rh_indoor_max) & is.na(rh_indoor_min) &
                             is.na(temp_indoor_current) & is.na(temp_indoor_max) & is.na(temp_indoor_min),
                             1, 0)) %>% 
  filter(na_filter == 0) %>% 
  group_by(area, home) %>% 
  summarize(n())

indoor_temp_rh_filtered <- indoor_temp_rh_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(rh_indoor_current) & is.na(rh_indoor_max) & is.na(rh_indoor_min) &
                             is.na(temp_indoor_current) & is.na(temp_indoor_max) & is.na(temp_indoor_min),
                             1, 0)) %>% 
  filter(na_filter == 0) %>% 
  ungroup() %>% 
  arrange(area, home)

# Quick data check for odd values
check_vars <- list(indoor_temp_rh_filtered$rh_indoor_current, indoor_temp_rh_filtered$rh_indoor_max, 
                   indoor_temp_rh_filtered$rh_indoor_min, indoor_temp_rh_filtered$temp_indoor_current, 
                   indoor_temp_rh_filtered$temp_indoor_max, indoor_temp_rh_filtered$temp_indoor_min)

lapply(check_vars, min, na.rm = TRUE)
lapply(check_vars, max, na.rm = TRUE)
```

# Stove use data
## 92/93 placebo homes from Winter 1 have stove use data
## 86/88 education homes from Winter 1 have stove use data
## 85/86 filter homes from Winter 1 have stove use data
#### Some further missing data within individual homes/rows ####
#### 5 homes reported using 20+ cords per year; confirmed by checking paper logs ####
```{r}
stove_use_check <- stove_use_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(burn_level) & is.na(wood_type) & 
                             is.na(wood_collect) & is.na(wood_collect_method),
                             1, 0)) %>% 
  filter(na_filter == 0) %>%
  group_by(area, home) %>% 
  summarize(n())

stove_use_filtered <- stove_use_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  mutate(na_filter = if_else(is.na(burn_level) & is.na(wood_type) & 
                             is.na(wood_collect) & is.na(wood_collect_method),
                             1, 0)) %>% 
  filter(na_filter == 0) %>%
  ungroup() %>% 
  arrange(area, home)

# Quick data check for odd values
min(stove_use_filtered$wood_cords, na.rm = TRUE)
max(stove_use_filtered$wood_cords, na.rm = TRUE)
```

# Stove grades data
## 90/93 placebo homes have stove grades data
## 83/88 education homes have stove grades data
## 79/86 filter homes have stove grades data
#### If Winter 1 was missing, the grade for Winter 2 was used ####
```{r}
stove_grades_check <- stove_grades_clean %>% 
  # change to correct joining data of interest
 left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  filter(!is.na(stove_grade)) %>% 
  group_by(area, home) %>% 
  summarize(n())

stove_grades_filtered <- stove_grades_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  filter(!is.na(stove_grade)) %>% 
  ungroup() %>% 
  arrange(area, home)

table(stove_grades_filtered$stove_grade)
```

# Home characteristics data
## Using answers from Winter 2 if Winter 1 = NA
## 92/93 placebo homes have home characteristics data
## 86/88 education homes have home characteristics data
## 85/86 filter homes have home characteristics data
#### Need to assess differences between Winter 1 and Winter 2 ####
```{r}
home_char_check <- home_char_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  group_by(area, home) %>% 
  mutate_all(funs(if_else(is.na(.), lead(.), .))) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  mutate(na_filter = if_else(is.na(residents_smoke) & is.na(home_type) & 
                             is.na(primary_heat) & is.na(stove_age),
                             1, 0)) %>%
  filter(na_filter == 0) %>%
  group_by(area, home) %>% 
  summarize(n())

home_char_filtered <- home_char_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(area, home) %>% 
  mutate_all(funs(if_else(is.na(.), lead(.), .))) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  mutate(na_filter = if_else(is.na(residents_smoke) & is.na(home_type) & 
                             is.na(primary_heat) & is.na(stove_age),
                             1, 0)) %>%
  filter(na_filter == 0) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Home activity data
## 93/93 placebo homes from Winter 1 have home activity data
### 92/93 have 6 observations; 1 has 5 observations
## 88/88 education homes from Winter 1 have home activity data
### 84/88 have 6 observations; 4 have 5 observations
## 86/86 filter homes from Winter 1 have home activity data
### 85/86 have 6 observations; 1 has 5 observations
```{r}
home_act_check <- home_act_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  group_by(area, home) %>% 
  summarize(n())

home_act_filtered <- home_act_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>%
  ungroup() %>% 
  arrange(area, home, home_act_day)
```

### 2020-01-23 home_char_act_cleaning.Rmd
data cleaning

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 15 Jan 2020"
+date: "Started 13 Jan 2020, Updated 23 Jan 2020"
 output: pdf_document
@@
   mutate(home_winter_id = as.factor(HomeWinterID),
-         residents_total = as.factor(Residents),
-         residents_under5 = as.factor(Under5),
-         residents_under18 = as.factor(Under18),
+         residents_total = as.numeric(Residents),
+         residents_under5 = as.numeric(Under5),
+         residents_under18 = as.numeric(Under18),
          residents_smoke = as.factor(Smoke),
@@
          home_type = as.factor(HomeType),
-         home_year_built = as.factor(YearHomeBuilt),
+         home_year_built = as.numeric(YearHomeBuilt),
          home_sqft = as.numeric(SqFootage),
-         home_floors = as.factor(NumFloors),
-         home_bedrooms = as.factor(NumBedrooms),
-         home_windows = as.factor(NumWindows),
-         home_pets = as.factor(NumPets),
-         home_furry = as.factor(NumFurryPets),
+         home_floors = as.numeric(NumFloors),
+         home_bedrooms = as.numeric(NumBedrooms),
+         home_windows = as.numeric(NumWindows),
+         home_pets = as.numeric(NumPets),
+         home_furry = as.numeric(NumFurryPets),
          primary_heat = as.factor(PrimaryHeat),
@@
    mutate(home_winter_id = as.factor(HomeWinterID),
-         residents_total = as.factor(Residents),
-         residents_under5 = as.factor(Under5),
-         residents_under18 = as.factor(Under18),
+         residents_total = as.numeric(Residents),
+         residents_under5 = as.numeric(Under5),
+         residents_under18 = as.numeric(Under18),
          residents_smoke = as.factor(Smoke),
@@
 home_char_clean <- kids_home_char %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  # cleaning based on checking odd values with paper logs
+  mutate(residents_under5 = if_else(residents_under5 == 10, 1, residents_under5)) %>% 
+  replace_with_na(replace = list(home_year_built = 0)) %>% 
+  replace_with_na(replace = list(home_floors = 0)) %>% 
+  replace_with_na(replace = list(home_windows = 0)) %>% 
+  replace_with_na(replace = list(residents_total = 0)) %>% 
+  replace_with_na(replace = list(residents_under5 = 0)) %>% 
+  replace_with_na(replace = list(residents_under18 = 0)) %>% 
+  replace_with_na(replace = list(home_sqft = 0)) %>% 
+  mutate(home_year_built = if_else(home_year_built == 15, 2002, home_year_built),
+         home_year_built = if_else(home_year_built == 874, 1980, home_year_built)) %>% 
+  replace_with_na(replace = list(home_year_built = 1400)) %>% 
+  mutate(home_sqft = if_else(home_sqft == 14 & home == "CH102", 360, home_sqft),
+         home_sqft = if_else(home_sqft == 14 & home == "CH107", 1152, home_sqft)) %>% 
+  replace_with_na(replace = list(home_sqft = 12)) %>% 
+  replace_with_na(replace = list(home_sqft = 14)) %>% 
+  replace_with_na(replace = list(home_sqft = 20)) %>% 
+  replace_with_na(replace = list(home_sqft = 24)) %>% 
+  replace_with_na(replace = list(home_sqft = 90)) %>% 
+  mutate(home_furry = if_else(home_furry == 4 & home == "WMA368", 1, home_furry),
+         home_furry = if_else(home_furry == 10 & home == "WMA345", 14, home_furry),
+         home_furry = if_else(home == "WMA309", 3, home_furry),
+         home_pets = if_else(home == "WMA309", 3, home_pets),
+         home_pets = if_else(home == "WMA437", 3, home_pets),
+         home_furry = if_else(home == "WMA437", 3, home_furry)) %>% 
+  replace_with_na(replace = list(home_furry = 14)) %>% 
+  ungroup() %>% 
+  arrange(area, home, winter_id)
 

### 2020-01-23 home_char_act_summary.Rmd
data cleaning

--- 2020-01-16 home_char_act_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 15 Jan 2020"
-output: pdf_document
+date: "Started 15 Jan 2020, Updated 23 Jan 2020"
+output:
+  pdf_document: default
+  word_document: default
 ---
@@
 home_char_summary <- home_char_clean %>% 
-  filter(winter_id == 1 | winter_id == 2) %>% 
-  filter(home_sqft != 0) %>% 
-  group_by(area, winter_id) %>% 
+  #filter(winter_id == 1 | winter_id == 2) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area) %>% 
   summarize("N" = n(),
@@
 home_plot <- data %>% 
-  filter(winter_id == 1) %>% 
-  filter(home_sqft != 0) %>%
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   select(area, variable) %>% 
@@
 home_plot <- data %>% 
-  filter(winter_id == 1) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   select(area, variable) %>% 

### 2020-01-23 indoor_temp_rh_summary.Rmd
data cleaning

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 16 Jan 2020"
+date: "Started 15 Jan 2020, Updated 23 Jan 2020"
 output: pdf_document
@@
 temp_rh_summary <- indoor_temp_rh_clean %>%
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area) %>% 
@@
 boxplots_temp_rh <- data %>% 
-  # change input to plot different temp/RH variable
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   select(area, variable) %>% 
@@
     theme_minimal() +
-    #ylim(0, 100) +
-    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
-    #labs(y = "Max indoor RH (%)") +
-    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
     theme(axis.text.x = element_text(size = 16, colour = "black"),

### 2020-01-23 moisture_summary.Rmd
data cleaning

--- 2020-01-14 moisture_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 14 Jan 2020"
+date: "Started 22 Nov 2019, Updated 23 Jan 2020"
 output: pdf_document
@@
 
-# Summary stats
+# Moisture content summary stats
+## Average of 3 measurements
 ```{r}
 summary_moisture <- moisture_clean %>%
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area) %>% 
-  summarize(mean(moisture_ave, na.rm = TRUE), sd(moisture_ave, na.rm = TRUE),
-            median(moisture_ave, na.rm = TRUE),
-            min(moisture_ave, na.rm = TRUE), max(moisture_ave, na.rm = TRUE)) %>% 
+  summarize(mean = mean(moisture_ave, na.rm = TRUE), 
+            SD = sd(moisture_ave, na.rm = TRUE),
+            median = median(moisture_ave, na.rm = TRUE),
+            min = min(moisture_ave, na.rm = TRUE), 
+            max = max(moisture_ave, na.rm = TRUE)) %>%
+  select(-area) %>% 
   t() 
-summary_moisture
+kable(summary_moisture, col.names = c("Alaska", "Navajo", "Montana"), 
+      digits = 1, align = "c")
+```
+
+# Was wood split?
+```{r}
+wood_split <- moisture_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(!is.na(moisture_split)) %>% 
+  group_by(area, moisture_split) %>% 
+  summarise(n = n())
+kable(wood_split, align = "c")
 ```
@@
 boxplots_moisture <- moisture_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   select(area, moisture_ave) %>% 

### 2020-01-23 pm_summary.Rmd
data cleaning

--- 2020-01-16 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 15 Jan 2020"
+date: "Started 22 Nov 2019, Updated 23 Jan 2020"
 output: pdf_document
@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(pm < 500) %>% 
   group_by(area) %>% 
@@
 boxplots_kids_pm <- pm_clean %>% 
-  #filter(pm < 500) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(pm < 500) %>% 
   ggplot(aes(area, pm)) +

### 2020-01-23 stove_grade_usage_cleaning.Rmd
data cleaning

@@
 stove_use_clean <- kids_stove_use %>% 
-  replace_with_na(replace = list(wood_cords = -999, wood_cords = 0)) %>% 
+  replace_with_na(replace = list(wood_cords = -999)) %>%
+  replace_with_na(replace = list(wood_cords = 0)) %>%
   mutate(wood_collect = factor(wood_collect,

### 2020-01-23 stove_grade_usage_summary.Rmd
data cleaning

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 15 Jan 2020"
+date: "Started 15 Jan 2020, Updated 23 Jan 2020"
 output: pdf_document
@@
   #filter(area == "WMT") %>% 
-  filter(winter_id == 1) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>%  
   select(-home_winter_id, -wood_cords, -wood_type_other,
@@
 stove_grade_summary <- stove_grades_clean %>% 
-  filter(winter_id == 1 | winter_id == 2) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   select(area, home, stove_grade) %>% 
-  filter(!is.na(stove_grade)) %>% 
+  #filter(!is.na(stove_grade)) %>% 
   group_by(area, stove_grade) %>% 
@@
 stove_use_summary <- stove_use_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area) %>% 
@@
 stove_use_plot <- data %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   select(area, variable) %>% 

### 2020-01-23 sums_data_cleaning.Rmd
data cleaning

@@
 sums_time_trends_plot <- sums_clean %>% 
-  filter(home == "WMA102") %>% 
+  filter(home == "CH405") %>% 
   ggplot() + 
@@
          x = "Datetime",
-         color = "Month of Year") +
+         color = "Logtag ID") +
     theme(axis.title.y = element_text(size = 12,
@@
 ```
+
+# Thresholds for heating events
+```{r}
+# Load SUMs data
+sums_clean <- read_rds("Output/sums_clean.rds")
+
+thresholds <- sums_clean %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(temp_sums_40 = if_else(temp_c_sums >= 40, 1, 0),
+         temp_sums_40_event = if_else(temp_sums_40 == 1 & lag(temp_sums_40) == 0, "Start", "Null"),
+         temp_sums_40_event = if_else(is.na(temp_sums_40_event), "Start", temp_sums_40_event),
+         temp_sums_40_event = if_else(temp_sums_40 == 1 & temp_sums_40_event == "Null",
+                                      "Event", temp_sums_40_event),
+         temp_sums_40_event = if_else(temp_sums_40_event == "Event" &
+                                      lead(temp_sums_40_event) == "Null" &
+                                      lead(temp_sums_40_event, 2) == "Null" &
+                                      lead(temp_sums_40_event, 3) == "Null", "End", temp_sums_40_event)) %>% 
+  group_by(area, home, winter_id, date_sums, temp_sums_40_event) %>% 
+  summarize(n = n()) %>% 
+  group_by(temp_sums_40_event) %>% 
+  summarize(mean(n))
+
+table(thresholds$temp_sums_40_event)
+```

### 2020-01-23 sums_summary.Rmd
data cleaning

@@
 author: "Ethan Walker"
-date: "Started 9 Jan 2020, Updated 15 Jan 2020"
+date: "Started 9 Jan 2020, Updated 23 Jan 2020"
 output: pdf_document
@@
   # group by: area, treatment, dtid, day_of_week, Sampling_day
-  group_by(area, winter_id) %>% 
+  #filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(temp_c_sums <= 0) %>% 
+  group_by(area, home, ibid) %>% 
   summarize("n" = n(),
@@
             "Max Temp" = max(temp_c_sums, na.rm = TRUE)) %>% 
-  select(-area, -winter_id) %>% 
+  select(-area) %>% 
   t()
 
-kable(sums_summary, col.names = c("Alaska (1)", "Alaska (2)",
-                                  "Navajo (1)", "Navajo (2)",
-                                  "Montana (1)", "Montana (2)"), 
-      digits = 1, align = "c")
+kable(sums_summary, col.names = c("Alaska", "Navajo", "Montana"), digits = 1, align = "c")
+
+sums_summary <- sums_clean %>% 
+  # group by: area, treatment, dtid, day_of_week, Sampling_day
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(temp_c_sums > 0) %>% 
+  group_by(area) %>% 
+  summarize("n" = n(),
+            "Mean Temp" = mean(temp_c_sums, na.rm = TRUE), 
+            "SD Temp" = sd(temp_c_sums, na.rm = TRUE),
+            "Min Temp" = min(temp_c_sums, na.rm = TRUE), 
+            "Median Temp" = median(temp_c_sums, na.rm = TRUE),
+            "Max Temp" = max(temp_c_sums, na.rm = TRUE)) %>% 
+  select(-area) %>% 
+  t()
+
+kable(sums_summary, col.names = c("Alaska", "Navajo", "Montana"), digits = 1, align = "c")
 ```
@@
 boxplots_kids_sums <- sums_clean %>%
-  filter(winter_id == 1) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   ggplot(aes(area, temp_c_sums)) +

### 2020-01-24 sums_data_cleaning.Rmd
new functions for SUMs data summary

@@
 
-thresholds <- sums_clean %>% 
+thresholds_test <- sums_clean %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
   group_by(area, home, winter_id) %>%
-  mutate(temp_sums_40 = if_else(temp_c_sums >= 40, 1, 0),
-         temp_sums_40_event = if_else(temp_sums_40 == 1 & lag(temp_sums_40) == 0, "Start", "Null"),
-         temp_sums_40_event = if_else(is.na(temp_sums_40_event), "Start", temp_sums_40_event),
-         temp_sums_40_event = if_else(temp_sums_40 == 1 & temp_sums_40_event == "Null",
-                                      "Event", temp_sums_40_event),
-         temp_sums_40_event = if_else(temp_sums_40_event == "Event" &
-                                      lead(temp_sums_40_event) == "Null" &
-                                      lead(temp_sums_40_event, 2) == "Null" &
-                                      lead(temp_sums_40_event, 3) == "Null", "End", temp_sums_40_event)) %>% 
-  group_by(area, home, winter_id, date_sums, temp_sums_40_event) %>% 
-  summarize(n = n()) %>% 
-  group_by(temp_sums_40_event) %>% 
-  summarize(mean(n))
-
-table(thresholds$temp_sums_40_event)
-```
+  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event = lead(event),
+         time_diff_min = (event - start),
+         time_diff_min = as.numeric(time_diff_min/60)) %>% 
+  filter(!is.na(time_diff_min)) %>% 
+  mutate(next_event = (lead(start)-event),
+         next_event = as.numeric(next_event)/3600,
+         event2 = if_else(next_event <= 1.5, lead(event), event),
+         time_diff_min2 = (event - start),
+         time_diff_min2 = as.numeric(time_diff_min2/60),
+         event3 = if_else(next_event <= 1.5 & lead(next_event) <= 1.5, lead(event, 2), event2),
+         time_diff_min3 = (event3 - start),
+         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
+  distinct(event3, .keep_all = TRUE) %>% 
+  mutate(time_diff_min4 = (event3 - start),
+         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
+  filter(!is.na(time_diff_min4)) %>% 
+  group_by(area, home, winter_id, date_sums) %>% 
+  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
+  group_by(area) %>% 
+  summarize("Mean events per day" = round(mean(n), digits = 2),
+            "Mean event length (min)" = round(mean(mean_time), digits = 0))
+thresholds_test
+
+# Function for calculating heating events using specified threshold temps
+# Also specify the "gap" time between events
+# Try another function for % day above threshold
+thresholds_function <- function(data, temp, gap) {
+
+thresholds <- data %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event = lead(event),
+         time_diff_min = (event - start),
+         time_diff_min = as.numeric(time_diff_min/60)) %>% 
+  filter(!is.na(time_diff_min)) %>% 
+  mutate(next_event = (lead(start)-event),
+         next_event = as.numeric(next_event)/3600,
+         event2 = if_else(next_event <= gap, lead(event), event),
+         time_diff_min2 = (event - start),
+         time_diff_min2 = as.numeric(time_diff_min2/60),
+         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
+         time_diff_min3 = (event3 - start),
+         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
+  distinct(event3, .keep_all = TRUE) %>% 
+  mutate(time_diff_min4 = (event3 - start),
+         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
+  filter(!is.na(time_diff_min4)) %>% 
+  group_by(area, home, winter_id, date_sums) %>% 
+  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
+  group_by(area) %>% 
+  summarize("Mean events per day" = round(mean(n), digits = 2),
+            "Mean event length (min)" = round(mean(mean_time), digits = 0))
+thresholds
+
+}
+
+thresholds_function(sums_clean, 40, 2)
+
+# Function to look at percentage of time spent above a threshold
+thresholds_percent <- function(data, temp) {
+
+thresholds <- data %>% 
+  group_by(area, home, winter_id, date_sums) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
+  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
+  group_by(area) %>% 
+  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
+thresholds
+
+}
+
+thresholds_percent(sums_clean, 40)
+```

### 2020-01-24 sums_summary.Rmd
new functions for SUMs data summary

@@
 author: "Ethan Walker"
-date: "Started 9 Jan 2020, Updated 23 Jan 2020"
+date: "Started 9 Jan 2020, Updated 24 Jan 2020"
 output: pdf_document
@@
 ```
+
+# Function for calculating heating events using specified threshold temps
+## Also specify the "gap" time between events
+```{r}
+thresholds_function <- function(data, temp, gap) {
+
+thresholds <- data %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event = lead(event),
+         time_diff_min = (event - start),
+         time_diff_min = as.numeric(time_diff_min/60)) %>% 
+  filter(!is.na(time_diff_min)) %>% 
+  mutate(next_event = (lead(start)-event),
+         next_event = as.numeric(next_event)/3600,
+         event2 = if_else(next_event <= gap, lead(event), event),
+         time_diff_min2 = (event - start),
+         time_diff_min2 = as.numeric(time_diff_min2/60),
+         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
+         time_diff_min3 = (event3 - start),
+         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
+  distinct(event3, .keep_all = TRUE) %>% 
+  mutate(time_diff_min4 = (event3 - start),
+         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
+  filter(!is.na(time_diff_min4)) %>% 
+  group_by(area, home, winter_id, date_sums) %>% 
+  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
+  group_by(area) %>% 
+  summarize("Mean events per day" = round(mean(n), digits = 2),
+            "Mean event length (min)" = round(mean(mean_time), digits = 0))
+thresholds
+
+}
+
+thresholds_function(sums_clean, 40, 2)
+```
+
+# Function to look at percentage of time spent above a threshold
+```{r}
+thresholds_percent <- function(data, temp) {
+
+thresholds <- data %>% 
+  group_by(area, home, winter_id, date_sums) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
+  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
+  group_by(area) %>% 
+  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
+thresholds
+
+}
+
+thresholds_percent(sums_clean, 40)
+```

### 2020-01-27 demographics_summary.Rmd
Summary tables for abstacts/papers

---
title: "KidsAIR: household demographics summary"
author: "Ethan Walker"
date: "Started 27 Jan 2020, Updated 27 Jan 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")
```

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

demographics_filtered <- demographics_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Summarize demographics data
```{r}
# Summarize numeric vars
kids_dems <- demographics_filtered %>%  
  mutate(total_residents = as.numeric(total_residents),
         residents_under_five = as.numeric(residents_under_five)) %>% 
  distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(total_residents, na.rm = TRUE), 
            "SD" = sd(total_residents, na.rm = TRUE),
            "min" = min(total_residents, na.rm = TRUE), 
            "max" = max(total_residents, na.rm = TRUE))
kable(kids_dems, digits = 1, align = "c")

# Summarize factor vars
alaska_dems <- demographics_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "AK") %>% 
  group_by(age_child_nov) %>% 
  #summarize(n = n(), percent = (n()/10)*100)
  summarize(n = n(), percent = (n()/18)*100)
alaska_dems

navajo_dems <- demographics_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "NN") %>% 
  group_by(age_child_nov) %>% 
  #summarize(n = n(), percent = (n()/23)*100)
  summarize(n = n(), percent = (n()/32)*100)
navajo_dems

montana_dems <- demographics_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>% 
  filter(area == "WMT") %>% 
  group_by(age_child_nov) %>% 
  #summarize(n = n(), percent = (n()/60)*100)
  summarize(n = n(), percent = (n()/87)*100)
montana_dems

## Select specific var to assess individually/group by area
table(demographics_filtered$education, demographics_filtered$area, useNA = "ifany")
```

### 2020-01-27 home_char_act_summary.Rmd
Summary tables for abstacts/papers

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 23 Jan 2020"
+date: "Started 15 Jan 2020, Updated 27 Jan 2020"
 output:
@@
 home_char_clean <- read_rds("Output/home_char_clean.rds") 
-
 home_act_clean <- read_rds("Output/home_act_clean.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
 ```
+
+#### Abstract/manuscript tables ####
+
+# Filter for homes that have PM data
+```{r}
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+home_char_filtered <- home_char_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+home_act_filtered <- home_act_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+# Summarize home data
+```{r}
+# Summarize numeric vars
+kids_char <- home_char_filtered %>%  
+  distinct(home, .keep_all = TRUE) %>%  
+  group_by(area) %>% 
+  summarize("Mean" = mean(home_furry, na.rm = TRUE), 
+            "SD" = sd(home_furry, na.rm = TRUE),
+            "min" = min(home_furry, na.rm = TRUE), 
+            "max" = max(home_furry, na.rm = TRUE))
+kable(kids_char, digits = 1, align = "c")
+
+# Summarize factor vars
+alaska_char <- home_act_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>% 
+  filter(area == "AK") %>% 
+  group_by(home_act_sweep) %>% 
+  #summarize(n = n(), percent = (n()/10)*100)
+  summarize(n = n(), percent = (n()/60)*100)
+alaska_char
+
+navajo_char <- home_act_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>% 
+  filter(area == "NN") %>% 
+  group_by(home_act_sweep) %>% 
+  #summarize(n = n(), percent = (n()/23)*100)
+  summarize(n = n(), percent = (n()/138)*100)
+navajo_char
+
+montana_char <- home_act_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>% 
+  filter(area == "WMT") %>% 
+  group_by(home_act_sweep) %>% 
+  #summarize(n = n(), percent = (n()/60)*100)
+  summarize(n = n(), percent = (n()/360)*100)
+montana_char
+
+kids_char <- rbind(alaska_char, navajo_char, montana_char)
+kids_char
+```

### 2020-01-27 home_child_data_cleaning.Rmd
Summary tables for abstacts/papers

@@
   mutate(intervention_datetime = ymd_hms(intervention_datetime)) %>% 
-  select(-trash)
+  select(-trash) %>% 
+  mutate(home_winter_id = as.numeric(home_winter_id),
+         home_winter_id = if_else(home == "CH218" & winter_id == 1, 77, home_winter_id),
+         home_winter_id = if_else(home == "CH218" & winter_id == 2, 111, home_winter_id),
+         home_winter_id = as.factor(home_winter_id))
 

### 2020-01-27 indoor_temp_rh_summary.Rmd
Summary tables for abstacts/papers

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 23 Jan 2020"
+date: "Started 15 Jan 2020, Updated 27 Jan 2020"
 output: pdf_document
@@
 indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
 ```
+
+#### Abstract/manuscript tables ####
+
+# Filter for homes that have PM data
+```{r}
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+temp_rh_filtered <- indoor_temp_rh_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+# Summarize home data
+```{r}
+# Summarize numeric vars
+kids_temp_rh <- temp_rh_filtered %>%  
+  distinct(home, .keep_all = TRUE) %>%  
+  group_by(area) %>% 
+  summarize("Mean" = mean(rh_indoor_min, na.rm = TRUE), 
+            "SD" = sd(rh_indoor_min, na.rm = TRUE),
+            "min" = min(rh_indoor_min, na.rm = TRUE), 
+            "med" = median(rh_indoor_min, na.rm = TRUE), 
+            "max" = max(rh_indoor_min, na.rm = TRUE))
+kable(kids_temp_rh, digits = 1, align = "c")
+
+```

### 2020-01-27 moisture_summary.Rmd
Summary tables for abstacts/papers

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 23 Jan 2020"
+date: "Started 22 Nov 2019, Updated 27 Jan 2020"
 output: pdf_document
@@
 moisture_clean <- read_rds("Output/moisture_clean.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
 ```
+
+#### Abstract/manuscript tables ####
+
+# Filter for homes that have PM data
+```{r}
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+moisture_filtered <- moisture_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+# Summarize home data
+```{r}
+# Summarize numeric vars
+kids_moisture <- moisture_filtered %>%  
+  group_by(area) %>% 
+  summarize("Mean" = mean(moisture_ave, na.rm = TRUE), 
+            "SD" = sd(moisture_ave, na.rm = TRUE),
+            "min" = min(moisture_ave, na.rm = TRUE), 
+            "med" = median(moisture_ave, na.rm = TRUE), 
+            "max" = max(moisture_ave, na.rm = TRUE))
+kable(kids_moisture, digits = 1, align = "c")
+
+table(moisture_filtered$area, moisture_filtered$moisture_split)
+```

### 2020-01-27 pm_summary.Rmd
Summary tables for abstacts/papers

@@
             "Min PM" = min(pm, na.rm = TRUE), 
+            "25%" = quantile(pm, 0.25),
             "Median PM" = median(pm, na.rm = TRUE),
+            "75%" = quantile(pm, 0.75),
             "Max PM" = max(pm, na.rm = TRUE),

### 2020-01-27 stove_grade_usage_summary.Rmd
Summary tables for abstacts/papers

@@
 stove_use_clean <- read_rds("Output/stove_use_clean.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
 ```
+
+#### Abstract/manuscript tables ####
+
+# Filter for homes that have PM data
+```{r}
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+stove_use_filtered <- stove_use_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+stove_grades_filtered <- stove_grades_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+# Summarize home data
+```{r}
+# Summarize numeric vars
+kids_use <- stove_use_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>%  
+  group_by(area) %>% 
+  summarize("Mean" = mean(wood_cords, na.rm = TRUE), 
+            "SD" = sd(wood_cords, na.rm = TRUE),
+            "min" = min(wood_cords, na.rm = TRUE), 
+            "med" = median(wood_cords, na.rm = TRUE), 
+            "max" = max(wood_cords, na.rm = TRUE))
+kable(kids_use, digits = 1, align = "c")
+
+# Summarize factor vars
+alaska_use <- stove_use_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>% 
+  filter(area == "AK") %>% 
+  group_by(wood_collect_method) %>% 
+  summarize(n = n(), percent = (n()/10)*100)
+  #summarize(n = n(), percent = (n()/60)*100)
+alaska_use
+
+navajo_use <- stove_use_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>% 
+  filter(area == "NN") %>% 
+  group_by(wood_collect_method) %>% 
+  summarize(n = n(), percent = (n()/23)*100)
+  #summarize(n = n(), percent = (n()/138)*100)
+navajo_use
+
+montana_use <- stove_use_filtered %>%  
+  #distinct(home, .keep_all = TRUE) %>% 
+  filter(area == "WMT") %>% 
+  group_by(wood_collect_method) %>% 
+  summarize(n = n(), percent = (n()/60)*100)
+  #summarize(n = n(), percent = (n()/360)*100)
+montana_use
+
+kids_use <- rbind(alaska_use, navajo_use, montana_use)
+kable(kids_use)
+
+# stove grade table
+stove_grades_filtered <- stove_grades_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(home) %>% 
+  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>%
+  filter(!is.na(stove_grade)) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+table(stove_grades_filtered$stove_grade, stove_grades_filtered$area)
+```

### 2020-01-28 moisture_summary.Rmd
summary stat updates

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 27 Jan 2020"
+date: "Started 22 Nov 2019, Updated 28 Jan 2020"
 output: pdf_document
@@
 library(tidyverse)
+library(lubridate)
 library(readr)
@@
   mutate(pm_home = "true") %>% 
-  select(area, home, home_winter_id, pm_home) %>% 
+  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
+  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
   ungroup() %>% 
@@
   arrange(area, home)
+
+moisture_filtered_time <- moisture_filtered %>% 
+  # fix wrong date
+  mutate(moisture_date = as.character(moisture_date),
+         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
+                                 "2017-04-10", moisture_date),
+         moisture_date = ymd(moisture_date)) %>% 
+  mutate(moisture_date = ymd(moisture_date),
+         first_date = ymd(first_date),
+         time_diff = moisture_date - first_date,
+         time_diff = as.numeric(abs(time_diff))) %>% 
+  group_by(home) %>% 
+  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
+  ungroup() %>% 
+  filter(time_diff_rank == 1)
 ```
@@
 
+# Summarize numeric vars
+kids_moisture <- moisture_filtered_time %>%  
+  group_by(area) %>% 
+  summarize("Mean" = mean(time_diff, na.rm = TRUE), 
+            "SD" = sd(time_diff, na.rm = TRUE),
+            "min" = min(time_diff, na.rm = TRUE), 
+            "med" = median(time_diff, na.rm = TRUE), 
+            "max" = max(time_diff, na.rm = TRUE))
+kable(kids_moisture, digits = 1, align = "c")
+
 table(moisture_filtered$area, moisture_filtered$moisture_split)

### 2020-01-28 pm_summary.Rmd
summary stat updates

--- 2020-01-23 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 23 Jan 2020"
+date: "Started 22 Nov 2019, Updated 28 Jan 2020"
 output: pdf_document
@@
 pm_summary <- pm_clean %>% 
-  # group by: area, treatment, dtid, day_of_week, Sampling_day
   filter(winter_id == 1 & treatment == "Placebo") %>% 
@@
             "Min PM" = min(pm, na.rm = TRUE), 
+            "25%" = quantile(pm, 0.25),
             "Median PM" = median(pm, na.rm = TRUE),
+            "75%" = quantile(pm, 0.75),
             "Max PM" = max(pm, na.rm = TRUE),
             "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
+            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
             "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
@@
       digits = 1, align = "c")
+
+pm_by_day <- pm_clean %>% 
+  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
+  mutate(day_of_week = factor(day_of_week,
+                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
+                                            "Friday", "Saturday", "Sunday"),
+                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(pm < 500) %>% 
+  group_by(area, sampling_month) %>% 
+  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "25%" = quantile(pm, 0.25),
+            "Median PM" = median(pm, na.rm = TRUE),
+            "75%" = quantile(pm, 0.75),
+            "Max PM" = max(pm, na.rm = TRUE))
+pm_by_day
 ```

### 2020-01-28 sums_summary.Rmd
summary stat updates

--- 2020-01-23 sums_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 9 Jan 2020, Updated 23 Jan 2020"
+date: "Started 9 Jan 2020, Updated 28 Jan 2020"
 output: pdf_document
@@
 ```
+
+# Filter for homes that have PM data
+```{r}
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home, pm_datetime_new) %>% 
+  separate(pm_datetime_new, c("sample_date", "sample_time"), sep = " ", remove = TRUE) %>% 
+  mutate(sample_date = ymd(sample_date)) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+sums_filtered <- sums_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+sums_filtered_date <- sums_clean %>% 
+  mutate(sample_date = date_sums,
+         sample_date = ymd(sample_date)) %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id", "sample_date")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+# Function for calculating heating events using specified threshold temps
+## Also specify the "gap" time between events
+```{r}
+thresholds_function <- function(data, temp, gap) {
+
+thresholds <- data %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event = lead(event),
+         time_diff_min = (event - start),
+         time_diff_min = as.numeric(time_diff_min/60)) %>% 
+  filter(!is.na(time_diff_min)) %>% 
+  mutate(next_event = (lead(start)-event),
+         next_event = as.numeric(next_event)/3600,
+         event2 = if_else(next_event <= gap, lead(event), event),
+         time_diff_min2 = (event - start),
+         time_diff_min2 = as.numeric(time_diff_min2/60),
+         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
+         time_diff_min3 = (event3 - start),
+         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
+  distinct(event3, .keep_all = TRUE) %>% 
+  mutate(time_diff_min4 = (event3 - start),
+         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
+  filter(!is.na(time_diff_min4)) %>% 
+  group_by(area, home, winter_id, date_sums) %>% 
+  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
+  group_by(area) %>% 
+  summarize("Mean events per day" = round(mean(n), digits = 2),
+            "SD events per day" = round(sd(n), digits = 2),
+            "Mean event length (min)" = round(mean(mean_time), digits = 0))
+thresholds
+
+}
+
+thresholds_function(sums_filtered_date, 25, 1.5)
+thresholds_function(sums_filtered_date, 30, 1.5)
+thresholds_function(sums_filtered_date, 35, 1.5)
+thresholds_function(sums_filtered_date, 40, 1.5)
+thresholds_function(sums_filtered_date, 45, 1.5)
+thresholds_function(sums_filtered_date, 50, 1.5)
+thresholds_function(sums_filtered_date, 55, 1.5)
+```
+
+# Function to look at percentage of time spent above a threshold
+```{r}
+thresholds_percent <- function(data, temp) {
+
+thresholds <- data %>% 
+  group_by(area, home, winter_id, date_sums) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
+  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
+  group_by(area) %>% 
+  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
+thresholds
+
+}
+
+thresholds_percent(sums_filtered_date, 25)
+thresholds_percent(sums_filtered_date, 30)
+thresholds_percent(sums_filtered_date, 35)
+thresholds_percent(sums_filtered_date, 40)
+thresholds_percent(sums_filtered_date, 45)
+thresholds_percent(sums_filtered_date, 50)
+thresholds_percent(sums_filtered_date, 55)
+```

### 2020-01-30 filter_treatment_obs_check.Rmd
SUMs cleaning and analysis

@@
 author: "Ethan Walker"
-date: "Started 21 Jan 2020, Updated 23 Jan 2020"
+date: "Started 21 Jan 2020, Updated 30 Jan 2020"
 output: pdf_document
@@
 # SUMs data
-## 80/93 placebo homes from Winter 1 have SUMs data
-## 77/88 education homes from Winter 1 have SUMs data
+## 77/93 placebo homes from Winter 1 have SUMs data
+## 73/88 education homes from Winter 1 have SUMs data
 ## 70/86 filter homes from Winter 1 have SUMs data

### 2020-01-30 moisture_data_cleaning.Rmd
data cleaning
SUMs cleaning and analysis

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 14 Jan 2020"
+date: "Started 19 Nov 2019, Updated 30 Jan 2020"
 output: pdf_document
@@
 moisture_clean <- kids_moisture %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  replace_with_na(replace = list(moisture_ave = "NaN")) %>% 
+  # lower range on instrument = 5%
+  # therefore, changing values < 5 to 5
+  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
+         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
+         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
+         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))
 

### 2020-01-30 moisture_summary.Rmd
SUMs cleaning and analysis

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 28 Jan 2020"
+date: "Started 22 Nov 2019, Updated 30 Jan 2020"
 output: pdf_document
@@
 # Summarize numeric vars
-kids_moisture <- moisture_filtered %>%  
+kids_moisture <- moisture_filtered_time %>%  
   group_by(area) %>% 

### 2020-01-30 sums_data_cleaning.Rmd
SUMs cleaning and analysis

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 14 Jan 2020"
+date: "Started 23 Dec 2019, Updated 30 Jan 2020"
 output: pdf_document
@@
   separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
-  mutate(datetime_sums = datetime,
-         date_sums = date,
-         time_sums = time,
+  mutate(datetime_sums = ymd_hms(datetime),
+         date_sums = ymd(date),
+         time_sums = hms(time),
          temp_c = if_else(is.na(temp_c), trash2, temp_c),
@@
   arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  filter(!is.na(temp_c_sums))
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  arrange(home, datetime_sums)
 
@@
   separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
-  mutate(date_sums = ymd(date_sums),
+  mutate(datetime_sums = ymd_hms(datetime_sums),
+         date_sums = ymd(date_sums),
+         time_sums = hms(time_sums),
          temp_c = if_else(is.na(temp_c), trash2, temp_c),
@@
          date_sums = mdy(date),
-         time_sums = time,
+         time_sums = hms(time),
          temp_f = if_else(is.na(temp_f), trash2, temp_f),
@@
   select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  filter(!is.na(temp_c_sums))
+  mutate(datetime_sums = ymd_hms(datetime_sums),
+         date_sums = ymd(date_sums)) %>% 
+  arrange(home, datetime_sums) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  arrange(home, datetime_sums)
 
@@
   separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>%
-  mutate(datetime_sums = datetime,
-         date_sums = date,
-         time_sums = time,
+  mutate(datetime_sums = ymd_hms(datetime),
+         date_sums = ymd(date),
+         time_sums = hms(time),
          temp_c = if_else(is.na(temp_c), trash2, temp_c),
@@
          date_sums = ymd(date),
-         time_sums = time,
+         time_sums = hms(time),
          temp_c = if_else(is.na(temp_c), trash2, temp_c),
@@
          file_type = "xlsx_c") %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  filter(!is.na(temp_c_sums))
+  select(home, area, datetime_sums, temp_c_sums, file_type) %>% 
+  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums),
+         date_sums = ymd(date_sums),
+         time_sums = hms(time_sums)) %>% 
+  arrange(home, datetime_sums) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  arrange(home, datetime_sums)
 
@@
          date_sums = ymd(date),
-         time_sums = time,
+         time_sums = hms(time),
          temp_c_sums = (temp_f-32)/1.8,
@@
   select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  filter(!is.na(temp_c_sums))
+  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums),
+         date_sums = ymd(date_sums),
+         time_sums = hms(time_sums)) %>% 
+  arrange(home, datetime_sums) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  arrange(home, datetime_sums)
 
@@
                       wmt_sums_xlsx4, wmt_sums_xlsx5) %>% 
-  select(-datetime_sums) %>% 
-  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
-  filter(!is.na(date_sums)) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% # remove duplicate lines that have same info from different files
+  filter(!is.na(datetime_sums)) %>% 
   mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
                                                   "temp_f", "xlsx_c",
-                                                  "xlsx_f")))
-
-# This filtering will keep only the first instance when there are multiple
-# observations for a specific datetime_sums
-wmt_sums_filter <- wmt_sums_raw %>% 
-  group_by(home) %>% 
-  distinct(datetime_sums, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(home, datetime_sums)
-
-# write_rds(wmt_sums_filter, "Output/wmt_sums_raw.rds")
+                                                  "xlsx_f"))) %>% 
+  arrange(home, datetime_sums, file_type) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  arrange(home, datetime_sums)
+
+# write_rds(wmt_sums_raw, "Output/wmt_sums_raw.rds")
 ```
@@
 ## No header, 3 columns: date, time, temp (F)
+##### NO LONGER USING THESE FILES!
+##### COMPARED TO CORRESPONDING LOGTAG FILES, THE TIMESTAMPS ARE OFF
 list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")
@@
 # Bind all WMT SUMs data together
-nn_sums_raw <- rbind(nn_sums_csv, nn_sums_csv_2, nn_sums_csv_3) %>% 
-  select(-datetime_sums) %>% 
-  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
-  filter(!is.na(date_sums)) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
-                                                  "temp_f")))
-
-# This filtering will keep only the first instance when there are multiple
-# observations for a specific datetime_sums
-nn_sums_filter <- nn_sums_raw %>% 
-  group_by(home) %>% 
-  distinct(datetime_sums, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(home, datetime_sums)
-
-# write_rds(nn_sums_filter, "Output/nn_sums_raw.rds")
+nn_sums_raw <- rbind(nn_sums_csv, nn_sums_csv_2) %>% 
+  filter(!is.na(datetime_sums)) %>% 
+  mutate(file_type = factor(file_type, levels = c("logtag", "temp_c"))) %>% 
+  arrange(home, datetime_sums, file_type) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  arrange(home, datetime_sums)
+
+# write_rds(nn_sums_raw, "Output/nn_sums_raw.rds")
 ```
@@
   mutate(datetime_sums = ymd_hms(datetime),
-         date_sums = date,
-         time_sums = time,
+         date_sums = ymd(date),
+         time_sums = hms(time),
          temp_c = if_else(is.na(temp_c), trash2, temp_c),
@@
 ## 5 columns: line, date, time, temp (F), notes
+##### NO LONGER USING THESE FILES!
+##### COMPARED TO CORRESPONDING LOGTAG FILES, THE TIMESTAMPS ARE OFF
 list_csv_files_2 <- list.files("Input/AK/sums_csv_old")
@@
   mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  filter(!is.na(temp_c_sums))
+  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(date_sums = ymd(date_sums),
+         time_sums = hms(time_sums)) %>% 
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
+  arrange(home, datetime_sums) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  arrange(home, datetime_sums)
 
@@
   mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
-  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  filter(!is.na(temp_c_sums))
+  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
+  mutate(date_sums = ymd(date_sums),
+         time_sums = hms(time_sums)) %>%
+  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
+  arrange(home, datetime_sums) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  arrange(home, datetime_sums)
 
 # Bind all AK SUMs data together
-ak_sums_raw <- rbind(ak_sums_csv, ak_sums_csv_2, 
+ak_sums_raw <- rbind(ak_sums_csv,  
                      ak_sums_xlsx, ak_sums_xlsx2) %>% 
-  select(-datetime_sums) %>% 
-  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
-  filter(!is.na(date_sums)) %>% 
-  arrange(home, datetime_sums) %>% 
-  distinct() %>% 
-  mutate(file_type = factor(file_type, levels = c("logtag", "temp_f",
-                                                  "xlsx_c", "xlsx_f")))
-
-# This filtering will keep only the first instance when there are multiple
-# observations for a specific datetime_sums
-ak_sums_filter <- ak_sums_raw %>% 
-  group_by(home) %>% 
-  distinct(datetime_sums, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(home, datetime_sums)
-
-# write_rds(ak_sums_filter, "Output/ak_sums_raw.rds")
+  filter(!is.na(datetime_sums)) %>% 
+  mutate(file_type = factor(file_type, levels = c("logtag",
+                                                  "xlsx_c", "xlsx_f"))) %>% 
+  arrange(home, datetime_sums, file_type) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>%  # remove duplicate lines that have same info from different files
+  ungroup() %>% 
+  arrange(home, datetime_sums)
+
+# write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
 ```

### 2020-01-30 sums_summary.Rmd
SUMs cleaning and analysis

--- 2020-01-23 sums_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 9 Jan 2020, Updated 23 Jan 2020"
+date: "Started 9 Jan 2020, Updated 30 Jan 2020"
 output: pdf_document
@@
 sums_clean <- read_rds("Output/sums_clean.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
+
 ```
@@
 ```
+
+# Filter for homes that have PM data
+```{r}
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home, pm_datetime_new, pm) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+pm_homes_placebo_date <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home, pm_datetime_new, pm) %>% 
+  separate(pm_datetime_new, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  mutate(sample_date = ymd(sample_date)) %>% 
+  group_by(area, home) %>% 
+  distinct(sample_date, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+sums_filtered <- sums_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+sums_filtered_date <- sums_clean %>% 
+  mutate(sample_date = date_sums,
+         sample_date = ymd(sample_date)) %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo_date, by = c("area", "home", "home_winter_id", "sample_date")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+# Function for calculating heating events using specified threshold temps
+## Also specify the "gap" time between events
+```{r}
+thresholds_function <- function(data, temp, gap) {
+
+thresholds <- data %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+  mutate(time_sums = as.character(time_sums),
+         date_sums = as.character(date_sums)) %>%  
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event = lead(event),
+         time_diff_min = (event - start),
+         time_diff_min = as.numeric(time_diff_min/60)) %>% 
+  filter(!is.na(time_diff_min)) %>% 
+  mutate(next_event = (lead(start)-event),
+         next_event = as.numeric(next_event)/3600,
+         event2 = if_else(next_event <= gap, lead(event), event),
+         time_diff_min2 = (event - start),
+         time_diff_min2 = as.numeric(time_diff_min2/60),
+         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
+         time_diff_min3 = (event3 - start),
+         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
+  distinct(event3, .keep_all = TRUE) %>% 
+  mutate(time_diff_min4 = (event3 - start),
+         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
+  filter(!is.na(time_diff_min4)) %>% 
+  group_by(area, home, winter_id, date_sums) %>% 
+  summarize(n = n(), mean_time = mean(time_diff_min4))
+
+thresholds_by_home <<- thresholds %>% 
+  group_by(home) %>% 
+  summarize("Mean events per day" = round(mean(n), digits = 2),
+            "SD events per day" = round(sd(n), digits = 2),
+            "Mean event length (min)" = round(mean(mean_time), digits = 0),
+            "SD event length (min)" = round(sd(mean_time), digits = 2))
+thresholds_by_home
+
+thresholds_by_area <<- thresholds %>% 
+  group_by(area) %>% 
+  summarize("Mean events per day" = round(mean(n), digits = 2),
+            "SD events per day" = round(sd(n), digits = 2),
+            "Mean event length (min)" = round(mean(mean_time), digits = 0),
+            "SD event length (min)" = round(sd(mean_time), digits = 2))
+thresholds_by_area
+
+}
+
+thresholds_function(sums_filtered_date, 30, 1.5)
+thresholds_function(sums_filtered_date, 40, 1.5)
+thresholds_function(sums_filtered_date, 50, 1.5)
+```
+
+# Function to look at percentage of time spent above a threshold
+```{r}
+thresholds_percent <- function(data, temp) {
+
+thresholds <- sums_filtered_date %>% 
+  group_by(area, home, winter_id, date_sums) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0)) %>% 
+  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
+  group_by(area) %>% 
+  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
+thresholds
+
+}
+
+thresholds_percent(sums_filtered_date, 25)
+thresholds_percent(sums_filtered_date, 30)
+thresholds_percent(sums_filtered_date, 35)
+thresholds_percent(sums_filtered_date, 40)
+thresholds_percent(sums_filtered_date, 45)
+thresholds_percent(sums_filtered_date, 50)
+thresholds_percent(sums_filtered_date, 55)
+```

### 2020-01-31 pm_summary.Rmd
more SUMs cleaning

@@
 pm_summary <- pm_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(winter_id == 1 & treatment == "Placebo") %>% 
   filter(pm < 500) %>% 
-  group_by(area) %>% 
+  group_by(treatment) %>% 
   summarize("n" = n(),
@@
             "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
-  select(-area) %>% 
+  #select(-treatment) %>% 
   t()
 
-kable(pm_summary, col.names = c("Alaska", "Navajo", "Montana"), 
+kable(pm_summary, 
       digits = 1, align = "c")

### 2020-01-31 sums_data_cleaning.Rmd
more SUMs cleaning

@@
 # write_rds(sums_clean, "Output/sums_clean.rds")
+
+sums_clean <- read_rds("Output/sums_clean.rds")
+
+time_diff_check <- sums_clean %>% 
+  group_by(area, home) %>% 
+  mutate(time_diff = lead(datetime_sums) - datetime_sums,
+         time_diff_indicator = if_else(time_diff < 20, 1, 0)) %>% 
+  filter(time_diff_indicator == 1) %>% 
+  summarise(n())
+
+time_diff_check_home <- sums_clean %>% 
+  group_by(area, home) %>% 
+  mutate(time_diff = lead(datetime_sums) - datetime_sums,
+         time_diff_indicator = if_else(time_diff < 20, 1, 0)) %>% 
+  filter(home == "WMA238")
 ```

### 2020-02-06 exposure_initial_analysis.Rmd
SUMs cleaning updates

---
title: "KidsAIR: initial exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 2 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Combine datasets
```{r}
pm_home_mean <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, home) %>% 
  summarize(mean_pm = mean(pm, na.rm = TRUE))

pm_demographics <- pm_home_mean %>% 
  left_join(demographics_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(income = factor(income,
                         levels = c("Less than $20,000", "$20,000 to $29,999",
                                    "$30,000 to $39,999", "$40,000 to $49,999",
                                    "$50,000 to $74,999", "$75,000 to $99,999",
                                    "$100,000 or more"))) %>% 
  mutate(education = factor(education,
                            levels = c("Less than high school", 
                                       "High school diploma or GED",
                                       "Some college",
                                       "College degree")))

pm_home_char <- pm_home_mean %>% 
  left_join(home_char_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")))

pm_home_char <- pm_home_mean %>% 
  left_join(home_char_clean, by = c("area", "home")) %>% 
  filter(winter_id == 1) %>% 
  mutate(stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")))
```

# Simple associations between PM and covariates
```{r}
pm_demographics_results <- lm(mean_pm ~ education, data = pm_demographics)
summary(pm_demographics_results)

pm_home_char_results <- lm(mean_pm ~ stove_age, data = pm_home_char)
summary(pm_home_char_results)
```

### 2020-02-06 pm_summary.Rmd
SUMs cleaning updates

--- 2020-01-27 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 27 Jan 2020"
+date: "Started 22 Nov 2019, Updated 2 Feb 2020"
 output: pdf_document
@@
 pm_summary <- pm_clean %>% 
-  # group by: area, treatment, dtid, day_of_week, Sampling_day
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(pm < 500) %>% 
+  #filter(pm < 500) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
   group_by(area) %>% 
@@
             "Min PM" = min(pm, na.rm = TRUE), 
-            "25%" = quantile(pm, 0.25),
+            "25%" = quantile(pm, 0.25, na.rm = TRUE),
             "Median PM" = median(pm, na.rm = TRUE),
-            "75%" = quantile(pm, 0.75),
+            "75%" = quantile(pm, 0.75, na.rm = TRUE),
             "Max PM" = max(pm, na.rm = TRUE),
             "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
+            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
             "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
             "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
-  select(-area) %>% 
+  #select(-treatment) %>% 
   t()
 
-kable(pm_summary, col.names = c("Alaska", "Navajo", "Montana"), 
+kable(pm_summary, 
       digits = 1, align = "c")
+
+pm_by_day <- pm_clean %>% 
+  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
+  mutate(day_of_week = factor(day_of_week,
+                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
+                                            "Friday", "Saturday", "Sunday"),
+                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(pm < 500) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  group_by(area, day_of_week) %>% 
+  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
+            "SD PM" = sd(pm, na.rm = TRUE),
+            "Min PM" = min(pm, na.rm = TRUE), 
+            "25%" = quantile(pm, 0.25, na.rm = TRUE),
+            "Median PM" = median(pm, na.rm = TRUE),
+            "75%" = quantile(pm, 0.75, na.rm = TRUE),
+            "Max PM" = max(pm, na.rm = TRUE))
+pm_by_day
 ```

### 2020-02-06 sums_data_cleaning.Rmd
SUMs cleaning updates

--- 2020-01-30 sums_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 30 Jan 2020"
+date: "Started 23 Dec 2019, Updated 6 Feb 2020"
 output: pdf_document
@@
 # write_rds(sums_clean, "Output/sums_clean.rds")
+
+########## New cleaning on 6 Feb 2020 ##########
+
+sums_clean <- read_rds("Output/sums_clean.rds")
+
+# This cleaning is a bit of a desperation attempt to try and salvage most of the
+# SUMs data. Because of the way data was collected and files were saved, it
+# became very difficult ot piece files together for each home. For some of the homes,
+# data is fine - particularly if they used LogTags for the entire Winter and there
+# is one continuous file from one instrument for the whole Winter. However, if an
+# IButton was used (which held only around 1 month of data), or if a LogTag was
+# switched out during sampling, piecing the various files together becomes a nightmare.
+# All of the files had heads and tails of pre- or post-sampling data that needed
+# to be cut off, but there wasn't a good indication of when the sampling exactly
+# stopped and started when an instrument was switched out in the field. There was
+# typically overlap in the files when an instrument was switched out: a tail from
+# the old instrument before it was turned off, and a head from the new instrument
+# from the time it was turned on, but prior to sampling. During these overlap
+# periods in the data, there was no good indication of what instrument was actually
+# sampling within the home. I tried to write some rules/algorithms to pull out what
+# I thought was the correct data, but there inconsistencies were inconsistent. 
+# Nothing I did made the data seem trustworthy for most of the homes, so here I have
+# resorted to cutting out those overlapping periods in the data. I am filtering out
+# around 161,000 observations (of 3.5 million, so < 5% of the data), but I can't come
+# up with any other way to have cleaned data that I actually trust, short of 
+# pasting each individual file together by hand. Even then, there are formatting
+# issues that would take days to resolve. So, this is the best way I have found
+# to get what I think is reliable SUMs data: filtering out the observations
+# with overlapping datapoints by pulling out instances when the time between
+# sampling points is not consistent.
+
+sums_new <- sums_clean %>% 
+  group_by(area, home) %>% 
+  mutate(time_diff = lead(datetime_sums) - datetime_sums,
+         time_diff_indicator = if_else(time_diff != rollmean(time_diff, 4), 1, 0)) %>% 
+  filter(time_diff_indicator != 1) %>%
+  ungroup() %>% 
+  arrange(area, home, winter_id, datetime_sums)
+
+# write_rds(sums_new, "Output/sums_clean.rds")
 ```

### 2020-02-06 sums_summary.Rmd
SUMs cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 9 Jan 2020, Updated 30 Jan 2020"
+date: "Started 9 Jan 2020, Updated 7 Feb 2020"
 output: pdf_document
@@
 
+thresholds_function(sums_filtered_date, 25, 1.5)
+thresholds_function(sums_filtered_date, 27, 1.5)
 thresholds_function(sums_filtered_date, 30, 1.5)
-thresholds_function(sums_filtered_date, 40, 1.5)
-thresholds_function(sums_filtered_date, 50, 1.5)
 ```
@@
 
-thresholds <- sums_filtered_date %>% 
+thresholds <- data %>% 
   group_by(area, home, winter_id, date_sums) %>%
-  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0)) %>% 
+  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
   summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
@@
 thresholds_percent(sums_filtered_date, 25)
+thresholds_percent(sums_filtered_date, 27)
 thresholds_percent(sums_filtered_date, 30)
-thresholds_percent(sums_filtered_date, 35)
-thresholds_percent(sums_filtered_date, 40)
-thresholds_percent(sums_filtered_date, 45)
-thresholds_percent(sums_filtered_date, 50)
-thresholds_percent(sums_filtered_date, 55)
 ```

### 2020-02-07 sums_summary.Rmd
small updates to SUMs functions

@@
   arrange(area, home)
+
+sums_winter1_placebo <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  ungroup() %>% 
+  arrange(area, home)
 ```
@@
                                       lead(threshold_temp_event) == "Temp < threshold", 
-                                      "event", threshold_temp_event)) %>% 
-  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+                                      "event_end", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
   mutate(time_sums = as.character(time_sums),
@@
   spread(threshold_temp_event, datetime_sums) %>% 
-  mutate(event = lead(event),
-         time_diff_min = (event - start),
-         time_diff_min = as.numeric(time_diff_min/60)) %>% 
-  filter(!is.na(time_diff_min)) %>% 
-  mutate(next_event = (lead(start)-event),
-         next_event = as.numeric(next_event)/3600,
-         event2 = if_else(next_event <= gap, lead(event), event),
-         time_diff_min2 = (event - start),
-         time_diff_min2 = as.numeric(time_diff_min2/60),
-         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
-         time_diff_min3 = (event3 - start),
-         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
-  distinct(event3, .keep_all = TRUE) %>% 
-  mutate(time_diff_min4 = (event3 - start),
-         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
-  filter(!is.na(time_diff_min4)) %>% 
+  mutate(event_end = lead(event_end),
+         event_length_min = (event_end - start),
+         event_length_min = as.numeric(event_length_min/60)) %>% 
+  filter(!is.na(event_length_min)) %>% 
+  mutate(hrs_till_next_event = (lead(start)-event_end),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
+         event_end2 = if_else(hrs_till_next_event <= gap, lead(event_end), event_end),
+         event_end3 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap, 
+                              lead(event_end, 2), event_end2),
+         event_end4 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap
+                              & lead(hrs_till_next_event, 2) <= gap, 
+                              lead(event_end, 3), event_end3),
+         event_end5 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap
+                              & lead(hrs_till_next_event, 2) <= gap & lead(hrs_till_next_event, 3) <= gap, 
+                              lead(event_end, 4), event_end4),
+         event_end6 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap
+                              & lead(hrs_till_next_event, 2) <= gap & lead(hrs_till_next_event, 3) <= gap
+                              & lead(hrs_till_next_event, 4) <= gap, 
+                              lead(event_end, 5), event_end5)) %>% 
+  distinct(event_end6, .keep_all = TRUE) %>% 
+  mutate(event_length_min = (event_end6 - start),
+         event_length_min = as.numeric(event_length_min/60),
+         hrs_till_next_event = (lead(start)-event_end),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
+  filter(!is.na(event_length_min)) %>% 
   group_by(area, home, winter_id, date_sums) %>% 
-  summarize(n = n(), mean_time = mean(time_diff_min4))
+  summarize(n = n(), mean_time = mean(event_length_min))
 

### 2020-02-13 home_char_act_summary.Rmd
summary stats update

@@
 table(table_list$winter_id, table_list$area, table_list$home_type, useNA = "ifany")
+
+table(home_char_summary$stove_age)
 ```
@@
 
+home_char_filtered %>% 
+  count(stove_age)
+
 # Summarize factor vars

### 2020-02-13 sums_summary.Rmd
summary stats update

@@
 ```
+
+# Closer look at negative temperatures
+```{r}
+total_homes <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  count(home)
+
+home_neg <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  mutate(neg_temp = if_else(temp_c_sums < 0, 1, 0)) %>% 
+  filter(neg_temp == 1) %>% 
+  count(home) %>% 
+  mutate(home_neg_temp = 1) %>% 
+  select(home, home_neg_temp)
+
+total_instruments <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  count(ibid)
+
+instrument_neg <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  mutate(neg_temp = if_else(temp_c_sums < 0, 1, 0)) %>% 
+  filter(neg_temp == 1) %>% 
+  count(ibid) %>% 
+  mutate(instrument_neg_temp = 1) %>% 
+  select(ibid, instrument_neg_temp)
+
+sums_neg_check <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  left_join(home_neg, by = "home") %>% 
+  left_join(instrument_neg, by = "ibid") %>% 
+  mutate(home_neg_temp = if_else(is.na(home_neg_temp), 0, home_neg_temp),
+         instrument_neg_temp = if_else(is.na(instrument_neg_temp), 0, instrument_neg_temp)) %>% 
+  # group by both home_neg_temp and instrument_neg_temp to check summaries
+  group_by(home_neg_temp) %>% 
+  summarize("mean" = mean(temp_c_sums, na.rm = TRUE),
+            "n" = n(),
+            "min" = min(temp_c_sums, na.rm = TRUE),
+            "median" = median(temp_c_sums, na.rm = TRUE),
+            "max" = max(temp_c_sums, na.rm = TRUE))
+sums_neg_check
+```
+
+The summaries above indicate that homes that had negative temperatures have similar
+summary stats to those that did not have negative temps. The same is true for
+the instruments used to measure temperature: instruments that had negative readings
+have similar summary stats to those without negative readings. This gives some
+reassurance that there are not any obvious trends associated with homes or
+instruments that recorded negative temperatures during sampling.
 
@@
 ```
+
+# Define events by rate of change in temperature
+```{r}
+sums_temp_change <- sums_clean %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, home) %>% 
+  mutate(time_diff = lead(datetime_sums) - datetime_sums) %>% 
+  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
+         lead_temp = if_else(time_diff > 25, temp_c_sums, lead_temp),
+         temp_diff = lead_temp - temp_c_sums,
+         temp_diff_av = abs(lead_temp - temp_c_sums),
+         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
+         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
+                              temp_diff_check)) %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_sums)
+
+summary_by_area <- sums_temp_change %>% 
+  group_by(area, home, date_sums) %>% 
+  summarize("events" = sum(heat_event, na.rm = TRUE)) %>% 
+  group_by(area) %>% 
+  summarize("Mean events" = mean(events),
+            "SD events" = sd(events),
+            "Min events" = min(events),
+            "Median events" = median(events),
+            "Max events" = max(events))
+summary_by_area
+
+temp_diff_by_area <- sums_temp_change %>% 
+  mutate(time_diff = as.numeric(time_diff)) %>% 
+  filter(time_diff >= 5 & time_diff < 30) %>% 
+  group_by(area) %>% 
+  summarize("mean temp diff" = mean(temp_diff_av, na.rm = TRUE),
+            "mean time diff" = mean(time_diff, na.rm = TRUE))
+temp_diff_by_area
+```

### 2020-02-14 exposure_initial_analysis.Rmd
exposure analysis - simple associations

@@
 library(zoo)
+library(lme4)
+library(lmerTest)
 ```
@@
 ```{r}
+# Mean PM for each placebo home over winter 1
 pm_home_mean <- pm_clean %>% 
@@
 
+# Join demographics with mean PM and format for analysis
 pm_demographics <- pm_home_mean %>% 
@@
 
+# Join home characteristics with mean PM and format for analysis
 pm_home_char <- pm_home_mean %>% 
@@
                                     "11-15 years old", "16 + years old",
-                                    "Do not know")))
+                                    "Do not know")),
+         stove_cert = factor(stove_cert,
+                             levels = c("Yes", "No", "Do not know")),
+         home_sqft_cat = cut(home_sqft, breaks = c(0, 500, 1000, 2000, 3000, 5000),
+                         labels = c("<500", "500-999", "1000-1999", 
+                                    "2000-2999", "3000+"),
+                         right = FALSE),
+         home_floors = factor(home_floors, levels = c(1,2,3)),
+         home_bedrooms_cat = cut(home_bedrooms, breaks = c(0, 2, 4, 10),
+                                 labels = c("0 or 1", "2 or 3", "4+"),
+                                 right = FALSE),
+         home_windows_cat = cut(home_windows, breaks = c(0, 5, 10, 20, 30, 50),
+                                labels = c("0-5", "6-10", "11-20", "21-30", "30+")))
 
-pm_home_char <- pm_home_mean %>% 
-  left_join(home_char_clean, by = c("area", "home")) %>% 
+# Join stove use with mean PM and format for analysis
+pm_stove_use <- pm_home_mean %>% 
+  left_join(stove_use_clean, by = c("area", "home")) %>% 
   filter(winter_id == 1) %>% 
-  mutate(stove_age = factor(stove_age,
-                         levels = c("0-5 years old", "6-10 years old",
-                                    "11-15 years old", "16 + years old",
-                                    "Do not know")))
+  mutate(burn_level = factor(burn_level,
+                         levels = c("No burning", "Light burning",
+                                    "Average burning", "Heavy burning"),
+                         labels = c("No burning", "Light", "Average", "Heavy"))) %>% 
+  mutate(wood_cords = if_else(wood_cords > 10, 10, wood_cords))
+
+# Join demographics with mean PM and format for analysis
+stove_grades <- stove_grades_clean %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(home) %>% 
+  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade))
+
+pm_stove_grades <- pm_home_mean %>% 
+  left_join(stove_grades, by = c("area", "home")) %>% 
+  filter(winter_id == 1)
+
+# Join home activity with mean PM and format for analysis
+pm_home_act <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(pm < 500) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  arrange(area, home, sampling_day) %>% 
+  group_by(area, home, sampling_day) %>% 
+  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  mutate(home_act_day = as.factor(sampling_day)) %>% 
+  arrange(area, home, home_act_day) %>% 
+  left_join(home_act_clean, by = c("area", "home", "home_act_day")) %>% 
+  filter(winter_id == 1)
+
+# Join moisture content with mean PM and format for analyzis
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
+  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+moisture_filtered <- moisture_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>%
+  arrange(area, home)
+
+moisture_filtered_time <- moisture_filtered %>% 
+  # fix wrong date
+  mutate(moisture_date = as.character(moisture_date),
+         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
+                                 "2017-04-10", moisture_date),
+         moisture_date = ymd(moisture_date)) %>% 
+  mutate(moisture_date = ymd(moisture_date),
+         first_date = ymd(first_date),
+         time_diff = moisture_date - first_date,
+         time_diff = as.numeric(abs(time_diff))) %>% 
+  group_by(home) %>% 
+  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
+  ungroup() %>% 
+  filter(time_diff_rank == 1)
+
+pm_moisture <- pm_home_mean %>% 
+  left_join(moisture_filtered_time, by = c("area", "home")) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(moisture_ave_5 = moisture_ave/5) %>% 
+  filter(!is.na(moisture_ave_5))
 ```
@@
 ```{r}
+# *income*, *education*
 pm_demographics_results <- lm(mean_pm ~ education, data = pm_demographics)
@@
 
-pm_home_char_results <- lm(mean_pm ~ stove_age, data = pm_home_char)
+# stove_age, stove_cert, *home_sqft_cat*, *secondary_heat*, *home_floors*
+# home_pets, home_furry, *home_windows_cat*, *home_bedrooms_cat*, *residents_smoke*
+pm_home_char_results <- lm(mean_pm ~ home_bedrooms_cat, data = pm_home_char)
 summary(pm_home_char_results)
+plot(pm_home_char_results)
+
+# burn_level, wood_collect, wood_cords
+pm_stove_use_results <- lm(mean_pm ~ wood_collect, data = pm_stove_use)
+summary(pm_stove_use_results)
+
+pm_stove_use_plot <- pm_stove_use %>% 
+  ggplot() + 
+    geom_boxplot(aes(burn_level, mean_pm))
+pm_stove_use_plot
+
+# *stove_grade*
+pm_stove_grades_results <- lm(mean_pm ~ stove_grade, data = pm_stove_grades)
+summary(pm_stove_grades_results)
+anova(pm_stove_grades_results)
+
+pm_stove_grades_plot <- pm_stove_grades %>% 
+  ggplot() + 
+    geom_boxplot(aes(stove_grade, mean_pm))
+pm_stove_grades_plot
+
+# home_act_sweep, home_act_windows, home_act_door, *home_act_smoking*
+# home_act_incense, home_act_candle
+pm_home_act_results <- lmer(mean_pm ~ home_act_sweep + (1 | home), data = pm_home_act)
+summary(pm_home_act_results)
+
+# *moisture_ave_5*
+pm_moisture_results <- lm(mean_pm ~ moisture_ave_5, data = pm_moisture)
+summary(pm_moisture_results)
 ```

### 2020-02-14 sums_summary.Rmd
exposure analysis - simple associations

@@
   filter(winter_id == 1) %>% 
+  #filter(treatment == "Placebo") %>% 
   group_by(area, home) %>% 
@@
          temp_diff_av = abs(lead_temp - temp_c_sums),
-         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
+         # change temp in the following line
+         temp_diff_check = if_else(temp_diff >= 2, 1, 0),
          heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
@@
 summary_by_area <- sums_temp_change %>% 
-  group_by(area, home, date_sums) %>% 
+  group_by(area, treatment, home, date_sums) %>% 
   summarize("events" = sum(heat_event, na.rm = TRUE)) %>% 
-  group_by(area) %>% 
+  group_by(area, treatment) %>% 
   summarize("Mean events" = mean(events),
@@
             "Min events" = min(events),
+            "25%" = quantile(events, 0.25),
             "Median events" = median(events),
+            "75%" = quantile(events, 0.75),
             "Max events" = max(events))
@@
   filter(time_diff >= 5 & time_diff < 30) %>% 
-  group_by(area) %>% 
+  group_by(area, treatment) %>% 
   summarize("mean temp diff" = mean(temp_diff_av, na.rm = TRUE),

### 2020-02-27 child_health_data_cleaning.Rmd
new RMDs for data cleaning

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 27 Feb 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")
```

################# At Home Log ###############

# Load and merge files from different locations
```{r}
wmt_athomelog_initial <- read_xlsx("Input/WMT/wmt_athomelog_20200227.xlsx") %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "WMT",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

nn_athomelog_initial <- read_xlsx("Input/NN/nn_athomelog_20200227.xlsx") %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "NN",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

ak_athomelog_initial <- read_xlsx("Input/AK/ak_athomelog_20200227.xlsx") %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "AK",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

# Combine logs from all areas
kids_at_home_log <- rbind(wmt_athomelog_initial, nn_athomelog_initial,
                          ak_athomelog_initial) %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_at_home_log, "Output/kids_at_home_log.rds")
```

################# Medical Visit ###############

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         lethargy = as.character(Lethargy),
         retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         resp_rate = as.numeric(RR),
         heart_rate = as.numeric(HR),
         wheeze = as.character(Wheeze),
         stridor = as.character(Stridor),
         crackles = as.character(Crackles),
         sp02 = as.numeric(SP02),
         hr_sp02 = as.numeric(HRPulseOx),
         referred = as.character(Referred),
         med_visit_notes = as.character(Notes)) %>% 
  select(home_winter_id:med_visit_notes) %>% 
  replace_with_na_all(condition = ~.x == -999)

nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         lethargy = as.character(Lethargy),
         retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         resp_rate = as.numeric(RR),
         heart_rate = as.numeric(HR),
         wheeze = as.character(Wheeze),
         stridor = as.character(Stridor),
         crackles = as.character(Crackles),
         sp02 = as.numeric(SP02),
         hr_sp02 = as.numeric(HRPulseOx),
         referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  select(home_winter_id:med_visit_notes) %>% 
  replace_with_na_all(condition = ~.x == -999)

ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "AK",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         lethargy = as.character(Lethargy),
         retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         resp_rate = as.numeric(RR),
         heart_rate = as.numeric(HR),
         wheeze = as.character(Wheeze),
         stridor = as.character(Stridor),
         crackles = as.character(Crackles),
         sp02 = as.numeric(SP02),
         hr_sp02 = as.numeric(HRPulseOx),
         referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  select(home_winter_id:med_visit_notes) %>% 
  replace_with_na_all(condition = ~.x == -999)

# Combine logs from all areas
kids_med_visit <- rbind(wmt_medicalvisit_initial, nn_medicalvisit_initial,
                          ak_medicalvisit_initial) %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

# write_rds(kids_med_visit, "Output/kids_med_visit.rds")
```

################# Birth Infancy ###############

# Load and merge files from different locations
```{r}
wmt_birthinfancy_initial <- read_xlsx("Input/WMT/wmt_birthinfancy_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         gender = as.character(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.character(CurrentlyBreastfeeding),
         breastfeed_past = as.character(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.character(FluShot),
         flu_shot_planned = as.character(FluShotPlanned),
         asthma_meds = as.character(AsthmaMeds),
         child_school = as.character(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.character(SibSchool)) %>% 
  select(home_winter_id:sibling_school)

# Waiting on files from NN and AK
```

################# Child Home Winter  ###############

# Load and merge files from different locations
```{r}
wmt_childhomewinter_initial <- read_xlsx("Input/WMT/wmt_childhomewinter_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         home_id = as.character(HomeID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         dr_relook = DrRelook,
         ari_status = as.character(Status),
         symptom_free_date = ymd(SymptomFreeDate),
         csq_duedate = ymd(CSQDueDate),
         csq_date = ymd(CSQDate),
         csq_time = as.character(CSQTime),
         dr_comments = as.character(DrComments),
         dr_history = as.character(DrHistory),
         notify_parents = as.character(NotifyParents),
         notes_to_parents = as.character(NotesToParents),
         any_ari = as.character(AnyARI),
         ari_comments = as.character(WinterARIComments)) %>% 
  select(home_winter_id:ari_comments)

# Waiting on files from NN and AK
```

################# CSQ  ###############

# Load and merge files from different locations
```{r}
wmt_csq_initial <- read_xlsx("Input/WMT/wmt_csq_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         csq_date = ymd(Date),
         sick_2_weeks = as.character(Illin2Weeks),
         sick = as.character(Sick),
         ari = as.character(ARI),
         still_sick = as.character(StillSick),
         first_symptom_date = ymd(FirstSymptomDate),
         see_dr = as.character(SeeDr),
         dr_diagnosis = as.character(DrDiagnosis),
         new_meds = as.character(NewMeds),
         fever = as.character(Fever),
         temp = as.numeric(Temperature),
         rapid_breathing = as.character(RapidBreathing),
         difficult_breathing = as.character(DifficultBreathing),
         noisy_breathing = as.character(NoisyBreathing),
         runny_nose = as.character(RunnyNose),
         cough = as.character(Coughing),
         wheeze = as.character(Wheezing),
         symptom_free_date = ymd(SymptomFreeDate),
         comments = as.character(Comments)) %>% 
  select(home_winter_id:comments)

# Waiting on files from NN and AK
```

################# Medical Records ###############

# Load and merge files from different locations
```{r}
wmt_medicalrecords_initial <- read_xlsx("Input/WMT/wmt_medicalrecords_20200227.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         visit_date = ymd(VisitDate),
         dr_diagnosis = as.character(DrDiag),
         ari = as.character(ARI),
         comments = as.character(DrComments)) %>% 
  select(home_winter_id:comments)

# Waiting on files from NN and AK
```

### 2020-02-27 stove_grade_usage_summary.Rmd
summary stat updates
new RMDs for data cleaning

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 27 Jan 2020"
+date: "Started 15 Jan 2020, Updated 27 Feb 2020"
 output: pdf_document 
@@
 kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") 
+demographics_clean <- read_rds("Output/demographics_clean.rds")
 pm_clean <- read_rds("Output/pm_clean.rds")
 ```
+
+############ Use only most recent stove grades for analysis ###############
 
@@
 stove_grade_summary <- stove_grades_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  #filter(!is.na(stove_grade)) %>% 
+  group_by(area, winter_id, stove_grade) %>% 
+  distinct(home) %>% 
+  summarize(n()) %>% 
+  filter(!is.na(stove_grade))
+kable(stove_grade_summary, align = "c")
+
+stove_grades_total <- stove_grades_clean %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   select(area, home, stove_grade) %>% 
-  #filter(!is.na(stove_grade)) %>% 
-  group_by(area, stove_grade) %>% 
-  distinct(home) %>% 
-  summarize(n())
-kable(stove_grade_summary, align = "c")
+  arrange(area, home) 
+```
+
+# Generate random sample of stoves to re-grade and validate grading methods
+```{r}
+# 10% of homes from each area, earlier winters
+# send new ID linked to new ID to Kathrene
+
+random_sample_wmt <- stove_grades_clean %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  mutate(winter_id = as.factor(winter_id)) %>% 
+  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
+  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  filter(cutoff_date == 1) %>% 
+  filter(area == "WMT") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
+  arrange(area, home) %>% 
+  group_by(area) %>% 
+  sample_n(17)
+write_csv(random_sample_wmt, "Output/random_sample_wmt.csv")
+
+random_sample_nn <- stove_grades_clean %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  mutate(winter_id = as.factor(winter_id)) %>% 
+  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
+  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  filter(cutoff_date == 1) %>% 
+  filter(area == "NN") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
+  arrange(area, home) %>% 
+  group_by(area) %>% 
+  sample_n(8)
+write_csv(random_sample_nn, "Output/random_sample_nn.csv")
+
+random_sample_ak <- stove_grades_clean %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  mutate(winter_id = as.factor(winter_id)) %>% 
+  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
+  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  filter(cutoff_date == 1) %>% 
+  filter(area == "AK") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
+  arrange(area, home) %>% 
+  group_by(area) %>% 
+  sample_n(5)
+write_csv(random_sample_ak, "Output/random_sample_ak.csv")
 ```

### 2020-02-28 child_health_data_cleaning.Rmd
data cleaning updates

@@
 # write_rds(kids_med_visit, "Output/kids_med_visit.rds")
+
+save_var_names <- data.frame(t(kids_med_visit)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+
+write_csv(save_var_names, "Output/medvisit_var_names.csv")
 ```
@@
 # Waiting on files from NN and AK
+
+save_var_names <- data.frame(t(wmt_birthinfancy_initial)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+
+write_csv(save_var_names, "Output/birthinfancy_var_names.csv")
 ```
@@
 
-################# CSQ  ###############
+################# CSQ - child symptom questionnaire  ###############
 
@@
 # Waiting on files from NN and AK
+
+save_var_names <- data.frame(t(wmt_csq_initial)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+
+write_csv(save_var_names, "Output/csq_var_names.csv")
 ```

### 2020-03-12 demographics_summary.Rmd
Updates to summary stats/plots

@@
 author: "Ethan Walker"
-date: "Started 27 Jan 2020, Updated 27 Jan 2020"
+date: "Started 27 Jan 2020, Updated 12 March 2020"
 output: pdf_document
@@
 ```
+
+```{r}
+# Filter for Winter 1; remove NAs
+
+demographics_filtered <- demographics_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(!is.na(hispanic_child)) %>% 
+  mutate(income = factor(income,
+                         levels = c("Less than $20,000", "$20,000 to $29,999",
+                                    "$30,000 to $39,999", "$40,000 to $49,999",
+                                    "$50,000 to $74,999", "$75,000 to $99,999",
+                                    "$100,000 or more")),
+         education = factor(education,
+                            levels = c("Less than high school",
+                                       "High school diploma or GED",
+                                       "Some college",
+                                       "College degree")))
+```
+
+```{r}
+# Demographics summary stats function
+
+demographics_function <- function(data, var, title_label, y_label) {
+
+  demographics_plot <- data %>% 
+  filter(winter_id == 1) %>% 
+  select(area, var) %>% 
+  rename(plot_var = var) %>% 
+  filter(!is.na(plot_var)) %>% 
+  ggplot() +
+    geom_bar(aes(plot_var, fill = area)) +
+    theme_minimal() +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 40, vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 18, colour = "black"),
+          legend.text = element_text(size = 14, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    labs(title = title_label) +
+    ylab(y_label) +
+    scale_fill_manual(values = jv_palette)
+  demographics_plot
+
+}
+
+demographics_function(demographics_filtered, "gender_parent", "Parent sex", "count")
+demographics_function(demographics_filtered, "hispanic_parent", "Parent Hispanic ethnicity", "count")
+demographics_function(demographics_filtered, "race_parent", "Parent race", "count")
+demographics_function(demographics_filtered, "hispanic_child", "Child Hispanic ethnicity", "count")
+demographics_function(demographics_filtered, "race_child", "Child race", "count")
+demographics_function(demographics_filtered, "age_child_nov", "Child age Nov 1st", "count")
+demographics_function(demographics_filtered, "income", "Household income", "count")
+demographics_function(demographics_filtered, "education", "Parent education", "count")
+demographics_function(demographics_filtered, "total_residents", "Total residents", "count")
+demographics_function(demographics_filtered, "residents_under_five", "Residents under 5", "count")
+```

### 2020-03-12 home_char_act_cleaning.Rmd
Updates to summary stats/plots

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 23 Jan 2020"
+date: "Started 13 Jan 2020, Updated 4 March 2020"
 output: pdf_document
@@
   # cleaning based on checking odd values with paper logs
-  mutate(residents_under5 = if_else(residents_under5 == 10, 1, residents_under5)) %>% 
+  mutate(residents_under5 = if_else(residents_under5 == 10, 1, residents_under5),
+         residents_under5 = if_else(home == "CH307" | home == "CH323", 1, residents_under5)) %>% 
   replace_with_na(replace = list(home_year_built = 0)) %>% 

### 2020-03-16 child_health_data_summary.Rmd
New summary plots

---
title: "Kids health and child data summary"
author: "Ethan Walker"
date: "Started 16 March 2020, Updated 16 March 2020"
output: 
  powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 5, fig.width = 7)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")

kids_med_visit <- read_rds("Output/kids_med_visit.rds") %>% 
  group_by(area, home_winter_id, child_id_num, med_visit_date) %>% 
  distinct(med_visit_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(temp1 = if_else(temp1 < 90, 0, temp1)) %>% 
  mutate(temp2 = if_else(temp2 < 90, 0, temp2)) %>% 
  mutate(temp3 = if_else(temp3 < 90, 0, temp3)) %>% 
  mutate(sp02 = if_else(sp02 < 80, 0, sp02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-9999, -99, 0)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  arrange(home_winter_id, med_visit_date) %>% 
  group_by(area, home_winter_id, med_visit_date) %>% 
  mutate(temp_ave = mean(value, na.rm = TRUE)) %>% 
  # spread 3 moisture samples back to wide format
  group_by(area, home_winter_id, med_visit_date, temp_ave) %>% 
  spread(sample, value) %>% 
  ungroup()

```

```{r}
medvisit_plot <- kids_med_visit %>% 
  filter(winter_id == 1) %>% 
  select(area, resp_rate, heart_rate, sp02, hr_sp02, temp_ave) %>% 
  pivot_longer(cols = c("resp_rate":"temp_ave"), names_to = "var") %>% 
  filter(!is.na(value)) %>% 
  ggplot(aes(area, value)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    theme_minimal() +
    theme(strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(~var, scales = "free_y")
medvisit_plot
```

```{r}
# Cardio plot function
medvisit_summary_function <- function(data, var, title_label, y_label) {

  medvisit_summary <- data %>% 
    filter(winter_id == 1) %>% 
    #select(var) %>% 
    select(area, var) %>% 
    group_by(area) %>% 
    summarize_all(list(mean=mean, sd=sd, min=min, med=median, max=max), na.rm = TRUE)
  medvisit_summary <<- medvisit_summary

  medvisit_plot <- data %>% 
    filter(winter_id == 1) %>% 
    select(area, var) %>% 
    rename(plot_var = var) %>% 
    filter(!is.na(plot_var)) %>% 
    ggplot(aes(area, plot_var)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
    ylab(y_label) +
    theme(title = element_text(size = 20, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) 
  medvisit_plot

}
```

```{r}
medvisit_summary_function(kids_med_visit, "resp_rate", "Respiratory rate", "breaths/min")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "heart_rate", "Heart rate", "beats/min")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "sp02", "Sp02", "%")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "hr_sp02", "Heart rate from Sp02", "beats/min")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "temp_ave", "Temperature", "Degrees F")
```

\pagebreak

```{r}
medvisit_plot <- kids_med_visit %>% 
  filter(winter_id == 1) %>% 
  select(area, med_visit_ari, lethargy, retractions, wheeze, stridor, 
         crackles, referred) %>% 
  pivot_longer(cols = c("med_visit_ari":"referred"), names_to = "var") %>% 
  filter(!is.na(value)) %>% 
  ggplot() +
    geom_bar(aes(value, fill = area)) +
    theme_minimal() +
    theme(strip.text = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black", 
                                     angle = 25, vjust = 0.6),
          legend.text = element_text(size = 16, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    facet_wrap(~var, scales = "free_y")
medvisit_plot
```

```{r}
# Cardio plot function
medvisit_summary_function <- function(data, var, title_label, y_label) {

  medvisit_summary <- data %>% 
    filter(winter_id == 1) %>% 
    #select(var) %>% 
    select(area, var) %>% 
    rename(plot_var = var) %>% 
    group_by(area, plot_var) %>% 
    summarize("n" = n())
  medvisit_summary <<- medvisit_summary

  medvisit_plot <- data %>% 
    filter(winter_id == 1) %>% 
    select(area, var) %>% 
    rename(plot_var = var) %>% 
    filter(!is.na(plot_var)) %>% 
    ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    theme(strip.text = element_text(size = 16, colour = "black"),
          title = element_text(size = 20, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black", 
                                     angle = 25, vjust = 0.6),
          legend.text = element_text(size = 16, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16, colour = "black"),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank()) +
    labs(title = title_label) +
    ylab(y_label) +
    scale_fill_manual(values = jv_palette)
  medvisit_plot

}
```

```{r}
medvisit_summary_function(kids_med_visit, "med_visit_ari", "Med visit ARI", "count")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "lethargy", "Lethargy", "count")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "retractions", "Retractions", "count")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "wheeze", "Wheeze", "count")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "stridor", "Stridor", "count")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "crackles", "Crackles", "count")
```

\pagebreak

```{r}
medvisit_summary_function(kids_med_visit, "referred", "Referred", "count")
```

### 2020-03-16 demographics_summary.Rmd
New summary plots

@@
 author: "Ethan Walker"
-date: "Started 27 Jan 2020, Updated 12 March 2020"
-output: pdf_document
+date: "Started 27 Jan 2020, Updated 13 March 2020"
+output: 
+  powerpoint_presentation
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
+                      include = TRUE, warning = FALSE,
+                      fig.height = 6, fig.width = 8)
 ```
@@
 
+```{r}
 # Load individual datasets
-```{r}
+
 kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
@@
 
+```{r}
 # Filter for homes that have PM data
-```{r}
+
 # 93 Winter 1 placebo homes have PM data
@@
 # Summarize demographics data
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars
@@
                                        "Some college",
-                                       "College degree")))
+                                       "College degree")),
+         total_residents = as.numeric(total_residents),
+         race_parent = factor(race_parent,
+                              levels = c("American Indian/Alaskan Native",
+                                         "Asian", "White",
+                                         "More than one race"),
+                              labels = c("AI/AN", "Asian", "White", "Multiple")),
+         race_child = factor(race_child,
+                              levels = c("American Indian/Alaskan Native",
+                                         "White", "More than one race"),
+                              labels = c("AI/AN", "White", "Multiple")))
 ```
@@
 }
+```
 
+```{r}
 demographics_function(demographics_filtered, "gender_parent", "Parent sex", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "hispanic_parent", "Parent Hispanic ethnicity", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "race_parent", "Parent race", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "hispanic_child", "Child Hispanic ethnicity", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "race_child", "Child race", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "age_child_nov", "Child age Nov 1st", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "income", "Household income", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "education", "Parent education", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "total_residents", "Total residents", "count")
+```
+
+\pagebreak
+
+```{r}
 demographics_function(demographics_filtered, "residents_under_five", "Residents under 5", "count")

### 2020-03-16 home_char_act_summary.Rmd
New summary plots

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 27 Jan 2020"
+date: "Started 15 Jan 2020, Updated 16 March 2020"
 output:
   pdf_document: default
+  powerpoint_presentation: default
   word_document: default
@@
 knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
-                      include = TRUE, warning = FALSE)
+                      include = TRUE, warning = FALSE,
+                      fig.height = 5, fig.width = 7)
 ```
@@
 library(tidyverse)
-library(readr)
+library(readxl)
+library(naniar)
+library(lubridate)
+library(zoo)
 library(knitr)
-library(readxl)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 # Load dataset
-home_char_clean <- read_rds("Output/home_char_clean.rds") 
+home_char_clean <- read_rds("Output/home_char_clean.rds") %>% 
+  mutate(chimney_clean = factor(chimney_clean,
+                                levels = c("Less than 6 months ago",
+                                           "6-12 months ago", "12-18 months ago",
+                                           "More than 18 months ago"),
+                                labels = c("< 6 months ago",
+                                           "6-12 months ago", "12-18 months ago",
+                                           "> 18 months ago")),
+         stove_age = factor(stove_age,
+                            levels = c("0-5 years old", "6-10 years old",
+                                       "11-15 years old", "16 + years old",
+                                       "Do not know")))
+
 home_act_clean <- read_rds("Output/home_act_clean.rds")
@@
 
-# Summary data for home characteristics
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars
@@
 
-# Summary data for home activity
-```{r}
+```{r, include=FALSE}
 # Summarize factor vars
@@
 
-# Home characteristics summary plots
+```{r, fig.height=8}
+home_char_plot <- home_char_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, residents_smoke, residents_smoke_inside, home_damp:home_insert) %>% 
+  pivot_longer(cols = c("residents_smoke":"home_insert"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  ggplot() +
+    geom_bar(aes(value, fill = area)) +
+    theme_minimal() +
+    theme(strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 14, colour = "black", 
+                                     angle = 25, vjust = 0.6),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, ncol = 2)
+home_char_plot
+```
+
 ```{r}
@@
 
-# Home activity summary plots
+```{r, fig.height=8}
+home_act_plot <- home_act_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home_act_gas:home_act_sweep) %>% 
+  pivot_longer(cols = c("home_act_gas":"home_act_sweep"), names_to = "var") %>% 
+  filter(!is.na(value)) %>% 
+  ggplot() +
+    geom_bar(aes(value, fill = area)) +
+    theme_minimal() +
+    theme(strip.text = element_text(size = 16, colour = "black"),
+          axis.text.x = element_text(size = 14, colour = "black", 
+                                     angle = 25, vjust = 0.6),
+          legend.text = element_text(size = 16, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16, colour = "black"),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank()) +
+    facet_wrap(~var, ncol = 2)
+home_act_plot
+```
+
 ```{r}
@@
 
+```{r, include=FALSE}
 #### Abstract/manuscript tables ####
@@
 # Filter for homes that have PM data
-```{r}
+
 # 93 Winter 1 placebo homes have PM data
@@
 
-# Summarize home data
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars

### 2020-03-16 indoor_temp_rh_summary.Rmd
New summary plots

--- 2020-01-23 indoor_temp_rh_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 23 Jan 2020"
-output: pdf_document
+date: "Started 15 Jan 2020, Updated 16 March 2020"
+output:
+  powerpoint_presentation: default
+  pdf_document: default
 ---
@@
 library(tidyverse)
-library(readr)
+library(readxl)
+library(naniar)
+library(lubridate)
+library(zoo)
 library(knitr)
+jv_palette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 ```
@@
 indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+pm_clean <- read_rds("Output/pm_clean.rds")
 ```
 
-# Summary stats
-```{r}
+```{r, include=FALSE}
 temp_rh_summary <- indoor_temp_rh_clean %>%
@@
 # Boxplot function
-boxplots_temp_rh_function <- function(data, variable){
+boxplots_temp_rh_function <- function(data, var, title_label, y_label){
 
-boxplots_temp_rh <- data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  select(area, variable) %>% 
-  rename(plot_var = variable) %>% 
-  na.exclude() %>% 
+temp_rh_plot <- data %>% 
+    filter(winter_id == 1) %>% 
+    select(area, var) %>% 
+    rename(plot_var = var) %>% 
+    filter(!is.na(plot_var)) %>% 
     ggplot(aes(area, plot_var)) +
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
-    theme(axis.text.x = element_text(size = 16, colour = "black"),
+    labs(title = title_label, subtitle = "Mean = Red, Median = Blue") +
+    ylab(y_label) +
+    theme(title = element_text(size = 20, colour = "black"),
+          axis.text.x = element_text(size = 16, colour = "black"),
           axis.text.y = element_text(size = 16, colour = "black"),
-          axis.title.y = element_blank(),
+          axis.title.y = element_text(size = 16, colour = "black"),
           axis.title.x = element_blank(),
@@
           axis.ticks = element_blank()) 
-boxplots_temp_rh
+  temp_rh_plot
 
@@
 
-## Current indoor humidity (%)
 ```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current")
+boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_current", 
+                          "Current indoor humidity", "%")
 ```
 
-## Max indoor humidity (%)
 ```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max")
+boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_max", 
+                          "Max indoor humidity", "%")
 ```
 
-## Min indoor humidity (%)
 ```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min")
+boxplots_temp_rh_function(indoor_temp_rh_clean, "rh_indoor_min", 
+                          "Min indoor humidity", "%")
 ```
 
-## Current indoor temperature (C)
 ```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current")
+boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_current", 
+                          "Current indoor temperature", "Degrees C")
 ```
 
-## Max indoor temperature (C)
 ```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max")
+boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_max", 
+                          "Max indoor temperature", "Degrees C")
 ```
 
-## Min indoor temperature (C)
 ```{r}
-boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min")
+boxplots_temp_rh_function(indoor_temp_rh_clean, "temp_indoor_min", 
+                          "Min indoor temperature", "Degrees C")
 ```
+
+```{r}
+#### Abstract/manuscript tables ####
+
+# Filter for homes that have PM data
+
+# 93 Winter 1 placebo homes have PM data
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+temp_rh_filtered <- indoor_temp_rh_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+```
+
+```{r, include=FALSE}
+# Summarize numeric vars
+kids_temp_rh <- temp_rh_filtered %>%  
+  distinct(home, .keep_all = TRUE) %>%  
+  group_by(area) %>% 
+  summarize("Mean" = mean(rh_indoor_min, na.rm = TRUE), 
+            "SD" = sd(rh_indoor_min, na.rm = TRUE),
+            "min" = min(rh_indoor_min, na.rm = TRUE), 
+            "med" = median(rh_indoor_min, na.rm = TRUE), 
+            "max" = max(rh_indoor_min, na.rm = TRUE))
+kable(kids_temp_rh, digits = 1, align = "c")
+
+```

### 2020-03-16 moisture_summary.Rmd
New summary plots

--- 2020-01-28 moisture_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 28 Jan 2020"
-output: pdf_document
+date: "Started 22 Nov 2019, Updated 16 March 2020"
+output: powerpoint_presentation
 ---
@@
 library(tidyverse)
+library(readxl)
+library(naniar)
 library(lubridate)
-library(readr)
+library(zoo)
 library(knitr)
+jv_palette <- c("#330099","#CC0066","#FF6633", 
+                 "#0099CC", "#FF9900","#CC6633",
+                  "#FF3366", "#33CC99", "#33999")
 ```
@@
 
+```{r, include=FALSE}
 # Moisture content summary stats
 ## Average of 3 measurements
-```{r}
+
 summary_moisture <- moisture_clean %>%
@@
 
-# Was wood split?
-```{r}
+```{r, include=FALSE}
 wood_split <- moisture_clean %>% 
@@
 
-# Boxplots
-```{r, fig.height=6, fig.width=5}
+```{r, fig.height=6, fig.width=7}
 boxplots_moisture <- moisture_clean %>% 
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
     ylim(0, 50) +
-    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
-    labs(y = "Mean moisture content (%)") +
-    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
+    labs(title = "Mean wood moisture content") +
+    labs(y = "%") +
     theme(axis.text.x = element_text(size = 16, colour = "black"),
+          title = element_text(size = 20, colour = "black"),
           axis.text.y = element_text(size = 16, colour = "black"),
@@
 
+```{r, fig.height=6, fig.width=7}
+bar_plot <- moisture_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, moisture_split) %>% 
+  ggplot() +
+    geom_bar(aes(moisture_split, fill = area)) +
+    theme_minimal() +
+    labs(title = "Was wood split?") +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 40, vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 20, colour = "black"),
+          legend.text = element_text(size = 14, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    scale_fill_manual(values = jv_palette)
+bar_plot
+```
+
+```{r}
 #### Abstract/manuscript tables ####
@@
 
-# Summarize home data
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars
-kids_moisture <- moisture_filtered %>%  
+kids_moisture <- moisture_filtered_time %>%  
   group_by(area) %>% 

### 2020-03-16 pm_summary.Rmd
New summary plots

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 2 Feb 2020"
-output: pdf_document
+date: "Started 22 Nov 2019, Updated 16 March 2020"
+output: powerpoint_presentation
 ---
@@
             "Max PM" = max(pm, na.rm = TRUE))
-pm_by_day
+#pm_by_day
 ```
@@
 # Boxplots for PM by study area
-```{r, fig.height=6, fig.width=5}
+```{r, fig.height=6, fig.width=8}
 boxplots_kids_pm <- pm_clean %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(pm < 500) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
   ggplot(aes(area, pm)) +
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +

### 2020-03-16 stove_grade_usage_summary.Rmd
New summary plots

--- 2020-01-27 stove_grade_usage_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 27 Jan 2020"
+date: "Started 15 Jan 2020, Updated 16 March 2020"
 output: pdf_document 
@@
 library(readr)
+library(readxl)
 library(knitr)
+library(lubridate)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") 
+demographics_clean <- read_rds("Output/demographics_clean.rds")
 pm_clean <- read_rds("Output/pm_clean.rds")
@@
 
-# Summary of stove grades
 ```{r}
@@
 stove_grade_summary <- stove_grades_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  #filter(!is.na(stove_grade)) %>% 
+  group_by(area, winter_id, stove_grade) %>% 
+  distinct(home) %>% 
+  summarize(n()) %>% 
+  filter(!is.na(stove_grade))
+kable(stove_grade_summary, align = "c")
+
+stove_grades_total <- stove_grades_clean %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   select(area, home, stove_grade) %>% 
-  #filter(!is.na(stove_grade)) %>% 
-  group_by(area, stove_grade) %>% 
-  distinct(home) %>% 
-  summarize(n())
-kable(stove_grade_summary, align = "c")
+  arrange(area, home) 
 ```
 
+```{r}
+# Generate random sample of stoves to re-grade and validate grading methods
+
+# 10% of homes from each area, earlier winters
+# send new ID linked to new ID to Kathrene
+
+random_sample_wmt <- stove_grades_clean %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  mutate(winter_id = as.factor(winter_id)) %>% 
+  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
+  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  filter(cutoff_date == 1) %>% 
+  filter(area == "WMT") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
+  arrange(area, home) %>% 
+  group_by(area) %>% 
+  sample_n(17)
+#write_csv(random_sample_wmt, "Output/random_sample_wmt.csv")
+
+random_sample_nn <- stove_grades_clean %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  mutate(winter_id = as.factor(winter_id)) %>% 
+  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
+  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  filter(cutoff_date == 1) %>% 
+  filter(area == "NN") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
+  arrange(area, home) %>% 
+  group_by(area) %>% 
+  sample_n(8)
+#write_csv(random_sample_nn, "Output/random_sample_nn.csv")
+
+random_sample_ak <- stove_grades_clean %>% 
+  select(area, home, stove_grade, winter_id) %>% 
+  mutate(winter_id = as.factor(winter_id)) %>% 
+  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
+  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(!is.na(home)) %>% 
+  filter(cutoff_date == 1) %>% 
+  filter(area == "AK") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
+  arrange(area, home) %>% 
+  group_by(area) %>% 
+  sample_n(5)
+#write_csv(random_sample_ak, "Output/random_sample_ak.csv")
+```
+
+```{r, include=FALSE}
 ## 18 homes with grades from 2 winters
@@
 ### Was there an upgrade/downgrade to the stove? Check for comments/notes.
-```{r}
+
 multiple_grades <- stove_grades_clean %>% 
@@
 
-# Summary of stove use
-```{r}
+```{r, include=FALSE}
 # Summarize numeric vars
@@
 
-# Stove use summary plots
 ```{r}
+stove_grade_plot <- stove_grades_clean %>% 
+  filter(winter_id == 1) %>% 
+  select(area, stove_grade) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  ggplot() +
+    geom_bar(aes(stove_grade, fill = area)) +
+    theme_minimal() +
+    labs(title = "Stove grades") +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     angle = 40, vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 18, colour = "black"),
+          legend.text = element_text(size = 14, colour = "black"),
+          legend.title = element_text(size = 16, colour = "black"),
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    scale_fill_manual(values = jv_palette)
+stove_grade_plot
+```
+
+```{r}
+stove_use_data <- read_rds("Output/stove_use_clean.rds") %>% 
+  mutate(wood_cords = as.factor(wood_cords))
+
 stove_use_plot_function <- function(data, variable) {
@@
           axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 18, colour = "black"),
           legend.text = element_text(size = 14, colour = "black"),
           legend.title = element_text(size = 16, colour = "black"),
-          title = element_text(size = 18, colour = "black"),
           axis.title.y = element_blank(),
@@
 
-stove_use_plot_function(stove_use_clean, "burn_level")
-stove_use_plot_function(stove_use_clean, "wood_type")
-stove_use_plot_function(stove_use_clean, "wood_collect")
-stove_use_plot_function(stove_use_clean, "wood_collect_method")
+stove_use_plot_function(stove_use_data, "burn_level")
+stove_use_plot_function(stove_use_data, "wood_type")
+stove_use_plot_function(stove_use_data, "wood_collect")
+stove_use_plot_function(stove_use_data, "wood_collect_method")
+stove_use_plot_function(stove_use_data, "wood_cords")
 ```
 
+```{r}
 #### Abstract/manuscript tables ####

### 2020-03-17 exposure_initial_analysis.Rmd
exposure analysis updates

--- 2020-02-06 exposure_initial_analysis.Rmd
@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 2 Feb 2020"
+date: "Started 2 Feb 2020, Updated 17 March 2020"
 output: pdf_document
@@
 home_act_clean <- read_rds("Output/home_act_clean.rds")
-```
 
-# Combine datasets
-```{r}
+# Mean PM for each placebo home over winter 1
 pm_home_mean <- pm_clean %>% 
@@
   summarize(mean_pm = mean(pm, na.rm = TRUE))
+```
 
+# Demographics
+```{r}
 pm_demographics <- pm_home_mean %>% 
@@
   filter(winter_id == 1) %>% 
-  mutate(income = factor(income,
-                         levels = c("Less than $20,000", "$20,000 to $29,999",
-                                    "$30,000 to $39,999", "$40,000 to $49,999",
-                                    "$50,000 to $74,999", "$75,000 to $99,999",
-                                    "$100,000 or more"))) %>% 
-  mutate(education = factor(education,
-                            levels = c("Less than high school", 
-                                       "High school diploma or GED",
+  mutate(income = fct_collapse(income,
+                               "Less than $20,000" = "Less than $20,000",
+                               "$20,000-$49,999" = 
+                                 c("$20,000 to $29,999", "$30,000 to $39,999",
+                                   "$40,000 to $49,999"),
+                                "$50,000+" = 
+                                 c("$50,000 to $74,999", "$75,000 to $99,999",
+                                   "$100,000 or more")),
+         income = factor(income, levels = c("Less than $20,000", 
+                                            "$20,000-$49,999", 
+                                            "$50,000+"))) %>% 
+  mutate(education = fct_collapse(education,
+                                  "High school or less" = 
+                                    c("Less than high school", 
+                                      "High school diploma or GED"),
+                                  "Some college" = "Some college",
+                                  "College degree" = "College degree"),
+         education = factor(education, levels = c("High school or less", 
                                             "Some college", 
-                                       "College degree")))
+                                            "College degree"))) %>% 
+  mutate(race_parent = fct_collapse(race_parent,
+                                    "White" = "White",
+                                    "Non-white" = 
+                                      c("American Indian/Alaskan Native",
+                                        "Asian", "More than one race")),
+         race_parent = factor(race_parent, levels = c("White", "Non-white"))) %>% 
+  mutate(race_child = fct_collapse(race_child,
+                                    "White" = "White",
+                                    "Non-white" = 
+                                      c("American Indian/Alaskan Native",
+                                        "More than one race")),
+         race_child = factor(race_child, levels = c("White", "Non-white"))) %>%
+  mutate(total_residents = fct_collapse(total_residents,
+                                    "<5" = c("1", "2", "3", "4"),
+                                    "5+" = c("5", "6", "7", "8", "9", "10",
+                                             "11", "12", "13", "14", "15")),
+         total_residents = factor(total_residents, levels = c("<5", "5+"))) %>%
+  mutate(residents_under_five = fct_collapse(residents_under_five,
+                                    "1" = c("0", "1"),
+                                    "2+" = c("2", "3", "4", "5")),
+         residents_under_five = factor(residents_under_five, levels = c("1", "2+"))) %>% 
+  distinct(mean_pm, .keep_all = TRUE)
 
+table(pm_demographics$income)
+
+pm_demographics_results <- lm(mean_pm ~ residents_under_five, data = pm_demographics)
+summary(pm_demographics_results)
+plot(pm_demographics_results)
+```
+
+# Home characteristics
+```{r}
 pm_home_char <- pm_home_mean %>% 
@@
   filter(winter_id == 1) %>% 
-  mutate(stove_age = factor(stove_age,
+  mutate(home_type = fct_collapse(home_type,
+                                  "House" = "House",
+                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
+         home_type = factor(home_type,
+                            levels = c("House", "Other")),
+         home_year_built_cat = cut(home_year_built, breaks = c(0, 1985, 2020),
+                                 labels = c("<1985", "1985+"),
+                                 right = FALSE),
+         stove_age = factor(stove_age,
                          levels = c("0-5 years old", "6-10 years old",
                                     "11-15 years old", "16 + years old",
-                                    "Do not know")))
+                                    "Do not know")),
+         stove_age = fct_collapse(stove_age,
+                                  "0-5 years old" = "0-5 years old",
+                                  "6-15 years old" = c("6-10 years old",
+                                                       "11-15 years old"),
+                                  "16 + years old" = "16 + years old",
+                                  "Do not know" = "Do not know"),
+         stove_cert = factor(stove_cert,
+                             levels = c("Yes", "No", "Do not know")),
+         home_sqft_cat = cut(home_sqft, breaks = c(0, 1400, 5000),
+                         labels = c("<1400", "1400+"),
+                         right = FALSE),
+         home_floors_cat = cut(home_floors, breaks = c(0, 2, 4),
+                                 labels = c("1", "2+"),
+                                 right = FALSE),
+         home_bedrooms_cat = cut(home_bedrooms, breaks = c(0, 3, 10),
+                                 labels = c("<3", "3+"),
+                                 right = FALSE),
+         home_windows_cat = cut(home_windows, breaks = c(0, 10, 50),
+                                labels = c("<10", "10+"),
+                                 right = FALSE),
+         home_pets_cat = cut(home_pets, breaks = c(0, 2, 50),
+                                labels = c("<2", "2+"),
+                                 right = FALSE),
+         home_furry_cat = cut(home_furry, breaks = c(0, 2, 50),
+                                labels = c("<2", "2+"),
+                                 right = FALSE),
+         secondary_heat = fct_collapse(secondary_heat,
+                                       "Electric" = "Electrical",
+                                       "Wood stove" = "Wood stove",
+                                       "Other" = c("Natural gas furnace",
+                                                   "Oil", "Other", "Propane")),
+         chimney_clean = factor(chimney_clean,
+                         levels = c("Less than 6 months ago", "6-12 months ago",
+                                    "12-18 months ago", "More than 18 months ago")),
+         chimney_clean = fct_collapse(chimney_clean,
+                                  "Less than 6 months ago" = "Less than 6 months ago",
+                                  "6-12 months ago" = "6-12 months ago",
+                                  "12 + months ago" = c("12-18 months ago",
+                                                        "More than 18 months ago")),
+         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
+         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
+         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
+         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
+         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know")))
 
-pm_home_char <- pm_home_mean %>% 
-  left_join(home_char_clean, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(stove_age = factor(stove_age,
-                         levels = c("0-5 years old", "6-10 years old",
-                                    "11-15 years old", "16 + years old",
-                                    "Do not know")))
+table(pm_home_char$home_insert)
+median(pm_home_char$home_furry, na.rm = TRUE)
+
+pm_home_char_results <- lm(mean_pm ~ home_insert, data = pm_home_char)
+summary(pm_home_char_results)
+plot(pm_home_char_results)
 ```
 
-# Simple associations between PM and covariates
+# Stove use
 ```{r}
-pm_demographics_results <- lm(mean_pm ~ education, data = pm_demographics)
-summary(pm_demographics_results)
+pm_stove_use <- pm_home_mean %>% 
+  left_join(stove_use_clean, by = c("area", "home")) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(burn_level = fct_collapse(burn_level,
+                                   "No/Light burning" = c("No burning", 
+                                                          "Light burning"),
+                                   "Ave burning" = "Average burning",
+                                   "Heavy burning" = "Heavy burning"),
+         burn_level = factor(burn_level,
+                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
+         wood_collect = fct_collapse(wood_collect,
+                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
+                                                   "1 mo to 3 mo"),
+                                   "3+ months" = c("3 mo to 6 mo", 
+                                                   "6 mo to 1 year", "> 1 yr")),
+         wood_collect = factor(wood_collect,
+                         levels = c("<3 months", "3+ months")),
+         wood_collect_method = fct_collapse(wood_collect_method,
+                                   "Harvest yourself" = "Harvest yourself",
+                                   "Purchase/other" = c("Mix", "Other", 
+                                                        "Purchase it")),
+         wood_cords = if_else(wood_cords > 10, 10, wood_cords),
+         wood_cords_cat = cut(wood_cords, breaks = c(0, 5, 50),
+                                labels = c("<5", "5+"),
+                                 right = FALSE))
 
-pm_home_char_results <- lm(mean_pm ~ stove_age, data = pm_home_char)
-summary(pm_home_char_results)
+table(pm_stove_use$wood_cords_cat)
+median(pm_stove_use$wood_cords, na.rm = TRUE)
+
+pm_stove_use_results <- lm(mean_pm ~ wood_cords_cat, data = pm_stove_use)
+summary(pm_stove_use_results)
+plot(pm_stove_use_results)
+pm_stove_use_plot <- pm_stove_use %>% 
+  ggplot() + 
+    geom_boxplot(aes(burn_level, mean_pm))
+pm_stove_use_plot
 ```
+
+# Stove grades
+```{r}
+stove_grades <- stove_grades_clean %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(home) %>% 
+  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade))
+
+pm_stove_grades <- pm_home_mean %>% 
+  left_join(stove_grades, by = c("area", "home")) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(stove_grade = fct_collapse(stove_grade,
+                                    "A" = "A",
+                                    "C" = "C",
+                                    "D or F" = c("D", "F")))
+
+table(pm_stove_grades$stove_grade)
+
+pm_stove_grades_results <- lm(mean_pm ~ stove_grade, data = pm_stove_grades)
+summary(pm_stove_grades_results)
+anova(pm_stove_grades_results)
+plot(pm_stove_grades_results)
+pm_stove_grades_plot <- pm_stove_grades %>% 
+  ggplot() + 
+    geom_boxplot(aes(stove_grade, mean_pm))
+pm_stove_grades_plot
+```
+
+# Home activity
+```{r}
+# Join home activity with mean PM and format for analysis
+pm_home_act <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  arrange(area, home, sampling_day) %>% 
+  group_by(area, home, sampling_day) %>% 
+  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  mutate(home_act_day = as.factor(sampling_day)) %>% 
+  arrange(area, home, home_act_day) %>% 
+  left_join(home_act_clean, by = c("area", "home", "home_act_day")) %>% 
+  filter(winter_id == 1)
+
+table(pm_home_act$home_act_sweep)
+
+pm_home_act_results <- lmer(mean_pm ~ home_act_sweep + (1 | home), data = pm_home_act)
+summary(pm_home_act_results)
+plot(pm_home_act_results)
+```
+
+# Wood moisture content
+```{r}
+# Join moisture content with mean PM and format for analyzis
+pm_homes_placebo <- pm_clean %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
+  filter(initial_obs == 1) %>% 
+  mutate(pm_home = "true") %>% 
+  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
+  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+moisture_filtered <- moisture_clean %>% 
+  # change to correct joining data of interest
+  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
+  filter(pm_home == "true") %>% 
+  filter(winter_id == 1) %>% 
+  ungroup() %>%
+  arrange(area, home)
+
+moisture_filtered_time <- moisture_filtered %>% 
+  # fix wrong date
+  mutate(moisture_date = as.character(moisture_date),
+         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
+                                 "2017-04-10", moisture_date),
+         moisture_date = ymd(moisture_date)) %>% 
+  mutate(moisture_date = ymd(moisture_date),
+         first_date = ymd(first_date),
+         time_diff = moisture_date - first_date,
+         time_diff = as.numeric(abs(time_diff))) %>% 
+  group_by(home) %>% 
+  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
+  ungroup() %>% 
+  filter(time_diff_rank == 1)
+
+pm_moisture <- pm_home_mean %>% 
+  left_join(moisture_filtered_time, by = c("area", "home")) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(moisture_ave_5 = moisture_ave/5,
+         moisture_cat = cut(moisture_ave, breaks = c(0, 10, 50),
+                                labels = c("<10", "10+"),
+                                 right = FALSE))
+
+table(pm_moisture$moisture_cat)
+median(pm_moisture$moisture_ave, na.rm = TRUE)
+
+pm_moisture_results <- lm(mean_pm ~ moisture_ave_5, data = pm_moisture)
+summary(pm_moisture_results)
+plot(pm_moisture_results)
+```
+
+# Indoor temp/Rh
+```{r}
+pm_temp_rh <- pm_home_mean %>% 
+  left_join(indoor_temp_rh_clean, by = c("area", "home")) %>% 
+  filter(winter_id == 1) %>% 
+  mutate(temp_max_5 = temp_indoor_max/5,
+         temp_min_5 = temp_indoor_min/5,
+         rh_max_5 = rh_indoor_max/5,
+         rh_min_5 = rh_indoor_min/5,
+         temp_max_cat = cut(temp_indoor_max, breaks = c(0, 30, 45),
+                                labels = c("<30", "30+"),
+                                 right = FALSE),
+         temp_min_cat = cut(temp_indoor_min, breaks = c(0, 16.5, 50),
+                                labels = c("<16.5", "16.5+"),
+                                 right = FALSE),
+         rh_max_cat = cut(rh_indoor_max, breaks = c(0, 40, 100),
+                                labels = c("<40", "40+"),
+                                 right = FALSE),
+         rh_min_cat = cut(rh_indoor_min, breaks = c(0, 20, 100),
+                                labels = c("<20", "20+"),
+                                 right = FALSE))
+
+table(pm_temp_rh$rh_min_cat)
+median(pm_temp_rh$rh_indoor_min, na.rm = TRUE)
+range(pm_temp_rh$rh_indoor_max, na.rm = TRUE)
+
+pm_home_act_results <- lm(mean_pm ~ rh_min_cat, data = pm_temp_rh)
+summary(pm_home_act_results)
+plot(pm_home_act_results)
+```

### 2020-03-18 exposure_initial_analysis.Rmd
exposure analysis updates

@@
 
-pm_home_char_results <- lm(mean_pm ~ home_insert, data = pm_home_char)
+pm_home_char_results <- lm(log(mean_pm) ~ home_insert, data = pm_home_char)
 summary(pm_home_char_results)
@@
 
-pm_stove_use_results <- lm(mean_pm ~ wood_cords_cat, data = pm_stove_use)
+pm_stove_use_results <- lm(log(mean_pm) ~ wood_collect_method, data = pm_stove_use)
 summary(pm_stove_use_results)
@@
 
-pm_stove_grades_results <- lm(mean_pm ~ stove_grade, data = pm_stove_grades)
+pm_stove_grades_results <- lm(log(mean_pm) ~ stove_grade, data = pm_stove_grades)
 summary(pm_stove_grades_results)
@@
 
-pm_home_act_results <- lmer(mean_pm ~ home_act_sweep + (1 | home), data = pm_home_act)
+pm_home_act_results <- lmer(log(mean_pm) ~ home_act_sweep + (1 | home), data = pm_home_act)
 summary(pm_home_act_results)
-plot(pm_home_act_results)
+plot(pm_home_act_results, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
 ```
@@
 
-pm_moisture_results <- lm(mean_pm ~ moisture_ave_5, data = pm_moisture)
+pm_moisture_results <- lm(log(mean_pm) ~ moisture_cat, data = pm_moisture)
 summary(pm_moisture_results)
@@
 
-pm_home_act_results <- lm(mean_pm ~ rh_min_cat, data = pm_temp_rh)
+pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = pm_temp_rh)
 summary(pm_home_act_results)
@@
 ```
+
+# SUMs 
+```{r}
+#### Join PM dataset with full SUMs dataset
+
+pm_join <- read_rds("Output/pm_clean.rds") %>% 
+  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
+  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  group_by(area, home) %>% 
+  arrange(sample_datetime) %>% 
+  mutate(pm_last_datetime = last(sample_datetime),
+         pm_first_datetime = first_datetime_new,
+         sample_datetime = round_date(sample_datetime, unit = "minute"),
+         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime)
+
+sums_join <- read_rds("Output/sums_clean.rds") %>% 
+  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
+  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime)
+
+pm_sums_joined <- sums_join %>% 
+  left_join(pm_join, by = c("area", "home", "sample_datetime")) %>% 
+  filter(pm_rolling_20 < 1000) %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  filter(!is.na(pm_rolling_20)) %>% 
+  mutate(temp_c_sums_5 = temp_c_sums/5)
+
+pm_sums_count <- pm_sums_joined %>% 
+  count(home)
+
+cor.test(pm_sums_joined$temp_c_sums, pm_sums_joined$pm_rolling_20, method = "spearman")
+pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), data = pm_sums_joined)
+
+summary(pm_sums_model)
+exp((0.146291)-1)*100
+plot(pm_sums_model, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_sums_model), main = "QQ Plot")
+
+pm_sums_plot <- pm_sums_joined %>% 
+  ggplot() + 
+    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
+    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
+    theme_minimal()
+pm_sums_plot
+
+#### Join PM dataset with SUMs events datasets
+
+# Mean PM for each placebo home over winter 1
+pm_home_mean_date <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  separate(pm_datetime_new, c("date_sums", "pm_time"), sep = " ", remove = TRUE) %>% 
+  group_by(area, home, date_sums) %>% 
+  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
+  group_by(area, home, date_sums) %>% 
+  distinct(date_sums, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home, date_sums)
+
+pm_sums_events_joined <- pm_home_mean_date %>% 
+  left_join(sums_events_27, by = c("area", "home", "date_sums")) %>% 
+  filter(sums_event == 1) %>% 
+  arrange(area, home, date_sums) %>% 
+  group_by(area, home, mean_pm) %>% 
+  summarize(sums_events = n()) %>% 
+  #mutate(sums_events = factor(sums_events))
+  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 1, 10),
+                           labels = c("1", ">1")))
+
+median(pm_sums_events_joined$sums_events)
+
+pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = pm_sums_events_joined)
+pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = pm_sums_events_joined)
+
+summary(pm_sums_model)
+#exp((0.146291)-1)*100
+plot(pm_sums_model, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_sums_model), main = "QQ Plot")
+```

### 2020-03-18 sums_data_cleaning.Rmd
exposure analysis updates

@@
 
-thresholds_test <- sums_clean %>% 
+sums_events_25 <- sums_clean %>% 
   mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
   group_by(area, home, winter_id) %>%
-  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0),
+  mutate(threshold_temp = if_else(temp_c_sums >= 25, 1, 0),
          threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
@@
                                       lead(threshold_temp_event) == "Temp < threshold", 
-                                      "event", threshold_temp_event)) %>% 
-  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+                                      "event_end", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
+  mutate(time_sums = as.character(time_sums),
+         date_sums = as.character(date_sums)) %>%  
   spread(threshold_temp_event, datetime_sums) %>% 
-  mutate(event = lead(event),
-         time_diff_min = (event - start),
-         time_diff_min = as.numeric(time_diff_min/60)) %>% 
-  filter(!is.na(time_diff_min)) %>% 
-  mutate(next_event = (lead(start)-event),
-         next_event = as.numeric(next_event)/3600,
-         event2 = if_else(next_event <= 1.5, lead(event), event),
-         time_diff_min2 = (event - start),
-         time_diff_min2 = as.numeric(time_diff_min2/60),
-         event3 = if_else(next_event <= 1.5 & lead(next_event) <= 1.5, lead(event, 2), event2),
-         time_diff_min3 = (event3 - start),
-         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
-  distinct(event3, .keep_all = TRUE) %>% 
-  mutate(time_diff_min4 = (event3 - start),
-         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
-  filter(!is.na(time_diff_min4)) %>% 
+  mutate(event_end = lead(event_end)) %>% 
+  filter(!is.na(start)) %>% 
+  mutate(hrs_till_next_event = (lead(start)-event_end),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
+         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
+         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
+                              lead(event_end, 2), event_end2),
+         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5, 
+                              lead(event_end, 3), event_end3),
+         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
+                              lead(event_end, 4), event_end4),
+         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
+                              & lead(hrs_till_next_event, 4) <= 1.5, 
+                              lead(event_end, 5), event_end5)) %>% 
+  distinct(event_end6, .keep_all = TRUE) %>% 
+  mutate(event_length_min = (event_end6 - start),
+         event_length_min = as.numeric(event_length_min/60),
+         hrs_till_next_event = (lead(start)-event_end6),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
+  filter(!is.na(event_length_min)) %>% 
   group_by(area, home, winter_id, date_sums) %>% 
-  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
-  group_by(area) %>% 
-  summarize("Mean events per day" = round(mean(n), digits = 2),
-            "Mean event length (min)" = round(mean(mean_time), digits = 0))
-thresholds_test
-
-# Function for calculating heating events using specified threshold temps
-# Also specify the "gap" time between events
-# Try another function for % day above threshold
-thresholds_function <- function(data, temp, gap) {
-
-thresholds <- data %>% 
+  mutate(event_start_datetime = start,
+         event_end_datetime = event_end6,
+         sums_event = as.numeric(1)) %>% 
+  select(area, home, home_winter_id, winter_id, date_sums, 
+         event_start_datetime, event_end_datetime, sums_event,
+         event_length_min, hrs_till_next_event)
+
+sums_events_27 <- sums_clean %>% 
   mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
   group_by(area, home, winter_id) %>%
-  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
+  mutate(threshold_temp = if_else(temp_c_sums >= 27, 1, 0),
          threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
@@
                                       lead(threshold_temp_event) == "Temp < threshold", 
-                                      "event", threshold_temp_event)) %>% 
-  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
+                                      "event_end", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
+  mutate(time_sums = as.character(time_sums),
+         date_sums = as.character(date_sums)) %>%  
   spread(threshold_temp_event, datetime_sums) %>% 
-  mutate(event = lead(event),
-         time_diff_min = (event - start),
-         time_diff_min = as.numeric(time_diff_min/60)) %>% 
-  filter(!is.na(time_diff_min)) %>% 
-  mutate(next_event = (lead(start)-event),
-         next_event = as.numeric(next_event)/3600,
-         event2 = if_else(next_event <= gap, lead(event), event),
-         time_diff_min2 = (event - start),
-         time_diff_min2 = as.numeric(time_diff_min2/60),
-         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
-         time_diff_min3 = (event3 - start),
-         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
-  distinct(event3, .keep_all = TRUE) %>% 
-  mutate(time_diff_min4 = (event3 - start),
-         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
-  filter(!is.na(time_diff_min4)) %>% 
+  mutate(event_end = lead(event_end)) %>% 
+  filter(!is.na(start)) %>% 
+  mutate(hrs_till_next_event = (lead(start)-event_end),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
+         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
+         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
+                              lead(event_end, 2), event_end2),
+         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5, 
+                              lead(event_end, 3), event_end3),
+         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
+                              lead(event_end, 4), event_end4),
+         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
+                              & lead(hrs_till_next_event, 4) <= 1.5, 
+                              lead(event_end, 5), event_end5)) %>% 
+  distinct(event_end6, .keep_all = TRUE) %>% 
+  mutate(event_length_min = (event_end6 - start),
+         event_length_min = as.numeric(event_length_min/60),
+         hrs_till_next_event = (lead(start)-event_end6),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
+  filter(!is.na(event_length_min)) %>% 
   group_by(area, home, winter_id, date_sums) %>% 
-  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
-  group_by(area) %>% 
-  summarize("Mean events per day" = round(mean(n), digits = 2),
-            "Mean event length (min)" = round(mean(mean_time), digits = 0))
-thresholds
-
-}
-
-thresholds_function(sums_clean, 40, 2)
-
-# Function to look at percentage of time spent above a threshold
-thresholds_percent <- function(data, temp) {
-
-thresholds <- data %>% 
+  mutate(event_start_datetime = start,
+         event_end_datetime = event_end6,
+         sums_event = as.numeric(1)) %>% 
+  select(area, home, home_winter_id, winter_id, date_sums, 
+         event_start_datetime, event_end_datetime, sums_event,
+         event_length_min, hrs_till_next_event)
+
+sums_events_30 <- sums_clean %>% 
+  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
+  group_by(area, home, winter_id) %>%
+  mutate(threshold_temp = if_else(temp_c_sums >= 30, 1, 0),
+         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
+                                        "start", "Temp < threshold"),
+         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
+                                      "Temp >= threshold", threshold_temp_event),
+         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
+                                      lead(threshold_temp_event) == "Temp < threshold", 
+                                      "event_end", threshold_temp_event)) %>% 
+  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
+  mutate(time_sums = as.character(time_sums),
+         date_sums = as.character(date_sums)) %>%  
+  spread(threshold_temp_event, datetime_sums) %>% 
+  mutate(event_end = lead(event_end)) %>% 
+  filter(!is.na(start)) %>% 
+  mutate(hrs_till_next_event = (lead(start)-event_end),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
+         event_end2 = if_else(hrs_till_next_event <= 1.5, lead(event_end), event_end),
+         event_end3 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5, 
+                              lead(event_end, 2), event_end2),
+         event_end4 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5, 
+                              lead(event_end, 3), event_end3),
+         event_end5 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5, 
+                              lead(event_end, 4), event_end4),
+         event_end6 = if_else(hrs_till_next_event <= 1.5 & lead(hrs_till_next_event) <= 1.5
+                              & lead(hrs_till_next_event, 2) <= 1.5 & lead(hrs_till_next_event, 3) <= 1.5
+                              & lead(hrs_till_next_event, 4) <= 1.5, 
+                              lead(event_end, 5), event_end5)) %>% 
+  distinct(event_end6, .keep_all = TRUE) %>% 
+  mutate(event_length_min = (event_end6 - start),
+         event_length_min = as.numeric(event_length_min/60),
+         hrs_till_next_event = (lead(start)-event_end6),
+         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
+  filter(!is.na(event_length_min)) %>% 
   group_by(area, home, winter_id, date_sums) %>% 
-  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
-  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
-  group_by(area) %>% 
-  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
-thresholds
-
-}
-
-thresholds_percent(sums_clean, 40)
-```
+  mutate(event_start_datetime = start,
+         event_end_datetime = event_end6,
+         sums_event = as.numeric(1)) %>% 
+  select(area, home, home_winter_id, winter_id, date_sums, 
+         event_start_datetime, event_end_datetime, sums_event,
+         event_length_min, hrs_till_next_event)
+
+#write_rds(sums_events_25, "Output/sums_events_25.rds")
+#write_rds(sums_events_27, "Output/sums_events_27.rds")
+#write_rds(sums_events_30, "Output/sums_events_30.rds")
+```

### 2020-03-19 exposure_initial_analysis.Rmd
exposure analysis updates

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 17 March 2020"
+date: "Started 2 Feb 2020, Updated 19 March 2020"
 output: pdf_document
@@
 
-# SUMs 
+# SUMs - experimenting with several methods of looking at associations
 ```{r}
-#### Join PM dataset with full SUMs dataset
+#### Join PM dataset full SUMs dataset by SUMs sampling times
 
-pm_join <- read_rds("Output/pm_clean.rds") %>% 
+pm_join <- pm_clean %>% 
   mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
@@
 
-sums_join <- read_rds("Output/sums_clean.rds") %>% 
+sums_join <- sums_clean %>% 
   mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
@@
 
+#### Join PM dataset with SUMs dataset by sampling day
+
+pm_join <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  mutate(pm = if_else(pm > 500, 500, pm),
+         pm_mean = mean(pm, na.rm = TRUE)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm, na.rm = TRUE)) %>% 
+  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  arrange(sample_datetime) %>% 
+  mutate(pm_last_datetime = last(sample_datetime),
+         pm_first_datetime = first(sample_datetime)) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime) %>% 
+  select(area, home, home_winter_id, dtid, day_of_week, pm_total_observations,
+         pm_mean, sampling_day, pm_mean_daily, pm_first_datetime, pm_last_datetime) %>% 
+  group_by(area, home) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home, sampling_day)
+
+sums_join <- sums_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, winter_id, ibid, treatment, datetime_sums, temp_c_sums,)
+
+pm_sums_joined <- sums_join %>% 
+  left_join(pm_join, by = c("area", "home")) %>% 
+  arrange(area, home, datetime_sums) %>% 
+  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
+                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
+  filter(filter_var == 1) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(temp_perc_25 = if_else(temp_c_sums >= 30, 1, 0)) %>% 
+  arrange(area, home, sampling_day) %>% 
+  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
+         sums_mean_daily_5 = sums_mean_daily/5) %>% 
+  group_by(area, home, home_winter_id, sampling_day, pm_mean, 
+           pm_mean_daily, sums_mean_daily) %>% 
+  summarize(temp_perc_25 = (sum(temp_perc_25)/n()*100)) %>% 
+  mutate(temp_perc_25_cat = cut(temp_perc_25, breaks = c(0, 10, 200),
+                                labels = c("< 50%", "50% +"),
+                                 right = FALSE)) %>%
+  distinct(pm_mean_daily, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home, sampling_day)
+
+pm_sums_count <- pm_sums_joined %>% 
+  count(home)
+median(pm_sums_joined$temp_perc_25)
+table(pm_sums_joined$temp_perc_25_cat)
+
+cor.test(pm_sums_joined$sums_mean_daily, pm_sums_joined$pm_mean_daily, method = "spearman")
+pm_sums_model <- lmer(pm_mean_daily ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)
+pm_sums_model <- lmer(log(pm_mean_daily) ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)
+
+summary(pm_sums_model)
+exp((0.146291)-1)*100
+plot(pm_sums_model, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_sums_model), main = "QQ Plot")
+
 #### Join PM dataset with SUMs events datasets
@@
 
+sums_temp_change <- sums_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home, date_sums) %>% 
+  mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
+  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
+         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
+  replace_with_na(replace = list(lead_temp = 999)) %>% 
+  mutate(temp_diff = lead_temp - temp_c_sums,
+         temp_diff_abs = abs(lead_temp - temp_c_sums),
+         # change temp in the following line
+         temp_diff_check = if_else(temp_diff >= 8, 1, 0),
+         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
+                              temp_diff_check),
+         heat_event = if_else(is.na(temp_diff_check), 1,
+                              heat_event)) %>% 
+  ungroup() %>% 
+  mutate(date_sums = as.character(date_sums)) %>% 
+  group_by(area, home, date_sums) %>% 
+  summarize(sums_events = sum(heat_event, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, date_sums)
+
 pm_sums_events_joined <- pm_home_mean_date %>% 
-  left_join(sums_events_27, by = c("area", "home", "date_sums")) %>% 
-  filter(sums_event == 1) %>% 
+  left_join(sums_temp_change, by = c("area", "home", "date_sums")) %>% 
   arrange(area, home, date_sums) %>% 
-  group_by(area, home, mean_pm) %>% 
-  summarize(sums_events = n()) %>% 
   #mutate(sums_events = factor(sums_events))
-  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 1, 10),
-                           labels = c("1", ">1")))
+  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 2, 20),
+                           labels = c("<2", "2+"),
+                                 right = FALSE))
 
-median(pm_sums_events_joined$sums_events)
+median(pm_sums_events_joined$sums_events, na.rm = TRUE)
+table(pm_sums_events_joined$sums_events_cat)
 
@@
 qqnorm(residuals(pm_sums_model), main = "QQ Plot")
+
+pm_sums_plot <- pm_sums_events_joined %>% 
+  filter(!is.na(sums_events)) %>% 
+  ggplot() + 
+    geom_boxplot(aes(sums_events_cat, log(mean_pm))) +
+    theme_minimal()
+pm_sums_plot
+
 ```
+
+# Create new dataset with mean PM/SUMs by PM sampling day; join with covariates
+```{r}
+# Prep PM data to join with SUMs data by sampling day
+pm_join <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm),
+         pm_mean = mean(pm, na.rm = TRUE)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(pm_mean_daily = mean(pm, na.rm = TRUE)) %>% 
+  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  arrange(sample_datetime) %>% 
+  mutate(pm_last_datetime = last(sample_datetime),
+         pm_first_datetime = first(sample_datetime)) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime) %>% 
+  select(area, home, home_winter_id, dtid, day_of_week, pm_total_observations,
+         pm_mean, sampling_day, pm_mean_daily, pm_first_datetime, pm_last_datetime) %>% 
+  group_by(area, home) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home, sampling_day)
+
+# Prep SUMs data to join with PM data
+sums_join <- sums_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, winter_id, ibid, treatment, datetime_sums, temp_c_sums,)
+
+# Join PM/SUMs data and filter and arrange by sampling day
+pm_sums_joined <- sums_join %>% 
+  left_join(pm_join, by = c("area", "home")) %>% 
+  arrange(area, home, datetime_sums) %>% 
+  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
+                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
+  filter(filter_var == 1) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0)) %>% 
+  arrange(area, home, sampling_day) %>% 
+  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
+         sums_mean_daily_5 = sums_mean_daily/5) %>% 
+  group_by(area, home, home_winter_id, sampling_day, 
+           sums_mean_daily, sums_mean_daily_5) %>% 
+  summarize(temp_perc_30 = (sum(temp_perc_30)/n()*100)) %>% 
+  mutate(temp_perc_30_cat = cut(temp_perc_30, breaks = c(0, 10, 200),
+                                labels = c("< 10%", "10% +"),
+                                 right = FALSE)) %>%
+  distinct(pm_mean_daily, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home, sampling_day)
+
+# Join SUMs data with original PM data, by sampling day
+pm_sums_joined2 <- pm_sums_joined %>% 
+  right_join(pm_join, by = c("area", "home", "home_winter_id", "sampling_day")) %>% 
+  mutate(sampling_day = as.factor(sampling_day)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+# Join PM/SUMs data with other covariate datasets
+analysis_data1 <- home_act_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  mutate(sampling_day = home_act_day) %>% 
+  select(-home_act_day, -home_act_comments, -home_id) %>% 
+  left_join(pm_sums_joined2, by = c("area", "home", "home_winter_id", "sampling_day")) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+analysis_data2 <- demographics_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, gender_parent:residents_under_five) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  left_join(analysis_data1, by = c("area", "home", "home_winter_id")) %>% 
+  filter(!is.na(gender_parent)) %>% 
+  mutate(income_3level = fct_collapse(income,
+                               "Less than $20,000" = "Less than $20,000",
+                               "$20,000-$49,999" = 
+                                 c("$20,000 to $29,999", "$30,000 to $39,999",
+                                   "$40,000 to $49,999"),
+                                "$50,000+" = 
+                                 c("$50,000 to $74,999", "$75,000 to $99,999",
+                                   "$100,000 or more")),
+         income_3level = factor(income_3level, levels = c("Less than $20,000", 
+                                            "$20,000-$49,999", 
+                                            "$50,000+")), 
+         education_3level = fct_collapse(education,
+                                  "High school or less" = 
+                                    c("Less than high school", 
+                                      "High school diploma or GED"),
+                                  "Some college" = "Some college",
+                                  "College degree" = "College degree"),
+         education_3level = factor(education_3level, levels = c("High school or less", 
+                                            "Some college", 
+                                            "College degree")), 
+         race_parent_2level = fct_collapse(race_parent,
+                                    "White" = "White",
+                                    "Non-white" = 
+                                      c("American Indian/Alaskan Native",
+                                        "Asian", "More than one race")),
+         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
+         total_residents_2level = fct_collapse(total_residents,
+                                    "<5" = c("1", "2", "3", "4"),
+                                    "5+" = c("5", "6", "7", "8", "9", "10",
+                                             "11", "12", "13", "14", "15")),
+         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
+         residents_under5_2level = fct_collapse(residents_under_five,
+                                    "1" = c("0", "1"),
+                                    "2+" = c("2", "3", "4", "5")),
+         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+analysis_data3 <- home_char_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, residents_smoke:home_insert) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data2, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(home_type_2level = fct_collapse(home_type,
+                                  "House" = "House",
+                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
+         home_type_2level = factor(home_type_2level,
+                            levels = c("House", "Other")),
+         home_year_built_2level = cut(home_year_built, breaks = c(0, 1985, 2020),
+                                 labels = c("<1985", "1985+"),
+                                 right = FALSE),
+         stove_age = factor(stove_age,
+                         levels = c("0-5 years old", "6-10 years old",
+                                    "11-15 years old", "16 + years old",
+                                    "Do not know")),
+         stove_age_4level = fct_collapse(stove_age,
+                                  "0-5 years old" = "0-5 years old",
+                                  "6-15 years old" = c("6-10 years old",
+                                                       "11-15 years old"),
+                                  "16 + years old" = "16 + years old",
+                                  "Do not know" = "Do not know"),
+         stove_cert = factor(stove_cert,
+                             levels = c("Yes", "No", "Do not know")),
+         home_sqft_2level = cut(home_sqft, breaks = c(0, 1400, 5000),
+                         labels = c("<1400", "1400+"),
+                         right = FALSE),
+         home_floors_2level = cut(home_floors, breaks = c(0, 2, 10),
+                                 labels = c("1", "2+"),
+                                 right = FALSE),
+         home_bedrooms_2level = cut(home_bedrooms, breaks = c(0, 3, 50),
+                                 labels = c("<3", "3+"),
+                                 right = FALSE),
+         home_windows_2level = cut(home_windows, breaks = c(0, 10, 100),
+                                labels = c("<10", "10+"),
+                                 right = FALSE),
+         home_pets_2level = cut(home_pets, breaks = c(0, 2, 50),
+                                labels = c("<2", "2+"),
+                                 right = FALSE),
+         home_furry_2level = cut(home_furry, breaks = c(0, 2, 50),
+                                labels = c("<2", "2+"),
+                                 right = FALSE),
+         secondary_heat_3level = fct_collapse(secondary_heat,
+                                       "Electric" = "Electrical",
+                                       "Wood stove" = "Wood stove",
+                                       "Other" = c("Natural gas furnace",
+                                                   "Oil", "Other", "Propane")),
+         chimney_clean = factor(chimney_clean,
+                         levels = c("Less than 6 months ago", "6-12 months ago",
+                                    "12-18 months ago", "More than 18 months ago")),
+         chimney_clean_3level = fct_collapse(chimney_clean,
+                                  "Less than 6 months ago" = "Less than 6 months ago",
+                                  "6-12 months ago" = "6-12 months ago",
+                                  "12 + months ago" = c("12-18 months ago",
+                                                        "More than 18 months ago")),
+         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
+         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
+         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
+         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
+         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+analysis_data4 <- indoor_temp_rh_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, rh_indoor_current:temp_indoor_min) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data3, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(temp_max_5 = temp_indoor_max/5,
+         temp_min_5 = temp_indoor_min/5,
+         rh_max_5 = rh_indoor_max/5,
+         rh_min_5 = rh_indoor_min/5,
+         temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
+                                labels = c("<30", "30+"),
+                                 right = FALSE),
+         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
+                                labels = c("<16.5", "16.5+"),
+                                 right = FALSE),
+         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 200),
+                                labels = c("<40", "40+"),
+                                 right = FALSE),
+         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 200),
+                                labels = c("<20", "20+"),
+                                 right = FALSE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+analysis_data5 <- moisture_filtered_time %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, moisture_date, moisture_ave, moisture_split) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data4, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(moisture_ave_5 = moisture_ave/5,
+         moisture_2level = cut(moisture_ave, breaks = c(0, 10, 100),
+                                labels = c("<10", "10+"),
+                                 right = FALSE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+analysis_data6 <- stove_grades_clean %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(home) %>% 
+  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, stove_grade) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data5, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(stove_grade_3level = fct_collapse(stove_grade,
+                                    "A" = "A",
+                                    "C" = "C",
+                                    "D or F" = c("D", "F"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+analysis_data7 <- stove_use_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, burn_level:wood_cords) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data6, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(burn_level_3level = fct_collapse(burn_level,
+                                   "No/Light burning" = c("No burning", 
+                                                          "Light burning"),
+                                   "Ave burning" = "Average burning",
+                                   "Heavy burning" = "Heavy burning"),
+         burn_level_3level = factor(burn_level_3level,
+                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
+         wood_collect_2level = fct_collapse(wood_collect,
+                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
+                                                   "1 mo to 3 mo"),
+                                   "3+ months" = c("3 mo to 6 mo", 
+                                                   "6 mo to 1 year", "> 1 yr")),
+         wood_collect_2level = factor(wood_collect_2level,
+                         levels = c("<3 months", "3+ months")),
+         wood_collect_method_2level = fct_collapse(wood_collect_method,
+                                   "Harvest yourself" = "Harvest yourself",
+                                   "Purchase/other" = c("Mix", "Other", 
+                                                        "Purchase it")),
+         wood_cords = if_else(wood_cords > 10, 10, wood_cords),
+         wood_cords_2level = cut(wood_cords, breaks = c(0, 5, 50),
+                                labels = c("<5", "5+"),
+                                 right = FALSE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sampling_day)
+
+#write_rds(analysis_data7, "Output/exposure_analysis_data.rds")
+#write_csv(analysis_data7, "Output/exposure_analysis_data.csv")
+```

### 2020-03-19 exposure_models.Rmd
exposure analysis updates

---
title: "KidsAIR: exposure model"
author: "Ethan Walker"
date: "Started 19 March 2020, Updated 19 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(emmeans)
library(broom)
library(MuMIn)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
exposure_analysis_data <- read_rds("Output/exposure_analysis_data.rds")
```

```{r}
model_results <- lmer(log(pm_mean_daily) ~ education_3level + 
                        area + sampling_day + day_of_week +
                        (1 | home), data = exposure_analysis_data)

summary(model_results)
r.squaredGLMM(model_results)

tidy_results <- tidy(model_results)
tidy_results

plot(model_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(model_results), main = "QQ Plot")
```

### 2020-03-20 exposure_initial_analysis.Rmd
new exposure analysis best subsets

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 19 March 2020"
+date: "Started 2 Feb 2020, Updated 20 March 2020"
 output: pdf_document
@@
 # Create new dataset with mean PM/SUMs by PM sampling day; join with covariates
+## Short data averaged over sampling period
+```{r}
+# Prep PM data to join with SUMs data by sampling day
+pm_join <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm),
+         pm_mean = mean(pm, na.rm = TRUE)) %>% 
+  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
+  group_by(area, home) %>% 
+  arrange(sample_datetime) %>% 
+  mutate(pm_last_datetime = last(sample_datetime),
+         pm_first_datetime = first(sample_datetime)) %>% 
+  ungroup() %>% 
+  arrange(area, home) %>% 
+  separate(pm_first_datetime, c("pm_date_start", "pm_time"), sep = " ", remove = FALSE) %>% 
+  select(area, home, home_winter_id, dtid, pm_total_observations,
+         pm_mean, pm_first_datetime, pm_last_datetime, pm_date_start) %>% 
+  group_by(area, home) %>% 
+  distinct(pm_mean, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+# Prep SUMs data to join with PM data
+sums_join <- sums_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, winter_id, ibid, treatment, datetime_sums, temp_c_sums,)
+
+# Join PM/SUMs data and filter and arrange by sampling day
+pm_sums_joined <- sums_join %>% 
+  left_join(pm_join, by = c("area", "home")) %>% 
+  arrange(area, home, datetime_sums) %>% 
+  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
+                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
+  filter(filter_var == 1) %>% 
+  group_by(area, home) %>% 
+  mutate(temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0)) %>% 
+  arrange(area, home) %>% 
+  mutate(sums_mean = mean(temp_c_sums, na.rm = TRUE),
+         sums_mean_5 = sums_mean/5) %>% 
+  group_by(area, home, home_winter_id, sums_mean, sums_mean_5) %>% 
+  summarize(temp_perc_30 = (sum(temp_perc_30)/n()*100)) %>% 
+  mutate(temp_perc_30_cat = cut(temp_perc_30, breaks = c(0, 15, 200),
+                                labels = c("< 15%", "15% +"),
+                                 right = FALSE)) %>%
+  ungroup() %>% 
+  arrange(area, home)
+
+# Join SUMs data with original PM data, by sampling day
+pm_sums_joined2 <- pm_sums_joined %>% 
+  right_join(pm_join, by = c("area", "home", "home_winter_id")) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+# Join PM/SUMs data with other covariate datasets
+analysis_data1 <- home_act_clean %>% 
+  arrange(area, home) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(-home_act_day, -home_act_comments, -home_id, -home_act_day) %>% 
+  group_by(area, home, home_winter_id, treatment, winter_id) %>% 
+  mutate_if(is.factor, factor, levels = c("Yes", "No"), labels = c(1, 0)) %>% 
+  mutate_if(is.factor, as.character) %>% 
+  mutate_if(is.character, as.numeric) %>% 
+  summarize_if(is.numeric, sum, na.rm = TRUE) %>% 
+  right_join(pm_sums_joined2, by = c("area", "home", "home_winter_id")) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+analysis_data2 <- demographics_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, gender_parent:residents_under_five) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data1, by = c("area", "home", "home_winter_id")) %>% 
+  filter(!is.na(gender_parent)) %>% 
+  mutate(income_3level = fct_collapse(income,
+                               "Less than $20,000" = "Less than $20,000",
+                               "$20,000-$49,999" = 
+                                 c("$20,000 to $29,999", "$30,000 to $39,999",
+                                   "$40,000 to $49,999"),
+                                "$50,000+" = 
+                                 c("$50,000 to $74,999", "$75,000 to $99,999",
+                                   "$100,000 or more")),
+         income_3level = factor(income_3level, levels = c("Less than $20,000", 
+                                            "$20,000-$49,999", 
+                                            "$50,000+")), 
+         education_3level = fct_collapse(education,
+                                  "High school or less" = 
+                                    c("Less than high school", 
+                                      "High school diploma or GED"),
+                                  "Some college" = "Some college",
+                                  "College degree" = "College degree"),
+         education_3level = factor(education_3level, levels = c("High school or less", 
+                                            "Some college", 
+                                            "College degree")), 
+         race_parent_2level = fct_collapse(race_parent,
+                                    "White" = "White",
+                                    "Non-white" = 
+                                      c("American Indian/Alaskan Native",
+                                        "Asian", "More than one race")),
+         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
+         total_residents_2level = fct_collapse(total_residents,
+                                    "<5" = c("1", "2", "3", "4"),
+                                    "5+" = c("5", "6", "7", "8", "9", "10",
+                                             "11", "12", "13", "14", "15")),
+         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
+         residents_under5_2level = fct_collapse(residents_under_five,
+                                    "1" = c("0", "1"),
+                                    "2+" = c("2", "3", "4", "5")),
+         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+analysis_data3 <- home_char_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, residents_smoke:home_insert) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data2, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(home_type_2level = fct_collapse(home_type,
+                                  "House" = "House",
+                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
+         home_type_2level = factor(home_type_2level,
+                            levels = c("House", "Other")),
+         home_year_built_2level = cut(home_year_built, breaks = c(0, 1985, 2020),
+                                 labels = c("<1985", "1985+"),
+                                 right = FALSE),
+         stove_age = factor(stove_age,
+                         levels = c("0-5 years old", "6-10 years old",
+                                    "11-15 years old", "16 + years old",
+                                    "Do not know")),
+         stove_age_3level = fct_collapse(stove_age,
+                                  "0-5 years old" = "0-5 years old",
+                                  "6-15 years old" = c("6-10 years old",
+                                                       "11-15 years old"),
+                                  "16 + years old" = "16 + years old",
+                                  "Do not know" = "Do not know"),
+         stove_cert = factor(stove_cert,
+                             levels = c("Yes", "No", "Do not know")),
+         home_sqft_2level = cut(home_sqft, breaks = c(0, 1400, 5000),
+                         labels = c("<1400", "1400+"),
+                         right = FALSE),
+         home_floors_2level = cut(home_floors, breaks = c(0, 2, 10),
+                                 labels = c("1", "2+"),
+                                 right = FALSE),
+         home_bedrooms_2level = cut(home_bedrooms, breaks = c(0, 3, 50),
+                                 labels = c("<3", "3+"),
+                                 right = FALSE),
+         home_windows_2level = cut(home_windows, breaks = c(0, 10, 100),
+                                labels = c("<10", "10+"),
+                                 right = FALSE),
+         home_pets_2level = cut(home_pets, breaks = c(0, 2, 50),
+                                labels = c("<2", "2+"),
+                                 right = FALSE),
+         home_furry_2level = cut(home_furry, breaks = c(0, 2, 50),
+                                labels = c("<2", "2+"),
+                                 right = FALSE),
+         secondary_heat_3level = fct_collapse(secondary_heat,
+                                       "Electric" = "Electrical",
+                                       "Wood stove" = "Wood stove",
+                                       "Other" = c("Natural gas furnace",
+                                                   "Oil", "Other", "Propane")),
+         chimney_clean = factor(chimney_clean,
+                         levels = c("Less than 6 months ago", "6-12 months ago",
+                                    "12-18 months ago", "More than 18 months ago")),
+         chimney_clean_3level = fct_collapse(chimney_clean,
+                                  "Less than 6 months ago" = "Less than 6 months ago",
+                                  "6-12 months ago" = "6-12 months ago",
+                                  "12 + months ago" = c("12-18 months ago",
+                                                        "More than 18 months ago")),
+         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
+         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
+         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
+         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
+         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+analysis_data4 <- indoor_temp_rh_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, rh_indoor_current:temp_indoor_min) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data3, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(temp_max_5 = temp_indoor_max/5,
+         temp_min_5 = temp_indoor_min/5,
+         rh_max_5 = rh_indoor_max/5,
+         rh_min_5 = rh_indoor_min/5,
+         temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
+                                labels = c("<30", "30+"),
+                                 right = FALSE),
+         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
+                                labels = c("<16.5", "16.5+"),
+                                 right = FALSE),
+         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 200),
+                                labels = c("<40", "40+"),
+                                 right = FALSE),
+         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 200),
+                                labels = c("<20", "20+"),
+                                 right = FALSE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+analysis_data5 <- moisture_filtered_time %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, moisture_date, moisture_ave, moisture_split) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data4, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(moisture_ave_5 = moisture_ave/5,
+         moisture_2level = cut(moisture_ave, breaks = c(0, 10, 100),
+                                labels = c("<10", "10+"),
+                                 right = FALSE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+analysis_data6 <- stove_grades_clean %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(home) %>% 
+  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, stove_grade) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data5, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(stove_grade_3level = fct_collapse(stove_grade,
+                                    "A" = "A",
+                                    "C" = "C",
+                                    "D or F" = c("D", "F"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+analysis_data7 <- stove_use_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  select(area, home, home_winter_id, burn_level:wood_cords) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  right_join(analysis_data6, by = c("area", "home", "home_winter_id")) %>% 
+  mutate(burn_level_3level = fct_collapse(burn_level,
+                                   "No/Light burning" = c("No burning", 
+                                                          "Light burning"),
+                                   "Ave burning" = "Average burning",
+                                   "Heavy burning" = "Heavy burning"),
+         burn_level_3level = factor(burn_level_3level,
+                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
+         wood_collect_2level = fct_collapse(wood_collect,
+                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
+                                                   "1 mo to 3 mo"),
+                                   "3+ months" = c("3 mo to 6 mo", 
+                                                   "6 mo to 1 year", "> 1 yr")),
+         wood_collect_2level = factor(wood_collect_2level,
+                         levels = c("<3 months", "3+ months")),
+         wood_collect_method_2level = fct_collapse(wood_collect_method,
+                                   "Harvest yourself" = "Harvest yourself",
+                                   "Purchase/other" = c("Mix", "Other", 
+                                                        "Purchase it")),
+         wood_cords = if_else(wood_cords > 10, 10, wood_cords),
+         wood_cords_2level = cut(wood_cords, breaks = c(0, 5, 50),
+                                labels = c("<5", "5+"),
+                                 right = FALSE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id)
+
+#write_rds(analysis_data7, "Output/exposure_analysis_data_short.rds")
+#write_csv(analysis_data7, "Output/exposure_analysis_data_short.csv")
+```
+
+# Create new dataset with mean PM/SUMs by PM sampling day; join with covariates
+## Long/daily data
 ```{r}
@@
                                     "Do not know")),
-         stove_age_4level = fct_collapse(stove_age,
+         stove_age_3level = fct_collapse(stove_age,
                                   "0-5 years old" = "0-5 years old",
@@
 
-#write_rds(analysis_data7, "Output/exposure_analysis_data.rds")
-#write_csv(analysis_data7, "Output/exposure_analysis_data.csv")
+#write_rds(analysis_data7, "Output/exposure_analysis_data_long.rds")
+#write_csv(analysis_data7, "Output/exposure_analysis_data_long.csv")
 ```

### 2020-03-20 exposure_models.Rmd
new exposure analysis best subsets

@@
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 19 March 2020"
+date: "Started 19 March 2020, Updated 20 March 2020"
 output: pdf_document
@@
 library(MuMIn)
+library(leaps)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 ```{r}
-exposure_analysis_data <- read_rds("Output/exposure_analysis_data.rds")
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds") %>% 
+  filter(!is.na(pm_mean_daily))
+
+exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") %>% 
+  filter(!is.na(pm_mean))
 ```
 
+# Linear model
 ```{r}
-model_results <- lmer(log(pm_mean_daily) ~ education_3level + 
-                        area + sampling_day + day_of_week +
-                        (1 | home), data = exposure_analysis_data)
+subset_model <- lm(log(pm_mean) ~ 
+                        income_3level + home_act_smoking + 
+                        area, data = exposure_analysis_data_short)
 
-summary(model_results)
-r.squaredGLMM(model_results)
+full_model <- lm(log(pm_mean) ~ 
+                        education_3level + income_3level + 
+                        total_residents_2level + home_floors_2level + home_windows_2level + 
+                        stove_age_3level + chimney_clean_3level + 
+                        wood_collect_2level + 
+                        stove_grade_3level + 
+                        home_act_smoking + 
+                        moisture_2level + 
+                        temp_perc_30_cat +
+                        area, data = exposure_analysis_data_short)
 
-tidy_results <- tidy(model_results)
+summary(subset_model)
+r.squaredGLMM(subset_model)
+
+tidy_results <- tidy(full_model)
 tidy_results
 
-plot(model_results, 
+plot(full_model, 
      xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(model_results), main = "QQ Plot")
+qqnorm(residuals(full_model), main = "QQ Plot")
 ```
+
+# Best subsets selection using dredge() from MuMIn
+```{r}
+options(na.action = "na.fail")
+
+all_subsets <- dredge(full_model, rank = "AIC", extra = c("R^2"))
+```
+
+# Linear mixed model with repeated measures
+```{r}
+subset_model <- lmer(log(pm_mean_daily) ~ 
+                        education_3level + income_3level + race_parent_2level + 
+                        area + sampling_day + day_of_week +
+                        (1 | home), data = exposure_analysis_data_long)
+
+full_model <- lmer(log(pm_mean_daily) ~ 
+                        education_3level + income_3level + race_parent_2level + 
+                        total_residents_2level + residents_smoke + 
+                        home_type_2level + home_floors_2level + home_windows_2level + 
+                        stove_age_3level + chimney_clean_3level + home_mold + 
+                        home_fireplace + 
+                        wood_collect_2level + 
+                        stove_grade_3level + 
+                        home_act_gas + home_act_propane + home_act_smoking + 
+                        moisture_2level + 
+                        temp_max_5 + 
+                        temp_perc_30_cat +
+                        area + sampling_day + day_of_week +
+                        (1 | home), data = exposure_analysis_data_long)
+
+summary(full_model)
+r.squaredGLMM(full_model)
+
+tidy_results <- tidy(full_model)
+tidy_results
+
+plot(full_model, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(full_model), main = "QQ Plot")
+```

### 2020-03-24 exposure_initial_analysis.Rmd
merge datasets

--- 2020-03-18 exposure_initial_analysis.Rmd
@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 17 March 2020"
+date: "Started 2 Feb 2020, Updated 20 March 2020"
 output: pdf_document
@@
   summarize(mean_pm = mean(pm, na.rm = TRUE))
+
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+
+exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") 
 ```
@@
 ```{r}
-pm_demographics <- pm_home_mean %>% 
-  left_join(demographics_clean, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(income = fct_collapse(income,
-                               "Less than $20,000" = "Less than $20,000",
-                               "$20,000-$49,999" = 
-                                 c("$20,000 to $29,999", "$30,000 to $39,999",
-                                   "$40,000 to $49,999"),
-                                "$50,000+" = 
-                                 c("$50,000 to $74,999", "$75,000 to $99,999",
-                                   "$100,000 or more")),
-         income = factor(income, levels = c("Less than $20,000", 
-                                            "$20,000-$49,999", 
-                                            "$50,000+"))) %>% 
-  mutate(education = fct_collapse(education,
-                                  "High school or less" = 
-                                    c("Less than high school", 
-                                      "High school diploma or GED"),
-                                  "Some college" = "Some college",
-                                  "College degree" = "College degree"),
-         education = factor(education, levels = c("High school or less", 
-                                            "Some college", 
-                                            "College degree"))) %>% 
-  mutate(race_parent = fct_collapse(race_parent,
-                                    "White" = "White",
-                                    "Non-white" = 
-                                      c("American Indian/Alaskan Native",
-                                        "Asian", "More than one race")),
-         race_parent = factor(race_parent, levels = c("White", "Non-white"))) %>% 
-  mutate(race_child = fct_collapse(race_child,
-                                    "White" = "White",
-                                    "Non-white" = 
-                                      c("American Indian/Alaskan Native",
-                                        "More than one race")),
-         race_child = factor(race_child, levels = c("White", "Non-white"))) %>%
-  mutate(total_residents = fct_collapse(total_residents,
-                                    "<5" = c("1", "2", "3", "4"),
-                                    "5+" = c("5", "6", "7", "8", "9", "10",
-                                             "11", "12", "13", "14", "15")),
-         total_residents = factor(total_residents, levels = c("<5", "5+"))) %>%
-  mutate(residents_under_five = fct_collapse(residents_under_five,
-                                    "1" = c("0", "1"),
-                                    "2+" = c("2", "3", "4", "5")),
-         residents_under_five = factor(residents_under_five, levels = c("1", "2+"))) %>% 
-  distinct(mean_pm, .keep_all = TRUE)
+filter_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean))
 
-table(pm_demographics$income)
+# Variables to use:
+## education_3level, income_3level, race_parent_2level, race_child_2level
+## total_residents_2level, residents_under5_2level, resident_smokes
 
-pm_demographics_results <- lm(mean_pm ~ residents_under_five, data = pm_demographics)
+table(filter_data$residents_smoke)
+median(filter_data$home_furry, na.rm = TRUE)
+
+pm_demographics_results <- lm(log(pm_mean) ~ residents_smoke, data = filter_data)
 summary(pm_demographics_results)
@@
 ```{r}
-pm_home_char <- pm_home_mean %>% 
-  left_join(home_char_clean, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(home_type = fct_collapse(home_type,
-                                  "House" = "House",
-                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
-         home_type = factor(home_type,
-                            levels = c("House", "Other")),
-         home_year_built_cat = cut(home_year_built, breaks = c(0, 1985, 2020),
-                                 labels = c("<1985", "1985+"),
-                                 right = FALSE),
-         stove_age = factor(stove_age,
-                         levels = c("0-5 years old", "6-10 years old",
-                                    "11-15 years old", "16 + years old",
-                                    "Do not know")),
-         stove_age = fct_collapse(stove_age,
-                                  "0-5 years old" = "0-5 years old",
-                                  "6-15 years old" = c("6-10 years old",
-                                                       "11-15 years old"),
-                                  "16 + years old" = "16 + years old",
-                                  "Do not know" = "Do not know"),
-         stove_cert = factor(stove_cert,
-                             levels = c("Yes", "No", "Do not know")),
-         home_sqft_cat = cut(home_sqft, breaks = c(0, 1400, 5000),
-                         labels = c("<1400", "1400+"),
-                         right = FALSE),
-         home_floors_cat = cut(home_floors, breaks = c(0, 2, 4),
-                                 labels = c("1", "2+"),
-                                 right = FALSE),
-         home_bedrooms_cat = cut(home_bedrooms, breaks = c(0, 3, 10),
-                                 labels = c("<3", "3+"),
-                                 right = FALSE),
-         home_windows_cat = cut(home_windows, breaks = c(0, 10, 50),
-                                labels = c("<10", "10+"),
-                                 right = FALSE),
-         home_pets_cat = cut(home_pets, breaks = c(0, 2, 50),
-                                labels = c("<2", "2+"),
-                                 right = FALSE),
-         home_furry_cat = cut(home_furry, breaks = c(0, 2, 50),
-                                labels = c("<2", "2+"),
-                                 right = FALSE),
-         secondary_heat = fct_collapse(secondary_heat,
-                                       "Electric" = "Electrical",
-                                       "Wood stove" = "Wood stove",
-                                       "Other" = c("Natural gas furnace",
-                                                   "Oil", "Other", "Propane")),
-         chimney_clean = factor(chimney_clean,
-                         levels = c("Less than 6 months ago", "6-12 months ago",
-                                    "12-18 months ago", "More than 18 months ago")),
-         chimney_clean = fct_collapse(chimney_clean,
-                                  "Less than 6 months ago" = "Less than 6 months ago",
-                                  "6-12 months ago" = "6-12 months ago",
-                                  "12 + months ago" = c("12-18 months ago",
-                                                        "More than 18 months ago")),
-         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
-         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
-         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
-         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
-         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know")))
+filter_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean))
 
-table(pm_home_char$home_insert)
-median(pm_home_char$home_furry, na.rm = TRUE)
+# Variables to use:
+## home_type_2level, home_year_built_2level, home_sqft_2level, home_floors_2level
+## home_bedrooms_2level, home_windows_2level, home_pets_2level, home_furry_2level
+## secondary_heat_3level, stove_age_3level, stove_cert, chimney_clean_3level,
+## home_damp, home_mold, home_fireplace, home_furnace, home_insert
 
-pm_home_char_results <- lm(log(mean_pm) ~ home_insert, data = pm_home_char)
+table(filter_data$home_insert)
+median(filter_data$home_furry, na.rm = TRUE)
+
+pm_home_char_results <- lm(log(pm_mean) ~ home_insert, data = filter_data)
 summary(pm_home_char_results)
@@
 ```{r}
-pm_stove_use <- pm_home_mean %>% 
-  left_join(stove_use_clean, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(burn_level = fct_collapse(burn_level,
-                                   "No/Light burning" = c("No burning", 
-                                                          "Light burning"),
-                                   "Ave burning" = "Average burning",
-                                   "Heavy burning" = "Heavy burning"),
-         burn_level = factor(burn_level,
-                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
-         wood_collect = fct_collapse(wood_collect,
-                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
-                                                   "1 mo to 3 mo"),
-                                   "3+ months" = c("3 mo to 6 mo", 
-                                                   "6 mo to 1 year", "> 1 yr")),
-         wood_collect = factor(wood_collect,
-                         levels = c("<3 months", "3+ months")),
-         wood_collect_method = fct_collapse(wood_collect_method,
-                                   "Harvest yourself" = "Harvest yourself",
-                                   "Purchase/other" = c("Mix", "Other", 
-                                                        "Purchase it")),
-         wood_cords = if_else(wood_cords > 10, 10, wood_cords),
-         wood_cords_cat = cut(wood_cords, breaks = c(0, 5, 50),
-                                labels = c("<5", "5+"),
-                                 right = FALSE))
+filter_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean))
 
-table(pm_stove_use$wood_cords_cat)
-median(pm_stove_use$wood_cords, na.rm = TRUE)
+# Variables to use:
+## burn_level_3level, wood_collect_2level, wood_collect_method_2level,
+## wood_cords
 
-pm_stove_use_results <- lm(log(mean_pm) ~ wood_collect_method, data = pm_stove_use)
+table(filter_data$home_insert)
+median(filter_data$wood_cords, na.rm = TRUE)
+
+pm_stove_use_results <- lm(log(pm_mean) ~ wood_collect_method_2level, data = filter_data)
 summary(pm_stove_use_results)
@@
 ```{r}
-stove_grades <- stove_grades_clean %>% 
-  arrange(area, home, winter_id) %>% 
-  group_by(home) %>% 
-  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade))
+filter_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean))
 
-pm_stove_grades <- pm_home_mean %>% 
-  left_join(stove_grades, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(stove_grade = fct_collapse(stove_grade,
-                                    "A" = "A",
-                                    "C" = "C",
-                                    "D or F" = c("D", "F")))
+# Variables to use:
+## stove_grade_3level
 
-table(pm_stove_grades$stove_grade)
+table(filter_data$stove_grade_3level)
 
-pm_stove_grades_results <- lm(log(mean_pm) ~ stove_grade, data = pm_stove_grades)
+pm_stove_grades_results <- lm(log(pm_mean) ~ stove_grade_3level, data = filter_data)
 summary(pm_stove_grades_results)
@@
 ```{r}
-# Join home activity with mean PM and format for analysis
-pm_home_act <- pm_clean %>% 
+filter_data <- exposure_analysis_data_long %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  arrange(area, home, sampling_day) %>% 
-  group_by(area, home, sampling_day) %>% 
-  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
-  ungroup() %>% 
-  mutate(home_act_day = as.factor(sampling_day)) %>% 
-  arrange(area, home, home_act_day) %>% 
-  left_join(home_act_clean, by = c("area", "home", "home_act_day")) %>% 
-  filter(winter_id == 1)
+  group_by(home) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean_sampling_day))
 
-table(pm_home_act$home_act_sweep)
+# Variables to use:
+## home_act_gas, home_act_elect, home_act_propane, home_act_oil, home_act_other,
+## home_act_smoke, home_act_incense, home_act_candle, home_act_windows,
+## home_act_door, home_act_sweep
 
-pm_home_act_results <- lmer(log(mean_pm) ~ home_act_sweep + (1 | home), data = pm_home_act)
+table(filter_data$home_act_sweep)
+
+pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
+                              (1 | home), data = filter_data)
 summary(pm_home_act_results)
@@
 ```{r}
-# Join moisture content with mean PM and format for analyzis
-pm_homes_placebo <- pm_clean %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
+filter_data <- exposure_analysis_data_long %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(!is.na(moisture_ave)) %>% 
+  select(sample_date, pm_first_datetime, area, home, 
+         moisture_ave, pm_mean_sampling_period) %>% 
+  separate(pm_first_datetime, c("pm_first_date", "first_time"), sep = " ") %>% 
   group_by(area, home) %>% 
-  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
-  filter(initial_obs == 1) %>% 
-  mutate(pm_home = "true") %>% 
-  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
-  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
-moisture_filtered <- moisture_clean %>% 
-  # change to correct joining data of interest
-  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
-  filter(pm_home == "true") %>% 
-  filter(winter_id == 1) %>% 
-  ungroup() %>%
-  arrange(area, home)
-
-moisture_filtered_time <- moisture_filtered %>% 
-  # fix wrong date
-  mutate(moisture_date = as.character(moisture_date),
-         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
-                                 "2017-04-10", moisture_date),
-         moisture_date = ymd(moisture_date)) %>% 
-  mutate(moisture_date = ymd(moisture_date),
-         first_date = ymd(first_date),
+  mutate(moisture_date = ymd(sample_date),
+         first_date = ymd(pm_first_date),
          time_diff = moisture_date - first_date,
@@
   ungroup() %>% 
-  filter(time_diff_rank == 1)
+  filter(time_diff_rank == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean_sampling_period))
 
-pm_moisture <- pm_home_mean %>% 
-  left_join(moisture_filtered_time, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(moisture_ave_5 = moisture_ave/5,
-         moisture_cat = cut(moisture_ave, breaks = c(0, 10, 50),
-                                labels = c("<10", "10+"),
-                                 right = FALSE))
+# Variables to use:
+## moisture_ave, mean_moisture_5
 
-table(pm_moisture$moisture_cat)
-median(pm_moisture$moisture_ave, na.rm = TRUE)
+table(filter_data$moisture_cat)
+median(filter_data$moisture_ave, na.rm = TRUE)
 
-pm_moisture_results <- lm(log(mean_pm) ~ moisture_cat, data = pm_moisture)
+pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
+                            moisture_ave, data = filter_data)
 summary(pm_moisture_results)
@@
 
-pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = pm_temp_rh)
+pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = filter_data)
 summary(pm_home_act_results)
@@
 
+sums_temp_change <- sums_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home, date_sums) %>% 
+  mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
+  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
+         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
+  replace_with_na(replace = list(lead_temp = 999)) %>% 
+  mutate(temp_diff = lead_temp - temp_c_sums,
+         temp_diff_abs = abs(lead_temp - temp_c_sums),
+         # change temp in the following line
+         temp_diff_check = if_else(temp_diff >= 8, 1, 0),
+         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
+                              temp_diff_check),
+         heat_event = if_else(is.na(temp_diff_check), 1,
+                              heat_event)) %>% 
+  ungroup() %>% 
+  mutate(date_sums = as.character(date_sums)) %>% 
+  group_by(area, home, date_sums) %>% 
+  summarize(sums_events = sum(heat_event, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  arrange(area, home, date_sums)
+
 pm_sums_events_joined <- pm_home_mean_date %>% 
-  left_join(sums_events_27, by = c("area", "home", "date_sums")) %>% 
-  filter(sums_event == 1) %>% 
+  left_join(sums_temp_change, by = c("area", "home", "date_sums")) %>% 
   arrange(area, home, date_sums) %>% 
-  group_by(area, home, mean_pm) %>% 
-  summarize(sums_events = n()) %>% 
   #mutate(sums_events = factor(sums_events))
-  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 1, 10),
-                           labels = c("1", ">1")))
+  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 2, 20),
+                           labels = c("<2", "2+"),
+                                 right = FALSE))
 
-median(pm_sums_events_joined$sums_events)
+median(pm_sums_events_joined$sums_events, na.rm = TRUE)
+table(pm_sums_events_joined$sums_events_cat)
 
-pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = pm_sums_events_joined)
-pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = pm_sums_events_joined)
+pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = filter_data)
+pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = filter_data)
 
@@
 qqnorm(residuals(pm_sums_model), main = "QQ Plot")
+
+pm_sums_plot <- pm_sums_events_joined %>% 
+  filter(!is.na(sums_events)) %>% 
+  ggplot() + 
+    geom_boxplot(aes(sums_events_cat, log(mean_pm))) +
+    theme_minimal()
+pm_sums_plot
+
 ```

### 2020-03-24 exposure_models.Rmd
merge datasets

@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds") %>% 
-  filter(!is.na(pm_mean_daily))
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
 
-exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") %>% 
-  filter(!is.na(pm_mean))
+exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") 
 ```

### 2020-03-24 merge_datasets.Rmd
merge datasets

---
title: "KidsAIR: merge datasets"
author: "Ethan Walker"
date: "Started 24 March 2020, Updated 24 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
```

# Load individual datasets
```{r}
pm_clean <- read_rds("Output/pm_clean.rds")
sums_clean <- read_rds("Output/sums_clean.rds")

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
demographics_clean <- read_rds("Output/demographics_clean.rds")
moisture_clean <- read_rds("Output/moisture_clean.rds")
indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
home_char_clean <- read_rds("Output/home_char_clean.rds")
home_act_clean <- read_rds("Output/home_act_clean.rds")
```

# Create new dataset with mean PM/SUMs; join with covariates
## Short data averaged over sampling period
```{r}
# Prep PM data to join with SUMs data
pm_join <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm = if_else(pm > 500, 500, pm),
         pm_mean = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  group_by(area, home) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime),
         winter_id = as.character("1")) %>% 
  ungroup() %>% 
  arrange(area, home) %>% 
  separate(pm_first_datetime, c("pm_date_start", "pm_time"), sep = " ", remove = FALSE) %>% 
  select(area, home, dtid, pm_total_observations, winter_id, home_winter_id, treatment,
         pm_mean, pm_first_datetime, pm_last_datetime, pm_date_start) %>% 
  group_by(area, home) %>% 
  distinct(pm_mean, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home)

# Prep SUMs data to join with PM data
sums_join <- sums_clean %>% 
  select(area, home, home_winter_id, winter_id, ibid, treatment, datetime_sums, temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(sums_mean = mean(temp_c_sums, na.rm = TRUE),
         sums_mean_5 = sums_mean/5,
         temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0),
         temp_perc_30 = (sum(temp_perc_30)/n()*100),
         temp_perc_27 = if_else(temp_c_sums >= 27, 1, 0),
         temp_perc_27 = (sum(temp_perc_27)/n()*100),
         temp_perc_25 = if_else(temp_c_sums >= 25, 1, 0),
         temp_perc_25 = (sum(temp_perc_25)/n()*100)) %>% 
  arrange(area, home, winter_id) %>%
  ungroup() %>% 
  distinct(home_winter_id, .keep_all = TRUE)

# Join PM/SUMs data
pm_sums_joined <- sums_join %>% 
  full_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  arrange(area, home, winter_id)

# Join PM/SUMs data with other covariate datasets
analysis_data1 <- home_act_clean %>% 
  arrange(area, home) %>% 
  select(-home_act_day, -home_act_comments, -home_id) %>% 
  group_by(area, home, home_winter_id, treatment, winter_id) %>% 
  # Sum home activity vars over entire sampling period to make 1 obs per sampling period
  mutate_if(is.factor, factor, levels = c("Yes", "No"), labels = c(1, 0)) %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate_if(is.character, as.numeric) %>% 
  summarize_if(is.numeric, sum, na.rm = TRUE) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  right_join(pm_sums_joined, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

analysis_data2 <- demographics_clean %>% 
  left_join(analysis_data1, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  mutate(income_3level = fct_collapse(income,
                               "Less than $20,000" = "Less than $20,000",
                               "$20,000-$49,999" = 
                                 c("$20,000 to $29,999", "$30,000 to $39,999",
                                   "$40,000 to $49,999"),
                                "$50,000+" = 
                                 c("$50,000 to $74,999", "$75,000 to $99,999",
                                   "$100,000 or more")),
         income_3level = factor(income_3level, levels = c("Less than $20,000", 
                                            "$20,000-$49,999", 
                                            "$50,000+")), 
         education_3level = fct_collapse(education,
                                  "High school or less" = 
                                    c("Less than high school", 
                                      "High school diploma or GED"),
                                  "Some college" = "Some college",
                                  "College degree" = "College degree"),
         education_3level = factor(education_3level, levels = c("High school or less", 
                                            "Some college", 
                                            "College degree")), 
         race_parent_2level = fct_collapse(race_parent,
                                    "White" = "White",
                                    "Non-white" = 
                                      c("American Indian/Alaskan Native",
                                        "Asian", "More than one race")),
         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
         total_residents_2level = fct_collapse(total_residents,
                                    "<5" = c("1", "2", "3", "4"),
                                    "5+" = c("5", "6", "7", "8", "9", "10",
                                             "11", "12", "13", "14", "15")),
         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
         residents_under5_2level = fct_collapse(residents_under_five,
                                    "1" = c("0", "1"),
                                    "2+" = c("2", "3", "4", "5")),
         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

analysis_data3 <- home_char_clean %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  filter(winter_id == 1 | winter_id == 2) %>% 
  select(area, home, home_winter_id, winter_id, treatment, residents_smoke:home_insert) %>% 
  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(home_type_2level = fct_collapse(home_type,
                                  "House" = "House",
                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
         home_type_2level = factor(home_type_2level,
                            levels = c("House", "Other")),
         home_year_built_2level = cut(home_year_built, breaks = c(0, 1985, 2020),
                                 labels = c("<1985", "1985+"),
                                 right = FALSE),
         stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")),
         stove_age_3level = fct_collapse(stove_age,
                                  "0-5 years old" = "0-5 years old",
                                  "6-15 years old" = c("6-10 years old",
                                                       "11-15 years old"),
                                  "16 + years old" = "16 + years old",
                                  "Do not know" = "Do not know"),
         stove_cert = factor(stove_cert,
                             levels = c("Yes", "No", "Do not know")),
         home_sqft_2level = cut(home_sqft, breaks = c(0, 1500, 5000),
                         labels = c("<1500", "1500+"),
                         right = FALSE),
         home_floors_2level = cut(home_floors, breaks = c(0, 2, 10),
                                 labels = c("1", "2+"),
                                 right = FALSE),
         home_bedrooms_2level = cut(home_bedrooms, breaks = c(0, 3, 50),
                                 labels = c("<3", "3+"),
                                 right = FALSE),
         home_windows_2level = cut(home_windows, breaks = c(0, 10, 100),
                                labels = c("<10", "10+"),
                                 right = FALSE),
         home_pets_2level = cut(home_pets, breaks = c(0, 2, 50),
                                labels = c("<2", "2+"),
                                 right = FALSE),
         home_furry_2level = cut(home_furry, breaks = c(0, 2, 50),
                                labels = c("<2", "2+"),
                                 right = FALSE),
         secondary_heat_3level = fct_collapse(secondary_heat,
                                       "Electric" = "Electrical",
                                       "Wood stove" = "Wood stove",
                                       "Other" = c("Natural gas furnace",
                                                   "Oil", "Other", "Propane")),
         chimney_clean = factor(chimney_clean,
                         levels = c("Less than 6 months ago", "6-12 months ago",
                                    "12-18 months ago", "More than 18 months ago")),
         chimney_clean_3level = fct_collapse(chimney_clean,
                                  "Less than 6 months ago" = "Less than 6 months ago",
                                  "6-12 months ago" = "6-12 months ago",
                                  "12 + months ago" = c("12-18 months ago",
                                                        "More than 18 months ago")),
         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know"))) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

analysis_data4 <- indoor_temp_rh_clean %>% 
  filter(winter_id == 1) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment, rh_indoor_current:temp_indoor_min) %>% 
  ungroup() %>% 
  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(temp_max_5 = temp_indoor_max/5,
         temp_min_5 = temp_indoor_min/5,
         rh_max_5 = rh_indoor_max/5,
         rh_min_5 = rh_indoor_min/5) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

analysis_data5 <- moisture_clean %>% 
  select(area, home, home_winter_id, winter_id, treatment, moisture_ave, moisture_split) %>% 
  # Average moisture over entire winter for each house
  group_by(area, home, winter_id) %>% 
  mutate(mean_moisture = mean(moisture_ave, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  replace_with_na(replace = list(mean_moisture = "NaN")) %>% 
  arrange(area, home, winter_id) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  select(-moisture_ave) %>% 
  right_join(analysis_data4, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(mean_moisture_5 = mean_moisture/5) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

analysis_data6 <- stove_grades_clean %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment, stove_grade) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  right_join(analysis_data5, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(stove_grade_3level = fct_collapse(stove_grade,
                                    "A" = "A",
                                    "C" = "C",
                                    "D or F" = c("D", "F"))) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

analysis_data7 <- stove_use_clean %>% 
  filter(winter_id == 1) %>% 
  select(-home_id) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  right_join(analysis_data6, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(burn_level_3level = fct_collapse(burn_level,
                                   "No/Light burning" = c("No burning", 
                                                          "Light burning"),
                                   "Ave burning" = "Average burning",
                                   "Heavy burning" = "Heavy burning"),
         burn_level_3level = factor(burn_level_3level,
                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
         wood_collect_2level = fct_collapse(wood_collect,
                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
                                                   "1 mo to 3 mo"),
                                   "3+ months" = c("3 mo to 6 mo", 
                                                   "6 mo to 1 year", "> 1 yr")),
         wood_collect_2level = factor(wood_collect_2level,
                         levels = c("<3 months", "3+ months")),
         wood_collect_method_2level = fct_collapse(wood_collect_method,
                                   "Harvest yourself" = "Harvest yourself",
                                   "Purchase/other" = c("Mix", "Other", 
                                                        "Purchase it")),
         wood_cords = if_else(wood_cords > 10, 10, wood_cords)) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

#write_rds(analysis_data7, "Output/exposure_analysis_data_short.rds")
#write_csv(analysis_data7, "Output/exposure_analysis_data_short.csv")
```

# Create new dataset with mean PM/SUMs; join with covariates
## Long data averaged over sampling day
```{r}
# Prep PM data to join with SUMs data by sampling day
pm_join <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm = if_else(pm > 500, 500, pm),
         pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ",
           remove = FALSE) %>% 
  group_by(area, home, sample_date) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE),
         pm_daily_obs = n()) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date) %>% 
  select(area, home, home_winter_id, winter_id, treatment, dtid, day_of_week, 
         pm_total_observations, pm_mean_sampling_period, sampling_day, 
         pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
         sample_date, pm_mean_daily, pm_daily_obs) %>% 
  group_by(area, home) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

# Prep SUMs data to join with PM data
sums_join <- sums_clean %>% 
  mutate(sample_date = as.character(date_sums)) %>% 
  select(area, home, home_winter_id, winter_id, ibid, treatment, 
         datetime_sums, sample_date, temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(sums_mean_winter = mean(temp_c_sums, na.rm = TRUE),
         sums_mean_5 = sums_mean_winter/5) %>% 
  group_by(area, home, winter_id, sample_date) %>% 
  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
         temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0),
         temp_perc_30 = (sum(temp_perc_30)/n()*100),
         temp_perc_27 = if_else(temp_c_sums >= 27, 1, 0),
         temp_perc_27 = (sum(temp_perc_27)/n()*100),
         temp_perc_25 = if_else(temp_c_sums >= 25, 1, 0),
         temp_perc_25 = (sum(temp_perc_25)/n()*100),
         sums_daily_obs = n()) %>% 
  #distinct(sums_mean_daily, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

# Join PM/SUMs data and filter/arrange by sampling day
pm_sums_joined1 <- sums_join %>% 
  left_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  arrange(area, home, datetime_sums) %>% 
  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
  filter(filter_var == 1) %>% 
  group_by(area, home, winter_id) %>%
  mutate(sums_mean_sampling_period = mean(temp_c_sums, na.rm = TRUE)) %>% 
  group_by(area, home, winter_id, sampling_day) %>% 
  mutate(sums_mean_sampling_day = mean(temp_c_sums, na.rm = TRUE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sums_mean_sampling_day, sums_mean_sampling_period)

pm_sums_joined2 <- pm_join %>% 
  left_join(pm_sums_joined1, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment", "sampling_day"))

# Join SUMs data with original PM data, by sampling day
pm_sums_joined3 <- sums_join %>% 
  select(-datetime_sums, -temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  arrange(area, home, winter_id, sample_date) %>% 
  full_join(pm_sums_joined2, by = c("area", "home", "home_winter_id", "winter_id", 
                                    "treatment", "sample_date")) %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

# Join PM/SUMs data with other covariate datasets
analysis_data1 <- home_act_clean %>% 
  mutate(sampling_day = home_act_day) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  select(-home_act_day, -home_act_comments, -home_id) %>% 
  full_join(pm_sums_joined3, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment", "sampling_day")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id, sample_date)

analysis_data2 <- demographics_clean %>% 
  full_join(analysis_data1, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  mutate(income_3level = fct_collapse(income,
                               "Less than $20,000" = "Less than $20,000",
                               "$20,000-$49,999" = 
                                 c("$20,000 to $29,999", "$30,000 to $39,999",
                                   "$40,000 to $49,999"),
                                "$50,000+" = 
                                 c("$50,000 to $74,999", "$75,000 to $99,999",
                                   "$100,000 or more")),
         income_3level = factor(income_3level, levels = c("Less than $20,000", 
                                            "$20,000-$49,999", 
                                            "$50,000+")), 
         education_3level = fct_collapse(education,
                                  "High school or less" = 
                                    c("Less than high school", 
                                      "High school diploma or GED"),
                                  "Some college" = "Some college",
                                  "College degree" = "College degree"),
         education_3level = factor(education_3level, levels = c("High school or less", 
                                            "Some college", 
                                            "College degree")), 
         race_parent_2level = fct_collapse(race_parent,
                                    "White" = "White",
                                    "Non-white" = 
                                      c("American Indian/Alaskan Native",
                                        "Asian", "More than one race")),
         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
         total_residents_2level = fct_collapse(total_residents,
                                    "<5" = c("1", "2", "3", "4"),
                                    "5+" = c("5", "6", "7", "8", "9", "10",
                                             "11", "12", "13", "14", "15")),
         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
         residents_under5_2level = fct_collapse(residents_under_five,
                                    "1" = c("0", "1"),
                                    "2+" = c("2", "3", "4", "5")),
         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data3 <- home_char_clean %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  filter(winter_id == 1 | winter_id == 2) %>% 
  select(area, home, home_winter_id, winter_id, treatment, residents_smoke:home_insert) %>% 
  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(home_type_2level = fct_collapse(home_type,
                                  "House" = "House",
                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
         home_type_2level = factor(home_type_2level,
                            levels = c("House", "Other")),
         home_year_built_2level = cut(home_year_built, breaks = c(0, 1985, 2020),
                                 labels = c("<1985", "1985+"),
                                 right = FALSE),
         stove_age = factor(stove_age,
                         levels = c("0-5 years old", "6-10 years old",
                                    "11-15 years old", "16 + years old",
                                    "Do not know")),
         stove_age_3level = fct_collapse(stove_age,
                                  "0-5 years old" = "0-5 years old",
                                  "6-15 years old" = c("6-10 years old",
                                                       "11-15 years old"),
                                  "16 + years old" = "16 + years old",
                                  "Do not know" = "Do not know"),
         stove_cert = factor(stove_cert,
                             levels = c("Yes", "No", "Do not know")),
         home_sqft_2level = cut(home_sqft, breaks = c(0, 1500, 5000),
                         labels = c("<1500", "1500+"),
                         right = FALSE),
         home_floors_2level = cut(home_floors, breaks = c(0, 2, 10),
                                 labels = c("1", "2+"),
                                 right = FALSE),
         home_bedrooms_2level = cut(home_bedrooms, breaks = c(0, 3, 50),
                                 labels = c("<3", "3+"),
                                 right = FALSE),
         home_windows_2level = cut(home_windows, breaks = c(0, 10, 100),
                                labels = c("<10", "10+"),
                                 right = FALSE),
         home_pets_2level = cut(home_pets, breaks = c(0, 2, 50),
                                labels = c("<2", "2+"),
                                 right = FALSE),
         home_furry_2level = cut(home_furry, breaks = c(0, 2, 50),
                                labels = c("<2", "2+"),
                                 right = FALSE),
         secondary_heat_3level = fct_collapse(secondary_heat,
                                       "Electric" = "Electrical",
                                       "Wood stove" = "Wood stove",
                                       "Other" = c("Natural gas furnace",
                                                   "Oil", "Other", "Propane")),
         chimney_clean = factor(chimney_clean,
                         levels = c("Less than 6 months ago", "6-12 months ago",
                                    "12-18 months ago", "More than 18 months ago")),
         chimney_clean_3level = fct_collapse(chimney_clean,
                                  "Less than 6 months ago" = "Less than 6 months ago",
                                  "6-12 months ago" = "6-12 months ago",
                                  "12 + months ago" = c("12-18 months ago",
                                                        "More than 18 months ago")),
         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data4 <- indoor_temp_rh_clean %>% 
  filter(winter_id == 1) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment, rh_indoor_current:temp_indoor_min) %>% 
  ungroup() %>% 
  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(temp_max_5 = temp_indoor_max/5,
         temp_min_5 = temp_indoor_min/5,
         rh_max_5 = rh_indoor_max/5,
         rh_min_5 = rh_indoor_min/5) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data5 <- moisture_clean %>% 
  mutate(sample_date = as.character(moisture_date)) %>% 
  filter(!is.na(moisture_ave)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, moisture_ave, 
         moisture_split, sample_date) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  full_join(analysis_data4, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment", "sample_date")) %>% 
  mutate(moisture_ave_5 = moisture_ave/5) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data6 <- stove_grades_clean %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment, stove_grade) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  right_join(analysis_data5, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  mutate(stove_grade_3level = fct_collapse(stove_grade,
                                    "A" = "A",
                                    "C" = "C",
                                    "D or F" = c("D", "F"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data7 <- stove_use_clean %>% 
  filter(winter_id == 1) %>% 
  select(-home_id) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  right_join(analysis_data6, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  mutate(burn_level_3level = fct_collapse(burn_level,
                                   "No/Light burning" = c("No burning", 
                                                          "Light burning"),
                                   "Ave burning" = "Average burning",
                                   "Heavy burning" = "Heavy burning"),
         burn_level_3level = factor(burn_level_3level,
                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
         wood_collect_2level = fct_collapse(wood_collect,
                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
                                                   "1 mo to 3 mo"),
                                   "3+ months" = c("3 mo to 6 mo", 
                                                   "6 mo to 1 year", "> 1 yr")),
         wood_collect_2level = factor(wood_collect_2level,
                         levels = c("<3 months", "3+ months")),
         wood_collect_method_2level = fct_collapse(wood_collect_method,
                                   "Harvest yourself" = "Harvest yourself",
                                   "Purchase/other" = c("Mix", "Other", 
                                                        "Purchase it")),
         wood_cords = if_else(wood_cords > 10, 10, wood_cords)) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

#write_rds(analysis_data7, "Output/exposure_analysis_data_long.rds")
#write_csv(analysis_data7, "Output/exposure_analysis_data_long.csv")
```

### 2020-03-26 demographics_summary.Rmd
data cleaning/merging updates

@@
 ---
-title: "KidsAIR: household demographics summary"
+title: 'KidsAIR: household demographics summary'
 author: "Ethan Walker"
-date: "Started 27 Jan 2020, Updated 13 March 2020"
+date: "Started 27 Jan 2020, Updated 25 March 2020"
 output:
-  powerpoint_presentation
+  beamer_presentation: default
+  powerpoint_presentation: default
 ---
@@
 
-# Summarize demographics data
 ```{r, include=FALSE}
@@
   filter(winter_id == 1) %>% 
-  filter(!is.na(hispanic_child)) %>% 
+  filter(residents_under_five != 0) %>% 
   mutate(income = factor(income,
@@
          total_residents = as.numeric(total_residents),
+         total_residents = if_else(total_residents == 1, 2, total_residents),
          race_parent = factor(race_parent,
@@
                                          "White", "More than one race"),
-                              labels = c("AI/AN", "White", "Multiple")))
+                              labels = c("AI/AN", "White", "Multiple")),
+         age_child_nov = factor(age_child_nov,
+                                levels = c("Less than 1", "1", "2", "3", "4")))
 ```
@@
           axis.text.y = element_text(size = 16, colour = "black"),
-          title = element_text(size = 18, colour = "black"),
-          legend.text = element_text(size = 14, colour = "black"),
-          legend.title = element_text(size = 16, colour = "black"),
-          axis.title.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 20, colour = "black"),
+          legend.text = element_text(size = 18, colour = "black"),
+          legend.title = element_text(size = 18, colour = "black"),
+          axis.title.y = element_text(size = 18, colour = "black"),
           axis.title.x = element_blank(),
@@
 ```{r}
-demographics_function(demographics_filtered, "residents_under_five", "Residents under 5", "count")
+demographics_function(demographics_filtered, "residents_under_five", 
+                      "Residents under 5 years of age", "count")
 ```

### 2020-03-26 exposure_initial_analysis.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 20 March 2020"
+date: "Started 2 Feb 2020, Updated 26 March 2020"
 output: pdf_document
@@
 library(lubridate)
+library(broom)
+library(broom.mixed)
 library(zoo)
@@
 ```{r}
-pm_clean <- read_rds("Output/pm_clean.rds")
-sums_clean <- read_rds("Output/sums_clean.rds")
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-demographics_clean <- read_rds("Output/demographics_clean.rds")
-moisture_clean <- read_rds("Output/moisture_clean.rds")
-indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
-stove_use_clean <- read_rds("Output/stove_use_clean.rds")
-stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
-home_char_clean <- read_rds("Output/home_char_clean.rds")
-home_act_clean <- read_rds("Output/home_act_clean.rds")
-
-# Mean PM for each placebo home over winter 1
-pm_home_mean <- pm_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  #filter(pm < 500) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  group_by(area, home) %>% 
-  summarize(mean_pm = mean(pm, na.rm = TRUE))
-
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
 
 exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")
+
+sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
 ```
@@
 ```{r}
-filter_data <- exposure_analysis_data_short %>% 
+analysis_data <- exposure_analysis_data_short %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
   distinct(home, .keep_all = TRUE) %>% 
-  filter(!is.na(pm_mean))
+  ungroup()
 
@@
 
-table(filter_data$residents_smoke)
-median(filter_data$home_furry, na.rm = TRUE)
+table(analysis_data$residents_smoke)
+median(analysis_data$home_furry, na.rm = TRUE)
 
-pm_demographics_results <- lm(log(pm_mean) ~ residents_smoke, data = filter_data)
+pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ residents_smoke, 
+                              data = analysis_data)
 summary(pm_demographics_results)
+tidy(pm_demographics_results)
 plot(pm_demographics_results)
@@
 
-table(filter_data$home_insert)
-median(filter_data$home_furry, na.rm = TRUE)
+table(analysis_data$home_insert)
+median(analysis_data$home_furry, na.rm = TRUE)
 
-pm_home_char_results <- lm(log(pm_mean) ~ home_insert, data = filter_data)
+pm_home_char_results <- lm(log(pm_mean_sampling_period) ~ home_insert, 
+                           data = analysis_data)
 summary(pm_home_char_results)
+tidy(pm_home_char_results)
 plot(pm_home_char_results)
@@
 
-table(filter_data$home_insert)
-median(filter_data$wood_cords, na.rm = TRUE)
+table(analysis_data$burn_level_3level)
+median(analysis_data$wood_cords, na.rm = TRUE)
 
-pm_stove_use_results <- lm(log(pm_mean) ~ wood_collect_method_2level, data = filter_data)
+pm_stove_use_results <- lm(log(pm_mean_sampling_period) ~ wood_collect_method_2level, 
+                           data = analysis_data)
 summary(pm_stove_use_results)
+tidy(pm_stove_use_results)
 plot(pm_stove_use_results)
@@
 
-table(filter_data$stove_grade_3level)
+table(analysis_data$stove_grade_3level)
 
-pm_stove_grades_results <- lm(log(pm_mean) ~ stove_grade_3level, data = filter_data)
+pm_stove_grades_results <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level, 
+                              data = analysis_data)
 summary(pm_stove_grades_results)
+tidy(pm_stove_grades_results)
 anova(pm_stove_grades_results)
@@
 ```{r}
-filter_data <- exposure_analysis_data_long %>% 
+analysis_data <- exposure_analysis_data_medium %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(home) %>% 
+  group_by(area, home) %>% 
   distinct(sampling_day, .keep_all = TRUE) %>% 
-  filter(!is.na(pm_mean_sampling_day))
+  ungroup()
 
@@
 
-table(filter_data$home_act_sweep)
+table(analysis_data$home_act_sweep)
 
 pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
-                              (1 | home), data = filter_data)
+                              (1 | home), data = analysis_data)
 summary(pm_home_act_results)
+tidy(pm_home_act_results)
 plot(pm_home_act_results, 
@@
 ```{r}
-filter_data <- exposure_analysis_data_long %>% 
+analysis_data <- exposure_analysis_data_short %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(!is.na(moisture_ave)) %>% 
-  select(sample_date, pm_first_datetime, area, home, 
-         moisture_ave, pm_mean_sampling_period) %>% 
-  separate(pm_first_datetime, c("pm_first_date", "first_time"), sep = " ") %>% 
   group_by(area, home) %>% 
-  mutate(moisture_date = ymd(sample_date),
-         first_date = ymd(pm_first_date),
-         time_diff = moisture_date - first_date,
-         time_diff = as.numeric(abs(time_diff))) %>% 
-  group_by(home) %>% 
-  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
+  distinct(home, .keep_all = TRUE) %>% 
   ungroup() %>% 
-  filter(time_diff_rank == 1) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  filter(!is.na(pm_mean_sampling_period))
+  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
+                                 labels = c("<10.5", "10.5+"),
+                                 right = FALSE),
+         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
+                                 labels = c("<10.5", "10.5+"),
+                                 right = FALSE))
 
 # Variables to use:
-## moisture_ave, mean_moisture_5
+## moisture_closest, moisture_closest_2level
+## mean_moisture_winter, moisture_winter_2level
 
-table(filter_data$moisture_cat)
-median(filter_data$moisture_ave, na.rm = TRUE)
+table(analysis_data$moisture_closest_2level)
+median(analysis_data$mean_moisture_winter, na.rm = TRUE)
 
 pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
-                            moisture_ave, data = filter_data)
+                            moisture_winter_2level, data = analysis_data)
 summary(pm_moisture_results)
@@
 
-# Indoor temp/Rh
+# Indoor temp/RH
 ```{r}
-pm_temp_rh <- pm_home_mean %>% 
-  left_join(indoor_temp_rh_clean, by = c("area", "home")) %>% 
-  filter(winter_id == 1) %>% 
-  mutate(temp_max_5 = temp_indoor_max/5,
-         temp_min_5 = temp_indoor_min/5,
-         rh_max_5 = rh_indoor_max/5,
-         rh_min_5 = rh_indoor_min/5,
-         temp_max_cat = cut(temp_indoor_max, breaks = c(0, 30, 45),
+analysis_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                  labels = c("<30", "30+"),
                                  right = FALSE),
-         temp_min_cat = cut(temp_indoor_min, breaks = c(0, 16.5, 50),
+         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                  labels = c("<16.5", "16.5+"),
                                  right = FALSE),
-         rh_max_cat = cut(rh_indoor_max, breaks = c(0, 40, 100),
+         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                  labels = c("<40", "40+"),
                                  right = FALSE),
-         rh_min_cat = cut(rh_indoor_min, breaks = c(0, 20, 100),
+         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                  labels = c("<20", "20+"),
@@
 
-table(pm_temp_rh$rh_min_cat)
-median(pm_temp_rh$rh_indoor_min, na.rm = TRUE)
-range(pm_temp_rh$rh_indoor_max, na.rm = TRUE)
+# Variables to use:
+## rh_min_5, rh_max_5, temp_min_5, temp_max_5
+## temp_max_2level, temp_min_2level,rh_max_2level ,rh_min_2level
 
-pm_home_act_results <- lm(log(mean_pm) ~ rh_min_cat, data = filter_data)
-summary(pm_home_act_results)
-plot(pm_home_act_results)
+table(analysis_data$temp_max_2level)
+median(analysis_data$rh_indoor_min, na.rm = TRUE)
+
+pm_temp_rh_results <- lm(log(pm_mean_sampling_period) ~ temp_max_2level, 
+                         data = analysis_data)
+summary(pm_temp_rh_results)
+tidy(pm_temp_rh_results)
+plot(pm_temp_rh_results)
 ```
 
-# SUMs - experimenting with several methods of looking at associations
+# SUMs - by sampling day
 ```{r}
-#### Join PM dataset full SUMs dataset by SUMs sampling times
+analysis_data <- exposure_analysis_data_medium %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  mutate(sums_mean_sampling_day_5 = sums_mean_sampling_day/5,
+         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
+                                 labels = c("<50", "50+"),
+                                 right = FALSE),
+         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
+                                 labels = c("<35", "35+"),
+                                 right = FALSE),
+         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
+                                 labels = c("<10", "10+"),
+                                 right = FALSE)) %>% 
+  ungroup()
 
-pm_join <- pm_clean %>% 
-  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
-  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+# Variables to use:
+## sums_mean_sampling_day_5, temp_perc_25_2level
+## temp_perc_27_2level, temp_perc_30_2level
+
+table(analysis_data$temp_perc_25_2level)
+median(analysis_data$temp_perc_30, na.rm = TRUE)
+
+pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_25_2level + 
+                              (1 | home), data = analysis_data)
+summary(pm_sums_results)
+tidy(pm_sums_results)
+plot(pm_sums_results, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_sums_results), main = "QQ Plot")
+```
+
+# SUMs - by sampling period
+```{r}
+analysis_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area, home) %>% 
-  arrange(sample_datetime) %>% 
-  mutate(pm_last_datetime = last(sample_datetime),
-         pm_first_datetime = first_datetime_new,
-         sample_datetime = round_date(sample_datetime, unit = "minute"),
-         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
-  ungroup() %>% 
-  arrange(area, home, sample_datetime)
+  distinct(home, .keep_all = TRUE) %>% 
+  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
+         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
+                                 labels = c("<50", "50+"),
+                                 right = FALSE),
+         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
+                                 labels = c("<35", "35+"),
+                                 right = FALSE),
+         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
+                                 labels = c("<10", "10+"),
+                                 right = FALSE)) %>% 
+  ungroup()
 
-sums_join <- sums_clean %>% 
-  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
-  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
-  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
-  ungroup() %>% 
-  arrange(area, home, sample_datetime)
+# Variables to use:
+## sums_mean_sampling_period_5, temp_perc_25_2level
+## temp_perc_27_2level, temp_perc_30_2level
 
-pm_sums_joined <- sums_join %>% 
-  left_join(pm_join, by = c("area", "home", "sample_datetime")) %>% 
-  filter(pm_rolling_20 < 1000) %>% 
-  filter(!is.na(temp_c_sums)) %>% 
-  filter(!is.na(pm_rolling_20)) %>% 
+table(analysis_data$temp_perc_25_2level)
+median(analysis_data$temp_perc_30, na.rm = TRUE)
+
+pm_sums_results <- lm(log(pm_mean_sampling_period) ~ sums_mean_sampling_period_5, 
+                      data = analysis_data)
+summary(pm_sums_results)
+tidy(pm_sums_results)
+plot(pm_sums_results, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_sums_results), main = "QQ Plot")
+```
+
+# SUMs - mean PM per SUMs interval
+```{r}
+analysis_data <- sums_rolling_pm_data %>% 
   mutate(temp_c_sums_5 = temp_c_sums/5)
 
-pm_sums_count <- pm_sums_joined %>% 
-  count(home)
-
-cor.test(pm_sums_joined$temp_c_sums, pm_sums_joined$pm_rolling_20, method = "spearman")
-pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), data = pm_sums_joined)
+cor.test(analysis_data$temp_c_sums, analysis_data$pm_rolling_20, method = "spearman")
+pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), 
+                      data = analysis_data)
 
 summary(pm_sums_model)
+tidy(pm_sums_model)
 exp((0.146291)-1)*100
@@
 
-pm_sums_plot <- pm_sums_joined %>% 
+pm_sums_plot <- analysis_data %>% 
   ggplot() + 
@@
 pm_sums_plot
+```
 
-#### Join PM dataset with SUMs dataset by sampling day
-
-pm_join <- pm_clean %>% 
+# SUMs - events per sampling day - KEEP WORKING ON THIS
+```{r}
+sums_temp_change <- sums_rolling_pm_data %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  mutate(pm = if_else(pm > 500, 500, pm),
-         pm_mean = mean(pm, na.rm = TRUE)) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(pm_mean_daily = mean(pm, na.rm = TRUE)) %>% 
-  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
-  group_by(area, home, sampling_day) %>% 
-  arrange(sample_datetime) %>% 
-  mutate(pm_last_datetime = last(sample_datetime),
-         pm_first_datetime = first(sample_datetime)) %>% 
-  ungroup() %>% 
-  arrange(area, home, sample_datetime) %>% 
-  select(area, home, home_winter_id, dtid, day_of_week, pm_total_observations,
-         pm_mean, sampling_day, pm_mean_daily, pm_first_datetime, pm_last_datetime) %>% 
-  group_by(area, home) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(area, home, sampling_day)
-
-sums_join <- sums_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  select(area, home, winter_id, ibid, treatment, datetime_sums, temp_c_sums,)
-
-pm_sums_joined <- sums_join %>% 
-  left_join(pm_join, by = c("area", "home")) %>% 
-  arrange(area, home, datetime_sums) %>% 
-  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
-                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
-  filter(filter_var == 1) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(temp_perc_25 = if_else(temp_c_sums >= 30, 1, 0)) %>% 
-  arrange(area, home, sampling_day) %>% 
-  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
-         sums_mean_daily_5 = sums_mean_daily/5) %>% 
-  group_by(area, home, home_winter_id, sampling_day, pm_mean, 
-           pm_mean_daily, sums_mean_daily) %>% 
-  summarize(temp_perc_25 = (sum(temp_perc_25)/n()*100)) %>% 
-  mutate(temp_perc_25_cat = cut(temp_perc_25, breaks = c(0, 10, 200),
-                                labels = c("< 50%", "50% +"),
-                                 right = FALSE)) %>%
-  distinct(pm_mean_daily, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(area, home, sampling_day)
-
-pm_sums_count <- pm_sums_joined %>% 
-  count(home)
-median(pm_sums_joined$temp_perc_25)
-table(pm_sums_joined$temp_perc_25_cat)
-
-cor.test(pm_sums_joined$sums_mean_daily, pm_sums_joined$pm_mean_daily, method = "spearman")
-pm_sums_model <- lmer(pm_mean_daily ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)
-pm_sums_model <- lmer(log(pm_mean_daily) ~ temp_perc_25_cat + (1 | home), data = pm_sums_joined)
-
-summary(pm_sums_model)
-exp((0.146291)-1)*100
-plot(pm_sums_model, 
-     xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(pm_sums_model), main = "QQ Plot")
-
-#### Join PM dataset with SUMs events datasets
-
-# Mean PM for each placebo home over winter 1
-pm_home_mean_date <- pm_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
-  separate(pm_datetime_new, c("date_sums", "pm_time"), sep = " ", remove = TRUE) %>% 
-  group_by(area, home, date_sums) %>% 
-  summarize(mean_pm = mean(pm, na.rm = TRUE)) %>% 
-  group_by(area, home, date_sums) %>% 
-  distinct(date_sums, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(area, home, date_sums)
-
-sums_temp_change <- sums_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home, date_sums) %>% 
+  group_by(area, home, sample_date) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
   mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
@@
   ungroup() %>% 
-  mutate(date_sums = as.character(date_sums)) %>% 
-  group_by(area, home, date_sums) %>% 
-  summarize(sums_events = sum(heat_event, na.rm = TRUE)) %>% 
+  mutate(sample_date = as.character(sample_date)) %>% 
+  group_by(area, home, sample_date) %>% 
+  mutate(sums_events = sum(heat_event, na.rm = TRUE),
+         sums_events_cat = cut(sums_events, breaks = c(0, 5, 50),
+                           labels = c("<4", "4+"),
+                                 right = FALSE)) %>% 
   ungroup() %>% 
-  arrange(area, home, date_sums)
+  arrange(area, home, sample_date)
 
-pm_sums_events_joined <- pm_home_mean_date %>% 
-  left_join(sums_temp_change, by = c("area", "home", "date_sums")) %>% 
-  arrange(area, home, date_sums) %>% 
-  #mutate(sums_events = factor(sums_events))
-  mutate(sums_events_cat = cut(sums_events, breaks = c(0, 2, 20),
-                           labels = c("<2", "2+"),
-                                 right = FALSE))
+median(sums_temp_change$sums_events, na.rm = TRUE)
+table(sums_temp_change$sums_events_cat)
 
-median(pm_sums_events_joined$sums_events, na.rm = TRUE)
-table(pm_sums_events_joined$sums_events_cat)
-
-pm_sums_model <- lmer(mean_pm ~ sums_events_cat + (1 | home), data = filter_data)
-pm_sums_model <- lmer(log(mean_pm) ~ sums_events_cat + (1 | home), data = filter_data)
+pm_sums_model <- lmer(mean_pm ~ sums_events + (1 | home), data = sums_temp_change)
+pm_sums_model <- lmer(log(pm_mean_daily) ~ sums_events + (1 | home), data = sums_temp_change)
 
@@
 pm_sums_plot
-
 ```

### 2020-03-26 home_char_act_summary.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 16 March 2020"
+date: "Started 15 Jan 2020, Updated 25 March 2020"
 output:
@@
   powerpoint_presentation: default
-  word_document: default
 ---
@@
 # Load dataset
-home_char_clean <- read_rds("Output/home_char_clean.rds") %>% 
-  mutate(chimney_clean = factor(chimney_clean,
-                                levels = c("Less than 6 months ago",
-                                           "6-12 months ago", "12-18 months ago",
-                                           "More than 18 months ago"),
-                                labels = c("< 6 months ago",
-                                           "6-12 months ago", "12-18 months ago",
-                                           "> 18 months ago")),
-         stove_age = factor(stove_age,
-                            levels = c("0-5 years old", "6-10 years old",
-                                       "11-15 years old", "16 + years old",
-                                       "Do not know")))
-
+home_char_clean <- read_rds("Output/home_char_clean.rds") 
 home_act_clean <- read_rds("Output/home_act_clean.rds")
 kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-pm_clean <- read_rds("Output/pm_clean.rds")
+#pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
 
+```{r}
+table(home_char_clean$primary_heat, home_char_clean$secondary_heat, home_char_clean$area,
+      dnn = c("primary heating", "secondary heating", "area"))
+```
+
 ```{r, include=FALSE}
@@
 
+```{r}
+home_char_filtered <- read_rds("Output/home_char_clean.rds") %>% 
+  mutate(chimney_clean = factor(chimney_clean,
+                                levels = c("Less than 6 months ago",
+                                           "6-12 months ago", "12-18 months ago",
+                                           "More than 18 months ago"),
+                                labels = c("< 6 months ago",
+                                           "6-12 months ago", "12-18 months ago",
+                                           "> 18 months ago")),
+         stove_age = factor(stove_age,
+                            levels = c("0-5 years old", "6-10 years old",
+                                       "11-15 years old", "16 + years old",
+                                       "Do not know")))
+```
+
 ```{r, fig.height=8}
-home_char_plot <- home_char_clean %>% 
+home_char_plot <- home_char_filtered %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
@@
           axis.ticks = element_blank()) +
-    facet_wrap(~var, ncol = 2)
+    facet_wrap(~var, ncol = 2) +
+    scale_fill_manual(values = jv_palette)
 home_char_plot
@@
 ```{r}
-home_char_function <- function(data, variable) {
+home_char_function <- function(data, var, title_label, y_label) {
 
-home_plot <- data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  select(area, variable) %>% 
-  rename(plot_var = variable) %>% 
-  #na.exclude() %>% 
+  home_char_plot <- data %>% 
+  filter(winter_id == 1) %>% 
+  select(area, var) %>% 
+  rename(plot_var = var) %>% 
+  filter(!is.na(plot_var)) %>% 
   ggplot() +
@@
     theme_minimal() +
-    labs(title = variable) +
     theme(axis.text.x = element_text(size = 16, colour = "black", 
@@
           axis.text.y = element_text(size = 16, colour = "black"),
-          title = element_text(size = 18, colour = "black"),
-          legend.text = element_text(size = 14, colour = "black"),
-          legend.title = element_text(size = 16, colour = "black"),
-          axis.title.y = element_blank(),
+          title = element_text(size = 20, colour = "black"),
+          legend.text = element_text(size = 18, colour = "black"),
+          legend.title = element_text(size = 18, colour = "black"),
+          axis.title.y = element_text(size = 18, colour = "black"),
           axis.title.x = element_blank(),
@@
           panel.grid = element_blank()) +
+    labs(title = title_label) +
+    ylab(y_label) +
     scale_fill_manual(values = jv_palette)
-home_plot
+  home_char_plot
 
@@
 
-home_char_function(home_char_clean, "residents_total")
-home_char_function(home_char_clean, "residents_under5")
-home_char_function(home_char_clean, "residents_under18")
-home_char_function(home_char_clean, "residents_smoke")
-home_char_function(home_char_clean, "residents_smoke_inside")
-home_char_function(home_char_clean, "home_type")
-home_char_function(home_char_clean, "home_floors")
-home_char_function(home_char_clean, "home_windows")
-home_char_function(home_char_clean, "home_bedrooms")
-home_char_function(home_char_clean, "home_pets")
-home_char_function(home_char_clean, "home_furry")
-home_char_function(home_char_clean, "primary_heat")
-home_char_function(home_char_clean, "secondary_heat")
-home_char_function(home_char_clean, "stove_age")
-home_char_function(home_char_clean, "stove_cert")
-home_char_function(home_char_clean, "chimney_clean")
-home_char_function(home_char_clean, "home_mold")
-home_char_function(home_char_clean, "home_damp")
-home_char_function(home_char_clean, "home_fireplace")
-home_char_function(home_char_clean, "home_furnace")
-home_char_function(home_char_clean, "home_woodstove")
-home_char_function(home_char_clean, "home_insert")
+home_char_function(home_char_filtered, "residents_smoke", "Residents smoke", "count")
+home_char_function(home_char_filtered, "residents_smoke_inside", "Residents smoke inside", "count")
+home_char_function(home_char_filtered, "home_type", "Home type", "count")
+home_char_function(home_char_filtered, "home_floors", "Floors in home", "count")
+home_char_function(home_char_filtered, "home_windows", "Windows in home", "count")
+home_char_function(home_char_filtered, "home_bedrooms", "Bedrooms in home", "count")
+home_char_function(home_char_filtered, "home_pets", "Pets in home", "count")
+home_char_function(home_char_filtered, "home_furry", "Furry pet in home", "count")
+home_char_function(home_char_filtered, "primary_heat", "Primary heat", "count")
+home_char_function(home_char_filtered, "secondary_heat", "Secondary heat", "count")
+home_char_function(home_char_filtered, "stove_age", "Stove age", "count")
+home_char_function(home_char_filtered, "stove_cert", "Stove EPA certified", "count")
+home_char_function(home_char_filtered, "chimney_clean", "Chimney last cleaned", "count")
+home_char_function(home_char_filtered, "home_mold", "Home mold", "count")
+home_char_function(home_char_filtered, "home_damp", "Home damp", "count")
+home_char_function(home_char_filtered, "home_fireplace", "Home fireplace", "count")
+home_char_function(home_char_filtered, "home_furnace", "Home furnace", "count")
 ```
@@
           axis.ticks = element_blank()) +
-    facet_wrap(~var, ncol = 2)
+    facet_wrap(~var, ncol = 2) +
+    scale_fill_manual(values = jv_palette)
 home_act_plot
@@
 ```{r}
-home_act_function <- function(data, variable) {
+home_act_function <- function(data, var, title_label, y_label) {
 
-home_plot <- data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  select(area, variable) %>% 
-  rename(plot_var = variable) %>% 
-  #na.exclude() %>% 
+  home_act_plot <- data %>% 
+  filter(winter_id == 1) %>% 
+  select(area, var) %>% 
+  rename(plot_var = var) %>% 
+  filter(!is.na(plot_var)) %>% 
   ggplot() +
@@
           panel.grid = element_blank()) +
+    labs(title = title_label) +
+    ylab(y_label) +
     scale_fill_manual(values = jv_palette)
-home_plot
+  home_act_plot
 
@@
 
-home_act_function(home_act_clean, "home_act_gas")
-home_act_function(home_act_clean, "home_act_elect")
-home_act_function(home_act_clean, "home_act_propane")
-home_act_function(home_act_clean, "home_act_oil")
-home_act_function(home_act_clean, "home_act_smoking")
-home_act_function(home_act_clean, "home_act_incense")
-home_act_function(home_act_clean, "home_act_candle")
-home_act_function(home_act_clean, "home_act_kerosene")
-home_act_function(home_act_clean, "home_act_windows")
-home_act_function(home_act_clean, "home_act_door")
-home_act_function(home_act_clean, "home_act_sweep")
+home_act_function(home_act_clean, "home_act_gas", "Gas appliance used", "count")
+home_act_function(home_act_clean, "home_act_elect", "Electric appliance used", "count")
+home_act_function(home_act_clean, "home_act_propane", "Propane appliance used", "count")
+home_act_function(home_act_clean, "home_act_oil", "Oil appliance used", "count")
+home_act_function(home_act_clean, "home_act_smoking", "Smoking in home", "count")
+home_act_function(home_act_clean, "home_act_incense", "Incense burned", "count")
+home_act_function(home_act_clean, "home_act_candle", "Candle burned", "count")
+home_act_function(home_act_clean, "home_act_kerosene", "Kerosene burned", "count")
+home_act_function(home_act_clean, "home_act_windows", "Windows open", "count")
+home_act_function(home_act_clean, "home_act_door", "Doors open", "count")
+home_act_function(home_act_clean, "home_act_sweep", "Sweeping/cleaning/vacuuming", "count")
 ```
 
-```{r, include=FALSE}
+```{r, include=FALSE, eval=FALSE}
 #### Abstract/manuscript tables ####
@@
 
-```{r, include=FALSE}
+```{r, include=FALSE, eval=FALSE}
 # Summarize numeric vars

### 2020-03-26 merge_datasets.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 24 March 2020, Updated 24 March 2020"
+date: "Started 24 March 2020, Updated 26 March 2020"
 output: pdf_document
@@
 ```{r}
-pm_clean <- read_rds("Output/pm_clean.rds")
-sums_clean <- read_rds("Output/sums_clean.rds")
+#pm_clean <- read_rds("Output/pm_clean.rds")
+#sums_clean <- read_rds("Output/sums_clean.rds")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+#kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 demographics_clean <- read_rds("Output/demographics_clean.rds")
@@
 # Create new dataset with mean PM/SUMs; join with covariates
-## Short data averaged over sampling period
-```{r}
-# Prep PM data to join with SUMs data
-pm_join <- pm_clean %>% 
-  group_by(area, home) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm),
-         pm_mean = mean(pm, na.rm = TRUE)) %>% 
-  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
-  group_by(area, home) %>% 
-  arrange(sample_datetime) %>% 
-  mutate(pm_last_datetime = last(sample_datetime),
-         pm_first_datetime = first(sample_datetime),
-         winter_id = as.character("1")) %>% 
-  ungroup() %>% 
-  arrange(area, home) %>% 
-  separate(pm_first_datetime, c("pm_date_start", "pm_time"), sep = " ", remove = FALSE) %>% 
-  select(area, home, dtid, pm_total_observations, winter_id, home_winter_id, treatment,
-         pm_mean, pm_first_datetime, pm_last_datetime, pm_date_start) %>% 
-  group_by(area, home) %>% 
-  distinct(pm_mean, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  arrange(area, home)
-
-# Prep SUMs data to join with PM data
-sums_join <- sums_clean %>% 
-  select(area, home, home_winter_id, winter_id, ibid, treatment, datetime_sums, temp_c_sums) %>% 
-  group_by(area, home, winter_id) %>% 
-  mutate(sums_mean = mean(temp_c_sums, na.rm = TRUE),
-         sums_mean_5 = sums_mean/5,
-         temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0),
-         temp_perc_30 = (sum(temp_perc_30)/n()*100),
-         temp_perc_27 = if_else(temp_c_sums >= 27, 1, 0),
-         temp_perc_27 = (sum(temp_perc_27)/n()*100),
-         temp_perc_25 = if_else(temp_c_sums >= 25, 1, 0),
-         temp_perc_25 = (sum(temp_perc_25)/n()*100)) %>% 
-  arrange(area, home, winter_id) %>%
-  ungroup() %>% 
-  distinct(home_winter_id, .keep_all = TRUE)
-
-# Join PM/SUMs data
-pm_sums_joined <- sums_join %>% 
-  full_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  arrange(area, home, winter_id)
-
-# Join PM/SUMs data with other covariate datasets
-analysis_data1 <- home_act_clean %>% 
-  arrange(area, home) %>% 
-  select(-home_act_day, -home_act_comments, -home_id) %>% 
-  group_by(area, home, home_winter_id, treatment, winter_id) %>% 
-  # Sum home activity vars over entire sampling period to make 1 obs per sampling period
-  mutate_if(is.factor, factor, levels = c("Yes", "No"), labels = c(1, 0)) %>% 
-  mutate_if(is.factor, as.character) %>% 
-  mutate_if(is.character, as.numeric) %>% 
-  summarize_if(is.numeric, sum, na.rm = TRUE) %>% 
-  mutate(winter_id = as.character(winter_id)) %>% 
-  right_join(pm_sums_joined, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-analysis_data2 <- demographics_clean %>% 
-  left_join(analysis_data1, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  filter(!is.na(home_winter_id)) %>% 
-  mutate(income_3level = fct_collapse(income,
-                               "Less than $20,000" = "Less than $20,000",
-                               "$20,000-$49,999" = 
-                                 c("$20,000 to $29,999", "$30,000 to $39,999",
-                                   "$40,000 to $49,999"),
-                                "$50,000+" = 
-                                 c("$50,000 to $74,999", "$75,000 to $99,999",
-                                   "$100,000 or more")),
-         income_3level = factor(income_3level, levels = c("Less than $20,000", 
-                                            "$20,000-$49,999", 
-                                            "$50,000+")), 
-         education_3level = fct_collapse(education,
-                                  "High school or less" = 
-                                    c("Less than high school", 
-                                      "High school diploma or GED"),
-                                  "Some college" = "Some college",
-                                  "College degree" = "College degree"),
-         education_3level = factor(education_3level, levels = c("High school or less", 
-                                            "Some college", 
-                                            "College degree")), 
-         race_parent_2level = fct_collapse(race_parent,
-                                    "White" = "White",
-                                    "Non-white" = 
-                                      c("American Indian/Alaskan Native",
-                                        "Asian", "More than one race")),
-         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
-         total_residents_2level = fct_collapse(total_residents,
-                                    "<5" = c("1", "2", "3", "4"),
-                                    "5+" = c("5", "6", "7", "8", "9", "10",
-                                             "11", "12", "13", "14", "15")),
-         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
-         residents_under5_2level = fct_collapse(residents_under_five,
-                                    "1" = c("0", "1"),
-                                    "2+" = c("2", "3", "4", "5")),
-         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-analysis_data3 <- home_char_clean %>% 
-  mutate(winter_id = as.character(winter_id)) %>%
-  filter(winter_id == 1 | winter_id == 2) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, residents_smoke:home_insert) %>% 
-  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  mutate(home_type_2level = fct_collapse(home_type,
-                                  "House" = "House",
-                                  "Other" = c("Duplex/Apt", "Mobile Home", "Other")),
-         home_type_2level = factor(home_type_2level,
-                            levels = c("House", "Other")),
-         home_year_built_2level = cut(home_year_built, breaks = c(0, 1985, 2020),
-                                 labels = c("<1985", "1985+"),
-                                 right = FALSE),
-         stove_age = factor(stove_age,
-                         levels = c("0-5 years old", "6-10 years old",
-                                    "11-15 years old", "16 + years old",
-                                    "Do not know")),
-         stove_age_3level = fct_collapse(stove_age,
-                                  "0-5 years old" = "0-5 years old",
-                                  "6-15 years old" = c("6-10 years old",
-                                                       "11-15 years old"),
-                                  "16 + years old" = "16 + years old",
-                                  "Do not know" = "Do not know"),
-         stove_cert = factor(stove_cert,
-                             levels = c("Yes", "No", "Do not know")),
-         home_sqft_2level = cut(home_sqft, breaks = c(0, 1500, 5000),
-                         labels = c("<1500", "1500+"),
-                         right = FALSE),
-         home_floors_2level = cut(home_floors, breaks = c(0, 2, 10),
-                                 labels = c("1", "2+"),
-                                 right = FALSE),
-         home_bedrooms_2level = cut(home_bedrooms, breaks = c(0, 3, 50),
-                                 labels = c("<3", "3+"),
-                                 right = FALSE),
-         home_windows_2level = cut(home_windows, breaks = c(0, 10, 100),
-                                labels = c("<10", "10+"),
-                                 right = FALSE),
-         home_pets_2level = cut(home_pets, breaks = c(0, 2, 50),
-                                labels = c("<2", "2+"),
-                                 right = FALSE),
-         home_furry_2level = cut(home_furry, breaks = c(0, 2, 50),
-                                labels = c("<2", "2+"),
-                                 right = FALSE),
-         secondary_heat_3level = fct_collapse(secondary_heat,
-                                       "Electric" = "Electrical",
-                                       "Wood stove" = "Wood stove",
-                                       "Other" = c("Natural gas furnace",
-                                                   "Oil", "Other", "Propane")),
-         chimney_clean = factor(chimney_clean,
-                         levels = c("Less than 6 months ago", "6-12 months ago",
-                                    "12-18 months ago", "More than 18 months ago")),
-         chimney_clean_3level = fct_collapse(chimney_clean,
-                                  "Less than 6 months ago" = "Less than 6 months ago",
-                                  "6-12 months ago" = "6-12 months ago",
-                                  "12 + months ago" = c("12-18 months ago",
-                                                        "More than 18 months ago")),
-         home_damp = factor(home_damp, levels = c("Yes", "No", "Do not know")),
-         home_mold = factor(home_mold, levels = c("Yes", "No", "Do not know")),
-         home_fireplace = factor(home_fireplace, levels = c("Yes", "No", "Do not know")),
-         home_furnace = factor(home_furnace, levels = c("Yes", "No", "Do not know")),
-         home_insert = factor(home_insert, levels = c("Yes", "No", "Do not know"))) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-analysis_data4 <- indoor_temp_rh_clean %>% 
-  filter(winter_id == 1) %>% 
-  mutate(winter_id = as.character(winter_id)) %>%
-  select(area, home, home_winter_id, winter_id, treatment, rh_indoor_current:temp_indoor_min) %>% 
-  ungroup() %>% 
-  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  mutate(temp_max_5 = temp_indoor_max/5,
-         temp_min_5 = temp_indoor_min/5,
-         rh_max_5 = rh_indoor_max/5,
-         rh_min_5 = rh_indoor_min/5) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-analysis_data5 <- moisture_clean %>% 
-  select(area, home, home_winter_id, winter_id, treatment, moisture_ave, moisture_split) %>% 
-  # Average moisture over entire winter for each house
-  group_by(area, home, winter_id) %>% 
-  mutate(mean_moisture = mean(moisture_ave, na.rm = TRUE)) %>% 
-  ungroup() %>% 
-  mutate(winter_id = as.character(winter_id)) %>%
-  replace_with_na(replace = list(mean_moisture = "NaN")) %>% 
-  arrange(area, home, winter_id) %>% 
-  distinct(home_winter_id, .keep_all = TRUE) %>% 
-  select(-moisture_ave) %>% 
-  right_join(analysis_data4, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  mutate(mean_moisture_5 = mean_moisture/5) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-analysis_data6 <- stove_grades_clean %>% 
-  arrange(area, home, winter_id) %>% 
-  group_by(home) %>% 
-  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
-  mutate(winter_id = as.character(winter_id)) %>%
-  select(area, home, home_winter_id, winter_id, treatment, stove_grade) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  right_join(analysis_data5, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  mutate(stove_grade_3level = fct_collapse(stove_grade,
-                                    "A" = "A",
-                                    "C" = "C",
-                                    "D or F" = c("D", "F"))) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-analysis_data7 <- stove_use_clean %>% 
-  filter(winter_id == 1) %>% 
-  select(-home_id) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  mutate(winter_id = as.character(winter_id)) %>%
-  right_join(analysis_data6, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
-  mutate(burn_level_3level = fct_collapse(burn_level,
-                                   "No/Light burning" = c("No burning", 
-                                                          "Light burning"),
-                                   "Ave burning" = "Average burning",
-                                   "Heavy burning" = "Heavy burning"),
-         burn_level_3level = factor(burn_level_3level,
-                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
-         wood_collect_2level = fct_collapse(wood_collect,
-                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
-                                                   "1 mo to 3 mo"),
-                                   "3+ months" = c("3 mo to 6 mo", 
-                                                   "6 mo to 1 year", "> 1 yr")),
-         wood_collect_2level = factor(wood_collect_2level,
-                         levels = c("<3 months", "3+ months")),
-         wood_collect_method_2level = fct_collapse(wood_collect_method,
-                                   "Harvest yourself" = "Harvest yourself",
-                                   "Purchase/other" = c("Mix", "Other", 
-                                                        "Purchase it")),
-         wood_cords = if_else(wood_cords > 10, 10, wood_cords)) %>% 
-  ungroup() %>% 
-  arrange(area, home, winter_id)
-
-#write_rds(analysis_data7, "Output/exposure_analysis_data_short.rds")
-#write_csv(analysis_data7, "Output/exposure_analysis_data_short.csv")
-```
-
-# Create new dataset with mean PM/SUMs; join with covariates
-## Long data averaged over sampling day
+## Long data averaged over sampling date, including SUMs dates, repeated by child_id
 ```{r}
@@
 
+#write_rds(pm_join, "Output/pm_join.rds")
+pm_join <- read_rds("Output/pm_join.rds")
+
 # Prep SUMs data to join with PM data
@@
 
+#write_rds(sums_join, "Output/sums_join.rds")
+sums_join <- read_rds("Output/sums_join.rds")
+
 # Join PM/SUMs data and filter/arrange by sampling day
@@
          sums_mean_sampling_day, sums_mean_sampling_period)
+
+#write_rds(pm_sums_joined1, "Output/pm_sums_joined1.rds")
+pm_sums_joined1 <- read_rds("Output/pm_sums_joined1.rds")
 
@@
 
-analysis_data5 <- moisture_clean %>% 
-  mutate(sample_date = as.character(moisture_date)) %>% 
+# Format and merge wood moisture content data
+pm_join2 <- pm_join %>% 
+  filter(sampling_day == 1) %>% 
+  separate(pm_first_datetime, c("pm_first_date", "trash"), sep = " ") %>% 
+  select(area, home, winter_id, pm_first_date) %>% 
+  group_by(area, home, winter_id) %>% 
+  distinct(pm_first_date, .keep_all = TRUE)
+
+pm_moisture <- moisture_clean %>% 
+  mutate(moisture_date = as.character(moisture_date),
+         winter_id = as.character(winter_id)) %>% 
   filter(!is.na(moisture_ave)) %>% 
   select(area, home, home_winter_id, winter_id, treatment, moisture_ave, 
-         moisture_split, sample_date) %>% 
-  mutate(winter_id = as.character(winter_id)) %>%
-  full_join(analysis_data4, by = c("area", "home", "home_winter_id", 
-                                    "winter_id", "treatment", "sample_date")) %>% 
-  mutate(moisture_ave_5 = moisture_ave/5) %>% 
+         moisture_split, moisture_date) %>% 
+  left_join(pm_join2, by = c("area", "home", "winter_id")) %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(area, home, winter_id) %>% 
+  mutate(mean_moisture_winter = mean(moisture_ave, na.rm = TRUE),
+         moisture_obs_winter = n(),
+         moisture_date = ymd(moisture_date),
+         pm_first_date = ymd(pm_first_date),
+         time_diff = moisture_date - pm_first_date,
+         time_diff_rank = rank(abs(time_diff)),
+         moisture_split_numeric = factor(moisture_split,
+                                 levels = c("Yes", "No"),
+                                 labels = c(1, 0)),
+         moisture_split_numeric = as.numeric(moisture_split_numeric),
+         moisture_split_numeric = if_else(moisture_split_numeric == 2, 0, moisture_split_numeric),
+         moisture_split_percent_winter = (sum(moisture_split_numeric)/n())*100,
+         moisture_split_closest = moisture_split) %>% 
+  filter(time_diff_rank == 1) %>% 
+  mutate(moisture_closest = if_else(!is.na(pm_first_date), moisture_ave, 999),
+         moisture_closest_date = moisture_date,
+         moisture_closest_date = as.character(moisture_closest_date),
+         moisture_closest_date = if_else(is.na(time_diff), "NA", moisture_closest_date),
+         pm_moisture_days_diff = time_diff) %>% 
+  replace_with_na(replace = list(moisture_closest = 999)) %>% 
+  replace_with_na(replace = list(moisture_closest_date = "NA")) %>% 
+  ungroup() %>% 
+  arrange(area, home, winter_id) %>% 
+  select(-moisture_ave, -time_diff, -time_diff_rank, -moisture_split, 
+         -moisture_split_numeric)
+
+analysis_data5 <- pm_moisture %>% 
+  right_join(analysis_data4, by = c("area", "home", "home_winter_id", 
+                                    "winter_id", "treatment")) %>% 
   ungroup() %>% 
@@
 #write_csv(analysis_data7, "Output/exposure_analysis_data_long.csv")
+
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+
+save_var_names <- data.frame(t(exposure_analysis_data_long)) %>% 
+  rownames_to_column() %>% 
+  select(rowname)
+write_csv(save_var_names, "Output/kids_data_var_names.csv")
 ```
+
+# Update long dataset -- REMEMBER TO UPDATE SHORT DATA BELOW
+```{r}
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+
+exposure_analysis_data_long_new <- exposure_analysis_data_long %>%
+  # fix incorrect data after checking paper logs
+  mutate(total_residents = if_else(home == "WMA438", "6", total_residents),
+         total_residents = if_else(home == "WMA480" & winter_id == 1, "13", total_residents),
+         total_residents_2level = fct_collapse(total_residents,
+                                    "<5" = c("1", "2", "3", "4"),
+                                    "5+" = c("5", "6", "7", "8", "9", "10",
+                                             "11", "12", "13", "14")),
+         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
+         income = as.character(income),
+         income = if_else(home == "WMA480" & winter_id == 1, "$40,000 to $49,999", income),
+         income_3level = fct_collapse(income,
+                               "Less than $20,000" = "Less than $20,000",
+                               "$20,000-$49,999" = 
+                                 c("$20,000 to $29,999", "$30,000 to $39,999",
+                                   "$40,000 to $49,999"),
+                                "$50,000+" = 
+                                 c("$50,000 to $74,999", "$75,000 to $99,999",
+                                   "$100,000 or more")),
+         income_3level = factor(income_3level, levels = c("Less than $20,000", 
+                                            "$20,000-$49,999", 
+                                            "$50,000+"))) %>% 
+  ungroup() %>% 
+  arrange(area, home, home_winter_id, sample_date)
+
+#write_rds(exposure_analysis_data_long_new, "Output/exposure_analysis_data_long_new.rds")
+#write_csv(exposure_analysis_data_long_new, "Output/exposure_analysis_data_long_new.csv")
+```
+
+# Convert long data to medium: only keeping days with PM sampling, repeated by child_id
+```{r}
+exposure_analysis_data_long_new <- read_rds("Output/exposure_analysis_data_long_new.rds")
+
+exposure_analysis_data_medium <- exposure_analysis_data_long_new %>% 
+  select(area, home, home_winter_id, winter_id, treatment, 
+         sampling_day, sample_date, burn_level:wood_collect_method_2level) %>% 
+  group_by(area, home, winter_id, child_id_char) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, winter_id, treatment, sampling_day,
+         child_id_char, burn_level:wood_collect_method_2level) %>% 
+  mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
+         filter_var = if_else(is.na(filter_var), 0, filter_var)) %>% 
+  select(filter_var, area:wood_collect_method_2level) %>% 
+  filter(filter_var == 0)
+
+#write_rds(exposure_analysis_data_medium, "Output/exposure_analysis_data_medium.rds")
+#write_csv(exposure_analysis_data_medium, "Output/exposure_analysis_data_medium.csv")
+```
+
+# Convert medium data to short: only keeping 1 observation per sampling period, repeated by child_id
+```{r}
+exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
+
+exposure_analysis_data_short <- exposure_analysis_data_medium %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(area, home, winter_id, child_id_char) %>% 
+  # Sum home activity vars over entire sampling period to make 1 obs per sampling period
+  mutate(home_act_gas = factor(home_act_gas, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_gas = as.character(home_act_gas),
+         home_act_gas = as.numeric(home_act_gas),
+         home_act_gas_sum = sum(home_act_gas),
+         home_act_elect = factor(home_act_elect, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_elect = as.character(home_act_elect),
+         home_act_elect = as.numeric(home_act_elect),
+         home_act_elect_sum = sum(home_act_elect),
+         home_act_propane = factor(home_act_propane, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_propane = as.character(home_act_propane),
+         home_act_propane = as.numeric(home_act_propane),
+         home_act_propane_sum = sum(home_act_propane),
+         home_act_oil = factor(home_act_oil, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_oil = as.character(home_act_oil),
+         home_act_oil = as.numeric(home_act_oil),
+         home_act_oil_sum = sum(home_act_oil),
+         home_act_other = factor(home_act_other, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_other = as.character(home_act_other),
+         home_act_other = as.numeric(home_act_other),
+         home_act_other_sum = sum(home_act_other),
+         home_act_smoking = factor(home_act_smoking, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_smoking = as.character(home_act_smoking),
+         home_act_smoking = as.numeric(home_act_smoking),
+         home_act_smoking_sum = sum(home_act_smoking),
+         home_act_incense = factor(home_act_incense, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_incense = as.character(home_act_incense),
+         home_act_incense = as.numeric(home_act_incense),
+         home_act_incense_sum = sum(home_act_incense),
+         home_act_candle = factor(home_act_candle, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_candle = as.character(home_act_candle),
+         home_act_candle = as.numeric(home_act_candle),
+         home_act_candle_sum = sum(home_act_candle),
+         home_act_kerosene = factor(home_act_kerosene, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_kerosene = as.character(home_act_kerosene),
+         home_act_kerosene = as.numeric(home_act_kerosene),
+         home_act_kerosene_sum = sum(home_act_kerosene),
+         home_act_windows = factor(home_act_windows, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_windows = as.character(home_act_windows),
+         home_act_windows = as.numeric(home_act_windows),
+         home_act_windows_sum = sum(home_act_windows),
+         home_act_door = factor(home_act_door, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_door = as.character(home_act_door),
+         home_act_door = as.numeric(home_act_door),
+         home_act_door_sum = sum(home_act_door),
+         home_act_sweep = factor(home_act_sweep, levels = c("Yes", "No"), labels = c(1, 0)), 
+         home_act_sweep = as.character(home_act_sweep),
+         home_act_sweep = as.numeric(home_act_sweep),
+         home_act_sweep_sum = sum(home_act_sweep)) %>% 
+  group_by(area, home, winter_id) %>% 
+  distinct(child_id_char, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  arrange(area, home, winter_id)
+
+#write_rds(exposure_analysis_data_short, "Output/exposure_analysis_data_short.rds")
+#write_csv(exposure_analysis_data_short, "Output/exposure_analysis_data_short.csv")
+```
+
+# Dataset with PM and SUMs matched/average over short interval
+```{r}
+pm_join <- pm_clean %>% 
+  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
+  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
+  mutate(winter_id = as.character(winter_id)) %>%
+  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  group_by(area, home) %>% 
+  arrange(sample_datetime) %>% 
+  mutate(pm_last_datetime = last(sample_datetime),
+         pm_first_datetime = first_datetime_new,
+         sample_datetime = round_date(sample_datetime, unit = "minute"),
+         pm_rolling_20 = rollmean(pm, 20, fill = "extend")) %>% 
+  ungroup() %>% 
+  select(-home_id, -treatment) %>% 
+  arrange(area, home, sample_datetime)
+
+sums_join <- sums_clean %>% 
+  mutate(sample_datetime = ymd_hms(datetime_sums)) %>% 
+  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  mutate(sample_datetime = round_date(sample_datetime, unit = "minute")) %>% 
+  ungroup() %>% 
+  arrange(area, home, sample_datetime)
+
+pm_sums_joined <- sums_join %>% 
+  full_join(pm_join, by = c("area", "home", "home_winter_id", 
+                            "winter_id", "sample_datetime")) %>% 
+  filter(!is.na(temp_c_sums)) %>% 
+  filter(!is.na(pm_rolling_20))
+
+pm_sums_joined2 <- pm_sums_joined %>% 
+  rename(sample_date = sample_date.x,
+         day_of_week = day_of_week.x) %>% 
+  select(area, home, home_winter_id, winter_id, datetime_sums, 
+         sample_date, temp_c_sums, pm_rolling_20)
+
+exposure_analysis_data_long_new <- read_rds("Output/exposure_analysis_data_long_new.rds")
+
+pm_sums_joined3 <- pm_sums_joined2 %>% 
+  left_join(exposure_analysis_data_long_new, 
+            by = c("area", "home", "home_winter_id", "winter_id", "sample_date")) %>% 
+  select(area, home, home_winter_id, winter_id, treatment, sampling_day,
+         child_id_char, sample_date, datetime_sums, temp_c_sums:wood_collect_method_2level) %>% 
+  ungroup() %>% 
+  arrange(area, home, datetime_sums)
+
+#write_rds(pm_sums_joined3, "Output/sums_rolling_pm_data.rds")
+#write_csv(pm_sums_joined3, "Output/sums_rolling_pm_data.csv")
+```

### 2020-03-26 stove_grade_usage_summary.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 16 March 2020"
+date: "Started 15 Jan 2020, Updated 25 March 2020"
 output: pdf_document 
@@
 demographics_clean <- read_rds("Output/demographics_clean.rds")
-pm_clean <- read_rds("Output/pm_clean.rds")
+#pm_clean <- read_rds("Output/pm_clean.rds")
 ```
@@
   filter(!is.na(stove_grade))
-kable(stove_grade_summary, align = "c")
+#kable(stove_grade_summary, align = "c")
 
@@
 
-```{r}
+```{r, eval=FALSE}
 # Generate random sample of stoves to re-grade and validate grading methods
@@
 
-```{r, include=FALSE}
+```{r, include=FALSE, eval=FALSE}
 ## 18 homes with grades from 2 winters
@@
 ```{r}
-stove_grade_plot <- stove_grades_clean %>% 
+stove_grade_filtered <- stove_grades_clean %>% 
+  arrange(area, home, winter_id) %>% 
+  group_by(home) %>% 
+  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
+  ungroup()
+
+stove_grade_plot <- stove_grade_filtered %>% 
   filter(winter_id == 1) %>% 
@@
 stove_use_data <- read_rds("Output/stove_use_clean.rds") %>% 
-  mutate(wood_cords = as.factor(wood_cords))
+  mutate(wood_cords = as.factor(wood_cords),
+         burn_level = factor(burn_level,
+                             levels = c("No burning", "Light burning",
+                                        "Average burning", "Heavy burning")))
 
-stove_use_plot_function <- function(data, variable) {
+stove_use_plot_function <- function(data, var, title_label, y_label) {
 
   stove_use_plot <- data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  select(area, variable) %>% 
-  rename(plot_var = variable) %>% 
-  #na.exclude() %>% 
+  filter(winter_id == 1) %>% 
+  select(area, var) %>% 
+  rename(plot_var = var) %>% 
+  filter(!is.na(plot_var)) %>% 
   ggplot() +
@@
           panel.grid = element_blank()) +
+    labs(title = title_label) +
+    ylab(y_label) +
     scale_fill_manual(values = jv_palette)
@@
 
-stove_use_plot_function(stove_use_data, "burn_level")
-stove_use_plot_function(stove_use_data, "wood_type")
-stove_use_plot_function(stove_use_data, "wood_collect")
-stove_use_plot_function(stove_use_data, "wood_collect_method")
-stove_use_plot_function(stove_use_data, "wood_cords")
+stove_use_plot_function(stove_use_data, "burn_level", "Burn level", "count")
+stove_use_plot_function(stove_use_data, "wood_type", "Wood type", "count")
+stove_use_plot_function(stove_use_data, "wood_collect", "Wood collection", "count")
+stove_use_plot_function(stove_use_data, "wood_collect_method", "Wood collect method", "count")
+stove_use_plot_function(stove_use_data, "wood_cords", "Wood cords burned", "count")
 ```
 
-```{r}
+```{r, eval=FALSE}
 #### Abstract/manuscript tables ####

### 2020-03-27 child_health_data_cleaning.Rmd
New summary plots
merge health data

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 27 Feb 2020"
+date: "Started 27 Feb 2020, Updated 27 March 2020"
 output: pdf_document 
@@
 
+# Format at-home-log and merge with PM data
+```{r}
+kids_at_home_log <- read_rds("Output/kids_at_home_log.rds") %>% 
+  arrange(area, home_winter_id, at_home_day)
+
+kids_at_home_log_new <- kids_at_home_log %>% 
+  mutate(sampling_day = at_home_day) %>% 
+  group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
+  mutate("00:00:00" = 1,
+         "01:00:00" = 1,
+         "02:00:00" = 1,
+         "03:00:00" = 1,
+         "04:00:00" = 1,
+         "05:00:00" = 1,
+         "06:00:00" = 1,
+         "07:00:00" = as.numeric(at_home_7) - 1,
+         "08:00:00" = as.numeric(at_home_8) - 1,
+         "09:00:00" = as.numeric(at_home_9) - 1,
+         "10:00:00" = as.numeric(at_home_10) - 1,
+         "11:00:00" = as.numeric(at_home_11) - 1,
+         "12:00:00" = as.numeric(at_home_12) - 1,
+         "13:00:00" = as.numeric(at_home_13) - 1,
+         "14:00:00" = as.numeric(at_home_14) - 1,
+         "15:00:00" = as.numeric(at_home_15) - 1,
+         "16:00:00" = as.numeric(at_home_16) - 1,
+         "17:00:00" = as.numeric(at_home_17) - 1,
+         "18:00:00" = as.numeric(at_home_18) - 1,
+         "19:00:00" = as.numeric(at_home_19) - 1,
+         "20:00:00" = as.numeric(at_home_20) - 1,
+         "21:00:00" = as.numeric(at_home_21) - 1,
+         "22:00:00" = as.numeric(at_home_22) - 1,
+         "23:00:00" = 1) %>% 
+  pivot_longer(cols = c("00:00:00":"23:00:00"), names_to = "sample_time",
+                        values_to = "at_home") %>% 
+  mutate("00:00:00" = "home",
+         "01:00:00" = "home",
+         "02:00:00" = "home",
+         "03:00:00" = "home",
+         "04:00:00" = "home",
+         "05:00:00" = "home",
+         "06:00:00" = "home",
+         "07:00:00" = location_7,
+         "08:00:00" = location_8,
+         "09:00:00" = location_9,
+         "10:00:00" = location_10,
+         "11:00:00" = location_11,
+         "12:00:00" = location_12,
+         "13:00:00" = location_13,
+         "14:00:00" = location_14,
+         "15:00:00" = location_15,
+         "16:00:00" = location_16,
+         "17:00:00" = location_17,
+         "18:00:00" = location_18,
+         "19:00:00" = location_19,
+         "20:00:00" = location_20,
+         "21:00:00" = location_21,
+         "22:00:00" = location_22,
+         "23:00:00" = "home") %>% 
+  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
+         sampling_day, sample_time, at_home, "00:00:00":"23:00:00") %>% 
+  group_by(area, home_winter_id, child_id_num, sampling_day, sample_time) %>% 
+  pivot_longer(cols = c("00:00:00":"23:00:00"), names_to = "location_time",
+                        values_to = "at_home_location") %>% 
+  mutate(filter_var = if_else(sample_time == location_time, 1, 0)) %>% 
+  filter(filter_var == 1) %>%  
+  arrange(area, home_winter_id, child_id_num, sampling_day, sample_time) %>% 
+  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
+         sampling_day, sample_time, at_home, at_home_location)  %>% 
+  group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
+  mutate(at_home = if_else(sample_time == "06:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "05:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "04:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "03:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "02:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "01:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "00:00:00", lead(at_home), at_home),
+         at_home = if_else(sample_time == "23:00:00", lag(at_home), at_home),
+         at_home_location = if_else(sample_time == "06:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "05:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "04:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "03:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "02:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "01:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "00:00:00", 
+                                    lead(at_home_location), at_home_location),
+         at_home_location = if_else(sample_time == "23:00:00", 
+                                    lag(at_home_location), at_home_location),
+         at_home_location = if_else(at_home == 1, "home", at_home_location)) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, child_id_num, sampling_day, sample_time)
+
+at_home_pm_merged <- pm_clean %>% 
+  group_by(area, home) %>% 
+  mutate(pm = if_else(pm > 500, 500, pm),
+         pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
+         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
+         pm_obs_sampling_day = n()) %>% 
+  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
+  group_by(area, home, sampling_day, sample_time) %>% 
+  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
+  distinct(sample_time, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  mutate(sampling_day = as.character(sampling_day)) %>% 
+  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
+         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
+         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day) %>% 
+  left_join(kids_at_home_log_new, 
+            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
+                   "sampling_day", "sample_time")) %>% 
+  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
+  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
+  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
+  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
+  distinct(at_home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  filter(!is.na(at_home)) %>% 
+  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)
+
+#write_rds(at_home_pm_merged, "Output/at_home_pm_merged.rds")
+#write_csv(at_home_pm_merged, "Output/at_home_pm_merged.csv")
+
+at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")
+```
+
 ################# Medical Visit ###############
@@
 
+kids_med_visit <- read_rds("Output/kids_med_visit.rds")
+
 save_var_names <- data.frame(t(kids_med_visit)) %>% 
@@
 
+# Combine logs from all areas
+kids_birth_infancy <- wmt_birthinfancy_initial %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id"))
+
+# write_rds(kids_birth_infancy, "Output/kids_birth_infancy.rds")
+
+kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds")
+
 save_var_names <- data.frame(t(wmt_birthinfancy_initial)) %>% 
@@
 # Waiting on files from NN and AK
+
+# Combine logs from all areas
+child_home_winter <- wmt_childhomewinter_initial %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id"))
+
+# write_rds(child_home_winter, "Output/child_home_winter.rds")
+
+child_home_winter <- read_rds("Output/child_home_winter.rds")
+
 ```
@@
 
+# Combine logs from all areas
+kids_csq <- wmt_csq_initial %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id"))
+
+# write_rds(kids_csq, "Output/kids_csq.rds")
+
+kids_csq <- read_rds("Output/kids_csq.rds")
+
 save_var_names <- data.frame(t(wmt_csq_initial)) %>% 
@@
 # Waiting on files from NN and AK
+
+# Combine logs from all areas
+kids_medical_records <- wmt_medicalrecords_initial %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id"))
+
+# write_rds(kids_medical_records, "Output/kids_medical_records.rds")
+
+kids_medical_records <- read_rds("Output/kids_medical_records.rds")
 ```

### 2020-03-27 exposure_models.Rmd
merge health data

@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")
+
+exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
 
 exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")
+
+sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
 ```

### 2020-03-27 merge_datasets.Rmd
merge health data

@@
 ---
-title: "KidsAIR: merge datasets"
+title: "KidsAIR: merge exposure datasets"
 author: "Ethan Walker"
-date: "Started 24 March 2020, Updated 26 March 2020"
+date: "Started 24 March 2020, Updated 27 March 2020"
 output: pdf_document
@@
 
-#kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
 demographics_clean <- read_rds("Output/demographics_clean.rds")
@@
   select(rowname)
-write_csv(save_var_names, "Output/kids_data_var_names.csv")
+
+#write_csv(save_var_names, "Output/kids_data_var_names.csv")
 ```
@@
   select(filter_var, area:wood_collect_method_2level) %>% 
-  filter(filter_var == 0)
+  filter(filter_var == 0) %>% 
+  select(-filter_var)
 

### 2020-03-27 merge_health_datasets.Rmd
merge health data

---
title: "KidsAIR: merge health datasets"
author: "Ethan Walker"
date: "Started 27 March 2020, Updated 27 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
```

# Load individual datasets
```{r}
at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")
kids_med_visit <- read_rds("Output/kids_med_visit.rds")
kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds")
child_home_winter <- read_rds("Output/child_home_winter.rds")
kids_csq <- read_rds("Output/kids_csq.rds")
kids_medical_records <- read_rds("Output/kids_medical_records.rds")
```

# Initial data merge - long format
```{r}
kids_med_visit <- read_rds("Output/kids_med_visit.rds") %>% 
  mutate(visit_date = med_visit_date) %>% 
  rename(wheeze_med_visit = wheeze)

health_merged1 <- kids_csq %>% 
  select(-home_id) %>% 
  mutate(visit_date = csq_date) %>% 
  rename(wheeze_csq = wheeze) %>% 
  full_join(kids_med_visit, by = c("area", "home", "home_winter_id", "treatment",
                                   "winter_id", "child_id_num", "visit_date")) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, csq_date:med_visit_notes) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

child_home_winter <- read_rds("Output/child_home_winter.rds") %>% 
  select(-csq_date, -home_id.x, -home_id.y, -symptom_free_date)

health_merged2 <- health_merged1 %>% 
  left_join(child_home_winter, by = c("area", "home", "home_winter_id", "treatment",
                                      "winter_id", "child_id_num")) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds") %>% 
  select(-home_id)

health_merged3 <- health_merged2 %>% 
  left_join(kids_birth_infancy, by = c("area", "home", "home_winter_id", "treatment",
                                      "winter_id", "child_id_num")) %>% 
  mutate(sample_date = as.character(visit_date)) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  ungroup()

at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")

health_merged4 <- health_merged3 %>% 
  full_join(at_home_pm_merged, by = c("area", "home", "home_winter_id", "winter_id", 
                                      "child_id_num", "sample_date", "treatment")) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         treatment, sample_date, sampling_day, at_home_location, 
         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:at_home) %>% 
  ungroup()

#write_rds(health_merged4, "Output/health_data_merged.rds")
#write_csv(health_merged4, "Output/health_data_merged.csv")
```

# Merge full health data with exposure data
```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")

exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")

sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")

health_data_merged <- read_rds("Output/health_data_merged.rds")

health_exposure_data_sampling_day <- health_data_merged %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  left_join(exposure_analysis_data_medium,
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "child_id_num", "sampling_day"))

#write_rds(health_exposure_data_sampling_day, "Output/health_exposure_data_sampling_day.rds")
#write_csv(health_exposure_data_sampling_day, "Output/health_exposure_data_sampling_day.csv")

health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")

save_var_names <- data.frame(t(health_exposure_data_sampling_day)) %>% 
  rownames_to_column() %>% 
  select(rowname)

#write_csv(save_var_names, "Output/kids_data_var_names.csv")
```

### 2020-03-30 ambient_initial.Rmd
ambient data updates

---
title: "KidsAIR: ambient data"
author: "Ethan Walker"
date: "Started 30 March 2020, Updated 30 March 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(ggmap)
library(maps)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data

wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
  rename_all(tolower)

nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
  rename_all(tolower)

ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
  rename_all(tolower)

us_zipcodes <- read_tsv("Input/us_zipcodes/us_zipcodes.txt", 
                        col_names = c("country", "zip", "city", "state",
                                      "state_abbr", "trash1", "trash2", "trash3", 
                                      "trash4", "lat", "long", "trash5")) %>% 
  select(-trash1:-trash4, -trash5)

wmt_map <- map_data("state") %>% 
  filter(region == "montana" | region == "idaho")

nn_map <- map_data("state") %>% 
  filter(region == "arizona" | region == "utah" |
         region == "new mexico" | region == "colorado")

ak_map <- map_data("state") %>% 
  filter(region == "alaska")

montana_temps <- read_csv("Input/ambient/montana_temp.txt") %>% 
  rename_all(tolower) %>% 
  mutate(station = `stn---`,
         temp_date = yearmoda,
         mean_temp = temp,
         mean_dewp = dewp,
         mean_pressure = stp,
         mean_visib_miles = visib,
         mean_wind_knots = wdsp,
         max_wind_knots = mxspd,
         max_temp = max,
         min_temp = min,
         precip = prcp,
         snow_depth = sndp) %>% 
  select(station:snow_depth)

us_temp_stations <- read_table("Input/ambient/temp_station_list.txt", skip = 20,
                            col_names = c("station", "wban", "station_names",
                                          "country", "state", "call", "lat",
                                          "long", "elevation", "begin_date",
                                          "end_date")) %>% 
  filter(country == "US") %>% 
  mutate(lat = as.numeric(lat),
         long = as.numeric(long),
         elevation = as.numeric(elevation),
         begin_date = ymd(begin_date),
         end_date = ymd(end_date))

pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
```

# Mapping zipcodes, weather stations, pm stations - WMT
```{r}
wmt_zipcodes_new <- wmt_zipcodes %>% 
  left_join(us_zipcodes, by = "zip") %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  mutate(lat_zip = lat,
         long_zip = long)

#Plot temp stations
wmt_temp_stations <- us_temp_stations %>% 
  filter(state == "MT" | state == "ID") %>% 
  mutate(lat_temp = lat,
         long_temp = long) %>% 
  distinct(station, .keep_all = TRUE) %>% 
  full_join(wmt_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_temp),
         long_diff = abs(long_zip - long_temp),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

wmt_plot <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=wmt_temp_stations, 
             aes(long_zip, lat_zip, color = station_names), size = 3) +
    geom_point(data=wmt_temp_stations, 
             aes(long_temp, lat_temp, color = station_names), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
wmt_plot

#Plot PM stations
wmt_pm_stations <- pm_stations %>% 
  filter(area == "wmt") %>% 
  full_join(wmt_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_pm),
         long_diff = abs(long_zip - long_pm),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

wmt_plot <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=wmt_pm_stations, 
             aes(long_zip, lat_zip, color = site), size = 3) +
    geom_point(data=wmt_pm_stations, 
             aes(long_pm, lat_pm, color = site), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
wmt_plot
```

# Mapping zipcodes, weather stations, pm stations - NN
```{r}
nn_zipcodes_new <- nn_zipcodes %>% 
  left_join(us_zipcodes, by = "zip") %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  mutate(lat_zip = lat,
         long_zip = long)

#Plot temp stations
nn_temp_stations <- us_temp_stations %>% 
  filter(state == "AZ" | state == "NM" |
         state == "UT" | state == "CO") %>% 
  mutate(lat_temp = lat,
         long_temp = long) %>% 
  distinct(station, .keep_all = TRUE) %>% 
  full_join(nn_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_temp),
         long_diff = abs(long_zip - long_temp),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

nn_plot_temp_stations <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=nn_temp_stations, 
             aes(long_zip, lat_zip, color = station_names), size = 3) +
    geom_point(data=nn_temp_stations, 
             aes(long_temp, lat_temp, color = station_names), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
nn_plot_temp_stations

#Plot PM stations
nn_pm_stations <- pm_stations %>% 
  filter(area == "nn") %>% 
  full_join(nn_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_pm),
         long_diff = abs(long_zip - long_pm),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

nn_plot <- nn_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=nn_pm_stations, 
             aes(long_zip, lat_zip, color = site), size = 3) +
    geom_point(data=nn_pm_stations, 
             aes(long_pm, lat_pm, color = site), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
nn_plot
```

# Mapping zipcodes, weather stations, pm stations - WMT
```{r}
ak_zipcodes_new <- ak_zipcodes %>% 
  left_join(us_zipcodes, by = "zip") %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  mutate(lat_zip = lat,
         long_zip = long)

#Plot temp stations
ak_temp_stations <- us_temp_stations %>% 
  filter(state == "AK") %>% 
  mutate(lat_temp = lat,
         long_temp = long) %>% 
  distinct(station, .keep_all = TRUE) %>% 
  full_join(ak_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_temp),
         long_diff = abs(long_zip - long_temp),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE)

zipcodes_map <- wmt_zipcodes_new %>% 
  ggplot() +
    geom_point(data=ak_temp_stations, 
             aes(long_zip, lat_zip, color = station_names), size = 3) +
    geom_point(data=ak_temp_stations, 
             aes(long_temp, lat_temp, color = station_names), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette) 
zipcodes_map

#Plot PM stations
ak_pm_stations <- pm_stations %>% 
  filter(area == "ak") %>% 
  full_join(ak_zipcodes_new, by = "country") %>% 
  mutate(lat_diff = abs(lat_zip - lat_pm),
         long_diff = abs(long_zip - long_pm),
         sum_diff = lat_diff + long_diff) %>% 
  arrange(zip, sum_diff) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(home))

ak_plot <- ak_zipcodes_new %>% 
  ggplot() +
    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
               colour="black", fill = "white") +
    geom_point(data=ak_pm_stations, 
             aes(long_zip, lat_zip, color = site), size = 3) +
    geom_point(data=ak_pm_stations, 
             aes(long_pm, lat_pm, color = site), 
             shape = 15, size = 3) +
    theme_minimal() +
    scale_color_manual(values = jv_palette)
ak_plot
```

### 2020-03-30 exposure_initial_analysis.Rmd
ambient data updates

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 26 March 2020"
+date: "Started 2 Feb 2020, Updated 30 March 2020"
 output: pdf_document
@@
 library(lmerTest)
+library(ggmap)
 ```
@@
 sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
+
+health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
 ```
@@
 
-table(analysis_data$temp_perc_25_2level)
+table(analysis_data$temp_perc_30_2level)
 median(analysis_data$temp_perc_30, na.rm = TRUE)
 
-pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_25_2level + 
+pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_30_2level + 
+                              (1 | home), data = analysis_data)
+pm_sums_results <- lmer(pm_mean_sampling_day ~ temp_perc_30_2level + 
                               (1 | home), data = analysis_data)
@@
                       data = analysis_data)
+pm_sums_results <- lm(pm_mean_sampling_period ~ sums_mean_sampling_period_5, 
+                      data = analysis_data)
 summary(pm_sums_results)
@@
                       data = analysis_data)
+pm_sums_model <- lmer(pm_rolling_20 ~ temp_c_sums_5 + (1 | home), 
+                      data = analysis_data)
 
@@
 tidy(pm_sums_model)
-exp((0.146291)-1)*100
+exp((0.1496794)-1)*100
 plot(pm_sums_model, 
@@
 
-# SUMs - events per sampling day - KEEP WORKING ON THIS
+# SUMs - events per sampling day 
 ```{r}
@@
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home, sample_date) %>% 
+  group_by(area, home, sampling_day) %>% 
   distinct(datetime_sums, .keep_all = TRUE) %>% 
-  mutate(time_diff = (lead(datetime_sums) - datetime_sums)/60) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
   mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
@@
   mutate(temp_diff = lead_temp - temp_c_sums,
-         temp_diff_abs = abs(lead_temp - temp_c_sums),
          # change temp in the following line
-         temp_diff_check = if_else(temp_diff >= 8, 1, 0),
+         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
          heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
@@
                               heat_event)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(sums_events = sum(heat_event, na.rm = TRUE),
+         sums_events_cat = cut(sums_events, breaks = c(0, 2, 50),
+                           labels = c("<2", "2+"),
+                                 right = FALSE)) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
   ungroup() %>% 
-  mutate(sample_date = as.character(sample_date)) %>% 
-  group_by(area, home, sample_date) %>% 
-  mutate(sums_events = sum(heat_event, na.rm = TRUE),
-         sums_events_cat = cut(sums_events, breaks = c(0, 5, 50),
-                           labels = c("<4", "4+"),
-                                 right = FALSE)) %>% 
-  ungroup() %>% 
+  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
+         time_diff:sums_events_cat) %>% 
   arrange(area, home, sample_date)
@@
 
-pm_sums_model <- lmer(mean_pm ~ sums_events + (1 | home), data = sums_temp_change)
-pm_sums_model <- lmer(log(pm_mean_daily) ~ sums_events + (1 | home), data = sums_temp_change)
+pm_sums_model <- lmer(pm_mean_sampling_day ~ sums_events_cat + 
+                        (1 | home), data = sums_temp_change)
+pm_sums_model <- lmer(log(pm_mean_sampling_day) ~ sums_events_cat + 
+                        (1 | home), data = sums_temp_change)
 
 summary(pm_sums_model)
+tidy(pm_sums_model)
 #exp((0.146291)-1)*100
@@
 
-pm_sums_plot <- pm_sums_events_joined %>% 
+pm_sums_plot <- sums_temp_change %>% 
   filter(!is.na(sums_events)) %>% 
   ggplot() + 
-    geom_boxplot(aes(sums_events_cat, log(mean_pm))) +
+    geom_boxplot(aes(sums_events_cat, log(pm_mean_sampling_day))) +
     theme_minimal()

### 2020-03-30 exposure_models.Rmd
ambient data updates

@@
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 20 March 2020"
+date: "Started 19 March 2020, Updated 30 March 2020"
 output: pdf_document

### 2020-03-30 merge_health_datasets.Rmd
ambient data updates

@@
 health_exposure_data_sampling_day <- health_data_merged %>% 
+  select(-pm_mean_sampling_day, -pm_mean_sampling_period, day_of_week) %>% 
   mutate(winter_id = as.character(winter_id)) %>% 

### 2020-03-31 ambient_initial.Rmd
ambient data analysis

@@
 
-montana_temps <- read_csv("Input/ambient/montana_temp.txt") %>% 
-  rename_all(tolower) %>% 
-  mutate(station = `stn---`,
-         temp_date = yearmoda,
-         mean_temp = temp,
-         mean_dewp = dewp,
-         mean_pressure = stp,
-         mean_visib_miles = visib,
-         mean_wind_knots = wdsp,
-         max_wind_knots = mxspd,
-         max_temp = max,
-         min_temp = min,
-         precip = prcp,
-         snow_depth = sndp) %>% 
-  select(station:snow_depth)
-
 us_temp_stations <- read_table("Input/ambient/temp_station_list.txt", skip = 20,
@@
          long_temp = long) %>% 
-  distinct(station, .keep_all = TRUE) %>% 
+  filter(begin_date < "2014-06-01") %>% 
+  filter(end_date > "2020-03-21") %>% 
+  #distinct(station, .keep_all = TRUE) %>% 
   full_join(wmt_zipcodes_new, by = "country") %>% 
@@
          long_diff = abs(long_zip - long_temp),
-         sum_diff = lat_diff + long_diff) %>% 
+         sum_diff = lat_diff + long_diff,
+         area = "WMT") %>% 
   arrange(zip, sum_diff) %>% 
   distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home))
+  filter(!is.na(home)) %>% 
+  mutate(station_char = factor(station_names, 
+                               levels = c("BERT MOONEY AIRPORT", "GLACIER PARK INTERNATIONAL AI",
+                                          "GOLF (SIMMS)", "HELENA REGIONAL AIRPORT",
+                                          "LEMHI COUNTY AIRPORT", "MISSOULA INTERNATIONAL AIRPOR",
+                                          "MULLAN PASS"),
+                               labels = c("bertmooneyairport", "glacierparkairport",
+                                          "golfsimms", "helenaairport",
+                                          "lemhicountyairport", "missoulaairport",
+                                          "mullanpass"))) %>% 
+  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
+         long_zip, station_names)
 
@@
              shape = 15, size = 3) +
-    theme_minimal() +
+    theme_void() +
     scale_color_manual(values = jv_palette)
@@
 wmt_pm_stations <- pm_stations %>% 
-  filter(area == "wmt") %>% 
+  filter(area == "WMT") %>% 
   full_join(wmt_zipcodes_new, by = "country") %>% 
@@
   distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home))
+  filter(!is.na(home)) %>% 
+  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
 
@@
     geom_point(data=wmt_pm_stations, 
-             aes(long_zip, lat_zip, color = site), size = 3) +
+             aes(long_zip, lat_zip, color = local_site), size = 3) +
     geom_point(data=wmt_pm_stations, 
-             aes(long_pm, lat_pm, color = site), 
+             aes(long_pm, lat_pm, color = local_site), 
              shape = 15, size = 3) +
-    theme_minimal() +
+    theme_void() +
     scale_color_manual(values = jv_palette)
@@
          long_temp = long) %>% 
-  distinct(station, .keep_all = TRUE) %>% 
+  filter(begin_date < "2014-06-01") %>% 
+  filter(end_date > "2020-03-21") %>% 
+  #distinct(station, .keep_all = TRUE) %>% 
   full_join(nn_zipcodes_new, by = "country") %>% 
@@
          long_diff = abs(long_zip - long_temp),
-         sum_diff = lat_diff + long_diff) %>% 
+         sum_diff = lat_diff + long_diff,
+         area = "NN") %>% 
   arrange(zip, sum_diff) %>% 
   distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home))
+  filter(!is.na(home)) %>% 
+  mutate(station_char = factor(station_names, 
+                               levels = c("FOUR CORNERS REGIONAL ARPT", "GALLUP MUNICIPAL AIRPORT",
+                                          "PAGE MUNICIPAL AIRPORT", "WINDOW ROCK AIRPORT",
+                                          "WINSLOW-LINDBERGH REGIONAL AI"),
+                               labels = c("fourcornersairport", "gallopairport",
+                                          "pageairport", "windowrockairport",
+                                          "winslowlindberghairport"))) %>% 
+  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
+         long_zip, station_names)
 
@@
 nn_pm_stations <- pm_stations %>% 
-  filter(area == "nn") %>% 
+  filter(area == "NN") %>% 
   full_join(nn_zipcodes_new, by = "country") %>% 
@@
     geom_point(data=nn_pm_stations, 
-             aes(long_zip, lat_zip, color = site), size = 3) +
+             aes(long_zip, lat_zip, color = local_site), size = 3) +
     geom_point(data=nn_pm_stations, 
-             aes(long_pm, lat_pm, color = site), 
+             aes(long_pm, lat_pm, color = local_site), 
              shape = 15, size = 3) +
-    theme_minimal() +
+    theme_void() +
     scale_color_manual(values = jv_palette)
@@
          long_temp = long) %>% 
-  distinct(station, .keep_all = TRUE) %>% 
+  filter(begin_date < "2014-06-01") %>% 
+  filter(end_date > "2020-03-21") %>% 
+  #distinct(station, .keep_all = TRUE) %>% 
   full_join(ak_zipcodes_new, by = "country") %>% 
@@
          long_diff = abs(long_zip - long_temp),
-         sum_diff = lat_diff + long_diff) %>% 
+         sum_diff = lat_diff + long_diff,
+         area = "AK") %>% 
   arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE)
+  distinct(zip, .keep_all = TRUE) %>% 
+  mutate(station_char = factor(station_names, 
+                               levels = c("KIPNUK AIRPORT", "BETHEL AIRPORT",
+                                          "QUINHAGAK AIRPORT"),
+                               labels = c("kipnukairport", "bethelairport",
+                                          "quinhagakairport"))) %>% 
+  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
+         long_zip, station_names)
 
@@
              shape = 15, size = 3) +
-    theme_minimal() +
+    theme_void() +
     scale_color_manual(values = jv_palette) 
@@
 ak_pm_stations <- pm_stations %>% 
-  filter(area == "ak") %>% 
+  filter(area == "AK") %>% 
   full_join(ak_zipcodes_new, by = "country") %>% 
@@
     geom_point(data=ak_pm_stations, 
-             aes(long_zip, lat_zip, color = site), size = 3) +
+             aes(long_zip, lat_zip, color = local_site), size = 3) +
     geom_point(data=ak_pm_stations, 
-             aes(long_pm, lat_pm, color = site), 
+             aes(long_pm, lat_pm, color = local_site), 
              shape = 15, size = 3) +
-    theme_minimal() +
+    theme_void() +
     scale_color_manual(values = jv_palette)
@@
 ```
+
+# Compile temperature data and filter for closest stations
+```{r}
+list_files <- list.files("Input/ambient/temp_data")
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd("Input/ambient/temp_data") # set new working directory
+initial_data = tibble(file_name = list_files) %>% # use list of files from above
+  mutate(save_data = lapply(file_name, read_csv, 
+                            col_names = c("station", "wban", "temp_date",
+                                          "mean_temp", "t1", "mean_dewp", "t2",
+                                          "t3", "t4", "mean_pressure", "t5",
+                                          "mean_visib_miles", "t6", "mean_wind_knots",
+                                          "t7", "max_wind_knots", "t8", "max_temp",
+                                          "min_temp", "precip", "snow_depth", "t9"))) %>%
+  unnest(save_data)
+
+# Format combined file from above; combine with other PM data below
+temp_data <- initial_data %>% 
+  filter(!is.na(t1)) %>% 
+  separate(file_name, c("area", "station_char"), sep = "_") %>%
+  mutate(temp_date = ymd(temp_date),
+         mean_temp = as.numeric(mean_temp),
+         mean_dewp = as.numeric(mean_dewp),
+         mean_pressure = as.numeric(mean_pressure),
+         mean_visib_miles = as.numeric(mean_visib_miles),
+         mean_wind_knots = as.numeric(mean_wind_knots),
+         max_wind_knots = as.numeric(max_wind_knots)) %>% 
+  select(station_char, station, temp_date, mean_temp, mean_dewp,
+         mean_pressure, mean_visib_miles, mean_wind_knots, max_wind_knots) %>% 
+  replace_with_na_all(condition = ~.x %in% c(9999.9, 999.9))
+
+temp_stations <- rbind(wmt_temp_stations, nn_temp_stations, ak_temp_stations)
+
+temp_data_joined <- temp_data %>% 
+  left_join(temp_stations, by = c("station", "station_char"))
+
+#write_rds(temp_data_joined, "Output/kids_temp_data.rds")
+#write_csv(temp_data_joined, "Output/kids_temp_data.csv")
+
+kids_temp_data <- read_rds("Output/kids_temp_data.rds")
+```
+
+# Compile PM data and filter for correct dates/variables
+```{r}
+pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
+
+list_files <- list.files("Input/ambient/pm_data")
+# Set working directory and load files in list; extract file name and add as column
+## run next 4 lines together
+setwd("Input/ambient/pm_data") # set new working directory
+initial_data = tibble(file_name = list_files) %>% # use list of files from above
+   mutate(save_data = lapply(file_name, read_csv, skip = 1,
+                            col_names = c("t1", "t2", "t3", "t4", "parameter",
+                                          "pm_lat", "pm_long", "t5", "parameter_name",
+                                          "duration", "pollutant", "amb_pm_date",
+                                          "year", "t6", "units", "t7", "t8", "t9",
+                                          "t10", "t11", "mean_pollution", "t12",
+                                          "t13", "t14", "aqi", "t15", "t16",
+                                          "t17", "t18", "t19", "local_site",
+                                          "t20", "t21", "t22"),
+                            col_types = list(col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_double(), 
+                                          col_double(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(),
+                                          col_character(), col_character(), 
+                                          col_double(), col_character(),
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character(), 
+                                          col_character(), col_character()))) %>%
+  unnest(save_data)
+
+pm_data <- initial_data %>% 
+  filter(t1 != "State Code") %>% 
+  filter(pollutant == "PM25 24-hour 2012") %>% 
+  select(local_site, pollutant, amb_pm_date, mean_pollution) %>% 
+  mutate(amb_pm_24hr = as.numeric(mean_pollution),
+         date1 = ymd(amb_pm_date),
+         date2 = mdy(amb_pm_date)) %>% 
+  unite(amb_pm_date, c("date1", "date2"), sep = "") %>% 
+  separate(amb_pm_date, c("trash", "amb_pm_date"), sep = "NA") %>% 
+  unite(amb_pm_date, c("trash", "amb_pm_date"), sep = "") %>% 
+  group_by(local_site, amb_pm_date) %>% 
+  distinct(pollutant, .keep_all = TRUE) %>% 
+  select(local_site, amb_pm_date, amb_pm_24hr)
+  arrange(local_site, amb_pm_date)
+
+pm_stations <- rbind(wmt_pm_stations, nn_pm_stations, ak_pm_stations)
+
+pm_data_joined <- pm_data %>% 
+  left_join(pm_stations, by = "local_site")
+
+#write_rds(pm_data_joined, "Output/kids_amb_pm_data.rds")
+#write_csv(pm_data_joined, "Output/kids_amb_pm_data.csv")
+
+kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")
+```
+
+# Add new variables to pm/temp datasets
+```{r}
+kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")
+kids_temp_data <- read_rds("Output/kids_temp_data.rds")
+
+# PM data
+kids_amb_pm_data_new <- kids_amb_pm_data %>% 
+  arrange(area, local_site, amb_pm_date, zip) %>% 
+  group_by(area, local_site, zip) %>% 
+  mutate(amb_pm_24hr_lag1 = lag(amb_pm_24hr),
+         amb_pm_24hr_lag2 = lag(amb_pm_24hr, 2),
+         amb_pm_24hr_lag3 = lag(amb_pm_24hr, 3),
+         amb_pm_roll_mean_2day = rollmean(amb_pm_24hr, 2, fill = NA, align = "right"),
+         amb_pm_roll_mean_3day = rollmean(amb_pm_24hr, 3, fill = NA, align = "right"),
+         amb_pm_roll_mean_4day = rollmean(amb_pm_24hr, 4, fill = NA, align = "right")) %>% 
+  ungroup()
+
+#write_rds(kids_amb_pm_data_new, "Output/kids_amb_pm_data_new.rds")
+#write_csv(kids_amb_pm_data_new, "Output/kids_amb_pm_data_new.csv")
+
+# Temp data
+kids_amb_temp_data_new <- kids_temp_data %>% 
+  arrange(area, station, temp_date, zip) %>% 
+  group_by(area, station, zip) %>% 
+  mutate(mean_temp_lag1 = lag(mean_temp),
+         mean_temp_lag2 = lag(mean_temp, 2),
+         mean_temp_lag3 = lag(mean_temp, 3),
+         mean_temp_roll_2day = rollmean(mean_temp, 2, fill = NA, align = "right"),
+         mean_temp_roll_3day = rollmean(mean_temp, 3, fill = NA, align = "right"),
+         mean_temp_roll_4day = rollmean(mean_temp, 4, fill = NA, align = "right")) %>% 
+  ungroup()
+
+#write_rds(kids_amb_temp_data_new, "Output/kids_amb_temp_data_new.rds")
+#write_csv(kids_amb_temp_data_new, "Output/kids_amb_temp_data_new.csv")
+```

### 2020-03-31 merge_datasets.Rmd
ambient data analysis

@@
 exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
+  mutate(sample_date = amb_pm_date)
+kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
+  mutate(sample_date = temp_date) %>% 
+  mutate(sample_date = as.character(sample_date)) %>% 
+  select(-lat_zip, -long_zip)
+wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
+  rename_all(tolower)
+nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
+  rename_all(tolower)
+ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
+  rename_all(tolower)
+kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
+  mutate(home = gsub(home, pattern = "_", replacement = ""))
 
@@
   ungroup() %>% 
+  left_join(kids_zipcodes, by = "home") %>% 
+  left_join(kids_amb_pm_data, by = c("area", "zip", "sample_date")) %>% 
+  left_join(kids_amb_temp_data, by = c("area", "zip", "sample_date")) %>%
   arrange(area, home, home_winter_id, sample_date)
@@
   select(area, home, home_winter_id, winter_id, treatment, 
-         sampling_day, sample_date, burn_level:wood_collect_method_2level) %>% 
+         sampling_day, sample_date, burn_level:mean_temp_roll_4day) %>% 
   group_by(area, home, winter_id, child_id_char) %>% 
@@
   select(area, home, home_winter_id, winter_id, treatment, sampling_day,
-         child_id_char, burn_level:wood_collect_method_2level) %>% 
+         child_id_char, burn_level:mean_temp_roll_4day) %>% 
   mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
          filter_var = if_else(is.na(filter_var), 0, filter_var)) %>% 
-  select(filter_var, area:wood_collect_method_2level) %>% 
+  select(filter_var, area:mean_temp_roll_4day) %>% 
   filter(filter_var == 0) %>% 

### 2020-04-01 ambient_initial.Rmd
ambient analysis updates

@@
 library(maps)
+library(tigris)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 
+# Zipcode data from: http://download.geonames.org/export/dump/ 
 us_zipcodes <- read_tsv("Input/us_zipcodes/us_zipcodes.txt", 
@@
 
-ak_map <- map_data("state") %>% 
-  filter(region == "alaska")
+ak_map <- map_data("world") %>% 
+  filter(subregion == "Alaska") %>% 
+  filter(long < 0)
 
+# Weather data and stations comes from the National Climatic Data Center:
+# https://www7.ncdc.noaa.gov/CDO/cdoselect.cmd?datasetabbv=GSOD&countryabbv=&georegionabbv=
 us_temp_stations <- read_table("Input/ambient/temp_station_list.txt", skip = 20,
@@
 
+# Ambient PM data comes from EPA Air Quality Monitors:
+# https://epa.maps.arcgis.com/apps/webappviewer/index.html?id=5f239fd3e72f424f98ef3d5def547eb5&extent=-146.2334,13.1913,-46.3896,56.5319 
 pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
@@
 # Mapping zipcodes, weather stations, pm stations - WMT
+# These steps are repeated for NN and AK below
 ```{r}
+# Zipcodes for each KidsAIR study home were downloaded from the server
+# If there was no zipcode provided, the zipcode from the nearest town was used
+# Here, WMT zipcodes are joined with the US zipcode data
+# This step gives lat/long coordinates for each WMT zipcode
 wmt_zipcodes_new <- wmt_zipcodes %>% 
@@
 
-#Plot temp stations
+# US weather stations list was filtered for MT and ID locations
+# Stations were further filtered for data availability during study dates
+# Temp data was joined with WMT zipcode data
+# Lat/long differences were calculated
+# The weather station closest to each zipcode was selected and plotted below
 wmt_temp_stations <- us_temp_stations %>% 
@@
 
+# Here, zipcodes and weather stations are plotted on a map
+# Colors match for weather stations and zips that are closest together
 wmt_plot <- wmt_zipcodes_new %>% 
@@
     theme_void() +
+    labs(title = "WMT zipcodes and weather stations") +
     scale_color_manual(values = jv_palette)
@@
 
-#Plot PM stations
+# Lat/long coordinates for PM stations was manually extracted from the EPA site
+# Similar to weather stations, each WMT zipcode was matched with the closest PM monitor
 wmt_pm_stations <- pm_stations %>% 
@@
 
+# PM monitors and zipcodes are plotted below
+# Colors match for PM monitors and zips that are closest together
 wmt_plot <- wmt_zipcodes_new %>% 
@@
     theme_void() +
+    labs(title = "WMT zipcodes and PM monitors") +
     scale_color_manual(values = jv_palette)
@@
     theme_void() +
+    labs(title = "NN zipcodes and weather stations") +
     scale_color_manual(values = jv_palette)
@@
     theme_void() +
+    labs(title = "NN zipcodes and PM monitors") +
     scale_color_manual(values = jv_palette)
@@
   ggplot() +
+    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
+               colour="black", fill = "white") +
     geom_point(data=ak_temp_stations, 
@@
     theme_void() +
+    labs(title = "AK zipcodes and weather stations") +
     scale_color_manual(values = jv_palette) 
@@
     theme_void() +
+    labs(title = "AK zipcodes and PM monitors") +
     scale_color_manual(values = jv_palette)
@@
 ```{r}
+# Read in ambient weather files and bind together
 list_files <- list.files("Input/ambient/temp_data")
 # Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
+## run next 10 lines together
 setwd("Input/ambient/temp_data") # set new working directory
@@
 
-# Format combined file from above; combine with other PM data below
+# Format combined file from above
 temp_data <- initial_data %>% 
@@
 
+# Bind together weather station names from each location
 temp_stations <- rbind(wmt_temp_stations, nn_temp_stations, ak_temp_stations)
 
+# Join weather data with station names
 temp_data_joined <- temp_data %>% 
@@
 
+# Read in ambient PM files and bind together
 list_files <- list.files("Input/ambient/pm_data")
 # Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
+## run next 28 lines together
 setwd("Input/ambient/pm_data") # set new working directory
@@
 
+# Format combined file from above
 pm_data <- initial_data %>% 
@@
 
+# Bind together PM monitor names from each location
 pm_stations <- rbind(wmt_pm_stations, nn_pm_stations, ak_pm_stations)
 
+# Join PM data with monitor names
 pm_data_joined <- pm_data %>% 

### 2020-04-01 ambient_summary.Rmd
ambient analysis updates

---
title: "KidsAIR: ambient temperature and PM analysis"
author: "Ethan Walker"
date: "Started 1 April 2020, Updated 1 April 2020"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(ggmap)
library(maps)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")

exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")

exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")

sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")

health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
```

```{r, eval=FALSE, include=FALSE}
# data from each sampling day
pm_summary <- exposure_analysis_data_long %>% 
  group_by(area) %>% 
  summarize(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE))
pm_summary

temp_summary <- exposure_analysis_data_long %>% 
  group_by(area) %>% 
  summarize(mean_amb_temp = mean(mean_temp, na.rm = TRUE))
temp_summary

# data from indoor PM sampling
pm_summary <- exposure_analysis_data_medium %>% 
  group_by(area) %>% 
  summarize(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE))
pm_summary

temp_summary <- exposure_analysis_data_medium %>% 
  group_by(area) %>% 
  summarize(mean_amb_temp = mean(mean_temp, na.rm = TRUE))
temp_summary
```

```{r}
boxplots_amb_pm <- exposure_analysis_data_medium %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, amb_pm_24hr)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    labs(y = expression(paste("Ambient PM "[2.5], " (", mu, g/m^3, ")"))) +
    labs(title = "Ambient PM levels during indoor PM sampling",
          subtitle = "Data from closest EPA monitor to each home's zipcode") +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          title = element_text(size = 16, colour = "black"))
boxplots_amb_pm
```

```{r}
boxplots_amb_temp <- exposure_analysis_data_medium %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, mean_temp)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    labs(y = "Degrees (F)") +
    labs(title = "Ambient mean temperatures during indoor PM sampling",
          subtitle = "Data from closest site to each home's zipcode") +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          title = element_text(size = 16, colour = "black"))
boxplots_amb_temp
```

```{r}
wmt_map <- map_data("state") %>% 
  filter(region == "montana")

nn_map <- map_data("state") %>% 
  filter(region == "arizona")

ak_map <- map_data("world") %>% 
  filter(subregion == "Alaska") %>% 
  filter(long < 0)

wmt_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "WMT") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "tan") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
               shape = 21, color = "black") +
    geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    theme_void() +
    labs(title = "WMT mean ambient PM at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = EPA monitors") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 8)
wmt_plot
```

```{r}
nn_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "NN") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "tan") +
    geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "NN mean ambient PM at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = EPA monitors") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 1.5)
nn_plot
```

```{r}
ak_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "AK") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_pm = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "tan") +
    geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "AK mean ambient PM at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = EPA monitors") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 6)
ak_plot
```

```{r}
wmt_map <- map_data("state") %>% 
  filter(region == "montana")

nn_map <- map_data("state") %>% 
  filter(region == "arizona")

ak_map <- map_data("world") %>% 
  filter(subregion == "Alaska") %>% 
  filter(long < 0)

wmt_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "WMT") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_temp = mean(mean_temp, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "tan") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
               shape = 21, color = "black") +
    geom_point(aes(long_temp, lat_temp), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    theme_void() +
    labs(title = "WMT mean ambient Temp at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = weather stations") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 20)
wmt_plot
```

```{r}
nn_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "NN") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_temp = mean(mean_temp, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "tan") +
    geom_point(aes(long_temp, lat_temp), shape = 24, size = 3, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "NN mean ambient Temp at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = weather stations") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 45)
nn_plot
```

```{r}
ak_plot <- exposure_analysis_data_medium %>%
  ungroup() %>% 
  filter(area == "AK") %>% 
  group_by(zip) %>% 
  mutate(mean_amb_temp = mean(mean_temp, na.rm = TRUE)) %>% 
  distinct(zip, .keep_all = TRUE) %>% 
  filter(!is.na(zip)) %>% 
  ggplot() +
    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
               color="black", fill = "tan") +
    geom_point(aes(long_temp, lat_temp), shape = 24, size = 5, fill = "yellow",
               color = "black") +
    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
               shape = 21, color = "black") +
    theme_void() +
    labs(title = "AK mean ambient Temp at each zipcode during indoor PM sampling",
         subtitle = "Yellow triangles = weather stations") +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 30)
ak_plot
```

### 2020-04-01 merge_datasets.Rmd
ambient analysis updates

@@
 kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
-  mutate(sample_date = amb_pm_date)
+  mutate(sample_date = amb_pm_date) %>% 
+  distinct(zip, .keep_all = TRUE)
 kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
@@
   mutate(sample_date = as.character(sample_date)) %>% 
-  select(-lat_zip, -long_zip)
+  distinct(zip, .keep_all = TRUE)
 wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
@@
 kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
-  mutate(home = gsub(home, pattern = "_", replacement = ""))
+  mutate(home = gsub(home, pattern = "_", replacement = ""),
+         home = gsub(home, pattern = " ", replacement = "")) %>% 
+  left_join(kids_amb_pm_data, by = "zip") %>% 
+  select(home, zip, lat_pm, long_pm) %>% 
+  left_join(kids_amb_temp_data, by = "zip") %>% 
+  select(home, zip, lat_zip, long_zip, lat_pm, long_pm, lat_temp, long_temp)
+kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
+  mutate(sample_date = amb_pm_date) %>% 
+  select(-lat_zip, -long_zip, -lat_pm, -long_pm)
+kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
+  mutate(sample_date = temp_date) %>% 
+  mutate(sample_date = as.character(sample_date)) %>% 
+  select(-lat_zip, -long_zip, -lat_temp, -long_temp)
 
@@
   filter(filter_var == 0) %>% 
-  select(-filter_var)
+  select(-filter_var) %>% 
+  ungroup()
 

### 2020-04-03 ambient_initial.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 30 March 2020, Updated 30 March 2020"
+date: "Started 30 March 2020, Updated 3 April 2020"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
+                      include = TRUE, warning = FALSE, message = FALSE,
+                      fig.height = 6, fig.width = 8)
 ```
@@
                   "#FF3366", "#33CC99", "#33999")
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+               "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 
+```{r}
 # Mapping zipcodes, weather stations, pm stations - WMT
 # These steps are repeated for NN and AK below
-```{r}
+
 # Zipcodes for each KidsAIR study home were downloaded from the server
@@
     geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
+               colour="black", fill = "azure", size = 1.2) +
     geom_point(data=wmt_temp_stations, 
@@
     theme_void() +
-    labs(title = "WMT zipcodes and weather stations") +
-    scale_color_manual(values = jv_palette)
+    labs(title = "WMT zipcodes and weather stations", color = "Weather station") +
+    scale_color_manual(values = cbPalette)
 wmt_plot
@@
     geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
+               colour="black", fill = "azure", size = 1.2) +
     geom_point(data=wmt_pm_stations, 
@@
     theme_void() +
-    labs(title = "WMT zipcodes and PM monitors") +
-    scale_color_manual(values = jv_palette)
+    labs(title = "WMT zipcodes and PM monitors", color = "EPA Monitor") +
+    scale_color_manual(values = cbPalette)
 wmt_plot
@@
 
+```{r}
 # Mapping zipcodes, weather stations, pm stations - NN
-```{r}
+
 nn_zipcodes_new <- nn_zipcodes %>% 
@@
     geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
+               colour="black", fill = "azure", size = 1.2) +
     geom_point(data=nn_temp_stations, 
@@
     theme_void() +
-    labs(title = "NN zipcodes and weather stations") +
-    scale_color_manual(values = jv_palette)
+    labs(title = "NN zipcodes and weather stations", color = "Weather station") +
+    scale_color_manual(values = cbPalette)
 nn_plot_temp_stations
@@
     geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
+               colour="black", fill = "azure", size = 1.2) +
     geom_point(data=nn_pm_stations, 
@@
     theme_void() +
-    labs(title = "NN zipcodes and PM monitors") +
-    scale_color_manual(values = jv_palette)
+    labs(title = "NN zipcodes and PM monitors", color = "EPA monitor") +
+    scale_color_manual(values = cbPalette)
 nn_plot
@@
 
+```{r}
 # Mapping zipcodes, weather stations, pm stations - WMT
-```{r}
+
 ak_zipcodes_new <- ak_zipcodes %>% 
@@
     geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
+               colour="black", fill = "azure", size = 1.2) +
     geom_point(data=ak_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 3) +
+             aes(long_zip, lat_zip, color = station_names), size = 5) +
     geom_point(data=ak_temp_stations, 
              aes(long_temp, lat_temp, color = station_names), 
-             shape = 15, size = 3) +
+             shape = 15, size = 5) +
     theme_void() +
-    labs(title = "AK zipcodes and weather stations") +
-    scale_color_manual(values = jv_palette) 
+    labs(title = "AK zipcodes and weather stations", color = "Weather station") +
+    scale_color_manual(values = cbPalette) 
 zipcodes_map
@@
     geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
+               colour="black", fill = "azure", size = 1.2) +
     geom_point(data=ak_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 3) +
+             aes(long_zip, lat_zip, color = local_site), size = 5) +
     geom_point(data=ak_pm_stations, 
              aes(long_pm, lat_pm, color = local_site), 
-             shape = 15, size = 3) +
+             shape = 15, size = 5) +
     theme_void() +
-    labs(title = "AK zipcodes and PM monitors") +
-    scale_color_manual(values = jv_palette)
+    labs(title = "AK zipcodes and PM monitors", color = "EPA monitor") +
+    scale_color_manual(values = cbPalette)
 ak_plot
@@
 
+```{r, echo=FALSE, include=FALSE, eval=FALSE}
 # Compile temperature data and filter for closest stations
-```{r}
+
 # Read in ambient weather files and bind together
@@
 
+```{r, echo=FALSE, include=FALSE, eval=FALSE}
 # Compile PM data and filter for correct dates/variables
-```{r}
+
 pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
@@
 
+```{r, echo=FALSE, include=FALSE, eval=FALSE}
 # Add new variables to pm/temp datasets
-```{r}
+
 kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")

### 2020-04-03 ambient_summary.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 1 April 2020, Updated 1 April 2020"
-output: powerpoint_presentation
+date: "Started 1 April 2020, Updated 3 April 2020"
+output: pdf_document
 ---
@@
     geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
-               color="black", fill = "tan") +
+               color="black", fill = "azure", size = 1.2) +
     geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
@@
     labs(title = "WMT mean ambient PM at each zipcode during indoor PM sampling",
-         subtitle = "Yellow triangles = EPA monitors") +
+         subtitle = "Yellow triangles = EPA monitors",
+         fill = "Ambient PM", size = "Ambient PM") +
     scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 8)
@@
     geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
-               color="black", fill = "tan") +
+               color="black", fill = "azure", size = 1.2) +
     geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
@@
     labs(title = "NN mean ambient PM at each zipcode during indoor PM sampling",
-         subtitle = "Yellow triangles = EPA monitors") +
+         subtitle = "Yellow triangles = EPA monitors",
+         fill = "Ambient PM", size = "Ambient PM") +
     scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 1.5)
@@
     geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
-               color="black", fill = "tan") +
-    geom_point(aes(long_pm, lat_pm), shape = 24, size = 3, fill = "yellow",
+               color="black", fill = "azure", size = 1.2) +
+    geom_point(aes(long_pm, lat_pm), shape = 24, size = 5, fill = "yellow",
                color = "black") +
@@
     labs(title = "AK mean ambient PM at each zipcode during indoor PM sampling",
-         subtitle = "Yellow triangles = EPA monitors") +
+         subtitle = "Yellow triangle = EPA monitor",
+         fill = "Ambient PM", size = "Ambient PM") +
     scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 6)
@@
     geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
-               color="black", fill = "tan") +
+               color="black", fill = "azure", size = 1.2) +
     geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
@@
     labs(title = "WMT mean ambient Temp at each zipcode during indoor PM sampling",
-         subtitle = "Yellow triangles = weather stations") +
+         subtitle = "Yellow triangles = weather stations",
+         fill = "Ambient temp", size = "Ambient temp") +
     scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 20)
@@
     geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
-               color="black", fill = "tan") +
+               color="black", fill = "azure", size = 1.2) +
     geom_point(aes(long_temp, lat_temp), shape = 24, size = 3, fill = "yellow",
@@
     labs(title = "NN mean ambient Temp at each zipcode during indoor PM sampling",
-         subtitle = "Yellow triangles = weather stations") +
+         subtitle = "Yellow triangles = weather stations",
+         fill = "Ambient temp", size = "Ambient temp") +
     scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 45)
@@
     geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
-               color="black", fill = "tan") +
+               color="black", fill = "azure", size = 1.2) +
     geom_point(aes(long_temp, lat_temp), shape = 24, size = 5, fill = "yellow",
@@
     labs(title = "AK mean ambient Temp at each zipcode during indoor PM sampling",
-         subtitle = "Yellow triangles = weather stations") +
+         subtitle = "Yellow triangles = weather stations",
+         fill = "Ambient temp", size = "Ambient temp") +
     scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 30)

### 2020-04-03 exposure_initial_analysis.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 30 March 2020"
+date: "Started 2 Feb 2020, Updated 3 April 2020"
 output: pdf_document
@@
 ```
+
+# Ambient temp/PM - by sampling date
+```{r}
+analysis_data <- exposure_analysis_data_long %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(area == "AK") %>% 
+  filter(!is.na(sampling_day)) %>% 
+  filter(!is.na(sample_date)) %>%
+  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
+         zip:mean_temp_roll_4day) %>% 
+  mutate(amb_pm_5 = amb_pm_24hr/5,
+         mean_temp_5 = mean_temp/5,
+         mean_temp_cat = cut(mean_temp, breaks = c(0, 30, 150),
+                           labels = c("<30", "30+"),
+                                 right = FALSE),
+         amb_pm_cat = cut(amb_pm_24hr, breaks = c(0, 7, 100),
+                           labels = c("<7", "7+"),
+                                 right = FALSE))
+
+median(analysis_data$amb_pm_24hr, na.rm = TRUE)
+table(analysis_data$amb_pm_cat)
+
+# Variables to use:
+## amb_pm_5, amb_pm_cat,
+## amb_pm_24hr, amb_pm_24hr_lag1, amb_pm_24hr_lag2, amb_pm_24hr_lag3
+## amb_pm_roll_mean_2day, amb_pm_roll_mean_3day, amb_pm_roll_mean_4day
+## mean_temp_5, mean_temp_cat,
+## mean_temp, mean_temp_lag1, mean_temp_lag2, mean_temp_lag3
+## mean_temp_roll_2day, mean_temp_roll_3day, mean_temp_roll_4day
+
+cor.test(analysis_data$mean_temp, analysis_data$pm_mean_daily, method = "spearman")
+pm_ambient_model <- lmer(log(pm_mean_daily) ~ amb_pm_cat + (1 | home), 
+                      data = analysis_data)
+pm_ambient_model <- lmer(pm_mean_daily ~ amb_pm_cat + (1 | home), 
+                      data = analysis_data)
+
+#summary(pm_ambient_model)
+tidy(pm_ambient_model)
+exp((-0.06723228)-1)*100
+plot(pm_ambient_model, 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(pm_ambient_model), main = "QQ Plot")
+
+pm_sums_plot <- analysis_data %>% 
+  ggplot() + 
+    geom_point(aes(log(pm_mean_daily), mean_temp)) +
+    geom_smooth(aes(log(pm_mean_daily), mean_temp)) +
+    theme_minimal()
+pm_sums_plot
+```

### 2020-04-03 wmt_addresses.Rmd
analysis updates

--- 2020-03-31 ambient_initial.Rmd
@@
 ---
-title: "KidsAIR: ambient data"
+title: "KidsAIR: WMT addresses"
 author: "Ethan Walker"
-date: "Started 30 March 2020, Updated 30 March 2020"
+date: "Started 3 April 2020, Updated 3 April 2020"
 output: pdf_document
@@
 ```{r}
-# Load data
+wmt_addresses <- read_xlsx("Input/WMT/wmt_addresses.xlsx")
 
-wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
-  rename_all(tolower)
-
-nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
-  rename_all(tolower)
-
-ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
-  rename_all(tolower)
-
-us_zipcodes <- read_tsv("Input/us_zipcodes/us_zipcodes.txt", 
-                        col_names = c("country", "zip", "city", "state",
-                                      "state_abbr", "trash1", "trash2", "trash3", 
-                                      "trash4", "lat", "long", "trash5")) %>% 
-  select(-trash1:-trash4, -trash5)
-
-wmt_map <- map_data("state") %>% 
-  filter(region == "montana" | region == "idaho")
-
-nn_map <- map_data("state") %>% 
-  filter(region == "arizona" | region == "utah" |
-         region == "new mexico" | region == "colorado")
-
-ak_map <- map_data("state") %>% 
-  filter(region == "alaska")
-
-us_temp_stations <- read_table("Input/ambient/temp_station_list.txt", skip = 20,
-                            col_names = c("station", "wban", "station_names",
-                                          "country", "state", "call", "lat",
-                                          "long", "elevation", "begin_date",
-                                          "end_date")) %>% 
-  filter(country == "US") %>% 
-  mutate(lat = as.numeric(lat),
-         long = as.numeric(long),
-         elevation = as.numeric(elevation),
-         begin_date = ymd(begin_date),
-         end_date = ymd(end_date))
-
-pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
+register_google(key = "")
 ```
 
-# Mapping zipcodes, weather stations, pm stations - WMT
 ```{r}
-wmt_zipcodes_new <- wmt_zipcodes %>% 
-  left_join(us_zipcodes, by = "zip") %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  mutate(lat_zip = lat,
-         long_zip = long)
-
-#Plot temp stations
-wmt_temp_stations <- us_temp_stations %>% 
-  filter(state == "MT" | state == "ID") %>% 
-  mutate(lat_temp = lat,
-         long_temp = long) %>% 
-  filter(begin_date < "2014-06-01") %>% 
-  filter(end_date > "2020-03-21") %>% 
-  #distinct(station, .keep_all = TRUE) %>% 
-  full_join(wmt_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_temp),
-         long_diff = abs(long_zip - long_temp),
-         sum_diff = lat_diff + long_diff,
-         area = "WMT") %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
-  mutate(station_char = factor(station_names, 
-                               levels = c("BERT MOONEY AIRPORT", "GLACIER PARK INTERNATIONAL AI",
-                                          "GOLF (SIMMS)", "HELENA REGIONAL AIRPORT",
-                                          "LEMHI COUNTY AIRPORT", "MISSOULA INTERNATIONAL AIRPOR",
-                                          "MULLAN PASS"),
-                               labels = c("bertmooneyairport", "glacierparkairport",
-                                          "golfsimms", "helenaairport",
-                                          "lemhicountyairport", "missoulaairport",
-                                          "mullanpass"))) %>% 
-  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
-         long_zip, station_names)
-
-wmt_plot <- wmt_zipcodes_new %>% 
-  ggplot() +
-    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
-    geom_point(data=wmt_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 3) +
-    geom_point(data=wmt_temp_stations, 
-             aes(long_temp, lat_temp, color = station_names), 
-             shape = 15, size = 3) +
-    theme_void() +
-    scale_color_manual(values = jv_palette)
-wmt_plot
-
-#Plot PM stations
-wmt_pm_stations <- pm_stations %>% 
-  filter(area == "WMT") %>% 
-  full_join(wmt_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_pm),
-         long_diff = abs(long_zip - long_pm),
-         sum_diff = lat_diff + long_diff) %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
-  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
-
-wmt_plot <- wmt_zipcodes_new %>% 
-  ggplot() +
-    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
-    geom_point(data=wmt_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 3) +
-    geom_point(data=wmt_pm_stations, 
-             aes(long_pm, lat_pm, color = local_site), 
-             shape = 15, size = 3) +
-    theme_void() +
-    scale_color_manual(values = jv_palette)
-wmt_plot
+addresses_new <- wmt_addresses %>% 
+  unite(address, c("Address", "City", "Zip"), sep = ", ", remove = FALSE) %>% 
+  select(Home, address) 
 ```
 
-# Mapping zipcodes, weather stations, pm stations - NN
 ```{r}
-nn_zipcodes_new <- nn_zipcodes %>% 
-  left_join(us_zipcodes, by = "zip") %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  mutate(lat_zip = lat,
-         long_zip = long)
-
-#Plot temp stations
-nn_temp_stations <- us_temp_stations %>% 
-  filter(state == "AZ" | state == "NM" |
-         state == "UT" | state == "CO") %>% 
-  mutate(lat_temp = lat,
-         long_temp = long) %>% 
-  filter(begin_date < "2014-06-01") %>% 
-  filter(end_date > "2020-03-21") %>% 
-  #distinct(station, .keep_all = TRUE) %>% 
-  full_join(nn_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_temp),
-         long_diff = abs(long_zip - long_temp),
-         sum_diff = lat_diff + long_diff,
-         area = "NN") %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
-  mutate(station_char = factor(station_names, 
-                               levels = c("FOUR CORNERS REGIONAL ARPT", "GALLUP MUNICIPAL AIRPORT",
-                                          "PAGE MUNICIPAL AIRPORT", "WINDOW ROCK AIRPORT",
-                                          "WINSLOW-LINDBERGH REGIONAL AI"),
-                               labels = c("fourcornersairport", "gallopairport",
-                                          "pageairport", "windowrockairport",
-                                          "winslowlindberghairport"))) %>% 
-  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
-         long_zip, station_names)
-
-nn_plot_temp_stations <- wmt_zipcodes_new %>% 
-  ggplot() +
-    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
-    geom_point(data=nn_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 3) +
-    geom_point(data=nn_temp_stations, 
-             aes(long_temp, lat_temp, color = station_names), 
-             shape = 15, size = 3) +
-    theme_void() +
-    scale_color_manual(values = jv_palette)
-nn_plot_temp_stations
-
-#Plot PM stations
-nn_pm_stations <- pm_stations %>% 
-  filter(area == "NN") %>% 
-  full_join(nn_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_pm),
-         long_diff = abs(long_zip - long_pm),
-         sum_diff = lat_diff + long_diff) %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
-  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
-
-nn_plot <- nn_zipcodes_new %>% 
-  ggplot() +
-    geom_polygon(data=nn_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
-    geom_point(data=nn_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 3) +
-    geom_point(data=nn_pm_stations, 
-             aes(long_pm, lat_pm, color = local_site), 
-             shape = 15, size = 3) +
-    theme_void() +
-    scale_color_manual(values = jv_palette)
-nn_plot
+addresses_geocoded <- geocode(location = addresses_new$address, output = "latlon",
+                              source = "google")
 ```
-
-# Mapping zipcodes, weather stations, pm stations - WMT
-```{r}
-ak_zipcodes_new <- ak_zipcodes %>% 
-  left_join(us_zipcodes, by = "zip") %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  mutate(lat_zip = lat,
-         long_zip = long)
-
-#Plot temp stations
-ak_temp_stations <- us_temp_stations %>% 
-  filter(state == "AK") %>% 
-  mutate(lat_temp = lat,
-         long_temp = long) %>% 
-  filter(begin_date < "2014-06-01") %>% 
-  filter(end_date > "2020-03-21") %>% 
-  #distinct(station, .keep_all = TRUE) %>% 
-  full_join(ak_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_temp),
-         long_diff = abs(long_zip - long_temp),
-         sum_diff = lat_diff + long_diff,
-         area = "AK") %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  mutate(station_char = factor(station_names, 
-                               levels = c("KIPNUK AIRPORT", "BETHEL AIRPORT",
-                                          "QUINHAGAK AIRPORT"),
-                               labels = c("kipnukairport", "bethelairport",
-                                          "quinhagakairport"))) %>% 
-  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
-         long_zip, station_names)
-
-zipcodes_map <- wmt_zipcodes_new %>% 
-  ggplot() +
-    geom_point(data=ak_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 3) +
-    geom_point(data=ak_temp_stations, 
-             aes(long_temp, lat_temp, color = station_names), 
-             shape = 15, size = 3) +
-    theme_void() +
-    scale_color_manual(values = jv_palette) 
-zipcodes_map
-
-#Plot PM stations
-ak_pm_stations <- pm_stations %>% 
-  filter(area == "AK") %>% 
-  full_join(ak_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_pm),
-         long_diff = abs(long_zip - long_pm),
-         sum_diff = lat_diff + long_diff) %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
-  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
-
-ak_plot <- ak_zipcodes_new %>% 
-  ggplot() +
-    geom_polygon(data=ak_map, aes(x=long, y=lat, group=group), 
-               colour="black", fill = "white") +
-    geom_point(data=ak_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 3) +
-    geom_point(data=ak_pm_stations, 
-             aes(long_pm, lat_pm, color = local_site), 
-             shape = 15, size = 3) +
-    theme_void() +
-    scale_color_manual(values = jv_palette)
-ak_plot
-```
-
-# Compile temperature data and filter for closest stations
-```{r}
-list_files <- list.files("Input/ambient/temp_data")
-# Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
-setwd("Input/ambient/temp_data") # set new working directory
-initial_data = tibble(file_name = list_files) %>% # use list of files from above
-  mutate(save_data = lapply(file_name, read_csv, 
-                            col_names = c("station", "wban", "temp_date",
-                                          "mean_temp", "t1", "mean_dewp", "t2",
-                                          "t3", "t4", "mean_pressure", "t5",
-                                          "mean_visib_miles", "t6", "mean_wind_knots",
-                                          "t7", "max_wind_knots", "t8", "max_temp",
-                                          "min_temp", "precip", "snow_depth", "t9"))) %>%
-  unnest(save_data)
-
-# Format combined file from above; combine with other PM data below
-temp_data <- initial_data %>% 
-  filter(!is.na(t1)) %>% 
-  separate(file_name, c("area", "station_char"), sep = "_") %>%
-  mutate(temp_date = ymd(temp_date),
-         mean_temp = as.numeric(mean_temp),
-         mean_dewp = as.numeric(mean_dewp),
-         mean_pressure = as.numeric(mean_pressure),
-         mean_visib_miles = as.numeric(mean_visib_miles),
-         mean_wind_knots = as.numeric(mean_wind_knots),
-         max_wind_knots = as.numeric(max_wind_knots)) %>% 
-  select(station_char, station, temp_date, mean_temp, mean_dewp,
-         mean_pressure, mean_visib_miles, mean_wind_knots, max_wind_knots) %>% 
-  replace_with_na_all(condition = ~.x %in% c(9999.9, 999.9))
-
-temp_stations <- rbind(wmt_temp_stations, nn_temp_stations, ak_temp_stations)
-
-temp_data_joined <- temp_data %>% 
-  left_join(temp_stations, by = c("station", "station_char"))
-
-#write_rds(temp_data_joined, "Output/kids_temp_data.rds")
-#write_csv(temp_data_joined, "Output/kids_temp_data.csv")
-
-kids_temp_data <- read_rds("Output/kids_temp_data.rds")
-```
-
-# Compile PM data and filter for correct dates/variables
-```{r}
-pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
-
-list_files <- list.files("Input/ambient/pm_data")
-# Set working directory and load files in list; extract file name and add as column
-## run next 4 lines together
-setwd("Input/ambient/pm_data") # set new working directory
-initial_data = tibble(file_name = list_files) %>% # use list of files from above
-   mutate(save_data = lapply(file_name, read_csv, skip = 1,
-                            col_names = c("t1", "t2", "t3", "t4", "parameter",
-                                          "pm_lat", "pm_long", "t5", "parameter_name",
-                                          "duration", "pollutant", "amb_pm_date",
-                                          "year", "t6", "units", "t7", "t8", "t9",
-                                          "t10", "t11", "mean_pollution", "t12",
-                                          "t13", "t14", "aqi", "t15", "t16",
-                                          "t17", "t18", "t19", "local_site",
-                                          "t20", "t21", "t22"),
-                            col_types = list(col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_double(), 
-                                          col_double(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(),
-                                          col_character(), col_character(), 
-                                          col_double(), col_character(),
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character(), 
-                                          col_character(), col_character()))) %>%
-  unnest(save_data)
-
-pm_data <- initial_data %>% 
-  filter(t1 != "State Code") %>% 
-  filter(pollutant == "PM25 24-hour 2012") %>% 
-  select(local_site, pollutant, amb_pm_date, mean_pollution) %>% 
-  mutate(amb_pm_24hr = as.numeric(mean_pollution),
-         date1 = ymd(amb_pm_date),
-         date2 = mdy(amb_pm_date)) %>% 
-  unite(amb_pm_date, c("date1", "date2"), sep = "") %>% 
-  separate(amb_pm_date, c("trash", "amb_pm_date"), sep = "NA") %>% 
-  unite(amb_pm_date, c("trash", "amb_pm_date"), sep = "") %>% 
-  group_by(local_site, amb_pm_date) %>% 
-  distinct(pollutant, .keep_all = TRUE) %>% 
-  select(local_site, amb_pm_date, amb_pm_24hr)
-  arrange(local_site, amb_pm_date)
-
-pm_stations <- rbind(wmt_pm_stations, nn_pm_stations, ak_pm_stations)
-
-pm_data_joined <- pm_data %>% 
-  left_join(pm_stations, by = "local_site")
-
-#write_rds(pm_data_joined, "Output/kids_amb_pm_data.rds")
-#write_csv(pm_data_joined, "Output/kids_amb_pm_data.csv")
-
-kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")
-```
-
-# Add new variables to pm/temp datasets
-```{r}
-kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")
-kids_temp_data <- read_rds("Output/kids_temp_data.rds")
-
-# PM data
-kids_amb_pm_data_new <- kids_amb_pm_data %>% 
-  arrange(area, local_site, amb_pm_date, zip) %>% 
-  group_by(area, local_site, zip) %>% 
-  mutate(amb_pm_24hr_lag1 = lag(amb_pm_24hr),
-         amb_pm_24hr_lag2 = lag(amb_pm_24hr, 2),
-         amb_pm_24hr_lag3 = lag(amb_pm_24hr, 3),
-         amb_pm_roll_mean_2day = rollmean(amb_pm_24hr, 2, fill = NA, align = "right"),
-         amb_pm_roll_mean_3day = rollmean(amb_pm_24hr, 3, fill = NA, align = "right"),
-         amb_pm_roll_mean_4day = rollmean(amb_pm_24hr, 4, fill = NA, align = "right")) %>% 
-  ungroup()
-
-#write_rds(kids_amb_pm_data_new, "Output/kids_amb_pm_data_new.rds")
-#write_csv(kids_amb_pm_data_new, "Output/kids_amb_pm_data_new.csv")
-
-# Temp data
-kids_amb_temp_data_new <- kids_temp_data %>% 
-  arrange(area, station, temp_date, zip) %>% 
-  group_by(area, station, zip) %>% 
-  mutate(mean_temp_lag1 = lag(mean_temp),
-         mean_temp_lag2 = lag(mean_temp, 2),
-         mean_temp_lag3 = lag(mean_temp, 3),
-         mean_temp_roll_2day = rollmean(mean_temp, 2, fill = NA, align = "right"),
-         mean_temp_roll_3day = rollmean(mean_temp, 3, fill = NA, align = "right"),
-         mean_temp_roll_4day = rollmean(mean_temp, 4, fill = NA, align = "right")) %>% 
-  ungroup()
-
-#write_rds(kids_amb_temp_data_new, "Output/kids_amb_temp_data_new.rds")
-#write_csv(kids_amb_temp_data_new, "Output/kids_amb_temp_data_new.csv")
-```

### 2020-04-06 child_health_data_cleaning.Rmd
health data update, LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 27 March 2020"
+date: "Started 27 Feb 2020, Updated 6 April 2020"
 output: pdf_document 
@@
 wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
+  replace_with_na_all(condition = ~.x == -999) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          med_visit_ari = as.character(ARI),
-         lethargy = as.character(Lethargy),
-         retractions = as.character(Retractions),
+         med_visit_lethargy = as.character(Lethargy),
+         med_visit_retractions = as.character(Retractions),
          temp1 = as.numeric(Temp1),
@@
          temp3 = as.numeric(Temp3),
-         resp_rate = as.numeric(RR),
-         heart_rate = as.numeric(HR),
-         wheeze = as.character(Wheeze),
-         stridor = as.character(Stridor),
-         crackles = as.character(Crackles),
-         sp02 = as.numeric(SP02),
-         hr_sp02 = as.numeric(HRPulseOx),
-         referred = as.character(Referred),
+         med_visit_resp_rate = as.numeric(RR),
+         med_visit_heart_rate = as.numeric(HR),
+         med_visit_wheeze = as.character(Wheeze),
+         med_visit_stridor = as.character(Stridor),
+         med_visit_crackles = as.character(Crackles),
+         med_visit_sp02 = as.numeric(SP02),
+         med_visit_hr_sp02 = as.numeric(HRPulseOx),
+         med_visit_referred = as.character(Referred),
          med_visit_notes = as.character(Notes)) %>% 
-  select(home_winter_id:med_visit_notes) %>% 
-  replace_with_na_all(condition = ~.x == -999)
+  # gather 3 temperatures to calculate average
+  gather("sample", "value", c("temp1":"temp3")) %>% 
+  group_by(home_winter_id, med_visit_date) %>% 
+  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  # spread 3 moisture samples back to wide format
+  spread(sample, value) %>%
+  select(home_winter_id:med_visit_mean_temp)
 
 nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>% 
+  replace_with_na_all(condition = ~.x == -999) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          temp3 = as.numeric(Temp3),
-         resp_rate = as.numeric(RR),
-         heart_rate = as.numeric(HR),
-         wheeze = as.character(Wheeze),
-         stridor = as.character(Stridor),
-         crackles = as.character(Crackles),
-         sp02 = as.numeric(SP02),
-         hr_sp02 = as.numeric(HRPulseOx),
-         referred = as.character(Referred),
+         med_visit_resp_rate = as.numeric(RR),
+         med_visit_heart_rate = as.numeric(HR),
+         med_visit_wheeze = as.character(Wheeze),
+         med_visit_stridor = as.character(Stridor),
+         med_visit_crackles = as.character(Crackles),
+         med_visit_sp02 = as.numeric(SP02),
+         med_visit_hr_sp02 = as.numeric(HRPulseOx),
+         med_visit_referred = as.character(Referred),
          med_visit_notes = as.character(Level)) %>% 
-  select(home_winter_id:med_visit_notes) %>% 
-  replace_with_na_all(condition = ~.x == -999)
+  # gather 3 temperatures to calculate average
+  gather("sample", "value", c("temp1":"temp3")) %>% 
+  group_by(home_winter_id, med_visit_date) %>% 
+  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  # spread 3 moisture samples back to wide format
+  spread(sample, value) %>%
+  select(home_winter_id:med_visit_mean_temp)
 
 ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
+  replace_with_na_all(condition = ~.x == -999) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          temp3 = as.numeric(Temp3),
-         resp_rate = as.numeric(RR),
-         heart_rate = as.numeric(HR),
-         wheeze = as.character(Wheeze),
-         stridor = as.character(Stridor),
-         crackles = as.character(Crackles),
-         sp02 = as.numeric(SP02),
-         hr_sp02 = as.numeric(HRPulseOx),
-         referred = as.character(Referred),
+         med_visit_resp_rate = as.numeric(RR),
+         med_visit_heart_rate = as.numeric(HR),
+         med_visit_wheeze = as.character(Wheeze),
+         med_visit_stridor = as.character(Stridor),
+         med_visit_crackles = as.character(Crackles),
+         med_visit_sp02 = as.numeric(SP02),
+         med_visit_hr_sp02 = as.numeric(HRPulseOx),
+         med_visit_referred = as.character(Referred),
          med_visit_notes = as.character(Level)) %>% 
-  select(home_winter_id:med_visit_notes) %>% 
-  replace_with_na_all(condition = ~.x == -999)
+  # gather 3 temperatures to calculate average
+  gather("sample", "value", c("temp1":"temp3")) %>% 
+  group_by(home_winter_id, med_visit_date) %>% 
+  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  # spread 3 moisture samples back to wide format
+  spread(sample, value) %>%
+  select(home_winter_id:med_visit_mean_temp)
 
@@
          area = "WMT",
-         dr_relook = DrRelook,
-         ari_status = as.character(Status),
-         symptom_free_date = ymd(SymptomFreeDate),
-         csq_duedate = ymd(CSQDueDate),
-         csq_date = ymd(CSQDate),
-         csq_time = as.character(CSQTime),
-         dr_comments = as.character(DrComments),
-         dr_history = as.character(DrHistory),
-         notify_parents = as.character(NotifyParents),
-         notes_to_parents = as.character(NotesToParents),
-         any_ari = as.character(AnyARI),
-         ari_comments = as.character(WinterARIComments)) %>% 
-  select(home_winter_id:ari_comments)
+         chw_dr_relook = DrRelook,
+         chw_ari_status = as.character(Status),
+         chw_symptom_free_date = ymd(SymptomFreeDate),
+         chw_csq_duedate = ymd(CSQDueDate),
+         chw_csq_date = ymd(CSQDate),
+         chw_csq_time = as.character(CSQTime),
+         chw_dr_comments = as.character(DrComments),
+         chw_dr_history = as.character(DrHistory),
+         chw_notify_parents = as.character(NotifyParents),
+         chw_notes_to_parents = as.character(NotesToParents),
+         chw_any_ari = as.character(AnyARI),
+         chw_ari_comments = as.character(WinterARIComments)) %>% 
+  select(home_winter_id:chw_ari_comments)
 
@@
          csq_date = ymd(Date),
-         sick_2_weeks = as.character(Illin2Weeks),
-         sick = as.character(Sick),
-         ari = as.character(ARI),
-         still_sick = as.character(StillSick),
-         first_symptom_date = ymd(FirstSymptomDate),
-         see_dr = as.character(SeeDr),
-         dr_diagnosis = as.character(DrDiagnosis),
-         new_meds = as.character(NewMeds),
-         fever = as.character(Fever),
-         temp = as.numeric(Temperature),
-         rapid_breathing = as.character(RapidBreathing),
-         difficult_breathing = as.character(DifficultBreathing),
-         noisy_breathing = as.character(NoisyBreathing),
-         runny_nose = as.character(RunnyNose),
-         cough = as.character(Coughing),
-         wheeze = as.character(Wheezing),
-         symptom_free_date = ymd(SymptomFreeDate),
-         comments = as.character(Comments)) %>% 
-  select(home_winter_id:comments)
+         csq_collect = as.character(HowCollected),
+         csq_sick_2_weeks = as.character(Illin2Weeks),
+         csq_still_sick = as.character(StillSick),
+         csq_first_symptom_date = ymd(FirstSymptomDate),
+         csq_symptom_free_date = ymd(SymptomFreeDate),
+         csq_miss_work = as.character(MissWork),
+         csq_miss_work_days = as.character(DaysMissed),
+         csq_see_dr = as.character(SeeDr),
+         csq_dr_diagnosis = as.character(DrDiagnosis),
+         csq_new_meds = as.character(NewMeds),
+         csq_fever_2weeks = as.character(Fever),
+         csq_temp_2weeks = as.numeric(Temperature),
+         csq_rapid_breathing_2weeks = as.character(RapidBreathing),
+         csq_difficult_breathing_2weeks = as.character(DifficultBreathing),
+         csq_noisy_breathing_2weeks = as.character(NoisyBreathing),
+         csq_runny_nose_2weeks = as.character(RunnyNose),
+         csq_cough_2weeks = as.character(Coughing),
+         csq_wheeze_2weeks = as.character(Wheezing),
+         csq_fever_current = as.character(FeverStill),
+         csq_rapid_breathing_current = as.character(RapidBreathingStill),
+         csq_difficult_breathing_current = as.character(DifficultBreathingStill),
+         csq_noisy_breathing_current = as.character(NoisyBreathingStill),
+         csq_runny_nose_current = as.character(RunnyNoseStill),
+         csq_cough_current = as.character(CoughingStill),
+         csq_wheeze_current = as.character(WheezingStill),
+         csq_comments = as.character(Comments)) %>% 
+  select(home_winter_id:csq_comments)
 
@@
          visit_date = ymd(VisitDate),
-         dr_diagnosis = as.character(DrDiag),
-         ari = as.character(ARI),
-         comments = as.character(DrComments)) %>% 
-  select(home_winter_id:comments)
+         med_rec_visit_date = ymd(VisitDate),
+         med_rec_dr_diagnosis = as.character(DrDiag),
+         med_rec_ari = as.character(ARI),
+         med_rec_comments = as.character(DrComments)) %>% 
+  select(home_winter_id:med_rec_comments)
 

### 2020-04-06 lrti_tracking.Rmd
health data update, LRTI tracking

---
title: "Kids LRTI tracking"
author: "Ethan Walker"
date: "Started 6 April 2020, Updated 6 April 2020"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
```

```{r}
lrti_data <- health_exposure_data_sampling_day %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
         csq_date:csq_comments, med_rec_visit_date:med_rec_comments) %>% 
  filter(!is.na(visit_date)) %>% 
  filter(area == "WMT") %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  mutate(trash_date = ymd("2099-01-01"),
         still_sick_date = if_else(csq_still_sick == "Yes", csq_date, trash_date),
         med_rec_ari_date = if_else(med_rec_ari != "No ARI", med_rec_visit_date, trash_date),
         med_visit_ari_date = if_else(med_visit_ari != "No", med_visit_date, trash_date)) %>% 
  pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
               names_to = "lrti_visits", values_to = "lrti_dates") %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  mutate(lrti_start_date_new = if_else(!is.na(csq_first_symptom_date),
                                       csq_first_symptom_date, lrti_dates)) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
  distinct(lrti_start_date_new, .keep_all = TRUE)

table(lrti_data$med_visit_ari)
```

### 2020-04-06 merge_health_datasets.Rmd
health data update, LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 27 March 2020"
+date: "Started 27 March 2020, Updated 6 April 2020"
 output: pdf_document
@@
 kids_med_visit <- read_rds("Output/kids_med_visit.rds") %>% 
-  mutate(visit_date = med_visit_date) %>% 
-  rename(wheeze_med_visit = wheeze)
+  mutate(visit_date = med_visit_date)
 
@@
   mutate(visit_date = csq_date) %>% 
-  rename(wheeze_csq = wheeze) %>% 
   full_join(kids_med_visit, by = c("area", "home", "home_winter_id", "treatment",
@@
 child_home_winter <- read_rds("Output/child_home_winter.rds") %>% 
-  select(-csq_date, -home_id.x, -home_id.y, -symptom_free_date)
+  select(-home_id.x, -home_id.y, -chw_symptom_free_date)
 
@@
 
-#write_rds(health_merged4, "Output/health_data_merged.rds")
-#write_csv(health_merged4, "Output/health_data_merged.csv")
+kids_medical_records <- read_rds("Output/kids_medical_records.rds") %>% 
+  select(-visit_date)
+
+health_merged5 <- health_merged4 %>% 
+  left_join(kids_medical_records, by = c("area", "home", "home_winter_id", "winter_id", 
+                                      "child_id_num", "treatment")) %>% 
+  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
+  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
+         treatment, sample_date, sampling_day, at_home_location, 
+         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:med_rec_comments) %>% 
+  ungroup()
+
+#write_rds(health_merged5, "Output/health_data_merged.rds")
+#write_csv(health_merged5, "Output/health_data_merged.csv")
 ```

### 2020-04-07 child_health_data_cleaning.Rmd
LRTI tracking updates

@@
 wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
-  replace_with_na_all(condition = ~.x == -999) %>% 
+  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
+         HR = if_else(HR < 50, -999, HR),
+         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          med_visit_referred = as.character(Referred),
-         med_visit_notes = as.character(Notes)) %>% 
+         med_visit_notes = as.character(Level)) %>% 
   # gather 3 temperatures to calculate average
   gather("sample", "value", c("temp1":"temp3")) %>% 
-  group_by(home_winter_id, med_visit_date) %>% 
+  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
+  mutate(value = if_else(value < 90, -99, value)) %>% 
+  replace_with_na(replace = list(value = -99)) %>% 
   mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
@@
 nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>%
-  replace_with_na_all(condition = ~.x == -999) %>% 
+  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
+         HR = if_else(HR < 50, -999, HR),
+         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
   gather("sample", "value", c("temp1":"temp3")) %>% 
-  group_by(home_winter_id, med_visit_date) %>% 
+  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
+  mutate(value = if_else(value < 90, -99, value)) %>% 
+  replace_with_na(replace = list(value = -99)) %>% 
   mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
@@
 ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
-  replace_with_na_all(condition = ~.x == -999) %>% 
+  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
+         HR = if_else(HR < 50, -999, HR),
+         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),

### 2020-04-07 lrti_tracking.Rmd
LRTI tracking updates

@@
 author: "Ethan Walker"
-date: "Started 6 April 2020, Updated 6 April 2020"
+date: "Started 6 April 2020, Updated 7 April 2020"
 output: pdf_document 
@@
 library(knitr)
-jv_palette <- c("#330099","#CC0066","#FF6633", 
-                 "#0099CC", "#FF9900","#CC6633",
-                  "#FF3366", "#33CC99", "#33999")
-cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
-               "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 
+# Diagnosing LRTI from the data
+
+Is there a specific protocol Paul uses for diagnosing LRTI, or is it case-by-case
+based on the given information for each child? We'll need to have some kind of
+protocol/algorithm for publications, I would think.
+
+Other questions:
+- How far has the team made it (date/winter) in diagnosing LRTI? Winter 17/18?
+- Has this process been done/started for NN/AK, or just WMT?
+
 ```{r}
+# LRTI Diagnosis data
+lrti_diagnosis_data <- health_exposure_data_sampling_day %>% 
+  # select relevant health data
+  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
+         med_visit_date:med_visit_mean_temp,
+         csq_date:csq_comments, 
+         med_rec_visit_date, med_rec_dr_diagnosis, med_rec_ari, med_rec_comments) %>% 
+  # Currently only have this data for WMT - need to get new files and updates dataset
+  filter(area == "WMT") %>% 
+  # This is the most broad way of distinguishing if a child was sick
+  # Need to get Paul's protocol for diagnosis to filter down from here
+  filter(csq_sick_2_weeks == "Yes")
+```
+
+The filtering below narrows down the LRTI definition too much; we are missing cases
+where children had LRTI but did not visit a healthcare facility and have a med rec.
+If we only use these cases, it could bias the results.
+
+There are instances of LRTI in our records that weren’t confirmed in the medical record. 
+There are other instances where the medical record diagnosed LRTI, but our records 
+overturned it. What is our criteria for diagnosing LRTI, where is it documented, 
+and how can we be sure it is consistent in diagnosing LRTI?
+
+```{r}
+# Further filtering and re-ordering of variables
+lrti_diagnosis_data_filtered <- lrti_diagnosis_data %>% 
+  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
+         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
+  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
+  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
+  # If a home CSQ was done on same visit date as initial phone CSQ, filter for only the home CSQ
+  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
+  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
+         csq_first_symptom_date:csq_comments) %>% 
+  # Were they referred (med visit) or did they see a Dr (CSQ)?
+  filter(med_visit_referred == "Yes" | csq_see_dr == "Yes") %>% 
+  # If yes, did we get medical records?
+  filter(!is.na(med_rec_visit_date)) %>% 
+  # If yes, was there an ARI diagnosis?
+  filter(med_rec_ari == "ARI" | med_rec_ari == "Possible ARI") %>% 
+  group_by(child_id_num) %>% 
+  distinct(med_rec_visit_date, .keep_all = TRUE)
+```
+
+# Defining start date, end date, and at-risk period for each diagnosed LRTI case
+
+First, I need to know (or make) exactly which variable is a final diagnosed case 
+of LRTI in the data. Then, I can create a method in R to identify start/end/at-risk.
+The code below is a start at this, using Nick's handoff documents/algorithm.
+
+```{r}
+# LRTI start date, end date, and at-risk period identification
 lrti_data <- health_exposure_data_sampling_day %>% 
@@
          csq_date:csq_comments, med_rec_visit_date:med_rec_comments) %>% 
-  filter(!is.na(visit_date)) %>% 
   filter(area == "WMT") %>% 
+  filter(csq_sick_2_weeks == "Yes") %>% 
   select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
@@
   group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
-  distinct(visit_date, .keep_all = TRUE) %>% 
+  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
+  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
+  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
+         csq_first_symptom_date:csq_comments)
   mutate(trash_date = ymd("2099-01-01"),
@@
   select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date) %>% 
+         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date,
+         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari) %>% 
   arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
@@
 table(lrti_data$med_visit_ari)
+table(health_exposure_data_sampling_day$csq_sick_2_weeks)
 ```

### 2020-04-07 merge_health_datasets.Rmd
LRTI tracking updates

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 6 April 2020"
+date: "Started 27 March 2020, Updated 7 April 2020"
 output: pdf_document
@@
   select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         treatment, csq_date:med_visit_notes) %>% 
+         treatment, csq_date:med_visit_mean_temp) %>% 
   arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")
-
 exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
-
-exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")
-
-sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
 

### 2020-04-09 child_health_data_cleaning.Rmd
Update LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 6 April 2020"
+date: "Started 27 Feb 2020, Updated 8 April 2020"
 output: pdf_document 
@@
          csq_comments = as.character(Comments)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
   select(home_winter_id:csq_comments)
@@
 ```
+
+################# Confirmed LRTI ###############
+
+# Load and merge files from different locations
+```{r}
+wmt_ari_initial <- read_xlsx("Input/WMT/wmt_ari_20200408.xlsx") %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "WMT",
+         lrti_number = as.character(ARINum),
+         lrti_basis = as.character(ARIBasis),
+         lrti_start_date_final = ymd(ProfessionalStartDate),
+         lrti_end_date_final = ymd(ProfessionalEndDate),
+         lrti_professional_comments = as.character(ProfessionalComments)) %>% 
+  select(home_winter_id:lrti_professional_comments)
+
+# Waiting on files from NN and AK
+
+# Combine logs from all areas
+kids_lrti <- wmt_ari_initial %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id"))
+
+# write_rds(kids_lrti, "Output/kids_lrti.rds")
+
+kids_lrti <- read_rds("Output/kids_lrti.rds")
+
+```

### 2020-04-09 child_health_data_summary.Rmd
Update LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 16 March 2020, Updated 16 March 2020"
+date: "Started 16 March 2020, Updated 8 April 2020"
 output: 
@@
                       include = TRUE, warning = FALSE,
-                      fig.height = 5, fig.width = 7)
+                      fig.height = 5, fig.width = 8)
 ```
@@
 
-kids_med_visit <- read_rds("Output/kids_med_visit.rds") %>% 
-  group_by(area, home_winter_id, child_id_num, med_visit_date) %>% 
-  distinct(med_visit_date, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  mutate(temp1 = if_else(temp1 < 90, 0, temp1)) %>% 
-  mutate(temp2 = if_else(temp2 < 90, 0, temp2)) %>% 
-  mutate(temp3 = if_else(temp3 < 90, 0, temp3)) %>% 
-  mutate(sp02 = if_else(sp02 < 80, 0, sp02)) %>% 
-  replace_with_na_all(condition = ~.x %in% c(-9999, -99, 0)) %>% 
-  # gather 3 moisture samples to calculate average
-  gather("sample", "value", c("temp1":"temp3")) %>% 
-  arrange(home_winter_id, med_visit_date) %>% 
-  group_by(area, home_winter_id, med_visit_date) %>% 
-  mutate(temp_ave = mean(value, na.rm = TRUE)) %>% 
-  # spread 3 moisture samples back to wide format
-  group_by(area, home_winter_id, med_visit_date, temp_ave) %>% 
-  spread(sample, value) %>% 
-  ungroup()
-
+kids_med_visit <- read_rds("Output/kids_med_visit.rds") 
 ```
@@
   filter(winter_id == 1) %>% 
-  select(area, resp_rate, heart_rate, sp02, hr_sp02, temp_ave) %>% 
-  pivot_longer(cols = c("resp_rate":"temp_ave"), names_to = "var") %>% 
+  select(area, med_visit_resp_rate, med_visit_heart_rate, med_visit_sp02, 
+         med_visit_hr_sp02, med_visit_mean_temp) %>% 
+  pivot_longer(cols = c("med_visit_resp_rate":"med_visit_mean_temp"), names_to = "var") %>% 
   filter(!is.na(value)) %>% 
@@
     theme_minimal() +
-    theme(strip.text = element_text(size = 16, colour = "black"),
+    theme(strip.text = element_text(size = 14, colour = "black"),
           axis.text.x = element_text(size = 16, colour = "black"),
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "resp_rate", "Respiratory rate", "breaths/min")
+medvisit_summary_function(kids_med_visit, "med_visit_resp_rate", "Respiratory rate", "breaths/min")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "heart_rate", "Heart rate", "beats/min")
+medvisit_summary_function(kids_med_visit, "med_visit_heart_rate", "Heart rate", "beats/min")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "sp02", "Sp02", "%")
+medvisit_summary_function(kids_med_visit, "med_visit_sp02", "Sp02", "%")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "hr_sp02", "Heart rate from Sp02", "beats/min")
+medvisit_summary_function(kids_med_visit, "med_visit_hr_sp02", "Heart rate from Sp02", "beats/min")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "temp_ave", "Temperature", "Degrees F")
+medvisit_summary_function(kids_med_visit, "med_visit_mean_temp", "Temperature", "Degrees F")
 ```
@@
   filter(winter_id == 1) %>% 
-  select(area, med_visit_ari, lethargy, retractions, wheeze, stridor, 
-         crackles, referred) %>% 
-  pivot_longer(cols = c("med_visit_ari":"referred"), names_to = "var") %>% 
+  select(area, med_visit_lethargy, med_visit_retractions, med_visit_wheeze, 
+         med_visit_stridor, med_visit_crackles, med_visit_referred) %>% 
+  pivot_longer(cols = c("med_visit_lethargy":"med_visit_referred"), names_to = "var") %>% 
   filter(!is.na(value)) %>% 
@@
     theme_minimal() +
-    theme(strip.text = element_text(size = 16, colour = "black"),
+    theme(strip.text = element_text(size = 12, colour = "black"),
           axis.text.x = element_text(size = 14, colour = "black", 
@@
     ylab(y_label) +
-    scale_fill_manual(values = jv_palette)
+    scale_fill_manual(values = cbPalette)
   medvisit_plot
@@
 
+\pagebreak
+
 ```{r}
-medvisit_summary_function(kids_med_visit, "med_visit_ari", "Med visit ARI", "count")
+medvisit_summary_function(kids_med_visit, "med_visit_lethargy", "Lethargy", "count")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "lethargy", "Lethargy", "count")
+medvisit_summary_function(kids_med_visit, "med_visit_retractions", "Retractions", "count")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "retractions", "Retractions", "count")
+medvisit_summary_function(kids_med_visit, "med_visit_wheeze", "Wheeze", "count")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "wheeze", "Wheeze", "count")
+medvisit_summary_function(kids_med_visit, "med_visit_stridor", "Stridor", "count")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "stridor", "Stridor", "count")
+medvisit_summary_function(kids_med_visit, "med_visit_crackles", "Crackles", "count")
 ```
@@
 ```{r}
-medvisit_summary_function(kids_med_visit, "crackles", "Crackles", "count")
+medvisit_summary_function(kids_med_visit, "med_visit_referred", "Referred", "count")
 ```
-
-\pagebreak
-
-```{r}
-medvisit_summary_function(kids_med_visit, "referred", "Referred", "count")
-```

### 2020-04-09 lrti_tracking.Rmd
Update LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 6 April 2020, Updated 7 April 2020"
+date: "Started 6 April 2020, Updated 8 April 2020"
 output: pdf_document 
@@
          csq_date:csq_comments, 
-         med_rec_visit_date, med_rec_dr_diagnosis, med_rec_ari, med_rec_comments) %>% 
+         med_rec_visit_date, med_rec_dr_diagnosis, med_rec_ari, med_rec_comments,
+         lrti_number:lrti_professional_comments) %>% 
   # Currently only have this data for WMT - need to get new files and updates dataset
@@
   select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
-         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
+         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments,
+         lrti_number:lrti_professional_comments) %>% 
   group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
@@
   select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
-         csq_first_symptom_date:csq_comments) %>% 
+         csq_first_symptom_date:lrti_professional_comments) %>% 
   # Were they referred (med visit) or did they see a Dr (CSQ)?
@@
 
+The initial attempt at having R select start dates did not match up well with
+the cases in which Paul had already finalized the start/end dates. There seems
+to be a good amount of professional opinion by assessing symptoms and medical
+records. It may be best to filter a dataset down to only confirmed LRTI cases
+and useful health/medical variables, then decide start/end/at-risk times 
+on a case-by-case basis.
+
 ```{r}
 # LRTI start date, end date, and at-risk period identification
-lrti_data <- health_exposure_data_sampling_day %>% 
+lrti_date_data <- health_exposure_data_sampling_day %>% 
   select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
          med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
-         csq_date:csq_comments, med_rec_visit_date:med_rec_comments) %>% 
+         csq_date:csq_comments, med_rec_visit_date:med_rec_comments,
+         lrti_number:lrti_professional_comments) %>% 
   filter(area == "WMT") %>% 
@@
   select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
-         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments) %>% 
+         med_rec_visit_date:med_rec_comments, 
+         med_visit_lethargy:lrti_professional_comments) %>% 
   group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
@@
   select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
-         csq_first_symptom_date:csq_comments)
+         csq_first_symptom_date:lrti_professional_comments) %>% 
   mutate(trash_date = ymd("2099-01-01"),
-         still_sick_date = if_else(csq_still_sick == "Yes", csq_date, trash_date),
-         med_rec_ari_date = if_else(med_rec_ari != "No ARI", med_rec_visit_date, trash_date),
-         med_visit_ari_date = if_else(med_visit_ari != "No", med_visit_date, trash_date)) %>% 
+         still_sick_date = if_else(csq_still_sick == "Yes", 
+                                   csq_date, trash_date),
+         med_rec_ari_date = if_else(!is.na(med_rec_visit_date), 
+                                    med_rec_visit_date, trash_date),
+         med_visit_ari_date = if_else(!is.na(med_visit_date), 
+                                      med_visit_date, trash_date)) %>% 
   pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
@@
          csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date,
-         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari) %>% 
+         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari,
+         lrti_number:lrti_professional_comments) %>% 
   arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
@@
   group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
-  distinct(lrti_start_date_new, .keep_all = TRUE)
-
-table(lrti_data$med_visit_ari)
-table(health_exposure_data_sampling_day$csq_sick_2_weeks)
+  distinct(lrti_start_date_new, .keep_all = TRUE) %>% 
+  distinct(child_id_num, lrti_number, .keep_all = TRUE) %>% 
+  filter(!is.na(lrti_number)) %>% 
+  select(child_id_num, home_winter_id, lrti_number, lrti_start_date_new:lrti_dates, 
+         lrti_start_date_final, lrti_end_date_final, area:lrti_professional_comments)
 ```

### 2020-04-09 merge_health_datasets.Rmd
Update LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 7 April 2020"
+date: "Started 27 March 2020, Updated 8 April 2020"
 output: pdf_document
@@
 kids_csq <- read_rds("Output/kids_csq.rds")
+kids_lrti <- read_rds("Output/kids_lrti.rds")
 kids_medical_records <- read_rds("Output/kids_medical_records.rds")
@@
 
-#write_rds(health_merged5, "Output/health_data_merged.rds")
-#write_csv(health_merged5, "Output/health_data_merged.csv")
+kids_lrti <- read_rds("Output/kids_lrti.rds") %>% 
+  select(-home_id)
+
+health_merged6 <- health_merged5 %>% 
+  left_join(kids_lrti, by = c("area", "home", "home_winter_id", "winter_id", 
+                                      "child_id_num", "treatment")) %>% 
+  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
+  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
+         treatment, sample_date, sampling_day, at_home_location, 
+         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:lrti_professional_comments) %>% 
+  ungroup()
+
+#write_rds(health_merged6, "Output/health_data_merged.rds")
+#write_csv(health_merged6, "Output/health_data_merged.csv")
 ```

### 2020-04-09 pm_summary.Rmd
Update LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 16 March 2020"
+date: "Started 22 Nov 2019, Updated Apri 8 2020"
 output: powerpoint_presentation
@@
   summarize("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "25%" = quantile(pm, 0.25, na.rm = TRUE),
-            "Median PM" = median(pm, na.rm = TRUE),
-            "75%" = quantile(pm, 0.75, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
-            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
-            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
-            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
-            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
+            "Mean PM" = round(mean(pm, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm, na.rm = TRUE)),
+            "Min PM" = round(min(pm, na.rm = TRUE)), 
+            "25%" = round(quantile(pm, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm, na.rm = TRUE)),
+            "75%" = round(quantile(pm, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm, na.rm = TRUE)),
+            "Mean Sample Days" = round(mean(pm_sample_interval, na.rm = TRUE)),
+            "SD Sample Days" = round(sd(pm_sample_interval, na.rm = TRUE))) %>% 
   #select(-treatment) %>% 
@@
 
-kable(pm_summary, 
-      digits = 1, align = "c")
+kable(pm_summary, align = "c")
 

### 2020-04-09 exposure_initial_analysis.Rmd
exposure and lrti analysis updates

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 3 April 2020"
+date: "Started 2 Feb 2020, Updated 9 April 2020"
 output: pdf_document
@@
 ```{r, message=FALSE}
-library(tidyverse)
 library(readxl)
@@
 library(ggmap)
+library(mctest)
+library(ppcor)
+library(GGally)
+library(tidyverse)
 ```
@@
 ```
+
+# Run correlations/Chi-square tests between different covariates
+Goal is to check for vars that have a lot of overlap and eliminate some from
+consideration to include in models
+```{r}
+pm_mean_sampling_period <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1) %>% 
+  dplyr::select(pm_mean_sampling_period, education_3level, income_3level,
+                race_parent, race_child, total_residents_2level, 
+                residents_under5_2level, residents_smoke,
+                stove_grade_3level, burn_level_3level, wood_cords,
+                mean_moisture_winter) %>% 
+  mutate_all(as.numeric) %>% 
+  na.exclude() %>% 
+  dplyr::select(pm_mean_sampling_period)
+
+other_vars <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1) %>% 
+  dplyr::select(pm_mean_sampling_period, education_3level, income_3level,
+                race_parent, race_child, total_residents_2level, 
+                residents_under5_2level, residents_smoke,
+                stove_grade_3level, burn_level_3level, wood_cords,
+                mean_moisture_winter) %>% 
+  mutate_all(as.numeric) %>% 
+  na.exclude() %>% 
+  dplyr::select(-pm_mean_sampling_period)
+
+omcdiag(other_vars, pm_mean_sampling_period)
+
+imcdiag(other_vars, pm_mean_sampling_period)
+imcdiag(other_vars, pm_mean_sampling_period, method = "Wi")
+```
+
+# Try individual chi-square tests and correlations b/w two vars
+```{r}
+chisqu_results <- chisq.test(chisqu_data_demographics$education_3level, 
+                             chisqu_data_demographics$income_3level)
+chisqu_results
+
+chisqu_results <- cor.test(chisqu_data_demographics$education_3level, 
+                             chisqu_data_demographics$income_3level)
+chisqu_results
+
+ggpairs(chisqu_data_demographics)
+```

### 2020-04-09 lrti_tracking.Rmd
exposure and lrti analysis updates

@@
 author: "Ethan Walker"
-date: "Started 6 April 2020, Updated 8 April 2020"
+date: "Started 6 April 2020, Updated 9 April 2020"
 output: pdf_document 
@@
 Other questions:
+- LRTI and ARI have been used interchangably. Our actual outcome is LRTI, yes?
+  - Is there a clinical definition we/Paul have been following when diagnosing cases?
+  - LRTI: bronchiolitis, pneumonia, croupe? Case definition up to Paul.
+  - Primary diagnosis is based on med recs, but secondary outcomes could be from
+    surveys. Also see how well survey data predicts Pauls diagnosis.
+- How does Paul initially narrow down which cases to look at? Or does he look
+  at each CSQ and med visit form?
 - How far has the team made it (date/winter) in diagnosing LRTI? Winter 17/18?
+  - Could touch base with Nick, but probably through winter 17/18.
 - Has this process been done/started for NN/AK, or just WMT?
+  - No. Surveys may be worse for these areas, but med rec data could be better
+    through Indian Health Services.
+  - At risk period could be difficult for these sites
+- Where did the med_visit_ari variable come from?
+
+Notes and specific documentation for steps to identify LRTI cases:
+- Double check the filter method of "yes" for sick/still sick
+- Ask Kathrene about med rec ARI vs med visit ARI and where they came from
+- Was there a variable where parents called us to report sickness?
+-
+
+Next steps
+- Ask Kathrene some questions from above
+  - Have her upload final health data for NN/AK
+- Push to get the final data input, QA/QC, uploaded to Box
+- Go through winter 18/19 data line by line and flag cases that could be 
+  respiratory.
+  - Do this for all sites, and start on other sites for all years.
+- Also flag other potential secondary outcomes that we could look at.
 
@@
   # select relevant health data
-  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         med_visit_date:med_visit_mean_temp,
-         csq_date:csq_comments, 
-         med_rec_visit_date, med_rec_dr_diagnosis, med_rec_ari, med_rec_comments,
+  select(area, home, home_winter_id, winter_id, child_id_num, 
+         visit_date, med_visit_date, med_visit_referred,
+         csq_date:csq_symptom_free_date,
+         csq_see_dr, csq_dr_diagnosis,
+         med_rec_visit_date:med_rec_comments,
+         csq_miss_work:csq_comments,
+         med_visit_ari:med_visit_mean_temp,
          lrti_number:lrti_professional_comments) %>% 
@@
   # This is the most broad way of distinguishing if a child was sick
-  # Need to get Paul's protocol for diagnosis to filter down from here
-  filter(csq_sick_2_weeks == "Yes") 
+  # Careful when filtering here. Could still make cases that these vars didn't catch
+  # Best to have a look through all rows, then use this to filter data down so it
+  # is easier to work with to assess dates/at-risk times
+  filter(csq_sick_2_weeks == "Yes" | csq_still_sick == "Yes") %>% 
+  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
+  ungroup()
+
+#  filter(visit_date > "2018-06-01") # filter for newer data by date
+
+# write_rds(lrti_diagnosis_data, "Output/lrti_diagnosis_data.rds")
+# write_csv(lrti_diagnosis_data, "Output/lrti_diagnosis_data.csv", na = " ")
 ```

### 2020-04-10 exposure_initial_analysis.Rmd
new exposure analysis

--- 2020-04-03 exposure_initial_analysis.Rmd
@@
 library(ggmap)
+library(mctest)
+library(GGally)
+library(tidyverse)
 ```
@@
 ```
+
+# Run tests for multicollinearity between different covariates
+Goal is to check for vars that have a lot of overlap and eliminate some from
+consideration to include in models
+
+First, use `mctest` package to check for multicollinearity among similar variables
+https://cran.r-project.org/web/packages/mctest/mctest.pdf
+
+Notes on multicollinearity:
+
+https://datascienceplus.com/multicollinearity-in-r/ 
+https://www.theanalysisfactor.com/eight-ways-to-detect-multicollinearity/
+
+```{r}
+# Filter data for specific variables; select pm variable
+pm_mean_sampling_period <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(pm_mean_sampling_period, income_3level, residents_smoke) %>% 
+  mutate_all(as.numeric) %>% 
+  na.exclude() %>% 
+  select(pm_mean_sampling_period)
+
+# Filter data for specific variables; select predictor variables
+other_vars <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(pm_mean_sampling_period, income_3level, residents_smoke) %>% 
+  mutate_all(as.numeric) %>% 
+  na.exclude() %>% 
+  select(-pm_mean_sampling_period)
+
+# Use `omcdiag` to assess overall multicollinearity
+omcdiag(other_vars, pm_mean_sampling_period)
+
+# Use `imcdiag` to assess multicollinearity between individual variables
+imcdiag(other_vars, pm_mean_sampling_period)
+
+# Specify a method to assess
+imcdiag(other_vars, pm_mean_sampling_period, method = "Fi")
+```
+
+# Initial variables to use for model selection (these have less overlap):
+Demographics: income_3level, total_residents_2level, residents_smoke
+Home Characteristics: home_type_2level, home_floors_2level, stove_age_3level,
+                      chimney_clean_3level, home_mold, home_fireplace
+Home Activity: home_act_gas, home_act_smoking
+Others (make some categorical): amb_pm_24hr, mean_temp, sums_mean_sampling_day, 
+       moisture_closest, stove_grade_3level
+
+# Also try individual chi-square tests and correlations b/w two vars
+
+```{r}
+test_data <- exposure_analysis_data_medium %>% 
+  filter(winter_id == 1) %>% 
+  filter(!is.na(pm_mean_sampling_day))
+
+chisqu_results <- chisq.test(test_data$home_floors_2level, 
+                             test_data$home_windows_2level)
+chisqu_results
+
+cor_results <- cor.test(test_data$temp_indoor_max, 
+                        test_data$mean_temp)
+cor_results
+```
+
+If variables have very high overlap, then only use the one with the strongest
+association with PM for consideration in the model.
+
+Once variables are narrowed down, use best subsets selection (exposure_models.RMD) 
+to help inform a final model.
+
+Run sensitivity analyses to see if adding other variables back in makes a difference.

### 2020-04-13 ambient_summary.Rmd
update file path method/backup

@@
 author: "Ethan Walker"
-date: "Started 1 April 2020, Updated 3 April 2020"
+date: "Started 1 April 2020, Updated 13 April 2020"
 output: pdf_document
@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")
+input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+input_file <- c("exposure_analysis_data_long_new.rds")
+output_file <- c("exposure_analysis_data_long_new.rds")
 
-exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
+exposure_analysis_data_long <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
 
-exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")
+exposure_analysis_data_medium <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
 
-sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
+exposure_analysis_data_short <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))
 
-health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
+sums_rolling_pm_data <- 
+  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
 ```

### 2020-04-13 exposure_initial_analysis.Rmd
update file path method/backup

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 9 April 2020"
+date: "Started 2 Feb 2020, Updated 13 April 2020"
 output: pdf_document

### 2020-04-13 exposure_models.Rmd
update file path method/backup

--- 2020-03-24 exposure_models.Rmd
@@
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 20 March 2020"
+date: "Started 19 March 2020, Updated 13 April 2020"
 output: pdf_document
@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+input_file <- c("exposure_analysis_data_long_new.rds")
+output_file <- c("exposure_analysis_data_long_new.rds")
 
-exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds") 
+exposure_analysis_data_long <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
+
+exposure_analysis_data_medium <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+
+exposure_analysis_data_short <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))
+
+sums_rolling_pm_data <- 
+  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
 ```

### 2020-04-14 child_health_data_cleaning.Rmd
Update PM data

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 8 April 2020"
+date: "Started 27 Feb 2020, Updated 14 April 2020"
 output: pdf_document 
@@
 # Load data
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-demographics_clean <- read_rds("Output/demographics_clean.rds")
-pm_clean <- read_rds("Output/pm_clean.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_linked_ids <- read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
+demographics_clean <- 
+  read_rds(paste0(output_path, "Output/demographics_clean.rds"))
+pm_clean <- read_rds(paste0(output_path, "Output/pm_clean.rds"))
 ```
@@
 ```{r}
-kids_at_home_log <- read_rds("Output/kids_at_home_log.rds") %>% 
+kids_at_home_log <- 
+  read_rds(paste0(output_path, "Output/kids_at_home_log.rds")) %>% 
   arrange(area, home_winter_id, at_home_day)
@@
   group_by(area, home) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm),
-         pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
+  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
   group_by(area, home, sampling_day) %>% 
@@
   mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
+  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
+          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
   distinct(sample_time, .keep_all = TRUE) %>% 
@@
          sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
-         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day) %>% 
+         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
+         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
   left_join(kids_at_home_log_new, 
@@
   mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
+  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
+          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
   distinct(at_home, .keep_all = TRUE) %>% 
@@
 
-#write_rds(at_home_pm_merged, "Output/at_home_pm_merged.rds")
-#write_csv(at_home_pm_merged, "Output/at_home_pm_merged.csv")
+#write_rds(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.rds"))
+#write_csv(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.csv"))
 
-at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")
+at_home_pm_merged <- 
+  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
 ```

### 2020-04-14 dustrak_data_cleaning.Rmd
Update PM data

@@
 pm_time_plot_function <- function(id, location) {
-pm_time_plot <- kids_pm %>% 
+pm_time_plot <- kids_pm_new %>% 
   filter(home_winter_id == id & area == location) %>% 
@@
 
-gaps <<- kids_pm %>% 
+gaps <<- kids_pm_new %>% 
   filter(home_winter_id == id & area == location) %>% 
@@
 
-pm_desc <<- kids_pm %>% 
+pm_desc <<- kids_pm_new %>% 
   filter(home_winter_id == id & area == location) %>% 
@@
 # Load data 
-kids_pm <- read_rds("Output/kids_pm_new.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_pm_new <- 
+  read_rds(paste0(output_path, "Output/kids_pm_new.rds"))
 
@@
 # Run all of these lines together
-pm_time_plot_function(id = "129", location = "WMT")
+pm_time_plot_function(id = "273", location = "WMT")
 head(gaps, 25)
@@
 
-# Save cleaned dataset
+# Check high PM values
 ```{r}
-kids_pm <- read_rds("Output/kids_pm_new4.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_pm <- 
+  read_rds(paste0(output_path, "Output/kids_pm_new4.rds"))
+
+kids_pm_check <- kids_pm %>% 
+  filter(pm > 2000) %>% 
+  group_by(area) %>% 
+  count(home_winter_id)
+
+# 0.1% of the PM data is greater than 2000
+# Nothing stood out as being unreasonable when looking (plotting) at homes
+# with values over 2000. Plan to exclude values > 10,000, leave the rest
+# as is, and do sensitivity analyses excluding home with values over specific
+# thresholds. Also plan to see how PM in homes with PM spikes is associated with
+# health and behavioral variables.
+```
+
+# Join PM data with linked IDs and make new variables for spikes in PM
+```{r}
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_pm <- 
+  read_rds(paste0(output_path, "Output/kids_pm_new4.rds")) 
+kids_linked_ids <- 
+  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
 
@@
   select(-home, -treatment) %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  group_by(home, sampling_day) %>% 
+  mutate(pm_spike_1000 = if_else(pm > 1000, "Yes", "No"),
+         pm_spike_2000 = if_else(pm > 2000, "Yes", "No"),
+         pm_spike_5000 = if_else(pm > 5000, "Yes", "No")) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, pm_datetime_new)
+```
 
-# write_rds(kids_pm_new5, "Output/pm_clean.rds")
+# Save dataset
+```{r}
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+write_rds(kids_pm_new5, paste0(output_path, "Output/pm_clean.rds"))
 ```

### 2020-04-14 merge_datasets.Rmd
Update PM data

@@
 author: "Ethan Walker"
-date: "Started 24 March 2020, Updated 27 March 2020"
+date: "Started 24 March 2020, Updated 14 April 2020"
 output: pdf_document
@@
 ```{r}
-#pm_clean <- read_rds("Output/pm_clean.rds")
-#sums_clean <- read_rds("Output/sums_clean.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-demographics_clean <- read_rds("Output/demographics_clean.rds")
-moisture_clean <- read_rds("Output/moisture_clean.rds")
-indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
-stove_use_clean <- read_rds("Output/stove_use_clean.rds")
-stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
-home_char_clean <- read_rds("Output/home_char_clean.rds")
-home_act_clean <- read_rds("Output/home_act_clean.rds")
+pm_clean <- 
+  read_rds(paste0(output_path, "Output/pm_clean.rds"))
+sums_clean <- 
+  read_rds(paste0(output_path, "Output/sums_clean.rds"))
+
+kids_linked_ids <- 
+  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
+demographics_clean <- 
+  read_rds(paste0(output_path, "Output/demographics_clean.rds"))
+moisture_clean <- 
+  read_rds(paste0(output_path, "Output/moisture_clean.rds"))
+indoor_temp_rh_clean <- 
+  read_rds(paste0(output_path, "Output/indoor_temp_rh_clean.rds"))
+stove_use_clean <- 
+  read_rds(paste0(output_path, "Output/stove_use_clean.rds"))
+stove_grades_clean <- 
+  read_rds(paste0(output_path, "Output/stove_grades_clean.rds"))
+home_char_clean <- 
+  read_rds(paste0(output_path, "Output/home_char_clean.rds"))
+home_act_clean <- 
+  read_rds(paste0(output_path, "Output/home_act_clean.rds"))
 ```
@@
   ungroup() %>% 
-  arrange(area, home, sample_date) %>% 
+  arrange(area, home, sample_date, desc(pm_spike_1000),
+          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
   select(area, home, home_winter_id, winter_id, treatment, dtid, day_of_week, 
@@
          pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
-         sample_date, pm_mean_daily, pm_daily_obs) %>% 
+         sample_date, pm_mean_daily, pm_daily_obs, pm_spike_1000,
+         pm_spike_2000, pm_spike_5000) %>% 
   group_by(area, home) %>% 
@@
 
-#write_rds(pm_join, "Output/pm_join.rds")
-pm_join <- read_rds("Output/pm_join.rds")
+#write_rds(pm_join, paste0(output_path, "Output/pm_join.rds"))
+pm_join <- read_rds(paste0(output_path, "Output/pm_join.rds"))
 
@@
 
-#write_rds(sums_join, "Output/sums_join.rds")
-sums_join <- read_rds("Output/sums_join.rds")
+#write_rds(sums_join, paste0(output_path, "Output/sums_join.rds"))
+sums_join <- read_rds(paste0(output_path, "Output/sums_join.rds"))
 
@@
 
-#write_rds(pm_sums_joined1, "Output/pm_sums_joined1.rds")
-pm_sums_joined1 <- read_rds("Output/pm_sums_joined1.rds")
+#write_rds(pm_sums_joined1, paste0(output_path, "Output/pm_sums_joined1.rds"))
+pm_sums_joined1 <- read_rds(paste0(output_path, "Output/pm_sums_joined1.rds"))
 
@@
 
-#write_rds(analysis_data7, "Output/exposure_analysis_data_long.rds")
-#write_csv(analysis_data7, "Output/exposure_analysis_data_long.csv")
+#write_rds(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.rds"))
+#write_csv(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.csv"))
 
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
+exposure_analysis_data_long <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))
 
@@
 
-#write_csv(save_var_names, "Output/kids_data_var_names.csv")
+#write_csv(save_var_names, paste0(output_path, "Output/kids_data_var_names.csv"))
 ```
@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long.rds")
-kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+exposure_analysis_data_long <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))
+
+kids_amb_pm_data <- 
+  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
   mutate(sample_date = amb_pm_date) %>% 
   distinct(zip, .keep_all = TRUE)
-kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
+
+kids_amb_temp_data <- 
+  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
   mutate(sample_date = temp_date) %>% 
@@
   distinct(zip, .keep_all = TRUE)
-wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
+
+wmt_zipcodes <- read_csv(paste0(output_path, "Input/WMT/wmt_zipcodes.csv")) %>% 
   rename_all(tolower)
-nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
+
+nn_zipcodes <- read_csv(paste0(output_path, "Input/NN/nn_zipcodes.csv")) %>% 
   rename_all(tolower)
-ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
+
+ak_zipcodes <- read_csv(paste0(output_path, "Input/AK/ak_zipcodes.csv")) %>% 
   rename_all(tolower)
+
 kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
@@
   select(home, zip, lat_zip, long_zip, lat_pm, long_pm, lat_temp, long_temp)
-kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data_new.rds") %>% 
+
+kids_amb_pm_data <- 
+  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
   mutate(sample_date = amb_pm_date) %>% 
   select(-lat_zip, -long_zip, -lat_pm, -long_pm)
-kids_amb_temp_data <- read_rds("Output/kids_amb_temp_data_new.rds") %>% 
+
+kids_amb_temp_data <- 
+  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
   mutate(sample_date = temp_date) %>% 
@@
 
-#write_rds(exposure_analysis_data_long_new, "Output/exposure_analysis_data_long_new.rds")
-#write_csv(exposure_analysis_data_long_new, "Output/exposure_analysis_data_long_new.csv")
+#write_rds(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
+#write_csv(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.csv"))
 ```
@@
 ```{r}
-exposure_analysis_data_long_new <- read_rds("Output/exposure_analysis_data_long_new.rds")
+exposure_analysis_data_long_new <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
 
@@
 
-#write_rds(exposure_analysis_data_medium, "Output/exposure_analysis_data_medium.rds")
-#write_csv(exposure_analysis_data_medium, "Output/exposure_analysis_data_medium.csv")
+#write_rds(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+#write_csv(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.csv"))
 ```
@@
 ```{r}
-exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
+exposure_analysis_data_medium <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
 
@@
 
-#write_rds(exposure_analysis_data_short, "Output/exposure_analysis_data_short.rds")
-#write_csv(exposure_analysis_data_short, "Output/exposure_analysis_data_short.csv")
+#write_rds(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.rds"))
+#write_csv(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.csv"))
 ```
@@
 ```{r}
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+pm_clean <- 
+  read_rds(paste0(output_path, "Output/pm_clean.rds"))
+sums_clean <- 
+  read_rds(paste0(output_path, "Output/sums_clean.rds"))
+
 pm_join <- pm_clean %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
   mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
@@
 
-exposure_analysis_data_long_new <- read_rds("Output/exposure_analysis_data_long_new.rds")
+exposure_analysis_data_long_new <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
 
@@
 
-#write_rds(pm_sums_joined3, "Output/sums_rolling_pm_data.rds")
-#write_csv(pm_sums_joined3, "Output/sums_rolling_pm_data.csv")
+#write_rds(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+#write_csv(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.csv"))
 ```

### 2020-04-14 merge_health_datasets.Rmd
Update PM data

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 8 April 2020"
+date: "Started 27 March 2020, Updated 14 April 2020"
 output: pdf_document
@@
 ```{r}
-at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")
-kids_med_visit <- read_rds("Output/kids_med_visit.rds")
-kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds")
-child_home_winter <- read_rds("Output/child_home_winter.rds")
-kids_csq <- read_rds("Output/kids_csq.rds")
-kids_lrti <- read_rds("Output/kids_lrti.rds")
-kids_medical_records <- read_rds("Output/kids_medical_records.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+at_home_pm_merged <- 
+  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
+kids_med_visit <- 
+  read_rds(paste0(output_path, "Output/kids_med_visit.rds")) %>% 
+  mutate(visit_date = med_visit_date) 
+kids_birth_infancy <- 
+  read_rds(paste0(output_path, "Output/kids_birth_infancy.rds")) %>% 
+  select(-home_id)
+child_home_winter <- 
+  read_rds(paste0(output_path, "Output/child_home_winter.rds")) %>% 
+  select(-home_id.x, -home_id.y, -chw_symptom_free_date)
+kids_csq <- 
+  read_rds(paste0(output_path, "Output/kids_csq.rds"))
+kids_lrti <- 
+  read_rds(paste0(output_path, "Output/kids_lrti.rds")) %>% 
+  select(-home_id)
+kids_medical_records <- 
+  read_rds(paste0(output_path, "Output/kids_medical_records.rds")) %>% 
+  select(-visit_date)
 ```
@@
 ```{r}
-kids_med_visit <- read_rds("Output/kids_med_visit.rds") %>% 
-  mutate(visit_date = med_visit_date)
-
 health_merged1 <- kids_csq %>% 
@@
 
-child_home_winter <- read_rds("Output/child_home_winter.rds") %>% 
-  select(-home_id.x, -home_id.y, -chw_symptom_free_date)
-
 health_merged2 <- health_merged1 %>% 
@@
   ungroup()
-
-kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds") %>% 
-  select(-home_id)
 
@@
   ungroup()
-
-at_home_pm_merged <- read_rds("Output/at_home_pm_merged.rds")
 
@@
 
-kids_medical_records <- read_rds("Output/kids_medical_records.rds") %>% 
-  select(-visit_date)
-
 health_merged5 <- health_merged4 %>% 
@@
   ungroup()
-
-kids_lrti <- read_rds("Output/kids_lrti.rds") %>% 
-  select(-home_id)
 
@@
 
-#write_rds(health_merged6, "Output/health_data_merged.rds")
-#write_csv(health_merged6, "Output/health_data_merged.csv")
+#write_rds(health_merged6, paste0(output_path, "Output/health_data_merged.rds"))
+#write_csv(health_merged6, paste0(output_path, "Output/health_data_merged.csv"))
 ```
@@
 ```{r}
-exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-health_data_merged <- read_rds("Output/health_data_merged.rds")
+exposure_analysis_data_medium <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+
+health_data_merged <- 
+  read_rds(paste0(output_path, "Output/health_data_merged.rds"))
 
@@
 
-#write_rds(health_exposure_data_sampling_day, "Output/health_exposure_data_sampling_day.rds")
-#write_csv(health_exposure_data_sampling_day, "Output/health_exposure_data_sampling_day.csv")
+#write_rds(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
+#write_csv(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.csv"))
 
-health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
 

### 2020-04-14 pm_summary.Rmd
Update PM data

--- 2020-02-06 pm_summary.Rmd
@@
 author: "Ethan Walker"
-date: "Started 22 Nov 2019, Updated 2 Feb 2020"
+date: "Started 22 Nov 2019, Updated April 14 2020"
 output: pdf_document
+header-includes:
+    - \usepackage[labelformat=empty]{caption}
 ---
@@
 library(lubridate)
+library(kableExtra)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 # Load data
-pm_clean <- read_rds("Output/pm_clean.rds")
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+pm_clean <- read_rds(paste0(output_path, "Output/pm_clean.rds"))
 ```
 
+```{r}
 # Summary stats
-```{r}
-pm_summary <- pm_clean %>% 
+
+pm_500 <- pm_clean %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  #filter(pm < 500) %>% 
   mutate(pm = if_else(pm > 500, 500, pm)) %>% 
@@
   summarize("n" = n(),
-            "Mean PM" = mean(pm, na.rm = TRUE), 
-            "SD PM" = sd(pm, na.rm = TRUE),
-            "Min PM" = min(pm, na.rm = TRUE), 
-            "25%" = quantile(pm, 0.25, na.rm = TRUE),
-            "Median PM" = median(pm, na.rm = TRUE),
-            "75%" = quantile(pm, 0.75, na.rm = TRUE),
-            "Max PM" = max(pm, na.rm = TRUE),
-            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
-            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
-            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
-            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
-  #select(-treatment) %>% 
-  t()
+            "Mean PM" = round(mean(pm, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm, na.rm = TRUE)),
+            "Min PM" = round(min(pm, na.rm = TRUE)), 
+            "25%" = round(quantile(pm, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm, na.rm = TRUE)),
+            "75%" = round(quantile(pm, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm, na.rm = TRUE)))
 
-kable(pm_summary, 
-      digits = 1, align = "c")
+pm_2000 <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  mutate(pm = if_else(pm > 2000, 2000, pm)) %>% 
+  group_by(area) %>% 
+  summarize("n" = n(),
+            "Mean PM" = round(mean(pm, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm, na.rm = TRUE)),
+            "Min PM" = round(min(pm, na.rm = TRUE)), 
+            "25%" = round(quantile(pm, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm, na.rm = TRUE)),
+            "75%" = round(quantile(pm, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm, na.rm = TRUE)))
 
+pm_summary <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area) %>% 
+  summarize("n" = n(),
+            "Mean PM" = round(mean(pm, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm, na.rm = TRUE)),
+            "Min PM" = round(min(pm, na.rm = TRUE)), 
+            "25%" = round(quantile(pm, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm, na.rm = TRUE)),
+            "75%" = round(quantile(pm, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm, na.rm = TRUE)))
+
+full_summary <- rbind(pm_summary, pm_2000, pm_500)
+
+kable(full_summary, 
+      caption = "Kids Indoor PM2.5 (Winter 1 Placebo)") %>% 
+   kable_styling() %>% 
+   pack_rows("No PM2.5 Cutoff", 1, 3) %>% 
+   pack_rows("PM2.5 Cutoff: 2000", 4, 6) %>% 
+   pack_rows("PM2.5 Cutoff: 500", 7, 9) 
+```
+
+\clearpage
+
+```{r}
 pm_by_day <- pm_clean %>% 
@@
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  #filter(pm < 500) %>% 
-  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
   group_by(area, day_of_week) %>% 
@@
             "Max PM" = max(pm, na.rm = TRUE))
-pm_by_day
+kable(pm_by_day, 
+      caption = "Kids Indoor PM2.5 by day of week (Winter 1 Placebo)", digits = 0)
 ```
 
+\clearpage
+
 # Boxplots for PM by study area
-```{r, fig.height=6, fig.width=5}
+```{r, fig.height=6, fig.width=8}
 boxplots_kids_pm <- pm_clean %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(pm < 500) %>% 
   ggplot(aes(area, pm)) +
@@
                fatten = 1, outlier.size = 1.5, width = 0.35) +
-    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
     theme_minimal() +
-    ylim(0, 500) +
-    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
+    #ylim(0, 500) +
     labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
-    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
     theme(axis.text.x = element_text(size = 16, colour = "black"),
@@
 ```
+
+\pagebreak
+
+```{r, fig.height=6, fig.width=8}
+boxplots_kids_pm <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  ggplot(aes(area, log(pm))) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
+                 position=position_nudge(x = -0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
+                 size=7, position=position_nudge(x = -0.35, y = 0)) +
+    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
+                 position=position_nudge(x = 0.1, y = 0)) +
+    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
+                 size=7, position=position_nudge(x = 0.35, y = 0)) +
+    theme_minimal() +
+    #ylim(0, 500) +
+    labs(y = expression(paste("Log PM"[2.5], " (", mu, g/m^3, ")"))) +
+    theme(axis.text.x = element_text(size = 16, colour = "black"),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_kids_pm
+```

### 2020-04-16 exposure_initial_analysis.Rmd
exposure analysis update

@@
 output: pdf_document
+header-includes:
+    - \usepackage[labelformat=empty]{caption}
 ---
@@
 library(lmerTest)
-library(ggmap)
 library(mctest)
-library(GGally)
 library(tidyverse)
+library(knitr)
+library(kableExtra)
 ```
@@
   distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
   ungroup()
@@
 # Variables to use:
-## education_3level, income_3level, race_parent_2level, race_child_2level
+## education_3level, income_3level, race_parent_2level, 
 ## total_residents_2level, residents_under5_2level, resident_smokes
@@
 
-pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ residents_smoke, 
+pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ education_3level, 
                               data = analysis_data)
+
 summary(pm_demographics_results)
-tidy(pm_demographics_results)
+glance(pm_demographics_results)
+
+tidy(pm_demographics_results) %>% 
+    filter(term != '(Intercept)') %>% 
+    mutate(estimate = round(estimate, digits = 2),
+           'p-value' = round(p.value, digits = 2),
+           term = gsub("education_3level", "Education:  ", term)) %>% 
+    select(term, estimate, 'p-value')
+
 plot(pm_demographics_results)
@@
 pm_stove_grades_plot
+```
+
+```{r}
+analysis_function <- function(covar, label, 
+                              dataset = exposure_analysis_data_short){
+
+  analysis_data <- dataset %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  rename(analysis_covar = covar) %>% 
+  ungroup() %>% 
+  select(pm_mean_sampling_period, analysis_covar)
+
+  analysis_results <- lm(log(pm_mean_sampling_period) ~ analysis_covar, analysis_data)
+
+  tidy_results <<- tidy(analysis_results) %>% 
+    filter(term != '(Intercept)') %>% 
+    mutate(estimate = round(estimate, digits = 2),
+           'p-value' = round(p.value, digits = 2),
+           term = gsub("analysis_covar", paste0(label, ": "), term)) %>% 
+    select(term, estimate, 'p-value')
+
+}
+
+d1 <- analysis_function("education_3level", "Education")
+d2 <- analysis_function("income_3level", "Income")
+d3 <- analysis_function("race_parent_2level", "Parent race")
+d4 <- analysis_function("total_residents_2level", "Total residents")
+d5 <- analysis_function("residents_under5_2level", "Residents under 5 y/o")
+d6 <- analysis_function("residents_smoke", "Residents smoke")
+
+d7 <- analysis_function("home_type_2level", "Home type")
+d8 <- analysis_function("home_year_built_2level", "Home age")
+d9 <- analysis_function("home_sqft_2level", "home sqft")
+d10 <- analysis_function("home_floors_2level", "Home floors")
+d11 <- analysis_function("home_bedrooms_2level", "Home bedrooms")
+d12 <- analysis_function("home_windows_2level", "Home windows")
+d13 <- analysis_function("home_pets_2level", "Pets - total")
+d14 <- analysis_function("home_furry_2level", "Pets - furry")
+d15 <- analysis_function("secondary_heat_3level", "Secondary heating")
+d16 <- analysis_function("stove_age_3level", "Stove age")
+d17 <- analysis_function("stove_cert", "Stove EPA certified")
+d18 <- analysis_function("chimney_clean_3level", "Chimney cleaned")
+d19 <- analysis_function("home_damp", "Home - damp")
+d20 <- analysis_function("home_mold", "Home - mold")
+d21 <- analysis_function("home_fireplace", "Home fireplace")
+d22 <- analysis_function("home_furnace", "Home furnace")
+
+d23 <- analysis_function("burn_level_3level", "Burn level")
+d24 <- analysis_function("wood_collect_2level", "Wood collect time")
+d25 <- analysis_function("wood_collect_method_2level", "Wood collect method")
+d26 <- analysis_function("wood_cords", "Wood cords")
+
+table_data <- rbind(d1, d2, d3, d4, d5, d6,
+                    d7, d8, d9, d10, d11, d12, d13, d14, d15, d16, d17, d18,
+                    d19, d20, d21, d22,
+                    d23, d24, d25, d26)
+
+kable(table_data, 
+      caption = "Indoor PM2.5 (log) and covariate associations") %>% 
+   kable_styling() %>% 
+   row_spec(0, bold = TRUE) %>% 
+   pack_rows("Demographic variables", 1, 8) %>%
+   pack_rows("Home characteristics", 9, 31) %>%
+   pack_rows("Stove use", 32, 36) %>%
+   footnote(c("PM2.5 = fine particulate matter"))
 ```

### 2020-04-17 collinearity_exposure_analysis.Rmd
exposure analysis updates

---
title: "KidsAIR: multicollinearity exposure analysis"
author: "Ethan Walker"
date: "Started 16 April 2020, Updated 16 April 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, include = FALSE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r, eval=TRUE, include=TRUE}
# Load individual datasets

input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
input_file <- c("exposure_analysis_data_long_new.rds")
output_file <- c("exposure_analysis_data_long_new.rds")

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))

exposure_analysis_data_short <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
```

```{r, eval=TRUE, include=TRUE}
# Prep and merge data

####### Format SUMs data to merge with analysis data below
sums_temp_change <- sums_rolling_pm_data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, sampling_day) %>% 
  distinct(datetime_sums, .keep_all = TRUE) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
                                            breaks = c(0, 2, 50),
                           labels = c("<2", "2+"),
                                 right = FALSE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
         time_diff:sums_events_sampling_day_cat) %>% 
  arrange(area, home, sample_date) %>% 
  group_by(home) %>% 
  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(area == "AK") %>% 
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(home, area, sampling_day, starts_with("home_act"))

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  left_join(sums_temp_change, by = c("area", "home")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home")) %>% 
  mutate(area = as.factor(area))
```

# Run tests for multicollinearity between different covariates
Goal is to check for vars that have a lot of overlap and eliminate some from
consideration to include in models

First, use `mctest` package to check for multicollinearity among similar variables

https://cran.r-project.org/web/packages/mctest/mctest.pdf

Notes on multicollinearity:

https://datascienceplus.com/multicollinearity-in-r/

https://www.theanalysisfactor.com/eight-ways-to-detect-multicollinearity/

```{r, eval=TRUE, include=TRUE}
# Filter data for specific variables; select pm variable
pm_mean_sampling_period <- analysis_data %>% 
  select(pm_mean_sampling_period, area, income_3level, 
         residents_smoke,
         home_floors_2level, 
         stove_age_3level, chimney_clean_3level, home_mold,
         wood_collect_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         home_act_smoking, 
         home_act_door, home_act_sweep,
         stove_grade_3level) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(pm_mean_sampling_period)

# Filter data for specific variables; select predictor variables
other_vars <- analysis_data %>% 
  select(pm_mean_sampling_period, area, income_3level, 
         residents_smoke,
         home_floors_2level, 
         stove_age_3level, chimney_clean_3level, home_mold,
         wood_collect_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         home_act_smoking, 
         home_act_door, home_act_sweep,
         stove_grade_3level) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(-pm_mean_sampling_period)

# Use `omcdiag` to assess overall multicollinearity
omcdiag(other_vars, pm_mean_sampling_period)

# Use `imcdiag` to assess multicollinearity between individual variables
imcdiag(other_vars, pm_mean_sampling_period)

# Specify a method to assess
imcdiag(other_vars, pm_mean_sampling_period, method = "VIF")
```

Initially assessed the following variables for collinearity, based on simple
associations with mean PM over the sampling period:
         pm_mean_sampling_period, area, income_3level, total_residents_2level,
         race_parent_2level, residents_smoke,
         home_type_2level, home_floors_2level, home_windows_2level,
         stove_age_3level, chimney_clean_3level, home_damp, home_mold,
         home_fireplace,
         wood_collect_2level,
         moisture_closest_2level,
         temp_max_5,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         home_act_gas, home_act_smoking, home_act_windows, 
         home_act_door, home_act_sweep,
         stove_grade_3level

The goal is to remove variables that have overlap, and parse the covariates
down to a reasonable number that a best-subsets selection method can handle.

Based on results and potential overlap, the following vars were removed:
         race_parent_2level, home_damp, moisture_closest_2level,
         temp_max_5, home_act_gas, total_residents_2level, home_type_2level, 
         home_windows_2level, home_fireplace, home_act_windows

With the above variables removed, the VIF is at 5 or below for all variables.
There is still a question of how to handle the "area" variable, but I will
run best-subsets next and go from there.

Initial variables to use for model selection (14):
Demographics: area, income_3level, residents_smoke
Home Characteristics: home_floors_2level, stove_age_3level,
                      chimney_clean_3level, home_mold, wood_collect_2level
Home Activity: home_act_door, home_act_smoking, home_act_sweep
Others: sums_events_sampling_period, mean_temp_sampling_period_5, stove_grade_3level

Run sensitivity analyses to see if adding other variables back in makes a difference.

Can also try individual chi-square tests and correlations b/w two vars:

```{r}
test_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(pm_mean_sampling_day))

chisqu_results <- chisq.test(test_data$home_floors_2level, 
                             test_data$home_windows_2level)
chisqu_results

cor_results <- cor.test(test_data$temp_indoor_max, 
                        test_data$mean_temp)
cor_results
```

### 2020-04-17 exposure_covariate_associations.Rmd
exposure analysis updates

---
title: "KidsAIR: covariate associations exposure analysis"
author: "Ethan Walker"
date: "Started 2 Feb 2020, Updated 16 April 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, include = FALSE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r, eval=TRUE, include=TRUE}
# Load individual datasets

input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
input_file <- c("exposure_analysis_data_long_new.rds")
output_file <- c("exposure_analysis_data_long_new.rds")

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))

exposure_analysis_data_short <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
```

```{r, eval=TRUE, include=TRUE}
####### Format SUMs data to merge with analysis data below
sums_temp_change <- sums_rolling_pm_data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, sampling_day) %>% 
  distinct(datetime_sums, .keep_all = TRUE) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
                                            breaks = c(0, 2, 50),
                           labels = c("<2", "2+"),
                                 right = FALSE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
         time_diff:sums_events_sampling_day_cat) %>% 
  arrange(area, home, sample_date) %>% 
  group_by(home) %>% 
  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(area == "AK") %>% 
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(home, area, sampling_day, starts_with("home_act"))

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  left_join(sums_temp_change, by = c("area", "home")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home"))

####### Function to run analysis
analysis_function <- function(covar, label, 
                              dataset = analysis_data){

  analysis_data <- dataset %>% 
    filter(winter_id == 1 & treatment == "Placebo") %>% 
    group_by(area, home) %>% 
    distinct(home, .keep_all = TRUE) %>% 
    rename(analysis_covar = covar) %>% 
    ungroup() %>% 
    select(pm_mean_sampling_period, analysis_covar)

  analysis_results <- lm(log(pm_mean_sampling_period) ~ analysis_covar, analysis_data)

  tidy_results <<- tidy(analysis_results) %>% 
    filter(term != '(Intercept)') %>% 
    mutate(estimate = round(estimate, digits = 2),
           'p-value' = round(p.value, digits = 2),
           term = gsub("analysis_covar", paste0(label, ": "), term)) %>% 
    select(term, estimate, 'p-value')

}

####### Variables to run through fuction and save results
d1 <- analysis_function("education_3level", "Education")
d2 <- analysis_function("income_3level", "Income")
d3 <- analysis_function("race_parent_2level", "Parent race")
d4 <- analysis_function("total_residents_2level", "Total residents")
d5 <- analysis_function("residents_under5_2level", "Residents under 5 y/o")
d6 <- analysis_function("residents_smoke", "Residents smoke")

d7 <- analysis_function("home_type_2level", "Home type")
d8 <- analysis_function("home_year_built_2level", "Home age")
d9 <- analysis_function("home_sqft_2level", "home sqft")
d10 <- analysis_function("home_floors_2level", "Home floors")
d11 <- analysis_function("home_bedrooms_2level", "Home bedrooms")
d12 <- analysis_function("home_windows_2level", "Home windows")
d13 <- analysis_function("home_pets_2level", "Pets - total")
d14 <- analysis_function("home_furry_2level", "Pets - furry")
d15 <- analysis_function("secondary_heat_3level", "Secondary heating")
d16 <- analysis_function("stove_age_3level", "Stove age")
d17 <- analysis_function("stove_cert", "Stove EPA certified")
d18 <- analysis_function("chimney_clean_3level", "Chimney cleaned")
d19 <- analysis_function("home_damp", "Home - damp")
d20 <- analysis_function("home_mold", "Home - mold")
d21 <- analysis_function("home_fireplace", "Home fireplace")
d22 <- analysis_function("home_furnace", "Home furnace")

d23 <- analysis_function("burn_level_3level", "Burn level")
d24 <- analysis_function("wood_collect_2level", "Wood collect time")
d25 <- analysis_function("wood_collect_method_2level", "Wood collect method")
d26 <- analysis_function("wood_cords", "Wood cords")

d27 <- analysis_function("moisture_closest", "Wood moisture - closest to sampling")
d28 <- analysis_function("moisture_closest_2level", "Wood moisture - closest to sampling")
d29 <- analysis_function("mean_moisture_winter", "Wood moisture - Winter mean")
d30 <- analysis_function("moisture_winter_2level", "Wood moisture - Winter mean")

d31 <- analysis_function("temp_min_5", "Minimum indoor temp - per 5 degree increase")
d32 <- analysis_function("temp_max_5", "Maximum indoor temp - per 5 degree increase")
d33 <- analysis_function("rh_min_5", "Minimum indoor RH - per 5% increase")
d34 <- analysis_function("rh_max_5", "Maximum indoor RH - per 5% increase")
d35 <- analysis_function("temp_min_2level", "Minimum indoor temp")
d36 <- analysis_function("temp_max_2level", "Maximum indoor temp")
d37 <- analysis_function("rh_min_2level", "Minimum indoor RH")
d38 <- analysis_function("rh_max_2level", "Maximum indoor RH")

d39 <- analysis_function("sums_mean_sampling_period_5", "SUMs mean temperature")
d40 <- analysis_function("temp_perc_25_2level", "SUMs - % obs > 25 degrees C")
d41 <- analysis_function("temp_perc_27_2level", "SUMs - % obs > 27 degrees C")
d42 <- analysis_function("temp_perc_30_2level", "SUMs - % obs > 30 degrees C")
d43 <- analysis_function("sums_events_sampling_period", 
                         "Mean daily SUMs events - 5+ degree C temp increase in 20 min")

d44 <- analysis_function("amb_pm_sampling_period_5", "Mean ambient PM2.5 - per 5ug/m3 increase")
d45 <- analysis_function("mean_temp_sampling_period_5", 
                         "Mean ambient temp - per 5 degree increase")

d46 <- analysis_function("home_act_gas", "Home activity - gas appliance")
d47 <- analysis_function("home_act_elect", "Home activity - electric appliance")
d48 <- analysis_function("home_act_propane", "Home activity - propane appliance")
d49 <- analysis_function("home_act_oil", "Home activity - oil appliance")
d50 <- analysis_function("home_act_other", "Home activity - other appliance")
d51 <- analysis_function("home_act_smoking", "Home activity - smoking")
d52 <- analysis_function("home_act_incense", "Home activity - incense")
d53 <- analysis_function("home_act_candle", "Home activity - candle")
d54 <- analysis_function("home_act_windows", "Home activity - windows open")
d55 <- analysis_function("home_act_door", "Home activity - doors open")
d56 <- analysis_function("home_act_sweep", "Home activity - cleaning")

d57 <- analysis_function("stove_grade_3level", "Stove grade")

####### Combine results to use in a table
table_data <- rbind(d1, d2, d3, d4, d5, d6,
                    d7, d8, d9, d10, d11, d12, d13, d14, d15, d16, d17, d18,
                    d19, d20, d21, d22,
                    d23, d24, d25, d26,
                    d27, d28, d29, d30,
                    d31, d32, d33, d34, d35, d36, d37, d38,
                    d39, d40, d41, d42, d43,
                    d44, d45,
                    d46, d47, d48, d49, d50, d51, d52, d53, d54, d55, d56,
                    d57)

####### Print full table of results
kable(table_data, longtable = TRUE,
      caption = "Indoor PM2.5 (natural log transformed) and covariate associations") %>% 
   kable_styling(latex_options = "repeat_header") %>% 
   row_spec(0, bold = TRUE) %>% 
   pack_rows("Demographic variables", 1, 8) %>%
   pack_rows("Home characteristics", 9, 31) %>%
   pack_rows("Stove use", 32, 36) %>%
   pack_rows("Wood moisture content", 37, 40) %>%
   pack_rows("Indoor temperature and relative humidity", 41, 48) %>%
   pack_rows("Stove Use Monitors (SUMs) temperature", 49, 53) %>%
   pack_rows("Ambient data", 54, 55) %>%
   pack_rows("Home activities - Sum of all sampling days", 56, 66) %>%
   pack_rows("Stove grade", 67, 68) %>%
   footnote(c("PM2.5 = fine particulate matter", 
              "PM2.5 and all covariates have been averaged or summed to produce 1 observation per sampling period"))
```

```{r}
###### Chunks below are useful for checking individual associations

# Demographics

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  ungroup()

# Variables to use:
## education_3level, income_3level, race_parent_2level, 
## total_residents_2level, residents_under5_2level, resident_smokes

table(analysis_data$residents_smoke)
median(analysis_data$home_furry, na.rm = TRUE)

pm_demographics_results <- lm(log(pm_mean_sampling_period) ~ education_3level, 
                              data = analysis_data)

summary(pm_demographics_results)
glance(pm_demographics_results)

tidy(pm_demographics_results) %>% 
    filter(term != '(Intercept)') %>% 
    mutate(estimate = round(estimate, digits = 2),
           'p-value' = round(p.value, digits = 2),
           term = gsub("education_3level", "Education:  ", term)) %>% 
    select(term, estimate, 'p-value')

plot(pm_demographics_results)
```

```{r}
# Home characteristics

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## home_type_2level, home_year_built_2level, home_sqft_2level, home_floors_2level
## home_bedrooms_2level, home_windows_2level, home_pets_2level, home_furry_2level
## secondary_heat_3level, stove_age_3level, stove_cert, chimney_clean_3level,
## home_damp, home_mold, home_fireplace, home_furnace, home_insert

table(analysis_data$home_insert)
median(analysis_data$home_furry, na.rm = TRUE)

pm_home_char_results <- lm(log(pm_mean_sampling_period) ~ home_insert, 
                           data = analysis_data)
summary(pm_home_char_results)
tidy(pm_home_char_results)
plot(pm_home_char_results)
```

```{r}
# Stove use

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## burn_level_3level, wood_collect_2level, wood_collect_method_2level,
## wood_cords

table(analysis_data$burn_level_3level)
median(analysis_data$wood_cords, na.rm = TRUE)

pm_stove_use_results <- lm(log(pm_mean_sampling_period) ~ wood_collect_method_2level, 
                           data = analysis_data)
summary(pm_stove_use_results)
tidy(pm_stove_use_results)
plot(pm_stove_use_results)
pm_stove_use_plot <- pm_stove_use %>% 
  ggplot() + 
    geom_boxplot(aes(burn_level, mean_pm))
pm_stove_use_plot
```

```{r}
# Stove grades

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## stove_grade_3level

table(analysis_data$stove_grade_3level)

pm_stove_grades_results <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level, 
                              data = analysis_data)
summary(pm_stove_grades_results)
tidy(pm_stove_grades_results)
anova(pm_stove_grades_results)
plot(pm_stove_grades_results)
pm_stove_grades_plot <- pm_stove_grades %>% 
  ggplot() + 
    geom_boxplot(aes(stove_grade, mean_pm))
pm_stove_grades_plot
```

```{r}
# Home activity

analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup()

# Variables to use:
## home_act_gas, home_act_elect, home_act_propane, home_act_oil, home_act_other,
## home_act_smoke, home_act_incense, home_act_candle, home_act_windows,
## home_act_door, home_act_sweep

table(analysis_data$home_act_sweep)

pm_home_act_results <- lmer(log(pm_mean_sampling_day) ~ home_act_sweep + 
                              (1 | home), data = analysis_data)
summary(pm_home_act_results)
tidy(pm_home_act_results)
plot(pm_home_act_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_home_act_results), main = "QQ Plot")
```

```{r}
# Wood moisture content

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE))

# Variables to use:
## moisture_closest, moisture_closest_2level
## mean_moisture_winter, moisture_winter_2level

table(analysis_data$moisture_closest_2level)
median(analysis_data$mean_moisture_winter, na.rm = TRUE)

pm_moisture_results <- lm(log(pm_mean_sampling_period) ~ 
                            moisture_winter_2level, data = analysis_data)
summary(pm_moisture_results)
plot(pm_moisture_results)
```

```{r}
# Indoor temp/RH

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE))

# Variables to use:
## rh_min_5, rh_max_5, temp_min_5, temp_max_5
## temp_max_2level, temp_min_2level,rh_max_2level ,rh_min_2level

table(analysis_data$temp_max_2level)
median(analysis_data$rh_indoor_min, na.rm = TRUE)

pm_temp_rh_results <- lm(log(pm_mean_sampling_period) ~ temp_max_2level, 
                         data = analysis_data)
summary(pm_temp_rh_results)
tidy(pm_temp_rh_results)
plot(pm_temp_rh_results)
```

```{r}
# SUMs - by sampling day

analysis_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_day_5 = sums_mean_sampling_day/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_day_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_30_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lmer(log(pm_mean_sampling_day) ~ temp_perc_30_2level + 
                              (1 | home), data = analysis_data)
pm_sums_results <- lmer(pm_mean_sampling_day ~ temp_perc_30_2level + 
                              (1 | home), data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

```{r}
# SUMs - by sampling period

analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  ungroup()

# Variables to use:
## sums_mean_sampling_period_5, temp_perc_25_2level
## temp_perc_27_2level, temp_perc_30_2level

table(analysis_data$temp_perc_25_2level)
median(analysis_data$temp_perc_30, na.rm = TRUE)

pm_sums_results <- lm(log(pm_mean_sampling_period) ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
pm_sums_results <- lm(pm_mean_sampling_period ~ sums_mean_sampling_period_5, 
                      data = analysis_data)
summary(pm_sums_results)
tidy(pm_sums_results)
plot(pm_sums_results, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_results), main = "QQ Plot")
```

```{r}
# SUMs - mean PM per SUMs interval

analysis_data <- sums_rolling_pm_data %>% 
  mutate(temp_c_sums_5 = temp_c_sums/5)

cor.test(analysis_data$temp_c_sums, analysis_data$pm_rolling_20, method = "spearman")
pm_sums_model <- lmer(log(pm_rolling_20) ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)
pm_sums_model <- lmer(pm_rolling_20 ~ temp_c_sums_5 + (1 | home), 
                      data = analysis_data)

summary(pm_sums_model)
tidy(pm_sums_model)
exp((0.1496794)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- analysis_data %>% 
  ggplot() + 
    geom_point(aes(log(pm_rolling_20), temp_c_sums)) +
    geom_smooth(aes(log(pm_rolling_20), temp_c_sums)) +
    theme_minimal()
pm_sums_plot
```

```{r}
# SUMs - events per sampling day

sums_temp_change <- sums_rolling_pm_data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home, sampling_day) %>% 
  distinct(datetime_sums, .keep_all = TRUE) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
  replace_with_na(replace = list(lead_temp = 999)) %>% 
  mutate(temp_diff = lead_temp - temp_c_sums,
         # change temp in the following line
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check),
         heat_event = if_else(is.na(temp_diff_check), 1,
                              heat_event)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
                                            breaks = c(0, 2, 50),
                           labels = c("<2", "2+"),
                                 right = FALSE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
         time_diff:sums_events_sampling_day_cat) %>% 
  arrange(area, home, sample_date)

median(sums_temp_change$sums_events, na.rm = TRUE)
table(sums_temp_change$sums_events_sampling_day_cat)

pm_sums_model <- lmer(pm_mean_sampling_day ~ sums_events_sampling_day_cat + 
                        (1 | home), data = sums_temp_change)
pm_sums_model <- lmer(log(pm_mean_sampling_day) ~ sums_events_sampling_day_cat + 
                        (1 | home), data = sums_temp_change)

summary(pm_sums_model)
tidy(pm_sums_model)
#exp((0.146291)-1)*100
plot(pm_sums_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_sums_model), main = "QQ Plot")

pm_sums_plot <- sums_temp_change %>% 
  filter(!is.na(sums_events)) %>% 
  ggplot() + 
    geom_boxplot(aes(sums_events_cat, log(pm_mean_sampling_day))) +
    theme_minimal()
pm_sums_plot
```

```{r}
# Ambient temp/PM - by sampling date

analysis_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(area == "AK") %>% 
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  mutate(amb_pm_5 = amb_pm_24hr/5,
         mean_temp_5 = mean_temp/5,
         mean_temp_cat = cut(mean_temp, breaks = c(0, 30, 150),
                           labels = c("<30", "30+"),
                                 right = FALSE),
         amb_pm_cat = cut(amb_pm_24hr, breaks = c(0, 7, 100),
                           labels = c("<7", "7+"),
                                 right = FALSE))

median(analysis_data$amb_pm_24hr, na.rm = TRUE)
table(analysis_data$amb_pm_cat)

# Variables to use:
## amb_pm_5, amb_pm_cat,
## amb_pm_24hr, amb_pm_24hr_lag1, amb_pm_24hr_lag2, amb_pm_24hr_lag3
## amb_pm_roll_mean_2day, amb_pm_roll_mean_3day, amb_pm_roll_mean_4day
## mean_temp_5, mean_temp_cat,
## mean_temp, mean_temp_lag1, mean_temp_lag2, mean_temp_lag3
## mean_temp_roll_2day, mean_temp_roll_3day, mean_temp_roll_4day

cor.test(analysis_data$mean_temp, analysis_data$pm_mean_daily, method = "spearman")
pm_ambient_model <- lmer(log(pm_mean_daily) ~ amb_pm_cat + (1 | home), 
                      data = analysis_data)
pm_ambient_model <- lmer(pm_mean_daily ~ amb_pm_cat + (1 | home), 
                      data = analysis_data)

#summary(pm_ambient_model)
tidy(pm_ambient_model)
exp((-0.06723228)-1)*100
plot(pm_ambient_model, 
     xlab = "fitted values", ylab = "residuals")
qqnorm(residuals(pm_ambient_model), main = "QQ Plot")

pm_sums_plot <- analysis_data %>% 
  ggplot() + 
    geom_point(aes(log(pm_mean_daily), mean_temp)) +
    geom_smooth(aes(log(pm_mean_daily), mean_temp)) +
    theme_minimal()
pm_sums_plot
```

### 2020-04-20 child_health_data_cleaning.Rmd
child data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 14 April 2020"
+date: "Started 27 Feb 2020, Updated 20 April 2020"
 output: pdf_document 
@@
 # Load data
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-kids_linked_ids <- read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 demographics_clean <- 
-  read_rds(paste0(output_path, "Output/demographics_clean.rds"))
-pm_clean <- read_rds(paste0(output_path, "Output/pm_clean.rds"))
+  read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
 ```
@@
 ```{r}
-wmt_athomelog_initial <- read_xlsx("Input/WMT/wmt_athomelog_20200227.xlsx") %>% 
+wmt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_athomelog_20200227.xlsx")) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
@@
 
-nn_athomelog_initial <- read_xlsx("Input/NN/nn_athomelog_20200227.xlsx") %>% 
+nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_athomelog_20200416.xlsx")) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
@@
 
-ak_athomelog_initial <- read_xlsx("Input/AK/ak_athomelog_20200227.xlsx") %>% 
+ak_athomelog_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_athomelog_20200227.xlsx")) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
@@
 kids_at_home_log <- 
-  read_rds(paste0(output_path, "Output/kids_at_home_log.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_at_home_log.rds")) %>% 
   arrange(area, home_winter_id, at_home_day)
@@
 
-#write_rds(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.rds"))
-#write_csv(at_home_pm_merged, paste0(output_path, "Output/at_home_pm_merged.csv"))
+#write_rds(at_home_pm_merged, paste0(file_path, "Output/at_home_pm_merged.rds"))
+#write_csv(at_home_pm_merged, paste0(file_path, "Output/at_home_pm_merged.csv"))
 
 at_home_pm_merged <- 
-  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
+  read_rds(paste0(file_path, "Output/at_home_pm_merged.rds"))
 ```
@@
 
-nn_medicalvisit_initial <- read_xlsx("Input/NN/nn_medicalvisit_20200227.xlsx") %>%
+nn_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200416.xlsx")) %>%
   mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
@@
   spread(sample, value) %>%
+  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
   select(home_winter_id:med_visit_mean_temp)
 
-ak_medicalvisit_initial <- read_xlsx("Input/AK/ak_medicalvisit_20200227.xlsx") %>% 
+ak_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_medicalvisit_20200416.xlsx")) %>% 
   mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
@@
   spread(sample, value) %>%
+  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
   select(home_winter_id:med_visit_mean_temp)
@@
 
-# write_rds(kids_med_visit, "Output/kids_med_visit.rds")
+# write_rds(kids_med_visit, paste0(file_path, "Output/kids_med_visit.rds"))
 
-kids_med_visit <- read_rds("Output/kids_med_visit.rds")
+kids_med_visit <- read_rds(paste0(file_path, "Output/kids_med_visit.rds"))
 
@@
 
-write_csv(save_var_names, "Output/medvisit_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/medvisit_var_names.csv"))
 ```
@@
 ```{r}
-wmt_birthinfancy_initial <- read_xlsx("Input/WMT/wmt_birthinfancy_20200227.xlsx") %>% 
+wmt_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_birthinfancy_20200227.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-# Waiting on files from NN and AK
+nn_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_birthinfancy_20200416.xlsx")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "NN",
+         gender = as.character(Gender),
+         bw_lbs = as.numeric(BirthweightLbs),
+         bw_oz = as.numeric(BirthWeightOz),
+         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
+         mom_age = as.numeric(AgeOfMom),
+         breastfeed_current = as.character(CurrentlyBreastfeeding),
+         breastfeed_past = as.character(DidBreastfeed),
+         bf_years = as.numeric(HowLongYears),
+         bf_months = as.numeric(HowLongMonths),
+         breastfeed_months = as.numeric((bf_years*12)+bf_months),
+         flu_shot = as.character(FluShot),
+         flu_shot_planned = as.character(FluShotPlanned),
+         asthma_meds = as.character(AsthmaMeds),
+         child_school = as.character(Attend),
+         school_hours = as.numeric(HoursDay),
+         school_days = as.numeric(DaysWeek),
+         school_hours_week = as.numeric(school_hours*school_days),
+         sibling_school = as.character(SibSchool)) %>% 
+  select(home_winter_id:sibling_school)
+
+ak_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_birthinfancy_20200416.xlsx")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "AK",
+         gender = as.character(Gender),
+         bw_lbs = as.numeric(BirthweightLbs),
+         bw_oz = as.numeric(BirthWeightOz),
+         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
+         mom_age = as.numeric(AgeOfMom),
+         breastfeed_current = as.character(CurrentlyBreastfeeding),
+         breastfeed_past = as.character(DidBreastfeed),
+         bf_years = as.numeric(HowLongYears),
+         bf_months = as.numeric(HowLongMonths),
+         breastfeed_months = as.numeric((bf_years*12)+bf_months),
+         flu_shot = as.character(FluShot),
+         flu_shot_planned = as.character(FluShotPlanned),
+         asthma_meds = as.character(AsthmaMeds),
+         child_school = as.character(Attend),
+         school_hours = as.numeric(HoursDay),
+         school_days = as.numeric(DaysWeek),
+         school_hours_week = as.numeric(school_hours*school_days),
+         sibling_school = as.character(SibSchool)) %>% 
+  select(home_winter_id:sibling_school)
 
 # Combine logs from all areas
-kids_birth_infancy <- wmt_birthinfancy_initial %>% 
+kids_birth_infancy <- rbind(wmt_birthinfancy_initial, nn_birthinfancy_initial,
+                            ak_birthinfancy_initial) %>% 
   left_join(kids_linked_ids, c("area", "home_winter_id"))
 
-# write_rds(kids_birth_infancy, "Output/kids_birth_infancy.rds")
+# write_rds(kids_birth_infancy, paste0(file_path, "Output/kids_birth_infancy.rds"))
 
-kids_birth_infancy <- read_rds("Output/kids_birth_infancy.rds")
+kids_birth_infancy <- read_rds(paste0(file_path, "Output/kids_birth_infancy.rds"))
 
-save_var_names <- data.frame(t(wmt_birthinfancy_initial)) %>% 
+save_var_names <- data.frame(t(kids_birth_infancy)) %>% 
   rownames_to_column() %>% 
@@
 
-write_csv(save_var_names, "Output/birthinfancy_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/birthinfancy_var_names.csv"))
 ```
@@
 ```{r}
-wmt_childhomewinter_initial <- read_xlsx("Input/WMT/wmt_childhomewinter_20200227.xlsx") %>% 
+wmt_childhomewinter_initial <- 
+  read_xlsx(paste0(file_path, "Input/WMT/wmt_childhomewinter_20200227.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-# write_rds(child_home_winter, "Output/child_home_winter.rds")
+# write_rds(child_home_winter, paste0(file_path, "Output/child_home_winter.rds"))
 
-child_home_winter <- read_rds("Output/child_home_winter.rds")
+child_home_winter <- read_rds(paste0(file_path, "Output/child_home_winter.rds"))
 
@@
 ```{r}
-wmt_csq_initial <- read_xlsx("Input/WMT/wmt_csq_20200227.xlsx") %>% 
+wmt_csq_initial <- 
+  read_xlsx(paste0(file_path, "Input/WMT/wmt_csq_20200227.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-# Waiting on files from NN and AK
+nn_csq_initial <- 
+  read_xlsx(paste0(file_path, "Input/NN/nn_csq_20200416.xlsx")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "NN",
+         csq_date = ymd(Date),
+         csq_collect = as.character(HowCollected),
+         csq_sick_2_weeks = as.character(Illin2Weeks),
+         csq_still_sick = as.character(StillSick),
+         csq_first_symptom_date = ymd(FirstSymptomDate),
+         csq_symptom_free_date = ymd(SymptomFreeDate),
+         csq_miss_work = as.character(MissWork),
+         csq_miss_work_days = as.character(DaysMissed),
+         csq_see_dr = as.character(SeeDr),
+         csq_dr_diagnosis = as.character(DrDiagnosis),
+         csq_new_meds = as.character(NewMeds),
+         csq_fever_2weeks = as.character(Fever),
+         csq_temp_2weeks = as.numeric(Temperature),
+         csq_rapid_breathing_2weeks = as.character(RapidBreathing),
+         csq_difficult_breathing_2weeks = as.character(DifficultBreathing),
+         csq_noisy_breathing_2weeks = as.character(NoisyBreathing),
+         csq_runny_nose_2weeks = as.character(RunnyNose),
+         csq_cough_2weeks = as.character(Coughing),
+         csq_wheeze_2weeks = as.character(Wheezing),
+         csq_fever_current = as.character(FeverStill),
+         csq_rapid_breathing_current = as.character(RapidBreathingStill),
+         csq_difficult_breathing_current = as.character(DifficultBreathingStill),
+         csq_noisy_breathing_current = as.character(NoisyBreathingStill),
+         csq_runny_nose_current = as.character(RunnyNoseStill),
+         csq_cough_current = as.character(CoughingStill),
+         csq_wheeze_current = as.character(WheezingStill),
+         csq_comments = as.character(`REPLACE(REPLACE(Comments, '\\n', ''), '\\r', '')`)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
+  select(home_winter_id:csq_comments)
+
+ak_csq_initial <- 
+  read_xlsx(paste0(file_path, "Input/AK/ak_csq_20200416.xlsx")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "AK",
+         csq_date = ymd(Date),
+         csq_collect = as.character(HowCollected),
+         csq_sick_2_weeks = as.character(Illin2Weeks),
+         csq_still_sick = as.character(StillSick),
+         csq_first_symptom_date = ymd(FirstSymptomDate),
+         csq_symptom_free_date = ymd(SymptomFreeDate),
+         csq_miss_work = as.character(MissWork),
+         csq_miss_work_days = as.character(DaysMissed),
+         csq_see_dr = as.character(SeeDr),
+         csq_dr_diagnosis = as.character(DrDiagnosis),
+         csq_new_meds = as.character(NewMeds),
+         csq_fever_2weeks = as.character(Fever),
+         csq_temp_2weeks = as.numeric(Temperature),
+         csq_rapid_breathing_2weeks = as.character(RapidBreathing),
+         csq_difficult_breathing_2weeks = as.character(DifficultBreathing),
+         csq_noisy_breathing_2weeks = as.character(NoisyBreathing),
+         csq_runny_nose_2weeks = as.character(RunnyNose),
+         csq_cough_2weeks = as.character(Coughing),
+         csq_wheeze_2weeks = as.character(Wheezing),
+         csq_fever_current = as.character(FeverStill),
+         csq_rapid_breathing_current = as.character(RapidBreathingStill),
+         csq_difficult_breathing_current = as.character(DifficultBreathingStill),
+         csq_noisy_breathing_current = as.character(NoisyBreathingStill),
+         csq_runny_nose_current = as.character(RunnyNoseStill),
+         csq_cough_current = as.character(CoughingStill),
+         csq_wheeze_current = as.character(WheezingStill),
+         csq_comments = as.character(`REPLACE(REPLACE(Comments, '\\n', ''), '\\r', '')`)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
+  select(home_winter_id:csq_comments)
 
 # Combine logs from all areas
-kids_csq <- wmt_csq_initial %>% 
+kids_csq <- rbind(wmt_csq_initial, nn_csq_initial, ak_csq_initial) %>% 
   left_join(kids_linked_ids, c("area", "home_winter_id"))
@@
 
-kids_csq <- read_rds("Output/kids_csq.rds")
+kids_csq <- read_rds(paste0(file_path, "Output/kids_csq.rds"))
 
@@
 
-write_csv(save_var_names, "Output/csq_var_names.csv")
+write_csv(save_var_names, paste0(file_path, "Output/csq_var_names.csv"))
 ```
@@
 ```{r}
-wmt_medicalrecords_initial <- read_xlsx("Input/WMT/wmt_medicalrecords_20200227.xlsx") %>% 
+wmt_medicalrecords_initial <- 
+  read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalrecords_20200227.xlsx")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-# Waiting on files from NN and AK
+nn_medicalrecords_initial <- 
+  read_xlsx(paste0(file_path, "Input/NN/nn_medicalrecords_20200416.xlsx")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "NN",
+         visit_date = ymd(VisitDate),
+         med_rec_visit_date = ymd(VisitDate),
+         med_rec_dr_diagnosis = as.character(DrDiag),
+         med_rec_ari = as.character(ARI),
+         med_rec_comments = as.character(DrComments)) %>% 
+  select(home_winter_id:med_rec_comments)
+
+ak_medicalrecords_initial <- 
+  read_xlsx(paste0(file_path, "Input/AK/ak_medicalrecords_20200416.xlsx")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         area = "AK",
+         visit_date = ymd(VisitDate),
+         med_rec_visit_date = ymd(VisitDate),
+         med_rec_dr_diagnosis = as.character(DrDiag),
+         med_rec_ari = as.character(ARI),
+         med_rec_comments = as.character(DrComments)) %>% 
+  select(home_winter_id:med_rec_comments)
 
@@
 
-# write_rds(kids_medical_records, "Output/kids_medical_records.rds")
+# write_rds(kids_medical_records, paste0(file_path, "Output/kids_medical_records.rds"))
 
-kids_medical_records <- read_rds("Output/kids_medical_records.rds")
+kids_medical_records <- read_rds(paste0(file_path, "Output/kids_medical_records.rds"))
 ```

### 2020-04-20 dustrak_data_cleaning.Rmd
child data cleaning updates

@@
 # Run all of these lines together
-pm_time_plot_function(id = "273", location = "WMT")
+pm_time_plot_function(id = "501", location = "WMT")
 head(gaps, 25)

### 2020-04-20 exposure_models.Rmd
child data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 13 April 2020"
+date: "Started 19 March 2020, Updated 16 April 2020"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
+                      eval = TRUE, include = TRUE)
 ```
@@
 ```{r}
+# Load individual datasets
+
 input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
@@
 
-# Linear model
 ```{r}
-subset_model <- lm(log(pm_mean) ~ 
-                        income_3level + home_act_smoking + 
-                        area, data = exposure_analysis_data_short)
+# Prep and merge data
 
-full_model <- lm(log(pm_mean) ~ 
-                        education_3level + income_3level + 
-                        total_residents_2level + home_floors_2level + home_windows_2level + 
-                        stove_age_3level + chimney_clean_3level + 
-                        wood_collect_2level + 
-                        stove_grade_3level + 
-                        home_act_smoking + 
-                        moisture_2level + 
-                        temp_perc_30_cat +
-                        area, data = exposure_analysis_data_short)
+####### Format SUMs data to merge with analysis data below
+sums_temp_change <- sums_rolling_pm_data %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home, sampling_day) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
+  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
+         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
+  replace_with_na(replace = list(lead_temp = 999)) %>% 
+  mutate(temp_diff = lead_temp - temp_c_sums,
+         # change temp in the following line
+         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
+         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
+                              temp_diff_check),
+         heat_event = if_else(is.na(temp_diff_check), 1,
+                              heat_event)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
+         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
+                                            breaks = c(0, 2, 50),
+                           labels = c("<2", "2+"),
+                                 right = FALSE)) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
+         time_diff:sums_events_sampling_day_cat) %>% 
+  arrange(area, home, sample_date) %>% 
+  group_by(home) %>% 
+  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
+  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup()
 
-summary(subset_model)
-r.squaredGLMM(subset_model)
+####### Format ambient data to merge with analysis data below
+ambient_data <- exposure_analysis_data_long %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  #filter(area == "AK") %>% 
+  filter(!is.na(sampling_day)) %>% 
+  filter(!is.na(sample_date)) %>%
+  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
+         zip:mean_temp_roll_4day) %>% 
+  group_by(home) %>% 
+  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
+         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup()
 
-tidy_results <- tidy(full_model)
-tidy_results
+####### Format home activity data to merge with analysis data below
+home_activity_data <- exposure_analysis_data_medium %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  group_by(area, home) %>% 
+  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
+  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
+  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
+  ungroup() %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  select(home, area, sampling_day, starts_with("home_act"))
 
-plot(full_model, 
-     xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(full_model), main = "QQ Plot")
+####### Format full data to use for analysis
+analysis_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
+                                 labels = c("<10.5", "10.5+"),
+                                 right = FALSE),
+         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
+                                 labels = c("<10.5", "10.5+"),
+                                 right = FALSE)) %>% 
+  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
+                                 labels = c("<30", "30+"),
+                                 right = FALSE),
+         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
+                                 labels = c("<16.5", "16.5+"),
+                                 right = FALSE),
+         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
+                                 labels = c("<40", "40+"),
+                                 right = FALSE),
+         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
+                                 labels = c("<20", "20+"),
+                                 right = FALSE)) %>% 
+  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
+         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
+                                 labels = c("<50", "50+"),
+                                 right = FALSE),
+         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
+                                 labels = c("<35", "35+"),
+                                 right = FALSE),
+         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
+                                 labels = c("<10", "10+"),
+                                 right = FALSE)) %>% 
+  left_join(sums_temp_change, by = c("area", "home")) %>% 
+  left_join(ambient_data, by = c("area", "home")) %>% 
+  select(-starts_with("home_act")) %>% 
+  left_join(home_activity_data, by = c("area", "home")) %>% 
+  select(pm_mean_sampling_period, area, income_3level,  
+                 residents_smoke, home_floors_2level, stove_age_3level,  
+                 chimney_clean_3level, home_mold, wood_collect_2level, 
+                 home_act_door, home_act_smoking, home_act_sweep,
+                 stove_grade_3level) %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  na.exclude()
+```
+
+Initial variables to use for model selection (14) from collinearity RMD:
+Demographics: area, income_3level, residents_smoke
+Home Characteristics: home_floors_2level, stove_age_3level,
+                      chimney_clean_3level, home_mold, wood_collect_2level
+Home Activity: home_act_door, home_act_smoking, home_act_sweep
+Others: sums_events_sampling_period, mean_temp_sampling_period_5, stove_grade_3level
+
+Note: sums_events_sampling_period and mean_temp_sampling_period_5 have a lot of
+missing data and reduce the sample size from 93 to 43. Not considering these
+variables for model selection for now.
+
+```{r, include=FALSE}
+full_model <- lm(log(pm_mean_sampling_period) ~ area + income_3level + 
+                 residents_smoke + home_floors_2level + stove_age_3level + 
+                 chimney_clean_3level + home_mold + wood_collect_2level +
+                 home_act_door + home_act_smoking + home_act_sweep + 
+                 stove_grade_3level,
+                 data = analysis_data)
+
+summary(full_model)
+r.squaredGLMM(full_model)
+tidy(full_model)
+
+plot(full_model)
 ```
@@
 # Best subsets selection using dredge() from MuMIn
-```{r}
+
+Make sure to run `na.exclude` on analysis dataset and `options(na.action = "na.fail")`
+before running the dredge best-subsets function below. This will ensure that the 
+same dataset is being used for each potential model, which is an assumption and
+a prerequisite for comparing different model ranking like AIC.
+
+```{r, eval=FALSE, include=FALSE}
 options(na.action = "na.fail")
 
-all_subsets <- dredge(full_model, rank = "AIC", extra = c("R^2"))
+all_subsets_aic <- dredge(full_model, rank = "AIC", extra = c("R^2"))
+all_subsets_aicc <- dredge(full_model, rank = "AICc", extra = c("R^2"))
+all_subsets_bic <- dredge(full_model, rank = "BIC", extra = c("R^2"))
+
+write_rds(all_subsets_aic, paste0(output_path, "Output/all_subsets_aic"))
+write_rds(all_subsets_aicc, paste0(output_path, "Output/all_subsets_aicc"))
+write_rds(all_subsets_bic, paste0(output_path, "Output/all_subsets_bic"))
 ```
 
-# Linear mixed model with repeated measures
 ```{r}
-subset_model <- lmer(log(pm_mean_daily) ~ 
-                        education_3level + income_3level + race_parent_2level + 
-                        area + sampling_day + day_of_week +
-                        (1 | home), data = exposure_analysis_data_long)
+all_subsets_aic <- read_rds(paste0(output_path, "Output/all_subsets_aic"))
+all_subsets_aicc <- read_rds(paste0(output_path, "Output/all_subsets_aicc"))
+all_subsets_bic <- read_rds(paste0(output_path, "Output/all_subsets_bic"))
 
-full_model <- lmer(log(pm_mean_daily) ~ 
-                        education_3level + income_3level + race_parent_2level + 
-                        total_residents_2level + residents_smoke + 
-                        home_type_2level + home_floors_2level + home_windows_2level + 
-                        stove_age_3level + chimney_clean_3level + home_mold + 
-                        home_fireplace + 
-                        wood_collect_2level + 
-                        stove_grade_3level + 
-                        home_act_gas + home_act_propane + home_act_smoking + 
-                        moisture_2level + 
-                        temp_max_5 + 
-                        temp_perc_30_cat +
-                        area + sampling_day + day_of_week +
-                        (1 | home), data = exposure_analysis_data_long)
+head(all_subsets_aic, 10)
+head(all_subsets_aicc, 10)
+head(all_subsets_bic, 10)
+```
 
-summary(full_model)
-r.squaredGLMM(full_model)
+```{r}
+subset_model <- lm(log(pm_mean_sampling_period) ~ 
+                   income_3level + home_act_door + residents_smoke, 
+                   data = analysis_data)
 
-tidy_results <- tidy(full_model)
-tidy_results
+summary(subset_model)
+#tidy(subset_model)
 
-plot(full_model, 
-     xlab = "fitted values", ylab = "residuals")
-qqnorm(residuals(full_model), main = "QQ Plot")
+plot(subset_model)
 ```

### 2020-04-20 wmt_addresses.Rmd
child data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 3 April 2020, Updated 3 April 2020"
+date: "Started 3 April 2020, Updated 17 April 2020"
 output: pdf_document
@@
 ```{r}
-wmt_addresses <- read_xlsx("Input/WMT/wmt_addresses.xlsx")
+input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_addresses <- read_xlsx(paste0(input_path, "Input/WMT/wmt_addresses.xlsx"))
 
@@
 ```{r}
-addresses_geocoded <- geocode(location = addresses_new$address, output = "latlon",
+addresses_geocoded <- addresses_new %>% 
+  mutate_geocode(location = address, output = "more",
                               source = "google")
 ```
+
+```{r}
+wmt_map <- map_data("state") %>% 
+  filter(region == "montana")
+
+wmt_homes <- addresses_geocoded %>% 
+  ggplot() +
+    geom_polygon(data=wmt_map, aes(x=long, y=lat, group=group), 
+               colour="black", fill = "azure", size = 1.2) +
+    geom_point(data=addresses_geocoded, 
+             aes(lon, lat), size = 3) +
+    theme_void() +
+    labs(title = "WMT Homes Mapped", subtitle = "Geocoded from addresses") 
+wmt_homes
+```

### 2020-04-22 child_health_data_cleaning.Rmd
health data cleaning

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 20 April 2020"
+date: "Started 27 Feb 2020, Updated 22 April 2020"
 output: pdf_document 
@@
 library(knitr)
-jv_palette <- c("#330099","#CC0066","#FF6633", 
-                 "#0099CC", "#FF9900","#CC6633",
-                  "#FF3366", "#33CC99", "#33999")
 ```
@@
 kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+
 demographics_clean <- 
   read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+
 pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
@@
 
-# write_rds(kids_at_home_log, "Output/kids_at_home_log.rds")
+# write_rds(kids_at_home_log, paste0(file_path, "Output/kids_at_home_log.rds"))
 ```
@@
 ```{r}
-wmt_medicalvisit_initial <- read_xlsx("Input/WMT/wmt_medicalvisit_20200227.xlsx") %>% 
+wmt_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalvisit_20200227.xlsx")) %>% 
   mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
@@
          sibling_school = as.character(SibSchool)) %>% 
+  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
   select(home_winter_id:sibling_school)
@@
          sibling_school = as.character(SibSchool)) %>% 
+  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>% 
   select(home_winter_id:sibling_school)
@@
 
-# Load and merge files from different locations
+As of 22 April 2020 we only have CHW files for WMT.
+The content isn't particularly necessary for data analysis, so leave out of the
+final dataset for now and see if files for NN and AK come up later.
 ```{r}
@@
          csq_comments = as.character(`REPLACE(REPLACE(Comments, '\\n', ''), '\\r', '')`)) %>% 
-  replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
+  replace_with_na_all(condition = ~.x %in% c(-9999, -999, -99, -99.9, 999, 0)) %>% 
   select(home_winter_id:csq_comments)
@@
 
-# write_rds(kids_csq, "Output/kids_csq.rds")
+# write_rds(kids_csq, paste0(file_path, "Output/kids_csq.rds"))
 
@@
 
-# Load and merge files from different locations
+These files are different for WMT compared to NN and AK
+The relevant information from NN/AK med recs seems to be in the "respiratory" 
+file. For now, I'll read those in and match with the WMT med rec file.
+
+See if WMT file for med rec/respiratory is updated once all data is input.
 ```{r}
@@
          med_rec_ari = as.character(ARI),
-         med_rec_comments = as.character(DrComments)) %>% 
-  select(home_winter_id:med_rec_comments)
+         med_rec_comments = as.character(DrComments),
+         med_rec_meds = "NA",
+         med_rec_tests = "NA",
+         med_rec_treatments = "NA",
+         med_rec_chart_notes = "NA") %>% 
+  select(home_winter_id:med_rec_chart_notes)
 
 nn_medicalrecords_initial <- 
-  read_xlsx(paste0(file_path, "Input/NN/nn_medicalrecords_20200416.xlsx")) %>% 
+  read_xlsx(paste0(file_path, "Input/NN/nn_respiratory_20200416.xlsx"),
+            skip = 1,
+            col_names = c("id", "mrid", "ChildID", "HomeWinterID", "VisitDate",
+                          "DrDiag", "tests", "treatments", "meds", "Status",
+                          "DrComments"),
+            col_types = c("numeric", "numeric", "text", "text", "date",
+                          "text", "text", "text", "text",
+                          "text", "text")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          med_rec_dr_diagnosis = as.character(DrDiag),
-         med_rec_ari = as.character(ARI),
-         med_rec_comments = as.character(DrComments)) %>% 
-  select(home_winter_id:med_rec_comments)
+         med_rec_ari = as.character(Status),
+         med_rec_comments = as.character(DrComments),
+         med_rec_meds = as.character(meds),
+         med_rec_tests = as.character(tests),
+         med_rec_treatments = as.character(treatments),
+         med_rec_chart_notes = "NA") %>% 
+  select(home_winter_id:med_rec_chart_notes)
 
 ak_medicalrecords_initial <- 
-  read_xlsx(paste0(file_path, "Input/AK/ak_medicalrecords_20200416.xlsx")) %>% 
+  read_xlsx(paste0(file_path, "Input/AK/ak_respiratory_20200416.xlsx"),
+            skip = 1,
+            col_names = c("id", "mrid", "ChildID", "HomeWinterID", "VisitDate",
+                          "DrDiag", "tests", "treatments", "meds", 
+                          "chartnotes", "prestudy", "Status",
+                          "DrComments"),
+            col_types = c("numeric", "numeric", "text", "text", "date",
+                          "text", "text", "text", "text", "text", "text",
+                          "text", "text")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          med_rec_dr_diagnosis = as.character(DrDiag),
-         med_rec_ari = as.character(ARI),
-         med_rec_comments = as.character(DrComments)) %>% 
-  select(home_winter_id:med_rec_comments)
+         med_rec_ari = as.character(Status),
+         med_rec_comments = as.character(DrComments),
+         med_rec_meds = as.character(meds),
+         med_rec_tests = as.character(tests),
+         med_rec_treatments = as.character(treatments),
+         med_rec_chart_notes = as.character(chartnotes)) %>% 
+  select(home_winter_id:med_rec_chart_notes)
 
 # Combine logs from all areas
-kids_medical_records <- wmt_medicalrecords_initial %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id"))
+kids_medical_records <- rbind(wmt_medicalrecords_initial, nn_medicalrecords_initial,
+                              ak_medicalrecords_initial) %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
+  arrange(area, visit_date)
 
@@
 
-# Load and merge files from different locations
+There isn't an LRTI file for NN/AK as of 22 April 2020. 
+Leaving WMT file for now and will check back on updated data/files later.
 ```{r}

### 2020-04-23 collinearity_exposure_analysis.Rmd
exposure analysis updates

@@
 Based on results and potential overlap, the following vars were removed:
-         race_parent_2level, home_damp, moisture_closest_2level,
+         race_parent_2level, home_damp, 
          temp_max_5, home_act_gas, total_residents_2level, home_type_2level, 
@@
 Home Activity: home_act_door, home_act_smoking, home_act_sweep
-Others: sums_events_sampling_period, mean_temp_sampling_period_5, stove_grade_3level
+Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
+        stove_grade_3level, moisture_closest_2level,
 

### 2020-04-23 exposure_models.Rmd
exposure analysis updates

@@
   filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(area == "WMT") %>% 
   group_by(area, home) %>% 
@@
   left_join(home_activity_data, by = c("area", "home")) %>% 
-  select(pm_mean_sampling_period, area, income_3level,  
-                 residents_smoke, home_floors_2level, stove_age_3level,  
-                 chimney_clean_3level, home_mold, wood_collect_2level, 
+  select(pm_mean_sampling_period, area, income_3level, home_act_windows, 
+                 stove_age_3level, home_windows_2level,
+                 chimney_clean_3level, wood_collect_2level, 
                  home_act_door, home_act_smoking, home_act_sweep,
-                 stove_grade_3level) %>% 
+                 stove_grade_3level, moisture_closest_2level,
+                 total_residents_2level, sums_events_sampling_period, mean_temp_sampling_period_5) %>% 
   filter(!is.na(pm_mean_sampling_period)) %>% 
@@
 
+Also try adding other variables in to check their impact on model selection:
+   total_residents_2level (seems impactful), home_type_2level (not impactful), 
+   home_windows_2level (seems impactful), home_act_windows (not very impactful),
+   education_3level (not impactful)
+
+Some vars not important in any models, so removing from consideration:
+   home_mold, home_floors_2level, home_act_windows, home_type_2level
+
+Next steps:
+Some vars (like residents_smoke), other demographics, and home activity vars
+have quite a bit of missing data. This is in addition to the missing ambient
+and SUMs data. The first step should be to account for as much of this as possible.
+See if there are ways to pull demographic data from Winter 2.
+Then, see if data for WMT is more complete in general. The idea is to minimize
+missing data for model selection and the final model to have as much power as
+possible. 
+Consider doing a more "full" model for WMT only (if there is less missing data),
+particularly since the ambient data is more reliable for WMT than the other areas.
+Then you could justify not doing a mixed model, and model selection would be easier.
+Then, using data from all areas, use the full model from WMT to work on a 
+mixed model that is likely reduced due to sparse data from the other areas.
+Since model selection is difficult with mixed models, use the WMT model to inform
+this model, while also doing sensitivity analysis to check other variables.
+Aside from the multivariable model, much of the paper could also be looking at
+simple associations between PM and some of the novel covars (SUMS, ambient data,
+stove grades, wood moisture), as well as methods for collecting and analysing 
+those vars.
+Also look into predict() function for teach/test and validating the model.
+Discussion can focus on what important predictors of indoor PM are in these homes - 
+if simple demographic data predicts as well as complex things like SUMs and wood
+moisture, or if there is a spectrum of predictors and complexity in their collection.
+
 ```{r, include=FALSE}
-full_model <- lm(log(pm_mean_sampling_period) ~ area + income_3level + 
-                 residents_smoke + home_floors_2level + stove_age_3level + 
-                 chimney_clean_3level + home_mold + wood_collect_2level +
+full_model <- lm(log(pm_mean_sampling_period) ~ income_3level + 
+                 stove_age_3level + home_windows_2level +
+                 chimney_clean_3level + wood_collect_2level + 
                  home_act_door + home_act_smoking + home_act_sweep + 
-                 stove_grade_3level,
+                 stove_grade_3level + moisture_closest_2level + 
+                 total_residents_2level + mean_temp_sampling_period_5 +
+                 sums_events_sampling_period, 
                  data = analysis_data)
@@
 all_subsets_aic <- dredge(full_model, rank = "AIC", extra = c("R^2"))
+
 all_subsets_aicc <- dredge(full_model, rank = "AICc", extra = c("R^2"))
+par(mar = c(3,5,6,4))
+plot(all_subsets_aicc)
+
 all_subsets_bic <- dredge(full_model, rank = "BIC", extra = c("R^2"))
@@
 ```
+
+#####################
+
+Looking at Lasso and Ridge regression
+May be beyond the scope of this paper, but keep code and references:
+
+https://cran.r-project.org/web/packages/MuMIn/MuMIn.pdf
+
+http://www.science.smith.edu/~jcrouser/SDS293/labs/lab10-r.html
+
+https://cran.r-project.org/web/packages/glmnet/glmnet.pdf
+
+https://cran.r-project.org/web/packages/glmmLasso/glmmLasso.pdf
+
+https://arxiv.org/pdf/1306.2427.pdf
+
+```{r} 
+library(MuMIn)
+library(leaps)
+library(glmnet)
+library(glmmLasso)
+
+# Range of lambdas
+grid = 10^seq(10, -2, length = 100)
+
+lasso_results <- glmmLasso(log(pm_mean_sampling_period) ~ income_3level + 
+                 residents_smoke + stove_age_3level + home_windows_2level +
+                 chimney_clean_3level + wood_collect_2level + 
+                 home_act_door + home_act_smoking + home_act_sweep + 
+                 stove_grade_3level + moisture_closest_2level + 
+                 total_residents_2level,
+                 rnd = list(area = ~ 1), 
+                 lambda = grid,
+                 data = analysis_data)
+
+summary(lasso_results)
+
+predict(lasso_results)
+```

### 2020-04-24 exposure_models.Rmd
exposure analysis updates

@@
 library(leaps)
+library(glmnet)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 ambient_data <- exposure_analysis_data_long %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  #filter(area == "AK") %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  #filter(area == "WMT") %>%
   filter(!is.na(sampling_day)) %>% 
@@
 home_activity_data <- exposure_analysis_data_medium %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home) %>% 
+  filter(treatment == "Placebo") %>% 
+  #filter(winter_id == 1) %>% 
+  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
+  group_by(area, home, home_winter_id) %>% 
   distinct(sampling_day, .keep_all = TRUE) %>% 
-  group_by(area, home) %>% 
+  group_by(area, home, home_winter_id) %>% 
   mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
@@
   ungroup() %>% 
-  distinct(area, home, .keep_all = TRUE) %>% 
-  select(home, area, sampling_day, starts_with("home_act"))
+  distinct(area, home,  home_winter_id, .keep_all = TRUE)
 
@@
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
   filter(area == "WMT") %>% 
   group_by(area, home) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
+  distinct(home, winter_id, .keep_all = TRUE) %>% 
   ungroup() %>% 
@@
                                  right = FALSE)) %>% 
-  left_join(sums_temp_change, by = c("area", "home")) %>% 
+  group_by(home) %>% 
+  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
+                                   lead(residents_smoke), residents_smoke)) %>% 
+  ungroup() %>% 
+  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
   left_join(ambient_data, by = c("area", "home")) %>% 
   select(-starts_with("home_act")) %>% 
-  left_join(home_activity_data, by = c("area", "home")) %>% 
-  select(pm_mean_sampling_period, area, income_3level, home_act_windows, 
+  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
+  select(pm_mean_sampling_period, area, income_3level, 
                  stove_age_3level, home_windows_2level,
                  chimney_clean_3level, wood_collect_2level, 
-                 home_act_door, home_act_smoking, home_act_sweep,
+                 home_act_door, home_act_smoking, home_act_sweep, home_act_windows, 
                  stove_grade_3level, moisture_closest_2level,
-                 total_residents_2level, sums_events_sampling_period, mean_temp_sampling_period_5) %>% 
-  filter(!is.na(pm_mean_sampling_period)) %>% 
-  na.exclude()
+                 total_residents_2level, mean_temp_sampling_period_5) %>% 
+  filter(!is.na(pm_mean_sampling_period))
+
+dep_var <- analysis_data$pm_mean_sampling_period
+
+ind_vars <- analysis_data %>% 
+  select(-pm_mean_sampling_period, -area) %>% 
+  data.matrix()
 ```
@@
 
-Looking at Lasso and Ridge regression
+Looking at Lasso regression
 May be beyond the scope of this paper, but keep code and references:
@@
 ```
+
+Looking at Ridge regression:
+
+https://drsimonj.svbtle.com/ridge-regression-with-glmnet
+
+```{r}
+ridge_results <- cv.glmnet(x = ind_vars, y = dep_var, alpha = 0)
+
+plot(ridge_results)
+
+opt_lambda <- ridge_results$lambda.min
+opt_lambda
+
+fit <- ridge_results$glmnet.fit
+summary(fit)
+
+y_predicted <- predict(fit, s = opt_lambda, newx = ind_vars)
+
+# Sum of Squares Total and Error
+sst <- sum((dep_var - mean(dep_var))^2)
+sse <- sum((y_predicted - dep_var)^2)
+
+# R squared
+rsq <- 1 - sse / sst
+rsq
+```
+
+# Train and test datasets
+
+```{r}
+# Random sample indexes
+train_index <- sample(1:nrow(adult), 0.8 * nrow(adult))
+test_index <- setdiff(1:nrow(adult), train_index)
+
+dep_var <- analysis_data$pm_mean_sampling_period
+
+ind_vars <- analysis_data %>% 
+  select(-pm_mean_sampling_period, -area) %>% 
+  data.matrix()
+```

### 2020-04-24 home_char_act_cleaning.Rmd
exposure analysis updates

@@
 library(zoo)
+
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 ```
@@
 # Load data
-home_act_initial_wmt <- read_csv("Input/WMT/wmt_home_activity_20200113.csv")
+home_act_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/wmt_home_activity_20200113.csv"))
 
@@
 # Save as RDS
-# write_rds(wmt_home_act, "Output/wmt_home_act.rds")
+# write_rds(wmt_home_act, paste0(file_path, "Output/wmt_home_act.rds"))
 
@@
 # Load data
-home_act_initial_nn <- read_xlsx("Input/NN/nn_home_activity_20200113.xlsx")
+home_act_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/nn_home_activity_20200113.xlsx"))
 
@@
 # Save as RDS
-# write_rds(nn_home_act, "Output/nn_home_act.rds")
+# write_rds(nn_home_act, paste0(file_path, "Output/nn_home_act.rds"))
 
@@
 # Load data
-home_act_initial_ak <- read_xlsx("Input/AK/ak_home_activity_20200113.xlsx")
+home_act_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/ak_home_activity_20200113.xlsx"))
 
@@
 # Save as RDS
-# write_rds(ak_home_act, "Output/ak_home_act.rds")
+# write_rds(ak_home_act, paste0(file_path, "Output/ak_home_act.rds"))
 ```
@@
 ```{r}
-wmt_home_act <- read_rds("Output/wmt_home_act.rds")
-nn_home_act <- read_rds("Output/nn_home_act.rds")
-ak_home_act <- read_rds("Output/ak_home_act.rds")
+wmt_home_act <- read_rds(paste0(file_path, "Output/wmt_home_act.rds"))
+nn_home_act <- read_rds(paste0(file_path, "Output/nn_home_act.rds"))
+ak_home_act <- read_rds(paste0(file_path, "Output/ak_home_act.rds"))
 
@@
 
-# write_rds(kids_home_act, "Output/kids_home_act.rds")
+# write_rds(kids_home_act, paste0(file_path, "Output/kids_home_act.rds"))
 ```
@@
 ```{r}
-kids_home_act <- read_rds("Output/kids_home_act.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+kids_home_act <- read_rds(paste0(file_path, "Output/kids_home_act.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 
@@
 
-# write_rds(home_act_clean, "Output/home_act_clean.rds")
+# write_rds(home_act_clean, paste0(file_path, "Output/home_act_clean.rds"))
 ```

### 2020-04-27 exposure_models.Rmd
exposure model updates

@@
   filter(treatment == "Placebo") %>% 
-  filter(area == "WMT") %>% 
+  #filter(area == "WMT") %>% 
   group_by(area, home) %>% 
@@
   left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
-  select(pm_mean_sampling_period, area, income_3level, 
+  select(pm_mean_sampling_period, income_3level, 
          stove_age_3level, home_windows_2level,
          chimney_clean_3level, wood_collect_2level,
-                 home_act_door, home_act_smoking, home_act_sweep, home_act_windows, 
-                 stove_grade_3level, moisture_closest_2level,
-                 total_residents_2level, mean_temp_sampling_period_5) %>% 
+         stove_grade_3level)
   filter(!is.na(pm_mean_sampling_period)) 
+  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN"))
+
+# Check % missing data
+#mean(is.na(analysis_data))
 
@@
 ```
+
+Looking at model selection and averaging after multiple imputation
+
+```{r}
+library(MAMI)
+#library(Amelia)
+library(mice)
+
+# Using mice package
+
+# Look for patterns in missing data
+md.pattern(analysis_data)
+
+# Impute missing data; m = 5 datasets (check % missing data and match)
+imputed_data <- mice(analysis_data, m=5)
+summary(imputed_data)
+
+# Sample one of the imputed datasets
+check1 <- complete(imputed_data, 3)
+
+# Using MAMI package, perform model averaging and selection
+model_ave <- mami(imputed_data, missing.data = "imputed", model = "gaussian",
+                  outcome = "pm_mean_sampling_period",
+                  method = "MA.criterion", criterion = "AIC",
+                  inference = "+boot", B = 15, X.org = analysis_data, print.time = TRUE)
+
+summary(model_ave)
+plot(model_ave)
+```

### 2020-04-28 exposure_models.Rmd
exposure analysis updates

--- 2020-04-24 exposure_models.Rmd
@@
   mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
+  mutate_at(vars(starts_with("home_act")), factor) %>%
   ungroup() %>% 
@@
   left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
-  select(pm_mean_sampling_period, area, income_3level, 
-                 stove_age_3level, home_windows_2level,
-                 chimney_clean_3level, wood_collect_2level, 
-                 home_act_door, home_act_smoking, home_act_sweep, home_act_windows, 
-                 stove_grade_3level, moisture_closest_2level,
-                 total_residents_2level, mean_temp_sampling_period_5) %>% 
-  filter(!is.na(pm_mean_sampling_period))
-
-dep_var <- analysis_data$pm_mean_sampling_period
-
-ind_vars <- analysis_data %>% 
-  select(-pm_mean_sampling_period, -area) %>% 
-  data.matrix()
+  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) 
 ```
@@
 ```
+
+Looking at model selection and averaging after multiple imputation
+
+Helpful manuals and vignettes:
+
+https://cran.r-project.org/web/packages/mice/mice.pdf
+
+```{r}
+library(mice)
+
+# Select analysis vars
+selected_vars <- analysis_data %>% 
+  select(area, pm_mean_sampling_period, income_3level, residents_smoke, 
+         stove_age_3level, home_windows_2level,
+         chimney_clean_3level, wood_collect_2level,
+         home_act_door, home_act_sweep, home_act_windows, 
+         sums_events_sampling_period, mean_temp_sampling_period_5, 
+         stove_grade_3level, moisture_closest_2level) %>% 
+  mutate(home_act_door = factor(home_act_door, levels = c(0,1,2,3,4,5,6)),
+         home_act_sweep = factor(home_act_sweep, levels = c(0,1,2,3,4,5,6)),
+         home_act_windows = factor(home_act_windows, levels = c(0,1,2,3,4,5,6)),
+         area = factor(area, levels = c("AK", "NN", "WMT")))
+
+# Check % missing data
+mean(is.na(selected_vars))
+
+# Using mice package
+
+# Look for patterns in missing data
+md.pattern(selected_vars)
+
+# Impute missing data; m = 20 datasets (check % missing data and match)
+imputed_data <- mice(selected_vars, m=5,
+                     method = c("polyreg", "pmm", "polr", "logreg", "polr",
+                                "logreg", "polr", "logreg", "polr", "polr",
+                                "polr", "pmm", "pmm", "polr", "logreg"))
+
+summary(imputed_data)
+
+# Can add "long" or "broad arguments
+complete(imputed_data, "long")
+
+# Sample one of the imputed datasets
+check1 <- complete(imputed_data, 3)
+
+densityplot(selected_vars$pm_mean_sampling_period)
+
+# Use seed to get the exact same results from random sampling
+# seed = 123
+
+# See vignette #2 for setting up predictor matrices and relationships b/w vars
+
+# Check for convergence
+## In general, we would like the streams to intermingle and be free of any trends 
+## at the later iterations.
+plot(imputed_data)
+
+# Check that the imputation method matches the data type correctly
+## particularly for PM that should be log transformed, and the factors
+
+stripplot(imputed_data)
+
+# Can do regression on the imputed data - either pooled or individual datasets
+
+```
+
+http://mami.r-forge.r-project.org/MAMI_manual.pdf
+
+```{r}
+library(MAMI)
+
+# Using MAMI package, perform model averaging and selection
+model_ave <- mami(imputed_data, missing.data = "imputed", model = "gaussian",
+                  outcome = "pm_mean_sampling_period", 
+                  method = "MA.criterion", criterion = "AIC", print.time = TRUE,
+                  add.transformation = c("log(pm_mean_sampling_period)"))
+
+                  #inference = "+boot", B = 2, X.org = selected_vars
+
+summary(model_ave)
+plot(model_ave)
+```

### 2020-04-29 exposure_models.Rmd
multiple imputation updates

@@
 library(mice)
+library(pan)
+library(multilevel)
 
@@
          home_act_windows = factor(home_act_windows, levels = c(0,1,2,3,4,5,6)),
-         area = factor(area, levels = c("AK", "NN", "WMT")))
+         area = factor(area, levels = c("AK", "NN", "WMT")),
+         log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
+  select(-pm_mean_sampling_period) %>% 
+  filter(!is.na(income_3level) & !is.na(home_windows_2level))
+
+# Look at structure of data and check variable types
+str(selected_vars)
 
@@
 # Look for patterns in missing data
-md.pattern(selected_vars)
+missing_patterns <- md.pattern(selected_vars)
+
+missing_patterns
+
+sum(missing_patterns[, "residents_smoke"] == 0)
+
+hist(selected_vars$pm_mean_sampling_period)
+
+# Check for distribution between vars and missingness - see if there are patterns
+## No obvious patterns in the missing data
+## Missing covars do not seem to be related to the outcomes
+missing_hist <- selected_vars %>% 
+  mutate(missing_var = if_else(is.na(moisture_closest_2level), "missing yes", "missing no")) %>% 
+  ggplot() +
+    geom_bar(aes(log_pm_mean_sampling_period), color = "red") +
+    facet_wrap(~missing_var)
+missing_hist
+
+# Check ICC 
+multilevel::ICC1(aov(mean_temp_sampling_period_5 ~ area, data = selected_vars))
 
 # Impute missing data; m = 20 datasets (check % missing data and match)
-imputed_data <- mice(selected_vars, m=5,
-                     method = c("polyreg", "pmm", "polr", "logreg", "polr",
+imputed_data <- mice(selected_vars, m=5, maxit = 5,
+                     method = c("polyreg", "polr", "logreg", "polr",
                                 "logreg", "polr", "logreg", "polr", "polr",
-                                "polr", "pmm", "pmm", "polr", "logreg"))
+                                "polr", "pmm", "pmm", "polr", "logreg", "pmm"))
 
-summary(imputed_data)
+# Compare imputed summary to original summary
+summary(selected_vars)
+summary(complete(imputed_data))
+
+# Look at imputation methods for each var; consider changing
+imputed_data$method
 
@@
 
-densityplot(selected_vars$pm_mean_sampling_period)
+densityplot(selected_vars$log_pm_mean_sampling_period)
 
@@
 # See vignette #2 for setting up predictor matrices and relationships b/w vars
+
+# Print predictor matrix
+## Keep considering, but I don't see a reason to alter this as of now
+imputed_data$pred
 
@@
 # Can do regression on the imputed data - either pooled or individual datasets
+```
 
-```
+The above seems to be a good method of imputing the missing data, so moving on 
+to model selection/averaging
 
@@
 # Using MAMI package, perform model averaging and selection
+## Will take ~20 to 30 mins with 5 iterations and bootstrap
 model_ave <- mami(imputed_data, missing.data = "imputed", model = "gaussian",
-                  outcome = "pm_mean_sampling_period", 
+                  outcome = "log_pm_mean_sampling_period", 
                   method = "MA.criterion", criterion = "AIC", print.time = TRUE,
-                  add.transformation = c("log(pm_mean_sampling_period)"))
-
-                  #inference = "+boot", B = 2, X.org = selected_vars
+                  inference = "+boot", B = 2, X.org = selected_vars)
 

### 2020-04-30 ambient_initial.Rmd
ambient temp/pm and exposure model udpates

@@
 author: "Ethan Walker"
-date: "Started 30 March 2020, Updated 3 April 2020"
+date: "Started 30 March 2020, Updated 30 April 2020"
 output: pdf_document
@@
 # Load data
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-wmt_zipcodes <- read_csv("Input/WMT/wmt_zipcodes.csv") %>% 
+# WMT homes now geocoded
+# Instead of zip location, pull home lat/lon from wmt_addresses.rmd and use below
+wmt_geocoded <- addresses_geocoded %>% 
+  mutate(lat_home = lat,
+         long_home = lon,
+         home = Home) %>% 
+  select(home, lat_home, long_home)
+
+wmt_zipcodes <- read_csv(paste0(file_path, "Input/WMT/wmt_zipcodes.csv")) %>% 
   rename_all(tolower)
 
-nn_zipcodes <- read_csv("Input/NN/nn_zipcodes.csv") %>% 
+nn_zipcodes <- read_csv(paste0(file_path, "Input/NN/nn_zipcodes.csv")) %>% 
   rename_all(tolower)
 
-ak_zipcodes <- read_csv("Input/AK/ak_zipcodes.csv") %>% 
+ak_zipcodes <- read_csv(paste0(file_path, "Input/AK/ak_zipcodes.csv")) %>% 
   rename_all(tolower)
@@
 # Zipcode data from: http://download.geonames.org/export/dump/ 
-us_zipcodes <- read_tsv("Input/us_zipcodes/us_zipcodes.txt", 
+us_zipcodes <- read_tsv(paste0(file_path, "Input/us_zipcodes/us_zipcodes.txt"), 
                         col_names = c("country", "zip", "city", "state",
@@
 # https://www7.ncdc.noaa.gov/CDO/cdoselect.cmd?datasetabbv=GSOD&countryabbv=&georegionabbv=
-us_temp_stations <- read_table("Input/ambient/temp_station_list.txt", skip = 20,
+us_temp_stations <- read_table(paste0(file_path, "Input/ambient/temp_station_list.txt"), skip = 20,
                             col_names = c("station", "wban", "station_names",
@@
 # https://epa.maps.arcgis.com/apps/webappviewer/index.html?id=5f239fd3e72f424f98ef3d5def547eb5&extent=-146.2334,13.1913,-46.3896,56.5319 
-pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
+pm_stations <- read_xlsx(paste0(file_path, "Input/ambient/pm_station_list.xlsx"))
 ```
@@
   left_join(us_zipcodes, by = "zip") %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  mutate(lat_zip = lat,
-         long_zip = long)
+  left_join(wmt_geocoded, by = "home")
 
@@
   full_join(wmt_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_temp),
-         long_diff = abs(long_zip - long_temp),
+  mutate(lat_diff = abs(lat_home - lat_temp),
+         long_diff = abs(long_home - long_temp),
          sum_diff = lat_diff + long_diff,
          area = "WMT") %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
+  arrange(home, sum_diff) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   mutate(station_char = factor(station_names, 
@@
                                           "mullanpass"))) %>% 
-  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
-         long_zip, station_names)
+  select(home, area, station, station_char, zip, elevation, lat_temp, long_temp, lat_home, 
+         long_home, station_names)
 
@@
     geom_point(data=wmt_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 3) +
+             aes(long_home, lat_home, color = station_names), size = 3) +
     geom_point(data=wmt_temp_stations, 
@@
   full_join(wmt_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_pm),
-         long_diff = abs(long_zip - long_pm),
+  mutate(lat_diff = abs(lat_home - lat_pm),
+         long_diff = abs(long_home - long_pm),
          sum_diff = lat_diff + long_diff) %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  filter(!is.na(home)) %>% 
-  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
+  arrange(home, sum_diff) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(home, area, local_site, zip, lat_pm, long_pm, lat_home, long_home)
 
@@
     geom_point(data=wmt_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 3) +
+             aes(long_home, lat_home, color = local_site), size = 3) +
     geom_point(data=wmt_pm_stations, 
@@
   left_join(us_zipcodes, by = "zip") %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
-  mutate(lat_zip = lat,
-         long_zip = long)
+  mutate(home = gsub(home, pattern = "_", replacement = ""),
+         home = gsub(home, pattern = " ", replacement = "")) %>%
+  mutate(lat_home = lat,
+         long_home = long)
 
@@
   full_join(nn_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_temp),
-         long_diff = abs(long_zip - long_temp),
+  mutate(lat_diff = abs(lat_home - lat_temp),
+         long_diff = abs(long_home - long_temp),
          sum_diff = lat_diff + long_diff,
          area = "NN") %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
+  arrange(home, sum_diff) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   filter(!is.na(home)) %>% 
@@
                                           "winslowlindberghairport"))) %>% 
-  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
-         long_zip, station_names)
+  select(home, area, station, station_char, zip, elevation, lat_temp, long_temp, lat_home, 
+         long_home, station_names)
 
@@
     geom_point(data=nn_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 3) +
+             aes(long_home, lat_home, color = station_names), size = 3) +
     geom_point(data=nn_temp_stations, 
@@
   full_join(nn_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_pm),
-         long_diff = abs(long_zip - long_pm),
+  mutate(lat_diff = abs(lat_home - lat_pm),
+         long_diff = abs(long_home - long_pm),
          sum_diff = lat_diff + long_diff) %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
+  arrange(home, sum_diff) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   filter(!is.na(home)) %>% 
-  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
+  select(home, area, local_site, zip, lat_pm, long_pm, lat_home, long_home)
 
@@
     geom_point(data=nn_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 3) +
+             aes(long_home, lat_home, color = local_site), size = 3) +
     geom_point(data=nn_pm_stations, 
@@
   full_join(ak_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_temp),
-         long_diff = abs(long_zip - long_temp),
+  mutate(lat_diff = abs(lat_home - lat_temp),
+         long_diff = abs(long_home - long_temp),
          sum_diff = lat_diff + long_diff,
          area = "AK") %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
+  arrange(home, sum_diff) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   mutate(station_char = factor(station_names, 
@@
                                           "quinhagakairport"))) %>% 
-  select(area, station, station_char, zip, elevation, lat_temp, long_temp, lat_zip, 
-         long_zip, station_names)
+  select(home, area, station, station_char, zip, elevation, lat_temp, long_temp, lat_home, 
+         long_home, station_names)
 
@@
     geom_point(data=ak_temp_stations, 
-             aes(long_zip, lat_zip, color = station_names), size = 5) +
+             aes(long_home, lat_home, color = station_names), size = 5) +
     geom_point(data=ak_temp_stations, 
@@
   full_join(ak_zipcodes_new, by = "country") %>% 
-  mutate(lat_diff = abs(lat_zip - lat_pm),
-         long_diff = abs(long_zip - long_pm),
+  mutate(lat_diff = abs(lat_home - lat_pm),
+         long_diff = abs(long_home - long_pm),
          sum_diff = lat_diff + long_diff) %>% 
-  arrange(zip, sum_diff) %>% 
-  distinct(zip, .keep_all = TRUE) %>% 
+  arrange(home, sum_diff) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   filter(!is.na(home)) %>% 
-  select(area, local_site, zip, lat_pm, long_pm, lat_zip, long_zip)
+  select(home, area, local_site, zip, lat_pm, long_pm, lat_home, long_home)
 
@@
     geom_point(data=ak_pm_stations, 
-             aes(long_zip, lat_zip, color = local_site), size = 5) +
+             aes(long_home, lat_home, color = local_site), size = 5) +
     geom_point(data=ak_pm_stations, 
@@
 # Read in ambient weather files and bind together
-list_files <- list.files("Input/ambient/temp_data")
+list_files <- list.files(paste0(file_path, "Input/ambient/temp_data"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next 10 lines together
-setwd("Input/ambient/temp_data") # set new working directory
+setwd(paste0(file_path, "Input/ambient/temp_data")) # set new working directory
 initial_data = tibble(file_name = list_files) %>% # use list of files from above
@@
 
-#write_rds(temp_data_joined, "Output/kids_temp_data.rds")
-#write_csv(temp_data_joined, "Output/kids_temp_data.csv")
+#write_rds(temp_data_joined, paste0(file_path, "Output/kids_temp_data.rds"))
+#write_csv(temp_data_joined, paste0(file_path, "Output/kids_temp_data.csv"))
 
-kids_temp_data <- read_rds("Output/kids_temp_data.rds")
+kids_temp_data <- read_rds(paste0(file_path, "Output/kids_temp_data.rds"))
 ```
@@
 
-pm_stations <- read_xlsx("Input/ambient/pm_station_list.xlsx")
+pm_stations <- read_xlsx(paste0(file_path, "Input/ambient/pm_station_list.xlsx"))
 
 # Read in ambient PM files and bind together
-list_files <- list.files("Input/ambient/pm_data")
+list_files <- list.files(paste0(file_path, "Input/ambient/pm_data"))
 # Set working directory and load files in list; extract file name and add as column
 ## run next 28 lines together
-setwd("Input/ambient/pm_data") # set new working directory
+setwd(paste0(file_path, "Input/ambient/pm_data")) # set new working directory
 initial_data = tibble(file_name = list_files) %>% # use list of files from above
@@
   distinct(pollutant, .keep_all = TRUE) %>% 
-  select(local_site, amb_pm_date, amb_pm_24hr)
+  select(local_site, amb_pm_date, amb_pm_24hr) %>% 
   arrange(local_site, amb_pm_date)
@@
 
-#write_rds(pm_data_joined, "Output/kids_amb_pm_data.rds")
-#write_csv(pm_data_joined, "Output/kids_amb_pm_data.csv")
+#write_rds(pm_data_joined, paste0(file_path, "Output/kids_amb_pm_data.rds"))
+#write_csv(pm_data_joined, paste0(file_path, "Output/kids_amb_pm_data.csv"))
 
-kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")
+kids_amb_pm_data <- read_rds(paste0(file_path, "Output/kids_amb_pm_data.rds"))
 ```
@@
 # Add new variables to pm/temp datasets
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-kids_amb_pm_data <- read_rds("Output/kids_amb_pm_data.rds")
-kids_temp_data <- read_rds("Output/kids_temp_data.rds")
+kids_amb_pm_data <- read_rds(paste0(file_path, "Output/kids_amb_pm_data.rds"))
+kids_temp_data <- read_rds(paste0(file_path, "Output/kids_temp_data.rds"))
 
@@
 kids_amb_pm_data_new <- kids_amb_pm_data %>% 
-  arrange(area, local_site, amb_pm_date, zip) %>% 
-  group_by(area, local_site, zip) %>% 
+  arrange(area, local_site, amb_pm_date, zip, home) %>% 
+  group_by(area, local_site, zip, home) %>% 
   mutate(amb_pm_24hr_lag1 = lag(amb_pm_24hr),
@@
 
-#write_rds(kids_amb_pm_data_new, "Output/kids_amb_pm_data_new.rds")
-#write_csv(kids_amb_pm_data_new, "Output/kids_amb_pm_data_new.csv")
+#write_rds(kids_amb_pm_data_new, paste0(file_path, "Output/kids_amb_pm_data_new.rds"))
+#write_csv(kids_amb_pm_data_new, paste0(file_path, "Output/kids_amb_pm_data_new.csv"))
 
@@
 kids_amb_temp_data_new <- kids_temp_data %>% 
-  arrange(area, station, temp_date, zip) %>% 
-  group_by(area, station, zip) %>% 
+  arrange(area, station, temp_date, zip, home) %>% 
+  group_by(area, station, zip, home) %>% 
   mutate(mean_temp_lag1 = lag(mean_temp),
@@
 
-#write_rds(kids_amb_temp_data_new, "Output/kids_amb_temp_data_new.rds")
-#write_csv(kids_amb_temp_data_new, "Output/kids_amb_temp_data_new.csv")
+#write_rds(kids_amb_temp_data_new, paste0(file_path, "Output/kids_amb_temp_data_new.rds"))
+#write_csv(kids_amb_temp_data_new, paste0(file_path, "Output/kids_amb_temp_data_new.csv"))
 ```

### 2020-04-30 ambient_summary.Rmd
ambient temp/pm and exposure model udpates

--- 2020-04-03 ambient_summary.Rmd
@@
 ```{r}
-exposure_analysis_data_long <- read_rds("Output/exposure_analysis_data_long_new.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-exposure_analysis_data_medium <- read_rds("Output/exposure_analysis_data_medium.rds")
+exposure_analysis_data_long <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
 
-exposure_analysis_data_short <- read_rds("Output/exposure_analysis_data_short.rds")
+exposure_analysis_data_medium <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
 
-sums_rolling_pm_data <- read_rds("Output/sums_rolling_pm_data.rds")
+exposure_analysis_data_short <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds"))
 
-health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
+sums_rolling_pm_data <- 
+  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 ```
@@
                color="black", fill = "azure", size = 1.2) +
-    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
+    geom_point(aes(long_home, lat_home, fill = mean_amb_pm, size = mean_amb_pm), 
                shape = 21, color = "black") +
@@
                color = "black") +
-    geom_point(aes(long_zip, lat_zip, fill = mean_amb_pm, size = mean_amb_pm), 
+    geom_point(aes(long_home, lat_home, fill = mean_amb_pm, size = mean_amb_pm), 
                shape = 21, color = "black") +
@@
                color="black", fill = "azure", size = 1.2) +
-    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
+    geom_point(aes(long_home, lat_home, fill = mean_amb_temp, size = mean_amb_temp), 
                shape = 21, color = "black") +
@@
                color = "black") +
-    geom_point(aes(long_zip, lat_zip, fill = mean_amb_temp, size = mean_amb_temp), 
+    geom_point(aes(long_home, lat_home, fill = mean_amb_temp, size = mean_amb_temp), 
                shape = 21, color = "black") +

### 2020-04-30 collinearity_exposure_analysis.Rmd
ambient temp/pm and exposure model udpates

@@
   select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
-         zip:mean_temp_roll_4day) %>% 
+         amb_pm_24hr:mean_temp_roll_4day) %>% 
   group_by(home) %>% 

### 2020-04-30 exposure_covariate_associations.Rmd
ambient temp/pm and exposure model udpates

@@
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
   group_by(area, home) %>% 
@@
   select(-starts_with("home_act")) %>% 
-  left_join(home_activity_data, by = c("area", "home"))
+  left_join(home_activity_data, by = c("area", "home")) %>% 
+  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
+                                 labels = c("<3", "3+"),
+                                 right = FALSE))
 
@@
 
-  tidy_results <<- tidy(analysis_results) %>% 
+  tidy_results <<- tidy(analysis_results, conf.int = TRUE) %>% 
     filter(term != '(Intercept)') %>% 
-    mutate(estimate = round(estimate, digits = 2),
-           'p-value' = round(p.value, digits = 2),
-           term = gsub("analysis_covar", paste0(label, ": "), term)) %>% 
-    select(term, estimate, 'p-value')
+    mutate(Estimate = round(estimate, digits = 2),
+           'P-value' = round(p.value, digits = 2),
+           'Upper 95% CL' = round(conf.high, digits = 2),
+           'Lower 95% CL' = round(conf.low, digits = 2),
+           Term = gsub("analysis_covar", paste0(label, ": "), term)) %>% 
+    select(Term, Estimate, 'Lower 95% CL', 'Upper 95% CL', 'P-value')
 
@@
 d43 <- analysis_function("sums_events_sampling_period", 
-                         "Mean daily SUMs events - 5+ degree C temp increase in 20 min")
+                         "Daily SUMs events - 5C temp increase in 20 min")
 
@@
 d56 <- analysis_function("home_act_sweep", "Home activity - cleaning")
+d57 <- analysis_function("home_act_smoke_2level", "Home activity - smoking 2 level")
+d58 <- analysis_function("home_act_windows_2level", "Home activity - windows 2 level")
+d59 <- analysis_function("home_act_door_2level", "Home activity - doors 2 level")
+d60 <- analysis_function("home_act_sweep_2level", "Home activity - cleaning 2 level")
 
-d57 <- analysis_function("stove_grade_3level", "Stove grade")
+d61 <- analysis_function("stove_grade_3level", "Stove grade")
 
@@
                     d46, d47, d48, d49, d50, d51, d52, d53, d54, d55, d56,
-                    d57)
+                    d57, d58, d59, d60,
+                    d61)
 
@@
    pack_rows("Ambient data", 54, 55) %>%
-   pack_rows("Home activities - Sum of all sampling days", 56, 66) %>%
-   pack_rows("Stove grade", 67, 68) %>%
+   pack_rows("Home activities - Sum of all sampling days", 56, 69) %>%
+   pack_rows("Stove grade", 70, 71) %>%
    footnote(c("PM2.5 = fine particulate matter", 

### 2020-04-30 exposure_models.Rmd
ambient temp/pm and exposure model udpates

@@
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 16 April 2020"
+date: "Started 19 March 2020, Updated 30 April 2020"
 output: pdf_document
@@
 
-input_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/Input/")
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
-input_file <- c("exposure_analysis_data_long_new.rds")
-output_file <- c("exposure_analysis_data_long_new.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 exposure_analysis_data_long <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
 
 exposure_analysis_data_medium <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
 
 exposure_analysis_data_short <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_short.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds"))
 
 sums_rolling_pm_data <- 
-  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))
 
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 ```
@@
   mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
-  mutate_at(vars(starts_with("home_act")), factor) %>%
   ungroup() %>% 
@@
   left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
-  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) 
+  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) %>% 
+  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
+                                 labels = c("<3", "3+"),
+                                 right = FALSE))
 ```
@@
 selected_vars <- analysis_data %>% 
-  select(area, pm_mean_sampling_period, income_3level, residents_smoke, 
+  select(area, pm_mean_sampling_period, income_3level, 
          stove_age_3level, home_windows_2level,
-         chimney_clean_3level, wood_collect_2level,
-         home_act_door, home_act_sweep, home_act_windows, 
+         chimney_clean_3level, 
+         home_act_door_2level, home_act_windows_2level, home_act_smoke_2level,
          sums_events_sampling_period, mean_temp_sampling_period_5, 
-         stove_grade_3level, moisture_closest_2level) %>% 
-  mutate(home_act_door = factor(home_act_door, levels = c(0,1,2,3,4,5,6)),
-         home_act_sweep = factor(home_act_sweep, levels = c(0,1,2,3,4,5,6)),
-         home_act_windows = factor(home_act_windows, levels = c(0,1,2,3,4,5,6)),
-         area = factor(area, levels = c("AK", "NN", "WMT")),
+         amb_pm_sampling_period_5, stove_grade_3level, moisture_closest_2level) %>% 
+  mutate(area = factor(area, levels = c("AK", "NN", "WMT")),
          log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
   select(-pm_mean_sampling_period) %>% 
-  filter(!is.na(income_3level) & !is.na(home_windows_2level))
+  filter(!is.na(income_3level) & !is.na(home_windows_2level)) %>% 
+  filter(area == "WMT") %>% 
+  select(-area)
 
@@
 imputed_data <- mice(selected_vars, m=5, maxit = 5,
-                     method = c("polyreg", "polr", "logreg", "polr",
-                                "logreg", "polr", "logreg", "polr", "polr",
-                                "polr", "pmm", "pmm", "polr", "logreg", "pmm"))
+                     method = c("polr", "polr", "logreg", "polr", 
+                                "logreg", "logreg", "logreg", "pmm", "pmm",
+                                "pmm", "polr", "logreg", "pmm"))
 

### 2020-04-30 merge_datasets.Rmd
ambient temp/pm and exposure model udpates

@@
   read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
-  mutate(sample_date = amb_pm_date) %>% 
-  distinct(zip, .keep_all = TRUE)
+  mutate(sample_date = amb_pm_date)
 
@@
   mutate(sample_date = temp_date) %>% 
-  mutate(sample_date = as.character(sample_date)) %>% 
-  distinct(zip, .keep_all = TRUE)
-
-wmt_zipcodes <- read_csv(paste0(output_path, "Input/WMT/wmt_zipcodes.csv")) %>% 
-  rename_all(tolower)
-
-nn_zipcodes <- read_csv(paste0(output_path, "Input/NN/nn_zipcodes.csv")) %>% 
-  rename_all(tolower)
-
-ak_zipcodes <- read_csv(paste0(output_path, "Input/AK/ak_zipcodes.csv")) %>% 
-  rename_all(tolower)
-
-kids_zipcodes <- rbind(wmt_zipcodes, nn_zipcodes, ak_zipcodes) %>% 
-  mutate(home = gsub(home, pattern = "_", replacement = ""),
-         home = gsub(home, pattern = " ", replacement = "")) %>% 
-  left_join(kids_amb_pm_data, by = "zip") %>% 
-  select(home, zip, lat_pm, long_pm) %>% 
-  left_join(kids_amb_temp_data, by = "zip") %>% 
-  select(home, zip, lat_zip, long_zip, lat_pm, long_pm, lat_temp, long_temp)
-
-kids_amb_pm_data <- 
-  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
-  mutate(sample_date = amb_pm_date) %>% 
-  select(-lat_zip, -long_zip, -lat_pm, -long_pm)
-
-kids_amb_temp_data <- 
-  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
-  mutate(sample_date = temp_date) %>% 
-  mutate(sample_date = as.character(sample_date)) %>% 
-  select(-lat_zip, -long_zip, -lat_temp, -long_temp)
+  mutate(sample_date = as.character(sample_date))
 
@@
   ungroup() %>% 
-  left_join(kids_zipcodes, by = "home") %>% 
-  left_join(kids_amb_pm_data, by = c("area", "zip", "sample_date")) %>% 
-  left_join(kids_amb_temp_data, by = c("area", "zip", "sample_date")) %>%
+  left_join(kids_amb_pm_data, by = c("area", "home", "sample_date")) %>% 
+  select(-zip, -lat_home, -long_home) %>% 
+  left_join(kids_amb_temp_data, by = c("area", "home", "sample_date")) %>%
   arrange(area, home, home_winter_id, sample_date)

### 2020-05-01 collinearity_exposure_analysis.Rmd
exposure analysis updates

--- 2020-04-17 collinearity_exposure_analysis.Rmd
@@
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  filter(area == "WMT") %>% 
   group_by(area, home) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
+  distinct(home, winter_id, .keep_all = TRUE) %>% 
   ungroup() %>% 
@@
                                  right = FALSE)) %>% 
-  left_join(sums_temp_change, by = c("area", "home")) %>% 
+  group_by(home) %>% 
+  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
+                                   lead(residents_smoke), residents_smoke)) %>% 
+  ungroup() %>% 
+  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
   left_join(ambient_data, by = c("area", "home")) %>% 
   select(-starts_with("home_act")) %>% 
-  left_join(home_activity_data, by = c("area", "home")) %>% 
-  mutate(area = as.factor(area))
+  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
+  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) %>% 
+  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
+                                 labels = c("<1", "1+"),
+                                 right = FALSE),
+         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
+                                 labels = c("<3", "3+"),
+                                 right = FALSE))
 ```
@@
 pm_mean_sampling_period <- analysis_data %>% 
-  select(pm_mean_sampling_period, area, income_3level, 
+  select(pm_mean_sampling_period, income_3level, 
          residents_smoke,
          home_floors_2level, 
-         stove_age_3level, chimney_clean_3level, home_mold,
+         home_pets_2level,
+         chimney_clean_3level, 
+         stove_cert, 
          wood_collect_2level,
+         moisture_closest_2level,
          sums_events_sampling_period,
          mean_temp_sampling_period_5,
-         home_act_smoking, 
-         home_act_door, home_act_sweep,
+         amb_pm_sampling_period_5,
+         home_act_smoke_2level, 
+         home_act_door_2level,
          stove_grade_3level) %>% 
+  #mutate(area = factor(area)) %>% 
   mutate_all(as.numeric) %>% 
@@
 other_vars <- analysis_data %>% 
-  select(pm_mean_sampling_period, area, income_3level, 
+  select(pm_mean_sampling_period, income_3level, 
          residents_smoke,
          home_floors_2level, 
-         stove_age_3level, chimney_clean_3level, home_mold,
+         home_pets_2level,
+         chimney_clean_3level, 
+         stove_cert, 
          wood_collect_2level,
+         moisture_closest_2level,
          sums_events_sampling_period,
          mean_temp_sampling_period_5,
-         home_act_smoking, 
-         home_act_door, home_act_sweep,
+         amb_pm_sampling_period_5,
+         home_act_smoke_2level, 
+         home_act_door_2level,
          stove_grade_3level) %>% 
+  #mutate(area = factor(area)) %>% 
   mutate_all(as.numeric) %>% 
@@
 # Use `omcdiag` to assess overall multicollinearity
-omcdiag(other_vars, pm_mean_sampling_period)
+#omcdiag(other_vars, pm_mean_sampling_period)
 
 # Use `imcdiag` to assess multicollinearity between individual variables
-imcdiag(other_vars, pm_mean_sampling_period)
+#imcdiag(other_vars, pm_mean_sampling_period)
 
 # Specify a method to assess
-imcdiag(other_vars, pm_mean_sampling_period, method = "VIF")
+imcdiag(other_vars, log(pm_mean_sampling_period), method = "VIF")
 ```
@@
 associations with mean PM over the sampling period:
-         pm_mean_sampling_period, area, income_3level, total_residents_2level,
-         race_parent_2level, residents_smoke,
-         home_type_2level, home_floors_2level, home_windows_2level,
-         stove_age_3level, chimney_clean_3level, home_damp, home_mold,
-         home_fireplace,
+         pm_mean_sampling_period, income_3level, area,
+         residents_smoke,
+         home_floors_2level, 
+         home_pets_2level,
+         stove_age_3level, chimney_clean_3level, 
+         stove_cert, home_mold,
          wood_collect_2level,
          moisture_closest_2level,
-         temp_max_5,
          sums_events_sampling_period,
          mean_temp_sampling_period_5,
-         home_act_gas, home_act_smoking, home_act_windows, 
-         home_act_door, home_act_sweep,
-         stove_grade_3level
+         amb_pm_sampling_period_5,
+         home_act_smoke_2level, 
+         home_act_door_2level,
+         stove_grade_3level,
+         race_parent_2level, home_damp, 
+         temp_max_5, home_act_gas, home_type_2level, 
+         home_windows_2level, home_fireplace
 
@@
 
+This is the first sign that `area` has overlap with the other covariates that
+will be difficult to deal with. Removing `area` lowers the VIF for several 
+of the variables.
+
 Based on results and potential overlap, the following vars were removed:
-         race_parent_2level, home_damp, moisture_closest_2level,
-         temp_max_5, home_act_gas, total_residents_2level, home_type_2level, 
-         home_windows_2level, home_fireplace, home_act_windows
+         race_parent_2level, home_damp, home_mold, stove_age_3level,
+         temp_max_5, home_act_gas, home_type_2level, 
+         home_windows_2level, home_fireplace
 
@@
 Demographics: area, income_3level, residents_smoke
-Home Characteristics: home_floors_2level, stove_age_3level,
-                      chimney_clean_3level, home_mold, wood_collect_2level
-Home Activity: home_act_door, home_act_smoking, home_act_sweep
-Others: sums_events_sampling_period, mean_temp_sampling_period_5, stove_grade_3level
+Home Characteristics: home_floors_2level, home_pets_2level, stove_cert, 
+                      chimney_clean_3level, wood_collect_2level, 
+Home Activity: home_act_door_2level, home_act_smoke_2level
+Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
+        stove_grade_3level, moisture_closest_2level, amb_pm_sampling_period_5,
 

### 2020-05-01 exposure_models.Rmd
exposure analysis updates

@@
 ```
+
+**Note**
+
+Add some of the variables created below to the full dataset
 
@@
 
-Initial variables to use for model selection (14) from collinearity RMD:
+**Step 1**
+
+The model selection process begins with simple associations between mean
+PM2.5 over the 6-day sampling period and covariates in the dataset. See
+the `exposure_covariate_associations.rmd` file for this process.
+
+Variables that are associated with PM2.5 are selected and assessed for collinearity
+in the `collinearity_exposure_analysis.rmd`. Variables that have high overlap
+are selected out of the pool of potential covariates to parse down the options;
+however, these variables can be used in sensitivity analyses later.
+
+Initial variables to use for model selection (14):
 Demographics: area, income_3level, residents_smoke
-Home Characteristics: home_floors_2level, stove_age_3level,
-                      chimney_clean_3level, home_mold, wood_collect_2level
-Home Activity: home_act_door, home_act_smoking, home_act_sweep
+Home Characteristics: home_floors_2level, home_pets_2level, stove_cert, 
+                      chimney_clean_3level, wood_collect_2level, 
+Home Activity: home_act_door_2level, home_act_smoke_2level
 Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
-        stove_grade_3level, moisture_closest_2level,
+        stove_grade_3level, moisture_closest_2level, amb_pm_sampling_period_5,
 
-Note: sums_events_sampling_period and mean_temp_sampling_period_5 have a lot of
-missing data and reduce the sample size from 93 to 43. Not considering these
-variables for model selection for now.
-
-Also try adding other variables in to check their impact on model selection:
-   total_residents_2level (seems impactful), home_type_2level (not impactful), 
-   home_windows_2level (seems impactful), home_act_windows (not very impactful),
-   education_3level (not impactful)
-
-Some vars not important in any models, so removing from consideration:
-   home_mold, home_floors_2level, home_act_windows, home_type_2level
-
-Next steps:
+**Missing data notes**
 Some vars (like residents_smoke), other demographics, and home activity vars
@@
 Then, see if data for WMT is more complete in general. The idea is to minimize
-missing data for model selection and the final model to have as much power as
+missing data for model selection and the final model to have as many observations as
 possible. 
@@
 particularly since the ambient data is more reliable for WMT than the other areas.
+(ambient temp/PM data for WMT is based on geolocated lat/long, whereas NN and AK
+are based on zipcode and have much less spatial variation).
 Then you could justify not doing a mixed model, and model selection would be easier.
@@
 moisture, or if there is a spectrum of predictors and complexity in their collection.
+
+**Looking at model selection and averaging after multiple imputation**
+
+After assessing many model selection possibilities, using multiple imputation
+and model selection from imputed datasets seems to be the best option.
+There is not a huge amount of missing data, but it is spread out enough
+that it eliminates a larger percentage of the homes from consideration for
+model selection. Multiple imputation gives us a full dataset to work with,
+and there are documented techniques of using this data for model selection.
+
+Helpful manuals and vignettes:
+
+https://cran.r-project.org/web/packages/mice/mice.pdf
+
+First, select analysis variables
+```{r}
+library(mice)
+#library(pan)
+#library(multilevel)
+
+# Select analysis vars
+selected_vars <- analysis_data %>% 
+  filter(area == "WMT") %>% 
+  select(pm_mean_sampling_period, income_3level, 
+         residents_smoke, home_floors_2level, home_pets_2level,
+         chimney_clean_3level, stove_cert, wood_collect_2level,
+         moisture_closest_2level, sums_events_sampling_period,
+         mean_temp_sampling_period_5, amb_pm_sampling_period_5,
+         home_act_smoke_2level, home_act_door_2level, stove_grade_3level) %>% 
+  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
+  select(-pm_mean_sampling_period) 
+```
+
+**Note**
+
+When selecting data above, WMT has 63 observations
+When use `na.exclude()` to only use complete data (what is needed for AIC selection),
+the observations drops to 30. While there is only 10% missing data, it is spread
+out enough that excluding it to have complete observations means we can only use
+less than half for model selection. This would be a very biased version of model
+selection to exclude that much data, which justifies multiple imputation to 
+impute the 10% missing data and use that for model selection.
+The missing observations are even more extreme for NN and AK.
+
+Next, assess data structure and patterns of missing data
+Document missing data thoroughly for manuscript
+```{r}
+# Look at structure of data and check variable types
+str(selected_vars)
+
+# Check % missing data
+mean(is.na(selected_vars)) # 10%
+
+# Using mice package
+
+# Look for patterns in missing data
+missing_patterns <- md.pattern(selected_vars)
+
+missing_patterns
+
+sum(missing_patterns[, "residents_smoke"] == 0)
+
+hist(selected_vars$pm_mean_sampling_period)
+
+# Check for distribution between vars and missingness - see if there are patterns
+## No obvious patterns in the missing data
+## Missing covars do not seem to be related to the outcomes
+missing_hist <- selected_vars %>% 
+  mutate(missing_var = if_else(is.na(moisture_closest_2level), "missing yes", "missing no")) %>% 
+  ggplot() +
+    geom_bar(aes(log_pm_mean_sampling_period), color = "red") +
+    facet_wrap(~missing_var)
+missing_hist
+
+# Check ICC 
+multilevel::ICC1(aov(mean_temp_sampling_period_5 ~ area, data = selected_vars))
+```
+
+Next, use `mice()` to impute the missing data
+Assess imputed data for summary stats and convergence
+```{r}
+# Impute missing data; m = 10 datasets (check % missing data and match)
+imputed_data <- mice(selected_vars, m=5, maxit = 5,
+                     method = c("polr", "logreg", "logreg", "logreg", "polr",
+                                "polr", "logreg", "logreg", "pmm", "pmm",
+                                "pmm", "logreg", "logreg", "polr", "pmm"))
+
+# Compare imputed summary to original summary
+summary(selected_vars)
+summary(complete(imputed_data))
+
+# Look at imputation methods for each var; consider changing
+imputed_data$method
+
+# Can add "long" or "broad arguments
+complete(imputed_data, "long")
+
+# Sample one of the imputed datasets
+check1 <- complete(imputed_data, 3)
+
+densityplot(selected_vars$log_pm_mean_sampling_period)
+
+# Use seed to get the exact same results from random sampling
+# seed = 123
+
+# See vignette #2 for setting up predictor matrices and relationships b/w vars
+
+# Print predictor matrix
+## Keep considering, but I don't see a reason to alter this as of now
+imputed_data$pred
+
+# Check for convergence
+## In general, we would like the streams to intermingle and be free of any trends 
+## at the later iterations.
+plot(imputed_data)
+
+# Check that the imputation method matches the data type correctly
+## particularly for PM that should be log transformed, and the factors
+
+stripplot(imputed_data)
+
+# Can do regression on the imputed data - either pooled or individual datasets
+```
+
+The above seems to be a good method of imputing the missing data, so moving on 
+to model selection/averaging
+
+http://mami.r-forge.r-project.org/MAMI_manual.pdf
+
+```{r}
+library(MAMI)
+
+# Using MAMI package, perform model averaging and selection
+## Will take ~20 to 30 mins with 5 iterations and bootstrap
+model_ave <- mami(imputed_data, missing.data = "imputed", model = "gaussian",
+                  outcome = "log_pm_mean_sampling_period", 
+                  method = "MA.criterion", criterion = "AIC", print.time = TRUE,
+                  inference = "+boot", B = 2, X.org = selected_vars)
+
+summary(model_ave)
+plot(model_ave)
+```
+
+#####################
+
+Below are other options/ideas for model selection
+Not using for now, but keep for resources
 
@@
 ```
-
-#####################
 
@@
 ```
-
-Looking at model selection and averaging after multiple imputation
-
-Helpful manuals and vignettes:
-
-https://cran.r-project.org/web/packages/mice/mice.pdf
-
-```{r}
-library(mice)
-library(pan)
-library(multilevel)
-
-# Select analysis vars
-selected_vars <- analysis_data %>% 
-  select(area, pm_mean_sampling_period, income_3level, 
-         stove_age_3level, home_windows_2level,
-         chimney_clean_3level, 
-         home_act_door_2level, home_act_windows_2level, home_act_smoke_2level,
-         sums_events_sampling_period, mean_temp_sampling_period_5, 
-         amb_pm_sampling_period_5, stove_grade_3level, moisture_closest_2level) %>% 
-  mutate(area = factor(area, levels = c("AK", "NN", "WMT")),
-         log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
-  select(-pm_mean_sampling_period) %>% 
-  filter(!is.na(income_3level) & !is.na(home_windows_2level)) %>% 
-  filter(area == "WMT") %>% 
-  select(-area)
-
-# Look at structure of data and check variable types
-str(selected_vars)
-
-# Check % missing data
-mean(is.na(selected_vars))
-
-# Using mice package
-
-# Look for patterns in missing data
-missing_patterns <- md.pattern(selected_vars)
-
-missing_patterns
-
-sum(missing_patterns[, "residents_smoke"] == 0)
-
-hist(selected_vars$pm_mean_sampling_period)
-
-# Check for distribution between vars and missingness - see if there are patterns
-## No obvious patterns in the missing data
-## Missing covars do not seem to be related to the outcomes
-missing_hist <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(moisture_closest_2level), "missing yes", "missing no")) %>% 
-  ggplot() +
-    geom_bar(aes(log_pm_mean_sampling_period), color = "red") +
-    facet_wrap(~missing_var)
-missing_hist
-
-# Check ICC 
-multilevel::ICC1(aov(mean_temp_sampling_period_5 ~ area, data = selected_vars))
-
-# Impute missing data; m = 20 datasets (check % missing data and match)
-imputed_data <- mice(selected_vars, m=5, maxit = 5,
-                     method = c("polr", "polr", "logreg", "polr", 
-                                "logreg", "logreg", "logreg", "pmm", "pmm",
-                                "pmm", "polr", "logreg", "pmm"))
-
-# Compare imputed summary to original summary
-summary(selected_vars)
-summary(complete(imputed_data))
-
-# Look at imputation methods for each var; consider changing
-imputed_data$method
-
-# Can add "long" or "broad arguments
-complete(imputed_data, "long")
-
-# Sample one of the imputed datasets
-check1 <- complete(imputed_data, 3)
-
-densityplot(selected_vars$log_pm_mean_sampling_period)
-
-# Use seed to get the exact same results from random sampling
-# seed = 123
-
-# See vignette #2 for setting up predictor matrices and relationships b/w vars
-
-# Print predictor matrix
-## Keep considering, but I don't see a reason to alter this as of now
-imputed_data$pred
-
-# Check for convergence
-## In general, we would like the streams to intermingle and be free of any trends 
-## at the later iterations.
-plot(imputed_data)
-
-# Check that the imputation method matches the data type correctly
-## particularly for PM that should be log transformed, and the factors
-
-stripplot(imputed_data)
-
-# Can do regression on the imputed data - either pooled or individual datasets
-```
-
-The above seems to be a good method of imputing the missing data, so moving on 
-to model selection/averaging
-
-http://mami.r-forge.r-project.org/MAMI_manual.pdf
-
-```{r}
-library(MAMI)
-
-# Using MAMI package, perform model averaging and selection
-## Will take ~20 to 30 mins with 5 iterations and bootstrap
-model_ave <- mami(imputed_data, missing.data = "imputed", model = "gaussian",
-                  outcome = "log_pm_mean_sampling_period", 
-                  method = "MA.criterion", criterion = "AIC", print.time = TRUE,
-                  inference = "+boot", B = 2, X.org = selected_vars)
-
-summary(model_ave)
-plot(model_ave)
-```

### 2020-05-04 exposure_models.Rmd
exposure analysis updates

@@
 selected_vars <- analysis_data %>% 
-  filter(area == "WMT") %>% 
-  select(pm_mean_sampling_period, income_3level, 
+  #filter(area == "WMT") %>% 
+  select(home, area, pm_mean_sampling_period, income_3level, 
          residents_smoke, home_floors_2level, home_pets_2level,
@@
          home_act_smoke_2level, home_act_door_2level, stove_grade_3level) %>% 
-  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
-  select(-pm_mean_sampling_period) 
+  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period))  
+  #select(-pm_mean_sampling_period) 
 ```
@@
 
-hist(selected_vars$pm_mean_sampling_period)
+hist(selected_vars$log_pm_mean_sampling_period)
 
@@
 ## Missing covars do not seem to be related to the outcomes
+
+# Histogram
 missing_hist <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(moisture_closest_2level), "missing yes", "missing no")) %>% 
+  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
   ggplot() +
     geom_bar(aes(log_pm_mean_sampling_period), color = "red") +
-    facet_wrap(~missing_var)
+    facet_wrap(~missing_var) +
+    theme_minimal()
 missing_hist
+
+# Summary Statistics
+missing_summary <- selected_vars %>% 
+  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>%
+  group_by(area, missing_var) %>% 
+  summarize(n = n(),
+            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
+            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+  group_by(area) %>% 
+  mutate(total_n = sum(n),
+         perc_miss = n/total_n*100) %>% 
+  select(area, missing_var, n, perc_miss, mean_pm, sd_pm, median_pm)
+missing_summary
+
+# T-tests
+missing_test <- selected_vars %>% 
+  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
+  filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))
+
+ak_data <- missing_test %>% 
+  filter(area == "AK")
+
+nn_data <- missing_test %>% 
+  filter(area == "NN")
+
+wmt_data <- missing_test %>% 
+  filter(area == "WMT")
+
+t.test(log_pm_mean_sampling_period ~ missing_var, data = ak_data)
+t.test(log_pm_mean_sampling_period ~ missing_var, data = nn_data)
+t.test(log_pm_mean_sampling_period ~ missing_var, data = wmt_data)
+
+# Complete case vs missing data comparison
+complete_case <- selected_vars %>% 
+  na.exclude()
+
+missing_summary <- complete_case %>% 
+  group_by(area) %>% 
+  summarize(n = n(),
+            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
+            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+  group_by(area) %>% 
+  select(area, n, mean_pm, sd_pm, median_pm)
+missing_summary
+
+missing_data <- selected_vars %>% 
+  group_by(home) %>% 
+  filter_at(vars(pm_mean_sampling_period:stove_grade_3level), any_vars(is.na(.)))
+
+missing_summary <- missing_data %>% 
+  group_by(area) %>% 
+  summarize(n = n(),
+            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
+            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+  group_by(area) %>% 
+  select(area, n, mean_pm, sd_pm, median_pm)
+missing_summary
 
@@
 ```{r}
+# set seed to be able to replicate imputation
+set.seed(123)
+
 # Impute missing data; m = 10 datasets (check % missing data and match)
-imputed_data <- mice(selected_vars, m=5, maxit = 5,
+imputed_data <- mice(selected_vars, m=10, maxit = 5,
                      method = c("polr", "logreg", "logreg", "logreg", "polr",

### 2020-05-05 exposure_covariate_associations.Rmd
missing data summary updates

@@
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area, home) %>% 
@@
 
+analysis_function <- function(covar, label, 
+                              dataset = analysis_data){
+
+  analysis_data <- dataset %>% 
+    filter(winter_id == 1 & treatment == "Placebo") %>% 
+    rename(analysis_covar = covar)
+
+  table(analysis_data$analysis_covar)
+
+}
+
 ####### Variables to run through fuction and save results
@@
 d26 <- analysis_function("wood_cords", "Wood cords")
+data_check <- analysis_data %>% 
+  filter(area == "NN")
+table(data_check$education_3level)
+table(!is.na(data_check$mean_temp_sampling_period_5))
 

### 2020-05-05 exposure_models.Rmd
missing data summary updates

@@
 missing_summary <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>%
+  mutate(missing_var = if_else(is.na(stove_grade_3level), "missing yes", "missing no")) %>%
   group_by(area, missing_var) %>% 
@@
 missing_test <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
+  mutate(missing_var = if_else(is.na(home_act_door_2level), "missing yes", "missing no")) %>% 
   filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))
@@
   summarize(n = n(),
-            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+            mean_pm = mean(mean_temp_sampling_period_5, na.rm = TRUE), 
+            sd_pm = sd(mean_temp_sampling_period_5, na.rm = TRUE),
+            median_pm = median(mean_temp_sampling_period_5, na.rm = TRUE)) %>% 
   group_by(area) %>% 

### 2020-05-07 collinearity_exposure_analysis.Rmd
analysis updates

@@
 pm_mean_sampling_period <- analysis_data %>% 
-  select(pm_mean_sampling_period, income_3level, 
+  select(pm_mean_sampling_period, 
+         income_3level, 
          residents_smoke,
+         home_year_built_2level,
          home_floors_2level, 
          home_pets_2level,
+         stove_age_3level,
          chimney_clean_3level, 
-         stove_cert, 
+         home_fireplace,
          wood_collect_2level,
@@
          amb_pm_sampling_period_5,
-         home_act_smoke_2level, 
-         home_act_door_2level,
          stove_grade_3level) %>% 
@@
 other_vars <- analysis_data %>% 
-  select(pm_mean_sampling_period, income_3level, 
+  select(pm_mean_sampling_period, 
+         income_3level, 
          residents_smoke,
+         home_year_built_2level,
          home_floors_2level, 
          home_pets_2level,
+         stove_age_3level,
          chimney_clean_3level, 
-         stove_cert, 
+         home_fireplace,
          wood_collect_2level,

### 2020-05-07 exposure_models.Rmd
analysis updates

--- 2020-05-04 exposure_models.Rmd
@@
   ungroup() %>% 
-  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
-                                 labels = c("<10.5", "10.5+"),
+  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 12, 100),
+                                 labels = c("<12", "12+"),
                                  right = FALSE),
@@
 selected_vars <- analysis_data %>% 
-  #filter(area == "WMT") %>% 
-  select(home, area, pm_mean_sampling_period, income_3level, 
-         residents_smoke, home_floors_2level, home_pets_2level,
-         chimney_clean_3level, stove_cert, wood_collect_2level,
-         moisture_closest_2level, sums_events_sampling_period,
-         mean_temp_sampling_period_5, amb_pm_sampling_period_5,
-         home_act_smoke_2level, home_act_door_2level, stove_grade_3level) %>% 
-  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period))  
-  #select(-pm_mean_sampling_period) 
+  filter(area == "WMT") %>% 
+  select(pm_mean_sampling_period, 
+         income_3level, 
+         residents_smoke,
+         home_year_built_2level,
+         home_floors_2level, 
+         home_pets_2level,
+         stove_age_3level,
+         chimney_clean_3level, 
+         home_fireplace,
+         wood_collect_2level,
+         moisture_closest_2level,
+         sums_events_sampling_period,
+         mean_temp_sampling_period_5,
+         amb_pm_sampling_period_5,
+         stove_grade_3level) %>% 
+  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
+  select(-pm_mean_sampling_period) 
 ```
@@
 missing_summary <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>%
+  mutate(missing_var = if_else(is.na(stove_age_3level), "missing yes", "missing no")) %>%
   group_by(area, missing_var) %>% 
@@
 missing_test <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
+  mutate(missing_var = if_else(is.na(stove_age_3level), "missing yes", "missing no")) %>% 
   filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))
@@
   summarize(n = n(),
-            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+            mean_var = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            sd_var = sd(pm_mean_sampling_period, na.rm = TRUE),
+            median_var = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
   group_by(area) %>% 
-  select(area, n, mean_pm, sd_pm, median_pm)
+  select(area, n, mean_var, sd_var, median_var)
+missing_summary
+
+missing_summary <- complete_case %>% 
+  group_by(area, wood_collect_2level) %>% 
+  summarize(n = n()) %>% 
+  group_by(area) %>% 
+  mutate(total_n = sum(n),
+         perc = n/total_n*100) %>% 
+  select(area, wood_collect_2level, n, perc, total_n)
 missing_summary
@@
   summarize(n = n(),
-            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+            mean_var = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            sd_var = sd(pm_mean_sampling_period, na.rm = TRUE),
+            median_var = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
   group_by(area) %>% 
-  select(area, n, mean_pm, sd_pm, median_pm)
+  select(area, n, mean_var, sd_var, median_var)
+missing_summary
+
+missing_summary <- missing_data %>% 
+  group_by(area, wood_collect_2level) %>% 
+  summarize(n = n()) %>% 
+  group_by(area) %>% 
+  mutate(total_n = sum(n),
+         perc = n/total_n*100) %>% 
+  select(area, wood_collect_2level, n, perc, total_n)
 missing_summary
@@
 # Impute missing data; m = 10 datasets (check % missing data and match)
-imputed_data <- mice(selected_vars, m=10, maxit = 5,
-                     method = c("polr", "logreg", "logreg", "logreg", "polr",
-                                "polr", "logreg", "logreg", "pmm", "pmm",
-                                "pmm", "logreg", "logreg", "polr", "pmm"))
+imputed_data <- mice(selected_vars, m=10, maxit = 40,
+                     method = c("polr", "logreg", "logreg", "logreg", "logreg",
+                                "polr", "polr", "logreg", "logreg", "logreg",
+                                "pmm", "pmm", "pmm", "polr", "pmm"))
+
+write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
+
+imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
 
@@
                   outcome = "log_pm_mean_sampling_period", 
-                  method = "MA.criterion", criterion = "AIC", print.time = TRUE,
-                  inference = "+boot", B = 2, X.org = selected_vars)
+                  method = "MA.criterion", criterion = "AIC", print.time = TRUE)
 
-summary(model_ave)
-plot(model_ave)
+write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
+
+exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
+
+summary(test_model_results)
+
+# plot(model_ave) # only used when bootstrap used in mami
+
+print(exposure_model)
 ```

### 2020-05-11 demographics_summary.Rmd
update demographics data

@@
 # Load individual datasets
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-demographics_clean <- read_rds("Output/demographics_clean.rds") 
-pm_clean <- read_rds("Output/pm_clean.rds")
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+demographics_clean <- read_rds(paste0(file_path, "Output/demographics_clean.rds")) 
+pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
 ```
@@
 # Summarize numeric vars
-kids_dems <- demographics_filtered %>%  
-  mutate(total_residents = as.numeric(total_residents),
-         residents_under_five = as.numeric(residents_under_five)) %>% 
+kids_dems <- demographics_clean %>%  
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>%
   distinct(home, .keep_all = TRUE) %>%  
+  filter(!is.na(gender_parent)) %>% 
   group_by(area) %>% 

### 2020-05-11 exposure_models.Rmd
update demographics data

@@
   filter(treatment == "Placebo") %>% 
-  #filter(area == "WMT") %>% 
+  filter(area == "WMT") %>% 
   group_by(area, home) %>% 

### 2020-05-11 exposure_prediction.Rmd
update demographics data

--- 2020-05-11 exposure_models.Rmd
@@
 ---
-title: "KidsAIR: exposure model"
+title: "KidsAIR: exposure prediction"
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 30 April 2020"
+date: "Started 11 May 2020, Updated 11 May 2020"
 output: pdf_document
@@
 
+exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
+
+imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
+
 exposure_analysis_data_long <- 
@@
 
-**Note**
+```{r}
+winter_2_dates <- exposure_analysis_data_short %>% 
+  select(home, area, pm_first_date) %>% 
+  mutate(pm_winter_2_start = pm_first_date + 365,
+         pm_winter_2_end = pm_winter_2_start + 6)
 
-Add some of the variables created below to the full dataset
+winter_2_homes <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  filter(area == "WMT") %>% 
+  distinct(home, winter_id, .keep_all = TRUE) %>% 
+  select(home, area)
+
+winter_2_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 2) %>% 
+  filter(treatment == "Placebo") %>% 
+  filter(area == "WMT") %>% 
+  select(area, home, winter_id, income_3level, residents_smoke, home_floors_2level,
+         home_year_built_2level, chimney_clean_3level, wood_collect_2level,
+         home_fireplace) %>% 
+  distinct(home, winter_id, .keep_all = TRUE)  %>% 
+  right_join(winter_2_homes, by = c("area", "home"))
+```
+
+Currently (11 May 2020) waiting on final year field data to be entered and loaded
+to Box so I can clean it and form a full dataset. Until then, there is Winter 2
+data missing from Montana home and I can't predict PM with the Winter 1 model.
+
+Next steps:
+-Clean Winter 2 data
+-Join to form full dataset from both Winters
+-Use code above to look at missing data from variables used in the predictive model
+-Use code below to pull in further variables (sums_events_sampling_period, mean_temp_sampling_period_5)
+ -This is just example code - needs cleaned up to pull the vars for Winter 2
+ -Goal is to predict Winter 2 PM for 1 year after the Winter 1 PM was sampled
+ -So, need to pull out a 6 day mean for ambient temp and SUMs var
+ -Then join those vars with Household Winter 2 data from above
 
@@
 ```
-
-**Step 1**
-
-The model selection process begins with simple associations between mean
-PM2.5 over the 6-day sampling period and covariates in the dataset. See
-the `exposure_covariate_associations.rmd` file for this process.
-
-Variables that are associated with PM2.5 are selected and assessed for collinearity
-in the `collinearity_exposure_analysis.rmd`. Variables that have high overlap
-are selected out of the pool of potential covariates to parse down the options;
-however, these variables can be used in sensitivity analyses later.
-
-Initial variables to use for model selection (14):
-Demographics: area, income_3level, residents_smoke
-Home Characteristics: home_floors_2level, home_pets_2level, stove_cert, 
-                      chimney_clean_3level, wood_collect_2level, 
-Home Activity: home_act_door_2level, home_act_smoke_2level
-Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
-        stove_grade_3level, moisture_closest_2level, amb_pm_sampling_period_5,
-
-**Missing data notes**
-Some vars (like residents_smoke), other demographics, and home activity vars
-have quite a bit of missing data. This is in addition to the missing ambient
-and SUMs data. The first step should be to account for as much of this as possible.
-See if there are ways to pull demographic data from Winter 2.
-Then, see if data for WMT is more complete in general. The idea is to minimize
-missing data for model selection and the final model to have as many observations as
-possible. 
-Consider doing a more "full" model for WMT only (if there is less missing data),
-particularly since the ambient data is more reliable for WMT than the other areas.
-(ambient temp/PM data for WMT is based on geolocated lat/long, whereas NN and AK
-are based on zipcode and have much less spatial variation).
-Then you could justify not doing a mixed model, and model selection would be easier.
-Then, using data from all areas, use the full model from WMT to work on a 
-mixed model that is likely reduced due to sparse data from the other areas.
-Since model selection is difficult with mixed models, use the WMT model to inform
-this model, while also doing sensitivity analysis to check other variables.
-Aside from the multivariable model, much of the paper could also be looking at
-simple associations between PM and some of the novel covars (SUMS, ambient data,
-stove grades, wood moisture), as well as methods for collecting and analysing 
-those vars.
-Also look into predict() function for teach/test and validating the model.
-Discussion can focus on what important predictors of indoor PM are in these homes - 
-if simple demographic data predicts as well as complex things like SUMs and wood
-moisture, or if there is a spectrum of predictors and complexity in their collection.
-
-**Looking at model selection and averaging after multiple imputation**
-
-After assessing many model selection possibilities, using multiple imputation
-and model selection from imputed datasets seems to be the best option.
-There is not a huge amount of missing data, but it is spread out enough
-that it eliminates a larger percentage of the homes from consideration for
-model selection. Multiple imputation gives us a full dataset to work with,
-and there are documented techniques of using this data for model selection.
-
-Helpful manuals and vignettes:
-
-https://cran.r-project.org/web/packages/mice/mice.pdf
-
-First, select analysis variables
-```{r}
-library(mice)
-#library(pan)
-#library(multilevel)
-
-# Select analysis vars
-selected_vars <- analysis_data %>% 
-  filter(area == "WMT") %>% 
-  select(pm_mean_sampling_period, 
-         income_3level, 
-         residents_smoke,
-         home_year_built_2level,
-         home_floors_2level, 
-         home_pets_2level,
-         stove_age_3level,
-         chimney_clean_3level, 
-         home_fireplace,
-         wood_collect_2level,
-         moisture_closest_2level,
-         sums_events_sampling_period,
-         mean_temp_sampling_period_5,
-         amb_pm_sampling_period_5,
-         stove_grade_3level) %>% 
-  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>% 
-  select(-pm_mean_sampling_period) 
-```
-
-**Note**
-
-When selecting data above, WMT has 63 observations
-When use `na.exclude()` to only use complete data (what is needed for AIC selection),
-the observations drops to 30. While there is only 10% missing data, it is spread
-out enough that excluding it to have complete observations means we can only use
-less than half for model selection. This would be a very biased version of model
-selection to exclude that much data, which justifies multiple imputation to 
-impute the 10% missing data and use that for model selection.
-The missing observations are even more extreme for NN and AK.
-
-Next, assess data structure and patterns of missing data
-Document missing data thoroughly for manuscript
-```{r}
-# Look at structure of data and check variable types
-str(selected_vars)
-
-# Check % missing data
-mean(is.na(selected_vars)) # 10%
-
-# Using mice package
-
-# Look for patterns in missing data
-missing_patterns <- md.pattern(selected_vars)
-
-missing_patterns
-
-sum(missing_patterns[, "residents_smoke"] == 0)
-
-hist(selected_vars$log_pm_mean_sampling_period)
-
-# Check for distribution between vars and missingness - see if there are patterns
-## No obvious patterns in the missing data
-## Missing covars do not seem to be related to the outcomes
-
-# Histogram
-missing_hist <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
-  ggplot() +
-    geom_bar(aes(log_pm_mean_sampling_period), color = "red") +
-    facet_wrap(~missing_var) +
-    theme_minimal()
-missing_hist
-
-# Summary Statistics
-missing_summary <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(stove_age_3level), "missing yes", "missing no")) %>%
-  group_by(area, missing_var) %>% 
-  summarize(n = n(),
-            mean_pm = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_pm = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_pm = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
-  group_by(area) %>% 
-  mutate(total_n = sum(n),
-         perc_miss = n/total_n*100) %>% 
-  select(area, missing_var, n, perc_miss, mean_pm, sd_pm, median_pm)
-missing_summary
-
-# T-tests
-missing_test <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(stove_age_3level), "missing yes", "missing no")) %>% 
-  filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))
-
-ak_data <- missing_test %>% 
-  filter(area == "AK")
-
-nn_data <- missing_test %>% 
-  filter(area == "NN")
-
-wmt_data <- missing_test %>% 
-  filter(area == "WMT")
-
-t.test(log_pm_mean_sampling_period ~ missing_var, data = ak_data)
-t.test(log_pm_mean_sampling_period ~ missing_var, data = nn_data)
-t.test(log_pm_mean_sampling_period ~ missing_var, data = wmt_data)
-
-# Complete case vs missing data comparison
-complete_case <- selected_vars %>% 
-  na.exclude()
-
-missing_summary <- complete_case %>% 
-  group_by(area) %>% 
-  summarize(n = n(),
-            mean_var = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_var = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_var = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
-  group_by(area) %>% 
-  select(area, n, mean_var, sd_var, median_var)
-missing_summary
-
-missing_summary <- complete_case %>% 
-  group_by(area, wood_collect_2level) %>% 
-  summarize(n = n()) %>% 
-  group_by(area) %>% 
-  mutate(total_n = sum(n),
-         perc = n/total_n*100) %>% 
-  select(area, wood_collect_2level, n, perc, total_n)
-missing_summary
-
-missing_data <- selected_vars %>% 
-  group_by(home) %>% 
-  filter_at(vars(pm_mean_sampling_period:stove_grade_3level), any_vars(is.na(.)))
-
-missing_summary <- missing_data %>% 
-  group_by(area) %>% 
-  summarize(n = n(),
-            mean_var = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_var = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_var = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
-  group_by(area) %>% 
-  select(area, n, mean_var, sd_var, median_var)
-missing_summary
-
-missing_summary <- missing_data %>% 
-  group_by(area, wood_collect_2level) %>% 
-  summarize(n = n()) %>% 
-  group_by(area) %>% 
-  mutate(total_n = sum(n),
-         perc = n/total_n*100) %>% 
-  select(area, wood_collect_2level, n, perc, total_n)
-missing_summary
-
-# Check ICC 
-multilevel::ICC1(aov(mean_temp_sampling_period_5 ~ area, data = selected_vars))
-```
-
-Next, use `mice()` to impute the missing data
-Assess imputed data for summary stats and convergence
-```{r}
-# set seed to be able to replicate imputation
-set.seed(123)
-
-# Impute missing data; m = 10 datasets (check % missing data and match)
-imputed_data <- mice(selected_vars, m=10, maxit = 40,
-                     method = c("polr", "logreg", "logreg", "logreg", "logreg",
-                                "polr", "polr", "logreg", "logreg", "logreg",
-                                "pmm", "pmm", "pmm", "polr", "pmm"))
-
-write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
-
-imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
-
-# Compare imputed summary to original summary
-summary(selected_vars)
-summary(complete(imputed_data))
-
-# Look at imputation methods for each var; consider changing
-imputed_data$method
-
-# Can add "long" or "broad arguments
-complete(imputed_data, "long")
-
-# Sample one of the imputed datasets
-check1 <- complete(imputed_data, 3)
-
-densityplot(selected_vars$log_pm_mean_sampling_period)
-
-# Use seed to get the exact same results from random sampling
-# seed = 123
-
-# See vignette #2 for setting up predictor matrices and relationships b/w vars
-
-# Print predictor matrix
-## Keep considering, but I don't see a reason to alter this as of now
-imputed_data$pred
-
-# Check for convergence
-## In general, we would like the streams to intermingle and be free of any trends 
-## at the later iterations.
-plot(imputed_data)
-
-# Check that the imputation method matches the data type correctly
-## particularly for PM that should be log transformed, and the factors
-
-stripplot(imputed_data)
-
-# Can do regression on the imputed data - either pooled or individual datasets
-```
-
-The above seems to be a good method of imputing the missing data, so moving on 
-to model selection/averaging
-
-http://mami.r-forge.r-project.org/MAMI_manual.pdf
-
-```{r}
-library(MAMI)
-
-# Using MAMI package, perform model averaging and selection
-## Will take ~20 to 30 mins with 5 iterations and bootstrap
-model_ave <- mami(imputed_data, missing.data = "imputed", model = "gaussian",
-                  outcome = "log_pm_mean_sampling_period", 
-                  method = "MA.criterion", criterion = "AIC", print.time = TRUE)
-
-write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
-
-exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
-
-summary(test_model_results)
-
-# plot(model_ave) # only used when bootstrap used in mami
-
-print(exposure_model)
-```
-
-#####################
-
-Below are other options/ideas for model selection
-Not using for now, but keep for resources
-
-```{r, include=FALSE}
-full_model <- lm(log(pm_mean_sampling_period) ~ income_3level + 
-                 stove_age_3level + home_windows_2level +
-                 chimney_clean_3level + wood_collect_2level + 
-                 home_act_door + home_act_smoking + home_act_sweep + 
-                 stove_grade_3level + moisture_closest_2level + 
-                 total_residents_2level + mean_temp_sampling_period_5 +
-                 sums_events_sampling_period, 
-                 data = analysis_data)
-
-summary(full_model)
-r.squaredGLMM(full_model)
-tidy(full_model)
-
-plot(full_model)
-```
-
-# Best subsets selection using dredge() from MuMIn
-
-Make sure to run `na.exclude` on analysis dataset and `options(na.action = "na.fail")`
-before running the dredge best-subsets function below. This will ensure that the 
-same dataset is being used for each potential model, which is an assumption and
-a prerequisite for comparing different model ranking like AIC.
-
-```{r, eval=FALSE, include=FALSE}
-options(na.action = "na.fail")
-
-all_subsets_aic <- dredge(full_model, rank = "AIC", extra = c("R^2"))
-
-all_subsets_aicc <- dredge(full_model, rank = "AICc", extra = c("R^2"))
-par(mar = c(3,5,6,4))
-plot(all_subsets_aicc)
-
-all_subsets_bic <- dredge(full_model, rank = "BIC", extra = c("R^2"))
-
-write_rds(all_subsets_aic, paste0(output_path, "Output/all_subsets_aic"))
-write_rds(all_subsets_aicc, paste0(output_path, "Output/all_subsets_aicc"))
-write_rds(all_subsets_bic, paste0(output_path, "Output/all_subsets_bic"))
-```
-
-```{r}
-all_subsets_aic <- read_rds(paste0(output_path, "Output/all_subsets_aic"))
-all_subsets_aicc <- read_rds(paste0(output_path, "Output/all_subsets_aicc"))
-all_subsets_bic <- read_rds(paste0(output_path, "Output/all_subsets_bic"))
-
-head(all_subsets_aic, 10)
-head(all_subsets_aicc, 10)
-head(all_subsets_bic, 10)
-```
-
-```{r}
-subset_model <- lm(log(pm_mean_sampling_period) ~ 
-                   income_3level + home_act_door + residents_smoke, 
-                   data = analysis_data)
-
-summary(subset_model)
-#tidy(subset_model)
-
-plot(subset_model)
-```
-
-Looking at Lasso regression
-May be beyond the scope of this paper, but keep code and references:
-
-https://cran.r-project.org/web/packages/MuMIn/MuMIn.pdf
-
-http://www.science.smith.edu/~jcrouser/SDS293/labs/lab10-r.html
-
-https://cran.r-project.org/web/packages/glmnet/glmnet.pdf
-
-https://cran.r-project.org/web/packages/glmmLasso/glmmLasso.pdf
-
-https://arxiv.org/pdf/1306.2427.pdf
-
-```{r} 
-library(MuMIn)
-library(leaps)
-library(glmnet)
-library(glmmLasso)
-
-# Range of lambdas
-grid = 10^seq(10, -2, length = 100)
-
-lasso_results <- glmmLasso(log(pm_mean_sampling_period) ~ income_3level + 
-                 residents_smoke + stove_age_3level + home_windows_2level +
-                 chimney_clean_3level + wood_collect_2level + 
-                 home_act_door + home_act_smoking + home_act_sweep + 
-                 stove_grade_3level + moisture_closest_2level + 
-                 total_residents_2level,
-                 rnd = list(area = ~ 1), 
-                 lambda = grid,
-                 data = analysis_data)
-
-summary(lasso_results)
-
-predict(lasso_results)
-```
-
-Looking at Ridge regression:
-
-https://drsimonj.svbtle.com/ridge-regression-with-glmnet
-
-```{r}
-ridge_results <- cv.glmnet(x = ind_vars, y = dep_var, alpha = 0)
-
-plot(ridge_results)
-
-opt_lambda <- ridge_results$lambda.min
-opt_lambda
-
-fit <- ridge_results$glmnet.fit
-summary(fit)
-
-y_predicted <- predict(fit, s = opt_lambda, newx = ind_vars)
-
-# Sum of Squares Total and Error
-sst <- sum((dep_var - mean(dep_var))^2)
-sse <- sum((y_predicted - dep_var)^2)
-
-# R squared
-rsq <- 1 - sse / sst
-rsq
-```
-
-# Train and test datasets
-
-```{r}
-# Random sample indexes
-train_index <- sample(1:nrow(adult), 0.8 * nrow(adult))
-test_index <- setdiff(1:nrow(adult), train_index)
-
-dep_var <- analysis_data$pm_mean_sampling_period
-
-ind_vars <- analysis_data %>% 
-  select(-pm_mean_sampling_period, -area) %>% 
-  data.matrix()
-```

### 2020-05-11 home_child_data_cleaning.Rmd
update demographics data

--- 2020-01-22 home_child_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 17 Dec 2019"
+date: "Started 19 Nov 2019, Updated 11 May 2020"
 output: pdf_document
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 # Load files 
@@
 ## 2019-November-Data > WMT > child.xlsx
-wmt_child_initial <- read_xlsx("Input/WMT/wmt_child_20191126.xlsx") %>% 
+wmt_child_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_child_20191126.xlsx")) %>% 
   rename_all(tolower) %>% 
+  # filter(enrolled == 1) %>% # filter out kids who weren't enrolled
   select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
@@
 ## 2019-November-Data > WMT > home.xlsx
-wmt_home_initial <- read_xlsx("Input/WMT/wmt_home_20191126.xlsx") %>% 
+wmt_home_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_home_20191126.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## 2019-November-Data > WMT > homewinter.xlsx
-wmt_homewinter_initial <- read_xlsx("Input/WMT/wmt_homewinter_20191126.xlsx") %>% 
+wmt_homewinter_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_homewinter_20191126.xlsx")) %>% 
   rename_all(tolower) %>% 
   select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  filter(winterid != 3) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   mutate(home_winter_id = as.character(homewinterid),
@@
 wmt_parent_demographics_initial <-
-  read_xlsx("Input/WMT/wmt_parent_demographics_20191126.xlsx") %>% 
+  read_xlsx(paste0(file_path, "Input/WMT/wmt_parent_demographics_20191126.xlsx")) %>% 
   rename_all(tolower) %>% 
+  filter(whatmissing != "all" | is.na(whatmissing)) %>%  # filter out if missing == "all"
   select(homewinterid, gender, hispanic, race, income, education,
@@
   select(home_winter_id, gender_parent, hispanic_parent, race_parent,
-         income, education, total_residents, residents_under_five) %>% 
-  filter(!is.na(gender_parent))
+         income, education, total_residents, residents_under_five)
 
@@
   left_join(wmt_child_initial, by = "home_id") %>% 
-  mutate(area = "WMT")
+  mutate(area = "WMT") %>% 
+  filter(!is.na(enrolled))
 
@@
   #distinct(child_id_num, .keep_all = TRUE) %>% 
-  arrange(home_id, child_id_num, winter_id) %>% 
-  filter(winter_id == 1 | winter_id == 2)
+  arrange(home_id, child_id_num, winter_id)
 
 # Save as RDS and CSV
-write_rds(wmt_home_ids_linked, "Output/wmt_ids_linked.rds")
-write_rds(wmt_child_demographics, "Output/wmt_demographic_data.rds")
+write_rds(wmt_home_ids_linked, paste0(file_path, "Output/wmt_ids_linked.rds"))
+write_rds(wmt_child_demographics, paste0(file_path, "Output/wmt_demographic_data.rds"))
 ```
@@
 ## 2019-November-Data > Navajo > child.xlsx
-nn_child_initial <- read_xlsx("Input/NN/nn_child_20191216.xlsx") %>% 
+nn_child_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_child_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
+  # filter(enrolled == 1) %>% # filter out kids who weren't enrolled
   select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
@@
          child_id_char = as.character(id),
+         child_id_char = gsub(child_id_char, pattern = " ", replacement = ""),
+         child_id_char = gsub(child_id_char, pattern = "_", replacement = ""),
+         child_id_char = gsub(child_id_char, pattern = "-", replacement = ""),
          hispanic_child = as.factor(hispanic),
@@
 ## 2019-November-Data > Navajo > home.xlsx
-nn_home_initial <- read_xlsx("Input/NN/nn_home_20191216.xlsx") %>% 
+nn_home_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_home_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## 2019-November-Data > Navajo > homewinter.xlsx
-nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20191216.xlsx") %>% 
+nn_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_homewinter_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
   select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  filter(winterid != 3) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   mutate(home_winter_id = as.character(homewinterid),
@@
 nn_parent_demographics_initial <-
-  read_xlsx("Input/NN/nn_parent_demographics_20191216.xlsx") %>% 
+  read_xlsx(paste0(file_path, "Input/NN/nn_parent_demographics_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
+  filter(whatmissing != "all" | is.na(whatmissing)) %>%  # filter out if missing == "all"
   select(homewinterid, gender, hispanic, race, income, education,
@@
   left_join(nn_child_initial, by = "home_id") %>% 
-  mutate(area = "NN")
+  mutate(area = "NN") %>% 
+  filter(!is.na(enrolled))
 
@@
   #distinct(child_id_num, .keep_all = TRUE) %>% 
-  arrange(home_id, child_id_num, winter_id) %>% 
-  filter(winter_id == 1 | winter_id == 2)
+  arrange(home_id, child_id_num, winter_id)
 
 # Save as RDS and CSV
-write_rds(nn_home_ids_linked, "Output/nn_ids_linked.rds")
-write_rds(nn_child_demographics, "Output/nn_demographic_data.rds")
+write_rds(nn_home_ids_linked, paste0(file_path, "Output/nn_ids_linked.rds"))
+write_rds(nn_child_demographics, paste0(file_path, "Output/nn_demographic_data.rds"))
 ```
@@
 ## 2019-November-Data > AK > child.xlsx
-ak_child_initial <- read_xlsx("Input/AK/ak_child_20191216.xlsx") %>% 
+ak_child_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_child_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
+  # filter(enrolled == 1) %>% # filter out kids who weren't enrolled
   select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
@@
 ## 2019-November-Data > AK > home.xlsx
-ak_home_initial <- read_xlsx("Input/AK/ak_home_20191216.xlsx") %>% 
+ak_home_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_home_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
@@
 ## 2019-November-Data > AK > homewinter.xlsx
-ak_homewinter_initial <- read_xlsx("Input/AK/ak_homewinter_20191216.xlsx") %>% 
+ak_homewinter_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_homewinter_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
   select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  filter(winterid != 3) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   mutate(home_winter_id = as.character(homewinterid),
@@
 ak_parent_demographics_initial <-
-  read_xlsx("Input/AK/ak_parent_demographics_20191216.xlsx") %>% 
+  read_xlsx(paste0(file_path, "Input/AK/ak_parent_demographics_20191216.xlsx")) %>% 
   rename_all(tolower) %>% 
+  filter(whatmissing != "all" | is.na(whatmissing)) %>%  # filter out if missing == "all"
   select(homewinterid, gender, hispanic, race, income, education,
@@
   left_join(ak_child_initial, by = "home_id") %>% 
-  mutate(area = "AK")
+  mutate(area = "AK") %>% 
+  filter(!is.na(enrolled))
 
@@
   #distinct(child_id_num, .keep_all = TRUE) %>% 
-  arrange(home_id, child_id_num, winter_id) %>% 
-  filter(winter_id == 1 | winter_id == 2)
+  arrange(home_id, child_id_num, winter_id)
 
 # Save as RDS and CSV
-write_rds(ak_home_ids_linked, "Output/ak_ids_linked.rds")
-write_rds(ak_child_demographics, "Output/ak_demographic_data.rds")
+write_rds(ak_home_ids_linked, paste0(file_path, "Output/ak_ids_linked.rds"))
+write_rds(ak_child_demographics, paste0(file_path, "Output/ak_demographic_data.rds"))
 ```
@@
 ```{r}
-ak_ids <- read_rds("Output/ak_ids_linked.rds")
-nn_ids <- read_rds("Output/nn_ids_linked.rds")
-wmt_ids <- read_rds("Output/wmt_ids_linked.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+ak_ids <- read_rds(paste0(file_path, "Output/ak_ids_linked.rds"))
+nn_ids <- read_rds(paste0(file_path, "Output/nn_ids_linked.rds"))
+wmt_ids <- read_rds(paste0(file_path, "Output/wmt_ids_linked.rds"))
 
@@
 
-#write_rds(kids_ids, "Output/kids_linked_ids.rds")
+#write_rds(kids_ids, paste0(file_path, "Output/kids_linked_ids.rds")
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds")) %>% 
   mutate(winter_id = as.numeric(winter_id),
          winter_id = if_else(home == "CH218", winter_id + 1, winter_id)) %>% 
-  filter(winter_id > 0)
+  filter(winter_id == 1 | winter_id == 2)
 
-#write_rds(kids_linked_ids, "Output/kids_linked_ids.rds")
+#write_rds(kids_linked_ids, paste0(file_path, "Output/kids_linked_ids.rds"))
 ```
@@
 ```{r}
-ak_demographics <- read_rds("Output/ak_demographic_data.rds")
-nn_demographics <- read_rds("Output/nn_demographic_data.rds")
-wmt_demographics <- read_rds("Output/wmt_demographic_data.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+ak_demographics <- read_rds(paste0(file_path, "Output/ak_demographic_data.rds"))
+nn_demographics <- read_rds(paste0(file_path, "Output/nn_demographic_data.rds"))
+wmt_demographics <- read_rds(paste0(file_path, "Output/wmt_demographic_data.rds"))
 
@@
 
-# write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
+# write_rds(kids_demographics, paste0(file_path, "Output/kids_combined_demographics.rds"))
 
-kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
+kids_demographics <- read_rds(paste0(file_path, "Output/kids_combined_demographics.rds"))
 ```
@@
 ```{r}
-kids_demographics <- read_rds("Output/kids_combined_demographics.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_demographics <- read_rds(paste0(file_path, "Output/kids_combined_demographics.rds"))
 
@@
   mutate(intervention_datetime = ymd_hms(intervention_datetime)) %>% 
-  select(-trash)
+  select(-trash) %>% 
+  mutate(home_winter_id = as.numeric(home_winter_id),
+         home_winter_id = if_else(home == "CH218" & winter_id == 1, 77, home_winter_id),
+         home_winter_id = if_else(home == "CH218" & winter_id == 2, 111, home_winter_id),
+         home_winter_id = as.factor(home_winter_id)) %>% 
+  filter(winter_id == 1 | winter_id == 2)
 
-# write_rds(demographics_clean, "Output/demographics_clean.rds")
+# write_rds(demographics_clean, paste0(file_path, "Output/demographics_clean.rds"))
 ```

### 2020-05-12 demographics_summary.Rmd
demographics/id filtering and merging

@@
 author: "Ethan Walker"
-date: "Started 27 Jan 2020, Updated 25 March 2020"
+date: "Started 27 Jan 2020, Updated 12 May 2020"
 output:
@@
 kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
-demographics_clean <- read_rds(paste0(file_path, "Output/demographics_clean.rds")) 
+demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds")) 
 pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
+```
+
+```{r}
+# number of placebo households, total and by area
+placebo_homes <- kids_linked_ids %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  count(area)
+
+# number of placebo households, total and by area
+treatment_homes <- kids_linked_ids %>% 
+  filter(winter_id == 1) %>% 
+  count(treatment)
+
+# number of placebo households, total and by area
+treatment_kids <- demographics_initial %>% 
+  filter(winter_id == 2) %>% 
+  count(treatment)
+```
+
+```{r}
+# add new vars to dataset
+
+```
+
+```{r}
+# summary data for manuscript tables
+demo_funct_total <- function(var, data = demographics_clean){
+
+demographics_summary <- demographics_clean %>% 
+  rename(new_var = var) %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(area) %>% 
+  mutate(area_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = n()/total_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  select(new_var, n, percent)
+demographics_summary
+}
+
+demo_funct_total("gender_parent")
+
+demo_funct_area <- function(var, data = demographics_clean){
+
+demographics_summary <- demographics_clean %>% 
+  rename(new_var = var) %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(area) %>% 
+  mutate(area_n = n()) %>% 
+  group_by(area, new_var) %>% 
+  mutate(n = n(),
+         percent = n()/area_n) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  select(new_var, n, percent)
+demographics_summary
+}
+
+demo_funct_area("gender_parent")
 ```

### 2020-05-12 home_child_data_cleaning.Rmd
demographics/id filtering and merging

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 11 May 2020"
+date: "Started 19 Nov 2019, Updated 12 May 2020"
 output: pdf_document
@@
   rename_all(tolower) %>% 
-  filter(whatmissing != "all" | is.na(whatmissing)) %>%  # filter out if missing == "all"
+  #filter(whatmissing != "all" | is.na(whatmissing)) %>%  # filter out if missing == "all"
   select(homewinterid, gender, hispanic, race, income, education,
@@
 joined1 <- wmt_homewinter_initial %>% 
-  left_join(wmt_parent_demographics_initial, by = "home_winter_id") %>% 
-  left_join(wmt_home_initial, by = "home_id") %>% 
-  left_join(wmt_child_initial, by = "home_id") %>% 
-  mutate(area = "WMT") %>% 
-  filter(!is.na(enrolled))
+  full_join(wmt_parent_demographics_initial, by = "home_winter_id") %>% 
+  #mutate(home_id = if_else(home_winter_id == 275, "153", home_id)) %>% 
+  ### 1st big filter - remove those with no home_id
+  ### All other past this point should be considered "enrolled"
+  filter(!is.na(home_id)) %>% 
+  full_join(wmt_home_initial, by = "home_id") %>% 
+  full_join(wmt_child_initial, by = "home_id") %>% 
+  filter(!is.na(home_winter_id)) %>% 
+  mutate(area = "WMT") 
+  ### Once data is joined for all areas, filter for homes with treatment assignment
+  ### This will help further refine data for analysis and be a number in the flow chart
 
@@
   select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
-  filter(winterid != 3) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   mutate(home_winter_id = as.character(homewinterid),
@@
          intervention_time = interventiontime) %>% 
+  mutate(winter_id = as.numeric(winter_id),
+         winter_id = if_else(home_id == 62 | home_id == 45, winter_id + 1, winter_id),
+         winter_id = as.character(winter_id)) %>% 
+  filter(winter_id == 1 | winter_id ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
@@
 joined1 <- nn_homewinter_initial %>% 
-  left_join(nn_parent_demographics_initial, by = "home_winter_id") %>% 
-  left_join(nn_home_initial, by = "home_id") %>% 
-  left_join(nn_child_initial, by = "home_id") %>% 
-  mutate(area = "NN") %>% 
-  filter(!is.na(enrolled))
+  full_join(nn_parent_demographics_initial, by = "home_winter_id") %>% 
+  ### 1st big filter - remove those with no home_id
+  ### All other past this point should be considered "enrolled"
+  filter(!is.na(home_id)) %>% 
+  full_join(nn_home_initial, by = "home_id") %>% 
+  full_join(nn_child_initial, by = "home_id") %>% 
+  mutate(area = "NN") 
+  ### Once data is joined for all areas, filter for homes with treatment assignment
+  ### This will help further refine data for analysis and be a number in the flow chart
 
@@
 joined1 <- ak_homewinter_initial %>% 
-  left_join(ak_parent_demographics_initial, by = "home_winter_id") %>% 
-  left_join(ak_home_initial, by = "home_id") %>% 
-  left_join(ak_child_initial, by = "home_id") %>% 
-  mutate(area = "AK") %>% 
-  filter(!is.na(enrolled))
+  full_join(ak_parent_demographics_initial, by = "home_winter_id") %>% 
+  ### 1st big filter - remove those with no home_id
+  ### All other past this point should be considered "enrolled"
+  filter(!is.na(home_id)) %>% 
+  full_join(ak_home_initial, by = "home_id") %>% 
+  full_join(ak_child_initial, by = "home_id") %>% 
+  mutate(area = "AK") 
+  ### Once data is joined for all areas, filter for homes with treatment assignment
+  ### This will help further refine data for analysis and be a number in the flow chart
 
@@
 
+# Count total observations in dataset for homes and kids
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+ak_demographics <- read_rds(paste0(file_path, "Output/ak_demographic_data.rds"))
+nn_demographics <- read_rds(paste0(file_path, "Output/nn_demographic_data.rds"))
+wmt_demographics <- read_rds(paste0(file_path, "Output/wmt_demographic_data.rds"))
+
+kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics) %>% 
+  mutate(home_id = as.numeric(home_id)) %>% 
+  filter(home_id < 1000) %>%
+  mutate(home_id = as.character(home_id)) %>% 
+  # distinct(area, home_id)  # distinct homes in database = 360
+  filter(!is.na(child_id_num)) %>% 
+  # distinct(area, child_id_num, .keep_all = TRUE) %>%  # distinct kids in database = 512
+  # distinct(area, home_id, .keep_all = TRUE) %>% # distinct homes with kids in databse = 340
+  ### WMA222 and WMA 233 had a treatment but no kids - not counting in flow chart
+  filter(!is.na(treatment)) %>%  
+  distinct(area, child_id_num, .keep_all = TRUE) %>%   # distinct kids with treatment = 512
+  distinct(area, home_id, .keep_all = TRUE)  # distinct homes with kids with treatment = 340
+```
+
 # Combine linked id data for WMT, NN, AK
@@
 
-kids_ids <- rbind(ak_ids, nn_ids, wmt_ids)
+kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
+  mutate(home_id = as.numeric(home_id)) %>% 
+  filter(home_id < 1000) %>%  
+  mutate(home_id = as.character(home_id)) %>% 
+  # distinct(area, home_id)  # enrolled homes in database = 360
+  filter(!is.na(treatment)) %>% 
+  # distinct(area, home_id)  # randomized homes with treatment assignment = 342
+  ### WMA222 and WMA 233 had a treatment but no kids - not counting in flow chart
+  filter(home != "WMA222" & home != "WMA233") %>% 
+  #distinct(area, home_id)  # randomized homes with treatment assignment = 340
+  ungroup() %>% 
+  arrange(area, home)
 
-#write_rds(kids_ids, paste0(file_path, "Output/kids_linked_ids.rds")
-
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds")) %>% 
-  mutate(winter_id = as.numeric(winter_id),
-         winter_id = if_else(home == "CH218", winter_id + 1, winter_id)) %>% 
-  filter(winter_id == 1 | winter_id == 2)
-
-#write_rds(kids_linked_ids, paste0(file_path, "Output/kids_linked_ids.rds"))
+write_rds(kids_ids, paste0(file_path, "Output/kids_linked_ids.rds"))
 ```
@@
 
-kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics)
-
-# write_rds(kids_demographics, paste0(file_path, "Output/kids_combined_demographics.rds"))
-
-kids_demographics <- read_rds(paste0(file_path, "Output/kids_combined_demographics.rds"))
-```
-
-# Clean demographics data
-```{r}
-file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
-kids_demographics <- read_rds(paste0(file_path, "Output/kids_combined_demographics.rds"))
-
-demographics_clean <- kids_demographics %>% 
+kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics) %>% 
+  mutate(home_id = as.numeric(home_id)) %>% 
+  filter(home_id < 1000) %>%
+  mutate(home_id = as.character(home_id)) %>% 
+  filter(!is.na(child_id_num)) %>% 
+  filter(!is.na(treatment)) %>% 
   separate(intervention_time, c("trash", "intervention_time"), sep = " ") %>% 
@@
   select(-trash) %>% 
-  mutate(home_winter_id = as.numeric(home_winter_id),
-         home_winter_id = if_else(home == "CH218" & winter_id == 1, 77, home_winter_id),
-         home_winter_id = if_else(home == "CH218" & winter_id == 2, 111, home_winter_id),
-         home_winter_id = as.factor(home_winter_id)) %>% 
-  filter(winter_id == 1 | winter_id == 2)
+  ungroup() %>% 
+  arrange(area, home, child_id_char)
 
-# write_rds(demographics_clean, paste0(file_path, "Output/demographics_clean.rds"))
+write_rds(kids_demographics, paste0(file_path, "Output/kids_combined_demographics.rds"))
 ```

### 2020-05-13 child_health_data_cleaning.Rmd
data merging updates start on manuscript tables

@@
 
-# write_rds(kids_at_home_log, paste0(file_path, "Output/kids_at_home_log.rds"))
+write_rds(kids_at_home_log, paste0(file_path, "Output/kids_at_home_log.rds"))
 ```
@@
 
-#write_rds(at_home_pm_merged, paste0(file_path, "Output/at_home_pm_merged.rds"))
-#write_csv(at_home_pm_merged, paste0(file_path, "Output/at_home_pm_merged.csv"))
-
-at_home_pm_merged <- 
-  read_rds(paste0(file_path, "Output/at_home_pm_merged.rds"))
+write_rds(at_home_pm_merged, paste0(file_path, "Output/at_home_pm_merged.rds"))
+write_csv(at_home_pm_merged, paste0(file_path, "Output/at_home_pm_merged.csv"))
 ```
@@
 
-# write_rds(kids_med_visit, paste0(file_path, "Output/kids_med_visit.rds"))
+write_rds(kids_med_visit, paste0(file_path, "Output/kids_med_visit.rds"))
+```
 
+```{r}
 kids_med_visit <- read_rds(paste0(file_path, "Output/kids_med_visit.rds"))
@@
 
-# write_rds(kids_birth_infancy, paste0(file_path, "Output/kids_birth_infancy.rds"))
+write_rds(kids_birth_infancy, paste0(file_path, "Output/kids_birth_infancy.rds"))
+```
 
+```{r}
 kids_birth_infancy <- read_rds(paste0(file_path, "Output/kids_birth_infancy.rds"))
@@
 
-# write_rds(child_home_winter, paste0(file_path, "Output/child_home_winter.rds"))
-
-child_home_winter <- read_rds(paste0(file_path, "Output/child_home_winter.rds"))
-
+write_rds(child_home_winter, paste0(file_path, "Output/child_home_winter.rds"))
 ```
@@
 
-# write_rds(kids_csq, paste0(file_path, "Output/kids_csq.rds"))
+write_rds(kids_csq, paste0(file_path, "Output/kids_csq.rds"))
+```
 
+```{r}
 kids_csq <- read_rds(paste0(file_path, "Output/kids_csq.rds"))
@@
 
-# write_rds(kids_medical_records, paste0(file_path, "Output/kids_medical_records.rds"))
-
-kids_medical_records <- read_rds(paste0(file_path, "Output/kids_medical_records.rds"))
+write_rds(kids_medical_records, paste0(file_path, "Output/kids_medical_records.rds"))
 ```
@@
 
-# write_rds(kids_lrti, "Output/kids_lrti.rds")
-
-kids_lrti <- read_rds("Output/kids_lrti.rds")
-
+write_rds(kids_lrti, "Output/kids_lrti.rds")
 ```

### 2020-05-13 collinearity_exposure_analysis.Rmd
data merging updates start on manuscript tables

@@
 ```{r, eval=TRUE, include=TRUE}
-# Prep and merge data
-
-####### Format SUMs data to merge with analysis data below
-sums_temp_change <- sums_rolling_pm_data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home, sampling_day) %>% 
-  distinct(datetime_sums, .keep_all = TRUE) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
-  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
-         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
-  replace_with_na(replace = list(lead_temp = 999)) %>% 
-  mutate(temp_diff = lead_temp - temp_c_sums,
-         # change temp in the following line
-         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
-         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
-                              temp_diff_check),
-         heat_event = if_else(is.na(temp_diff_check), 1,
-                              heat_event)) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
-         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
-                                            breaks = c(0, 2, 50),
-                           labels = c("<2", "2+"),
-                                 right = FALSE)) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
-         time_diff:sums_events_sampling_day_cat) %>% 
-  arrange(area, home, sample_date) %>% 
-  group_by(home) %>% 
-  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
-  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup()
-
-####### Format ambient data to merge with analysis data below
-ambient_data <- exposure_analysis_data_long %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  #filter(area == "WMT") %>%
-  filter(!is.na(sampling_day)) %>% 
-  filter(!is.na(sample_date)) %>%
-  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
-         amb_pm_24hr:mean_temp_roll_4day) %>% 
-  group_by(home) %>% 
-  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
-         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup()
-
-####### Format home activity data to merge with analysis data below
-home_activity_data <- exposure_analysis_data_medium %>% 
-  filter(treatment == "Placebo") %>% 
-  #filter(winter_id == 1) %>% 
-  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
-  group_by(area, home, home_winter_id) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  group_by(area, home, home_winter_id) %>% 
-  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
-  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
-  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
-  ungroup() %>% 
-  distinct(area, home,  home_winter_id, .keep_all = TRUE)
-
 ####### Format full data to use for analysis
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  filter(area == "WMT") %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area, home) %>% 
-  distinct(home, winter_id, .keep_all = TRUE) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   ungroup() %>% 
+  # dichotomize variables by median value
   mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
@@
                                  right = FALSE)) %>% 
-  group_by(home) %>% 
-  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
-                                   lead(residents_smoke), residents_smoke)) %>% 
-  ungroup() %>% 
-  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
-  left_join(ambient_data, by = c("area", "home")) %>% 
-  select(-starts_with("home_act")) %>% 
-  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
-  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) %>% 
-  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
+  mutate(home_act_door_2level = cut(home_act_door_sum, breaks = c(0, 1, 7),
                                  labels = c("<1", "1+"),
                                  right = FALSE),
-         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
+         home_act_smoke_2level = cut(home_act_smoking_sum, breaks = c(0, 1, 7),
                                  labels = c("<1", "1+"),
                                  right = FALSE),
-         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
+         home_act_windows_2level = cut(home_act_windows_sum, breaks = c(0, 1, 7),
                                  labels = c("<1", "1+"),
                                  right = FALSE),
-         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
+         home_act_sweep_2level = cut(home_act_sweep_sum, breaks = c(0, 3, 7),
                                  labels = c("<3", "3+"),

### 2020-05-13 demographics_summary.Rmd
data merging updates start on manuscript tables

@@
 # summary data for manuscript tables
-demo_funct_total <- function(var, data = demographics_clean){
+demo_funct_total <- function(var, data = demographics_initial){
 
-demographics_summary <- demographics_clean %>% 
+demographics_summary <- data %>% 
   rename(new_var = var) %>% 
@@
 demo_funct_total("gender_parent")
+demo_funct_total("hispanic_parent")
+demo_funct_total("race_parent")
 
-demo_funct_area <- function(var, data = demographics_clean){
+demo_funct_area <- function(var, data = demographics_initial){
 
-demographics_summary <- demographics_clean %>% 
+demographics_summary <- data %>% 
   rename(new_var = var) %>% 
@@
 demo_funct_area("gender_parent")
+demo_funct_area("hispanic_parent")
+demo_funct_area("race_parent")
 ```

### 2020-05-13 exposure_covariate_associations.Rmd
data merging updates start on manuscript tables

--- 2020-05-01 exposure_covariate_associations.Rmd
@@
 ```{r, eval=TRUE, include=TRUE}
-####### Format SUMs data to merge with analysis data below
-sums_temp_change <- sums_rolling_pm_data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home, sampling_day) %>% 
-  distinct(datetime_sums, .keep_all = TRUE) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
-  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
-         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
-  replace_with_na(replace = list(lead_temp = 999)) %>% 
-  mutate(temp_diff = lead_temp - temp_c_sums,
-         # change temp in the following line
-         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
-         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
-                              temp_diff_check),
-         heat_event = if_else(is.na(temp_diff_check), 1,
-                              heat_event)) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
-         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
-                                            breaks = c(0, 2, 50),
-                           labels = c("<2", "2+"),
-                                 right = FALSE)) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
-         time_diff:sums_events_sampling_day_cat) %>% 
-  arrange(area, home, sample_date) %>% 
-  group_by(home) %>% 
-  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
-  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup()
-
-####### Format ambient data to merge with analysis data below
-ambient_data <- exposure_analysis_data_long %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  #filter(area == "AK") %>% 
-  filter(!is.na(sampling_day)) %>% 
-  filter(!is.na(sample_date)) %>%
-  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
-         amb_pm_24hr:mean_temp_roll_4day) %>% 
-  group_by(home) %>% 
-  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
-         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup()
-
-####### Format home activity data to merge with analysis data below
-home_activity_data <- exposure_analysis_data_medium %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  group_by(area, home) %>% 
-  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
-  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
-  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
-  ungroup() %>% 
-  distinct(area, home, .keep_all = TRUE) %>% 
-  select(home, area, sampling_day, starts_with("home_act"))
-
 ####### Format full data to use for analysis
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area, home) %>% 
@@
   ungroup() %>% 
+  # dichotomize variables by median value
   mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
@@
                                  right = FALSE)) %>% 
-  left_join(sums_temp_change, by = c("area", "home")) %>% 
-  left_join(ambient_data, by = c("area", "home")) %>% 
-  select(-starts_with("home_act")) %>% 
-  left_join(home_activity_data, by = c("area", "home")) %>% 
-  mutate(home_act_door_2level = cut(home_act_door, breaks = c(0, 1, 7),
+  mutate(home_act_door_2level = cut(home_act_door_sum, breaks = c(0, 1, 7),
                                  labels = c("<1", "1+"),
                                  right = FALSE),
-         home_act_smoke_2level = cut(home_act_smoking, breaks = c(0, 1, 7),
+         home_act_smoke_2level = cut(home_act_smoking_sum, breaks = c(0, 1, 7),
                                  labels = c("<1", "1+"),
                                  right = FALSE),
-         home_act_windows_2level = cut(home_act_windows, breaks = c(0, 1, 7),
+         home_act_windows_2level = cut(home_act_windows_sum, breaks = c(0, 1, 7),
                                  labels = c("<1", "1+"),
                                  right = FALSE),
-         home_act_sweep_2level = cut(home_act_sweep, breaks = c(0, 3, 7),
+         home_act_sweep_2level = cut(home_act_sweep_sum, breaks = c(0, 3, 7),
                                  labels = c("<3", "3+"),
@@
    pack_rows("Ambient data", 54, 55) %>%
-   pack_rows("Home activities - Sum of all sampling days", 56, 69) %>%
-   pack_rows("Stove grade", 70, 71) %>%
+   pack_rows("Home activities - Sum of all sampling days", 56, 70) %>%
+   pack_rows("Stove grade", 71, 72) %>%
    footnote(c("PM2.5 = fine particulate matter", 

### 2020-05-13 exposure_models.Rmd
data merging updates start on manuscript tables

@@
 
-**Note**
-
-Add some of the variables created below to the full dataset
-
 ```{r}
-# Prep and merge data
-
-####### Format SUMs data to merge with analysis data below
-sums_temp_change <- sums_rolling_pm_data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area, home, sampling_day) %>% 
-  distinct(datetime_sums, .keep_all = TRUE) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
-  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
-         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
-  replace_with_na(replace = list(lead_temp = 999)) %>% 
-  mutate(temp_diff = lead_temp - temp_c_sums,
-         # change temp in the following line
-         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
-         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
-                              temp_diff_check),
-         heat_event = if_else(is.na(temp_diff_check), 1,
-                              heat_event)) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
-         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
-                                            breaks = c(0, 2, 50),
-                           labels = c("<2", "2+"),
-                                 right = FALSE)) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
-         time_diff:sums_events_sampling_day_cat) %>% 
-  arrange(area, home, sample_date) %>% 
-  group_by(home) %>% 
-  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
-  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup()
-
-####### Format ambient data to merge with analysis data below
-ambient_data <- exposure_analysis_data_long %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  #filter(area == "WMT") %>%
-  filter(!is.na(sampling_day)) %>% 
-  filter(!is.na(sample_date)) %>%
-  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
-         amb_pm_24hr:mean_temp_roll_4day) %>% 
-  group_by(home) %>% 
-  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
-         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  ungroup()
-
-####### Format home activity data to merge with analysis data below
-home_activity_data <- exposure_analysis_data_medium %>% 
-  filter(treatment == "Placebo") %>% 
-  #filter(winter_id == 1) %>% 
-  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
-  group_by(area, home, home_winter_id) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  group_by(area, home, home_winter_id) %>% 
-  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
-  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
-  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
-  ungroup() %>% 
-  distinct(area, home,  home_winter_id, .keep_all = TRUE)
-
 ####### Format full data to use for analysis
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  filter(area == "WMT") %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
   group_by(area, home) %>% 
-  distinct(home, winter_id, .keep_all = TRUE) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
   ungroup() %>% 
-  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 12, 100),
-                                 labels = c("<12", "12+"),
+  # dichotomize variables by median value
+  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
+                                 labels = c("<10.5", "10.5+"),
                                  right = FALSE),

### 2020-05-13 home_char_act_cleaning.Rmd
data merging updates start on manuscript tables

@@
 ```{r}
-kids_home_char <- read_rds("Output/kids_home_char.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_home_char <- read_rds(paste0(file_path, "Output/kids_home_char.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 
@@
 
-# write_rds(home_char_clean, "Output/home_char_clean.rds")
+write_rds(home_char_clean, paste0(file_path, "Output/home_char_clean.rds"))
 ```
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 kids_home_act <- read_rds(paste0(file_path, "Output/kids_home_act.rds"))
@@
 
-# write_rds(home_act_clean, paste0(file_path, "Output/home_act_clean.rds"))
+write_rds(home_act_clean, paste0(file_path, "Output/home_act_clean.rds"))
 ```

### 2020-05-13 indoor_temp_rh_cleaning.Rmd
data merging updates start on manuscript tables

@@
 ```{r}
-kids_indoor_temp_rh <- read_rds("Output/kids_indoor_temp_rh.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_indoor_temp_rh <- read_rds(paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 
@@
 
-# write_rds(indoor_temp_rh_clean, "Output/indoor_temp_rh_clean.rds")
+write_rds(indoor_temp_rh_clean, paste0(file_path, "Output/indoor_temp_rh_clean.rds"))
 ```

### 2020-05-13 manuscript_tables_figures.Rmd
data merging updates start on manuscript tables

---
title: 'KidsAIR: manuscript tables/figures'
author: "Ethan Walker"
date: "Started 13 May 2020, Updated 13 May 2020"
output:
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE,
                      fig.height = 6, fig.width = 8)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# Load individual datasets

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds"))

exposure_analysis_data_long <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))

exposure_analysis_data_short <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
```

```{r}
# number of placebo households, total and by area
placebo_homes <- health_exposure_data_sampling_day %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  distinct(home, .keep_all = TRUE) %>%
  count(area)
placebo_homes

# number of placebo households, total and by area
treatment_homes <- health_exposure_data_sampling_day %>% 
  filter(winter_id == 1) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  count(treatment)
treatment_homes

# number of placebo households, total and by area
treatment_kids <- health_exposure_data_sampling_day %>% 
  filter(winter_id == 2) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  filter(!is.na(child_id_num)) %>% 
  count(treatment)
treatment_kids
```

```{r}
# add new vars to dataset

```

```{r}
# summary data for manuscript tables
demo_funct_total <- function(var, data = health_exposure_data_sampling_day){

demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  arrange(new_var) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  select(new_var, n, percent)
demographics_summary
}

demo_funct_area <- function(var, data = health_exposure_data_sampling_day){

demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  arrange(new_var) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(area, new_var) %>% 
  mutate(n = n(),
         percent = n()/area_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  select(new_var, n, percent) %>% 
  arrange(area, new_var)
demographics_summary
}

demo_funct_total("gender_parent")
demo_funct_area("gender_parent")
demo_funct_total("hispanic_parent")
demo_funct_area("hispanic_parent")
demo_funct_total("race_parent")
demo_funct_area("race_parent")
demo_funct_total("education_3level")
demo_funct_area("education_3level")
demo_funct_total("income_3level")
demo_funct_area("income_3level")
```

### 2020-05-13 merge_datasets.Rmd
data merging updates start on manuscript tables

@@
 library(zoo)
-library(lme4)
-library(lmerTest)
 ```
@@
                                     "C" = "C",
-                                    "D or F" = c("D", "F"))) %>% 
+                                    "F" = c("D", "F"))) %>% 
   ungroup() %>% 
@@
   read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+exposure_analysis_data_long <- 
+  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds")) 
+sums_rolling_pm_data <- 
+  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+
+####### Format SUMs data to merge with analysis data below
+sums_temp_change <- sums_rolling_pm_data %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  group_by(area, home, sampling_day) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(time_diff = (lead(datetime_sums) - datetime_sums)) %>% 
+  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
+         lead_temp = if_else(time_diff > 25, 999, lead_temp)) %>% 
+  replace_with_na(replace = list(lead_temp = 999)) %>% 
+  mutate(temp_diff = lead_temp - temp_c_sums,
+         # change temp in the following line
+         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
+         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
+                              temp_diff_check),
+         heat_event = if_else(is.na(temp_diff_check), 1,
+                              heat_event)) %>% 
+  group_by(area, home, sampling_day) %>% 
+  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
+         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
+                                            breaks = c(0, 2, 50),
+                           labels = c("<2", "2+"),
+                                 right = FALSE)) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
+         time_diff:sums_events_sampling_day_cat) %>% 
+  arrange(area, home, sample_date) %>% 
+  group_by(home) %>% 
+  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
+  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup()
+
+####### Format ambient data to merge with analysis data below
+ambient_data <- exposure_analysis_data_long %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  #filter(area == "WMT") %>%
+  filter(!is.na(sampling_day)) %>% 
+  filter(!is.na(sample_date)) %>%
+  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
+         amb_pm_24hr:mean_temp_roll_4day) %>% 
+  group_by(home) %>% 
+  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
+         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  replace_with_na(replace = list(mean_temp_sampling_period_5 = "NaN")) %>% 
+  replace_with_na(replace = list(amb_pm_sampling_period_5 = "NaN")) %>% 
+  select(area, home, mean_temp_sampling_period_5, amb_pm_sampling_period_5)
 
@@
   ungroup() %>% 
+  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
+  left_join(ambient_data, by = c("area", "home")) %>% 
   arrange(area, home, winter_id)

### 2020-05-13 merge_health_datasets.Rmd
health data cleaning
data merging updates start on manuscript tables

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 14 April 2020"
+date: "Started 27 March 2020, Updated 22 April 2020"
 output: pdf_document
@@
   read_rds(paste0(output_path, "Output/kids_lrti.rds")) %>% 
-  select(-home_id)
+  select(-home_id) %>% 
+  mutate(winter_id = as.character(winter_id))
 kids_medical_records <- 
@@
   mutate(winter_id = as.character(winter_id)) %>% 
-  left_join(exposure_analysis_data_medium,
+  right_join(exposure_analysis_data_medium,
             by = c("area", "home", "home_winter_id", "winter_id", "treatment",
@@
 
-#write_rds(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
-#write_csv(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.csv"))
+write_rds(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
+write_csv(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.csv"))
+```
 
+```{r}
 health_exposure_data_sampling_day <- 

### 2020-05-13 moisture_data_cleaning.Rmd
data merging updates start on manuscript tables

@@
 ```{r}
-kids_moisture <- read_rds("Output/kids_moisture.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_moisture <- read_rds(paste0(file_path, "Output/kids_moisture.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 
@@
 
-# write_rds(moisture_clean, "Output/moisture_clean.rds")
+write_rds(moisture_clean, paste0(file_path, "Output/moisture_clean.rds"))
 ```

### 2020-05-13 stove_grade_usage_cleaning.Rmd
data merging updates start on manuscript tables

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 ##### WMT #####
 # Load data
-stove_grades_initial_wmt <- read_csv("Input/WMT/wmt_stove_grades_20200113.csv")
+stove_grades_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/wmt_stove_grades_20200113.csv"))
 
@@
 # Load data
-stove_grades_initial_nn <- read_xlsx("Input/NN/nn_stove_grades_20200113.xlsx")
+stove_grades_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/nn_stove_grades_20200113.xlsx"))
 
@@
 # Load data
-stove_grades_initial_ak <- read_xlsx("Input/AK/ak_stove_grades_20200113.xlsx")
+stove_grades_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/ak_stove_grades_20200113.xlsx"))
 
@@
 # Save as RDS
-# write_rds(ak_stove_grades, "Output/ak_stove_grades.rds")
+write_rds(ak_stove_grades, paste0(file_path, "Output/ak_stove_grades.rds"))
 ```
@@
 ```{r}
-kids_stove_grades <- read_rds("Output/kids_stove_grades.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+kids_stove_grades <- read_rds(paste0(file_path, "Output/kids_stove_grades.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
@@
   filter(home_id != 0) %>% 
+  filter(stove_grade != "D") %>% 
   rename(stove_grade_comments = comments) %>% 
   select(-home_id) %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  arrange(area, home, winter_id)
 
-# write_rds(stove_grades_clean, "Output/stove_grades_clean.rds")
+write_rds(stove_grades_clean, paste0(file_path, "Output/stove_grades_clean.rds"))
 
@@
 ```{r}
-kids_stove_use <- read_rds("Output/kids_stove_use.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+kids_stove_use <- read_rds(paste0(file_path, "Output/kids_stove_use.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
@@
 
-# write_rds(stove_use_clean, "Output/stove_use_clean.rds")
+write_rds(stove_use_clean, paste0(file_path, "Output/stove_use_clean.rds"))
 ```

### 2020-05-13 sums_data_cleaning.Rmd
data merging updates start on manuscript tables

@@
 ```{r}
-all_sums_filtered <- read_rds("Output/all_sums_filtered.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+all_sums_filtered <- read_rds(paste0(file_path, "Output/all_sums_filtered.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds")) %>% 
   select(home_winter_id, area, treatment)
@@
 
-# write_rds(sums_clean, "Output/sums_clean.rds")
+write_rds(sums_clean, paste0(file_path, "Output/sums_clean.rds"))
 
@@
 
-sums_clean <- read_rds("Output/sums_clean.rds")
+sums_clean <- read_rds(paste0(file_path, "Output/sums_clean.rds"))
 
@@
 
-# write_rds(sums_new, "Output/sums_clean.rds")
+write_rds(sums_new, paste0(file_path, "Output/sums_clean.rds"))
 ```
@@
 ```{r}
-sums_clean <- read_rds("Output/sums_clean.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+sums_clean <- read_rds(paste0(file_path, "Output/sums_clean.rds"))
 
@@
 # Load SUMs data
-sums_clean <- read_rds("Output/sums_clean.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+sums_clean <- read_rds(paste0(file_path, "Output/sums_clean.rds"))
 

### 2020-05-14 demographics_summary.Rmd
summary stat updates

@@
 # summary data for manuscript tables
-demo_funct_total <- function(var, data = demographics_initial){
+char_funct_total <- function(var, data = demographics_initial){
 
@@
 
-demo_funct_total("gender_parent")
-demo_funct_total("hispanic_parent")
-demo_funct_total("race_parent")
+char_funct_total("gender_parent")
+char_funct_total("hispanic_parent")
+char_funct_total("race_parent")
 
-demo_funct_area <- function(var, data = demographics_initial){
+char_funct_area <- function(var, data = demographics_initial){
 
@@
 
-demo_funct_area("gender_parent")
-demo_funct_area("hispanic_parent")
-demo_funct_area("race_parent")
+char_funct_area("gender_parent")
+char_funct_area("hispanic_parent")
+char_funct_area("race_parent")
 ```

### 2020-05-14 manuscript_tables_figures.Rmd
summary stat updates

@@
 
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+
 kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
@@
   read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))
-
-health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 ```
@@
 # add new vars to dataset
-
+checkdata <- health_exposure_data_sampling_day %>% 
+  select(area, home, residents_under_five)
 ```
@@
 # summary data for manuscript tables
-demo_funct_total <- function(var, data = health_exposure_data_sampling_day){
+char_funct_total <- function(var, data = health_exposure_data_sampling_day){
 
@@
 
-demo_funct_area <- function(var, data = health_exposure_data_sampling_day){
+char_funct_area <- function(var, data = health_exposure_data_sampling_day){
 
@@
 
-demo_funct_total("gender_parent")
-demo_funct_area("gender_parent")
-demo_funct_total("hispanic_parent")
-demo_funct_area("hispanic_parent")
-demo_funct_total("race_parent")
-demo_funct_area("race_parent")
-demo_funct_total("education_3level")
-demo_funct_area("education_3level")
-demo_funct_total("income_3level")
-demo_funct_area("income_3level")
+char_funct_total("gender_parent")
+char_funct_area("gender_parent")
+char_funct_total("hispanic_parent")
+char_funct_area("hispanic_parent")
+char_funct_total("race_parent")
+char_funct_area("race_parent")
+char_funct_total("education_3level")
+char_funct_area("education_3level")
+char_funct_total("income_3level")
+char_funct_area("income_3level")
+char_funct_total("home_floors_2level")
+char_funct_area("home_floors_2level")
+char_funct_total("home_year_built_2level")
+char_funct_area("home_year_built_2level")
+char_funct_total("home_pets_2level")
+char_funct_area("home_pets_2level")
+char_funct_total("stove_age_3level")
+char_funct_area("stove_age_3level")
+char_funct_total("chimney_clean_3level")
+char_funct_area("chimney_clean_3level")
+char_funct_total("wood_collect_method_2level")
+char_funct_area("wood_collect_method_2level")
+char_funct_total("wood_collect_2level")
+char_funct_area("wood_collect_2level")
+char_funct_total("stove_grade_3level")
+char_funct_area("stove_grade_3level")
 ```
+
+```{r}
+# summary data for manuscript tables
+num_funct_total <- function(var, data = health_exposure_data_sampling_day){
+
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  arrange(new_var) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  select(area, home, new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_area <- function(var, data = health_exposure_data_sampling_day){
+
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  arrange(new_var) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  select(area, home, new_var) %>% 
+  group_by(area) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_total("residents_under_five")
+num_funct_area("residents_under_five")
+num_funct_total("home_bedrooms")
+num_funct_area("home_bedrooms")
+```

### 2020-05-14 stove_grade_usage_summary.Rmd
summary stat updates

@@
 # Load data
-stove_grades_clean <- read_rds("Output/stove_grades_clean.rds") 
-stove_use_clean <- read_rds("Output/stove_use_clean.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") 
-demographics_clean <- read_rds("Output/demographics_clean.rds")
-#pm_clean <- read_rds("Output/pm_clean.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+stove_grades_clean <- read_rds(paste0(file_path, "Output/stove_grades_clean.rds")) 
+stove_use_clean <- read_rds(paste0(file_path, "Output/stove_use_clean.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+demographics_clean <- read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+#pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
 ```

### 2020-05-14 child_health_data_summary.Rmd
update analysis, results, manuscript tables

@@
 # Load data
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-demographics_clean <- read_rds("Output/demographics_clean.rds")
-pm_clean <- read_rds("Output/pm_clean.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-kids_med_visit <- read_rds("Output/kids_med_visit.rds") 
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+demographics_clean <- read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
+
+kids_med_visit <- read_rds(paste0(file_path, "Output/kids_med_visit.rds")) 
 ```

### 2020-05-14 exposure_covariate_associations.Rmd
update analysis, results, manuscript tables

@@
   analysis_data <- dataset %>% 
-    filter(winter_id == 1 & treatment == "Placebo") %>% 
+    filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
     group_by(area, home) %>% 

### 2020-05-14 exposure_models.Rmd
update analysis, results, manuscript tables

@@
 missing_summary <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(stove_age_3level), "missing yes", "missing no")) %>%
+  mutate(missing_var = if_else(is.na(stove_grade_3level), "missing yes", "missing no")) %>%
   group_by(area, missing_var) %>% 
@@
 missing_test <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(stove_age_3level), "missing yes", "missing no")) %>% 
+  mutate(missing_var = if_else(is.na(stove_grade_3level), "missing yes", "missing no")) %>% 
   filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))
@@
   summarize(n = n(),
-            mean_var = mean(pm_mean_sampling_period, na.rm = TRUE), 
-            sd_var = sd(pm_mean_sampling_period, na.rm = TRUE),
-            median_var = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
+            mean_var = mean(mean_temp_sampling_period_5, na.rm = TRUE), 
+            sd_var = sd(mean_temp_sampling_period_5, na.rm = TRUE),
+            median_var = median(mean_temp_sampling_period_5, na.rm = TRUE)) %>% 
   group_by(area) %>% 
@@
 
-write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
+write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_14may2020.rds"))
 
-imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_7may2020.rds"))
+imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_14may2020.rds"))
 
@@
 
-write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
+write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_14may2020.rds"))
 
-exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_7may2020.rds"))
+exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_14may2020.rds"))
 
-summary(test_model_results)
+summary(model_ave)
 

### 2020-05-14 home_char_act_summary.Rmd
update analysis, results, manuscript tables

@@
 # Load dataset
-home_char_clean <- read_rds("Output/home_char_clean.rds") 
-home_act_clean <- read_rds("Output/home_act_clean.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+home_char_clean <- read_rds(paste0(file_path, "Output/home_char_clean.rds")) 
+home_act_clean <- read_rds(paste0(file_path, "Output/home_act_clean.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 #pm_clean <- read_rds("Output/pm_clean.rds")

### 2020-05-14 manuscript_tables_figures.Rmd
update analysis, results, manuscript tables

@@
 author: "Ethan Walker"
-date: "Started 13 May 2020, Updated 13 May 2020"
+date: "Started 13 May 2020, Updated 14 May 2020"
 output:
@@
 num_funct_area("home_bedrooms")
+num_funct_total("pm_mean_sampling_period")
+num_funct_area("pm_mean_sampling_period")
+num_funct_total("pm_sample_interval")
+num_funct_area("pm_sample_interval")
+num_funct_total("moisture_closest")
+num_funct_area("moisture_closest")
+num_funct_total("temp_indoor_max")
+num_funct_area("temp_indoor_max")
+num_funct_total("rh_indoor_max")
+num_funct_area("rh_indoor_max")
+num_funct_total("mean_temp")
+num_funct_area("mean_temp")
+num_funct_total("amb_pm_24hr")
+num_funct_area("amb_pm_24hr")
 ```
+
+```{r}
+other_vars_total <- exposure_analysis_data_short %>% 
+  rename(new_var = home_act_sweep_sum) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  arrange(new_var) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  select(area, home, new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+other_vars_total
+
+other_vars_area <- exposure_analysis_data_short %>% 
+  rename(new_var = home_act_sweep_sum) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  filter(winter_id == 1) %>% 
+  filter(treatment == "Placebo") %>% 
+  arrange(new_var) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  filter(!is.na(new_var)) %>% 
+  select(area, home, new_var) %>% 
+  group_by(area) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+other_vars_area
+```

### 2020-05-14 merge_datasets.Rmd
update analysis, results, manuscript tables

@@
          sample_date, pm_mean_daily, pm_daily_obs, pm_spike_1000,
-         pm_spike_2000, pm_spike_5000) %>% 
+         pm_spike_2000, pm_spike_5000, pm_sample_interval) %>% 
   group_by(area, home) %>% 
@@
 analysis_data2 <- demographics_clean %>% 
+  mutate(residents_under_five = if_else(residents_under_five == 0 & age_child_nov != 4, "NA", residents_under_five),
+         residents_under_five = if_else(residents_under_five == 0 & child_id_char == "WMA673B", "NA", residents_under_five)) %>% 
+  replace_with_na(replace = list(total_residents = "0")) %>% 
+  replace_with_na(replace = list(total_residents = "1")) %>% 
+  replace_with_na(replace = list(residents_under_five = "NA")) %>% 
   full_join(analysis_data1, by = c("area", "home", "home_winter_id", 
@@
 
-#write_rds(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.rds"))
-#write_csv(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.csv"))
+write_rds(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.rds"))
+write_csv(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.csv"))
 
@@
          total_residents_2level = fct_collapse(total_residents,
-                                    "<5" = c("1", "2", "3", "4"),
+                                    "<5" = c("2", "3", "4"),
                                     "5+" = c("5", "6", "7", "8", "9", "10",
@@
 
-#write_rds(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
-#write_csv(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.csv"))
+write_rds(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
+write_csv(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.csv"))
 ```
@@
 ```{r}
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 exposure_analysis_data_long_new <- 
@@
 
-#write_rds(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
-#write_csv(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.csv"))
+write_rds(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+write_csv(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.csv"))
 ```
@@
 ```{r}
+output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 exposure_analysis_data_medium <- 
@@
 
-#write_rds(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.rds"))
-#write_csv(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.csv"))
+write_rds(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.rds"))
+write_csv(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.csv"))
 ```
@@
 
-#write_rds(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.rds"))
-#write_csv(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.csv"))
+write_rds(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+write_csv(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.csv"))
 ```

### 2020-05-14 merge_health_datasets.Rmd
update analysis, results, manuscript tables

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 22 April 2020"
+date: "Started 27 March 2020, Updated 14 May 2020"
 output: pdf_document
@@
 
-#write_rds(health_merged6, paste0(output_path, "Output/health_data_merged.rds"))
-#write_csv(health_merged6, paste0(output_path, "Output/health_data_merged.csv"))
+write_rds(health_merged6, paste0(output_path, "Output/health_data_merged.rds"))
+write_csv(health_merged6, paste0(output_path, "Output/health_data_merged.csv"))
 ```

### 2020-05-15 pm_summary.Rmd
exposure analysis update
pm summary updates

@@
 ```
+
+\pagebreak
 
@@
 
+```{r, fig.height=, fig.width=8}
+boxplots_kids_pm <- pm_clean %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  ggplot(aes(area, pm)) +
+    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
+               fatten = 1, outlier.size = 1.5, width = 0.35) +
+    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
+    scale_y_log10() +
+    theme_minimal() +
+    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    theme(axis.text.x = element_text(size = 16, colour = "black",
+                                      margin = margin(t = 0, r = 0, b = 30, l = 0)),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank())
+boxplots_kids_pm
+
+ggsave("boxplots_kids_pm.jpg", height = 7, width = 8)
+```
+
+\pagebreak
+
 ```{r, fig.height=6, fig.width=8}

### 2020-05-20 exposure_covariate_associations.Rmd
minor exposure analysis updates

@@
 
-  analysis_results <- lm(log(pm_mean_sampling_period) ~ analysis_covar, analysis_data)
+  analysis_results <<- lm(log(pm_mean_sampling_period) ~ analysis_covar, analysis_data)
 
@@
     filter(term != '(Intercept)') %>% 
-    mutate(Estimate = round(estimate, digits = 2),
+    mutate(Estimate = round((exp(estimate)-1)*100, digits = 2),
            'P-value' = round(p.value, digits = 2),
-           'Upper 95% CL' = round(conf.high, digits = 2),
-           'Lower 95% CL' = round(conf.low, digits = 2),
+           'Upper 95% CL' = round((exp(conf.high)-1)*100, digits = 2),
+           'Lower 95% CL' = round((exp(conf.low)-1)*100, digits = 2),
            Term = gsub("analysis_covar", paste0(label, ": "), term)) %>% 

### 2020-05-20 exposure_models.Rmd
minor exposure analysis updates

@@
 print(exposure_model)
+
+# Exponentiate results for manuscript tables
+exp_results <- as.data.frame(exposure_model$coefficients.s) %>% 
+  rownames_to_column() %>% 
+  mutate(estimate = round((exp(Estimate)-1)*100, digits = 2),
+         ci_low = round((exp(`Lower CI`)-1)*100, digits = 2),
+         ci_hi = round((exp(`Upper CI`)-1)*100, digits = 2)) %>% 
+  select(rowname, estimate, ci_low, ci_hi)
+
+exp_results
 ```

### 2020-05-21 exposure_models.Rmd
small exposure model/figure updates

@@
 
+```{r}
+# Look at R^2 for final selected model
+model_results <- lm(log_pm_mean_sampling_period ~
+                      income_3level +
+                      residents_smoke +
+                      home_year_built_2level +
+                      home_floors_2level + 
+                      home_pets_2level +
+                      chimney_clean_3level +
+                      home_fireplace +
+                      wood_collect_2level +
+                      sums_events_sampling_period +
+                      mean_temp_sampling_period_5,
+                      data = imputed_exposure_data$data)
+
+summary(model_results)
+```
+
 #####################

### 2020-05-21 manuscript_tables_figures.Rmd
small exposure model/figure updates

@@
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
+  #filter(treatment == "Placebo") %>% 
   distinct(home, .keep_all = TRUE) %>%
@@
 treatment_homes <- health_exposure_data_sampling_day %>% 
-  filter(winter_id == 1) %>% 
+  filter(winter_id == 2) %>% 
   distinct(home, .keep_all = TRUE) %>% 
+  group_by(area) %>% 
   count(treatment)
@@
   filter(!is.na(child_id_num)) %>% 
+  group_by(area) %>% 
   count(treatment)

### 2020-05-21 pm_summary.Rmd
small exposure model/figure updates

@@
 # Load data
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
-pm_clean <- read_rds(paste0(output_path, "Output/pm_clean.rds"))
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
+
+exposure_analysis_data_long <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
+
+exposure_analysis_data_medium <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
+
+exposure_analysis_data_short <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds"))
+
+sums_rolling_pm_data <- 
+  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))
+```
+
+```{r}
+# home that exceed PM guidelines
+pm_guidelines <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(pm_mean_sampling_period > 35) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  group_by(area) %>% 
+  count()
+pm_guidelines
 ```

### 2020-05-22 collinearity_exposure_analysis.Rmd
exposure analysis/model updates

@@
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
+  filter(home != "WMA300") %>% 
   group_by(area, home) %>% 
@@
          residents_smoke,
-         home_year_built_2level,
          home_floors_2level, 
+         home_sqft_2level,
          home_pets_2level,
@@
          chimney_clean_3level, 
-         home_fireplace,
          wood_collect_2level,
@@
          mean_temp_sampling_period_5,
-         amb_pm_sampling_period_5,
-         stove_grade_3level) %>% 
+         stove_grade_3level,
+         home_fireplace) %>% 
   #mutate(area = factor(area)) %>% 
@@
          residents_smoke,
-         home_floors_2level, 
+         home_floors_2level, home_sqft_2level,
          home_pets_2level,
          stove_age_3level, chimney_clean_3level, 
-         stove_cert, home_mold,
+         stove_cert, 
          wood_collect_2level,
@@
          home_act_smoke_2level, 
-         home_act_door_2level,
          stove_grade_3level,
-         race_parent_2level, home_damp, 
-         temp_max_5, home_act_gas, home_type_2level, 
-         home_windows_2level, home_fireplace
+         race_parent_2level, residents_under5_2level,
+         home_type_2level, 
+         home_fireplace
 

### 2020-05-22 dustrak_data_cleaning.Rmd
exposure analysis/model updates

@@
 # Run all of these lines together
-pm_time_plot_function(id = "501", location = "WMT")
+pm_time_plot_function(id = "259", location = "WMT")
 head(gaps, 25)
@@
 
+ggsave("wmt_hwid259_high_pm.jpeg")
+
 # filter data for a single home
 check_data <- kids_pm %>% 
-  filter(area == "WMT" & home_winter_id == "129") %>% 
+  filter(area == "WMT" & home_winter_id == "137") %>% 
   filter(pm < 0)

### 2020-05-22 exposure_covariate_associations.Rmd
exposure analysis/model updates

@@
   filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
   group_by(area, home) %>% 
@@
   analysis_results <<- lm(log(pm_mean_sampling_period) ~ analysis_covar, analysis_data)
+
+  print(table(model.frame(analysis_results)$analysis_covar))
+  #print(nobs(analysis_results))
 

### 2020-05-22 exposure_models.Rmd
exposure analysis/model updates

--- 2020-05-20 exposure_models.Rmd
@@
 selected_vars <- analysis_data %>% 
-  filter(area == "WMT") %>% 
+  #filter(area == "NN") %>% 
   select(pm_mean_sampling_period,
@@
          mean_temp_sampling_period_5,
-         amb_pm_sampling_period_5,
-         stove_grade_3level) %>% 
+         stove_grade_3level,
+         home_fireplace) %>% 
   mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>%  
@@
 # Check % missing data
-mean(is.na(selected_vars)) # 10%
+mean(is.na(selected_vars)) # 10.7% for WMT
 
@@
 missing_summary
+```
 
+```{r}
 # T-tests
 missing_test <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(stove_grade_3level), "missing yes", "missing no")) %>% 
+  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
   filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))
@@
   summarize(n = n(),
-            mean_var = mean(mean_temp_sampling_period_5, na.rm = TRUE), 
-            sd_var = sd(mean_temp_sampling_period_5, na.rm = TRUE),
-            median_var = median(mean_temp_sampling_period_5, na.rm = TRUE)) %>% 
+            mean_var = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            sd_var = sd(pm_mean_sampling_period, na.rm = TRUE),
+            median_var = median(pm_mean_sampling_period, na.rm = TRUE)) %>% 
   group_by(area) %>% 
@@
                      method = c("polr", "logreg", "logreg", "logreg", "logreg",
-                                "polr", "polr", "logreg", "logreg", "logreg",
-                                "pmm", "pmm", "pmm", "polr", "pmm"))
+                                "polr", "polr", "logreg", "logreg", "pmm",
+                                "pmm", "polr", "logreg", "pmm"))
 
-write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_14may2020.rds"))
+write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_22may2020.rds"))
 
-imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_14may2020.rds"))
+imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_22may2020.rds"))
 
@@
 
-write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_14may2020.rds"))
+write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_22may2020.rds"))
 
-exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_14may2020.rds"))
+exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_22may2020.rds"))
 
@@
 exp_results
+```
+
+```{r}
+# Look at R^2 for final selected model
+model_results <- lm(log_pm_mean_sampling_period ~
+                      residents_smoke +
+                      chimney_clean_3level +
+                      stove_grade_3level,
+                      data = imputed_exposure_data$data)
+
+summary(model_results)
 ```

### 2020-05-22 manuscript_tables_figures.Rmd
exposure analysis/model updates

@@
 author: "Ethan Walker"
-date: "Started 13 May 2020, Updated 14 May 2020"
+date: "Started 13 May 2020, Updated 22 May 2020"
 output:
@@
   filter(treatment == "Placebo") %>%
+  filter(home != "WMA300") %>% 
   arrange(new_var) %>% 
@@
   filter(treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
   arrange(new_var) %>% 
@@
   mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
   filter(winter_id == 1) %>% 
   filter(treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
   arrange(new_var) %>% 
@@
   mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
   filter(winter_id == 1) %>% 
   filter(treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
   arrange(new_var) %>% 
@@
 other_vars_total <- exposure_analysis_data_short %>% 
-  rename(new_var = home_act_sweep_sum) %>% 
+  rename(new_var = sums_events_sampling_period) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
@@
 other_vars_area <- exposure_analysis_data_short %>% 
-  rename(new_var = home_act_sweep_sum) %>% 
+  rename(new_var = sums_events_sampling_period) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 

### 2020-05-22 pm_summary.Rmd
exposure analysis/model updates

@@
 ```{r}
-# home that exceed PM guidelines
+# homes that exceed PM guidelines
 pm_guidelines <- exposure_analysis_data_short %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
   filter(pm_mean_sampling_period > 35) %>% 
@@
 pm_guidelines
+
+#Pm by different levels of covariates
+pm_check <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  group_by(chimney_clean_3level) %>% 
+  summarize("n" = n(),
+            "Mean PM" = round(mean(pm_mean_sampling_period, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm_mean_sampling_period, na.rm = TRUE)),
+            "Min PM" = round(min(pm_mean_sampling_period, na.rm = TRUE)), 
+            "25%" = round(quantile(pm_mean_sampling_period, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm_mean_sampling_period, na.rm = TRUE)),
+            "75%" = round(quantile(pm_mean_sampling_period, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm_mean_sampling_period, na.rm = TRUE)))
+pm_check
+
+# closer look at stove grade and PM
+pm_stove_grade <- exposure_analysis_data_short %>%
+  select(area, home, pm_mean_sampling_period, stove_grade_3level, 
+         treatment, winter_id, home_winter_id) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  mutate(stove_grade_2level = fct_collapse(stove_grade_3level,
+                                           "A" = "A",
+                                           "C or F" = c("C", "F"))) %>% 
+  group_by(stove_grade_2level) %>% 
+  summarize("n" = n(),
+            "Mean PM" = round(mean(pm_mean_sampling_period, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm_mean_sampling_period, na.rm = TRUE)),
+            "Min PM" = round(min(pm_mean_sampling_period, na.rm = TRUE)), 
+            "25%" = round(quantile(pm_mean_sampling_period, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm_mean_sampling_period, na.rm = TRUE)),
+            "75%" = round(quantile(pm_mean_sampling_period, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm_mean_sampling_period, na.rm = TRUE)))
+pm_stove_grade
+
+model_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(area == "WMT") %>% 
+  filter(home != "WMA300") %>% 
+  distinct(home, .keep_all = TRUE)
+
+check_model <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level,
+                  data = model_data)
+
+summary(check_model)
 ```

### 2020-05-28 child_health_data_cleaning.Rmd
Small plot and code annotation updates

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 22 April 2020"
+date: "Started 27 Feb 2020, Updated 28 May 2020"
 output: pdf_document 
@@
          at_home_day = as.factor(Day),
+         # rename columns and recode as factor
          at_home_7 = as.factor(AtHome1),
@@
 ```{r}
+# load data that was saved above
 kids_at_home_log <- 
@@
   group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
+  # recode factors so 1=home, 0=not home; add overnight hours
   mutate("00:00:00" = 1,
@@
          "23:00:00" = 1) %>% 
+  # change data from wide to long, with row for hour of each day
   pivot_longer(cols = c("00:00:00":"23:00:00"), names_to = "sample_time",
                         values_to = "at_home") %>% 
+  # fill in overnight hours with assumption that child was home at night
+  # match hour of day with corresponding location from the log
   mutate("00:00:00" = "home",
@@
   group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
+  # recode overnight hours with assumption that child was at location they were
+  # at 10pm and 7am
   mutate(at_home = if_else(sample_time == "06:00:00", lead(at_home), at_home),
@@
 
+# start with cleaned PM dataset
 at_home_pm_merged <- pm_clean %>% 
   group_by(area, home) %>% 
+  # average the PM for sampling period, day, and hour
   mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
@@
           desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
+  # keep one observation per hour of sampling, with PM averaged
   distinct(sample_time, .keep_all = TRUE) %>% 
@@
          pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
+  # join with at home log from above
   left_join(kids_at_home_log_new, 
@@
   group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
+  # average PM during time child was at home
   mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
@@
 wmt_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalvisit_20200227.xlsx")) %>% 
+  # recode values to NA
   mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
@@
   group_by(home_winter_id, child_id_num, med_visit_date) %>%
+  # recode temps of <90 to NA
   mutate(value = if_else(value < 90, -99, value)) %>% 

### 2020-05-28 pm_summary.Rmd
Small plot and code annotation updates

@@
   filter(area == "WMT") %>% 
-  filter(home != "WMA300") %>% 
+  filter(home == "WMA300") %>% 
   distinct(home, .keep_all = TRUE)
@@
   filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
   ggplot(aes(area, pm)) +
@@
 ```{r, fig.height=, fig.width=8}
-boxplots_kids_pm <- pm_clean %>% 
+boxplots_kids_pm <- exposure_analysis_data_short %>% 
   filter(winter_id == 1 & treatment == "Placebo") %>% 
-  ggplot(aes(area, pm)) +
+  filter(home != "WMA300") %>% 
+  ggplot(aes(area, pm_mean_sampling_period)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
     stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red") +
-    scale_y_log10() +
+    #scale_y_log10() +
     theme_minimal() +
@@
 ```
+
+```{r}
+# check PM spikes
+spikes_data <- exposure_analysis_data_medium %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
+  select(area, home, sampling_day, pm_mean_sampling_day, pm_mean_sampling_period,
+         pm_spike_1000, pm_spike_2000, pm_spike_5000,
+         income_3level, stove_grade_3level, chimney_clean_3level, home_floors_2level) %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  group_by(area, home) %>% 
+  distinct(area, home, sampling_day, .keep_all = TRUE) %>% 
+  mutate(pm_spike_1000 = if_else(pm_spike_1000 == "Yes", 1, 0),
+         pm_spike_2000 = if_else(pm_spike_2000 == "Yes", 1, 0),
+         pm_spike_5000 = if_else(pm_spike_5000 == "Yes", 1, 0),
+         spike_sum_1000 = sum(pm_spike_1000),
+         spike_sum_2000 = sum(pm_spike_2000),
+         spike_sum_5000 = sum(pm_spike_5000)) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  mutate(spike_sum_1000 = as.factor(spike_sum_1000),
+         spike_sum_1000 = fct_collapse(spike_sum_1000, "0" = "0",
+                                       "1" = "1",
+                                       "2" = "2",
+                                       "3+" = c("3", "5"))) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+spikes_summary <- spikes_data %>% 
+  mutate(spike_sum_1000 = as.factor(spike_sum_1000),
+         spike_sum_1000 = fct_collapse(spike_sum_1000, "0" = "0",
+                                       "1" = "1",
+                                       "2" = "2",
+                                       "3+" = c("3", "5"))) %>% 
+  group_by(spike_sum_1000) %>% 
+  summarize("N" = n(),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE), 
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+spikes_summary
+
+table(spikes_data$spike_sum_1000)
+table(spikes_data$spike_sum_1000, spikes_data$stove_grade_3level)
+```

### 2020-06-17 dustrak_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 14 Jan 2020"
+date: "Started 19 Nov 2019, Updated 16 June 2020"
 output: pdf_document
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 kids_pm <- 
-  read_rds(paste0(output_path, "Output/kids_pm_new4.rds")) 
+  read_rds(paste0(file_path, "Output/kids_pm_new4.rds")) 
 kids_linked_ids <- 
-  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
+  read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
 
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-write_rds(kids_pm_new5, paste0(output_path, "Output/pm_clean.rds"))
+write_rds(kids_pm_new5, paste0(file_path, "Output/pm_clean.rds"))
 ```

### 2020-06-17 home_child_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 12 May 2020"
+date: "Started 19 Nov 2019, Updated 16 June 2020"
 output: pdf_document
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > WMT > child.xlsx
-wmt_child_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_child_20191126.xlsx")) %>% 
+## FINAL DATA SETS > WMT > child
+wmt_child_initial <- read_csv(paste0(file_path, "Input/WMT/child.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > WMT > home.xlsx
-wmt_home_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_home_20191126.xlsx")) %>% 
+## FINAL DATA SETS > WMT > home
+wmt_home_initial <- read_csv(paste0(file_path, "Input/WMT/home.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > WMT > homewinter.xlsx
-wmt_homewinter_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_homewinter_20191126.xlsx")) %>% 
+## FINAL DATA SETS > WMT > homewinter
+wmt_homewinter_initial <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > WMT > Home Data > parentdemographics.xlsx
+## FINAL DATA SETS > WMT > Home Data > parentdemographics
 wmt_parent_demographics_initial <-
-  read_xlsx(paste0(file_path, "Input/WMT/wmt_parent_demographics_20191126.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/WMT/parentdemographics.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > child.xlsx
-nn_child_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_child_20191216.xlsx")) %>% 
+## FINAL DATA SETS > Navajo > child
+nn_child_initial <- read_xlsx(paste0(file_path, "Input/NN/child.xlsx"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > home.xlsx
-nn_home_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_home_20191216.xlsx")) %>% 
+## FINAL DATA SETS > Navajo > home
+nn_home_initial <- read_xlsx(paste0(file_path, "Input/NN/home.xlsx"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > homewinter.xlsx
-nn_homewinter_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_homewinter_20191216.xlsx")) %>% 
+## FINAL DATA SETS > Navajo > homewinter
+nn_homewinter_initial <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
   mutate(winter_id = as.numeric(winter_id),
-         winter_id = if_else(home_id == 62 | home_id == 45, winter_id + 1, winter_id),
          winter_id = as.character(winter_id)) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > Navajo > Home Data > parentdemographics.xlsx
+## FINAL DATA SETS > Navajo > Home Data > parentdemographics
 nn_parent_demographics_initial <-
-  read_xlsx(paste0(file_path, "Input/NN/nn_parent_demographics_20191216.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/NN/parentdemographics.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > child.xlsx
-ak_child_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_child_20191216.xlsx")) %>% 
+## FINAL DATA SETS > AK > child
+ak_child_initial <- read_xlsx(paste0(file_path, "Input/AK/child.xlsx"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > home.xlsx
-ak_home_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_home_20191216.xlsx")) %>% 
+## FINAL DATA SETS > AK > home
+ak_home_initial <- read_xlsx(paste0(file_path, "Input/AK/home.xlsx"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > homewinter.xlsx
-ak_homewinter_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_homewinter_20191216.xlsx")) %>% 
+## FINAL DATA SETS > AK > homewinter
+ak_homewinter_initial <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
 ## Box location: All Files > NoonGroupData > KidsAIR > 
-## 2019-November-Data > AK > Home Data > parentdemographics.xlsx
+## FINAL DATA SETS > AK > Home Data > parentdemographics
 ak_parent_demographics_initial <-
-  read_xlsx(paste0(file_path, "Input/AK/ak_parent_demographics_20191216.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/AK/parentdemographics.csv"), na = "NULL") %>% 
   rename_all(tolower) %>% 
@@
   mutate(home_id = as.character(home_id)) %>% 
-  # distinct(area, home_id)  # distinct homes in database = 360
+  # distinct(area, home_id)  # distinct homes in database = 390
   filter(!is.na(child_id_num)) %>% 
@@
 kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
+  mutate(home = gsub("_", "", home)) %>% 
+  mutate(home = gsub("-", "", home)) %>% 
+  mutate(home = gsub(" ", "", home)) %>% 
   mutate(home_id = as.numeric(home_id)) %>% 
@@
   filter(!is.na(treatment)) %>% 
+  filter(winter_id != 99) %>% 
   # distinct(area, home_id)  # randomized homes with treatment assignment = 342
@@
 
-ak_demographics <- read_rds(paste0(file_path, "Output/ak_demographic_data.rds"))
-nn_demographics <- read_rds(paste0(file_path, "Output/nn_demographic_data.rds"))
-wmt_demographics <- read_rds(paste0(file_path, "Output/wmt_demographic_data.rds"))
+ak_demographics <- read_rds(paste0(file_path, "Output/ak_demographic_data.rds")) %>% 
+  mutate(intervention_date = mdy(intervention_date))
+nn_demographics <- read_rds(paste0(file_path, "Output/nn_demographic_data.rds")) %>% 
+  mutate(intervention_date = mdy(intervention_date))
+wmt_demographics <- read_rds(paste0(file_path, "Output/wmt_demographic_data.rds")) %>% 
+  mutate(intervention_date = ymd(intervention_date))
 
 kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics) %>% 
+  mutate(home = gsub("_", "", home)) %>% 
+  mutate(home = gsub("-", "", home)) %>% 
+  mutate(home = gsub(" ", "", home)) %>% 
   mutate(home_id = as.numeric(home_id)) %>% 
@@
   filter(!is.na(treatment)) %>% 
-  separate(intervention_time, c("trash", "intervention_time"), sep = " ") %>% 
   unite(intervention_datetime, c("intervention_date", "intervention_time"),
@@
   mutate(intervention_datetime = ymd_hms(intervention_datetime)) %>% 
-  select(-trash) %>% 
   ungroup() %>% 
@@
 
-write_rds(kids_demographics, paste0(file_path, "Output/kids_combined_demographics.rds"))
+write_rds(kids_demographics, paste0(file_path, "Output/demographics_clean.rds"))
 ```
+
+# Check observations against documented reasons for dropping from study
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+
+demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+
+# Checking observations against drop reasons from "KidsAIR reasons for dropping" excel file
+# Filter out observations below when participants had dropped and should not be in dataset
+demographics_check <- demographics_initial %>% 
+  filter(area == "AK") %>% 
+  arrange(home, child_id_char)
+
+# dropped before winter 1 - filter out
+filter_out_wmt <- c("WMA227", "WMA254", "WMA337", "WMA600", "WMA601")
+# missing winter 2 demographic data but had not dropped - don't filter out
+keep_winter_2_wmt <- c("WMA407", "WMA497", "WMA509")
+# NN homes with missing data but had not dropped - don't filter out
+keep_nn <- c("CH110", "TC359", "TC452", "TC453")
+
+demographics_new <- demographics_initial %>% 
+  filter(!home %in% filter_out_wmt) %>% 
+  mutate(keep_var_wmt = if_else(home %in% keep_winter_2_wmt & winter_id == 2, 1, 0),
+         filter_var_wmt = if_else(area == "WMT" & keep_var_wmt == 0 & winter_id == 2 & is.na(gender_parent), 1, 0)) %>% 
+  filter(filter_var_wmt == 0) %>% 
+  filter(home != "CH105") %>% 
+  mutate(keep_var_nn = if_else(home %in% keep_nn, 1, 0),
+         filter_var_nn = if_else(area == "NN" & keep_var_nn == 0 & is.na(gender_parent), 1, 0)) %>% 
+  filter(filter_var_nn == 0) %>% 
+  filter(home != "YKC19102") %>% 
+  filter(home != "YKC19104") %>% 
+  mutate(filter_var_ak = if_else(area == "AK" & is.na(gender_parent), 1, 0)) %>% 
+  filter(filter_var_ak == 0) %>%
+  mutate(treatment_assigned = as.character(treatment),
+         treatment_actual = as.character(treatment),
+         treatment_assigned = if_else(treatment == "Placebo - TX Education", "Placebo", treatment_actual)) %>% 
+  select(-treatment, -filter_var_wmt, -keep_var_wmt, -filter_var_nn, 
+         -keep_var_nn, -filter_var_ak)
+
+write_rds(demographics_new, paste0(file_path, "Output/demographics_final.rds"))
+
+kids_linked_ids_new <- demographics_new %>% 
+  select(area, home, home_id, winter_id, home_winter_id, treatment_assigned, treatment_actual) %>% 
+  distinct(area, home, winter_id, .keep_all = TRUE)
+
+write_rds(kids_linked_ids_new, paste0(file_path, "Output/kids_linked_ids_final.rds"))
+```

### 2020-06-17 manuscript_tables_figures.Rmd
data cleaning updates

@@
 kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+
 demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds"))
@@
   filter(winter_id == 1) %>% 
+  filter(area == "WMT") %>% 
+  select(area, home, home_id, winter_id, home_winter_id, treatment)
   #filter(treatment == "Placebo") %>% 

### 2020-06-17 merge_health_datasets.Rmd
data cleaning updates

@@
   arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
+  distinct(child_id_num, visit_date, .keep_all = TRUE) %>% 
   ungroup()
@@
 
+#health_merged5 <- health_merged4 %>% 
+# left_join(kids_medical_records, by = c("area", "home", "home_winter_id", "winter_id", 
+#                                      "child_id_num", "treatment")) %>% 
+#  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
+#  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
+#         treatment, sample_date, sampling_day, at_home_location, 
+#         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:med_rec_comments) %>% 
+#  ungroup()
+
 health_merged5 <- health_merged4 %>% 
-  left_join(kids_medical_records, by = c("area", "home", "home_winter_id", "winter_id", 
-                                      "child_id_num", "treatment")) %>% 
-  arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
-  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         treatment, sample_date, sampling_day, at_home_location, 
-         pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:med_rec_comments) %>% 
-  ungroup()
-
-health_merged6 <- health_merged5 %>% 
   left_join(kids_lrti, by = c("area", "home", "home_winter_id", "winter_id", 
@@
 
-write_rds(health_merged6, paste0(output_path, "Output/health_data_merged.rds"))
-write_csv(health_merged6, paste0(output_path, "Output/health_data_merged.csv"))
+write_rds(health_merged5, paste0(output_path, "Output/health_data_merged.rds"))
+write_csv(health_merged5, paste0(output_path, "Output/health_data_merged.csv"))
 ```

### 2020-06-19 dustrak_data_cleaning.Rmd
new data cleaning/merging

@@
 kids_linked_ids <- 
-  read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+  read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
@@
   ungroup() %>% 
+  filter(home != "WMA300") %>%  # filtering out home with sampling issues after discussion with team
   arrange(area, home_winter_id, pm_datetime_new)

### 2020-06-19 home_char_act_cleaning.Rmd
new data cleaning/merging

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 4 March 2020"
+date: "Started 13 Jan 2020, Updated 19 June 2020"
 output: pdf_document
@@
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) 
 ```
@@
 # Load data
-home_char_initial_wmt <- read_csv("Input/WMT/wmt_home_char_20200113.csv")
+home_char_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/homechar.csv"), 
+                                  na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(wmt_home_char, "Output/wmt_home_char.rds")
+write_rds(wmt_home_char, paste0(file_path, "Output/wmt_home_char.rds"))
 
@@
 # Load data
-home_char_initial_nn <- read_xlsx("Input/NN/nn_home_char_20200113.xlsx")
+home_char_initial_nn <- read_csv(paste0(file_path, "Input/NN/homechar.csv"), na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(nn_home_char, "Output/nn_home_char.rds")
+write_rds(nn_home_char, paste0(file_path, "Output/nn_home_char.rds"))
 
@@
 # Load data
-home_char_initial_ak <- read_xlsx("Input/AK/ak_home_char_20200113.xlsx")
+home_char_initial_ak <- read_csv(paste0(file_path, "Input/AK/homechar.csv"), na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(ak_home_char, "Output/ak_home_char.rds")
+write_rds(ak_home_char, paste0(file_path, "Output/ak_home_char.rds"))
 ```
@@
 ```{r}
-wmt_home_char <- read_rds("Output/wmt_home_char.rds")
-nn_home_char <- read_rds("Output/nn_home_char.rds")
-ak_home_char <- read_rds("Output/ak_home_char.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_home_char <- read_rds(paste0(file_path, "Output/wmt_home_char.rds"))
+nn_home_char <- read_rds(paste0(file_path, "Output/nn_home_char.rds"))
+ak_home_char <- read_rds(paste0(file_path, "Output/ak_home_char.rds"))
 
@@
 
-# write_rds(kids_home_char, "Output/kids_home_char.rds")
+# write_rds(kids_home_char, paste0(file_path, "Output/kids_home_char.rds"))
 ```
@@
 kids_home_char <- read_rds(paste0(file_path, "Output/kids_home_char.rds"))
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
 home_char_clean <- kids_home_char %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
   # cleaning based on checking odd values with paper logs
@@
 
+summary(home_char_clean)
+
 write_rds(home_char_clean, paste0(file_path, "Output/home_char_clean.rds"))
@@
 # Load data
-home_act_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/wmt_home_activity_20200113.csv"))
+home_act_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/homeactivity.csv"), na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(wmt_home_act, paste0(file_path, "Output/wmt_home_act.rds"))
+write_rds(wmt_home_act, paste0(file_path, "Output/wmt_home_act.rds"))
 
@@
 # Load data
-home_act_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/nn_home_activity_20200113.xlsx"))
+home_act_initial_nn <- read_csv(paste0(file_path, "Input/NN/homeactivity.csv"), na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(nn_home_act, paste0(file_path, "Output/nn_home_act.rds"))
+write_rds(nn_home_act, paste0(file_path, "Output/nn_home_act.rds"))
 
@@
 # Load data
-home_act_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/ak_home_activity_20200113.xlsx"))
+home_act_initial_ak <- read_csv(paste0(file_path, "Input/AK/homeactivity.csv"), na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(ak_home_act, paste0(file_path, "Output/ak_home_act.rds"))
+write_rds(ak_home_act, paste0(file_path, "Output/ak_home_act.rds"))
 ```
@@
 kids_home_act <- read_rds(paste0(file_path, "Output/kids_home_act.rds"))
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
 home_act_clean <- kids_home_act %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, by = c("area", "home_winter_id"))
+
+summary(home_act_clean)
 

### 2020-06-19 home_child_data_cleaning.Rmd
new data cleaning/merging

@@
   select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
-  filter(winterid != 3) %>% # filter for winterid == 1 or 2. What about missing intervention date?
+  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   mutate(home_winter_id = as.character(homewinterid),
@@
          winter_id = as.character(winter_id)) %>% 
-  filter(winter_id == 1 | winter_id ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
+  filter(winter_id == 0 | winter_id == 1 | winter_id ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
   select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)
@@
   filter(!is.na(treatment)) %>% 
-  filter(winter_id != 99) %>% 
   # distinct(area, home_id)  # randomized homes with treatment assignment = 342
@@
 demographics_check <- demographics_initial %>% 
-  filter(area == "AK") %>% 
+  filter(area == "NN") %>% 
   arrange(home, child_id_char)
@@
 # NN homes with missing data but had not dropped - don't filter out
-keep_nn <- c("CH110", "TC359", "TC452", "TC453")
+keep_nn <- c("CH110", "CH209", "TC359", "TC452", "TC453")
 
@@
   filter(filter_var_nn == 0) %>% 
+  mutate(filter_var_nn = if_else(home == "CH209" & winter_id == 2, 1, 0)) %>% 
+  filter(filter_var_nn == 0) %>% 
   filter(home != "YKC19102") %>% 
@@
          treatment_actual = as.character(treatment),
-         treatment_assigned = if_else(treatment == "Placebo - TX Education", "Placebo", treatment_actual)) %>% 
+         treatment_assigned = if_else(treatment == "Placebo - TX Education", "Placebo", treatment_actual),
+         winter_id = as.factor(winter_id),
+         treatment_assigned = as.factor(treatment_assigned),
+         treatment_actual = as.factor(treatment_actual),
+         area = as.factor(area)) %>% 
   select(-treatment, -filter_var_wmt, -keep_var_wmt, -filter_var_nn, 
@@
 ```
+
+# Final check for home and child observations by winter and treatment
+```{r}
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) %>% 
+  count(area, winter_id, treatment_assigned)
+
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
+  count(area, winter_id, treatment_assigned)
+```

### 2020-06-19 indoor_temp_rh_summary.Rmd
new data cleaning/merging

@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 # Load data
-indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-pm_clean <- read_rds("Output/pm_clean.rds")
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) 
+pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))
 ```

### 2020-06-19 manuscript_tables_figures.Rmd
new data cleaning/merging

@@
 
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
-demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
 

### 2020-06-22 child_health_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 28 May 2020"
+date: "Started 27 Feb 2020, Updated 22 June 2020"
 output: pdf_document 
@@
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
-
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
-
-demographics_clean <- 
-  read_rds(paste0(file_path, "Output/demographics_clean.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
 
@@
 ```{r}
-wmt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_athomelog_20200227.xlsx")) %>% 
+wmt_athomelog_initial <- read_csv(paste0(file_path, "Input/WMT/athomelog.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
@@
 
-nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_athomelog_20200416.xlsx")) %>% 
+nn_athomelog_initial <- read_csv(paste0(file_path, "Input/NN/athomelog.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
@@
 
-ak_athomelog_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_athomelog_20200227.xlsx")) %>% 
+ak_athomelog_initial <- read_csv(paste0(file_path, "Input/AK/athomelog.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.factor(HomeWinterID),
@@
                           ak_athomelog_initial) %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
+  filter(!is.na(child_id_num))
 
@@
          "23:00:00" = "home") %>% 
-  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
-         sampling_day, sample_time, at_home, "00:00:00":"23:00:00") %>% 
+  select(area, home, home_winter_id, child_id_num, winter_id, treatment_assigned,
+         treatment_actual, sampling_day, sample_time, at_home, "00:00:00":"23:00:00") %>% 
   group_by(area, home_winter_id, child_id_num, sampling_day, sample_time) %>% 
@@
   arrange(area, home_winter_id, child_id_num, sampling_day, sample_time) %>% 
-  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
-         sampling_day, sample_time, at_home, at_home_location)  %>% 
+  select(area, home, home_winter_id, child_id_num, winter_id, treatment_assigned,
+         treatment_actual, sampling_day, sample_time, at_home, at_home_location)  %>% 
   group_by(area, home_winter_id, child_id_num, sampling_day) %>% 
@@
   mutate(sampling_day = as.character(sampling_day)) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
-         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         sampling_day, sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
          pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
@@
   left_join(kids_at_home_log_new, 
-            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
-                   "sampling_day", "sample_time")) %>% 
+            by = c("area", "home", "home_winter_id", "winter_id", "treatment_assigned",
+                   "treatment_actual", "sampling_day", "sample_time")) %>% 
   group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
@@
 ```{r}
-wmt_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalvisit_20200227.xlsx")) %>% 
+wmt_medicalvisit_initial <- read_csv(paste0(file_path, "Input/WMT/medicalvisit.csv"), 
+                                  na = c("NULL", "")) %>% 
   # recode values to NA
@@
          area = "WMT",
-         med_visit_date = ymd(VisitDate),
-         med_visit_ari = as.character(ARI),
-         med_visit_lethargy = as.character(Lethargy),
-         med_visit_retractions = as.character(Retractions),
+         # fix data entry error for date
+         med_visit_date = as.character(VisitDate),
+         med_visit_date = if_else(med_visit_date == "2025-11-04", "2015-11-04", med_visit_date),
+         med_visit_date = ymd(med_visit_date),
+         med_visit_ari = as.factor(ARI),
+         med_visit_lethargy = as.factor(Lethargy),
+         med_visit_retractions = as.factor(Retractions),
          temp1 = as.numeric(Temp1),
@@
          med_visit_heart_rate = as.numeric(HR),
-         med_visit_wheeze = as.character(Wheeze),
-         med_visit_stridor = as.character(Stridor),
-         med_visit_crackles = as.character(Crackles),
+         med_visit_wheeze = as.factor(Wheeze),
+         med_visit_stridor = as.factor(Stridor),
+         med_visit_crackles = as.factor(Crackles),
          med_visit_sp02 = as.numeric(SP02),
          med_visit_hr_sp02 = as.numeric(HRPulseOx),
-         med_visit_referred = as.character(Referred),
-         med_visit_notes = as.character(Level)) %>% 
+         med_visit_referred = as.factor(Referred)) %>% 
   # gather 3 temperatures to calculate average
@@
   ungroup() %>% 
-  # spread 3 moisture samples back to wide format
+  # spread back to wide format
   spread(sample, value) %>%
+  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
   select(home_winter_id:med_visit_mean_temp)
 
-nn_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200416.xlsx")) %>%
+nn_medicalvisit_initial <- read_csv(paste0(file_path, "Input/NN/medicalvisit.csv"), 
+                                  na = c("NULL", "")) %>%
   mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
@@
          med_visit_date = ymd(VisitDate),
-         med_visit_ari = as.character(ARI),
-         med_visit_lethargy = as.character(Lethargy),
-         med_visit_retractions = as.character(Retractions),
+         med_visit_ari = as.factor(ARI),
+         med_visit_lethargy = as.factor(Lethargy),
+         med_visit_retractions = as.factor(Retractions),
          temp1 = as.numeric(Temp1),
@@
   ungroup() %>% 
-  # spread 3 moisture samples back to wide format
+  # spread back to wide format
   spread(sample, value) %>%
@@
 
-ak_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_medicalvisit_20200416.xlsx")) %>% 
+ak_medicalvisit_initial <- read_csv(paste0(file_path, "Input/AK/medicalvisit.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
@@
                           ak_medicalvisit_initial) %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, c("area", "home_winter_id"))
+
+summary(kids_med_visit)
 
@@
 ```{r}
-wmt_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_birthinfancy_20200227.xlsx")) %>% 
+wmt_birthinfancy_initial <- read_csv(paste0(file_path, "Input/WMT/birthinfancy.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          area = "WMT",
-         gender = as.character(Gender),
+         gender = as.factor(Gender),
          bw_lbs = as.numeric(BirthweightLbs),
@@
          mom_age = as.numeric(AgeOfMom),
-         breastfeed_current = as.character(CurrentlyBreastfeeding),
-         breastfeed_past = as.character(DidBreastfeed),
+         breastfeed_current = as.factor(CurrentlyBreastfeeding),
+         breastfeed_past = as.factor(DidBreastfeed),
          bf_years = as.numeric(HowLongYears),
@@
          breastfeed_months = as.numeric((bf_years*12)+bf_months),
-         flu_shot = as.character(FluShot),
-         flu_shot_planned = as.character(FluShotPlanned),
-         asthma_meds = as.character(AsthmaMeds),
-         child_school = as.character(Attend),
+         flu_shot = as.factor(FluShot),
+         flu_shot_planned = as.factor(FluShotPlanned),
+         asthma_meds = as.factor(AsthmaMeds),
+         child_school = as.factor(Attend),
          school_hours = as.numeric(HoursDay),
@@
          school_hours_week = as.numeric(school_hours*school_days),
-         sibling_school = as.character(SibSchool)) %>% 
+         sibling_school = as.factor(SibSchool)) %>% 
   replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
@@
 
-nn_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_birthinfancy_20200416.xlsx")) %>% 
+nn_birthinfancy_initial <- read_csv(paste0(file_path, "Input/NN/birthinfancy.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          area = "NN",
-         gender = as.character(Gender),
+         gender = as.factor(Gender),
          bw_lbs = as.numeric(BirthweightLbs),
@@
 
-ak_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_birthinfancy_20200416.xlsx")) %>% 
+ak_birthinfancy_initial <- read_csv(paste0(file_path, "Input/AK/birthinfancy.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          area = "AK",
-         gender = as.character(Gender),
+         gender = as.factor(Gender),
          bw_lbs = as.numeric(BirthweightLbs),
@@
          school_hours_week = as.numeric(school_hours*school_days),
-         sibling_school = as.character(SibSchool)) %>% 
+         sibling_school = as.factor(SibSchool)) %>% 
   replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>% 
@@
                             ak_birthinfancy_initial) %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, c("area", "home_winter_id"))
+
+summary(kids_birth_infancy)
 
@@
 
-As of 22 April 2020 we only have CHW files for WMT.
+As of 22 June 2020 we only have CHW files for WMT.
 The content isn't particularly necessary for data analysis, so leave out of the
@@
 wmt_childhomewinter_initial <- 
-  read_xlsx(paste0(file_path, "Input/WMT/wmt_childhomewinter_20200227.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/WMT/childhomewinter.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 wmt_csq_initial <- 
-  read_xlsx(paste0(file_path, "Input/WMT/wmt_csq_20200227.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/WMT/csq.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          area = "WMT",
-         csq_date = ymd(Date),
+         csq_date = mdy(Date),
          csq_collect = as.character(HowCollected),
@@
          csq_still_sick = as.character(StillSick),
-         csq_first_symptom_date = ymd(FirstSymptomDate),
-         csq_symptom_free_date = ymd(SymptomFreeDate),
+         csq_first_symptom_date = mdy(FirstSymptomDate),
+         csq_symptom_free_date = mdy(SymptomFreeDate),
          csq_miss_work = as.character(MissWork),
@@
          csq_cough_current = as.character(CoughingStill),
-         csq_wheeze_current = as.character(WheezingStill),
-         csq_comments = as.character(Comments)) %>% 
+         csq_wheeze_current = as.character(WheezingStill)) %>% 
   replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
-  select(home_winter_id:csq_comments)
+  select(home_winter_id:csq_wheeze_current)
 
 nn_csq_initial <- 
-  read_xlsx(paste0(file_path, "Input/NN/nn_csq_20200416.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/NN/csq.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          csq_cough_current = as.character(CoughingStill),
-         csq_wheeze_current = as.character(WheezingStill),
-         csq_comments = as.character(`REPLACE(REPLACE(Comments, '\\n', ''), '\\r', '')`)) %>% 
+         csq_wheeze_current = as.character(WheezingStill)) %>% 
   replace_with_na_all(condition = ~.x %in% c(-9999, -999, -99, -99.9, 999, 0)) %>% 
-  select(home_winter_id:csq_comments)
+  select(home_winter_id:csq_wheeze_current)
 
 ak_csq_initial <- 
-  read_xlsx(paste0(file_path, "Input/AK/ak_csq_20200416.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/AK/csq.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          area = "AK",
-         csq_date = ymd(Date),
+         csq_date = mdy(Date),
          csq_collect = as.character(HowCollected),
@@
          csq_cough_current = as.character(CoughingStill),
-         csq_wheeze_current = as.character(WheezingStill),
-         csq_comments = as.character(`REPLACE(REPLACE(Comments, '\\n', ''), '\\r', '')`)) %>% 
+         csq_wheeze_current = as.character(WheezingStill)) %>% 
   replace_with_na_all(condition = ~.x %in% c(-999, -99, -99.9, 999, 0)) %>% 
-  select(home_winter_id:csq_comments)
+  select(home_winter_id:csq_wheeze_current)
 
@@
 kids_csq <- rbind(wmt_csq_initial, nn_csq_initial, ak_csq_initial) %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
+  mutate_if(is.character, as.factor)
+
+summary(kids_csq)
 
@@
 wmt_medicalrecords_initial <- 
-  read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalrecords_20200227.xlsx")) %>% 
+  read_csv(paste0(file_path, "Input/WMT/medicalrecords.csv"), 
+                                  na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
                               ak_medicalrecords_initial) %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
   arrange(area, visit_date)
+
+summary(kids_medical_records)
 
@@
 
-There isn't an LRTI file for NN/AK as of 22 April 2020. 
+There isn't an LRTI file for NN/AK as of 22 June 2020. 
 Leaving WMT file for now and will check back on updated data/files later.
 ```{r}
-wmt_ari_initial <- read_xlsx("Input/WMT/wmt_ari_20200408.xlsx") %>% 
+wmt_ari_initial <- read_csv("Input/WMT/ari.csv") %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 kids_lrti <- wmt_ari_initial %>% 
-  left_join(kids_linked_ids, c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, c("area", "home_winter_id"))
 

### 2020-06-22 home_char_act_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 19 June 2020"
+date: "Started 13 Jan 2020, Updated 22 June 2020"
 output: pdf_document
@@
 
-kids_home_act <- rbind(wmt_home_act, nn_home_act, ak_home_act)
+kids_home_act <- rbind(wmt_home_act, nn_home_act, ak_home_act) %>% 
+  select(-home_act_comments)
 
@@
   mutate(area = as.factor(area)) %>% 
-  right_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  filter(!is.na(home_act_day))
 

### 2020-06-22 sums_data_cleaning.Rmd
data cleaning updates

@@
 author: "Ethan Walker"
-date: "Started 23 Dec 2019, Updated 6 Feb 2020"
+date: "Started 23 Dec 2019, Updated 22 June 2020"
 output: pdf_document
@@
                   "#FF3366", "#33CC99", "#33999")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 ```
@@
 # Input file from Box: ibutton.xlsx
-wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>%
+wmt_sums_data_log <- read_csv(paste0(file_path, "Input/WMT/ibutton.csv"), 
+                              na = c("NULL", "")) %>%
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) %>% 
   filter(area == "WMT") %>% 
@@
 wmt_sums_log <- wmt_sums_data_log %>% 
-  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  right_join(kids_linked_ids, by = "home_winter_id") %>% 
   select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
   arrange(home, date_log) %>% 
-  mutate(date_log = mdy(date_log))
-
-write_rds(wmt_sums_log, "Output/wmt_sums_data_log.rds")
-write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")
+  mutate(date_log = ymd(date_log))
+
+write_rds(wmt_sums_log, paste0(file_path, "Output/wmt_sums_data_log.rds"))
+#write_csv(wmt_sums_log, paste0(file_path, "Output/wmt_sums_data_log.csv"))
 
 # Input file from Box: ibuttonupd.xlsx
-wmt_sums_data_log_updated <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
+wmt_sums_data_log_updated <- read_csv(paste0(file_path, "Input/WMT/ibuttonupd.csv"), 
+                                      na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          alarms = Alarms,
-         file_name = ibutFile,
-         notes = `REPLACE(REPLACE(Notes`) %>% 
-  select(home_winter_id, date, time, alarms, file_name, notes)
+         file_name = ibutFile) %>% 
+  select(home_winter_id, date, time, alarms, file_name)
 
 wmt_sums_log <- wmt_sums_data_log_updated %>% 
-  left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
+  right_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, alarms, file_name) %>% 
   arrange(home, date)
 
-write_csv(wmt_sums_log, "Output/wmt_sums_data_log_updated.csv")
+# write_csv(wmt_sums_log, paste0(file_path, "Output/wmt_sums_data_log_updated.csv"))
 ```
@@
 # Input file from Box: ibutton.xlsx
-nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
+nn_sums_data_log <- read_csv(paste0(file_path, "Input/NN/ibutton.csv"), 
+                              na = c("NULL", "")) %>%
   mutate(home_winter_id = as.character(HomeWinterID),
@@
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) %>% 
   filter(area == "NN") %>% 
@@
 nn_sums_log <- nn_sums_data_log %>% 
-  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  right_join(kids_linked_ids, by = "home_winter_id") %>% 
   select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
   arrange(home, date_log) %>% 
-  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
-  select(-trash)
-
-write_rds(nn_sums_log, "Output/nn_sums_data_log.rds")
-write_csv(nn_sums_log, "Output/nn_sums_data_log.csv")
+  mutate(date_log = ymd(date_log))
+
+write_rds(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log.rds"))
+#write_csv(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log.csv"))
 
 # Input file from Box: ibuttonupd.xlsx
-nn_sums_data_log_updated <- read_xlsx("Input/NN/nn_sums_data_log_updated.xlsx") %>% 
+nn_sums_data_log_updated <- read_csv(paste0(file_path, "Input/NN/ibuttonupd.csv"), 
+                                      na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          alarms = Alarms,
-         file_name = ibutFile,
-         notes = IgnoreReason) %>% 
-  select(home_winter_id, date, time, alarms, file_name, notes)
+         file_name = ibutFile) %>% 
+  select(home_winter_id, date, time, alarms, file_name)
 
 nn_sums_log <- nn_sums_data_log_updated %>% 
-  left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
+  right_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, alarms, file_name) %>% 
   arrange(home, date)
 
-write_csv(nn_sums_log, "Output/nn_sums_data_log_updated.csv")
+# write_csv(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log_updated.csv"))
 ```
@@
 # Input file from Box: ibutton.xlsx
-ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
+ak_sums_data_log <- read_csv(paste0(file_path, "Input/AK/ibutton.csv"), 
+                              na = c("NULL", "")) %>%
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          time_log = InstallTime) %>% 
-  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
   select(home_winter_id, ibid, date_log, time_log)
 
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) %>% 
   filter(area == "AK") %>% 
@@
 ak_sums_log <- ak_sums_data_log %>% 
-  left_join(kids_linked_ids, by = "home_winter_id") %>% 
+  right_join(kids_linked_ids, by = "home_winter_id") %>% 
   select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
-  arrange(home, date_log)
-
-write_csv(ak_sums_log, "Output/ak_sums_data_log.csv")
-write_rds(ak_sums_log, "Output/ak_sums_data_log.rds")
-
-# Input file from Box: ibutton.xlsx
-ak_sums_data_log_updated <- read_xlsx("Input/AK/ak_sums_data_log_updated.xlsx") %>% 
+  arrange(home, date_log) %>% 
+  mutate(date_log = ymd(date_log))
+
+write_rds(ak_sums_log, paste0(file_path, "Output/ak_sums_data_log.rds"))
+#write_csv(ak_sums_log, paste0(file_path, "Output/ak_sums_data_log.csv"))
+
+# Input file from Box: ibuttonupd.xlsx
+ak_sums_data_log_updated <- read_csv(paste0(file_path, "Input/AK/ibuttonupd.csv"), 
+                                      na = c("NULL", "")) %>% 
   mutate(home_winter_id = as.character(HomeWinterID),
@@
          alarms = Alarms,
-         file_name = ibutFile,
-         notes = IgnoreReason) %>% 
-  select(home_winter_id, date, time, alarms, file_name, notes)
+         file_name = ibutFile) %>% 
+  select(home_winter_id, date, time, alarms, file_name)
 
 ak_sums_log <- ak_sums_data_log_updated %>% 
-  left_join(kids_linked_ids, by = "home_winter_id") %>% 
-  select(home_winter_id, home, home_id, date, time, alarms, file_name, notes) %>% 
+  right_join(kids_linked_ids, by = "home_winter_id") %>% 
+  select(home_winter_id, home, home_id, date, time, alarms, file_name) %>% 
   arrange(home, date)
 
-write_csv(ak_sums_log, "Output/ak_sums_data_log_updated.csv")
+# write_csv(nn_sums_log, paste0(file_path, "Output/nn_sums_data_log_updated.csv"))
 ```
@@
 ## The files have 5 columns: line, date, time, temp (C), notes
-list_csv_files <- list.files("Input/WMT/sums_csv_new")
+list_csv_files <- list.files(paste0(file_path, "Input/WMT/sums_csv_new"))
 
@@
 ## run next 13 lines together
-setwd("Input/WMT/sums_csv_new") # set new working directory
+setwd(paste0(file_path, "Input/WMT/sums_csv_new")) # set new working directory
 initial_data = tibble(file_list = list_csv_files) %>% # use list of files from above
@@
 ## a 21 row header, then columns of datetime and temp (C)
-list_csv_files_2 <- list.files("Input/WMT/sums_csv_old/format_1")
+list_csv_files_2 <- list.files(paste0(file_path, "Input/WMT/sums_csv_old/format_1"))
 
@@
 ## run next 9 lines together
-setwd("Input/WMT/sums_csv_old/format_1")
+setwd(paste0(file_path, "Input/WMT/sums_csv_old/format_1"))
 initial_data = tibble(file_list = list_csv_files_2) %>%
@@
 ## 5 columns: line, date, time, temp (F), notes
-list_csv_files_3 <- list.files("Input/WMT/sums_csv_old/format_2")
+list_csv_files_3 <- list.files(paste0(file_path, "Input/WMT/sums_csv_old/format_2"))
 
@@
 ## run next 12 lines together
-setwd("Input/WMT/sums_csv_old/format_2")
+setwd(paste0(file_path, "Input/WMT/sums_csv_old/format_2"))
 initial_data = tibble(file_list = list_csv_files_3) %>%
@@
 ## 2 columns: datetime, temp (C)
-list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip16")
+list_xlsx_files <- list.files(paste0(file_path, "Input/WMT/sums_xlsx_files/header_skip16"))
 
@@
 ## run next 6 lines together
-setwd("Input/WMT/sums_xlsx_files/header_skip16")
+setwd(paste0(file_path, "Input/WMT/sums_xlsx_files/header_skip16"))
 initial_data = tibble(file_list = list_xlsx_files) %>%
@@
 ## 2 columns: datetime, temp (C)
-list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip21")
+list_xlsx_files <- list.files(paste0(file_path, "Input/WMT/sums_xlsx_files/header_skip21"))
 
@@
 ## run next 6 lines together
-setwd("Input/WMT/sums_xlsx_files/header_skip21")
+setwd(paste0(file_path, "Input/WMT/sums_xlsx_files/header_skip21"))
 initial_data = tibble(file_list = list_xlsx_files) %>%
@@
 ## 2 columns: datetime, temp (C)
-list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_2col_c")
+list_xlsx_files <- list.files(paste0(file_path, "Input/WMT/sums_xlsx_files/no_header_2col_c"))
 
@@
 ## run next 6 lines together
-setwd("Input/WMT/sums_xlsx_files/no_header_2col_c")
+setwd(paste0(file_path, "Input/WMT/sums_xlsx_files/no_header_2col_c"))
 initial_data = tibble(file_list = list_xlsx_files) %>%
@@
 ## 4 columns: date, time, am/pm, temp (C)
-list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_c")
+list_xlsx_files <- list.files(paste0(file_path, "Input/WMT/sums_xlsx_files/no_header_4col_c"))
 
@@
 ## run next 6 lines together
-setwd("Input/WMT/sums_xlsx_files/no_header_4col_c")
+setwd(paste0(file_path, "Input/WMT/sums_xlsx_files/no_header_4col_c"))
 initial_data = tibble(file_list = list_xlsx_files) %>%
@@
 ## 4 columns: date, time, am/pm, temp (F)
-list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_f")
+list_xlsx_files <- list.files(paste0(file_path, "Input/WMT/sums_xlsx_files/no_header_4col_f"))
 
@@
 ## run next 6 lines together
-setwd("Input/WMT/sums_xlsx_files/no_header_4col_f")
+setwd(paste0(file_path, "Input/WMT/sums_xlsx_files/no_header_4col_f"))
 initial_data = tibble(file_list = list_xlsx_files) %>%
@@
 
-# write_rds(wmt_sums_raw, "Output/wmt_sums_raw.rds")
+# write_rds(wmt_sums_raw, paste0(file_path, "Output/wmt_sums_raw.rds"))
 ```
@@
 ## The files have 5 columns: line, date, time, temp (C), notes
-list_csv_files <- list.files("Input/NN/sums_csv_new")
+list_csv_files <- list.files(paste0(file_path, "Input/NN/sums_csv_new"))
 
@@
 ## run next 12 lines together
-setwd("Input/NN/sums_csv_new")
+setwd(paste0(file_path, "Input/NN/sums_csv_new"))
 initial_data = tibble(file_list = list_csv_files) %>%
@@
 # 1 file had this format: No header, columns of datetime and temp (C)
-list_csv_files_2 <- list.files("Input/NN/sums_csv_old/2col-c")
+list_csv_files_2 <- list.files(paste0(file_path, "Input/NN/sums_csv_old/2col-c"))
 
@@
 ## run next 7 lines together
-setwd("Input/NN/sums_csv_old/2col-c")
+setwd(paste0(file_path, "Input/NN/sums_csv_old/2col-c"))
 initial_data = tibble(file_list = list_csv_files_2) %>%
@@
 ##### COMPARED TO CORRESPONDING LOGTAG FILES, THE TIMESTAMPS ARE OFF
-list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")
+list_csv_files_3 <- list.files(paste0(file_path, "Input/NN/sums_csv_old/3col-f"))
 
@@
 ## run next 12 lines together
-setwd("Input/NN/sums_csv_old/3col-f")
+setwd(paste0(file_path, "Input/NN/sums_csv_old/3col-f"))
 initial_data = tibble(file_list = list_csv_files_3) %>%
@@
 
-# write_rds(nn_sums_raw, "Output/nn_sums_raw.rds")
+# write_rds(nn_sums_raw, paste0(file_path, "Output/nn_sums_raw.rds"))
 ```
@@
 ## The files have 5 columns: line, date, time, temp (C), notes
-list_csv_files <- list.files("Input/AK/sums_csv_new")
+list_csv_files <- list.files(paste0(file_path, "Input/AK/sums_csv_new"))
 
@@
 ## run next 12 lines together
-setwd("Input/AK/sums_csv_new")
+setwd(paste0(file_path, "Input/AK/sums_csv_new"))
 initial_data = tibble(file_list = list_csv_files) %>%
@@
 ##### COMPARED TO CORRESPONDING LOGTAG FILES, THE TIMESTAMPS ARE OFF
-list_csv_files_2 <- list.files("Input/AK/sums_csv_old")
+list_csv_files_2 <- list.files(paste0(file_path, "Input/AK/sums_csv_old"))
 
@@
 ## run next 7 lines together
-setwd("Input/AK/sums_csv_old")
+setwd(paste0(file_path, "Input/AK/sums_csv_old"))
 initial_data = tibble(file_list = list_csv_files_2) %>%
@@
 ## 5 columns: date, time, am or pm, C, temp
-list_xlsx_files <- list.files("Input/AK/sums_xlsx_temp_c")
+list_xlsx_files <- list.files(paste0(file_path, "Input/AK/sums_xlsx_temp_c"))
 
@@
 ## run next 6 lines together
-setwd("Input/AK/sums_xlsx_temp_c")
+setwd(paste0(file_path, "Input/AK/sums_xlsx_temp_c"))
 initial_data = tibble(file_list = list_xlsx_files) %>%
@@
 ## 5 columns: date, time, am or pm, F, temp
-list_xlsx_files2 <- list.files("Input/AK/sums_xlsx_temp_f")
+list_xlsx_files2 <- list.files(paste0(file_path, "Input/AK/sums_xlsx_temp_f"))
 
@@
 ## run next 6 lines together
-setwd("Input/AK/sums_xlsx_temp_f")
+setwd(paste0(file_path, "Input/AK/sums_xlsx_temp_f"))
 initial_data = tibble(file_list = list_xlsx_files2) %>%
@@
 
-# write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
+# write_rds(paste0(file_path, ak_sums_raw, "Output/ak_sums_raw.rds"))
 ```
@@
 ```{r}
-wmt_sums_raw <- read_rds("Output/wmt_sums_raw.rds")
-
-nn_sums_raw <- read_rds("Output/nn_sums_raw.rds")
-
-ak_sums_raw <- read_rds("Output/ak_sums_raw.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_sums_raw <- read_rds(paste0(file_path, "Output/wmt_sums_raw.rds"))
+
+nn_sums_raw <- read_rds(paste0(file_path, "Output/nn_sums_raw.rds"))
+
+ak_sums_raw <- read_rds(paste0(file_path, "Output/ak_sums_raw.rds"))
 
@@
 
-# write_rds(all_sums_raw, "Output/all_sums_raw.rds")
+# write_rds(all_sums_raw, paste0(file_path, "Output/all_sums_raw.rds"))
 ```
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 # Load all SUMs data 
-all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
+all_sums_raw <- read_rds(paste0(file_path, "Output/all_sums_raw.rds")) %>% 
   ungroup() %>% 
@@
 # Load SUMs data logs
-wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
-nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
-ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")
+wmt_sums_log <- read_rds(paste0(file_path, "Output/wmt_sums_data_log.rds"))
+nn_sums_log <- read_rds(paste0(file_path, "Output/nn_sums_data_log.rds"))
+ak_sums_log <- read_rds(paste0(file_path, "Output/ak_sums_data_log.rds"))
 
@@
 
-write_csv(sums_time_comparison, "Output/sums_time_comparison.csv")
-write_rds(sums_time_comparison, "Output/sums_time_comparison.rds")
+write_csv(sums_time_comparison, paste0(file_path, "Output/sums_time_comparison.csv"))
+write_rds(sums_time_comparison, paste0(file_path, "Output/sums_time_comparison.rds"))
 ```
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 # Load all SUMs data
-all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
+all_sums_raw <- read_rds(paste0(file_path, "Output/all_sums_raw.rds")) %>% 
   ungroup() %>% 
@@
 
-wmt_sums_stop <- read_csv("Input/WMT/wmt_medicalvisit_20200107.csv") %>% 
-  mutate(home_winter_id = HomeWinterID,
-         final_visit_date = if_else(VisitDate == "11/4/2025", "11/4/2015", VisitDate),
-         final_visit_date = mdy(final_visit_date)) %>% 
+wmt_sums_stop <- read_csv(paste0(file_path, "Input/WMT/medicalvisit.csv")) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         final_visit_date = as.character(VisitDate),
+         final_visit_date = if_else(final_visit_date == "2025-11-04", "2015-11-04", final_visit_date),
+         final_visit_date = ymd(final_visit_date)) %>% 
   group_by(home_winter_id) %>% 
@@
 
-wmt_sums_start <- read_xlsx("Input/WMT/wmt_sums_start_times_20200107.xlsx") %>% 
+wmt_sums_start <- read_rds(paste0(file_path, "Output/wmt_sums_data_log.rds")) %>% 
+  mutate(area = "WMT",
+         time_log = as.character(time_log),
+         time_log = if_else(is.na(time_log), "00:00:00", time_log)) %>% 
+  unite("datetime_start", c("date_log", "time_log"), sep = " ") %>% 
   mutate(datetime_start = ymd_hms(datetime_start))
@@
   filter(area == "WMT") %>% 
-  left_join(wmt_sums_start_stop, by = c("area", "home")) %>% 
+  right_join(wmt_sums_start_stop, by = c("area", "home")) %>% 
   mutate(filter_var = 
@@
 
-# write_rds(wmt_sums_filtered, "Output/wmt_sums_filtered.rds")
+# write_rds(wmt_sums_filtered, paste0(file_path, "Output/wmt_sums_filtered.rds"))
 
@@
 
-nn_sums_stop <- read_xlsx("Input/NN/nn_medicalvisit_20200108.xlsx") %>% 
+nn_sums_stop <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200108.xlsx")) %>% 
   mutate(home_winter_id = HomeWinterID,
@@
 
-nn_sums_start <- read_xlsx("Input/NN/nn_sums_start_times_20200108.xlsx") %>% 
+nn_sums_start <- read_xlsx(paste0(file_path, "Input/NN/nn_sums_start_times_20200108.xlsx")) %>% 
   mutate(datetime_start = ymd_hms(datetime_start))
@@
 
-# write_rds(nn_sums_filtered, "Output/nn_sums_filtered.rds")
+# write_rds(nn_sums_filtered, paste0(file_path, "Output/nn_sums_filtered.rds"))
 
@@
 
-ak_sums_stop <- read_xlsx("Input/AK/ak_medicalvisit_20200108.xlsx") %>% 
+ak_sums_stop <- read_xlsx(paste0(file_path, "Input/AK/ak_medicalvisit_20200108.xlsx")) %>% 
   mutate(home_winter_id = HomeWinterID,
@@
 
-ak_sums_start <- read_xlsx("Input/AK/ak_sums_start_times_20200108.xlsx") %>% 
+ak_sums_start <- read_xlsx(paste0(file_path, "Input/AK/ak_sums_start_times_20200108.xlsx")) %>% 
   mutate(datetime_start = ymd_hms(datetime_start))
@@
 
-# write_rds(ak_sums_filtered, "Output/ak_sums_filtered.rds")
+# write_rds(ak_sums_filtered, paste0(file_path, "Output/ak_sums_filtered.rds"))
 ```
@@
 ```{r}
-wmt_sums_filtered <- read_rds("Output/wmt_sums_filtered.rds")
-nn_sums_filtered <- read_rds("Output/nn_sums_filtered.rds")
-ak_sums_filtered <- read_rds("Output/ak_sums_filtered.rds")
-
-check_dates <- wmt_sums_filtered %>% 
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_sums_filtered <- read_rds(paste0(file_path, "Output/wmt_sums_filtered.rds"))
+nn_sums_filtered <- read_rds(paste0(file_path, "Output/nn_sums_filtered.rds"))
+ak_sums_filtered <- read_rds(paste0(file_path, "Output/ak_sums_filtered.rds"))
+
+check_dates <- ak_sums_filtered %>% 
   group_by(home) %>% 
@@
 ```{r}
-wmt_sums_filtered <- read_rds("Output/wmt_sums_filtered.rds")
-nn_sums_filtered <- read_rds("Output/nn_sums_filtered.rds")
-ak_sums_filtered <- read_rds("Output/ak_sums_filtered.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_sums_filtered <- read_rds(paste0(file_path, "Output/wmt_sums_filtered.rds")) %>% 
+  group_by(home) %>% 
+  distinct(datetime_sums, .keep_all = TRUE) %>% 
+  ungroup()
+nn_sums_filtered <- read_rds(paste0(file_path, "Output/nn_sums_filtered.rds"))
+ak_sums_filtered <- read_rds(paste0(file_path, "Output/ak_sums_filtered.rds"))
 
@@
 
-# write_rds(all_sums_filtered, "Output/all_sums_filtered.rds")
+write_rds(all_sums_filtered, paste0(file_path, "Output/all_sums_filtered.rds"))
 ```
@@
 all_sums_filtered <- read_rds(paste0(file_path, "Output/all_sums_filtered.rds"))
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds")) %>% 
-  select(home_winter_id, area, treatment)
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) %>% 
+  select(home_winter_id, area, treatment_assigned, treatment_actual)
 
@@
   arrange(area, home, winter_id, datetime_sums) %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  right_join(kids_linked_ids, by = c("area", "home_winter_id"))
 

### 2020-06-24 child_health_data_cleaning.Rmd
check home char across winters

@@
   replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
+  # checked young ages with field data - correcting here
+  mutate(mom_age = if_else(mom_age == 14, 20, mom_age),
+         mom_age = if_else(mom_age == 10 | mom_age == 12, 21, mom_age)) %>% 
   select(home_winter_id:sibling_school)

### 2020-06-24 demographics_summary.Rmd
check home char across winters

@@
 
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
-demographics_initial <- read_rds(paste0(file_path, "Output/demographics_clean.rds")) 
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
 pm_clean <- read_rds(paste0(file_path, "Output/pm_clean.rds"))

### 2020-06-24 home_char_act_summary.Rmd
check home char across winters

@@
 author: "Ethan Walker"
-date: "Started 15 Jan 2020, Updated 25 March 2020"
+date: "Started 15 Jan 2020, Updated 23 June 2020"
 output:
@@
 library(knitr)
+library(forcats)
 jv_palette <- c("#330099","#CC0066","#FF6633", 
@@
 #pm_clean <- read_rds("Output/pm_clean.rds")
+```
+
+# Check Winter 1 vs Winter 2 data for home characteristics
+```{r}
+home_char_check_num <- home_char_clean %>% 
+  group_by(home) %>% 
+  arrange(home, winter_id) %>% 
+  pivot_longer(cols = c(residents_total:residents_under18, home_year_built:home_furry), 
+               names_to = "var", values_to = "value") %>% 
+  group_by(home, var) %>% 
+  arrange(home, var, winter_id) %>% 
+  mutate(data_check = if_else(value == lead(value), 1, 0),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check)) %>% 
+  filter(data_check == 0) %>% 
+  select(area, home, winter_id, var, value) %>% 
+  mutate(value = as.numeric(value),
+         data_check = if_else(var == "residents_total" & abs(value - lead(value)) > 1, 1, 0),
+         data_check = if_else(var == "residents_under5" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "residents_under18" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "home_year_built" & abs(value - lead(value)) > 10, 1, data_check),
+         data_check = if_else(var == "home_sqft" & abs(value - lead(value)) > 200, 1, data_check),
+         data_check = if_else(var == "home_floors" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "home_bedrooms" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "home_windows" & abs(value - lead(value)) > 5, 1, data_check),
+         data_check = if_else(var == "home_pets" & abs(value - lead(value)) > 2, 1, data_check),
+         data_check = if_else(var == "home_furry" & abs(value - lead(value)) > 2, 1, data_check),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check),
+         value = as.character(value)) %>% 
+  filter(data_check == 1) %>% 
+  select(area, home, winter_id, var, value)
+
+write_csv(home_char_check_num, paste0(file_path, "Output/home_char_check_num.csv"))
+
+home_char_check_fact <- home_char_clean %>% 
+  group_by(home) %>% 
+  arrange(home, winter_id) %>% 
+  pivot_longer(cols = c(home_type, stove_age, stove_cert, home_fireplace:home_insert), 
+               names_to = "var", values_to = "value") %>% 
+  group_by(home, var) %>% 
+  arrange(home, var, winter_id) %>% 
+  mutate(data_check = if_else(value == lead(value), 1, 0),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check)) %>% 
+  filter(data_check == 0) %>% 
+  select(area, home, winter_id, var, value) %>% 
+  mutate(var2 = var,
+         value2 = value) %>% 
+  pivot_wider(names_from = "var2", values_from = "value2") %>% 
+  ungroup() %>% 
+  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
+              "stove_cert", "home_type", "home_fireplace"), as.character) %>% 
+  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
+              "stove_cert", "home_type", "home_fireplace"), as.factor) %>% 
+  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
+              "stove_cert", "home_type", "home_fireplace"), as.numeric)
+
+home_char_check_fact$value_num <- 
+  coalesce(home_char_check_fact$home_insert, 
+           home_char_check_fact$stove_age, 
+           home_char_check_fact$home_furnace, 
+           home_char_check_fact$home_woodstove, 
+           home_char_check_fact$stove_cert, 
+           home_char_check_fact$home_type, 
+           home_char_check_fact$home_fireplace)
+
+home_char_check_fact2 <- home_char_check_fact %>% 
+  select(area, home, winter_id, var, value, value_num) %>% 
+  group_by(home, var) %>% 
+  mutate(data_check = if_else(abs(value_num - lead(value_num)) > 1, 1, 0),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check),
+         value = as.character(value))
+
+write_csv(home_char_check_fact2, paste0(file_path, "Output/home_char_check_fact.csv"))
+
+home_char_check <- rbind(home_char_check_num, home_char_check_fact2) %>% 
+  arrange(home, var, winter_id) %>% 
+  mutate(data_check = if_else(is.na(value_num), 1, data_check))
+
+write_csv(home_char_check, paste0(file_path, "Output/home_char_check.csv"))
 ```

### 2020-06-24 indoor_temp_rh_cleaning.Rmd
check home char across winters

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 15 Jan 2020"
+date: "Started 13 Jan 2020, Updated 23 June 2020"
 output: pdf_document
@@
 library(zoo)
+
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
 ```
@@
 # Load data
-temp_rh_initial_wmt <- read_csv("Input/WMT/wmt_indoor_temp_rh_20200113.csv")
+temp_rh_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/qtrack.csv"), 
+                                  na = c("NULL", ""))
 
@@
          temp_indoor_min = as.numeric(TempCMin),
-         rh_temp_comments = as.character(`REPLACE(REPLACE(Comments`),
          area = "WMT") %>% 
@@
          rh_indoor_max, rh_indoor_min, temp_indoor_current, 
-         temp_indoor_max, temp_indoor_min, rh_temp_comments) %>% 
+         temp_indoor_max, temp_indoor_min) %>% 
   arrange(home_winter_id)
@@
 # Save as RDS 
-# write_rds(wmt_temp_rh, "Output/wmt_temp_rh.rds")
+write_rds(wmt_temp_rh, paste0(file_path, "Output/wmt_temp_rh.rds"))
 
@@
 # Load data
-temp_rh_initial_nn <- read_xlsx("Input/NN/nn_indoor_temp_rh_20200113.xlsx")
+temp_rh_initial_nn <- read_csv(paste0(file_path, "Input/NN/qtrack.csv"), 
+                                  na = c("NULL", ""))
 
@@
          temp_indoor_min = as.numeric(TempCMin),
-         rh_temp_comments = as.character(IgnoreReason),
          area = "NN") %>% 
@@
 # Save as RDS 
-# write_rds(nn_temp_rh, "Output/nn_temp_rh.rds")
+write_rds(nn_temp_rh, paste0(file_path, "Output/nn_temp_rh.rds"))
 
@@
 # Load data
-temp_rh_initial_ak <- read_xlsx("Input/AK/ak_indoor_temp_rh_20200113.xlsx")
+temp_rh_initial_ak <- read_csv(paste0(file_path, "Input/AK/qtrack.csv"), 
+                                  na = c("NULL", ""))
 
@@
          temp_indoor_min = as.numeric(TempCMin),
-         rh_temp_comments = as.character(IgnoreReason),
          area = "AK") %>% 
@@
 # Save as RDS 
-# write_rds(ak_temp_rh, "Output/ak_temp_rh.rds")
+write_rds(ak_temp_rh, paste0(file_path, "Output/ak_temp_rh.rds"))
 ```
@@
 ```{r}
-wmt_temp_rh <- read_rds("Output/wmt_temp_rh.rds")
-nn_temp_rh <- read_rds("Output/nn_temp_rh.rds")
-ak_temp_rh <- read_rds("Output/ak_temp_rh.rds")
+wmt_temp_rh <- read_rds(paste0(file_path, "Output/wmt_temp_rh.rds"))
+nn_temp_rh <- read_rds(paste0(file_path, "Output/nn_temp_rh.rds"))
+ak_temp_rh <- read_rds(paste0(file_path, "Output/ak_temp_rh.rds"))
 
@@
 
-# write_rds(kids_indoor_temp_rh, "Output/kids_indoor_temp_rh.rds")
+write_rds(kids_indoor_temp_rh, paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
 ```
@@
 kids_indoor_temp_rh <- read_rds(paste0(file_path, "Output/kids_indoor_temp_rh.rds"))
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
 indoor_temp_rh_clean <- kids_indoor_temp_rh %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, by = c("area", "home_winter_id"))
+
+summary(indoor_temp_rh_clean)
 

### 2020-06-24 moisture_data_cleaning.Rmd
check home char across winters

--- 2020-01-14 moisture_data_cleaning.Rmd
@@
 author: "Ethan Walker"
-date: "Started 19 Nov 2019, Updated 14 Jan 2020"
+date: "Started 19 Nov 2019, Updated 23 June 2020"
 output: pdf_document
@@
 # Load data
-moisture_initial_wmt <- read_xlsx("Input/WMT/wmt_kids_moisture_20191122.xlsx")
+moisture_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/moisture.csv"), 
+                                  na = c("NULL", ""))
 
@@
          moisture_ave, moisture_split) %>% 
+  replace_with_na(replace = list(moisture_ave = "NaN")) %>% 
   arrange(home_winter_id, moisture_date)
@@
 # Save as RDS 
-# write_rds(wmt_moisture, "Output/wmt_moisture.rds")
+write_rds(wmt_moisture, paste0(file_path, "Output/wmt_moisture.rds"))
 
@@
 # Load data
-moisture_initial_nn <- read_xlsx("Input/NN/nn_moisture_20200113.xlsx")
+moisture_initial_nn <- read_csv(paste0(file_path, "Input/NN/moisture.csv"), 
+                                  na = c("NULL", ""))
 
@@
 # Save as RDS 
-# write_rds(nn_moisture, "Output/nn_moisture.rds")
+write_rds(nn_moisture, paste0(file_path, "Output/nn_moisture.rds"))
 
@@
 # Load data
-moisture_initial_ak <- read_xlsx("Input/AK/ak_moisture_20200113.xlsx")
+moisture_initial_ak <- read_csv(paste0(file_path, "Input/AK/moisture.csv"), 
+                                  na = c("NULL", ""))
 
@@
 # Save as RDS 
-# write_rds(ak_moisture, "Output/ak_moisture.rds")
+write_rds(ak_moisture, paste0(file_path, "Output/ak_moisture.rds"))
 ```
@@
 ```{r}
-wmt_moisture <- read_rds("Output/wmt_moisture.rds")
-nn_moisture <- read_rds("Output/nn_moisture.rds")
-ak_moisture <- read_rds("Output/ak_moisture.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_moisture <- read_rds(paste0(file_path, "Output/wmt_moisture.rds"))
+nn_moisture <- read_rds(paste0(file_path, "Output/nn_moisture.rds"))
+ak_moisture <- read_rds(paste0(file_path, "Output/ak_moisture.rds"))
 
@@
 
-# write_rds(kids_moisture, "Output/kids_moisture.rds")
+write_rds(kids_moisture, paste0(file_path, "Output/kids_moisture.rds"))
 ```
@@
 ```{r}
-kids_moisture <- read_rds("Output/kids_moisture.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_moisture <- read_rds(paste0(file_path, "Output/kids_moisture.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
 moisture_clean <- kids_moisture %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  mutate(area = as.factor(area)) %>% 
+  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  # lower range on instrument = 5%
+  # therefore, changing values < 5 to 5
+  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
+         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
+         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
+         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))
 
-# write_rds(moisture_clean, "Output/moisture_clean.rds")
+summary(moisture_clean)
+
+write_rds(moisture_clean, paste0(file_path, "Output/moisture_clean.rds"))
 ```

### 2020-06-24 stove_grade_usage_cleaning.Rmd
check home char across winters

@@
 # Load data
-stove_grades_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/wmt_stove_grades_20200113.csv"))
+stove_grades_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/stove.csv"), 
+                                  na = c("NULL", ""))
 
@@
          stove_grade = as.factor(Score1),
-         comments = as.character(`REPLACE(REPLACE(Notes`),
          area = "WMT") %>% 
-  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
+  select(home_id, home_winter_id, area, stove_grade) %>% 
   ungroup() %>% 
@@
 # Save as RDS
-# write_rds(wmt_stove_grades, "Output/wmt_stove_grades.rds")
+write_rds(wmt_stove_grades, paste0(file_path, "Output/wmt_stove_grades.rds"))
 
@@
 # Load data
-stove_grades_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/nn_stove_grades_20200113.xlsx"))
+stove_grades_initial_nn <- read_csv(paste0(file_path, "Input/NN/stove.csv"), 
+                                  na = c("NULL", ""))
 
@@
          stove_grade = as.factor(Score1),
-         comments = as.character(IgnoreReason),
          area = "NN") %>% 
-  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
+  select(home_id, home_winter_id, area, stove_grade) %>% 
   ungroup() %>% 
@@
 # Save as RDS
-# write_rds(nn_stove_grades, "Output/nn_stove_grades.rds")
+write_rds(nn_stove_grades, paste0(file_path, "Output/nn_stove_grades.rds"))
 
@@
 # Load data
-stove_grades_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/ak_stove_grades_20200113.xlsx"))
+stove_grades_initial_ak <- read_csv(paste0(file_path, "Input/AK/stove.csv"), 
+                                  na = c("NULL", ""))
 
@@
          stove_grade = as.factor(Score1),
-         comments = as.character(IgnoreReason),
          area = "AK") %>% 
-  select(home_id, home_winter_id, area, stove_grade, comments) %>% 
+  select(home_id, home_winter_id, area, stove_grade) %>% 
   ungroup() %>% 
@@
 ```{r}
-wmt_stove_grades <- read_rds("Output/wmt_stove_grades.rds")
-nn_stove_grades <- read_rds("Output/nn_stove_grades.rds")
-ak_stove_grades <- read_rds("Output/ak_stove_grades.rds")
+wmt_stove_grades <- read_rds(paste0(file_path, "Output/wmt_stove_grades.rds"))
+nn_stove_grades <- read_rds(paste0(file_path, "Output/nn_stove_grades.rds"))
+ak_stove_grades <- read_rds(paste0(file_path, "Output/ak_stove_grades.rds"))
 
@@
 
-# write_rds(kids_stove_grades, "Output/kids_stove_grades.rds")
+write_rds(kids_stove_grades, paste0(file_path, "Output/kids_stove_grades.rds"))
 ```
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 kids_stove_grades <- read_rds(paste0(file_path, "Output/kids_stove_grades.rds"))
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
-file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
 stove_grades_clean <- kids_stove_grades %>% 
+  mutate(area = as.factor(area)) %>% 
   filter(home_id != 0) %>% 
+  # Mark gave a couple of "D" grades when there was not enough info to score the stove
+  # removing those "D" grades from final dataset
   filter(stove_grade != "D") %>% 
-  rename(stove_grade_comments = comments) %>% 
   select(-home_id) %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
   arrange(area, home, winter_id)
+
+summary(stove_grades_clean)
+
+# Some homes/stoves were graded during both winters, typically when new photos were added
+# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
+# We can't make the assumption that the stove was exactly the same for both winters,
+# so it will be best to use the grade corresponding to each winter for analysis.
 
@@
 
-# Dataset showing home with no stove grade
+# Dataset showing homes with no stove grade
 stove_grades_na <- stove_grades_clean %>% 
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 ##### WMT #####
 # Load data
-stove_use_initial_wmt <- read_csv("Input/WMT/wmt_stove_use_20200113.csv")
+stove_use_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/woodstoveusage.csv"), 
+                                  na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(wmt_stove_use, "Output/wmt_stove_use.rds")
+write_rds(wmt_stove_use, paste0(file_path, "Output/wmt_stove_use.rds"))
 
@@
 # Load data
-stove_use_initial_nn <- read_xlsx("Input/NN/nn_stove_use_20200113.xlsx")
+stove_use_initial_nn <- read_csv(paste0(file_path, "Input/NN/woodstoveusage.csv"), 
+                                  na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(nn_stove_use, "Output/nn_stove_use.rds")
+write_rds(nn_stove_use, paste0(file_path, "Output/nn_stove_use.rds"))
 
@@
 # Load data
-stove_use_initial_ak <- read_xlsx("Input/AK/ak_stove_use_20200113.xlsx")
+stove_use_initial_ak <- read_csv(paste0(file_path, "Input/AK/woodstoveusage.csv"), 
+                                  na = c("NULL", ""))
 
@@
 # Save as RDS
-# write_rds(ak_stove_use, "Output/ak_stove_use.rds")
+write_rds(ak_stove_use, paste0(file_path, "Output/ak_stove_use.rds"))
 ```
@@
 ```{r}
-wmt_stove_use <- read_rds("Output/wmt_stove_use.rds")
-nn_stove_use <- read_rds("Output/nn_stove_use.rds")
-ak_stove_use <- read_rds("Output/ak_stove_use.rds")
+wmt_stove_use <- read_rds(paste0(file_path, "Output/wmt_stove_use.rds"))
+nn_stove_use <- read_rds(paste0(file_path, "Output/nn_stove_use.rds"))
+ak_stove_use <- read_rds(paste0(file_path, "Output/ak_stove_use.rds"))
 
@@
 
-# write_rds(kids_stove_use, "Output/kids_stove_use.rds")
+write_rds(kids_stove_use, paste0(file_path, "Output/kids_stove_use.rds"))
 ```
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 kids_stove_use <- read_rds(paste0(file_path, "Output/kids_stove_use.rds"))
-kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))
-file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
 
 stove_use_clean <- kids_stove_use %>% 
+  mutate(area = as.factor(area)) %>% 
   replace_with_na(replace = list(wood_cords = -999)) %>%
@@
                                           "> 1 yr"))) %>% 
-  left_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  right_join(kids_linked_ids, by = c("area", "home_winter_id"))
+
+summary(stove_use_clean)
 

### 2020-06-26 child_health_data_cleaning.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 27 Feb 2020, Updated 22 June 2020"
+date: "Started 27 Feb 2020, Updated 26 June 2020"
 output: pdf_document 
@@
   mutate(area = as.factor(area)) %>% 
-  right_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
   filter(!is.na(child_id_num))
@@
   mutate(area = as.factor(area)) %>% 
-  right_join(kids_linked_ids, c("area", "home_winter_id"))
+  left_join(kids_linked_ids, c("area", "home_winter_id"))
 
@@
   mutate(area = as.factor(area)) %>% 
-  right_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
+  left_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
   mutate_if(is.character, as.factor)

### 2020-06-26 home_char_act_cleaning.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 22 June 2020"
+date: "Started 13 Jan 2020, Updated 26 June 2020"
 output: pdf_document
@@
 # Load data
-home_char_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/homechar.csv"), 
+home_char_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/homechar_updated_data_SAVE.csv"), 
                                   na = c("NULL", ""))
@@
 # Load data
-home_char_initial_nn <- read_csv(paste0(file_path, "Input/NN/homechar.csv"), na = c("NULL", ""))
+home_char_initial_nn <- read_csv(paste0(file_path, "Input/NN/homechar_updated_data_SAVE.csv"), na = c("NULL", ""))
 
@@
 # Load data
-home_char_initial_ak <- read_csv(paste0(file_path, "Input/AK/homechar.csv"), na = c("NULL", ""))
+home_char_initial_ak <- read_csv(paste0(file_path, "Input/AK/homechar_updated_data_SAVE.csv"), na = c("NULL", ""))
 
@@
 
-# write_rds(kids_home_char, paste0(file_path, "Output/kids_home_char.rds"))
+write_rds(kids_home_char, paste0(file_path, "Output/kids_home_char.rds"))
 ```

### 2020-06-26 home_char_act_summary.Rmd
data cleaning/merging updates

--- 2020-05-14 home_char_act_summary.Rmd
@@
 #pm_clean <- read_rds("Output/pm_clean.rds")
+```
+
+# Check Winter 1 vs Winter 2 data for home characteristics
+```{r}
+home_char_check_num <- home_char_clean %>% 
+  group_by(home) %>% 
+  arrange(home, winter_id) %>% 
+  pivot_longer(cols = c(residents_total:residents_under18, home_year_built:home_furry), 
+               names_to = "var", values_to = "value") %>% 
+  group_by(home, var) %>% 
+  arrange(home, var, winter_id) %>% 
+  mutate(data_check = if_else(value == lead(value), 1, 0),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check)) %>% 
+  filter(data_check == 0) %>% 
+  select(area, home, home_winter_id, winter_id, var, value) %>% 
+  mutate(value = as.numeric(value),
+         data_check = if_else(var == "residents_total" & abs(value - lead(value)) > 1, 1, 0),
+         data_check = if_else(var == "residents_under5" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "residents_under18" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "home_year_built" & abs(value - lead(value)) > 10, 1, data_check),
+         data_check = if_else(var == "home_sqft" & abs(value - lead(value)) > 200, 1, data_check),
+         data_check = if_else(var == "home_floors" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "home_bedrooms" & abs(value - lead(value)) > 1, 1, data_check),
+         data_check = if_else(var == "home_windows" & abs(value - lead(value)) > 5, 1, data_check),
+         data_check = if_else(var == "home_pets" & abs(value - lead(value)) > 2, 1, data_check),
+         data_check = if_else(var == "home_furry" & abs(value - lead(value)) > 2, 1, data_check),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check),
+         value = as.character(value)) %>% 
+  filter(data_check == 1) %>% 
+  select(area, home, home_winter_id, winter_id, var, value)
+
+write_csv(home_char_check_num, paste0(file_path, "Output/home_char_check_num.csv"))
+
+home_char_check_fact <- home_char_clean %>% 
+  group_by(home) %>% 
+  arrange(home, winter_id) %>% 
+  pivot_longer(cols = c(home_type, stove_age, stove_cert, home_fireplace:home_insert), 
+               names_to = "var", values_to = "value") %>% 
+  group_by(home, var) %>% 
+  arrange(home, var, winter_id) %>% 
+  mutate(data_check = if_else(value == lead(value), 1, 0),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check)) %>% 
+  filter(data_check == 0) %>% 
+  select(area, home, home_winter_id, winter_id, var, value) %>% 
+  mutate(var2 = var,
+         value2 = value) %>% 
+  pivot_wider(names_from = "var2", values_from = "value2") %>% 
+  ungroup() %>% 
+  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
+              "stove_cert", "home_type", "home_fireplace"), as.character) %>% 
+  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
+              "stove_cert", "home_type", "home_fireplace"), as.factor) %>% 
+  mutate_at(c("home_insert", "stove_age", "home_furnace", "home_woodstove", 
+              "stove_cert", "home_type", "home_fireplace"), as.numeric)
+
+home_char_check_fact$value_num <- 
+  coalesce(home_char_check_fact$home_insert, 
+           home_char_check_fact$stove_age, 
+           home_char_check_fact$home_furnace, 
+           home_char_check_fact$home_woodstove, 
+           home_char_check_fact$stove_cert, 
+           home_char_check_fact$home_type, 
+           home_char_check_fact$home_fireplace)
+
+home_char_check_fact2 <- home_char_check_fact %>% 
+  select(area, home,home_winter_id, winter_id, var, value, value_num) %>% 
+  group_by(home, var) %>% 
+  mutate(data_check = if_else(abs(value_num - lead(value_num)) > 1, 1, 0),
+         data_check = if_else(is.na(data_check), lag(data_check), data_check),
+         value = as.character(value))
+
+write_csv(home_char_check_fact2, paste0(file_path, "Output/home_char_check_fact.csv"))
+
+home_char_check <- rbind(home_char_check_num, home_char_check_fact2) %>% 
+  arrange(home, var, winter_id) %>% 
+  mutate(data_check = if_else(is.na(value_num), 1, data_check))
+
+write_csv(home_char_check, paste0(file_path, "Output/home_char_check.csv"))
 ```

### 2020-06-26 manuscript_tables_figures.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 13 May 2020, Updated 22 May 2020"
+date: "Started 13 May 2020, Updated 26 June 2020"
 output:
@@
 
+health_data_merged <- 
+  read_rds(paste0(file_path, "Output/health_data_merged.rds"))
+
 kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
@@
 ```{r}
-# number of placebo households, total and by area
-placebo_homes <- health_exposure_data_sampling_day %>% 
-  filter(winter_id == 1) %>% 
-  filter(area == "WMT") %>% 
-  select(area, home, home_id, winter_id, home_winter_id, treatment)
-  #filter(treatment == "Placebo") %>% 
-  distinct(home, .keep_all = TRUE) %>%
-  count(area)
-placebo_homes
+# number of households, total and by area
+study_homes <- health_exposure_data_sampling_day %>% 
+  distinct(home, winter_id, .keep_all = TRUE) %>%
+  count(area, winter_id, treatment_assigned)
+study_homes
 
-# number of placebo households, total and by area
-treatment_homes <- health_exposure_data_sampling_day %>% 
-  filter(winter_id == 2) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  group_by(area) %>% 
-  count(treatment)
-treatment_homes
+study_kids <- health_exposure_data_sampling_day %>% 
+  filter(!is.na(child_id_num)) %>% 
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>%
+  count(area, winter_id, treatment_assigned)
+study_kids
 
-# number of placebo households, total and by area
-treatment_kids <- health_exposure_data_sampling_day %>% 
-  filter(winter_id == 2) %>% 
-  distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  filter(!is.na(child_id_num)) %>% 
-  group_by(area) %>% 
-  count(treatment)
-treatment_kids
+nn_homes <- health_exposure_data_sampling_day %>% 
+  filter(home == "CH209" | home == "CH218")
 ```

### 2020-06-26 merge_datasets.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 24 March 2020, Updated 14 April 2020"
+date: "Started 24 March 2020, Updated 26 June 2020"
 output: pdf_document
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
 
 pm_clean <- 
-  read_rds(paste0(output_path, "Output/pm_clean.rds")) 
+  read_rds(paste0(file_path, "Output/pm_clean.rds")) 
 sums_clean <- 
-  read_rds(paste0(output_path, "Output/sums_clean.rds"))
+  read_rds(paste0(file_path, "Output/sums_clean.rds"))
 
-kids_linked_ids <- 
-  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
-demographics_clean <- 
-  read_rds(paste0(output_path, "Output/demographics_clean.rds")) 
 moisture_clean <- 
-  read_rds(paste0(output_path, "Output/moisture_clean.rds"))
+  read_rds(paste0(file_path, "Output/moisture_clean.rds"))
 indoor_temp_rh_clean <- 
-  read_rds(paste0(output_path, "Output/indoor_temp_rh_clean.rds"))
+  read_rds(paste0(file_path, "Output/indoor_temp_rh_clean.rds"))
 stove_use_clean <- 
-  read_rds(paste0(output_path, "Output/stove_use_clean.rds"))
+  read_rds(paste0(file_path, "Output/stove_use_clean.rds"))
 stove_grades_clean <- 
-  read_rds(paste0(output_path, "Output/stove_grades_clean.rds")) 
+  read_rds(paste0(file_path, "Output/stove_grades_clean.rds")) 
 home_char_clean <- 
-  read_rds(paste0(output_path, "Output/home_char_clean.rds"))
+  read_rds(paste0(file_path, "Output/home_char_clean.rds"))
 home_act_clean <- 
-  read_rds(paste0(output_path, "Output/home_act_clean.rds"))
+  read_rds(paste0(file_path, "Output/home_act_clean.rds"))
 ```
@@
           desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, dtid, day_of_week, 
-         pm_total_observations, pm_mean_sampling_period, sampling_day, 
-         pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         dtid, day_of_week, pm_total_observations, pm_mean_sampling_period, 
+         sampling_day, pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
          sample_date, pm_mean_daily, pm_daily_obs, pm_spike_1000,
@@
 
-#write_rds(pm_join, paste0(output_path, "Output/pm_join.rds"))
-pm_join <- read_rds(paste0(output_path, "Output/pm_join.rds"))
+#write_rds(pm_join, paste0(file_path, "Output/pm_join.rds"))
+pm_join <- read_rds(paste0(file_path, "Output/pm_join.rds"))
 
@@
   mutate(sample_date = as.character(date_sums)) %>% 
-  select(area, home, home_winter_id, winter_id, ibid, treatment, 
-         datetime_sums, sample_date, temp_c_sums) %>% 
+  select(area, home, home_winter_id, winter_id, ibid, treatment_assigned, 
+         treatment_actual, datetime_sums, sample_date, temp_c_sums) %>% 
   group_by(area, home, winter_id) %>% 
@@
 
-#write_rds(sums_join, paste0(output_path, "Output/sums_join.rds"))
-sums_join <- read_rds(paste0(output_path, "Output/sums_join.rds"))
+#write_rds(sums_join, paste0(file_path, "Output/sums_join.rds"))
+sums_join <- read_rds(paste0(file_path, "Output/sums_join.rds"))
 
@@
 pm_sums_joined1 <- sums_join %>% 
-  left_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
+  left_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", 
+                            "treatment_assigned", "treatment_actual")) %>% 
   arrange(area, home, datetime_sums) %>% 
@@
   arrange(area, home, sampling_day) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
-         sums_mean_sampling_day, sums_mean_sampling_period)
+  select(area, home, home_winter_id, winter_id, sampling_day, 
+         sums_mean_sampling_day, sums_mean_sampling_period,
+         treatment_assigned, treatment_actual)
 
-#write_rds(pm_sums_joined1, paste0(output_path, "Output/pm_sums_joined1.rds"))
-pm_sums_joined1 <- read_rds(paste0(output_path, "Output/pm_sums_joined1.rds"))
+#write_rds(pm_sums_joined1, paste0(file_path, "Output/pm_sums_joined1.rds"))
+pm_sums_joined1 <- read_rds(paste0(file_path, "Output/pm_sums_joined1.rds"))
 
@@
   left_join(pm_sums_joined1, 
-            by = c("area", "home", "home_winter_id", "winter_id", "treatment", "sampling_day"))
+            by = c("area", "home", "home_winter_id", "winter_id", 
+                   "treatment_assigned", "treatment_actual", "sampling_day"))
 
@@
   full_join(pm_sums_joined2, by = c("area", "home", "home_winter_id", "winter_id", 
-                                    "treatment", "sample_date")) %>% 
+                                    "treatment_assigned", "treatment_actual", 
+                                    "sample_date")) %>% 
   mutate(sampling_day = as.character(sampling_day)) %>% 
@@
   mutate(sampling_day = home_act_day) %>% 
-  mutate(winter_id = as.character(winter_id)) %>% 
-  select(-home_act_day, -home_act_comments, -home_id) %>% 
+  select(-home_act_day, -home_id) %>% 
   full_join(pm_sums_joined3, by = c("area", "home", "home_winter_id", 
-                                    "winter_id", "treatment", "sampling_day")) %>% 
+                                    "winter_id", "treatment_assigned", 
+                                    "treatment_actual", "sampling_day")) %>% 
   ungroup() %>% 
@@
 
-analysis_data2 <- demographics_clean %>% 
+analysis_data2 <- demographics %>% 
   mutate(residents_under_five = if_else(residents_under_five == 0 & age_child_nov != 4, "NA", residents_under_five),
@@
   full_join(analysis_data1, by = c("area", "home", "home_winter_id", 
-                                    "winter_id", "treatment")) %>% 
+                                    "winter_id", "treatment_assigned", 
+                                    "treatment_actual")) %>% 
   filter(!is.na(home_winter_id)) %>% 
@@
   filter(winter_id == 1 | winter_id == 2) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, residents_smoke:home_insert) %>% 
-  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, 
+         treatment_actual, residents_smoke:home_insert) %>% 
+  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", 
+                                    "treatment_assigned", "treatment_actual")) %>% 
   mutate(home_type_2level = fct_collapse(home_type,
@@
   mutate(winter_id = as.character(winter_id)) %>%
-  select(area, home, home_winter_id, winter_id, treatment, rh_indoor_current:temp_indoor_min) %>% 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         rh_indoor_current:temp_indoor_min) %>% 
   ungroup() %>% 
-  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
+  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", 
+                                    "treatment_assigned", "treatment_actual")) %>% 
   mutate(temp_max_5 = temp_indoor_max/5,
@@
   filter(!is.na(moisture_ave)) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, moisture_ave, 
+  select(area, home, home_winter_id, winter_id, treatment_actual, 
+         treatment_assigned, moisture_ave, 
          moisture_split, moisture_date) %>% 
@@
   right_join(analysis_data4, by = c("area", "home", "home_winter_id", 
-                                    "winter_id", "treatment")) %>% 
+                                    "winter_id", "treatment_assigned", 
+                                    "treatment_actual")) %>% 
   ungroup() %>% 
@@
   mutate(winter_id = as.character(winter_id)) %>%
-  select(area, home, home_winter_id, winter_id, treatment, stove_grade) %>% 
+  select(area, home, home_winter_id, winter_id, treatment_actual, 
+         treatment_assigned, stove_grade) %>% 
   distinct(home, .keep_all = TRUE) %>% 
   right_join(analysis_data5, by = c("area", "home", "home_winter_id", 
-                                    "winter_id", "treatment")) %>% 
+                                    "winter_id", "treatment_assigned", 
+                                    "treatment_actual")) %>% 
   mutate(stove_grade_3level = fct_collapse(stove_grade,
@@
   right_join(analysis_data6, by = c("area", "home", "home_winter_id", 
-                                    "winter_id", "treatment")) %>% 
+                                    "winter_id", "treatment_assigned", 
+                                    "treatment_actual")) %>% 
   mutate(burn_level_3level = fct_collapse(burn_level,
@@
 
-write_rds(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.rds"))
-write_csv(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.csv"))
+write_rds(analysis_data7, paste0(file_path, "Output/exposure_analysis_data_long.rds"))
+write_csv(analysis_data7, 
+          paste0(file_path, "Output/exposure_analysis_data_long.csv"), na = "")
 
 exposure_analysis_data_long <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long.rds"))
 
@@
 
-#write_csv(save_var_names, paste0(output_path, "Output/kids_data_var_names.csv"))
+#write_csv(save_var_names, paste0(file_path, "Output/kids_data_var_names.csv"))
 ```
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 exposure_analysis_data_long <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long.rds"))
 
 kids_amb_pm_data <- 
-  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_amb_pm_data_new.rds")) %>% 
   mutate(sample_date = amb_pm_date)
@@
 kids_amb_temp_data <- 
-  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_amb_temp_data_new.rds")) %>% 
   mutate(sample_date = temp_date) %>% 
@@
 
-write_rds(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
-write_csv(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.csv"))
+write_rds(exposure_analysis_data_long_new, paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
+write_csv(exposure_analysis_data_long_new, 
+          paste0(file_path, "Output/exposure_analysis_data_long_new.csv"), na = "")
 ```
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 exposure_analysis_data_long_new <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
 
 exposure_analysis_data_medium <- exposure_analysis_data_long_new %>% 
-  select(area, home, home_winter_id, winter_id, treatment, 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
          sampling_day, sample_date, burn_level:mean_temp_roll_4day) %>% 
@@
   distinct(sampling_day, .keep_all = TRUE) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, sampling_day,
-         child_id_char, burn_level:mean_temp_roll_4day) %>% 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         sampling_day, child_id_char, burn_level:mean_temp_roll_4day) %>% 
   mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
@@
 
-write_rds(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
-write_csv(exposure_analysis_data_medium, paste0(output_path, "Output/exposure_analysis_data_medium.csv"))
+write_rds(exposure_analysis_data_medium, paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
+write_csv(exposure_analysis_data_medium, 
+          paste0(file_path, "Output/exposure_analysis_data_medium.csv"), na = "")
 ```
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 exposure_analysis_data_medium <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
 exposure_analysis_data_long <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds")) 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds")) 
 sums_rolling_pm_data <- 
-  read_rds(paste0(output_path, "Output/sums_rolling_pm_data.rds"))
+  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))
 
@@
 sums_temp_change <- sums_rolling_pm_data %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1) %>% 
   group_by(area, home, sampling_day) %>% 
@@
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
+  #filter(treatment == "Placebo") %>% 
   #filter(area == "WMT") %>%
@@
 
-write_rds(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.rds"))
-write_csv(exposure_analysis_data_short, paste0(output_path, "Output/exposure_analysis_data_short.csv"))
+write_rds(exposure_analysis_data_short, paste0(file_path, "Output/exposure_analysis_data_short.rds"))
+write_csv(exposure_analysis_data_short, 
+          paste0(file_path, "Output/exposure_analysis_data_short.csv"), na = "")
 ```
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 pm_clean <- 
-  read_rds(paste0(output_path, "Output/pm_clean.rds"))
+  read_rds(paste0(file_path, "Output/pm_clean.rds"))
 sums_clean <- 
-  read_rds(paste0(output_path, "Output/sums_clean.rds"))
+  read_rds(paste0(file_path, "Output/sums_clean.rds"))
 
@@
   ungroup() %>% 
-  select(-home_id, -treatment) %>% 
+  select(-home_id, -treatment_assigned, - treatment_actual) %>% 
   arrange(area, home, sample_datetime)
@@
 exposure_analysis_data_long_new <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
 
@@
             by = c("area", "home", "home_winter_id", "winter_id", "sample_date")) %>% 
-  select(area, home, home_winter_id, winter_id, treatment, sampling_day,
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, sampling_day,
          child_id_char, sample_date, datetime_sums, temp_c_sums:wood_collect_method_2level) %>% 
@@
 
-write_rds(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.rds"))
-write_csv(pm_sums_joined3, paste0(output_path, "Output/sums_rolling_pm_data.csv"))
+write_rds(pm_sums_joined3, paste0(file_path, "Output/sums_rolling_pm_data.rds"))
+write_csv(pm_sums_joined3, 
+          paste0(file_path, "Output/sums_rolling_pm_data.csv"), na = "")
 ```

### 2020-06-26 merge_health_datasets.Rmd
data cleaning/merging updates

@@
 author: "Ethan Walker"
-date: "Started 27 March 2020, Updated 14 May 2020"
+date: "Started 27 March 2020, Updated 26 June 2020"
 output: pdf_document
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 at_home_pm_merged <- 
-  read_rds(paste0(output_path, "Output/at_home_pm_merged.rds"))
+  read_rds(paste0(file_path, "Output/at_home_pm_merged.rds"))
 kids_med_visit <- 
-  read_rds(paste0(output_path, "Output/kids_med_visit.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_med_visit.rds")) %>% 
   mutate(visit_date = med_visit_date) 
 kids_birth_infancy <- 
-  read_rds(paste0(output_path, "Output/kids_birth_infancy.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_birth_infancy.rds")) %>% 
   select(-home_id)
-child_home_winter <- 
-  read_rds(paste0(output_path, "Output/child_home_winter.rds")) %>% 
-  select(-home_id.x, -home_id.y, -chw_symptom_free_date)
 kids_csq <- 
-  read_rds(paste0(output_path, "Output/kids_csq.rds"))
+  read_rds(paste0(file_path, "Output/kids_csq.rds")) 
 kids_lrti <- 
-  read_rds(paste0(output_path, "Output/kids_lrti.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_lrti.rds")) %>% 
   select(-home_id) %>% 
@@
 kids_medical_records <- 
-  read_rds(paste0(output_path, "Output/kids_medical_records.rds")) %>% 
+  read_rds(paste0(file_path, "Output/kids_medical_records.rds")) %>% 
   select(-visit_date)
@@
   mutate(visit_date = csq_date) %>% 
-  full_join(kids_med_visit, by = c("area", "home", "home_winter_id", "treatment",
+  full_join(kids_med_visit, by = c("area", "home", "home_winter_id", 
+                                   "treatment_assigned", "treatment_actual",
                                    "winter_id", "child_id_num", "visit_date")) %>% 
   select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         treatment, csq_date:med_visit_mean_temp) %>% 
+         treatment_assigned, treatment_actual, csq_date:med_visit_mean_temp) %>% 
   arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
@@
 
+#health_merged2 <- health_merged1 %>% 
+#  left_join(child_home_winter, by = c("area", "home", "home_winter_id", 
+#                                      "treatment_assigned", "treatment_actual",
+#                                      "winter_id", "child_id_num")) %>% 
+#  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
+#  ungroup()
+
 health_merged2 <- health_merged1 %>% 
-  left_join(child_home_winter, by = c("area", "home", "home_winter_id", "treatment",
-                                      "winter_id", "child_id_num")) %>% 
-  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
-  ungroup()
-
-health_merged3 <- health_merged2 %>% 
-  left_join(kids_birth_infancy, by = c("area", "home", "home_winter_id", "treatment",
+  left_join(kids_birth_infancy, by = c("area", "home", "home_winter_id", 
+                                       "treatment_assigned", "treatment_actual",
                                        "winter_id", "child_id_num")) %>% 
@@
 
-health_merged4 <- health_merged3 %>% 
+health_merged3 <- health_merged2 %>% 
   full_join(at_home_pm_merged, by = c("area", "home", "home_winter_id", "winter_id", 
-                                      "child_id_num", "sample_date", "treatment")) %>% 
+                                      "child_id_num", "sample_date", 
+                                      "treatment_assigned", "treatment_actual")) %>% 
   arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
   select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         treatment, sample_date, sampling_day, at_home_location, 
+         treatment_assigned, treatment_actual, 
+         sample_date, sampling_day, at_home_location, 
          pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:at_home) %>% 
-  ungroup()
+  ungroup() %>% 
+  filter(!is.na(home))
 
@@
 
-health_merged5 <- health_merged4 %>% 
+health_merged4 <- health_merged3 %>% 
   left_join(kids_lrti, by = c("area", "home", "home_winter_id", "winter_id", 
-                                      "child_id_num", "treatment")) %>% 
+                              "child_id_num", "treatment_assigned", "treatment_actual")) %>% 
   arrange(area, home, home_winter_id, winter_id, child_id_num, sample_date) %>% 
   select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         treatment, sample_date, sampling_day, at_home_location, 
+         treatment_assigned, treatment_actual, sample_date, sampling_day, at_home_location, 
          pm_at_home_sampling_day, pm_at_home_sampling_period, csq_date:lrti_professional_comments) %>% 
@@
 
-write_rds(health_merged5, paste0(output_path, "Output/health_data_merged.rds"))
-write_csv(health_merged5, paste0(output_path, "Output/health_data_merged.csv"))
+write_rds(health_merged3, paste0(file_path, "Output/health_data_merged.rds"))
+write_csv(health_merged3, paste0(file_path, "Output/health_data_merged.csv"), na = "")
 ```
@@
 ```{r}
-output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
 exposure_analysis_data_medium <- 
-  read_rds(paste0(output_path, "Output/exposure_analysis_data_medium.rds"))
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
 
 health_data_merged <- 
-  read_rds(paste0(output_path, "Output/health_data_merged.rds"))
+  read_rds(paste0(file_path, "Output/health_data_merged.rds"))
 
 health_exposure_data_sampling_day <- health_data_merged %>% 
-  select(-pm_mean_sampling_day, -pm_mean_sampling_period, day_of_week) %>% 
+  select(-pm_mean_sampling_day, -pm_mean_sampling_period) %>% 
   mutate(winter_id = as.character(winter_id)) %>% 
-  right_join(exposure_analysis_data_medium,
-            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
-                   "child_id_num", "sampling_day"))
+  full_join(exposure_analysis_data_medium,
+            by = c("area", "child_id_num", "home", "home_winter_id", "winter_id", 
+                   "treatment_assigned", "treatment_actual",
+                   "sampling_day")) %>% 
+  filter(!is.na(home))
 
-write_rds(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
-write_csv(health_exposure_data_sampling_day, paste0(output_path, "Output/health_exposure_data_sampling_day.csv"))
+write_rds(health_exposure_data_sampling_day, 
+          paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+write_csv(health_exposure_data_sampling_day, 
+          paste0(file_path, "Output/health_exposure_data_sampling_day.csv"), na = "")
 ```
@@
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(output_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 
@@
 
-#write_csv(save_var_names, paste0(output_path, "Output/kids_data_var_names.csv"))
+#write_csv(save_var_names, paste0(file_path, "Output/kids_data_var_names.csv"), na = "")
 ```

### 2020-06-29 manuscript_tables_figures.Rmd
update data cleaning

@@
 author: "Ethan Walker"
-date: "Started 13 May 2020, Updated 26 June 2020"
+date: "Started 13 May 2020, Updated 29 June 2020"
 output:
@@
 
+# Count number of kids initially randomized
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+ak_demographics <- read_rds(paste0(file_path, "Output/ak_demographic_data.rds"))
+nn_demographics <- read_rds(paste0(file_path, "Output/nn_demographic_data.rds"))
+wmt_demographics <- read_rds(paste0(file_path, "Output/wmt_demographic_data.rds"))
+
+kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics) %>% 
+  filter(!is.na(treatment)) %>% 
+  filter(!is.na(child_id_num)) %>% 
+  count(area, winter_id, treatment) %>% 
+  filter(winter_id != 2)
+```
+
+# Count number of households initially randomized
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+ak_ids <- read_rds(paste0(file_path, "Output/ak_ids_linked.rds"))
+nn_ids <- read_rds(paste0(file_path, "Output/nn_ids_linked.rds"))
+wmt_ids <- read_rds(paste0(file_path, "Output/wmt_ids_linked.rds"))
+
+kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
+  filter(!is.na(treatment)) %>% 
+  ### WMA222 and WMA 233 had a treatment but no kids - not counting in flow chart
+  filter(home != "WMA222" & home != "WMA233") %>%
+  count(area, winter_id, treatment) %>% 
+  filter(winter_id != 2)
+```
+
+# Count number of households and kids that participated by area and winter
 ```{r}
@@
   count(area, winter_id, treatment_assigned)
-study_homes
 
 study_kids <- health_exposure_data_sampling_day %>% 
-  filter(!is.na(child_id_num)) %>% 
+  select(area, home, child_id_char, child_id_num, winter_id, treatment_assigned) %>% 
+  filter(!is.na(child_id_char)) %>% 
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
   count(area, winter_id, treatment_assigned)
-study_kids
-
-nn_homes <- health_exposure_data_sampling_day %>% 
-  filter(home == "CH209" | home == "CH218")
 ```

### 2020-06-29 merge_health_datasets.Rmd
update data cleaning

@@
 
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
+  select(child_id_num, child_id_char, area, winter_id)
+
 health_exposure_data_sampling_day <- health_data_merged %>% 
@@
                    "sampling_day")) %>% 
-  filter(!is.na(home))
+  filter(!is.na(home)) %>% 
+  filter(!is.na(child_id_num)) %>% 
+  select(-child_id_char) %>% 
+  left_join(demographics, by = c("area", "winter_id", "child_id_num")) %>% 
+  select(area:child_id_num, child_id_char, visit_date:rh_max_5)
 

### 2020-06-30 collinearity_exposure_analysis.Rmd
exposure analysis updates

@@
 author: "Ethan Walker"
-date: "Started 16 April 2020, Updated 16 April 2020"
+date: "Started 16 April 2020, Updated 30 June 2020"
 output: pdf_document
@@
 ####### Format full data to use for analysis
-analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
-  filter(home != "WMA300") %>% 
-  group_by(area, home) %>% 
+analysis_data <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo" & area == "WMT") %>% 
+  arrange(area, home, sampling_day) %>% 
   distinct(home, .keep_all = TRUE) %>% 
@@
 pm_mean_sampling_period <- analysis_data %>% 
-  select(pm_mean_sampling_period, 
-         income_3level,
+  select(pm_mean_sampling_period, income_3level, 
          residents_smoke,
-         home_floors_2level, 
-         home_sqft_2level,
+         home_floors_2level, home_sqft_2level,
          home_pets_2level,
-         stove_age_3level, 
-         chimney_clean_3level, 
+         stove_age_3level, chimney_clean_3level, 
          wood_collect_2level,
-         moisture_closest_2level,
          sums_events_sampling_period,
          mean_temp_sampling_period_5,
+         amb_pm_sampling_period_5,
          stove_grade_3level,
@@
 other_vars <- analysis_data %>% 
-  select(pm_mean_sampling_period, 
-         income_3level,
+  select(pm_mean_sampling_period, income_3level, 
          residents_smoke,
-         home_floors_2level, 
-         home_sqft_2level,
+         home_floors_2level, home_sqft_2level,
          home_pets_2level,
-         stove_age_3level, 
-         chimney_clean_3level, 
+         stove_age_3level, chimney_clean_3level, 
          wood_collect_2level,
-         moisture_closest_2level,
          sums_events_sampling_period,
          mean_temp_sampling_period_5,
+         amb_pm_sampling_period_5,
          stove_grade_3level,
@@
          wood_collect_2level,
-         moisture_closest_2level,
          sums_events_sampling_period,

### 2020-06-30 exposure_covariate_associations.Rmd
exposure analysis updates

@@
 author: "Ethan Walker"
-date: "Started 2 Feb 2020, Updated 16 April 2020"
+date: "Started 2 Feb 2020, Updated 30 June 2020"
 output: pdf_document
@@
 
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+
 exposure_analysis_data_long <- 
@@
 ####### Format full data to use for analysis
-analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(home != "WMA300") %>% 
-  group_by(area, home) %>% 
+analysis_data <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  arrange(area, home, sampling_day) %>% 
   distinct(home, .keep_all = TRUE) %>% 
@@
   analysis_data <- dataset %>% 
-    filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
+    filter(winter_id == 1 & treatment_assigned == "Placebo" & area == "WMT") %>% 
     group_by(area, home) %>% 
@@
 analysis_data <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
   group_by(area, home) %>% 

### 2020-06-30 exposure_models.Rmd
exposure analysis updates

@@
 author: "Ethan Walker"
-date: "Started 19 March 2020, Updated 30 April 2020"
+date: "Started 19 March 2020, Updated 30 June 2020"
 output: pdf_document
@@
 selected_vars <- analysis_data %>% 
-  #filter(area == "NN") %>% 
-  select(pm_mean_sampling_period,
-         income_3level,
+  #filter(area == "WMT") %>% 
+  select(pm_mean_sampling_period, income_3level, area, home,
          residents_smoke,
-         home_floors_2level, 
-         home_sqft_2level,
+         home_floors_2level, home_sqft_2level,
          home_pets_2level,
-         stove_age_3level, 
-         chimney_clean_3level, 
+         stove_age_3level, chimney_clean_3level, 
          wood_collect_2level,
-         moisture_closest_2level,
          sums_events_sampling_period,
          mean_temp_sampling_period_5,
+         amb_pm_sampling_period_5,
          stove_grade_3level,
          home_fireplace) %>% 
-  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>%  
-  select(-pm_mean_sampling_period) 
+  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period))   
+  select(-pm_mean_sampling_period) %>%
+
+  na.exclude()
 ```
@@
 When use `na.exclude()` to only use complete data (what is needed for AIC selection),
-the observations drops to 30. While there is only 10% missing data, it is spread
+the observations drops to 34. While there is only 7.5% missing data, it is spread
 out enough that excluding it to have complete observations means we can only use
@@
 selection to exclude that much data, which justifies multiple imputation to 
-impute the 10% missing data and use that for model selection.
+impute the 7.5% missing data and use that for model selection.
 The missing observations are even more extreme for NN and AK.
@@
 # Check % missing data
-mean(is.na(selected_vars)) # 10.7% for WMT
+mean(is.na(selected_vars)) # 7.5% for WMT
 
@@
 missing_test <- selected_vars %>% 
-  mutate(missing_var = if_else(is.na(residents_smoke), "missing yes", "missing no")) %>% 
+  mutate(missing_var = if_else(is.na(stove_grade_3level), "missing yes", "missing no")) %>% 
   filter(!is.na(log_pm_mean_sampling_period) & !is.na(missing_var))

### 2020-06-30 manuscript_tables_figures.Rmd
exposure analysis updates

@@
 author: "Ethan Walker"
-date: "Started 13 May 2020, Updated 29 June 2020"
+date: "Started 13 May 2020, Updated 30 June 2020"
 output:
@@
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>%
-  filter(home != "WMA300") %>% 
+  filter(treatment_assigned == "Placebo") %>%
   arrange(new_var) %>% 
@@
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  filter(home != "WMA300") %>% 
+  filter(treatment_assigned == "Placebo") %>% 
   arrange(new_var) %>% 
@@
 ```{r}
-other_vars_total <- exposure_analysis_data_short %>% 
-  rename(new_var = sums_events_sampling_period) %>% 
+other_vars_total <- health_exposure_data_sampling_day %>% 
+  rename(new_var = home_act_sweep_sum) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  filter(home != "WMA300") %>% 
+  filter(treatment_assigned == "Placebo") %>% 
   arrange(new_var) %>% 
@@
 
-other_vars_area <- exposure_analysis_data_short %>% 
-  rename(new_var = sums_events_sampling_period) %>% 
+other_vars_area <- health_exposure_data_sampling_day %>% 
+  rename(new_var = home_act_sweep_sum) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
   filter(winter_id == 1) %>% 
-  filter(treatment == "Placebo") %>% 
-  filter(home != "WMA300") %>% 
+  filter(treatment_assigned == "Placebo") %>% 
   arrange(new_var) %>% 

### 2020-06-30 merge_datasets.Rmd
exposure analysis updates

@@
 
-exposure_analysis_data_long_new <- 
-  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
-
-exposure_analysis_data_medium <- exposure_analysis_data_long_new %>% 
-  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
-         sampling_day, sample_date, burn_level:mean_temp_roll_4day) %>% 
-  group_by(area, home, winter_id, child_id_char) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
-         sampling_day, child_id_char, burn_level:mean_temp_roll_4day) %>% 
-  mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
-         filter_var = if_else(is.na(filter_var), 0, filter_var)) %>% 
-  select(filter_var, area:mean_temp_roll_4day) %>% 
-  filter(filter_var == 0) %>% 
-  select(-filter_var) %>% 
-  ungroup()
-
-write_rds(exposure_analysis_data_medium, paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
-write_csv(exposure_analysis_data_medium, 
-          paste0(file_path, "Output/exposure_analysis_data_medium.csv"), na = "")
-```
-
-# Convert medium data to short: only keeping 1 observation per sampling period, repeated by child_id
-```{r}
-file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
-
-exposure_analysis_data_medium <- 
-  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
 exposure_analysis_data_long <- 
@@
 
-exposure_analysis_data_short <- exposure_analysis_data_medium %>% 
+exposure_analysis_data_medium <- exposure_analysis_data_long %>% 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         sampling_day, sample_date, burn_level:mean_temp_roll_4day) %>% 
+  group_by(area, home, winter_id, child_id_char) %>% 
+  distinct(sampling_day, .keep_all = TRUE) %>% 
+  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         sampling_day, child_id_char, burn_level:mean_temp_roll_4day) %>% 
+  mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
+         filter_var = if_else(is.na(filter_var), 0, filter_var)) %>% 
+  select(filter_var, area:mean_temp_roll_4day) %>% 
+  filter(filter_var == 0) %>% 
+  select(-filter_var) %>% 
   arrange(area, home, winter_id) %>% 
@@
          home_act_sweep_sum = sum(home_act_sweep)) %>% 
+  ungroup() %>% 
+  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
+  left_join(ambient_data, by = c("area", "home"))
+
+write_rds(exposure_analysis_data_medium, paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
+write_csv(exposure_analysis_data_medium, 
+          paste0(file_path, "Output/exposure_analysis_data_medium.csv"), na = "")
+```
+
+# Convert medium data to short: only keeping 1 observation per sampling period, repeated by child_id
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+exposure_analysis_data_medium <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds"))
+
+exposure_analysis_data_short <- exposure_analysis_data_medium %>% 
   group_by(area, home, winter_id) %>% 
@@
   ungroup() %>% 
-  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
-  left_join(ambient_data, by = c("area", "home")) %>% 
   arrange(area, home, winter_id)

### 2020-06-30 merge_health_datasets.Rmd
exposure analysis updates

--- 2020-06-26 merge_health_datasets.Rmd
@@
                    "sampling_day")) %>% 
-  filter(!is.na(home))
+  filter(!is.na(home)) %>% 
+  filter(!is.na(child_id_num)) %>% 
+  select(-child_id_char) %>% 
+  left_join(demographics, by = c("area", "winter_id", "child_id_num")) %>% 
+  select(area:child_id_num, child_id_char, visit_date:amb_pm_sampling_period_5)
 

### 2020-07-01 exposure_models.Rmd
analysis updates

@@
   #filter(area == "WMT") %>% 
-  select(pm_mean_sampling_period, income_3level, area, home,
+  select(pm_mean_sampling_period, income_3level, 
          residents_smoke,
@@
          home_fireplace) %>% 
-  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period))   
-  select(-pm_mean_sampling_period) %>%
+  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>%  
+  select(-pm_mean_sampling_period)
 
-  na.exclude()
+  #na.exclude()
 ```

### 2020-07-01 manuscript_tables_figures.Rmd
analysis updates

@@
 char_funct_area("stove_grade_3level")
+char_funct_total("ruca_code")
+char_funct_area("ruca_code")
+char_funct_total("ruca_code_2")
+char_funct_area("ruca_code_2")
+char_funct_total("ruca_code_3")
+char_funct_area("ruca_code_3")
 ```

### 2020-07-01 merge_datasets.Rmd
analysis updates

@@
   arrange(area, home, winter_id) %>% 
-  group_by(home) %>% 
-  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
+  #group_by(home) %>% 
+  #mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
   mutate(winter_id = as.character(winter_id)) %>%
@@
   distinct(sampling_day, .keep_all = TRUE) %>% 
-  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
-         sampling_day, child_id_char, burn_level:mean_temp_roll_4day) %>% 
+  select(area, home, home_id, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
+         sampling_day, child_id_char, child_id_num, burn_level:mean_temp_roll_4day) %>% 
   mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
@@
   group_by(area, home, winter_id) %>% 
+  arrange(area, home, child_id_char, winter_id, sampling_day) %>% 
   distinct(child_id_char, .keep_all = TRUE) %>% 

### 2020-07-01 merge_health_datasets.Rmd
analysis updates

--- 2020-06-26 merge_health_datasets.Rmd
@@
 
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
+  select(child_id_num, child_id_char, area, winter_id)
+
+ruca_mt <- 
+  read_xls(paste0(file_path, "Input/ruca2/ruca2_MT.xls")) %>% 
+  rename(zip = ZIPN,
+         ruca_code = RUCA2) %>% 
+  select(zip, ruca_code)
+
+ruca_ak <- 
+  read_xls(paste0(file_path, "Input/ruca2/ruca2_AK.xls")) %>% 
+  rename(zip = ZIPN,
+         ruca_code = RUCA2) %>% 
+  select(zip, ruca_code)
+
+ruca_az <- 
+  read_xls(paste0(file_path, "Input/ruca2/ruca2_AZ.xls")) %>% 
+  rename(zip = ZIPN,
+         ruca_code = RUCA2) %>% 
+  select(zip, ruca_code)
+
+ruca_nm <- 
+  read_xls(paste0(file_path, "Input/ruca2/ruca2_NM.xls")) %>% 
+  rename(zip = ZIPN,
+         ruca_code = RUCA2) %>% 
+  select(zip, ruca_code)
+
+ruca_ut <- 
+  read_xls(paste0(file_path, "Input/ruca2/ruca2_UT.xls")) %>% 
+  rename(zip = ZIPN,
+         ruca_code = RUCA2) %>% 
+  select(zip, ruca_code)
+
+ruca_codes <- rbind(ruca_mt, ruca_ak, ruca_az, ruca_nm, ruca_ut)
+
 health_exposure_data_sampling_day <- health_data_merged %>% 
-  select(-pm_mean_sampling_day, -pm_mean_sampling_period) %>% 
+  select(-pm_mean_sampling_day, -pm_mean_sampling_period, -day_of_week,
+         -pm_spike_1000, -pm_spike_2000, -pm_spike_5000) %>% 
   mutate(winter_id = as.character(winter_id)) %>% 
@@
                    "sampling_day")) %>% 
-  filter(!is.na(home))
+  filter(!is.na(home)) %>% 
+  filter(!is.na(child_id_num)) %>% 
+  select(-child_id_char) %>% 
+  left_join(demographics, by = c("area", "winter_id", "child_id_num")) %>% 
+  select(area, home, home_id, home_winter_id, winter_id, child_id_num, child_id_char, 
+         treatment_assigned, treatment_actual, sampling_day, visit_date:amb_pm_sampling_period_5) %>% 
+  left_join(ruca_codes, by = "zip") %>% 
+  mutate(ruca_code_2 = if_else(ruca_code >= 10, 10, ruca_code),
+         ruca_code_2 = if_else(ruca_code_2 == 7.4, 7, ruca_code_2),
+         ruca_code = as.factor(ruca_code),
+         ruca_code_2 = as.factor(ruca_code_2),
+         ruca_code_3 = fct_collapse(ruca_code_2,
+                                    "metro" = c("1", "2"),
+                                    "small_town" = c("7", "8", "9"),
+                                    "rural" = c("10")))
 
@@
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  select(area, home, zip)
 

### 2020-07-01 stove_grade_usage_summary.Rmd
analysis updates

@@
 
-stove_grades_total <- stove_grades_clean %>% 
-  filter(!is.na(stove_grade)) %>% 
-  filter(!is.na(home)) %>% 
-  group_by(area, home) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
-  select(area, home, stove_grade) %>% 
-  arrange(area, home) 
+stove_grades_summary <- stove_grades_clean %>% 
+  count(area, winter_id, treatment_assigned, stove_grade) 
 ```

### 2020-07-02 collinearity_exposure_analysis.Rmd
exposure analysis and data cleaning updates

@@
          wood_collect_2level,
-         sums_events_sampling_period,
+         sums_events_sampling_period_7.5,
          mean_temp_sampling_period_5,

### 2020-07-02 exposure_covariate_associations.Rmd
exposure analysis and data cleaning updates

@@
   analysis_data <- dataset %>% 
-    filter(winter_id == 1 & treatment_assigned == "Placebo" & area == "WMT") %>% 
+    filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
     group_by(area, home) %>% 
@@
 
-  print(table(model.frame(analysis_results)$analysis_covar))
-  #print(nobs(analysis_results))
+  #print(table(model.frame(analysis_results)$analysis_covar))
+  print(nobs(analysis_results))
 
@@
 d42 <- analysis_function("temp_perc_30_2level", "SUMs - % obs > 30 degrees C")
-d43 <- analysis_function("sums_events_sampling_period", 
-                         "Daily SUMs events - 5C temp increase in 20 min")
+d43 <- analysis_function("sums_events_sampling_period_5", 
+                         "SUMs events - 5C temp increase in 20 min")
 
@@
 
+d62 <- analysis_function("ruca_code_3", "RUCA Code")
+
+d63 <- analysis_function("sums_events_sampling_period_2.5", 
+                         "SUMs events - 2.5C temp increase in 20 min")
+d64 <- analysis_function("sums_events_sampling_period_7.5", 
+                         "SUMs events - 7.5C temp increase in 20 min")
+
 ####### Combine results to use in a table
@@
                     d57, d58, d59, d60,
-                    d61)
+                    d61, d62, d63, d64)
 
@@
    pack_rows("Stove grade", 71, 72) %>%
+   pack_rows("RUCE Code", 73, 74) %>% 
+   pack_rows("SUMs other", 75, 76) %>% 
    footnote(c("PM2.5 = fine particulate matter", 

### 2020-07-02 exposure_models.Rmd
exposure analysis and data cleaning updates

--- 2020-06-30 exposure_models.Rmd
@@
 selected_vars <- analysis_data %>% 
-  #filter(area == "WMT") %>% 
-  select(pm_mean_sampling_period, income_3level, area, home,
+  filter(area == "WMT") %>% 
+  select(pm_mean_sampling_period, income_3level, 
          residents_smoke,
@@
          home_fireplace) %>% 
-  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period))   
-  select(-pm_mean_sampling_period) %>%
+  mutate(log_pm_mean_sampling_period = log(pm_mean_sampling_period)) %>%  
+  select(-pm_mean_sampling_period)
 
@@
 When use `na.exclude()` to only use complete data (what is needed for AIC selection),
-the observations drops to 34. While there is only 7.5% missing data, it is spread
+the observations drops to 27. While there is only 8.4% missing data, it is spread
 out enough that excluding it to have complete observations means we can only use
@@
 selection to exclude that much data, which justifies multiple imputation to 
-impute the 7.5% missing data and use that for model selection.
+impute the 8.4% missing data and use that for model selection.
 The missing observations are even more extreme for NN and AK.
@@
 # Check % missing data
-mean(is.na(selected_vars)) # 7.5% for WMT
+mean(is.na(selected_vars)) # 8.4% for WMT
 
@@
 missing_summary <- complete_case %>% 
-  group_by(area, wood_collect_2level) %>% 
+  group_by(area, income_3level) %>% 
   summarize(n = n()) %>% 
@@
          perc = n/total_n*100) %>% 
-  select(area, wood_collect_2level, n, perc, total_n)
+  select(area, income_3level, n, perc, total_n)
 missing_summary
@@
 missing_summary <- missing_data %>% 
-  group_by(area, wood_collect_2level) %>% 
+  group_by(area, income_3level) %>% 
   summarize(n = n()) %>% 
@@
                      method = c("polr", "logreg", "logreg", "logreg", "logreg",
-                                "polr", "polr", "logreg", "logreg", "pmm",
-                                "pmm", "polr", "logreg", "pmm"))
+                                "polr", "polr", "logreg", "pmm",
+                                "pmm", "pmm", "polr", "polr", "pmm"))
 
-write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_22may2020.rds"))
+write_rds(imputed_data, paste0(file_path, "Output/imputed_exposure_data_2july2020.rds"))
 
-imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_22may2020.rds"))
+imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_2july2020.rds"))
 
@@
 
-# Can add "long" or "broad arguments
+# Can add "long" or "broad" arguments
 complete(imputed_data, "long")
@@
 
-write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_22may2020.rds"))
+write_rds(model_ave, paste0(file_path, "Output/imputed_exposure_model_2july2020.rds"))
 
-exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_22may2020.rds"))
+exposure_model <- read_rds(paste0(file_path, "Output/imputed_exposure_model_2july2020.rds"))
 

### 2020-07-02 manuscript_tables_figures.Rmd
exposure analysis and data cleaning updates

--- 2020-06-30 manuscript_tables_figures.Rmd
@@
 char_funct_area("home_pets_2level")
+char_funct_total("home_pets")
+char_funct_area("home_pets")
 char_funct_total("stove_age_3level")
@@
 num_funct_area("amb_pm_24hr")
+num_funct_total("sums_events_sampling_period_2.5")
+num_funct_area("sums_events_sampling_period_2.5")
+num_funct_total("sums_events_sampling_period_5")
+num_funct_area("sums_events_sampling_period_5")
+num_funct_total("sums_events_sampling_period_7.5")
+num_funct_area("sums_events_sampling_period_7.5")
 ```

### 2020-07-02 merge_datasets.Rmd
exposure analysis and data cleaning updates

@@
          # change temp in the following line
-         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
-         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
-                              temp_diff_check),
-         heat_event = if_else(is.na(temp_diff_check), 1,
-                              heat_event)) %>% 
-  group_by(area, home, sampling_day) %>% 
-  mutate(sums_events_sampling_day = sum(heat_event, na.rm = TRUE),
-         sums_events_sampling_day_cat = cut(sums_events_sampling_day, 
-                                            breaks = c(0, 2, 50),
-                           labels = c("<2", "2+"),
-                                 right = FALSE)) %>% 
-  distinct(sampling_day, .keep_all = TRUE) %>% 
-  ungroup() %>% 
-  select(area:pm_rolling_20, sums_mean_winter:sums_mean_sampling_period,
-         time_diff:sums_events_sampling_day_cat) %>% 
+         temp_diff_check_2.5 = if_else(temp_diff >= 2.5, 1, 0),
+         heat_event_2.5 = if_else(temp_diff_check_2.5 == 1 & lag(temp_diff_check_2.5) == 1, 0,
+                              temp_diff_check_2.5),
+         heat_event_2.5 = if_else(is.na(temp_diff_check_2.5), 1,
+                              heat_event_2.5),
+         temp_diff_check_5 = if_else(temp_diff >= 5, 1, 0),
+         heat_event_5 = if_else(temp_diff_check_5 == 1 & lag(temp_diff_check_5) == 1, 0,
+                              temp_diff_check_5),
+         heat_event_5 = if_else(is.na(temp_diff_check_5), 1,
+                              heat_event_5),
+         temp_diff_check_7.5 = if_else(temp_diff >= 7.5, 1, 0),
+         heat_event_7.5 = if_else(temp_diff_check_7.5 == 1 & lag(temp_diff_check_7.5) == 1, 0,
+                              temp_diff_check_7.5),
+         heat_event_7.5 = if_else(is.na(temp_diff_check_7.5), 1,
+                              heat_event_7.5)) %>% 
   arrange(area, home, sample_date) %>% 
-  group_by(home) %>% 
-  mutate(sums_events_sampling_period = mean(sums_events_sampling_day, na.rm = TRUE)) %>% 
-  select(area, home, home_winter_id, sums_events_sampling_period) %>% 
+  group_by(area, home) %>% 
+  mutate(sums_events_sampling_period_2.5 = sum(heat_event_2.5, na.rm = TRUE),
+         sums_events_sampling_period_5 = sum(heat_event_5, na.rm = TRUE),
+         sums_events_sampling_period_7.5 = sum(heat_event_7.5, na.rm = TRUE)) %>% 
+  select(area, home, home_winter_id, 
+         sums_events_sampling_period_2.5,
+         sums_events_sampling_period_5,
+         sums_events_sampling_period_7.5) %>% 
   distinct(home, .keep_all = TRUE) %>% 

### 2020-07-02 pm_summary.Rmd
exposure analysis and data cleaning updates

--- 2020-05-21 pm_summary.Rmd
@@
 ```{r}
-# home that exceed PM guidelines
-pm_guidelines <- exposure_analysis_data_short %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(pm_mean_sampling_period > 35) %>% 
+# homes that exceed PM guidelines
+pm_guidelines <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  filter(pm_mean_sampling_period > 12) %>% 
   distinct(home, .keep_all = TRUE) %>% 
@@
 pm_guidelines
+
+pm_guidelines <- pm_clean %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  #group_by(area) %>% 
+  mutate(total_obs = n()) %>% 
+  filter(pm > 12) %>% 
+  mutate(cutoff_obs = n(),
+         cutoff_perc = cutoff_obs/total_obs*100) %>% 
+  distinct(area, .keep_all = TRUE)
+pm_guidelines
+
+#Pm by different levels of covariates
+pm_check <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  group_by(chimney_clean_3level) %>% 
+  summarize("n" = n(),
+            "Mean PM" = round(mean(pm_mean_sampling_period, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm_mean_sampling_period, na.rm = TRUE)),
+            "Min PM" = round(min(pm_mean_sampling_period, na.rm = TRUE)), 
+            "25%" = round(quantile(pm_mean_sampling_period, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm_mean_sampling_period, na.rm = TRUE)),
+            "75%" = round(quantile(pm_mean_sampling_period, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm_mean_sampling_period, na.rm = TRUE)))
+pm_check
+
+# closer look at stove grade and PM
+pm_stove_grade <- exposure_analysis_data_short %>%
+  select(area, home, pm_mean_sampling_period, stove_grade_3level, 
+         treatment_assigned, winter_id, home_winter_id) %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(home != "WMA300") %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  mutate(stove_grade_2level = fct_collapse(stove_grade_3level,
+                                           "A" = "A",
+                                           "C or F" = c("C", "F"))) %>% 
+  group_by(stove_grade_2level) %>% 
+  summarize("n" = n(),
+            "Mean PM" = round(mean(pm_mean_sampling_period, na.rm = TRUE)), 
+            "SD PM" = round(sd(pm_mean_sampling_period, na.rm = TRUE)),
+            "Min PM" = round(min(pm_mean_sampling_period, na.rm = TRUE)), 
+            "25%" = round(quantile(pm_mean_sampling_period, 0.25, na.rm = TRUE)),
+            "Median PM" = round(median(pm_mean_sampling_period, na.rm = TRUE)),
+            "75%" = round(quantile(pm_mean_sampling_period, 0.75, na.rm = TRUE)),
+            "Max PM" = round(max(pm_mean_sampling_period, na.rm = TRUE)))
+pm_stove_grade
+
+model_data <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(area == "WMT") %>% 
+  filter(home == "WMA300") %>% 
+  distinct(home, .keep_all = TRUE)
+
+check_model <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level,
+                  data = model_data)
+
+summary(check_model)
 ```
@@
 boxplots_kids_pm <- pm_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
   ggplot(aes(area, pm)) +
@@
 ```{r, fig.height=, fig.width=8}
-boxplots_kids_pm <- pm_clean %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  ggplot(aes(area, pm)) +
+boxplots_kids_pm <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  arrange(area, home, sampling_day) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  ggplot(aes(area, pm_mean_sampling_period)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
 ```
+
+```{r}
+# check PM spikes
+spikes_data <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  select(area, home, sampling_day, pm_mean_sampling_day, pm_mean_sampling_period,
+         pm_spike_1000, pm_spike_2000, pm_spike_5000,
+         income_3level, stove_grade_3level, chimney_clean_3level, home_floors_2level) %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  group_by(area, home) %>% 
+  distinct(area, home, sampling_day, .keep_all = TRUE) %>% 
+  mutate(pm_spike_1000 = if_else(pm_spike_1000 == "Yes", 1, 0),
+         pm_spike_2000 = if_else(pm_spike_2000 == "Yes", 1, 0),
+         pm_spike_5000 = if_else(pm_spike_5000 == "Yes", 1, 0),
+         spike_sum_1000 = sum(pm_spike_1000),
+         spike_sum_2000 = sum(pm_spike_2000),
+         spike_sum_5000 = sum(pm_spike_5000)) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  mutate(spike_sum_1000 = as.factor(spike_sum_1000),
+         spike_sum_1000 = fct_collapse(spike_sum_1000, "0" = "0",
+                                       "1" = "1",
+                                       "2" = "2",
+                                       "3+" = c("3", "5"))) %>% 
+  ungroup() %>% 
+  arrange(area, home)
+
+spikes_summary <- spikes_data %>% 
+  mutate(spike_sum_1000 = as.factor(spike_sum_1000),
+         spike_sum_1000 = fct_collapse(spike_sum_1000, "0" = "0",
+                                       "1" = "1",
+                                       "2" = "2",
+                                       "3+" = c("3", "5"))) %>% 
+  group_by(spike_sum_1000) %>% 
+  summarize("N" = n(),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE), 
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE), 
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+spikes_summary
+
+table(spikes_data$spike_sum_1000)
+table(spikes_data$spike_sum_1000, spikes_data$stove_grade_3level)
+```

### 2020-07-17 exposure_models.Rmd
minor analysis updates

@@
   na.exclude()
+
+  # Check % missing data
+mean(is.na(selected_vars)) # 8% for WMT, 10% for NN, 16% for AK, 10% total
 ```
@@
 summary(selected_vars)
-summary(complete(imputed_data))
+summary(complete(imputed_exposure_data))
 
@@
 # Can add "long" or "broad" arguments
-complete(imputed_data, "long")
+complete(imputed_exposure_data, "long")
 
@@
 # Look at R^2 for final selected model
+imputed_exposure_data <- read_rds(paste0(file_path, "Output/imputed_exposure_data_2july2020.rds"))
+
+model_data <- complete(imputed_exposure_data, "long")
+
 model_results <- lm(log_pm_mean_sampling_period ~
+                    income_3level + 
                     residents_smoke +
+                    home_floors_2level +
+                    home_sqft_2level +
+                    stove_age_3level + 
                     chimney_clean_3level + 
-                      stove_grade_3level,
-                      data = imputed_exposure_data$data)
+                    sums_events_sampling_period_7.5 + 
+                    mean_temp_sampling_period_5 + 
+                    amb_pm_sampling_period_5 + 
+                    stove_grade_3level + 
+                    home_fireplace,
+                    data = model_data)
 

### 2020-07-17 manuscript_tables_figures.Rmd
minor analysis updates

@@
 char_funct_area("stove_grade_3level")
+char_funct_total("burn_level_3level")
+char_funct_area("burn_level_3level")
 char_funct_total("ruca_code")

### 2020-07-17 pm_summary.Rmd
minor analysis updates

@@
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 
@@
   filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  arrange(pm_mean_sampling_period) %>% 
   distinct(home, .keep_all = TRUE) %>% 
@@
 # closer look at stove grade and PM
-pm_stove_grade <- exposure_analysis_data_short %>%
+pm_stove_grade <- health_exposure_data_sampling_day %>%
   select(area, home, pm_mean_sampling_period, stove_grade_3level, 
          treatment_assigned, winter_id, home_winter_id) %>% 
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  filter(home != "WMA300") %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  arrange(pm_mean_sampling_period) %>% 
   distinct(home, .keep_all = TRUE) %>% 

### 2020-07-22 manuscript_tables_figures.Rmd
summary stat and plot updates

@@
 ```
+
+```{r, fig.height=1.25, fig.width=4.15}
+# Poster figures
+
+plot_data <- health_exposure_data_sampling_day %>% 
+  mutate(income_3level = factor(income_3level, 
+                                labels = c("< $20,000", "$20,000-$49,999", "$50,000+")),
+         chimney_clean_3level = factor(chimney_clean_3level, 
+                                labels = c("< 6 months ago", "6-12 months ago", "12 + months ago")),
+         stove_grade_3level = factor(stove_grade_3level, 
+                                labels = c("High-quality", "Medium-quality", "Low-quality")),
+         area = factor(area,
+                       labels = c("Alaska", "Navajo Nation", "Western Montana")))
+
+home_char_function <- function(data, var, title_label) {
+
+  home_char_plot <- data %>% 
+  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  arrange(area, home, sampling_day) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  select(area, var) %>% 
+  rename(plot_var = var) %>% 
+  filter(!is.na(plot_var)) %>% 
+  ggplot() +
+    geom_bar(aes(fct_rev(plot_var), fill = area)) +
+    theme_minimal() +
+    coord_flip() +
+    theme(axis.text.x = element_text(size = 16, colour = "black", 
+                                     vjust = 0.7, hjust = 0.7),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          title = element_text(size = 16, colour = "black"),
+          legend.text = element_blank(),
+          legend.title = element_blank(),
+          legend.position = "",
+          axis.title.y = element_blank(),
+          axis.title.x = element_blank(),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(),
+          panel.grid = element_blank()) +
+    labs(title = title_label) +
+    scale_fill_manual(values = cbPalette)
+  home_char_plot
+
+}
+
+home_char_function(plot_data, "income_3level", "Household income")
+home_char_function(plot_data, "residents_under5_2level", "Kids < 5 years of age")
+home_char_function(plot_data, "home_floors_2level", "Home floors/levels")
+home_char_function(plot_data, "stove_age_3level", "Stove age")
+home_char_function(plot_data, "chimney_clean_3level", "Chimney cleaned")
+home_char_function(plot_data, "wood_collect_2level", "Wood collected")
+home_char_function(plot_data, "stove_grade_3level", "Stove grade")
+```

### 2020-07-22 pm_summary.Rmd
summary stat and plot updates

@@
 
-```{r, fig.height=, fig.width=8}
+```{r, fig.height=4, fig.width=3}
 boxplots_kids_pm <- health_exposure_data_sampling_day %>% 
@@
 
-ggsave("boxplots_kids_pm.jpg", height = 7, width = 8)
+ggsave("boxplots_kids_pm.jpg", height = 7, width = 4)
 ```

### 2020-07-24 lrti_tracking.Rmd
lrti data update

@@
 ```{r}
-health_exposure_data_sampling_day <- read_rds("Output/health_exposure_data_sampling_day.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+
+health_data_merged <- 
+  read_rds(paste0(file_path, "Output/health_data_merged.rds"))
+
+wmt_chw <- read_xlsx(paste0(file_path, "input/WMT/childhomewinter.xlsx")) %>% 
+  mutate(area = "WMT") %>% 
+  rename(paul_comments = 19)
+nn_chw <- read_xlsx(paste0(file_path, "input/NN/childhomewinter.xlsx")) %>% 
+  mutate(area = "NN") %>% 
+  rename(paul_comments = 19)
+ak_chw <- read_xlsx(paste0(file_path, "input/AK/childhomewinter.xlsx")) %>% 
+  mutate(area = "AK") %>% 
+  rename(paul_comments = 19)
+
+child_home_winter <- rbind(wmt_chw, nn_chw, ak_chw) %>% 
+  mutate(home_winter_id = as.character(HomeWinterID),
+         child_id_num = as.character(ChildID),
+         paul_lrti = AnyARI) %>% 
+  select(area, home_winter_id, child_id_num, paul_lrti, paul_comments) %>% 
+  filter(paul_lrti %in% c("Yes", "Debating"))
+
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
+  select(child_id_num, child_id_char, area, winter_id)
 ```
@@
 # LRTI Diagnosis data
-lrti_diagnosis_data <- health_exposure_data_sampling_day %>% 
+lrti_diagnosis_data <- health_data_merged %>% 
   # select relevant health data
   select(area, home, home_winter_id, winter_id, child_id_num, 
-         visit_date, med_visit_date, med_visit_referred,
-         csq_date:csq_symptom_free_date,
-         csq_see_dr, csq_dr_diagnosis,
-         med_rec_visit_date:med_rec_comments,
-         csq_miss_work:csq_comments,
-         med_visit_ari:med_visit_mean_temp,
-         lrti_number:lrti_professional_comments) %>% 
-  # Currently only have this data for WMT - need to get new files and updates dataset
-  filter(area == "WMT") %>% 
+         visit_date, csq_date:sibling_school, treatment_assigned) %>% 
+  filter(!is.na(visit_date)) %>% 
+  full_join(child_home_winter, by = c("area", "home_winter_id", "child_id_num")) %>% 
+  filter(home_winter_id == 19)
   # This is the most broad way of distinguishing if a child was sick

### 2020-07-24 pm_summary.Rmd
lrti data update

@@
     theme_minimal() +
-    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
+    labs(title = "Red triangle = mean",
+         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
     theme(axis.text.x = element_text(size = 16, colour = "black",

### 2020-07-28 lrti_tracking.Rmd
update LRTI tracking

@@
 author: "Ethan Walker"
-date: "Started 6 April 2020, Updated 9 April 2020"
+date: "Started 6 April 2020, Updated 28 July 2020"
 output: pdf_document 
@@
   full_join(child_home_winter, by = c("area", "home_winter_id", "child_id_num")) %>%
-  filter(home_winter_id == 19)
-  # This is the most broad way of distinguishing if a child was sick
-  # Careful when filtering here. Could still make cases that these vars didn't catch
-  # Best to have a look through all rows, then use this to filter data down so it
-  # is easier to work with to assess dates/at-risk times
-  filter(csq_sick_2_weeks == "Yes" | csq_still_sick == "Yes") %>% 
-  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
-  ungroup()
+
+# 6 homes have LRTI data, but actually dropped. Notes on those homes below:
+# AK HWID 19/Child ID 25: No record that they participated except initial med visits/CSQs. Likely dropped.
+# AK HWID 83/Child ID 70: Dropped prior to winter 1
+# NN HWID 32/Child ID 45: No record that they participated except initial med visits/CSQs. Likely dropped.
+# NN HWID 162/Child ID 171: Dropped prior to winter 1
+# NN HWID 164/Child ID 173: Dropped prior to winter 1
+# NN HWID 213/Child ID 197: Dropped prior to winter 1
+# I confirmed that all of the above homes dropped, so filtering them from dataset
+
+  filter(!is.na(home)) %>% 
+  filter(!is.na(paul_lrti)) %>%
+
+# This filtered data contains all observations for kids with LRTI diagnosis of "Yes" or "Debating"
+# Next steps will be to:
+  # Select necessary variables to identify dates and other info for analysis
+  # Identify where they left off assigning dates previously
+  # Look through LRTI "No" designations and make sure Paul didn't miss any potential diagnoses
+  # Assign inital dates for those that don't have them
+    # Make notes along the way for a meeting with Paul to confirm final dates
+
+  select(area, child_id_num, home_winter_id, paul_lrti, paul_comments) %>% 
+  arrange(area, child_id_num)
 

### 2020-07-29 lrti_tracking.Rmd
update LRTI tracking

@@
          paul_lrti = AnyARI) %>% 
-  select(area, home_winter_id, child_id_num, paul_lrti, paul_comments) %>% 
-  filter(paul_lrti %in% c("Yes", "Debating"))
+  select(area, home_winter_id, child_id_num, paul_lrti, paul_comments)  
+  #filter(paul_lrti %in% c("Yes", "Debating"))
 
@@
 
-  select(area, child_id_num, home_winter_id, paul_lrti, paul_comments) %>% 
-  arrange(area, child_id_num)
+  select(area, home, child_id_num, home_winter_id, winter_id, 
+         paul_lrti, paul_comments,
+         visit_date:treatment_assigned) %>% 
+  mutate(child_id_num = as.numeric(child_id_num),
+         home_winter_id = as.numeric(home_winter_id)) %>% 
+  arrange(area, child_id_num, home_winter_id)
 
-#  filter(visit_date > "2018-06-01") # filter for newer data by date
+write_rds(lrti_diagnosis_data, paste0(file_path, "Output/lrti_diagnosis_data.rds"))
+write_csv(lrti_diagnosis_data, paste0(file_path, "Output/lrti_diagnosis_data.csv", na = " "))
 
-# write_rds(lrti_diagnosis_data, "Output/lrti_diagnosis_data.rds")
-# write_csv(lrti_diagnosis_data, "Output/lrti_diagnosis_data.csv", na = " ")
+lrti_notes_decisions <- lrti_diagnosis_data %>% 
+  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
+  select(area, home, child_id_num, home_winter_id, winter_id, 
+         paul_lrti, paul_comments) %>% 
+  mutate(notes_initial = "",
+         start_date_initial = "",
+         end_date_initial = "",
+         notes_final = "",
+         lrti_final = "",
+         lrti_start_date_final = "",
+         lrti_end_date_final = "")
+
+write_rds(lrti_notes_decisions, paste0(file_path, "Output/lrti_notes_decisions.rds"))
+write_csv(lrti_notes_decisions, paste0(file_path, "Output/lrti_notes_decisions.csv", na = " "))
 ```

### 2020-07-30 manuscript_tables_figures.Rmd
write new csv - data for letters

@@
 ```
+
+```{r}
+# Summary data for results letters to send to participants
+
+home_summary <- health_exposure_data_sampling_day %>% 
+  group_by(home) %>% 
+  arrange(home, sampling_day) %>% 
+  select(area, home, winter_id, pm_mean_sampling_period, rh_indoor_min, rh_indoor_max) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  mutate("Min RH (%)" = if_else(rh_indoor_min > rh_indoor_max, rh_indoor_max, rh_indoor_min),
+         "Max RH (%)" = if_else(rh_indoor_max < rh_indoor_min, rh_indoor_min, rh_indoor_max)) %>% 
+  arrange(area, home) %>% 
+  mutate("Mean PM2.5 (ug/m3)" = round(pm_mean_sampling_period, digits = 1),
+         check = if_else(`Min RH (%)` > `Max RH (%)`, 1, 0)) %>% 
+  select(-rh_indoor_min, -rh_indoor_max, -pm_mean_sampling_period, -check) %>% 
+  rename(Area = area,
+         Home = home,
+         Winter = winter_id)
+
+write_csv(home_summary, paste0(file_path, "Output/kids_letter_data.csv"), na = "")
+```

### 2020-08-25 lrti_tracking.Rmd
lrti data dates

@@
 
+check_data <- health_exposure_data_sampling_day %>% 
+  filter(child_id_num == 68 & area == "NN")
+  filter(home == "CH110")
+
 health_data_merged <- 
@@
 demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
+  filter(child_id_num == 52)
   select(child_id_num, child_id_char, area, winter_id)
@@
   filter(!is.na(home)) %>% 
-  filter(!is.na(paul_lrti)) %>%
+  select(area, home, child_id_num, home_winter_id, winter_id, 
+         paul_lrti, paul_comments,
+         visit_date:treatment_assigned) %>% 
+  mutate(lrti_any_winter = if_else(!is.na(paul_lrti), 1, 0)) %>% 
+  group_by(home) %>% 
+  mutate(lrti_any_winter = sum(lrti_any_winter)) %>% 
+  select(area, home, lrti_any_winter, child_id_num, home_winter_id, winter_id, 
+         paul_lrti, paul_comments,
+         visit_date:treatment_assigned) %>% 
+  filter(lrti_any_winter > 0) %>%
 
@@
 
-  select(area, home, child_id_num, home_winter_id, winter_id, 
-         paul_lrti, paul_comments,
-         visit_date:treatment_assigned) %>% 
   mutate(child_id_num = as.numeric(child_id_num),
@@
 
-The filtering below narrows down the LRTI definition too much; we are missing cases
-where children had LRTI but did not visit a healthcare facility and have a med rec.
-If we only use these cases, it could bias the results.
+# Merge dataset with final LRTI notes and dates, then assign person-time
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
 
-There are instances of LRTI in our records that weren’t confirmed in the medical record. 
-There are other instances where the medical record diagnosed LRTI, but our records 
-overturned it. What is our criteria for diagnosing LRTI, where is it documented, 
-and how can we be sure it is consistent in diagnosing LRTI?
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 
-```{r}
-# Further filtering and re-ordering of variables
-lrti_diagnosis_data_filtered <- lrti_diagnosis_data %>% 
-  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
-         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments,
-         lrti_number:lrti_professional_comments) %>% 
-  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
-  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
-  # If a home CSQ was done on same visit date as initial phone CSQ, filter for only the home CSQ
-  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
-  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
-         csq_first_symptom_date:lrti_professional_comments) %>% 
-  # Were they referred (med visit) or did they see a Dr (CSQ)?
-  filter(med_visit_referred == "Yes" | csq_see_dr == "Yes") %>% 
-  # If yes, did we get medical records?
-  filter(!is.na(med_rec_visit_date)) %>% 
-  # If yes, was there an ARI diagnosis?
-  filter(med_rec_ari == "ARI" | med_rec_ari == "Possible ARI") %>% 
-  group_by(child_id_num) %>% 
-  distinct(med_rec_visit_date, .keep_all = TRUE)
+lrti_notes_decisions <- read_xlsx(paste0(file_path, "Input/lrti_notes_decisions.xlsx")) %>% 
+  filter(lrti_final == "Yes") %>% 
+  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
+         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter)
+
+wmt_homewinter_initial <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = ymd(interventiondate),
+         area = "WMT") %>% 
+  select(area, home_winter_id, intervention_date)
+
+nn_homewinter_initial <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         area = "NN") %>% 
+  select(area, home_winter_id, intervention_date)
+
+ak_homewinter_initial <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
+  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         area = "AK") %>% 
+  select(area, home_winter_id, intervention_date)
+
+intervention_dates <- rbind(wmt_homewinter_initial, 
+                            nn_homewinter_initial,
+                            ak_homewinter_initial)
+
+lrti_yes <- lrti_notes_decisions %>% 
+  mutate(child_id_num = as.character(child_id_num),
+         home_winter_id = as.character(home_winter_id),
+         winter_id = as.character(winter_id)) %>% 
+  # join lrti data with full dataset
+  full_join(health_exposure_data_sampling_day, 
+            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
+  filter(lrti_final == "Yes") %>% 
+  # select vars for calculating person-time
+  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
+         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
+         csq_date) %>% 
+  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
+  group_by(area, child_id_num, winter_id) %>% 
+  arrange(area, child_id_num, winter_id, csq_date) %>% 
+  # var for person-time start date for each winter/child combo
+  # this date is intervention date + 14 days
+  mutate(person_time_start = intervention_date + 14) %>% 
+  filter(!is.na(csq_date)) %>% 
+  # var for person-time end date for each winter/child combo
+  # this date is the earliest of the last csq or april 30 of the study year
+  mutate(person_time_end = last(csq_date),
+         latest_date = "04-30") %>% 
+  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
+  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
+  select(-pte_mo, -pte_day) %>% 
+  mutate(latest_date = ymd(latest_date),
+         person_time_end = if_else(person_time_end > latest_date,
+                                   latest_date, person_time_end),
+         person_time_possible = person_time_end - person_time_start) %>% 
+  # person-time possible is number of days from start to end date for each child/winter combo
+  group_by(area, child_id_num, winter_id, lrti_start_date_final) %>% 
+  # end of lrti is the lrti end date +14 days
+  mutate(lrti_end_date_plus14 = ymd(lrti_end_date_final) + 14,
+         lrti_end_date_plus14 = if_else(lrti_end_date_plus14 > latest_date,
+                                 latest_date, lrti_end_date_plus14)) %>% 
+  # filter out lrti cases that start later than person-time end date (2 cases)
+  filter(lrti_start_date_final <= person_time_end) %>% 
+  # calculate days with LRTI
+  mutate(person_time_lrti = ymd(lrti_end_date_plus14) - ymd(lrti_start_date_final)) %>% 
+  distinct(area, child_id_num, winter_id, lrti_start_date_final, .keep_all = TRUE) %>% 
+  group_by(area, child_id_num, winter_id) %>% 
+  # calculate total LRTI days per child/winter combo
+  mutate(person_time_lrti_total = sum(person_time_lrti)) %>% 
+  ungroup() %>% 
+  #distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  # new var for time without lrti
+  mutate(person_time_no_lrti_total = person_time_possible - person_time_lrti_total) %>% 
+  arrange(area, child_id_num, winter_id)
+
+lrti_no <- lrti_notes_decisions %>% 
+  mutate(child_id_num = as.character(child_id_num),
+         home_winter_id = as.character(home_winter_id),
+         winter_id = as.character(winter_id)) %>% 
+  # join lrti data with full dataset
+  full_join(health_exposure_data_sampling_day, 
+            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
+  filter(is.na(lrti_final)) %>% 
+  # select vars for calculating person-time
+  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
+         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
+         csq_date) %>% 
+  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
+  group_by(area, child_id_num, winter_id) %>% 
+  arrange(area, child_id_num, winter_id, csq_date) %>% 
+  # var for person-time start date for each winter/child combo
+  # this date is intervention date + 14 days
+  mutate(person_time_start = intervention_date + 14) %>% 
+  filter(!is.na(csq_date)) %>% 
+  # var for person-time end date for each winter/child combo
+  # this date is the earliest of the last csq or april 30 of the study year
+  mutate(person_time_end = last(csq_date),
+         latest_date = "04-30") %>% 
+  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
+  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
+  select(-pte_mo, -pte_day) %>% 
+  mutate(latest_date = ymd(latest_date),
+         person_time_end = if_else(person_time_end > latest_date,
+                                   latest_date, person_time_end),
+         person_time_end = if_else(person_time_end < person_time_start, last(csq_date), person_time_end),
+         person_time_possible = as.numeric(person_time_end - person_time_start),
+         person_time_possible = if_else(person_time_possible < 0, 0,
+                                        person_time_possible)) %>% 
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  arrange(area, child_id_num, winter_id)
+
+lrti_final_data <- plyr::rbind.fill(lrti_yes, lrti_no) %>% 
+  arrange(area, child_id_num, winter_id)
+
+write_csv(lrti_final_data, paste0(file_path, "Output/lrti_final_data.csv"), na = "")
 ```
-
-# Defining start date, end date, and at-risk period for each diagnosed LRTI case
-
-First, I need to know (or make) exactly which variable is a final diagnosed case 
-of LRTI in the data. Then, I can create a method in R to identify start/end/at-risk.
-The code below is a start at this, using Nick's handoff documents/algorithm.
-
-The initial attempt at having R select start dates did not match up well with
-the cases in which Paul had already finalized the start/end dates. There seems
-to be a good amount of professional opinion by assessing symptoms and medical
-records. It may be best to filter a dataset down to only confirmed LRTI cases
-and useful health/medical variables, then decide start/end/at-risk times 
-on a case-by-case basis.
-
-```{r}
-# LRTI start date, end date, and at-risk period identification
-lrti_date_data <- health_exposure_data_sampling_day %>% 
-  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
-         csq_date:csq_comments, med_rec_visit_date:med_rec_comments,
-         lrti_number:lrti_professional_comments) %>% 
-  filter(area == "WMT") %>% 
-  filter(csq_sick_2_weeks == "Yes") %>% 
-  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
-         med_rec_visit_date:med_rec_comments, 
-         med_visit_lethargy:lrti_professional_comments) %>% 
-  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
-  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
-  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
-  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
-         csq_first_symptom_date:lrti_professional_comments) %>% 
-  mutate(trash_date = ymd("2099-01-01"),
-         still_sick_date = if_else(csq_still_sick == "Yes", 
-                                   csq_date, trash_date),
-         med_rec_ari_date = if_else(!is.na(med_rec_visit_date), 
-                                    med_rec_visit_date, trash_date),
-         med_visit_ari_date = if_else(!is.na(med_visit_date), 
-                                      med_visit_date, trash_date)) %>% 
-  pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
-               names_to = "lrti_visits", values_to = "lrti_dates") %>% 
-  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
-  mutate(lrti_start_date_new = if_else(!is.na(csq_first_symptom_date),
-                                       csq_first_symptom_date, lrti_dates)) %>% 
-  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
-         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date,
-         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari,
-         lrti_number:lrti_professional_comments) %>% 
-  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
-  distinct(visit_date, .keep_all = TRUE) %>% 
-  group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
-  distinct(lrti_start_date_new, .keep_all = TRUE) %>% 
-  distinct(child_id_num, lrti_number, .keep_all = TRUE) %>% 
-  filter(!is.na(lrti_number)) %>% 
-  select(child_id_num, home_winter_id, lrti_number, lrti_start_date_new:lrti_dates, 
-         lrti_start_date_final, lrti_end_date_final, area:lrti_professional_comments)
-```

### 2020-08-26 lrti_tracking.Rmd
fig 1 summary stats

@@
 author: "Ethan Walker"
-date: "Started 6 April 2020, Updated 28 July 2020"
+date: "Started 6 April 2020, Updated 26 Aug 2020"
 output: pdf_document 
@@
 
+write_rds(lrti_final_data, paste0(file_path, "Output/lrti_final_data.rds"))
 write_csv(lrti_final_data, paste0(file_path, "Output/lrti_final_data.csv"), na = "")
 ```
+
+# Summary stats for LRTI time at-risk. Check for odd values to finish data cleaning.
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  select(area, home_winter_id, treatment_assigned) %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE)
+
+summary_stats <- lrti_final_data %>% 
+  filter(winter_id != 0) %>% 
+  group_by(area, winter_id) %>% 
+  summarize("N" = n(),
+            "Mean PT Possible" = mean(person_time_possible, na.rm = TRUE),
+            "SD PT Possible" = sd(person_time_possible, na.rm = TRUE),
+            "Min PT Possible" = min(person_time_possible, na.rm = TRUE),
+            "Med PT Possible" = median(person_time_possible, na.rm = TRUE),
+            "Max PT Possible" = max(person_time_possible, na.rm = TRUE),
+            "Mean PT LRTI" = mean(person_time_lrti_total, na.rm = TRUE),
+            "SD PT LRTI" = sd(person_time_lrti_total, na.rm = TRUE),
+            "Min PT LRTI" = min(person_time_lrti_total, na.rm = TRUE),
+            "Med PT LRTI" = median(person_time_lrti_total, na.rm = TRUE),
+            "Max PT LRTI" = max(person_time_lrti_total, na.rm = TRUE))
+
+summary_stats
+
+fig1_stats <- lrti_final_data %>% 
+  left_join(health_exposure_data_sampling_day, by = c("area", "home_winter_id")) %>% 
+  filter(winter_id != 0) %>% 
+  group_by(area, winter_id, treatment_assigned) %>% 
+  filter(!is.na(person_time_possible)) %>% 
+  mutate(child_weeks_obs = as.numeric(person_time_possible),
+         child_weeks_obs = sum(child_weeks_obs)/7) %>% 
+  distinct(area, winter_id, treatment_assigned, .keep_all = TRUE) %>% 
+  select(area, winter_id, treatment_assigned, child_weeks_obs)
+
+fig1_stats
+```

### 2020-08-26 pm_summary.Rmd
fig 1 summary stats

@@
 ```
+
+# Summary stats for Figure 1
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+exposure_analysis_data_short <- 
+  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds")) %>% 
+  filter(winter_id == 1) %>% 
+  select(area, home_winter_id, winter_id, treatment_assigned, 
+         pm_sample_interval, pm_mean_sampling_period) %>% 
+  arrange(area, home_winter_id, pm_mean_sampling_period) %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE)
+
+fig1_stats <- exposure_analysis_data_short %>% 
+  filter(winter_id == 1) %>% 
+  group_by(area, winter_id, treatment_assigned) %>% 
+  filter(!is.na(pm_sample_interval)) %>% 
+  mutate(pm_sample_days = as.numeric(pm_sample_interval),
+         pm_sample_days = sum(pm_sample_days)) %>% 
+  distinct(area, winter_id, treatment_assigned, .keep_all = TRUE) %>% 
+  select(area, winter_id, treatment_assigned, pm_sample_days)
+
+fig1_stats
+```

### 2020-09-16 exposure_covariate_associations.Rmd
data cleaning and model updates

@@
   filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  filter(area == "AK") %>% 
   arrange(area, home, sampling_day) %>% 
@@
 
-  #print(table(model.frame(analysis_results)$analysis_covar))
+  print(table(model.frame(analysis_results)$analysis_covar))
   print(nobs(analysis_results))
@@
     select(Term, Estimate, 'Lower 95% CL', 'Upper 95% CL', 'P-value')
+  view(tidy_results)
 
@@
 d29 <- analysis_function("mean_moisture_winter", "Wood moisture - Winter mean")
-d30 <- analysis_function("moisture_winter_2level", "Wood moisture - Winter mean")
+#d30 <- analysis_function("moisture_winter_2level", "Wood moisture - Winter mean")
 

### 2020-09-16 indoor_temp_rh_cleaning.Rmd
data cleaning and model updates

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 23 June 2020"
+date: "Started 13 Jan 2020, Updated 15 Sept 2020"
 output: pdf_document
@@
   # replace -999 values with NA
-  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
-                                 rh_indoor_min = -999, temp_indoor_current = -999,
+  replace_with_na(replace = list(rh_indoor_max = -999,
+                                 rh_indoor_min = -999, 
                                  temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
-  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
-                                 rh_indoor_min = 0, temp_indoor_current = 0,
+  replace_with_na(replace = list(rh_indoor_max = 0,
+                                 rh_indoor_min = 0, 
                                  temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
-  select(home_winter_id, area, rh_indoor_current, 
-         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+  select(home_winter_id, area, 
+         rh_indoor_max, rh_indoor_min, 
          temp_indoor_max, temp_indoor_min) %>% 
@@
   # replace 0 and -999 values with NA
-  replace_with_na(replace = list(rh_indoor_current = -999, rh_indoor_max = -999,
-                                 rh_indoor_min = -999, temp_indoor_current = -999,
+  replace_with_na(replace = list(rh_indoor_max = -999,
+                                 rh_indoor_min = -999, 
                                  temp_indoor_max = -999, temp_indoor_min = -999)) %>% 
-  replace_with_na(replace = list(rh_indoor_current = 0, rh_indoor_max = 0,
-                                 rh_indoor_min = 0, temp_indoor_current = 0,
+  replace_with_na(replace = list(rh_indoor_max = 0,
+                                 rh_indoor_min = 0, 
                                  temp_indoor_max = 0, temp_indoor_min = 0)) %>% 
-  select(home_winter_id, area, rh_indoor_current, 
-         rh_indoor_max, rh_indoor_min, temp_indoor_current, 
+  select(home_winter_id, area, 
+         rh_indoor_max, rh_indoor_min, 
          temp_indoor_max, temp_indoor_min) %>% 
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 wmt_temp_rh <- read_rds(paste0(file_path, "Output/wmt_temp_rh.rds"))
@@
   mutate(area = as.factor(area)) %>% 
-  right_join(kids_linked_ids, by = c("area", "home_winter_id"))
+  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  mutate(rh_indoor_min2 = if_else(rh_indoor_min > rh_indoor_max, rh_indoor_max, rh_indoor_min),
+         rh_indoor_max2 = if_else(rh_indoor_max < rh_indoor_min, rh_indoor_min, rh_indoor_max),
+         temp_indoor_min2 = if_else(temp_indoor_min > temp_indoor_max, temp_indoor_max, temp_indoor_min),
+         temp_indoor_max2 = if_else(temp_indoor_max < temp_indoor_min, temp_indoor_min, temp_indoor_max)) %>% 
+  select(-rh_indoor_max, -rh_indoor_min, -temp_indoor_max, -temp_indoor_min) %>% 
+  rename(rh_indoor_max = rh_indoor_max2,
+         rh_indoor_min = rh_indoor_min2,
+         temp_indoor_max = temp_indoor_max2,
+         temp_indoor_min = temp_indoor_min2)
 

### 2020-09-16 indoor_temp_rh_summary.Rmd
data cleaning and model updates

--- 2020-03-16 indoor_temp_rh_summary.Rmd
@@
 ```{r}
-# Load data
-indoor_temp_rh_clean <- read_rds("Output/indoor_temp_rh_clean.rds")
-kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
-pm_clean <- read_rds("Output/pm_clean.rds")
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 ```
@@
 ```{r, include=FALSE}
-temp_rh_summary <- indoor_temp_rh_clean %>%
-  filter(winter_id == 1 & treatment == "Placebo") %>% 
-  group_by(area) %>% 
-  summarize_if(is.numeric, list(mean = mean, sd = sd, min = min, 
-                                med = median, max = max), na.rm = TRUE) %>% 
-  select(-area) %>% 
-  t() 
-kable(temp_rh_summary, col.names = c("Alaska", "Navajo", "Montana"), 
-      digits = 1, align = "c")
+temp_rh_summary <- health_exposure_data_sampling_day %>%
+  select(area, home, home_winter_id, winter_id, 
+         temp_indoor_max, temp_indoor_min, rh_indoor_max, rh_indoor_min) %>% 
+  mutate(temp_diff = if_else(temp_indoor_min > temp_indoor_max, 1, 0),
+         rh_diff = if_else(rh_indoor_min > rh_indoor_max, 1, 0)) %>% 
+  filter(!is.na(temp_diff)) %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE)
 ```

### 2020-09-16 manuscript_tables_figures.Rmd
data cleaning and model updates

--- 2020-07-22 manuscript_tables_figures.Rmd
@@
 ```
+
+```{r}
+# Summary data for results letters to send to participants
+
+home_summary <- health_exposure_data_sampling_day %>% 
+  group_by(home) %>% 
+  arrange(home, sampling_day) %>% 
+  select(area, home, winter_id, pm_mean_sampling_period, pm_mean_sampling_day, 
+         rh_indoor_min, rh_indoor_max) %>% 
+  mutate(daily_min = min(pm_mean_sampling_day, na.rm = TRUE),
+         daily_max= max(pm_mean_sampling_day, na.rm = TRUE)) %>% 
+  distinct(home, .keep_all = TRUE) %>% 
+  replace_with_na_all(condition = ~.x %in% c("Inf", "-Inf")) %>% 
+  mutate("Min RH (%)" = if_else(rh_indoor_min > rh_indoor_max, rh_indoor_max, rh_indoor_min),
+         "Max RH (%)" = if_else(rh_indoor_max < rh_indoor_min, rh_indoor_min, rh_indoor_max)) %>% 
+  arrange(area, home) %>% 
+  mutate("Mean PM2.5 (ug/m3)" = round(pm_mean_sampling_period, digits = 1),
+         "Min PM2.5 (ug/m3)" = round(daily_min, digits = 1),
+         "Max PM2.5 (ug/m3)" = round(daily_max, digits = 1),
+         check = if_else(`Min RH (%)` > `Max RH (%)`, 1, 0)) %>% 
+  select(-rh_indoor_min, -rh_indoor_max, -pm_mean_sampling_period, -check,
+         -daily_min, -daily_max) %>% 
+  rename(Area = area,
+         Home = home,
+         Winter = winter_id)
+
+write_csv(home_summary, paste0(file_path, "Output/kids_letter_data.csv"), na = "")
+```

### 2020-09-16 merge_datasets.Rmd
data cleaning and model updates

@@
 author: "Ethan Walker"
-date: "Started 24 March 2020, Updated 26 June 2020"
+date: "Started 24 March 2020, Updated 15 Sept 2020"
 output: pdf_document
@@
 
+#write_rds(pm_sums_joined3, paste0(file_path, "Output/pm_sums_joined3.rds"))
+pm_sums_joined3 <- read_rds(paste0(file_path, "Output/pm_sums_joined3.rds"))
+
 # Join PM/SUMs data with other covariate datasets
@@
   select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
-         rh_indoor_current:temp_indoor_min) %>% 
+         rh_indoor_min:temp_indoor_max) %>% 
   ungroup() %>% 

### 2020-09-16 stove_grade_usage_cleaning.Rmd
data cleaning and model updates

@@
 # Load data
-stove_grades_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/stove.csv"), 
+stove_grades_initial_wmt <- read_xlsx(paste0(file_path, "Input/WMT/stove.xlsx"), 
                                   na = c("NULL", ""))
@@
   ungroup() %>% 
-  arrange(home_winter_id)
+  arrange(home_winter_id) %>% 
+  filter(!is.na(stove_grade)) %>% 
+  filter(home_id != 0 & home_id != 999)
 
@@
 # Load data
-stove_grades_initial_nn <- read_csv(paste0(file_path, "Input/NN/stove.csv"), 
+stove_grades_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/stove.xlsx"), 
                                   na = c("NULL", ""))
@@
 # Load data
-stove_grades_initial_ak <- read_csv(paste0(file_path, "Input/AK/stove.csv"), 
+stove_grades_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/stove.xlsx"), 
                                   na = c("NULL", ""))
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
 wmt_stove_grades <- read_rds(paste0(file_path, "Output/wmt_stove_grades.rds"))
@@
   mutate(area = as.factor(area)) %>% 
-  filter(home_id != 0) %>% 
   # Mark gave a couple of "D" grades when there was not enough info to score the stove
@@
   select(-home_id) %>% 
-  right_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>% 
   arrange(area, home, winter_id)

### 2020-09-17 exposure_covariate_associations.Rmd
minor analysis updates

@@
   filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
-  filter(area == "AK") %>% 
+  #filter(area == "AK") %>% 
   arrange(area, home, sampling_day) %>% 
@@
 }
+
+# Check for variance in random effect for area and see if simple/mixed models have different results
+analysis_results1 <- lm(log(pm_mean_sampling_period) ~ stove_grade_3level, analysis_data)
+analysis_results2 <- lmer(log(pm_mean_sampling_period) ~ stove_grade_3level + (1|area), analysis_data)
+#summary(analysis_results1)
+#summary(analysis_results2)
 

### 2020-09-23 main_analysis_initial.Rmd
New RMD for main analysis

---
title: "KidsAIR: Main analysis initial steps"
author: "Ethan Walker"
date: "Started 23 Sept 2020, Updated 23 Sept 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, include = FALSE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r, eval=TRUE, include=TRUE}
# Load individual datasets

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))

treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  group_by(area, home_winter_id, child_id_num) %>% 
  mutate(lrti_events = n(),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         lrti_incidence = lrti_events/person_time_no_lrti_total*7,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence))
```

```{r}
# Merge dataset with blinded treatments and final LRTI data, select analysis variables
analysis_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>% 
  right_join(lrti_final_data, by = c("area", "child_id_num", "home_winter_id")) %>% 
  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
  select(area, home, home_winter_id, )

```

### 2020-09-23 manuscript_tables_figures.Rmd
New RMD for main analysis

@@
 
+treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
+
 demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
@@
 
-ak_ids <- read_rds(paste0(file_path, "Output/ak_ids_linked.rds"))
-nn_ids <- read_rds(paste0(file_path, "Output/nn_ids_linked.rds"))
-wmt_ids <- read_rds(paste0(file_path, "Output/wmt_ids_linked.rds"))
+ak_ids <- read_rds(paste0(file_path, "Output/ak_ids_linked.rds")) %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
+nn_ids <- read_rds(paste0(file_path, "Output/nn_ids_linked.rds")) %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
+wmt_ids <- read_rds(paste0(file_path, "Output/wmt_ids_linked.rds")) %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
 
@@
   filter(winter_id != 2)
+
+kids_ids <- rbind(ak_ids, nn_ids, wmt_ids) %>% 
+  select(-treatment) %>% 
+  filter(winter_id == 1) %>% 
+  filter(!is.na(home)) %>% 
+  mutate(original = 1) %>% 
+  right_join(treatments_masked, by = "home") %>% 
+  filter(home != "WMA222" & home != "WMA233") %>%
+  filter(original == 1)
 ```

### 2020-09-28 main_analysis_initial.Rmd
main analysis updates

@@
 author: "Ethan Walker"
-date: "Started 23 Sept 2020, Updated 23 Sept 2020"
+date: "Started 23 Sept 2020, Updated 28 Sept 2020"
 output: pdf_document
@@
 lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
-  group_by(area, home_winter_id, child_id_num) %>% 
-  mutate(lrti_events = n(),
+  arrange(area, home, child_id_num, winter_id) %>% 
+  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
+  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
+  group_by(area, child_id_num, home_winter_id) %>% 
+  mutate(lrti_events = n()) %>% 
+  group_by(area, child_id_num) %>% 
+  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
+  mutate(lrti_events = as.numeric(lrti_events),
+         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
+         lrti_events_total = sum(lrti_events, na.rm = TRUE),
          person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
-         lrti_incidence = lrti_events/person_time_no_lrti_total*7,
-         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence))
+         time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
+         lrti_incidence = lrti_events_total/sum(time_at_risk_total)*7,
+         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
+  distinct(area, child_id_num, .keep_all = TRUE) %>% 
+  select(area, home, cohort, child_id_num, lrti_events_total, time_at_risk_total, lrti_incidence)
 ```
@@
   left_join(treatments_masked, by = "home") %>% 
-  right_join(lrti_final_data, by = c("area", "child_id_num", "home_winter_id")) %>% 
-  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
-  select(area, home, home_winter_id, )
+  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
+  distinct(area, child_id_num, .keep_all = TRUE) %>% 
+  select(area, home, cohort, child_id_num, lrti_events_total, time_at_risk_total, lrti_incidence, newtx) %>% 
+  rename(treatment = newtx)
+```
 
+```{r}
+# Initial summary data for LRTI
+incidence_summary <- analysis_data %>% 
+  group_by(treatment) %>% 
+  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
+            sd_inc = sd(lrti_incidence, na.rm = TRUE),
+            min_inc = min(lrti_incidence, na.rm = TRUE),
+            median_inc = median(lrti_incidence, na.rm = TRUE),
+            max_inc = max(lrti_incidence, na.rm = TRUE))
+incidence_summary
 ```
+
+```{r}
+model_results <- glmer(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data, family = poisson(link = "log"))
+
+summary(model_results)
+```

### 2020-09-30 main_analysis_initial.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 23 Sept 2020, Updated 28 Sept 2020"
+date: "Started 23 Sept 2020, Updated 30 Sept 2020"
 output: pdf_document
@@
 ```{r}
+# clean up data for table 1 summary
+table_1_data <- health_exposure_data_sampling_day %>% 
+  select(-treatment_assigned, -treatment_actual) %>% 
+  left_join(treatments_masked, by = "home") %>%
+  rename(treatment = newtx) %>% 
+  arrange(area, home, winter_id, sampling_day) %>% 
+  filter(winter_id == 1) %>% 
+  distinct(area, home, .keep_all = TRUE)
+
+# summary data for manuscript tables
+char_funct_total <- function(var, data = table_1_data){
+
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = round(n()/total_n*100)) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_treatment <- function(var, data = table_1_data){
+
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  group_by(treatment) %>% 
+  mutate(treatment_n = n()) %>% 
+  group_by(treatment, new_var) %>% 
+    mutate(n = n(),
+         percent = round(n()/treatment_n*100)) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  select(new_var, n, percent) %>% 
+  arrange(treatment, new_var)
+demographics_summary
+}
+
+char_funct_total("gender_parent")
+char_funct_treatment("gender_parent")
+char_funct_total("hispanic_parent")
+char_funct_treatment("hispanic_parent")
+char_funct_total("race_parent")
+char_funct_treatment("race_parent")
+char_funct_total("education_3level")
+char_funct_treatment("education_3level")
+char_funct_total("income_3level")
+char_funct_treatment("income_3level")
+char_funct_total("home_floors_2level")
+char_funct_treatment("home_floors_2level")
+char_funct_total("home_year_built_2level")
+char_funct_treatment("home_year_built_2level")
+char_funct_total("home_pets_2level")
+char_funct_treatment("home_pets_2level")
+char_funct_total("home_pets")
+char_funct_treatment("home_pets")
+char_funct_total("stove_age_3level")
+char_funct_treatment("stove_age_3level")
+char_funct_total("chimney_clean_3level")
+char_funct_treatment("chimney_clean_3level")
+char_funct_total("wood_collect_method_2level")
+char_funct_treatment("wood_collect_method_2level")
+char_funct_total("wood_collect_2level")
+char_funct_treatment("wood_collect_2level")
+char_funct_total("stove_grade_3level")
+char_funct_treatment("stove_grade_3level")
+char_funct_total("burn_level_3level")
+char_funct_treatment("burn_level_3level")
+char_funct_total("ruca_code")
+char_funct_treatment("ruca_code")
+char_funct_total("ruca_code_2")
+char_funct_treatment("ruca_code_2")
+char_funct_total("ruca_code_3")
+char_funct_treatment("ruca_code_3")
+```
+
+```{r}
+# clean up data for table 1 summary
+table_1_data <- health_exposure_data_sampling_day %>% 
+  select(-treatment_assigned, -treatment_actual) %>% 
+  left_join(treatments_masked, by = "home") %>%
+  rename(treatment = newtx) %>% 
+  arrange(area, home, winter_id, sampling_day) %>% 
+  filter(winter_id == 1) %>% 
+  distinct(area, home, .keep_all = TRUE)
+
+# summary data for manuscript tables
+num_funct_total <- function(var, data = table_1_data){
+
+demographics_summary <- data %>% 
+  rename(new_var = var) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
+  filter(!is.na(new_var)) %>% 
+  select(treatment, home, new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_treatment <- function(var, data = table_1_data){
+
+demographics_summary <- data %>% 
+  rename(new_var = var) %>%
+  mutate(new_var = as.numeric(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
+  filter(!is.na(new_var)) %>% 
+  select(treatment, home, new_var) %>% 
+  group_by(treatment) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_total("residents_under_five")
+num_funct_treatment("residents_under_five")
+num_funct_total("home_bedrooms")
+num_funct_treatment("home_bedrooms")
+num_funct_total("pm_mean_sampling_period")
+num_funct_treatment("pm_mean_sampling_period")
+num_funct_total("pm_sample_interval")
+num_funct_treatment("pm_sample_interval")
+num_funct_total("moisture_closest")
+num_funct_treatment("moisture_closest")
+num_funct_total("temp_indoor_max")
+num_funct_treatment("temp_indoor_max")
+num_funct_total("rh_indoor_max")
+num_funct_treatment("rh_indoor_max")
+num_funct_total("mean_temp")
+num_funct_treatment("mean_temp")
+num_funct_total("amb_pm_24hr")
+num_funct_treatment("amb_pm_24hr")
+num_funct_total("sums_events_sampling_period_2.5")
+num_funct_treatment("sums_events_sampling_period_2.5")
+num_funct_total("sums_events_sampling_period_5")
+num_funct_treatment("sums_events_sampling_period_5")
+num_funct_total("sums_events_sampling_period_7.5")
+num_funct_treatment("sums_events_sampling_period_7.5")
+```
+
+```{r}
 # Initial summary data for LRTI
@@
             sd_inc = sd(lrti_incidence, na.rm = TRUE),
+            n_inc = n(),
             min_inc = min(lrti_incidence, na.rm = TRUE),
@@
 incidence_summary
+
+# Histogram for distribution of data
+incidence_plot <- analysis_data %>% 
+  ggplot() +
+    geom_histogram(aes(lrti_incidence)) +
+    theme_light() +
+    facet_wrap(~treatment)
+incidence_plot
+
+# Check for overdispersion in data using likelihood ratio test
+library(pscl)
+library(MASS)
+
+nb_results <- glm.nb(lrti_incidence ~ treatment, data = analysis_data)
+odTest(nb_results)
+
+# According to these results, data is overdispersed
+# Variance is much higher than the mean for LRTI incidence
+# Assumption of Poisson is variance = mean, so recommend using negative binomial model
+
+summary(nb_results)
+
+# Link for analysis ideas and notes:
+# https://stats.idre.ucla.edu/r/dae/negative-binomial-regression/ 
+# https://stats.idre.ucla.edu/r/dae/poisson-regression/ 
+# https://stats.idre.ucla.edu/r/faq/random-coefficient-poisson-models/ 
 ```
@@
 ```{r}
-model_results <- glmer(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data, family = poisson(link = "log"))
+# Mixed model
+model_results <- glmer.nb(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data)
 

### 2020-10-12 lrti_tracking.Rmd
analysis updates

@@
   filter(!is.na(csq_date)) %>% 
+  mutate(person_time_start = if_else(is.na(person_time_start), first(csq_date) + 14, person_time_start)) %>% 
   # var for person-time end date for each winter/child combo

### 2020-10-12 main_analysis_initial.Rmd
analysis updates

@@
   separate(cohort, c("cohort", "trash"), sep = "-") %>% 
-  group_by(area, child_id_num, home_winter_id) %>% 
+  group_by(area, child_id_num, winter_id) %>% 
   mutate(lrti_events = n()) %>% 
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
   group_by(area, child_id_num) %>% 
-  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
   mutate(lrti_events = as.numeric(lrti_events),
          lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
-         lrti_events_total = sum(lrti_events, na.rm = TRUE),
+         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
          person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
-         time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
-         lrti_incidence = lrti_events_total/sum(time_at_risk_total)*7,
+         person_time_no_lrti_total = if_else(is.na(person_time_no_lrti_total), person_time_possible,
+                                             person_time_no_lrti_total),
+         person_time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
+         person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
+         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
          lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
+  arrange(area, home, child_id_num, winter_id) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  select(area, home, cohort, child_id_num, lrti_events_total, time_at_risk_total, lrti_incidence)
+  ungroup() %>% 
+  filter(person_time_at_risk_total > 0) %>% 
+  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence)
 ```
@@
   select(-treatment_assigned, -treatment_actual) %>% 
+  mutate(child_age = fct_collapse(age_child_nov,
+                                  "< 1" = "Less than 1",
+                                  "1+" = c(1, 2, 3, 4))) %>% 
   left_join(treatments_masked, by = "home") %>% 
+  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
   right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  select(area, home, cohort, child_id_num, lrti_events_total, time_at_risk_total, lrti_incidence, newtx) %>% 
-  rename(treatment = newtx)
+  select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
+         person_time_at_risk_total, lrti_incidence, newtx) %>% 
+  rename(treatment = newtx,
+         lrti_events = lrti_events_overall,
+         person_time_at_risk = person_time_at_risk_total) %>% 
+  mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
+  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
+  filter(lrti_incidence < 0.7) %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Dearborn", "Smith", "Blackfoot")))
 ```
@@
   group_by(treatment) %>% 
+  #filter(lrti_incidence < 0.7) %>% 
   summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
@@
 
+count_summary <- analysis_data %>% 
+  group_by(treatment) %>% 
+  filter(lrti_incidence < 0.7) %>% 
+  summarize(mean_events = mean(lrti_events, na.rm = TRUE),
+            sd_events = sd(lrti_events, na.rm = TRUE),
+            n_events = n(),
+            min_events = min(lrti_events, na.rm = TRUE),
+            median_events = median(lrti_events, na.rm = TRUE),
+            max_events = max(lrti_events, na.rm = TRUE))
+count_summary
+
 # Histogram for distribution of data
-incidence_plot <- analysis_data %>% 
+events_plot <- analysis_data %>% 
   ggplot() +
-    geom_histogram(aes(lrti_incidence)) +
+    geom_histogram(aes(lrti_events)) +
     theme_light() +
     facet_wrap(~treatment)
-incidence_plot
+events_plot
+
+events_di_plot <- analysis_data %>% 
+  ggplot() +
+    geom_boxplot(aes(lrti_events_di, person_time_at_risk)) +
+    facet_wrap(~treatment)
+events_di_plot
 
@@
 
-nb_results <- glm.nb(lrti_incidence ~ treatment, data = analysis_data)
+nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), data = analysis_data)
 odTest(nb_results)
@@
 # https://stats.idre.ucla.edu/r/faq/random-coefficient-poisson-models/
+
+# Other diagnostic plots and tests
+library(vcd)
+fit <- goodfit(analysis_data$lrti_events)
+summary(fit) 
+rootogram(fit)
+Ord_plot(analysis_data$lrti_events)
+
+library(car)
+influencePlot(nb_results)
+
+library(pscl)
+
 ```
@@
 ```{r}
-# Mixed model
-model_results <- glmer.nb(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data)
+# Mixed model variations
 
-summary(model_results)
+poisson_results <- glmer(lrti_events ~ treatment + child_age + (1 | home:cohort:area), 
+                         data = analysis_data, family = poisson(link = "log"), offset = log_person_time_at_risk)
+summary(poisson_results)
+
+negbin_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                           data = analysis_data)
+summary(negbin_results)
+
+plot(negbin_results)
+qqnorm(residuals(negbin_results))
+qqline(residuals(negbin_results))
+
+logreg_results <- glmer(lrti_events_di ~ treatment + child_age + (1 | home:cohort:area), 
+                        data = analysis_data, family = binomial, offset = log_person_time_at_risk)
+summary(logreg_results)
 ```

### 2020-10-19 main_analysis_initial.Rmd
analysis updates

@@
 ```{r, eval=TRUE, include=TRUE}
-# Load individual datasets
+# Load individual datasets, add new vars, and merge/select analysis dataset
 
@@
          home = gsub("_", "", x = home))
+
+kids_amb_pm_data <- 
+  read_rds(paste0(file_path, "Output/kids_amb_pm_data_new.rds")) %>% 
+  mutate(sample_date = ymd(amb_pm_date))
+
+kids_amb_temp_data <- 
+  read_rds(paste0(file_path, "Output/kids_amb_temp_data_new.rds")) %>% 
+  mutate(sample_date = ymd(temp_date))
 
@@
   arrange(area, home, child_id_num, winter_id) %>% 
+  left_join(kids_amb_pm_data, by = c("area", "home")) %>% 
+  mutate(pm_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
+  filter(pm_filter == 1) %>% 
+  group_by(area, child_id_num) %>% 
+  mutate(amb_pm_lrti = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+         amb_pm_lrti, person_time_start, person_time_end) %>% 
+  left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
+  mutate(temp_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
+  filter(temp_filter == 1 | child_id_num == 94) %>% 
+  group_by(area, child_id_num) %>% 
+  mutate(amb_temp_lrti = mean(mean_temp, na.rm = TRUE),
+         amb_temp_lrti = if_else(child_id_num == 94, NaN, amb_temp_lrti),
+         amb_dewp_lrti = mean(mean_dewp, na.rm = TRUE),
+         amb_dewp_lrti = if_else(child_id_num == 94, NaN, amb_dewp_lrti),
+         amb_pressure_lrti = mean(mean_pressure, na.rm = TRUE),
+         amb_pressure_lrti = if_else(child_id_num == 94, NaN, amb_pressure_lrti),
+         amb_wind_lrti = mean(mean_wind_knots, na.rm = TRUE),
+         amb_wind_lrti = if_else(child_id_num == 94, NaN, amb_wind_lrti)) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
   ungroup() %>% 
-  filter(person_time_at_risk_total > 0) %>% 
-  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence)
-```
+  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti)
 
-```{r}
 # Merge dataset with blinded treatments and final LRTI data, select analysis variables
-analysis_data <- health_exposure_data_sampling_day %>% 
+kids_analysis_data_1obs_per_child <- health_exposure_data_sampling_day %>% 
   select(-treatment_assigned, -treatment_actual) %>% 
@@
   select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
-         person_time_at_risk_total, lrti_incidence, newtx) %>% 
+         person_time_at_risk_total, lrti_incidence, newtx,
+         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti,
+         gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
+         home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
+         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
+         wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,
+         ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
+         residents_under_five, home_bedrooms, pm_mean_sampling_period, pm_sample_interval,
+         moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,
+         sums_events_sampling_period_5, sums_events_sampling_period_7.5) %>% 
   rename(treatment = newtx,
@@
   mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
-  filter(lrti_incidence < 0.7) %>% 
   mutate(treatment = factor(treatment,
-                            levels = c("Dearborn", "Smith", "Blackfoot")))
+                            levels = c("Blackfoot", "Dearborn", "Smith")))
+
+write_rds(kids_analysis_data_1obs_per_child, 
+          paste0(file_path, "Output/kids_analysis_data_1obs_per_child.rds"))
+write_csv(kids_analysis_data_1obs_per_child, 
+          paste0(file_path, "Output/kids_analysis_data_1obs_per_child.csv"), na = "")
+```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+analysis_data <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_1obs_per_child.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+
+treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
 ```
@@
 char_funct_treatment("ruca_code_3")
+char_funct_total("residents_smoke")
+char_funct_treatment("residents_smoke")
+char_funct_total("residents_smoke_inside")
+char_funct_treatment("residents_smoke_inside")
 ```
@@
 num_funct_treatment("sums_events_sampling_period_7.5")
+
+at_home_pm_summary <- health_exposure_data_sampling_day %>% 
+  select(-treatment_assigned, -treatment_actual) %>% 
+  left_join(treatments_masked, by = "home") %>%
+  rename(treatment = newtx) %>% 
+  arrange(area, home, winter_id, sampling_day) %>% 
+  filter(winter_id == 1) %>% 
+  distinct(area, child_id_num, .keep_all = TRUE) %>% 
+  mutate(new_var = pm_at_home_sampling_period) %>% 
+  filter(!is.na(new_var)) %>%
+  group_by(treatment) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+at_home_pm_summary
 ```
@@
 
+# amb data summary
+# amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti
+ambient_summary <- analysis_data %>% 
+  mutate(new_var = amb_pm_lrti) %>% 
+  #group_by(treatment) %>% 
+  filter(!is.na(new_var)) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
+  summarize(mean_amb = mean(new_var, na.rm = TRUE),
+            sd_amb = sd(new_var, na.rm = TRUE),
+            n_amb = n(),
+            min_amb = min(new_var, na.rm = TRUE),
+            median_amb = median(new_var, na.rm = TRUE),
+            max_amb = max(new_var, na.rm = TRUE))
+ambient_summary
+
 # Histogram for distribution of data
@@
 events_di_plot
+```
 
+```{r}
 # Check for overdispersion in data using likelihood ratio test
@@
 
-nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), data = analysis_data)
+nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), 
+                     data = analysis_data)
+
+poisson_results <- glm(lrti_events ~ treatment + offset(log_person_time_at_risk), 
+                       family = "poisson", data = analysis_data)
+
 odTest(nb_results)
@@
 
+inf_results <- influence(nb_results)$hat
+inf_results
+
+influence(nb_results)$coef
 ```

### 2020-10-19 stove_grade_usage_cleaning.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 13 Jan 2020, Updated 23 June 2020"
+date: "Started 13 Jan 2020, Updated 18 Oct 2020"
 output: pdf_document
@@
 ```
+
+#####################################
+
+# Load, format, and KW monitor data
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+##### WMT #####
+# Load data
+kw_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/filtretechange.csv"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
+
+kw_wmt <- kw_initial_wmt %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "WMT") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>% 
+  filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_wmt, paste0(file_path, "Output/kw_wmt.rds"))
+
+##### NN #####
+# Load data
+kw_initial_nn <- read_csv(paste0(file_path, "Input/NN/filtretechange.csv"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
+
+kw_nn <- kw_initial_nn %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "NN") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>% 
+  filter(!is.na(kw_hours)) %>%  
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_nn, paste0(file_path, "Output/kw_nn.rds"))
+
+##### AK #####
+# Load data
+kw_initial_ak <- read_csv(paste0(file_path, "Input/AK/filtretechange.csv"), 
+                                  na = c("NULL", "", "...", -99, -999, -9999))
+
+kw_ak <- kw_initial_ak %>% 
+  # rename variables
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         area = "AK") %>% 
+  select(home_winter_id, area, kw_date, kw_hours) %>%  
+  filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_ak, paste0(file_path, "Output/kw_ak.rds"))
+```
+
+# Join KW hour data from different areas
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kw_wmt <- read_rds(paste0(file_path, "Output/kw_wmt.rds"))
+kw_nn <- read_rds(paste0(file_path, "Output/kw_nn.rds"))
+kw_ak <- read_rds(paste0(file_path, "Output/kw_ak.rds"))
+
+kids_kw <- rbind(kw_wmt, kw_nn, kw_ak) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, kw_date)
+
+write_rds(kids_kw, paste0(file_path, "Output/kids_kw.rds"))
+```
+
+# Clean up all KW hour data
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_kw <- read_rds(paste0(file_path, "Output/kids_kw.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
+treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
+
+kids_kw_clean <- kids_kw %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
+  filter(!is.na(home)) %>% 
+  arrange(area, home_winter_id, kw_date) %>% 
+  group_by(area, home_winter_id) %>% 
+  mutate(kw_hours_total = sum(kw_hours))
+
+summary(kids_kw_clean)
+
+kw_summary <- kids_kw_clean %>% 
+  select(-treatment_assigned, -treatment_actual) %>% 
+  left_join(treatments_masked, by = "home") %>% 
+  distinct(area, home, kw_hours_total) %>% 
+  ungroup() %>% 
+  #group_by(newtx) %>% 
+  summarize(mean(kw_hours_total),
+            sd(kw_hours_total),
+            n(),
+            min(kw_hours_total),
+            median(kw_hours_total),
+            max(kw_hours_total))
+kw_summary
+
+# Some homes/stoves were graded during both winters, typically when new photos were added
+# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
+# We can't make the assumption that the stove was exactly the same for both winters,
+# so it will be best to use the grade corresponding to each winter for analysis.
+
+write_rds(kids_kw_clean, paste0(file_path, "Output/kids_kw_clean.rds"))
+```

### 2020-10-23 main_analysis_initial.Rmd
small data cleaning updates

@@
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  mutate(home_pets = as.factor(home_pets),
+         home_pets_3level = fct_collapse(home_pets,
+                                    "0" = "0",
+                                    "1" = "1",
+                                    "2+" = c("2","3","4","5","6","7","8","9","10","12","14")))
 
@@
   filter(winter_id == 1) %>% 
+  filter(area == "NN") %>% 
   distinct(area, home, .keep_all = TRUE)
@@
 char_funct_treatment("gender_parent")
-char_funct_total("hispanic_parent")
-char_funct_treatment("hispanic_parent")
 char_funct_total("race_parent")
@@
 char_funct_treatment("home_pets_2level")
-char_funct_total("home_pets")
-char_funct_treatment("home_pets")
+char_funct_total("home_pets_3level")
+char_funct_treatment("home_pets_3level")
+char_funct_total("residents_smoke")
+char_funct_treatment("residents_smoke")
+char_funct_total("residents_smoke_inside")
+char_funct_treatment("residents_smoke_inside")
 char_funct_total("stove_age_3level")
@@
 char_funct_treatment("ruca_code_3")
-char_funct_total("residents_smoke")
-char_funct_treatment("residents_smoke")
-char_funct_total("residents_smoke_inside")
-char_funct_treatment("residents_smoke_inside")
 ```
@@
   filter(!is.na(new_var)) %>%
+  filter(area == "NN") %>% 
   group_by(treatment) %>% 
@@
 incidence_summary <- analysis_data %>% 
-  group_by(treatment) %>% 
+  filter(area == "NN") %>% 
+  #group_by(treatment) %>% 
   #filter(lrti_incidence < 0.7) %>% 
@@
 ambient_summary <- analysis_data %>% 
+  filter(area == "NN") %>% 
   mutate(new_var = amb_pm_lrti) %>% 
-  #group_by(treatment) %>% 
+  group_by(treatment) %>% 
   filter(!is.na(new_var)) %>% 

### 2020-10-23 stove_grade_usage_cleaning.Rmd
small data cleaning updates

@@
   group_by(area, home_winter_id) %>% 
-  mutate(kw_hours_total = sum(kw_hours))
+  mutate(kw_hours_last = last(kw_hours),
+         diff_check = if_else(kw_hours_last < kw_hours, 1, 0),
+         diff_check2 = kw_hours_last - kw_hours)  
+  distinct(area, home_winter_id, diff_check, .keep_all = TRUE)
 
@@
   ungroup() %>% 
+  filter(area == "WMT") %>% 
   #group_by(newtx) %>% 

### 2020-10-26 main_analysis_initial.Rmd
analysis updates

@@
   ungroup() %>% 
-  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+  dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
          amb_pm_lrti, person_time_start, person_time_end) %>% 
@@
   ungroup() %>% 
-  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+  dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
          amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti)
@@
 kids_analysis_data_1obs_per_child <- health_exposure_data_sampling_day %>% 
-  select(-treatment_assigned, -treatment_actual) %>% 
+  #select(-treatment_assigned, -treatment_actual) %>% 
   mutate(child_age = fct_collapse(age_child_nov,
@@
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
-         person_time_at_risk_total, lrti_incidence, newtx,
+  dplyr::select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
+         person_time_at_risk_total, lrti_incidence, newtx, treatment_assigned,
          amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti,
@@
   mutate(treatment = factor(treatment,
-                            levels = c("Blackfoot", "Dearborn", "Smith")))
+                            levels = c("Blackfoot", "Dearborn", "Smith")),
+         treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education")))
 
@@
 analysis_data <- read_rds(paste0(file_path, 
-                                 "Output/kids_analysis_data_1obs_per_child.rds"))
+                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
+  rownames_to_column() %>% 
+  mutate(filter_outliers = if_else(rowname %in% c(14, 57, 58, 249, 280, 16, 19, 20, 118), 1, 0)) %>% 
+  filter(lrti_incidence < 0.7) %>% 
+  filter(filter_outliers == 0)
 
@@
 ambient_summary
+```
 
+```{r}
 # Histogram for distribution of data
@@
 
+inc_plot <- analysis_data %>% 
+  ggplot() +
+    geom_histogram(aes(lrti_incidence)) +
+    theme_light() +
+    facet_wrap(~treatment)
+inc_plot
+
 events_di_plot <- analysis_data %>% 
   ggplot() +
-    geom_boxplot(aes(lrti_events_di, person_time_at_risk)) +
+    geom_histogram(aes(lrti_events_di)) +
     facet_wrap(~treatment)
@@
 ```{r}
-# Check for overdispersion in data using likelihood ratio test
-library(pscl)
+# Summarize initial models, check overdispersion and other diagnostic criteria
 library(MASS)
 
+# Simple model with offset
 nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                      data = analysis_data)
+
+summary(nb_results)
+anova(nb_results, test="Chisq")
 
@@
 
-odTest(nb_results)
-
-# According to these results, data is overdispersed
-# Variance is much higher than the mean for LRTI incidence
-# Assumption of Poisson is variance = mean, so recommend using negative binomial model
-
-summary(nb_results)
+summary(poisson_results)
+anova(poisson_results, test="Chisq")
 
@@
 # Other diagnostic plots and tests
+# https://stats.stackexchange.com/questions/70558/diagnostic-plots-for-count-regression
+
 library(vcd)
@@
 Ord_plot(analysis_data$lrti_events)
+distplot(analysis_data$lrti_events, type = "nbinom")
 
-library(car)
-influencePlot(nb_results)
+library(AER)
+dispersiontest(poisson_results, trafo=1) # test indicates poisson model is overdispersed
 
 library(pscl)
+odTest(nb_results) # test indicates poisson model is overdispersed
+# According to these results, data is overdispersed
+# Variance is much higher than the mean for LRTI incidence
+# Assumption of Poisson is variance = mean, so recommend using negative binomial model
+
+mod2 <- zeroinfl(lrti_events ~ treatment + offset(log_person_time_at_risk), 
+                 data = analysis_data, dist="poisson")
+AIC(poisson_results, mod2)
+# Above results indicate zero-inflated poisson model is a better fit than poisson model
+
+mod2 <- zeroinfl(lrti_events ~ treatment + offset(log_person_time_at_risk), 
+                 data = analysis_data, dist="negbin")
+AIC(nb_results, mod2)
+# Above results indicate neg bin model is a better fit than zero-inflated neg bin model
+# Since data is also overdispersed (neg bin model better fit than poisson),
+# a neg bin model seems to be the best fit overall
+
+# plot residuals
+res <- residuals(nb_results, type="deviance")
+plot(predict(nb_results), res)
+abline(h=0, lty=2)
+qqnorm(res)
+qqline(res)
+
+library(faraway)
+halfnorm(residuals(nb_results))
+
+library(boot)
+glm.diag.plots(nb_results)
+
+# Check for influential datapoints
+library(car)
+nb_influence <- influencePlot(nb_results)
+influencePlot(poisson_results)
 
@@
 influence(nb_results)$coef
+
+qqnorm(residuals(nb_results))
+qqnorm(residuals(poisson_results))
 ```
@@
 
-plot(negbin_results)
-qqnorm(residuals(negbin_results))
-qqline(residuals(negbin_results))
-
 logreg_results <- glmer(lrti_events_di ~ treatment + child_age + (1 | home:cohort:area), 
@@
 ```
+
+```{r}
+# Run mixed model and diagnostic plots from above
+
+negbin_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                           offset(log_person_time_at_risk), data = analysis_data)
+
+summary(negbin_results)
+tidy(negbin_results, conf.int = TRUE)
+
+# plot residuals
+res <- residuals(negbin_results, type="deviance")
+plot(predict(negbin_results), res)
+abline(h=0, lty=2)
+qqnorm(res)
+qqline(res)
+
+plot(residuals(negbin_results) ~ fitted(negbin_results), xlab="Fitted", ylab="Residuals")
+abline(h=0)
+
+library(faraway)
+halfnorm(residuals(negbin_results))
+
+```

### 2020-10-27 main_analysis_initial.Rmd
analysis upates new RMD for primary models

@@
 ```{r}
+# Load analysis dataset
+
 file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
@@
 ```{r}
+# Summary analysis for categorical variables
+
 # clean up data for table 1 summary
@@
   filter(winter_id == 1) %>% 
-  filter(area == "NN") %>% 
+  filter(area == "NN") %>% # Change area filter as necessary
   distinct(area, home, .keep_all = TRUE)
@@
 ```{r}
+# Summary analysis for numeric variables
+
 # clean up data for table 1 summary
@@
 
+# At-home PM summary - different than function above due to distinct obs by child instead of home
 at_home_pm_summary <- health_exposure_data_sampling_day %>% 
@@
   filter(!is.na(new_var)) %>%
-  filter(area == "NN") %>% 
-  group_by(treatment) %>% 
+  filter(area == "NN") %>% # Change area filter as necessary 
+  group_by(treatment) %>% # Change treatment grouping as necessary 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
 ```{r}
+# Summary for outcome events and ambient data
+
 # Initial summary data for LRTI
 incidence_summary <- analysis_data %>% 
-  filter(area == "NN") %>% 
-  #group_by(treatment) %>% 
+  filter(area == "NN") %>% # Change area filter as necessary  
+  #group_by(treatment) %>% # Change treatment grouping as necessary  
   #filter(lrti_incidence < 0.7) %>% 
@@
 count_summary <- analysis_data %>% 
-  group_by(treatment) %>% 
-  filter(lrti_incidence < 0.7) %>% 
+  group_by(treatment) %>% # Change treatment grouping as necessary 
+  #filter(lrti_incidence < 0.7) %>% 
   summarize(mean_events = mean(lrti_events, na.rm = TRUE),
@@
 ambient_summary <- analysis_data %>% 
-  filter(area == "NN") %>% 
+  filter(area == "NN") %>% # Change area filter as necessary  
   mutate(new_var = amb_pm_lrti) %>% 
-  group_by(treatment) %>% 
+  group_by(treatment) %>% # Change treatment grouping as necessary 
   filter(!is.na(new_var)) %>% 
-  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
+  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% # convert degrees F to C 
   summarize(mean_amb = mean(new_var, na.rm = TRUE),
@@
 ```{r}
-# Histogram for distribution of data
+# Histograms for distribution of outcome data
+
 events_plot <- analysis_data %>% 
@@
 
+events_plot_pm <- analysis_data %>% 
+  mutate(pm_cat = cut(pm_mean_sampling_period,
+                      breaks = c(0, 9.85, 18.06, 34.88, 286),
+                      labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
+  filter(!is.na(pm_cat)) %>% 
+  ggplot() +
+    geom_col(aes(pm_cat, lrti_events)) +
+    theme_light() +
+    facet_wrap(~treatment)
+events_plot_pm
+
+scatter_plot_pm <- analysis_data %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  ggplot() +
+    geom_point(aes(lrti_events, pm_mean_sampling_period)) +
+    theme_light() +
+    facet_wrap(~treatment)
+scatter_plot_pm
+
 events_di_plot <- analysis_data %>% 
@@
     geom_histogram(aes(lrti_events_di)) +
+    facet_wrap(~treatment)
+events_di_plot
+
+events_di_plot <- analysis_data %>% 
+  ggplot() +
+    geom_boxplot(aes(lrti_events_di, pm_mean_sampling_period, group = lrti_events_di)) +
     facet_wrap(~treatment)
@@
 # Summarize initial models, check overdispersion and other diagnostic criteria
+
 library(MASS)
@@
 odTest(nb_results) # test indicates poisson model is overdispersed
+
 # According to these results, data is overdispersed
@@
 halfnorm(residuals(negbin_results))
-
 ```

### 2020-10-27 main_analysis_primary_models.Rmd
analysis upates new RMD for primary models

---
title: 'KidsAIR: Main analysis primary models'
author: "Ethan Walker"
date: "Started 27 Oct 2020, Updated 27 Oct 2020"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
```

```{r}
# Load analysis dataset

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  rownames_to_column() %>% 
  mutate(filter_outliers = if_else(rowname %in% c(14, 57, 58, 249, 280, 16, 19, 20, 118), 1, 0)) %>% 
  filter(lrti_incidence < 0.7) %>% 
  filter(filter_outliers == 0)

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  mutate(home_pets = as.factor(home_pets),
         home_pets_3level = fct_collapse(home_pets,
                                    "0" = "0",
                                    "1" = "1",
                                    "2+" = c("2","3","4","5","6","7","8","9","10","12","14")))

treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))
```

```{r}
# Mixed model variations

poisson_results <- glmer(lrti_events ~ treatment + child_age + (1 | home:cohort:area), 
                         data = analysis_data, family = poisson(link = "log"), offset = log_person_time_at_risk)
summary(poisson_results)

nb_treatment_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + 
                                 offset(log_person_time_at_risk), data = analysis_data)
summary(nb_treatment_results)

nb_pm_results <- glmer.nb(lrti_events ~ log(pm_mean_sampling_period) + child_age + 
                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
                          data = analysis_data)
summary(nb_pm_results)

logreg_results <- glmer(lrti_events_di ~ treatment + child_age + (1 | home:cohort:area), 
                        data = analysis_data, family = binomial, offset = log_person_time_at_risk)
summary(logreg_results)

linear_results <- glmer(pm_mean_sampling_period ~ treatment + child_age + (1 | home:cohort:area), 
                        data = analysis_data)
summary(linear_results)
```

Intent-to-treat model framework: LRTI outcome
```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + 
                          offset(log_person_time_at_risk), data = analysis_data)

summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  dplyr::select(term, estimate, p.value, conf.low, conf.high)
tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "Intervention vs Placebo") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
                                     hjust = .8, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
  scale_y_continuous(breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5), 
                     labels = c(-3, -2, -1, 0, 1, 2, 3, 4, 5)) 
plot_estimates
```

Intent-to-treat model framework: PM outcome
```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer(log(pm_mean_sampling_period) ~ treatment + child_age + (1 | home:cohort:area), 
                        data = analysis_data)

summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment", "", term)) %>% 
  dplyr::select(term, estimate, conf.low, conf.high)
#tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "Intervention vs Placebo") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
                                     hjust = .8, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
plot_estimates
```

Exposure-response model framework
```{r}
# Run mixed model and diagnostic plots from above

model_results <- glmer.nb(lrti_events ~ log(pm_mean_sampling_period) + child_age + 
                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
                          data = analysis_data)

summary(model_results)

# plot residuals/diagnostics
res <- residuals(model_results, type="deviance")
plot(predict(model_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)
halfnorm(residuals(model_results))

# plot results
tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("log", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = "Log PM2.5") %>% 
  dplyr::select(term, estimate, p.value, conf.low, conf.high)
tidy_results

plot_estimates <- tidy_results %>%
  ggplot() +
  geom_point(aes(x=term, y=estimate), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
                size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "LRTI cases per log unit PM2.5 increase") +
  labs(x = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "none") 
plot_estimates
```

### 2020-10-28 main_analysis_initial.Rmd
analysis updates new RMD for secondary analyses

@@
          ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
-         residents_under_five, home_bedrooms, pm_mean_sampling_period, pm_sample_interval,
+         residents_under_five, home_bedrooms, pm_mean_sampling_period, 
+         pm_sample_interval, pm_at_home_sampling_period,
          moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,

### 2020-10-28 main_analysis_primary_models.Rmd
analysis updates new RMD for secondary analyses

@@
 author: "Ethan Walker"
-date: "Started 27 Oct 2020, Updated 27 Oct 2020"
+date: "Started 27 Oct 2020, Updated 28 Oct 2020"
 output:
@@
 library(faraway)
+library(influence.ME)
 ```
@@
                                  "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) 
   rownames_to_column() %>% 
-  mutate(filter_outliers = if_else(rowname %in% c(14, 57, 58, 249, 280, 16, 19, 20, 118), 1, 0)) %>% 
-  filter(lrti_incidence < 0.7) %>% 
+  #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
+  mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
+  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk)) 
+  #filter(lrti_incidence < 0.7) %>% 
   filter(filter_outliers == 0)
@@
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
-  mutate(home_pets = as.factor(home_pets),
-         home_pets_3level = fct_collapse(home_pets,
-                                    "0" = "0",
-                                    "1" = "1",
-                                    "2+" = c("2","3","4","5","6","7","8","9","10","12","14")))
-
-treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
-  mutate(home = gsub(" ", "", x = home),
-         home = gsub("_", "", x = home))
-```
-
-```{r}
-# Mixed model variations
-
-poisson_results <- glmer(lrti_events ~ treatment + child_age + (1 | home:cohort:area), 
-                         data = analysis_data, family = poisson(link = "log"), offset = log_person_time_at_risk)
-summary(poisson_results)
-
-nb_treatment_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + 
-                                 offset(log_person_time_at_risk), data = analysis_data)
-summary(nb_treatment_results)
-
-nb_pm_results <- glmer.nb(lrti_events ~ log(pm_mean_sampling_period) + child_age + 
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
-                          data = analysis_data)
-summary(nb_pm_results)
-
-logreg_results <- glmer(lrti_events_di ~ treatment + child_age + (1 | home:cohort:area), 
-                        data = analysis_data, family = binomial, offset = log_person_time_at_risk)
-summary(logreg_results)
-
-linear_results <- glmer(pm_mean_sampling_period ~ treatment + child_age + (1 | home:cohort:area), 
-                        data = analysis_data)
-summary(linear_results)
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
 ```
@@
 Intent-to-treat model framework: LRTI outcome
+
+What's wrong with the QQ plot?
+What does the half-normal quantile plot mean?
+Does the residual vs fitted plot look appropriate?
 ```{r}
@@
 
-model_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + 
+model_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                           offset(log_person_time_at_risk), data = analysis_data)
 
-summary(model_results)
+#summary(model_results)
 
@@
 
+#influential <- influence(model_results, obs = TRUE)
+#cooks <- cooks.distance.estex(influential, sort = TRUE)
+#plot.estex(influential, which = "cook")
+# Most influential datapoints: 57, 102, 280, 396, 61, 58
+
 # plot results
@@
   filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(term = gsub("treatment_assigned", "", term)) %>% 
   dplyr::select(term, estimate, p.value, conf.low, conf.high)
-tidy_results
+#tidy_results
 
@@
   #ggtitle(label = "Systolic BP ITT models") +
-  labs(y = "Intervention vs Placebo") +
+  labs(y = "LRTI events per child-week at risk \n compared to placebo treatment") +
   labs(x = "") +
@@
           legend.position = "none") 
-  scale_y_continuous(breaks = c(-3, -2, -1, 0, 1, 2, 3, 4, 5), 
-                     labels = c(-3, -2, -1, 0, 1, 2, 3, 4, 5)) 
 plot_estimates
@@
 Intent-to-treat model framework: PM outcome
+
+What's wrong with the residual vs fitted plot?
+This more model is more influenced by additional covariates than the LRTI ITT model
 ```{r}
@@
 
-model_results <- glmer(log(pm_mean_sampling_period) ~ treatment + child_age + (1 | home:cohort:area), 
-                        data = analysis_data)
+model_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), data = analysis_data)
 
-summary(model_results)
+#summary(model_results)
 
@@
 
+#influential <- influence(model_results, obs = TRUE)
+#cooks <- cooks.distance.estex(influential, sort = TRUE)
+#plot.estex(influential, which = "cook")
+
 # plot results
@@
   filter(group_filter == 1) %>% 
-  mutate(term = gsub("treatment", "", term)) %>% 
+  mutate(term = gsub("treatment_assigned", "", term)) %>% 
   dplyr::select(term, estimate, conf.low, conf.high)
@@
   #ggtitle(label = "Systolic BP ITT models") +
-  labs(y = "Intervention vs Placebo") +
+  labs(y = "Difference in PM2.5 \n compared to placebo") +
   labs(x = "") +
@@
 Exposure-response model framework
+
+First assess simple associations between LRTI and covariates
+Start with covariates that are associated with PM2.5 from the exposure paper
+Run these simple models with Placebo homes to get an idea of associations at "baseline"
+Then select a priori confounders based on simple associations and other literature
+
+```{r}
+placebo_analysis_data <- analysis_data %>% 
+  filter(treatment_assigned == "Placebo")
+
+covar_model <- glmer(lrti_events ~ home_bedrooms + (1 | area), data = placebo_analysis_data,
+                   family = poisson(link = "log"), offset = log_person_time_at_risk)
+summary(covar_model)
+
+# Selected vars:
+# education_3level, residents_smoke
+```
+
 ```{r}
@@
 
-model_results <- glmer.nb(lrti_events ~ log(pm_mean_sampling_period) + child_age + 
+model_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
                           (1 | home:cohort:area) + offset(log_person_time_at_risk), 
@@
 
-summary(model_results)
+#summary(model_results)
 
@@
 
+#influential <- influence(model_results, obs = TRUE)
+#cooks <- cooks.distance.estex(influential, sort = TRUE)
+#plot.estex(influential, which = "cook")
+# Most influential data points: 58, 19, 420, 85, 301, 383, 413
+
 # plot results
 tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
-  mutate(group_filter = if_else(grepl("log", term), 1, 0)) %>% 
+  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
   filter(group_filter == 1) %>% 
-  mutate(term = "Log PM2.5") %>% 
+  mutate(term = "PM2.5") %>% 
   dplyr::select(term, estimate, p.value, conf.low, conf.high)
-tidy_results
+#tidy_results
 
@@
   #ggtitle(label = "Systolic BP ITT models") +
-  labs(y = "LRTI cases per log unit PM2.5 increase") +
+  labs(y = "LRTI cases per IQR increase in PM2.5") +
   labs(x = "") +
@@
 ```
+
+Logistic regression framework
+
+```{r}
+# Run mixed model and diagnostic plots from above
+
+model_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+#summary(model_results)
+
+# plot residuals/diagnostics
+res <- residuals(model_results, type="deviance")
+plot(predict(model_results), res)
+abline(h=0, lty=2)
+qqnorm(res)
+qqline(res)
+halfnorm(residuals(model_results))
+
+#influential <- influence(model_results, obs = TRUE)
+#cooks <- cooks.distance.estex(influential, sort = TRUE)
+#plot.estex(influential, which = "cook")
+
+# plot results
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_assigned", "", term)) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high)
+#tidy_results
+
+plot_estimates <- tidy_results %>%
+  ggplot() +
+  geom_point(aes(x=term, y=estimate), size = 4) +
+  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
+  geom_errorbar(aes(x=term, ymin=conf.low, ymax=conf.high), 
+                size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  #ggtitle(label = "Systolic BP ITT models") +
+  labs(y = "Risk of LRTI event compared to Placebo") +
+  labs(x = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_blank(),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "none") 
+plot_estimates
+```

### 2020-10-28 main_analysis_secondary_models.Rmd
analysis updates new RMD for secondary analyses

---
title: 'KidsAIR: Main analysis secondary models'
author: "Ethan Walker"
date: "Started 28 Oct 2020, Updated 28 Oct 2020"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
```

```{r}
# Load analysis dataset

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25)

analysis_data_outliers <- analysis_data %>% 
  #rownames_to_column() %>% 
  #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
  mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk)) 
  #filter(lrti_incidence < 0.7) %>% 
  filter(filter_outliers == 0)

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
```

Sensitivity analyses for ITT framework, LRTI outcome

```{r}
# Run model variations
itt_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_income_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                        income_3level + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_education_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                        education_3level + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_smoke_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                        residents_smoke + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_burn_level_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                        burn_level_3level + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_home_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home) + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_no_age_results <- glmer.nb(lrti_events ~ treatment_assigned + (1 | home:cohort:area) + 
                        offset(log_person_time_at_risk), data = analysis_data)

itt_outlier_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                        offset(log_person_time_at_risk), data = analysis_data_outliers)

# Run function to clean results
results_function <- function(model_results, model_name) {

tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_assigned", "", term),
         model = model_name) %>% 
  dplyr::select(model, term, estimate, p.value, conf.low, conf.high)
tidy_results

}

# Run results through function, save, combine for plotting
itt_tidy <- results_function(itt_results, "ITT Primary")
itt_income_tidy <- results_function(itt_income_results, "ITT + income")
itt_education_tidy <- results_function(itt_education_results, "ITT + education")
itt_smoke_tidy <- results_function(itt_smoke_results, "ITT + smoke hx")
itt_burn_level_tidy <- results_function(itt_burn_level_results, "ITT + burn level")
itt_home_tidy <- results_function(itt_home_results, "ITT home only")
itt_no_age_tidy <- results_function(itt_no_age_results, "ITT - age")
itt_outlier_tidy <- results_function(itt_outlier_results, "ITT outlier")

itt_results_combined <- rbind(itt_tidy, itt_income_tidy, itt_education_tidy, 
                              itt_smoke_tidy, itt_burn_level_tidy, itt_home_tidy,
                              itt_no_age_tidy, itt_outlier_tidy)

# Plot results
itt_plot_estimates <- itt_results_combined %>%
  mutate(model = factor(model,
                        levels = c("ITT Primary", "ITT + income", "ITT + education",
                                   "ITT + smoke hx", "ITT + burn level", "ITT home only",
                                   "ITT - age", "ITT outlier"))) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "LRTI events per child-week at risk \n compared to placebo treatment") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
                                     hjust = .8, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates
```

Sensitivity analyses for ITT framework, PM2.5 outcome

```{r}
# Run model variations
itt_pm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
                       (1 | home:cohort:area), data = analysis_data)

itt_pm_income_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                               child_age + (1 | home:cohort:area) + 
                               income_3level, data = analysis_data)

itt_pm_education_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                                  child_age + (1 | home:cohort:area) + 
                                  education_3level, data = analysis_data)

itt_pm_smoke_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                              child_age + (1 | home:cohort:area) + 
                              residents_smoke, data = analysis_data)

itt_pm_burn_level_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                                   child_age + (1 | home:cohort:area) + 
                                   burn_level_3level, data = analysis_data)

itt_pm_home_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                             child_age + (1 | home), data = analysis_data)

itt_pm_no_age_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                               (1 | home:cohort:area), data = analysis_data)

itt_pm_outlier_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                                child_age + (1 | home:cohort:area), data = analysis_data_outliers)

# Run function to clean results
results_function <- function(model_results, model_name) {

tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_assigned", "", term),
         model = model_name) %>% 
  dplyr::select(model, term, estimate, conf.low, conf.high)
tidy_results

}

# Run results through function, save, combine for plotting
itt_pm_tidy <- results_function(itt_pm_results, "ITT Primary")
itt_pm_income_tidy <- results_function(itt_pm_income_results, "ITT + income")
itt_pm_education_tidy <- results_function(itt_pm_education_results, "ITT + education")
itt_pm_smoke_tidy <- results_function(itt_pm_smoke_results, "ITT + smoke hx")
itt_pm_burn_level_tidy <- results_function(itt_pm_burn_level_results, "ITT + burn level")
itt_pm_home_tidy <- results_function(itt_pm_home_results, "ITT home only")
itt_pm_no_age_tidy <- results_function(itt_pm_no_age_results, "ITT - age")
itt_pm_outlier_tidy <- results_function(itt_pm_outlier_results, "ITT outlier")

itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_income_tidy, itt_pm_education_tidy, 
                              itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_home_tidy,
                              itt_pm_no_age_tidy, itt_pm_outlier_tidy)

# Plot results
itt_pm_plot_estimates <- itt_pm_results_combined %>%
  mutate(model = factor(model,
                        levels = c("ITT Primary", "ITT + income", "ITT + education",
                                   "ITT + smoke hx", "ITT + burn level", "ITT home only",
                                   "ITT - age", "ITT outlier"))) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  #ggtitle(label = "Systolic BP ITT models") +
  labs(y = "Difference in PM2.5 \n compared to placebo") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
                                     hjust = .8, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_pm_plot_estimates
```

### 2020-10-30 child_health_data_summary.Rmd
data cleaning and analysis updates

@@
 ```
+
+\pagebreak
+
+```{r}
+kids_at_home_log <- read_rds(paste0(file_path, "Output/kids_at_home_log.rds"))
+
+at_home_summary <- kids_at_home_log %>% 
+  mutate(treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education"))) %>% 
+  arrange(area, home, child_id_num, sampling_day, sample_time) %>% 
+  group_by(area, home, child_id_num) %>% 
+  mutate(total_n = n(),
+         time_home = sum(at_home, na.rm = TRUE),
+         at_home_percent = time_home/total_n*100) %>% 
+  ungroup() %>% 
+  filter(!is.na(treatment_assigned)) %>% 
+  group_by(area, treatment_assigned) %>% 
+  summarize(mean_at_home = mean(at_home_percent, na.rm = TRUE),
+            sd_at_home = sd(at_home_percent, na.rm = TRUE),
+            min_at_home = min(at_home_percent, na.rm = TRUE),
+            median_at_home = median(at_home_percent, na.rm = TRUE),
+            max_at_home = max(at_home_percent, na.rm = TRUE))
+at_home_summary
+
+at_home_n <- kids_at_home_log %>% 
+  mutate(treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education"))) %>% 
+  filter(!is.na(treatment_assigned)) %>% 
+  filter(!is.na(at_home)) %>% 
+  distinct(area, home, child_id_num, .keep_all = TRUE) %>% 
+  count(area, treatment_assigned)
+at_home_n
+```

### 2020-10-30 main_analysis_initial.Rmd
data cleaning and analysis updates

--- 2020-10-27 main_analysis_initial.Rmd
@@
          ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
-         residents_under_five, home_bedrooms, pm_mean_sampling_period, pm_sample_interval,
+         residents_under_five, home_bedrooms, pm_mean_sampling_period, 
+         pm_sample_interval, pm_at_home_sampling_period,
          moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,
-         sums_events_sampling_period_5, sums_events_sampling_period_7.5) %>% 
+         sums_events_sampling_period_5, sums_events_sampling_period_7.5,
+         home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum) %>% 
   rename(treatment = newtx,
@@
 table_1_data <- health_exposure_data_sampling_day %>% 
-  select(-treatment_assigned, -treatment_actual) %>% 
-  left_join(treatments_masked, by = "home") %>%
-  rename(treatment = newtx) %>% 
+  mutate(treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education"))) %>% 
+    #select(-treatment_assigned, -treatment_actual) %>% 
+  #left_join(treatments_masked, by = "home") %>%
+  #rename(treatment = newtx) %>% 
   arrange(area, home, winter_id, sampling_day) %>% 
   filter(winter_id == 1) %>% 
-  filter(area == "NN") %>% # Change area filter as necessary
+  #filter(area == "NN") %>% # Change area filter as necessary
   distinct(area, home, .keep_all = TRUE)
@@
   rename(new_var = var) %>% 
-  group_by(treatment) %>% 
+  group_by(treatment_assigned) %>% 
   mutate(treatment_n = n()) %>% 
-  group_by(treatment, new_var) %>% 
+  group_by(treatment_assigned, new_var) %>% 
     mutate(n = n(),
@@
   select(new_var, n, percent) %>% 
-  arrange(treatment, new_var)
+  arrange(treatment_assigned, new_var)
 demographics_summary
@@
 char_funct_treatment("ruca_code_3")
+char_funct_total("home_act_smoking_sum")
+char_funct_treatment("home_act_smoking_sum")
+char_funct_total("home_act_windows_sum")
+char_funct_treatment("home_act_windows_sum")
+char_funct_total("home_act_sweep_sum")
+char_funct_treatment("home_act_sweep_sum")
+char_funct_total("home_act_door_sum")
+char_funct_treatment("home_act_door_sum")
 ```
@@
 table_1_data <- health_exposure_data_sampling_day %>% 
-  select(-treatment_assigned, -treatment_actual) %>% 
-  left_join(treatments_masked, by = "home") %>%
-  rename(treatment = newtx) %>% 
+  mutate(treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education"))) %>% 
+  #select(-treatment_assigned, -treatment_actual) %>% 
+  #left_join(treatments_masked, by = "home") %>%
+  #rename(treatment = newtx) %>% 
   arrange(area, home, winter_id, sampling_day) %>% 
   filter(winter_id == 1) %>% 
-  filter(area == "NN") %>% # Change area filter as necessary
+  #filter(area == "AK") %>% # Change area filter as necessary
   distinct(area, home, .keep_all = TRUE)
@@
   filter(!is.na(new_var)) %>% 
-  select(treatment, home, new_var) %>% 
+  select(treatment_assigned, home, new_var) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
   filter(!is.na(new_var)) %>% 
-  select(treatment, home, new_var) %>% 
-  group_by(treatment) %>% 
+  select(treatment_assigned, home, new_var) %>% 
+  group_by(treatment_assigned) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
 num_funct_treatment("sums_events_sampling_period_7.5")
+num_funct_total("home_act_smoking_sum")
+num_funct_treatment("home_act_smoking_sum")
+num_funct_total("home_act_windows_sum")
+num_funct_treatment("home_act_windows_sum")
+num_funct_total("home_act_sweep_sum")
+num_funct_treatment("home_act_sweep_sum")
+num_funct_total("home_act_door_sum")
+num_funct_treatment("home_act_door_sum")
+num_funct_total("home_act_incense_sum")
+num_funct_treatment("home_act_candle_sum")
 

### 2020-10-30 main_analysis_secondary_models.Rmd
data cleaning and analysis updates

@@
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  dplyr::select(area, home, child_id_num, at_home, at_home_location)
 ```
@@
                         burn_level_3level + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+
+itt_smoking_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_smoking_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+
+itt_windows_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_windows_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+
+itt_door_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_door_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+
+itt_sweep_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_sweep_sum + 
                         offset(log_person_time_at_risk), data = analysis_data)
@@
 itt_burn_level_tidy <- results_function(itt_burn_level_results, "ITT + burn level")
+itt_smoking_tidy <- results_function(itt_smoking_results, "ITT + smoke act")
+itt_windows_tidy <- results_function(itt_windows_results, "ITT + window act")
+itt_door_tidy <- results_function(itt_door_results, "ITT + door act")
+itt_sweep_tidy <- results_function(itt_sweep_results, "ITT + sweep act")
 itt_home_tidy <- results_function(itt_home_results, "ITT home only")
@@
 itt_results_combined <- rbind(itt_tidy, itt_income_tidy, itt_education_tidy, 
-                              itt_smoke_tidy, itt_burn_level_tidy, itt_home_tidy,
-                              itt_no_age_tidy, itt_outlier_tidy)
+                              itt_smoke_tidy, itt_burn_level_tidy, itt_smoking_tidy,
+                              itt_windows_tidy, itt_door_tidy, itt_sweep_tidy,
+                              itt_home_tidy, itt_no_age_tidy, itt_outlier_tidy)
 
@@
                         levels = c("ITT Primary", "ITT + income", "ITT + education",
-                                   "ITT + smoke hx", "ITT + burn level", "ITT home only",
-                                   "ITT - age", "ITT outlier"))) %>% 
+                                   "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
+                                   "ITT + window act", "ITT + door act", "ITT + sweep act",
+                                   "ITT home only", "ITT - age", "ITT outlier"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
           axis.text.x = element_text(size = 16, colour = "black", angle = 25,
-                                     hjust = .8, vjust = .8),
+                                     hjust = 1, vjust = .8),
           axis.text.y = element_text(size = 16, colour = "black"),

### 2020-10-30 stove_grade_usage_cleaning.Rmd
data cleaning and analysis updates

@@
 ```
+
+#####################################
+
+# Load, format, and KW compliance report
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+##### WMT #####
+# Load data
+kw_report_wmt <- read_csv(paste0(file_path, "Input/WMT/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         area = "WMT") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_report_wmt, paste0(file_path, "Output/kw_report_wmt.rds"))
+
+##### NN #####
+# Load data
+kw_report_nn <- read_csv(paste0(file_path, "Input/NN/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         area = "NN") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_report_nn, paste0(file_path, "Output/kw_report_nn.rds"))
+
+##### AK #####
+# Load data
+kw_report_ak <- read_csv(paste0(file_path, "Input/AK/filtrete_report.csv"), 
+                                  na = c("NULL", "")) %>% 
+  mutate(home_winter_id = as.factor(HomeWinterID),
+         home = as.factor(Home),
+         kw_date = ymd(ChangeDate),
+         kw_hours = as.numeric(KWHours),
+         kw_notes = as.character(Notes),
+         hours_possible = as.numeric(Hours),
+         kw_expected = as.numeric(Expected),
+         kw_perc_expected = as.numeric(PercExp),
+         area = "AK") %>% 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, 
+         hours_possible, kw_expected, kw_perc_expected) %>% 
+  #filter(!is.na(kw_hours)) %>% 
+  ungroup() %>% 
+  arrange(home_winter_id, kw_date)
+
+# Save as RDS
+write_rds(kw_report_ak, paste0(file_path, "Output/kw_report_ak.rds"))
+```
+
+# Join KW report data from different areas
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kw_report_wmt <- read_rds(paste0(file_path, "Output/kw_report_wmt.rds"))
+kw_report_nn <- read_rds(paste0(file_path, "Output/kw_report_nn.rds"))
+kw_report_ak <- read_rds(paste0(file_path, "Output/kw_report_ak.rds"))
+
+kids_kw_report <- rbind(kw_report_wmt, kw_report_nn, kw_report_ak) %>% 
+  dplyr::select(-home) %>% 
+  ungroup() %>% 
+  arrange(area, home_winter_id, kw_date)
+
+write_rds(kids_kw_report, paste0(file_path, "Output/kids_kw_report.rds"))
+```
+
+# Clean up all KW report data
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_kw_report <- read_rds(paste0(file_path, "Output/kids_kw_report.rds"))
+kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
+treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
+  mutate(home = gsub(" ", "", x = home),
+         home = gsub("_", "", x = home))
+
+kids_kw_report_clean <- kids_kw_report %>% 
+  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
+  arrange(area, home_winter_id)
+
+summary(kids_kw_report_clean)
+
+kw_summary <- kids_kw_report_clean %>% 
+  filter(!is.na(kw_perc_expected)) %>% 
+  #filter(area == "WMT") %>% 
+  #group_by(treatment_assigned) %>% 
+  summarize(mean(kw_perc_expected),
+            sd(kw_perc_expected),
+            n(),
+            min(kw_perc_expected),
+            median(kw_perc_expected),
+            max(kw_perc_expected))
+kw_summary
+
+# Some homes/stoves were graded during both winters, typically when new photos were added
+# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
+# We can't make the assumption that the stove was exactly the same for both winters,
+# so it will be best to use the grade corresponding to each winter for analysis.
+
+write_rds(kids_kw_report_clean, paste0(file_path, "Output/kids_kw_report_clean.rds"))
+```

### 2020-11-02 main_analysis_initial.Rmd
data cleaning/analysis updates

@@
 
-lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
+lrti_1obs_per_winter <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
+  arrange(area, home, child_id_num, winter_id) %>% 
+  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
+  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
+  group_by(area, child_id_num, winter_id) %>% 
+  mutate(lrti_events_winter = n(),
+         lrti_events_winter = as.numeric(lrti_events_winter),
+         lrti_events_winter = if_else(is.na(person_time_lrti), 0, lrti_events_winter),
+         person_time_lrti_winter = sum(person_time_lrti),
+         person_time_no_lrti_winter = as.numeric(person_time_possible - person_time_lrti_winter),
+         person_time_no_lrti_winter = if_else(is.na(person_time_no_lrti_winter), person_time_possible,
+                                             person_time_no_lrti_winter),
+         person_time_at_risk_winter = as.numeric(person_time_no_lrti_winter)/7,
+         lrti_incidence_winter = lrti_events_winter/person_time_at_risk_winter,
+         lrti_incidence_winter = if_else(is.na(lrti_incidence_winter), 0, lrti_incidence_winter)) %>% 
+  arrange(area, home, child_id_num, winter_id) %>%distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  ungroup() %>% 
+  dplyr::select(area, home, cohort, child_id_num, winter_id, 
+                lrti_events_winter, person_time_at_risk_winter, lrti_incidence_winter)
+
+lrti_1obs_total <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
   arrange(area, home, child_id_num, winter_id) %>% 
@@
   dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
-         amb_pm_lrti, person_time_start, person_time_end) %>% 
+         amb_pm_lrti, person_time_start, person_time_end, home_winter_id, winter_id) %>% 
   left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
@@
          amb_wind_lrti = if_else(child_id_num == 94, NaN, amb_wind_lrti)) %>% 
+  ungroup() %>% 
+  dplyr::select(area, home, cohort, child_id_num, winter_id,
+                lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
@@
   arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
-  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
+  right_join(lrti_1obs_total, by = c("area", "home", "child_id_num")) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
@@
           paste0(file_path, "Output/kids_analysis_data_1obs_per_child.csv"), na = "")
+
+kids_analysis_data_2obs_per_child <- health_exposure_data_sampling_day %>% 
+  #select(-treatment_assigned, -treatment_actual) %>% 
+  mutate(child_age = fct_collapse(age_child_nov,
+                                  "< 1" = "Less than 1",
+                                  "1+" = c(1, 2, 3, 4))) %>% 
+  left_join(treatments_masked, by = "home") %>% 
+  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
+  right_join(lrti_1obs_per_winter, by = c("area", "home", "child_id_num", "winter_id")) %>% 
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, 
+                child_age, lrti_events_winter, 
+         person_time_at_risk_winter, lrti_incidence_winter, newtx, treatment_assigned,
+         gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
+         home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
+         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
+         wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,
+         ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
+         residents_under_five, home_bedrooms, pm_mean_sampling_period, 
+         pm_sample_interval, pm_at_home_sampling_period,
+         moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,
+         sums_events_sampling_period_5, sums_events_sampling_period_7.5,
+         home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum) %>% 
+  rename(treatment = newtx) %>% 
+  mutate(log_person_time_at_risk_winter = log(person_time_at_risk_winter)) %>% 
+  mutate(lrti_events_di_winter = if_else(lrti_events_winter == 0, 0, 1)) %>% 
+  mutate(treatment = factor(treatment,
+                            levels = c("Blackfoot", "Dearborn", "Smith")),
+         treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education")))
+
+write_rds(kids_analysis_data_2obs_per_child, 
+          paste0(file_path, "Output/kids_analysis_data_2obs_per_child.rds"))
+write_csv(kids_analysis_data_2obs_per_child, 
+          paste0(file_path, "Output/kids_analysis_data_2obs_per_child.csv"), na = "")
 ```
@@
   filter(winter_id == 1) %>% 
-  #filter(area == "NN") %>% # Change area filter as necessary
+  filter(area == "AK") %>% # Change area filter as necessary
   distinct(area, home, .keep_all = TRUE)
@@
   distinct(new_var, .keep_all = TRUE) %>% 
-  select(new_var, n, percent) %>% 
+  dplyr::select(new_var, n, percent) %>% 
   arrange(new_var)
@@
   distinct(new_var, .keep_all = TRUE) %>% 
-  select(new_var, n, percent) %>% 
+  dplyr::select(new_var, n, percent) %>% 
   arrange(treatment_assigned, new_var)
@@
 
-count_summary <- analysis_data %>% 
-  group_by(treatment) %>% # Change treatment grouping as necessary 
+count_summary <- kids_analysis_data_1obs_per_child %>% 
+  #group_by(treatment) %>% # Change treatment grouping as necessary 
   #filter(lrti_incidence < 0.7) %>% 
@@
             max_events = max(lrti_events, na.rm = TRUE))
+count_summary
+
+count_summary <- kids_analysis_data_2obs_per_child %>% 
+  #group_by(treatment) %>% # Change treatment grouping as necessary 
+  #filter(lrti_incidence < 0.7) %>% 
+  summarize(mean_events = mean(lrti_events_winter, na.rm = TRUE),
+            sd_events = sd(lrti_events_winter, na.rm = TRUE),
+            n_events = n(),
+            min_events = min(lrti_events_winter, na.rm = TRUE),
+            median_events = median(lrti_events_winter, na.rm = TRUE),
+            max_events = max(lrti_events_winter, na.rm = TRUE))
 count_summary

### 2020-11-02 main_analysis_secondary_models.Rmd
data cleaning/analysis updates

@@
              position = position_dodge(width = 0.5), size = 4) +
-  #scale_shape_manual(values = c(15, 16, 17, 18, 13, 9)) +
   geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
@@
   theme_bw() +  
-  #ggtitle(label = "Systolic BP ITT models") +
   labs(y = "LRTI events per child-week at risk \n compared to placebo treatment") +
@@
 itt_plot_estimates
+
+ggsave("itt_plot_estimates.jpg", width = 12, height = 6)
 ```
@@
                                    burn_level_3level, data = analysis_data)
+
+itt_pm_smoking_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_smoking_sum, data = analysis_data)
+
+itt_pm_windows_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_windows_sum, data = analysis_data)
+
+itt_pm_door_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_door_sum, data = analysis_data)
+
+itt_pm_sweep_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_sweep_sum, data = analysis_data)
 
@@
 itt_pm_burn_level_tidy <- results_function(itt_pm_burn_level_results, "ITT + burn level")
+itt_pm_smoking_tidy <- results_function(itt_pm_smoking_results, "ITT + smoke act")
+itt_pm_windows_tidy <- results_function(itt_pm_windows_results, "ITT + window act")
+itt_pm_door_tidy <- results_function(itt_pm_door_results, "ITT + door act")
+itt_pm_sweep_tidy <- results_function(itt_pm_sweep_results, "ITT + sweep act")
 itt_pm_home_tidy <- results_function(itt_pm_home_results, "ITT home only")
@@
 itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_income_tidy, itt_pm_education_tidy, 
-                              itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_home_tidy,
-                              itt_pm_no_age_tidy, itt_pm_outlier_tidy)
+                              itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
+                              itt_pm_windows_tidy, itt_pm_door_tidy, itt_pm_sweep_tidy,
+                              itt_pm_home_tidy, itt_pm_no_age_tidy, itt_pm_outlier_tidy)
 
@@
   theme_bw() +  
-  #ggtitle(label = "Systolic BP ITT models") +
   labs(y = "Difference in PM2.5 \n compared to placebo") +
@@
 itt_pm_plot_estimates
+
+ggsave("itt_pm_plot_estimates.jpg", width = 12, height = 6)
 ```
+
+Sensitivity analyses for ER framework, LRTI outcome
+
+```{r}
+# Run model variations
+er_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_income_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_burn_level_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + burn_level_3level +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_smoking_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + home_act_smoking_sum + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_windows_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_windows_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_door_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_door_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_sweep_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_sweep_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_home_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_windows_sum +
+                          (1 | home) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_no_age_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+
+er_outlier_results <- glmer.nb(lrti_events ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data_outliers)
+
+# Run function to clean results
+results_function <- function(model_results, model_name) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = "PM2.5",
+         model = model_name) %>% 
+  dplyr::select(model, term, estimate, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+er_tidy <- results_function(er_results, "ER Primary")
+er_income_tidy <- results_function(er_income_results, "ER + income")
+er_burn_level_tidy <- results_function(er_burn_level_results, "ER + burn level")
+er_smoking_tidy <- results_function(er_smoking_results, "ER + smoke act")
+er_windows_tidy <- results_function(er_windows_results, "ER + window act")
+er_door_tidy <- results_function(er_door_results, "ER + door act")
+er_sweep_tidy <- results_function(er_sweep_results, "ER + sweep act")
+er_home_tidy <- results_function(er_home_results, "ER home only")
+er_no_age_tidy <- results_function(er_no_age_results, "ER - age")
+er_outlier_tidy <- results_function(er_outlier_results, "ER outlier")
+
+er_results_combined <- rbind(er_tidy, er_income_tidy, er_burn_level_tidy, 
+                              er_smoking_tidy, er_windows_tidy, er_door_tidy,
+                              er_sweep_tidy, er_home_tidy, er_no_age_tidy,
+                              er_outlier_tidy)
+
+# Plot results
+er_plot_estimates <- er_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("ER Primary", "ER + income", "ER + burn level",
+                                   "ER + smoke act", "ER + window act", "ER + door act",
+                                   "ER + sweep act", "ER home only", "ER - age",
+                                   "ER outlier"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  labs(y = "LRTI cases per IQR increase in PM2.5") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "none",
+          legend.text = element_text(size = 14, colour = "black")) 
+er_plot_estimates
+
+ggsave("er_plot_estimates.jpg", width = 12, height = 6)
+```
+
+Sensitivity analyses for LogReg framework, LRTI outcome
+
+```{r}
+# Run model variations
+logreg_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_income_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + income_3level +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_education_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + education_3level +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_smoke_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + residents_smoke +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_burn_level_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + burn_level_3level +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_smoking_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + home_act_smoking_sum +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_windows_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + home_act_windows_sum +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_door_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + home_act_door_sum +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_sweep_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + home_act_sweep_sum +
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_home_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + 
+                       (1 | home), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_no_age_results <- glmer(lrti_events_di ~ treatment_assigned +  
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+logreg_outlier_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), data = analysis_data_outliers, 
+                       family = binomial, offset = log_person_time_at_risk)
+
+# Run function to clean results
+results_function <- function(model_results, model_name) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_assigned", "", term),
+         model = model_name) %>% 
+  dplyr::select(model, term, estimate, p.value, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+logreg_tidy <- results_function(logreg_results, "LR Primary")
+logreg_income_tidy <- results_function(logreg_income_results, "LR + income")
+logreg_education_tidy <- results_function(logreg_education_results, "LR + education")
+logreg_smoke_tidy <- results_function(logreg_smoke_results, "LR + smoke hx")
+logreg_burn_level_tidy <- results_function(logreg_burn_level_results, "LR + burn level")
+logreg_smoking_tidy <- results_function(logreg_smoking_results, "LR + smoke act")
+logreg_windows_tidy <- results_function(logreg_windows_results, "LR + window act")
+logreg_door_tidy <- results_function(logreg_door_results, "LR + door act")
+logreg_sweep_tidy <- results_function(logreg_sweep_results, "LR + sweep act")
+logreg_home_tidy <- results_function(logreg_home_results, "LR home only")
+logreg_no_age_tidy <- results_function(logreg_no_age_results, "LR - age")
+logreg_outlier_tidy <- results_function(logreg_outlier_results, "LR outlier")
+
+logreg_results_combined <- rbind(logreg_tidy, logreg_income_tidy, logreg_education_tidy, 
+                              logreg_smoke_tidy, logreg_burn_level_tidy, logreg_smoking_tidy,
+                              logreg_windows_tidy, logreg_door_tidy, logreg_sweep_tidy,
+                              logreg_home_tidy, logreg_no_age_tidy, logreg_outlier_tidy)
+
+# Plot results
+logreg_plot_estimates <- logreg_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("LR Primary", "LR + income", "LR + education",
+                                   "LR + smoke hx", "LR + burn level", "LR + smoke act",
+                                   "LR + window act", "LR + door act", "LR + sweep act",
+                                   "LR home only", "LR - age", "LR outlier"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  labs(y = "Risk of LRTI event compared to Placebo") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+logreg_plot_estimates
+
+ggsave("logreg_plot_estimates.jpg", width = 12, height = 6)
+```

### 2020-11-10 main_analysis_initial.Rmd
main analysis updates

@@
 author: "Ethan Walker"
-date: "Started 23 Sept 2020, Updated 30 Sept 2020"
+date: "Started 23 Sept 2020, Updated 6 Nov 2020"
 output: pdf_document
@@
 
+# Condense LRTI data to 1 obs per child/winter and select analysis vars
 lrti_1obs_per_winter <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
@@
          lrti_incidence_winter = if_else(is.na(lrti_incidence_winter), 0, lrti_incidence_winter)) %>% 
-  arrange(area, home, child_id_num, winter_id) %>%distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  arrange(area, home, child_id_num, winter_id) %>%
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
   ungroup() %>% 
@@
 
+# Condense LRTI data to 1 obs per child and select analysis vars
 lrti_1obs_total <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
@@
          lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
-         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
+         lrti_events_total = sum(lrti_events, na.rm = TRUE),
          person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
@@
          person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
-         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
-         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
+         lrti_incidence_total = lrti_events_total/person_time_at_risk_total,
+         lrti_incidence_total = if_else(is.na(lrti_incidence_total), 0, lrti_incidence_total)) %>% 
   arrange(area, home, child_id_num, winter_id) %>% 
@@
   ungroup() %>% 
-  dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+  dplyr::select(area, home, cohort, child_id_num, lrti_events_total, person_time_at_risk_total, lrti_incidence_total,
          amb_pm_lrti, person_time_start, person_time_end, home_winter_id, winter_id) %>% 
@@
   dplyr::select(area, home, cohort, child_id_num, winter_id,
-                lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+                lrti_events_total, person_time_at_risk_total, lrti_incidence_total,
          amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti) %>% 
@@
   ungroup() %>% 
-  dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
+  dplyr::select(area, home, cohort, child_id_num, lrti_events_total, person_time_at_risk_total, lrti_incidence_total,
          amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti)
@@
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
-         person_time_at_risk_total, lrti_incidence, newtx, treatment_assigned,
-         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti,
+  dplyr::select(area, home, cohort, child_id_num, child_age, lrti_events_total, 
+         person_time_at_risk_total, lrti_incidence_total, newtx, treatment_assigned,
+         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti, gender,
          gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
          home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
-         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
+         stove_age_3level, chimney_clean_3level, wood_collect_method_2level, home_sqft_2level,
          wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,
@@
   rename(treatment = newtx,
-         lrti_events = lrti_events_overall,
          person_time_at_risk = person_time_at_risk_total) %>% 
   mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
-  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
+  mutate(lrti_events_di_total = if_else(lrti_events_total == 0, 0, 1)) %>% 
   mutate(treatment = factor(treatment,
@@
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, 
-                child_age, lrti_events_winter, 
-         person_time_at_risk_winter, lrti_incidence_winter, newtx, treatment_assigned,
+  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, child_age, gender,
+         lrti_events_winter, person_time_at_risk_winter, lrti_incidence_winter, newtx, treatment_assigned,
          gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
          home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
-         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
+         stove_age_3level, chimney_clean_3level, wood_collect_method_2level, home_sqft_2level,
          wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,

### 2020-11-10 main_analysis_secondary_models.Rmd
main analysis updates

--- 2020-10-28 main_analysis_secondary_models.Rmd
@@
 ---
-title: 'KidsAIR: Main analysis secondary models'
+title: 'KidsAIR: Main analysis'
 author: "Ethan Walker"
-date: "Started 28 Oct 2020, Updated 28 Oct 2020"
-output:
-  html_document:
+date: "Started 28 Oct 2020, Updated 9 Nov 2020"
+output: pdf_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, eval = TRUE, include = TRUE,
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                       message = FALSE, warning = FALSE)
@@
 
-```{r, eval=TRUE, include=TRUE}
+```{r}
 library(readxl)
@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25)
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+  #filter(gender == "Female")
+  filter(home_sqft_2level == "1500+")
 
@@
   mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk))
-  #filter(lrti_incidence < 0.7) %>% 
-  filter(filter_outliers == 0)
+
+analysis_data_winter1 <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+  filter(person_time_at_risk_winter > 0) %>% 
+  filter(winter_id == 1)
+
+analysis_data_winter2 <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+  filter(person_time_at_risk_winter > 0) %>% 
+  filter(winter_id == 2)
 
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  dplyr::select(area, home, child_id_num, at_home, at_home_location)
 ```
 
-Sensitivity analyses for ITT framework, LRTI outcome
+# Intent-to-treat framework, LRTI outcome
+
+```{r, echo=TRUE}
+# Run primary model and print results
+itt_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + 
+                        (1 | home:cohort:area) + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+#write_rds(itt_results, paste0(file_path, "Output/itt_results.rds"))
+
+summary(itt_results)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Run and save model results
+
+itt_winter1_results <- glmer.nb(lrti_events_winter ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        offset(log_person_time_at_risk_winter), data = analysis_data_winter1)
+write_rds(itt_winter1_results, paste0(file_path, "Output/itt_winter1_results.rds"))
+
+itt_winter2_results <- glmer.nb(lrti_events_winter ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        offset(log_person_time_at_risk_winter), data = analysis_data_winter2)
+write_rds(itt_winter2_results, paste0(file_path, "Output/itt_winter2_results.rds"))
+
+itt_income_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        income_3level + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_income_results, paste0(file_path, "Output/itt_income_results.rds"))
+
+itt_education_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        education_3level + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_education_results, paste0(file_path, "Output/itt_education_results.rds"))
+
+itt_smoke_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        residents_smoke + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_smoke_results, paste0(file_path, "Output/itt_smoke_results.rds"))
+
+itt_burn_level_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        burn_level_3level + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_burn_level_results, paste0(file_path, "Output/itt_burn_level_results.rds"))
+
+itt_smoking_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_smoking_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_smoking_results, paste0(file_path, "Output/itt_smoking_results.rds"))
+
+itt_windows_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_windows_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_windows_results, paste0(file_path, "Output/itt_windows_results.rds"))
+
+itt_door_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_door_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_door_results, paste0(file_path, "Output/itt_door_results.rds"))
+
+itt_sweep_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        home_act_sweep_sum + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_sweep_results, paste0(file_path, "Output/itt_sweep_results.rds"))
+
+itt_home_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home) + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_home_results, paste0(file_path, "Output/itt_home_results.rds"))
+
+itt_no_age_results <- glmer.nb(lrti_events_total ~ treatment_assigned + (1 | home:cohort:area) + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_no_age_results, paste0(file_path, "Output/itt_no_age_results.rds"))
+
+itt_outlier_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        offset(log_person_time_at_risk), data = analysis_data_outliers)
+write_rds(itt_outlier_results, paste0(file_path, "Output/itt_outlier_results.rds"))
+```
 
 ```{r}
-# Run model variations
-itt_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
-                        offset(log_person_time_at_risk), data = analysis_data)
+# Load results
+itt_results <- read_rds(paste0(file_path, "Output/itt_results.rds"))
+itt_winter1_results <- read_rds(paste0(file_path, "Output/itt_winter1_results.rds"))
+itt_winter2_results <- read_rds(paste0(file_path, "Output/itt_winter2_results.rds"))
+itt_income_results <- read_rds(paste0(file_path, "Output/itt_income_results.rds"))
+itt_education_results <- read_rds(paste0(file_path, "Output/itt_education_results.rds"))
+itt_smoke_results <- read_rds(paste0(file_path, "Output/itt_smoke_results.rds"))
+itt_burn_level_results <- read_rds(paste0(file_path, "Output/itt_burn_level_results.rds"))
+itt_smoking_results <- read_rds(paste0(file_path, "Output/itt_smoking_results.rds"))
+itt_windows_results <- read_rds(paste0(file_path, "Output/itt_windows_results.rds"))
+itt_door_results <- read_rds(paste0(file_path, "Output/itt_door_results.rds"))
+itt_sweep_results <- read_rds(paste0(file_path, "Output/itt_sweep_results.rds"))
+itt_home_results <- read_rds(paste0(file_path, "Output/itt_home_results.rds"))
+itt_no_age_results <- read_rds(paste0(file_path, "Output/itt_no_age_results.rds"))
+itt_outlier_results <- read_rds(paste0(file_path, "Output/itt_outlier_results.rds"))
 
-itt_income_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
-                        income_3level + 
-                        offset(log_person_time_at_risk), data = analysis_data)
-
-itt_education_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
-                        education_3level + 
-                        offset(log_person_time_at_risk), data = analysis_data)
-
-itt_smoke_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
-                        residents_smoke + 
-                        offset(log_person_time_at_risk), data = analysis_data)
-
-itt_burn_level_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
-                        burn_level_3level + 
-                        offset(log_person_time_at_risk), data = analysis_data)
-
-itt_home_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home) + 
-                        offset(log_person_time_at_risk), data = analysis_data)
-
-itt_no_age_results <- glmer.nb(lrti_events ~ treatment_assigned + (1 | home:cohort:area) + 
-                        offset(log_person_time_at_risk), data = analysis_data)
-
-itt_outlier_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
-                        offset(log_person_time_at_risk), data = analysis_data_outliers)
-
-# Run function to clean results
+# Function to clean results
 results_function <- function(model_results, model_name) {
@@
 itt_tidy <- results_function(itt_results, "ITT Primary")
+itt_winter1_tidy <- results_function(itt_winter1_results, "ITT Winter 1")
+itt_winter2_tidy <- results_function(itt_winter2_results, "ITT Winter 2")
 itt_income_tidy <- results_function(itt_income_results, "ITT + income")
@@
 
-itt_results_combined <- rbind(itt_tidy, itt_income_tidy, itt_education_tidy, 
-                              itt_smoke_tidy, itt_burn_level_tidy, itt_home_tidy,
-                              itt_no_age_tidy, itt_outlier_tidy)
+# Combine cleaned results
+itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, itt_winter2_tidy,
+                              itt_income_tidy, itt_education_tidy, 
+                              itt_smoke_tidy, itt_burn_level_tidy, itt_smoking_tidy,
+                              itt_windows_tidy, itt_door_tidy, itt_sweep_tidy,
+                              itt_home_tidy, itt_no_age_tidy, itt_outlier_tidy)
 
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT + income", "ITT + education",
-                                   "ITT + smoke hx", "ITT + burn level", "ITT home only",
-                                   "ITT - age", "ITT outlier"))) %>% 
+                        levels = c("ITT Primary", "ITT Winter 1", "ITT Winter 2", 
+                                   "ITT + income", "ITT + education",
+                                   "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
+                                   "ITT + window act", "ITT + door act", "ITT + sweep act",
+                                   "ITT home only", "ITT - age", "ITT outlier"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
   theme(title = element_text(size = 16), 
-          axis.text.x = element_text(size = 16, colour = "black", angle = 25,
-                                     hjust = .8, vjust = .8),
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
           axis.text.y = element_text(size = 16, colour = "black"),
@@
 itt_plot_estimates
+
+# Save plot as jpg
+ggsave("itt_plot_estimates.jpg", width = 12, height = 6)
 ```
 
-Sensitivity analyses for ITT framework, PM2.5 outcome
+\pagebreak
 
-```{r}
-# Run model variations
+# Intent-to-treat framework, PM2.5 outcome
+
+```{r, echo=TRUE}
+# Run primary model and print results
 itt_pm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
                        (1 | home:cohort:area), data = analysis_data)
+#write_rds(itt_pm_results, paste0(file_path, "Output/itt_pm_results.rds"))
+
+summary(itt_pm_results)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Run and save model results
+
+itt_pm_winter1_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), data = analysis_data_winter1)
+write_rds(itt_pm_winter1_results, paste0(file_path, "Output/itt_pm_winter1_results.rds"))
+
+#itt_pm_winter2_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
+#                       (1 | home:cohort:area), data = analysis_data_winter2)
+#write_rds(itt_pm_winter2_results, paste0(file_path, "Output/itt_pm_winter2_results.rds"))
 
@@
                                income_3level, data = analysis_data)
+write_rds(itt_pm_income_results, paste0(file_path, "Output/itt_pm_income_results.rds"))
 
@@
                                   education_3level, data = analysis_data)
+write_rds(itt_pm_education_results, paste0(file_path, "Output/itt_pm_education_results.rds"))
 
@@
                               residents_smoke, data = analysis_data)
+write_rds(itt_pm_smoke_results, paste0(file_path, "Output/itt_pm_smoke_results.rds"))
 
@@
                                    burn_level_3level, data = analysis_data)
+write_rds(itt_pm_burn_level_results, paste0(file_path, "Output/itt_pm_burn_level_results.rds"))
+
+itt_pm_smoking_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_smoking_sum, data = analysis_data)
+write_rds(itt_pm_smoking_results, paste0(file_path, "Output/itt_pm_smoking_results.rds"))
+
+itt_pm_windows_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_windows_sum, data = analysis_data)
+write_rds(itt_pm_windows_results, paste0(file_path, "Output/itt_pm_windows_results.rds"))
+
+itt_pm_door_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_door_sum, data = analysis_data)
+write_rds(itt_pm_door_results, paste0(file_path, "Output/itt_pm_door_results.rds"))
+
+itt_pm_sweep_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | home:cohort:area) + 
+                                   home_act_sweep_sum, data = analysis_data)
+write_rds(itt_pm_sweep_results, paste0(file_path, "Output/itt_pm_sweep_results.rds"))
 
@@
                              child_age + (1 | home), data = analysis_data)
+write_rds(itt_pm_home_results, paste0(file_path, "Output/itt_pm_home_results.rds"))
 
@@
                                (1 | home:cohort:area), data = analysis_data)
+write_rds(itt_pm_no_age_results, paste0(file_path, "Output/itt_pm_no_age_results.rds"))
 
@@
                                 child_age + (1 | home:cohort:area), data = analysis_data_outliers)
+write_rds(itt_pm_outlier_results, paste0(file_path, "Output/itt_pm_outlier_results.rds"))
+```
+
+```{r}
+# Load results
+itt_pm_results <- read_rds(paste0(file_path, "Output/itt_pm_results.rds"))
+itt_pm_winter1_results <- read_rds(paste0(file_path, "Output/itt_pm_winter1_results.rds"))
+#itt_pm_winter2_results <- read_rds(paste0(file_path, "Output/itt_pm_winter2_results.rds"))
+itt_pm_income_results <- read_rds(paste0(file_path, "Output/itt_pm_income_results.rds"))
+itt_pm_education_results <- read_rds(paste0(file_path, "Output/itt_pm_education_results.rds"))
+itt_pm_smoke_results <- read_rds(paste0(file_path, "Output/itt_pm_smoke_results.rds"))
+itt_pm_burn_level_results <- read_rds(paste0(file_path, "Output/itt_pm_burn_level_results.rds"))
+itt_pm_smoking_results <- read_rds(paste0(file_path, "Output/itt_pm_smoking_results.rds"))
+itt_pm_windows_results <- read_rds(paste0(file_path, "Output/itt_pm_windows_results.rds"))
+itt_pm_door_results <- read_rds(paste0(file_path, "Output/itt_pm_door_results.rds"))
+itt_pm_sweep_results <- read_rds(paste0(file_path, "Output/itt_pm_sweep_results.rds"))
+itt_pm_home_results <- read_rds(paste0(file_path, "Output/itt_pm_home_results.rds"))
+itt_pm_no_age_results <- read_rds(paste0(file_path, "Output/itt_pm_no_age_results.rds"))
+itt_pm_outlier_results <- read_rds(paste0(file_path, "Output/itt_pm_outlier_results.rds"))
 
@@
 itt_pm_tidy <- results_function(itt_pm_results, "ITT Primary")
+itt_pm_winter1_tidy <- results_function(itt_pm_winter1_results, "ITT Winter 1")
+#itt_pm_winter2_results <- results_function(itt_pm_winter2_results, "ITT Winter 2")
 itt_pm_income_tidy <- results_function(itt_pm_income_results, "ITT + income")
@@
 
-itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_income_tidy, itt_pm_education_tidy, 
-                              itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_home_tidy,
-                              itt_pm_no_age_tidy, itt_pm_outlier_tidy)
+# Combine cleaned results
+itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_winter1_tidy, 
+                                 itt_pm_income_tidy, itt_pm_education_tidy, 
+                              itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
+                              itt_pm_windows_tidy, itt_pm_door_tidy, itt_pm_sweep_tidy,
+                              itt_pm_home_tidy, itt_pm_no_age_tidy, itt_pm_outlier_tidy)
 
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT + income", "ITT + education",
-                                   "ITT + smoke hx", "ITT + burn level", "ITT home only",
-                                   "ITT - age", "ITT outlier"))) %>% 
+                        levels = c("ITT Primary", "ITT Winter 1", 
+                                   "ITT + income", "ITT + education",
+                                   "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
+                                   "ITT + window act", "ITT + door act", "ITT + sweep act",
+                                   "ITT home only", "ITT - age", "ITT outlier"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
 itt_pm_plot_estimates
+
+# Save plot as jpg
+ggsave("itt_pm_plot_estimates.jpg", width = 12, height = 6)
 ```
+
+\pagebreak
+
+# Exposure-response framework, LRTI outcome
+
+```{r, echo=TRUE}
+# Run primary model and print results
+er_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+#write_rds(er_results, paste0(file_path, "Output/er_results.rds"))
+
+summary(er_results)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Run and save model results
+
+er_winter1_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_winter1_results, paste0(file_path, "Output/er_winter1_results.rds"))
+
+#er_winter2_results <- glmer.nb(lrti_events_total_winter ~ pm_mean_sampling_period_iqr + child_age + 
+#                          education_3level + residents_smoke + 
+#                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+#                          data = analysis_data_winter2)
+#write_rds(er_winter2_results, paste0(file_path, "Output/er_winter2_results.rds"))
+
+er_income_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_income_results, paste0(file_path, "Output/er_income_results.rds"))
+
+er_burn_level_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + burn_level_3level +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_burn_level_results, paste0(file_path, "Output/er_burn_level_results.rds"))
+
+er_smoking_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + home_act_smoking_sum + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_smoking_results, paste0(file_path, "Output/er_smoking_results.rds"))
+
+er_windows_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_windows_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_windows_results, paste0(file_path, "Output/er_windows_results.rds"))
+
+er_door_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_door_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_door_results, paste0(file_path, "Output/er_door_results.rds"))
+
+er_sweep_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_sweep_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_sweep_results, paste0(file_path, "Output/er_sweep_results.rds"))
+
+er_home_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_windows_sum +
+                          (1 | home) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_home_results, paste0(file_path, "Output/er_home_results.rds"))
+
+er_no_age_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_no_age_results, paste0(file_path, "Output/er_no_age_results.rds"))
+
+er_outlier_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data_outliers)
+write_rds(er_outlier_results, paste0(file_path, "Output/er_outlier_results.rds"))
+```
+
+```{r}
+# Load results
+er_results <- read_rds(paste0(file_path, "Output/er_results.rds"))
+er_winter1_results <- read_rds(paste0(file_path, "Output/er_winter1_results.rds"))
+#er_winter2_results <- read_rds(paste0(file_path, "Output/er_winter2_results.rds"))
+er_income_results <- read_rds(paste0(file_path, "Output/er_income_results.rds"))
+er_burn_level_results <- read_rds(paste0(file_path, "Output/er_burn_level_results.rds"))
+er_smoking_results <- read_rds(paste0(file_path, "Output/er_smoking_results.rds"))
+er_windows_results <- read_rds(paste0(file_path, "Output/er_windows_results.rds"))
+er_door_results <- read_rds(paste0(file_path, "Output/er_door_results.rds"))
+er_sweep_results <- read_rds(paste0(file_path, "Output/er_sweep_results.rds"))
+er_home_results <- read_rds(paste0(file_path, "Output/er_home_results.rds"))
+er_no_age_results <- read_rds(paste0(file_path, "Output/er_no_age_results.rds"))
+er_outlier_results <- read_rds(paste0(file_path, "Output/er_outlier_results.rds"))
+
+# Run function to clean results
+results_function <- function(model_results, model_name) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("pm", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = "PM2.5",
+         model = model_name) %>% 
+  dplyr::select(model, term, estimate, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+er_tidy <- results_function(er_results, "ER Primary")
+er_winter1_tidy <- results_function(er_winter1_results, "ER Winter 1")
+#er_winter2_tidy <- results_function(er_winter2_results, "ER Winter 2")
+er_income_tidy <- results_function(er_income_results, "ER + income")
+er_burn_level_tidy <- results_function(er_burn_level_results, "ER + burn level")
+er_smoking_tidy <- results_function(er_smoking_results, "ER + smoke act")
+er_windows_tidy <- results_function(er_windows_results, "ER + window act")
+er_door_tidy <- results_function(er_door_results, "ER + door act")
+er_sweep_tidy <- results_function(er_sweep_results, "ER + sweep act")
+er_home_tidy <- results_function(er_home_results, "ER home only")
+er_no_age_tidy <- results_function(er_no_age_results, "ER - age")
+er_outlier_tidy <- results_function(er_outlier_results, "ER outlier")
+
+# Combine cleaned results
+er_results_combined <- rbind(er_tidy, er_winter1_tidy, 
+                             er_income_tidy, er_burn_level_tidy, 
+                              er_smoking_tidy, er_windows_tidy, er_door_tidy,
+                              er_sweep_tidy, er_home_tidy, er_no_age_tidy,
+                              er_outlier_tidy)
+
+# Plot results
+er_plot_estimates <- er_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("ER Primary", "ER Winter 1",
+                                   "ER + income", "ER + burn level",
+                                   "ER + smoke act", "ER + window act", "ER + door act",
+                                   "ER + sweep act", "ER home only", "ER - age",
+                                   "ER outlier"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  labs(y = "LRTI cases per IQR increase n\ in PM2.5") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "none",
+          legend.text = element_text(size = 14, colour = "black")) 
+er_plot_estimates
+
+# Save plot as jpg
+ggsave("er_plot_estimates.jpg", width = 12, height = 6)
+```
+
+\pagebreak
+
+# Logistic Regression framework, LRTI outcome
+
+```{r}
+# Run primary model and print results
+logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), data = analysis_data, 
+                       family = binomial, offset = log_person_time_at_risk)
+#write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
+
+summary(logreg_results)
+```

### 2020-11-10 stove_grade_usage_cleaning.Rmd
main analysis updates

@@
          kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
          area = "WMT") %>% 
-  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
          hours_possible, kw_expected, kw_perc_expected) %>% 
@@
          kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
          area = "NN") %>% 
-  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
          hours_possible, kw_expected, kw_perc_expected) %>% 
@@
          kw_perc_expected = as.numeric(PercExp),
+         filter_type = as.character(FType),
          area = "AK") %>% 
-  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, 
+  dplyr::select(area, home, home_winter_id, kw_date, kw_hours, kw_notes, filter_type, 
          hours_possible, kw_expected, kw_perc_expected) %>% 

### 2020-11-11 main_analysis_primary_models.Rmd
Analysis updates

@@
 
-model_results <- glmer(lrti_events_di ~ treatment_assigned + child_age + 
-                       (1 | home:cohort:area), data = analysis_data, 
-                       family = binomial, offset = log_person_time_at_risk)
+model_results <- glmer(lrti_events_di_total ~ treatment_assigned + child_age + 
+                       (1 | home:cohort:area), 
+                       data = analysis_data, family = binomial)
 
@@
   dplyr::select(term, estimate, conf.low, conf.high)
-#tidy_results
+tidy_results
 

### 2020-11-11 main_analysis_secondary_models.Rmd
Analysis updates

@@
 author: "Ethan Walker"
-date: "Started 28 Oct 2020, Updated 9 Nov 2020"
+date: "Started 28 Oct 2020, Updated 10 Nov 2020"
 output: pdf_document
@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) 
   #filter(gender == "Female")
-  filter(home_sqft_2level == "1500+")
+  #filter(home_sqft_2level == "1500+")
 
@@
 summary(itt_results)
+
+res <- residuals(itt_results, type="deviance")
+plot(predict(itt_results), res)
+abline(h=0, lty=2)
+qqnorm(res)
+qqline(res)
+halfnorm(residuals(itt_results))
 ```
@@
 itt_winter1_results <- read_rds(paste0(file_path, "Output/itt_winter1_results.rds"))
-itt_winter2_results <- read_rds(paste0(file_path, "Output/itt_winter2_results.rds"))
+#itt_winter2_results <- read_rds(paste0(file_path, "Output/itt_winter2_results.rds"))
 itt_income_results <- read_rds(paste0(file_path, "Output/itt_income_results.rds"))
@@
 itt_winter1_tidy <- results_function(itt_winter1_results, "ITT Winter 1")
-itt_winter2_tidy <- results_function(itt_winter2_results, "ITT Winter 2")
+#itt_winter2_tidy <- results_function(itt_winter2_results, "ITT Winter 2")
 itt_income_tidy <- results_function(itt_income_results, "ITT + income")
@@
 # Combine cleaned results
-itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, itt_winter2_tidy,
+itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, 
                               itt_income_tidy, itt_education_tidy, 
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT Winter 1", "ITT Winter 2", 
+                        levels = c("ITT Primary", "ITT Winter 1", 
                                    "ITT + income", "ITT + education",
@@
   theme_bw() +  
-  labs(y = "LRTI events per child-week at risk \n compared to placebo treatment") +
+  labs(y = "LRTI events per child-week \n at risk compared to placebo") +
   labs(x = "", group = "", shape = "") +
@@
   theme_bw() +  
-  labs(y = "Difference in PM2.5 \n compared to placebo") +
+  labs(y = "Difference in log(PM2.5) \n compared to placebo") +
   labs(x = "", group = "", shape = "") +
@@
   theme_bw() +  
-  labs(y = "LRTI cases per IQR increase n\ in PM2.5") +
+  labs(y = "LRTI cases per IQR \n increase in PM2.5") +
   labs(x = "", group = "", shape = "") +
@@
 
-```{r}
+```{r, echo=TRUE}
 # Run primary model and print results
@@
                        (1 | home:cohort:area), data = analysis_data, 
-                       family = binomial, offset = log_person_time_at_risk)
+                       family = binomial)
 #write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
@@
 summary(logreg_results)
+
+at_risk_summary <- analysis_data %>% 
+  group_by(treatment_assigned) %>% 
+  summarize(mean(person_time_at_risk, na.rm = TRUE),
+            sd(person_time_at_risk, na.rm = TRUE),
+            median(person_time_at_risk, na.rm = TRUE))
+at_risk_summary
 ```

### 2020-11-13 main_analysis_initial.Rmd
analysis updates

@@
   mutate(sample_date = ymd(temp_date))
+
+kids_kw_report_1obs <- read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
+  group_by(home) %>% 
+  mutate(mean_kw_per_exp = mean(kw_perc_expected, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  mutate(kw_perc_exp_cat = cut(kw_perc_expected, 
+                               breaks = c(0, 42, 83, 129, 50000),
+                               labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
+  dplyr::select(-home_winter_id, -treatment_assigned, -treatment_actual, -home_id)
+
+kids_kw_report_2obs <- read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
+  group_by(area, home_winter_id) %>% 
+  mutate(mean_kw_per_exp = mean(kw_perc_expected, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE) %>% 
+  mutate(kw_perc_exp_cat = cut(kw_perc_expected, 
+                               breaks = c(0, 37, 74, 118, 50000),
+                               labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
+  dplyr::select(-home_winter_id, -treatment_assigned, -treatment_actual, -home_id)
 
@@
   right_join(lrti_1obs_total, by = c("area", "home", "child_id_num")) %>% 
+  left_join(kids_kw_report_1obs, by = c("area", "home")) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
@@
          sums_events_sampling_period_5, sums_events_sampling_period_7.5,
-         home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum) %>% 
+         home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum,
+         filter_type, mean_kw_per_exp, kw_perc_exp_cat) %>% 
   rename(treatment = newtx,
@@
   mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
-  mutate(lrti_events_di_total = if_else(lrti_events_total == 0, 0, 1)) %>% 
+  mutate(lrti_events_di_total = if_else(lrti_events_total == 0, 0, 1),
+         filter_type2 = as.character(filter_type),
+         filter_type2 = if_else(treatment_assigned == "Education", "Ed Home", filter_type2),
+         filter_type2 = factor(filter_type2,
+                              levels = c("Honeywell", "Large Filtrete", "Small Filtrete",
+                                         "Winix", "Ed Home")),
+         kw_perc_exp_cat2 = as.character(kw_perc_exp_cat),
+         kw_perc_exp_cat2 = if_else(treatment_assigned == "Education", "Ed Home", kw_perc_exp_cat2),
+         kw_perc_exp_cat2 = factor(kw_perc_exp_cat2,
+                                  levels = c("Q1", "Q2", "Q3", "Q4", "Ed Home"))) %>% 
   mutate(treatment = factor(treatment,
@@
   right_join(lrti_1obs_per_winter, by = c("area", "home", "child_id_num", "winter_id")) %>%  
+  left_join(kids_kw_report_1obs, by = c("area", "home", "winter_id")) %>% 
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
@@
          sums_events_sampling_period_5, sums_events_sampling_period_7.5,
-         home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum) %>% 
+         home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum,
+         filter_type, mean_kw_per_exp, kw_perc_exp_cat) %>% 
   rename(treatment = newtx) %>% 
   mutate(log_person_time_at_risk_winter = log(person_time_at_risk_winter)) %>% 
-  mutate(lrti_events_di_winter = if_else(lrti_events_winter == 0, 0, 1)) %>% 
+  mutate(lrti_events_di_winter = if_else(lrti_events_winter == 0, 0, 1),
+         filter_type2 = as.character(filter_type),
+         filter_type2 = if_else(treatment_assigned == "Education", "Ed Home", filter_type2),
+         filter_type2 = factor(filter_type2,
+                              levels = c("Honeywell", "Large Filtrete", "Small Filtrete",
+                                         "Winix", "Ed Home")),
+         kw_perc_exp_cat2 = as.character(kw_perc_exp_cat),
+         kw_perc_exp_cat2 = if_else(treatment_assigned == "Education", "Ed Home", kw_perc_exp_cat2),
+         kw_perc_exp_cat2 = factor(kw_perc_exp_cat2,
+                                  levels = c("Q1", "Q2", "Q3", "Q4", "Ed Home"))) %>% 
   mutate(treatment = factor(treatment,

### 2020-11-13 main_analysis_secondary_models.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Started 28 Oct 2020, Updated 10 Nov 2020"
+date: "Updated 12 Nov 2020"
 output: pdf_document
@@
 library(faraway)
+library(DHARMa)
 ```
@@
   filter(winter_id == 2)
-
-health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
-  dplyr::select(area, home, child_id_num, at_home, at_home_location)
 ```
@@
 summary(itt_results)
+```
 
-res <- residuals(itt_results, type="deviance")
-plot(predict(itt_results), res)
-abline(h=0, lty=2)
-qqnorm(res)
-qqline(res)
-halfnorm(residuals(itt_results))
+## Diagnostic plots
+```{r}
+#halfnorm(residuals(itt_results))
+
+itt_res <- simulateResiduals(itt_results)
+plot(itt_res, rank = FALSE, quantreg = TRUE)
 ```
+
+The function creates two plots.
+To the left, a qq-uniform plot to detect deviations from overall uniformity of the residuals (calling plotQQunif), and to the right, a plot of residuals against predicted values (calling plotResiduals).
+Outliers are highlighted in red (for more on outliers, see testOutliers).
+For a correctly specified model, we would expect
+
+a) a straight 1-1 line in the uniform qq-plot -> evidence for an overall uniform (flat) distribution of the residuals
+
+b) uniformity of residuals in the vertical direction in the res against predictor plot
+
+Deviations of this can be interpreted as for a linear regression. See the vignette for detailed examples.
+
+To provide a visual aid in detecting deviations from uniformity in y-direction, the plot of the residuals against the predicted values also performs an (optional) quantile regression, which provides 0.25, 0.5 and 0.75 quantile lines across the plots.
+These lines should be straight, horizontal, and at y-values of 0.25, 0.5 and 0.75.
+Note, however, that some deviations from this are to be expected by chance, even for a perfect model, especially if the sample size is small.
 
@@
 write_rds(itt_education_results, paste0(file_path, "Output/itt_education_results.rds"))
+
+itt_filter_type_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        filter_type +
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_filter_type_results, paste0(file_path, "Output/itt_filter_type_results.rds"))
+
+itt_filter_comp_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
+                        kw_perc_exp_cat2 + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+write_rds(itt_filter_comp_results, paste0(file_path, "Output/itt_filter_comp_results.rds"))
 
@@
 itt_education_results <- read_rds(paste0(file_path, "Output/itt_education_results.rds"))
+#itt_filter_type_results <- read_rds(paste0(file_path, "Output/itt_filter_type_results.rds"))
+itt_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_filter_comp_results.rds"))
 itt_smoke_results <- read_rds(paste0(file_path, "Output/itt_smoke_results.rds"))
@@
 itt_education_tidy <- results_function(itt_education_results, "ITT + education")
+#itt_filter_type_tidy <- results_function(itt_filter_type_results, "ITT + filt type")
+itt_filter_comp_tidy <- results_function(itt_filter_comp_results, "ITT + filt comp")
 itt_smoke_tidy <- results_function(itt_smoke_results, "ITT + smoke hx")
@@
 itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, 
-                              itt_income_tidy, itt_education_tidy, 
+                              itt_income_tidy, itt_education_tidy,  itt_filter_comp_tidy,
                               itt_smoke_tidy, itt_burn_level_tidy, itt_smoking_tidy,
@@
                         levels = c("ITT Primary", "ITT Winter 1", 
-                                   "ITT + income", "ITT + education",
+                                   "ITT + income", "ITT + education", "ITT + filt comp",
                                    "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
@@
   theme_bw() +  
+  ggtitle(label = "Results for ITT framework, LRTI outcome") +
   labs(y = "LRTI events per child-week \n at risk compared to placebo") +
@@
 
+ITT Winter 1 = Primary model with Winter 1 data only  
+ITT + income = Primary model plus categorical income variable  
+ITT + education = Primary model plus categorical education variable  
+ITT + filt comp = Primary model plus variable for filter compliance  
+ITT + smoke hx = Primary model plus self-reported smoking variable  
+ITT + burn level = Primary model plus self-reported burn-level compared to typical  
+ITT + smoke act = Primary model plus # days during sampling with smoking  
+ITT + window act = Primary model plus # days during sampling with windows open  
+ITT + door act = Primary model plus # days during sampling with doors open  
+ITT + sweep act = Primary model plus # days during sampling with sweeping/cleaning  
+ITT + home = Primary model with only home as random term  
+ITT - age = Primary model without age variable  
+ITT outlier = Primary model with data that fixes incidence outlier
+
 \pagebreak
@@
 summary(itt_pm_results)
+```
+
+## Diagnostic plots
+```{r}
+plot(itt_pm_results, main = "PM ITT Primary model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_pm_results), main = "PM ITT Primary model")
+qqline(residuals(itt_pm_results))
+
+#itt_pm_res <- simulateResiduals(itt_pm_results)
+#plot(itt_pm_res, rank = TRUE, quantreg = TRUE)
 ```
@@
 write_rds(itt_pm_education_results, paste0(file_path, "Output/itt_pm_education_results.rds"))
+
+itt_pm_filter_type_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                  child_age + (1 | home:cohort:area) + 
+                                  filter_type, data = analysis_data)
+write_rds(itt_pm_filter_type_results, paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
+
+itt_pm_filter_comp_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                  child_age + (1 | home:cohort:area) + 
+                                  kw_perc_exp_cat2, data = analysis_data)
+write_rds(itt_pm_filter_comp_results, paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
 
@@
 itt_pm_education_results <- read_rds(paste0(file_path, "Output/itt_pm_education_results.rds"))
+#itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
+itt_pm_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
 itt_pm_smoke_results <- read_rds(paste0(file_path, "Output/itt_pm_smoke_results.rds"))
@@
 itt_pm_education_tidy <- results_function(itt_pm_education_results, "ITT + education")
+#itt_pm_filter_type_tidy <- results_function(itt_pm_filter_type_results, "ITT + filt type")
+itt_pm_filter_comp_tidy <- results_function(itt_pm_filter_comp_results, "ITT + filt comp")
 itt_pm_smoke_tidy <- results_function(itt_pm_smoke_results, "ITT + smoke hx")
@@
 itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_winter1_tidy, 
-                                 itt_pm_income_tidy, itt_pm_education_tidy, 
+                                 itt_pm_income_tidy, itt_pm_education_tidy, itt_pm_filter_comp_tidy,
                               itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
@@
   theme_bw() +  
+  ggtitle(label = "Results for ITT framework, PM outcome") +
   labs(y = "Difference in log(PM2.5) \n compared to placebo") +
@@
 summary(er_results)
+```
+
+## Diagnostic plots
+```{r}
+
+er_res <- simulateResiduals(er_results)
+plot(er_res, rank = TRUE, quantreg = TRUE)
 ```
@@
 write_rds(er_income_results, paste0(file_path, "Output/er_income_results.rds"))
+
+er_filter_type_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + filter_type +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_filter_type_results, paste0(file_path, "Output/er_filter_type_results.rds"))
+
+er_filter_comp_results <- glmer.nb(lrti_events_total ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + kw_perc_exp_cat2 +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          data = analysis_data)
+write_rds(er_filter_comp_results, paste0(file_path, "Output/er_filter_comp_results.rds"))
 
@@
 er_income_results <- read_rds(paste0(file_path, "Output/er_income_results.rds"))
+#er_filter_type_results <- read_rds(paste0(file_path, "Output/er_filter_type_results.rds"))
+er_filter_comp_results <- read_rds(paste0(file_path, "Output/er_filter_comp_results.rds"))
 er_burn_level_results <- read_rds(paste0(file_path, "Output/er_burn_level_results.rds"))
@@
 er_income_tidy <- results_function(er_income_results, "ER + income")
+#er_filter_type_tidy <- results_function(er_filter_type_results, "ER + filt type")
+er_filter_comp_tidy <- results_function(er_filter_comp_results, "ER + filt comp")
 er_burn_level_tidy <- results_function(er_burn_level_results, "ER + burn level")
@@
 er_results_combined <- rbind(er_tidy, er_winter1_tidy, 
-                             er_income_tidy, er_burn_level_tidy, 
+                             er_income_tidy, er_filter_comp_tidy, er_burn_level_tidy, 
                               er_smoking_tidy, er_windows_tidy, er_door_tidy,
@@
                         levels = c("ER Primary", "ER Winter 1",
-                                   "ER + income", "ER + burn level",
+                                   "ER + income", "ER + filt comp", "ER + burn level",
                                    "ER + smoke act", "ER + window act", "ER + door act",
@@
   theme_bw() +  
+  ggtitle(label = "Results for ER framework, LRTI outcome") +
   labs(y = "LRTI cases per IQR \n increase in PM2.5") +
@@
 
+ER Winter 1 = Primary model with Winter 1 data only  
+ER + income = Primary model using income instead of education  
+ER + filt comp = Primary model plus variable for filter compliance  
+ER + burn level = Primary model plus self-reported burn-level compared to typical  
+ER + smoke act = Primary model plus # days during sampling with smoking  
+ER + window act = Primary model plus # days during sampling with windows open  
+ER + door act = Primary model plus # days during sampling with doors open  
+ER + sweep act = Primary model plus # days during sampling with sweeping/cleaning  
+ER + home = Primary model with only home as random term  
+ER - age = Primary model without age variable  
+ER outlier = Primary model with data that fixes incidence outlier
+
 \pagebreak

### 2020-11-13 stove_grade_usage_cleaning.Rmd
analysis updates

@@
   left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
-  arrange(area, home_winter_id)
+  arrange(area, home_winter_id) %>% 
+  mutate(filter_type = fct_collapse(filter_type,
+                                    "Honeywell" = "Honeywell",
+                                    "Large Filtrete" = "Large Filtrete",
+                                    "Small Filtrete" = "Small Filtrete",
+                                    "Winix" = c("Winix", "winix"))) %>% 
+  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
+         kw_perc_expected = abs(kw_perc_expected)) %>% 
+  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
+  mutate(kw_perc_expected = as.numeric(kw_perc_expected))
 
-summary(kids_kw_report_clean)
+#summary(kids_kw_report_clean)
 
-kw_summary <- kids_kw_report_clean %>% 
+kw_perc_exp_summary <- kids_kw_report_clean %>% 
   filter(!is.na(kw_perc_expected)) %>% 
   #filter(area == "WMT") %>% 
-  #group_by(treatment_assigned) %>% 
-  summarize(mean(kw_perc_expected),
-            sd(kw_perc_expected),
-            n(),
-            min(kw_perc_expected),
-            median(kw_perc_expected),
-            max(kw_perc_expected))
-kw_summary
+  group_by(filter_type) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_expected),
+            "SD KW % Exp" = sd(kw_perc_expected),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_expected),
+            "25Q KW % Exp" = quantile(kw_perc_expected, 0.25),
+            "Med KW % Exp" = median(kw_perc_expected),
+            "75Q KW % Exp" = quantile(kw_perc_expected, 0.75),
+            "Max KW % Exp" = max(kw_perc_expected))
+kw_perc_exp_summary
 
-# Some homes/stoves were graded during both winters, typically when new photos were added
-# Since photos were added in these cases, grades may have changed from Winter 1 to Winter 2
-# We can't make the assumption that the stove was exactly the same for both winters,
-# so it will be best to use the grade corresponding to each winter for analysis.
+boxplot(log(kw_perc_expected) ~ filter_type, data = kids_kw_report_clean)
+
+boxplot(kw_perc_expected ~ filter_type, data = kids_kw_report_clean)
+
+kw_exp_summary <- kids_kw_report_clean %>% 
+  filter(!is.na(kw_expected)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type) %>% 
+  summarize("Mean KW Exp" = mean(kw_expected),
+            "SD KW Exp" = sd(kw_expected),
+            "N KW Exp" = n(),
+            "Min KW Exp" = min(kw_expected),
+            "25Q KW Exp" = quantile(kw_expected, 0.25),
+            "Med KW Exp" = median(kw_expected),
+            "75Q KW Exp" = quantile(kw_expected, 0.75),
+            "Max KW Exp" = max(kw_expected))
+kw_exp_summary
+
+fivenum(kids_kw_report_clean$kw_perc_expected)
 

### 2020-11-16 main_analysis_initial.Rmd
small analysis updates new analysis for interaction terms

@@
 
-count_summary <- kids_analysis_data_1obs_per_child %>% 
+count_summary <- analysis_data %>% 
   #group_by(treatment) %>% # Change treatment grouping as necessary 
   #filter(lrti_incidence < 0.7) %>% 
-  summarize(mean_events = mean(lrti_events, na.rm = TRUE),
-            sd_events = sd(lrti_events, na.rm = TRUE),
+  summarize(mean_events = mean(lrti_events_total, na.rm = TRUE),
+            sd_events = sd(lrti_events_total, na.rm = TRUE),
             n_events = n(),
-            min_events = min(lrti_events, na.rm = TRUE),
-            median_events = median(lrti_events, na.rm = TRUE),
-            max_events = max(lrti_events, na.rm = TRUE))
+            min_events = min(lrti_events_total, na.rm = TRUE),
+            median_events = median(lrti_events_total, na.rm = TRUE),
+            max_events = max(lrti_events_total, na.rm = TRUE))
 count_summary
@@
   ggplot() +
-    geom_histogram(aes(lrti_events)) +
+    geom_histogram(aes(lrti_events_total)) +
     theme_light() +
@@
 # Simple model with offset
-nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), 
+nb_results <- glm.nb(lrti_events_total ~ treatment + offset(log_person_time_at_risk), 
                      data = analysis_data)
@@
 
-poisson_results <- glm(lrti_events ~ treatment + offset(log_person_time_at_risk), 
+poisson_results <- glm(lrti_events_total ~ treatment + offset(log_person_time_at_risk), 
                        family = "poisson", data = analysis_data)

### 2020-11-16 main_analysis_secondary_models.Rmd
small analysis updates new analysis for interaction terms

@@
 author: "Ethan Walker"
-date: "Updated 12 Nov 2020"
+date: "Updated 13 Nov 2020"
 output: pdf_document
@@
 library(lubridate)
+library(emmeans)
 library(broom)
@@
 plot(itt_res, rank = FALSE, quantreg = TRUE)
+
+plot(itt_results, main = "ITT Primary model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_results), main = "ITT Primary model")
+qqline(residuals(itt_results))
 ```
@@
 Note, however, that some deviations from this are to be expected by chance, even for a perfect model, especially if the sample size is small.
+
+```{r, eval=FALSE, include=FALSE}
+# Effect modification/interaction analysis
+# Try child_age, gender, home_floors_2level, home_sqft_2level
+
+# Run primary model and print results
+itt_results_int <- glmer.nb(lrti_events_total ~ treatment_assigned*child_age + 
+#                                  child_age +
+                        (1 | home:cohort:area) + 
+                        offset(log_person_time_at_risk), data = analysis_data)
+#write_rds(itt_results_int, paste0(file_path, "Output/itt_results_int_age.rds"))
+
+summary(itt_results_int)
+
+int_emmeans <- emmeans(itt_results_int, pairwise ~ treatment_assigned | child_age)
+summary(int_emmeans, conf.int = TRUE)
+confint(int_emmeans)
+# This line will plot the interaction term
+#emmip(itt_results_int, child_age ~ treatment_assigned, cov.reduce = range)
+table(model.frame(itt_results_int)$treatment_assigned)
+nobs(itt_results_int)
+```
 
@@
 ## Diagnostic plots
+### The correlation in the residuals vs fitted plot is due to the random term for "home"
 ```{r}
@@
 #plot(itt_pm_res, rank = TRUE, quantreg = TRUE)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Effect modification/interaction analysis
+# Try child_age, gender, home_floors_2level, home_sqft_2level
+
+# Run primary model and print results
+itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*home_sqft_2level + 
+                              child_age + 
+                       (1 | home:cohort:area), data = analysis_data)
+#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+
+summary(itt_pm_results_int)
+
+int_emtrends <- emtrends(itt_pm_results_int, home_sqft_2level ~ treatment_assigned, var = "log(pm_mean_sampling_period)")
+summary(int_emtrends, conf.int = TRUE)
+# This line will plot the interaction term
+#emmip(itt_pm_results_int, child_age ~ treatment_assigned, cov.reduce = range)
+table(model.frame(itt_pm_results_int)$treatment_assigned)
+nobs(itt_pm_results_int)
 ```
@@
 summary(logreg_results)
+```
 
+## At-risk period across treatment arms
+```{r}
 at_risk_summary <- analysis_data %>% 
   group_by(treatment_assigned) %>% 
-  summarize(mean(person_time_at_risk, na.rm = TRUE),
-            sd(person_time_at_risk, na.rm = TRUE),
-            median(person_time_at_risk, na.rm = TRUE))
+  summarize("Mean at-risk time" = mean(person_time_at_risk, na.rm = TRUE),
+            "SD at-risk time" = sd(person_time_at_risk, na.rm = TRUE),
+            "Med at-risk time" = median(person_time_at_risk, na.rm = TRUE))
 at_risk_summary

### 2020-11-18 main_analysis_final.Rmd
major model updates new RMD for final analysis

---
title: 'KidsAIR: Main analysis'
author: "Ethan Walker"
date: "Updated 17 Nov 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = TRUE,
                      message = FALSE, warning = FALSE)
```

```{r}
library(readxl)
library(naniar)
library(lubridate)
library(emmeans)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
library(MASS)
library(faraway)
library(DHARMa)
```

```{r}
# Load analysis dataset

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_total = as.factor(lrti_events_di_total),
         person_time_at_risk_iqr = person_time_at_risk/20) 
  #filter(gender == "Female")
  #filter(home_sqft_2level == "1500+")

analysis_data_outliers <- analysis_data %>% 
  #rownames_to_column() %>% 
  #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
  mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk),
         lrti_events_total = as.factor(lrti_events_total),
         lrti_events_di_total = as.factor(lrti_events_di_total))

analysis_data_winter1 <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
  filter(person_time_at_risk_winter > 0) %>% 
  filter(winter_id == 1)

analysis_data_winter2 <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
  filter(person_time_at_risk_winter > 0) %>% 
  filter(winter_id == 2)
```

# Options for primary model framework

## Poisson model
```{r, echo=TRUE, eval=FALSE}
poisson_results <- glmer(lrti_events_total ~ treatment_assigned + child_age + 
                         (1 | home:cohort:area), 
                         data = analysis_data, family = poisson(link = "log"), 
                         offset = log_person_time_at_risk)
#write_rds(poisson_results, paste0(file_path, "Output/poisson_results.rds"))
```

```{r}
poisson_results <- read_rds(paste0(file_path, "Output/poisson_results.rds"))
summary(poisson_results)
#tidy(poisson_results, conf.int = TRUE)

poisson_simres <- simulateResiduals(poisson_results)
testDispersion(poisson_simres, alternative = "less")
plot(poisson_simres, rank = FALSE, quantreg = TRUE)
```

## Negative binomial model
```{r, echo=TRUE, eval=FALSE}
negbin_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + 
                           (1 | home:cohort:area) + 
                           offset(log_person_time_at_risk), 
                           data = analysis_data)
#write_rds(negbin_results, paste0(file_path, "Output/negbin_results.rds"))
```

```{r}
negbin_results <- read_rds(paste0(file_path, "Output/negbin_results.rds"))
summary(negbin_results)
#tidy(negbin_results, conf.int = TRUE)

negbin_simres <- simulateResiduals(negbin_results)
testDispersion(negbin_simres, alternative = "less")
plot(negbin_simres, rank = FALSE, quantreg = TRUE)
```

## Logistic regression model
```{r, echo=TRUE, eval=FALSE}
logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
                        child_age + person_time_at_risk +
                        (1 | home:cohort:area),
                        data = analysis_data, family = binomial, nAGQ = 20)
#write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
```

```{r}
logreg_results <- read_rds(paste0(file_path, "Output/logreg_results.rds"))
summary(logreg_results)
#tidy(logreg_results, conf.int = TRUE)

logreg_simres <- simulateResiduals(logreg_results)
testDispersion(logreg_simres, alternative = "two.sided")
plot(logreg_simres, rank = FALSE, quantreg = TRUE)
```

```{r}
# Function to clean results
results_function <- function(model_results, model_name) {

tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_assigned", "", term),
         model = model_name) %>% 
  dplyr::select(model, term, estimate, p.value, conf.low, conf.high)
tidy_results

}

# Run results through function, save, combine for plotting
poisson_tidy <- results_function(poisson_results, "Poisson")
negbin_tidy <- results_function(negbin_results, "Neg Bin")
logreg_tidy <- results_function(logreg_results, "Log Reg")

# Combine cleaned results
itt_results_combined <- rbind(poisson_tidy, negbin_tidy, logreg_tidy)

# Plot results
itt_plot_estimates <- itt_results_combined %>%
  mutate(model = factor(model,
                        levels = c("Poisson", "Neg Bin", "Log Reg"))) %>% 
  ggplot(aes(group = term, shape = term)) +
  geom_point(aes(x=model, y=estimate), 
             position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
  geom_hline(yintercept = 0) +
  theme_bw() +  
  ggtitle(label = "Results for ITT framework, LRTI outcome") +
  labs(y = "Estimate compared to placebo") +
  labs(x = "", group = "", shape = "") +
  theme(title = element_text(size = 16), 
          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
                                     hjust = 1, vjust = .8),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(),
          legend.position = "top",
          legend.text = element_text(size = 14, colour = "black")) 
itt_plot_estimates

# Save plot as jpg
#ggsave("itt_plot_estimates.jpg", width = 12, height = 6)
```

# Interpretation:
As Jon and I discussed 11/16/20, the Poisson and Negative Binomial models appear to be
underdispersed (variance < mean), not overdispersed (variance > mean). These models 
don't appear to be the best fit for our data. Additionally, results from the models 
were essentially the same.  
The Logistic Regression model appears to be a better fit for the data (no concerns 
in the diagnostic plots). The good news is that the results were also very similar to 
the Poisson/NegBin models, so we don't have to explain any big differences in results 
if model selection concerns come up with reviewers. The bad news is that regardless 
of model, there is no difference in LRTI between treatment arms.   
Given these results, I'm leaning towards using the logistic regression model for the 
primary ITT framework. There are models that can be used for underdispersed count 
data (generalized Poisson regression?), but I can't find a good way to run those 
as mixed models like we need to do. I think the logistic regression model will be 
easier to explain/interpret, and we aren't really losing much information since there 
were so few children with 2+ cases of LRTI.

### 2020-11-18 main_analysis_initial.Rmd
major model updates new RMD for final analysis

@@
 library(kableExtra)
+library(AER)
+library(DHARMa)
 ```
@@
 
-poisson_results <- glm(lrti_events_total ~ treatment + offset(log_person_time_at_risk), 
+poisson_results <- glm(lrti_events_total ~ treatment_assigned + offset(log_person_time_at_risk), 
                        family = "poisson", data = analysis_data)
@@
 library(AER)
-dispersiontest(poisson_results, trafo=1) # test indicates poisson model is overdispersed
+dispersiontest(poisson_results, alternative = "greater") # test indicates poisson model is overdispersed
+
+itt_res <- simulateResiduals(poisson_results)
+testDispersion(itt_res, alternative = "less")
 
@@
 ```{r}
-# Mixed model variations
+# Mixed model option for primary model
 
-poisson_results <- glmer(lrti_events ~ treatment + child_age + (1 | home:cohort:area), 
-                         data = analysis_data, family = poisson(link = "log"), offset = log_person_time_at_risk)
+poisson_results <- glmer(lrti_events_total ~ treatment_assigned + child_age + 
+                         (1 | home:cohort:area), 
+                         data = analysis_data, family = poisson(link = "log"), 
+                         offset = log_person_time_at_risk)
 summary(poisson_results)
+tidy(poisson_results, conf.int = TRUE)
 
-negbin_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+poisson_simres <- simulateResiduals(poisson_results)
+testDispersion(poisson_simres, alternative = "less")
+plot(poisson_simres, rank = FALSE, quantreg = TRUE)
+
+negbin_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + 
+                           (1 | home:cohort:area) + 
+                           offset(log_person_time_at_risk), 
                            data = analysis_data)
 summary(negbin_results)
+tidy(negbin_results, conf.int = TRUE)
 
-logreg_results <- glmer(lrti_events_di ~ treatment + child_age + (1 | home:cohort:area), 
-                        data = analysis_data, family = binomial, offset = log_person_time_at_risk)
+negbin_simres <- simulateResiduals(negbin_results)
+testDispersion(negbin_simres, alternative = "less")
+plot(negbin_simres, rank = FALSE, quantreg = TRUE)
+
+logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                        child_age + person_time_at_risk +
+                        (1 | home:cohort:area),
+                        data = analysis_data, family = binomial, nAGQ = 20)
 summary(logreg_results)
+tidy(logreg_results, conf.int = TRUE)
+
+logreg_simres <- simulateResiduals(logreg_results)
+testDispersion(logreg_simres, alternative = "less")
+plot(logreg_simres, rank = FALSE, quantreg = TRUE)
 ```

### 2020-11-18 main_analysis_secondary_models.Rmd
major model updates new RMD for final analysis

@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) 
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         lrti_events_di_total = as.factor(lrti_events_di_total),
+         person_time_at_risk_iqr = person_time_at_risk/20) 
   #filter(gender == "Female")
@@
   mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
-  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk))
+  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk),
+         lrti_events_total = as.factor(lrti_events_total),
+         lrti_events_di_total = as.factor(lrti_events_di_total))
 
@@
 itt_res <- simulateResiduals(itt_results)
+testDispersion(itt_res, alternative = "less")
 plot(itt_res, rank = FALSE, quantreg = TRUE)
@@
                           education_3level + residents_smoke + 
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk), 
+                          (1 | cohort:area) + offset(log_person_time_at_risk), 
                           data = analysis_data)
@@
 # Run primary model and print results
-logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + child_age + 
-                       (1 | home:cohort:area), data = analysis_data, 
-                       family = binomial)
+logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + child_age + person_time_at_risk +
+                        (1 | home:cohort:area),
+                        data = analysis_data, family = binomial, nAGQ = 20)
+#write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
+
+summary(logreg_results)
+
+# Run primary model and print results
+logreg_results <- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr + child_age + 
+                        education_3level + residents_smoke + 
+                        (1 | cohort:area),
+                        data = analysis_data, family = binomial, nAGQ = 20)
 #write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
@@
 
-## At-risk period across treatment arms
+# At-risk period across treatment arms
 ```{r}

### 2020-11-18 main_analysis_final.Rmd
major analysis updates

@@
 author: "Ethan Walker"
-date: "Updated 17 Nov 2020"
+date: "Updated 18 Nov 2020"
 output: pdf_document
@@
 
-# Options for primary model framework
+# Intent-to-treat framework, LRTI outcome
 
-## Poisson model
-```{r, echo=TRUE, eval=FALSE}
-poisson_results <- glmer(lrti_events_total ~ treatment_assigned + child_age + 
+```{r, echo=TRUE}
+# Run primary model and print results
+itt_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk +
                      (1 | home:cohort:area),
-                         data = analysis_data, family = poisson(link = "log"), 
-                         offset = log_person_time_at_risk)
-#write_rds(poisson_results, paste0(file_path, "Output/poisson_results.rds"))
+                     data = analysis_data, family = binomial, nAGQ = 20)
+#write_rds(itt_results, paste0(file_path, "Output/itt_results.rds"))
+```
+
+## Model summary and diagnostic plots
+```{r}
+summary(itt_results)
+itt_res <- simulateResiduals(itt_results)
+testDispersion(itt_res, alternative = "two.sided")
+plot(itt_res, rank = FALSE, quantreg = TRUE)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Run and save model results
+
+itt_winter1_results <- glmer(lrti_events_di_winter ~ treatment_assigned + 
+                     child_age + person_time_at_risk_winter +
+                     (1 | home:cohort:area),
+                     data = analysis_data_winter1, family = binomial, nAGQ = 20)
+write_rds(itt_winter1_results, paste0(file_path, "Output/itt_winter1_results.rds"))
+
+itt_winter2_results <- glmer(lrti_events_di_winter ~ treatment_assigned + 
+                     child_age + person_time_at_risk_winter +
+                     (1 | home:cohort:area),
+                     data = analysis_data_winter2, family = binomial, nAGQ = 20)
+write_rds(itt_winter2_results, paste0(file_path, "Output/itt_winter2_results.rds"))
+
+itt_income_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + income_3level + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_income_results, paste0(file_path, "Output/itt_income_results.rds"))
+
+itt_filter_type_results <- glmer(lrti_events_di_total ~ treatment_filter_type +  
+                     child_age + person_time_at_risk + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_filter_type_results, paste0(file_path, "Output/itt_filter_type_results.rds"))
+
+itt_filter_comp_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + kw_perc_exp_cat2 + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_filter_comp_results, paste0(file_path, "Output/itt_filter_comp_results.rds"))
+
+itt_smoke_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + residents_smoke + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_smoke_results, paste0(file_path, "Output/itt_smoke_results.rds"))
+
+itt_burn_level_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + burn_level_3level + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_burn_level_results, paste0(file_path, "Output/itt_burn_level_results.rds"))
+
+itt_smoking_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + home_act_smoking_sum + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_smoking_results, paste0(file_path, "Output/itt_smoking_results.rds"))
+
+itt_windows_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + home_act_windows_sum + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_windows_results, paste0(file_path, "Output/itt_windows_results.rds"))
+
+itt_door_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + home_act_door_sum + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_door_results, paste0(file_path, "Output/itt_door_results.rds"))
+
+itt_sweep_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + home_act_sweep_sum + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_sweep_results, paste0(file_path, "Output/itt_sweep_results.rds"))
+
+itt_ambpm_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + amb_pm_lrti + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_ambpm_results, paste0(file_path, "Output/itt_ambpm_results.rds"))
+
+itt_ambtemp_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + amb_temp_lrti + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_ambtemp_results, paste0(file_path, "Output/itt_ambtemp_results.rds"))
+
+itt_home_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + 
+                     (1 | cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+write_rds(itt_home_results, paste0(file_path, "Output/itt_home_results.rds"))
+
+itt_outlier_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + 
+                     (1 | home:cohort:area),
+                     data = analysis_data_outliers, family = binomial, nAGQ = 20)
+write_rds(itt_outlier_results, paste0(file_path, "Output/itt_outlier_results.rds"))
 ```
@@
 ```{r}
-poisson_results <- read_rds(paste0(file_path, "Output/poisson_results.rds"))
-summary(poisson_results)
-#tidy(poisson_results, conf.int = TRUE)
+# Load results
+itt_results <- read_rds(paste0(file_path, "Output/itt_results.rds"))
+itt_winter1_results <- read_rds(paste0(file_path, "Output/itt_winter1_results.rds"))
+#itt_winter2_results <- read_rds(paste0(file_path, "Output/itt_winter2_results.rds"))
+itt_income_results <- read_rds(paste0(file_path, "Output/itt_income_results.rds"))
+#itt_filter_type_results <- read_rds(paste0(file_path, "Output/itt_filter_type_results.rds"))
+itt_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_filter_comp_results.rds"))
+itt_smoke_results <- read_rds(paste0(file_path, "Output/itt_smoke_results.rds"))
+itt_burn_level_results <- read_rds(paste0(file_path, "Output/itt_burn_level_results.rds"))
+itt_smoking_results <- read_rds(paste0(file_path, "Output/itt_smoking_results.rds"))
+itt_windows_results <- read_rds(paste0(file_path, "Output/itt_windows_results.rds"))
+itt_door_results <- read_rds(paste0(file_path, "Output/itt_door_results.rds"))
+itt_sweep_results <- read_rds(paste0(file_path, "Output/itt_sweep_results.rds"))
+itt_ambpm_results <- read_rds(paste0(file_path, "Output/itt_ambpm_results.rds"))
+itt_ambtemp_results <- read_rds(paste0(file_path, "Output/itt_ambtemp_results.rds"))
+itt_home_results <- read_rds(paste0(file_path, "Output/itt_home_results.rds"))
+itt_outlier_results <- read_rds(paste0(file_path, "Output/itt_outlier_results.rds"))
 
-poisson_simres <- simulateResiduals(poisson_results)
-testDispersion(poisson_simres, alternative = "less")
-plot(poisson_simres, rank = FALSE, quantreg = TRUE)
-```
-
-## Negative binomial model
-```{r, echo=TRUE, eval=FALSE}
-negbin_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + 
-                           (1 | home:cohort:area) + 
-                           offset(log_person_time_at_risk), 
-                           data = analysis_data)
-#write_rds(negbin_results, paste0(file_path, "Output/negbin_results.rds"))
-```
-
-```{r}
-negbin_results <- read_rds(paste0(file_path, "Output/negbin_results.rds"))
-summary(negbin_results)
-#tidy(negbin_results, conf.int = TRUE)
-
-negbin_simres <- simulateResiduals(negbin_results)
-testDispersion(negbin_simres, alternative = "less")
-plot(negbin_simres, rank = FALSE, quantreg = TRUE)
-```
-
-## Logistic regression model
-```{r, echo=TRUE, eval=FALSE}
-logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
-                        child_age + person_time_at_risk +
-                        (1 | home:cohort:area),
-                        data = analysis_data, family = binomial, nAGQ = 20)
-#write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
-```
-
-```{r}
-logreg_results <- read_rds(paste0(file_path, "Output/logreg_results.rds"))
-summary(logreg_results)
-#tidy(logreg_results, conf.int = TRUE)
-
-logreg_simres <- simulateResiduals(logreg_results)
-testDispersion(logreg_simres, alternative = "two.sided")
-plot(logreg_simres, rank = FALSE, quantreg = TRUE)
-```
-
-```{r}
 # Function to clean results
@@
 # Run results through function, save, combine for plotting
-poisson_tidy <- results_function(poisson_results, "Poisson")
-negbin_tidy <- results_function(negbin_results, "Neg Bin")
-logreg_tidy <- results_function(logreg_results, "Log Reg")
+itt_tidy <- results_function(itt_results, "ITT Primary")
+itt_winter1_tidy <- results_function(itt_winter1_results, "ITT Winter 1")
+#itt_winter2_tidy <- results_function(itt_winter2_results, "ITT Winter 2")
+itt_income_tidy <- results_function(itt_income_results, "ITT + income")
+#itt_filter_type_tidy <- results_function(itt_filter_type_results, "ITT + filt type")
+itt_filter_comp_tidy <- results_function(itt_filter_comp_results, "ITT + filt comp")
+itt_smoke_tidy <- results_function(itt_smoke_results, "ITT + smoke hx")
+itt_burn_level_tidy <- results_function(itt_burn_level_results, "ITT + burn level")
+itt_smoking_tidy <- results_function(itt_smoking_results, "ITT + smoke act")
+itt_windows_tidy <- results_function(itt_windows_results, "ITT + window act")
+itt_door_tidy <- results_function(itt_door_results, "ITT + door act")
+itt_sweep_tidy <- results_function(itt_sweep_results, "ITT + sweep act")
+itt_ambpm_tidy <- results_function(itt_ambpm_results, "ITT + amb PM")
+itt_ambtemp_tidy <- results_function(itt_ambtemp_results, "ITT + amb temp")
+itt_home_tidy <- results_function(itt_home_results, "ITT - home")
+itt_outlier_tidy <- results_function(itt_outlier_results, "ITT outlier")
 
 # Combine cleaned results
-itt_results_combined <- rbind(poisson_tidy, negbin_tidy, logreg_tidy)
+itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, 
+                              itt_income_tidy, itt_filter_comp_tidy,
+                              itt_smoke_tidy, itt_burn_level_tidy, itt_smoking_tidy,
+                              itt_windows_tidy, itt_door_tidy, itt_sweep_tidy,
+                              itt_ambpm_tidy, itt_ambtemp_tidy,
+                              itt_home_tidy, itt_outlier_tidy)
 
@@
   mutate(model = factor(model,
-                        levels = c("Poisson", "Neg Bin", "Log Reg"))) %>% 
+                        levels = c("ITT Primary", "ITT Winter 1", 
+                                   "ITT + income", "ITT + filt comp",
+                                   "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
+                                   "ITT + window act", "ITT + door act", "ITT + sweep act",
+                                   "ITT + amb PM", "ITT + amb temp",
+                                   "ITT - home", "ITT outlier"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
   ggtitle(label = "Results for ITT framework, LRTI outcome") +
-  labs(y = "Estimate compared to placebo") +
+  labs(y = "LRTI events per child-week \n at risk compared to placebo") +
   labs(x = "", group = "", shape = "") +
@@
 
-# Interpretation:
-As Jon and I discussed 11/16/20, the Poisson and Negative Binomial models appear to be
-underdispersed (variance < mean), not overdispersed (variance > mean). These models 
-don't appear to be the best fit for our data. Additionally, results from the models 
-were essentially the same.  
-The Logistic Regression model appears to be a better fit for the data (no concerns 
-in the diagnostic plots). The good news is that the results were also very similar to 
-the Poisson/NegBin models, so we don't have to explain any big differences in results 
-if model selection concerns come up with reviewers. The bad news is that regardless 
-of model, there is no difference in LRTI between treatment arms.   
-Given these results, I'm leaning towards using the logistic regression model for the 
-primary ITT framework. There are models that can be used for underdispersed count 
-data (generalized Poisson regression?), but I can't find a good way to run those 
-as mixed models like we need to do. I think the logistic regression model will be 
-easier to explain/interpret, and we aren't really losing much information since there 
-were so few children with 2+ cases of LRTI.
+ITT Winter 1 = Primary model with Winter 1 data only  
+ITT + income = Primary model plus categorical income variable  
+ITT + filt comp = Primary model plus variable for filter compliance  
+ITT + smoke hx = Primary model plus self-reported smoking variable  
+ITT + burn level = Primary model plus self-reported burn-level compared to typical  
+ITT + smoke act = Primary model plus # days during sampling with smoking  
+ITT + window act = Primary model plus # days during sampling with windows open  
+ITT + door act = Primary model plus # days during sampling with doors open  
+ITT + sweep act = Primary model plus # days during sampling with sweeping/cleaning  
+ITT + amb PM = Primary model plus amb PM during LRTI at-risk period  
+ITT + amb temp = Primary model plus amb temp during LRTI at-risk period  
+ITT - home = Primary model without home as random term  
+ITT outlier = Primary model with data that fixes incidence outlier
+
+\pagebreak
+
+# Intent-to-treat framework, PM2.5 outcome
+
+```{r, echo=TRUE}
+# Run primary model and print results
+itt_pm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
+                       (1 | cohort:area), data = analysis_data_winter1)
+#write_rds(itt_pm_results, paste0(file_path, "Output/itt_pm_results.rds"))
+```
+
+## Model summary and diagnostic plots
+```{r}
+summary(itt_pm_results)
+plot(itt_pm_results, main = "PM ITT Primary model", 
+     xlab = "fitted values", ylab = "residuals")
+qqnorm(residuals(itt_pm_results), main = "PM ITT Primary model")
+qqline(residuals(itt_pm_results))
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Effect modification/interaction analysis
+# Try child_age, gender, home_floors_2level, home_sqft_2level
+
+# Run primary model and print results
+itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*home_sqft_2level + 
+                              child_age + 
+                       (1 | cohort:area), data = analysis_data)
+#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+
+summary(itt_pm_results_int)
+
+int_emtrends <- emtrends(itt_pm_results_int, home_sqft_2level ~ treatment_assigned, var = "log(pm_mean_sampling_period)")
+summary(int_emtrends, conf.int = TRUE)
+# This line will plot the interaction term
+#emmip(itt_pm_results_int, child_age ~ treatment_assigned, cov.reduce = range)
+table(model.frame(itt_pm_results_int)$treatment_assigned)
+nobs(itt_pm_results_int)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Run and save model results
+
+itt_pm_income_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                               child_age + (1 | cohort:area) + 
+                               income_3level, data = analysis_data_winter1)
+write_rds(itt_pm_income_results, paste0(file_path, "Output/itt_pm_income_results.rds"))
+
+itt_pm_filter_type_results <- glmer(log(pm_mean_sampling_period) ~ treatment_filter_type + 
+                                  child_age + (1 | cohort:area), data = analysis_data_winter1)
+write_rds(itt_pm_filter_type_results, paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
+
+itt_pm_filter_comp_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                  child_age + (1 | cohort:area) + 
+                                  kw_perc_exp_cat2, data = analysis_data_winter1)
+write_rds(itt_pm_filter_comp_results, paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
+
+itt_pm_smoke_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                              child_age + (1 | cohort:area) + 
+                              residents_smoke, data = analysis_data_winter1)
+write_rds(itt_pm_smoke_results, paste0(file_path, "Output/itt_pm_smoke_results.rds"))
+
+itt_pm_burn_level_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   burn_level_3level, data = analysis_data_winter1)
+write_rds(itt_pm_burn_level_results, paste0(file_path, "Output/itt_pm_burn_level_results.rds"))
+
+itt_pm_smoking_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   home_act_smoking_sum, data = analysis_data_winter1)
+write_rds(itt_pm_smoking_results, paste0(file_path, "Output/itt_pm_smoking_results.rds"))
+
+itt_pm_windows_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   home_act_windows_sum, data = analysis_data_winter1)
+write_rds(itt_pm_windows_results, paste0(file_path, "Output/itt_pm_windows_results.rds"))
+
+itt_pm_door_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   home_act_door_sum, data = analysis_data_winter1)
+write_rds(itt_pm_door_results, paste0(file_path, "Output/itt_pm_door_results.rds"))
+
+itt_pm_sweep_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   home_act_sweep_sum, data = analysis_data_winter1)
+write_rds(itt_pm_sweep_results, paste0(file_path, "Output/itt_pm_sweep_results.rds"))
+
+itt_pm_ambpm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   amb_pm_lrti, data = analysis_data_winter1)
+write_rds(itt_pm_ambpm_results, paste0(file_path, "Output/itt_pm_ambpm_results.rds"))
+
+itt_pm_ambtemp_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                                   child_age + (1 | cohort:area) + 
+                                   amb_temp_lrti, data = analysis_data_winter1)
+write_rds(itt_pm_ambtemp_results, paste0(file_path, "Output/itt_pm_ambtemp_results.rds"))
+```
+
+```{r}
+# Load results
+itt_pm_results <- read_rds(paste0(file_path, "Output/itt_pm_results.rds"))
+itt_pm_income_results <- read_rds(paste0(file_path, "Output/itt_pm_income_results.rds"))
+itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
+itt_pm_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
+itt_pm_smoke_results <- read_rds(paste0(file_path, "Output/itt_pm_smoke_results.rds"))
+itt_pm_burn_level_results <- read_rds(paste0(file_path, "Output/itt_pm_burn_level_results.rds"))
+itt_pm_smoking_results <- read_rds(paste0(file_path, "Output/itt_pm_smoking_results.rds"))
+itt_pm_windows_results <- read_rds(paste0(file_path, "Output/itt_pm_windows_results.rds"))
+itt_pm_door_results <- read_rds(paste0(file_path, "Output/itt_pm_door_results.rds"))
+itt_pm_sweep_results <- read_rds(paste0(file_path, "Output/itt_pm_sweep_results.rds"))
+itt_pm_ambpm_results <- read_rds(paste0(file_path, "Output/itt_pm_ambpm_results.rds"))
+itt_pm_ambtemp_results <- read_rds(paste0(file_path, "Output/itt_pm_ambtemp_results.rds"))
+
+# Run function to clean results
+results_function <- function(model_results, model_name) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_assigned", "", term),
+         model = model_name) %>% 
+  dplyr::select(model, term, estimate, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+itt_pm_tidy <- results_function(itt_pm_results, "ITT Primary")
+itt_pm_income_tidy <- results_function(itt_pm_income_results, "ITT + income")
+itt_pm_filter_type_tidy <- results_function(itt_pm_filter_type_results, "ITT filter type")
+itt_pm_filter_comp_tidy <- results_function(itt_pm_filter_comp_results, "ITT + filt comp")
+itt_pm_smoke_tidy <- results_function(itt_pm_smoke_results, "ITT + smoke hx")
+itt_pm_burn_level_tidy <- results_function(itt_pm_burn_level_results, "ITT + burn level")
+itt_pm_smoking_tidy <- results_function(itt_pm_smoking_results, "ITT + smoke act")
+itt_pm_windows_tidy <- results_function(itt_pm_windows_results, "ITT + window act")
+itt_pm_door_tidy <- results_function(itt_pm_door_results, "ITT + door act")
+itt_pm_sweep_tidy <- results_function(itt_pm_sweep_results, "ITT + sweep act")
+itt_pm_ambpm_tidy <- results_function(itt_pm_ambpm_results, "ITT + amb PM")
+itt_pm_ambtemp_tidy <- results_function(itt_pm_ambtemp_results, "ITT + amb temp")
+
+# Combine cleaned results
+itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_income_tidy, itt_pm_filter_comp_tidy, 
+                              itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
+                              itt_pm_windows_tidy, itt_pm_door_tidy, itt_pm_sweep_tidy,
+                              itt_pm_ambpm_tidy, itt_pm_ambtemp_tidy)
+
+# Plot results
+itt_pm_plot_estimates <- itt_pm_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("ITT Primary", "ITT + income", "ITT + filt comp", 
+                                   "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
+                                   "ITT + window act", "ITT + door act", "ITT + sweep act",
+                                   "ITT + amb PM", "ITT + amb temp"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Results for ITT framework, PM outcome") +
+  labs(y = "Difference in log(PM2.5) \n compared to placebo") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_pm_plot_estimates
+
+# Save plot as jpg
+#ggsave("itt_pm_plot_estimates.jpg", width = 12, height = 6)
+```
+
+ITT + income = Primary model plus categorical income variable  
+ITT + filt comp = Primary model plus variable for filter compliance  
+ITT + smoke hx = Primary model plus self-reported smoking variable  
+ITT + burn level = Primary model plus self-reported burn-level compared to typical  
+ITT + smoke act = Primary model plus # days during sampling with smoking  
+ITT + window act = Primary model plus # days during sampling with windows open  
+ITT + door act = Primary model plus # days during sampling with doors open  
+ITT + sweep act = Primary model plus # days during sampling with sweeping/cleaning  
+ITT + amb PM = Primary model plus amb PM during LRTI at-risk period  
+ITT + amb temp = Primary model plus amb temp during LRTI at-risk period
+
+\pagebreak
+
+# Exposure-response framework, LRTI outcome
+
+```{r, echo=TRUE}
+# Run primary model and print results
+er_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+#write_rds(er_results, paste0(file_path, "Output/er_results.rds"))
+```
+
+## Model summary and diagnostic plots
+```{r}
+summary(er_results)
+er_res <- simulateResiduals(er_results)
+plot(er_res, rank = TRUE, quantreg = TRUE)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Run and save model results
+
+er_income_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_income_results, paste0(file_path, "Output/er_income_results.rds"))
+
+er_filter_comp_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + kw_perc_exp_cat2 +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_filter_comp_results, paste0(file_path, "Output/er_filter_comp_results.rds"))
+
+er_burn_level_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + burn_level_3level +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_burn_level_results, paste0(file_path, "Output/er_burn_level_results.rds"))
+
+er_smoking_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + home_act_smoking_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_smoking_results, paste0(file_path, "Output/er_smoking_results.rds"))
+
+er_windows_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_windows_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_windows_results, paste0(file_path, "Output/er_windows_results.rds"))
+
+er_door_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_door_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_door_results, paste0(file_path, "Output/er_door_results.rds"))
+
+er_sweep_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + home_act_sweep_sum +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_sweep_results, paste0(file_path, "Output/er_sweep_results.rds"))
+
+er_ambpm_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + amb_pm_lrti +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_ambpm_results, paste0(file_path, "Output/er_ambpm_results.rds"))
+
+er_ambtemp_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + amb_temp_lrti +
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_ambtemp_results, paste0(file_path, "Output/er_ambtemp_results.rds"))
+
+er_home_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+write_rds(er_home_results, paste0(file_path, "Output/er_home_results.rds"))
+```
+
+```{r}
+# Load results
+er_results <- read_rds(paste0(file_path, "Output/er_results.rds"))
+er_income_results <- read_rds(paste0(file_path, "Output/er_income_results.rds"))
+er_filter_comp_results <- read_rds(paste0(file_path, "Output/er_filter_comp_results.rds"))
+er_burn_level_results <- read_rds(paste0(file_path, "Output/er_burn_level_results.rds"))
+er_smoking_results <- read_rds(paste0(file_path, "Output/er_smoking_results.rds"))
+er_windows_results <- read_rds(paste0(file_path, "Output/er_windows_results.rds"))
+er_door_results <- read_rds(paste0(file_path, "Output/er_door_results.rds"))
+er_sweep_results <- read_rds(paste0(file_path, "Output/er_sweep_results.rds"))
+er_ambpm_results <- read_rds(paste0(file_path, "Output/er_ambpm_results.rds"))
+er_ambtemp_results <- read_rds(paste0(file_path, "Output/er_ambtemp_results.rds"))
+er_home_results <- read_rds(paste0(file_path, "Output/er_home_results.rds"))
+
+# Run function to clean results
+results_function <- function(model_results, model_name) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("pm_mean", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = "PM2.5",
+         model = model_name) %>% 
+  dplyr::select(model, term, estimate, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+er_tidy <- results_function(er_results, "ER Primary")
+er_income_tidy <- results_function(er_income_results, "ER + income")
+er_filter_comp_tidy <- results_function(er_filter_comp_results, "ER + filt comp")
+er_burn_level_tidy <- results_function(er_burn_level_results, "ER + burn level")
+er_smoking_tidy <- results_function(er_smoking_results, "ER + smoke act")
+er_windows_tidy <- results_function(er_windows_results, "ER + window act")
+er_door_tidy <- results_function(er_door_results, "ER + door act")
+er_sweep_tidy <- results_function(er_sweep_results, "ER + sweep act")
+er_ambpm_tidy <- results_function(er_ambpm_results, "ER + amb PM")
+er_ambtemp_tidy <- results_function(er_ambtemp_results, "ER + amb temp")
+er_home_tidy <- results_function(er_home_results, "ER - home")
+
+# Combine cleaned results
+er_results_combined <- rbind(er_tidy, er_income_tidy, er_filter_comp_tidy, er_burn_level_tidy, 
+                              er_smoking_tidy, er_windows_tidy, er_door_tidy,
+                              er_sweep_tidy, er_ambpm_tidy, er_ambtemp_tidy, er_home_tidy)
+
+# Plot results
+er_plot_estimates <- er_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("ER Primary", "ER + income", "ER + filt comp", "ER + burn level",
+                                   "ER + smoke act", "ER + window act", "ER + door act",
+                                   "ER + sweep act", "ER + amb PM", "ER + amb temp", "ER - home"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Results for ER framework, LRTI outcome") +
+  labs(y = "LRTI cases per IQR \n increase in PM2.5") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "none",
+          legend.text = element_text(size = 14, colour = "black")) 
+er_plot_estimates
+
+# Save plot as jpg
+#ggsave("er_plot_estimates.jpg", width = 12, height = 6)
+```
+
+ER + income = Primary model using income instead of education  
+ER + filt comp = Primary model plus variable for filter compliance  
+ER + burn level = Primary model plus self-reported burn-level compared to typical  
+ER + smoke act = Primary model plus # days during sampling with smoking  
+ER + window act = Primary model plus # days during sampling with windows open  
+ER + door act = Primary model plus # days during sampling with doors open  
+ER + sweep act = Primary model plus # days during sampling with sweeping/cleaning  
+ER + amb PM = Primary model plus amb PM during LRTI at-risk period  
+ER + amb temp = Primary model plus amb temp during LRTI at-risk period  
+ER - home = Primary model without home as random term

### 2020-11-18 main_analysis_initial.Rmd
major analysis updates

@@
   mutate(kw_perc_exp_cat = cut(kw_perc_expected, 
-                               breaks = c(0, 42, 83, 129, 50000),
+                               breaks = c(0, 30, 56, 106, 1000),
                                labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
@@
   mutate(kw_perc_exp_cat = cut(kw_perc_expected, 
-                               breaks = c(0, 37, 74, 118, 50000),
+                               breaks = c(0, 29, 53, 96, 1000),
                                labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
@@
   arrange(area, home, child_id_num, winter_id) %>%
+  left_join(kids_amb_pm_data, by = c("area", "home")) %>% 
+  mutate(pm_filter = if_else((sample_date >= person_time_start) & (sample_date <= person_time_end), 1, 0)) %>% 
+  filter(pm_filter == 1) %>% 
+  group_by(area, child_id_num, winter_id) %>% 
+  mutate(amb_pm_lrti = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
   ungroup() %>% 
+  dplyr::select(area, home, cohort, child_id_num, lrti_events_winter, person_time_at_risk_winter, lrti_incidence_winter,
+         amb_pm_lrti, person_time_start, person_time_end, home_winter_id, winter_id) %>% 
+  left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
+  mutate(temp_filter = if_else((sample_date >= person_time_start) & (sample_date <= person_time_end), 1, 0)) %>% 
+  filter(temp_filter == 1 | child_id_num == 94) %>% 
+  group_by(area, child_id_num, winter_id) %>% 
+  mutate(amb_temp_lrti = mean(mean_temp, na.rm = TRUE),
+         amb_temp_lrti = if_else(child_id_num == 94, NaN, amb_temp_lrti),
+         amb_dewp_lrti = mean(mean_dewp, na.rm = TRUE),
+         amb_dewp_lrti = if_else(child_id_num == 94, NaN, amb_dewp_lrti),
+         amb_pressure_lrti = mean(mean_pressure, na.rm = TRUE),
+         amb_pressure_lrti = if_else(child_id_num == 94, NaN, amb_pressure_lrti),
+         amb_wind_lrti = mean(mean_wind_knots, na.rm = TRUE),
+         amb_wind_lrti = if_else(child_id_num == 94, NaN, amb_wind_lrti)) %>% 
+  ungroup() %>% 
   dplyr::select(area, home, cohort, child_id_num, winter_id,
-                lrti_events_winter, person_time_at_risk_winter, lrti_incidence_winter)
+                lrti_events_winter, person_time_at_risk_winter, lrti_incidence_winter,
+         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti) %>% 
+  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
+  ungroup()
 
@@
   left_join(kids_amb_pm_data, by = c("area", "home")) %>% 
-  mutate(pm_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
+  mutate(pm_filter = if_else((sample_date >= person_time_start) & (sample_date <= person_time_end), 1, 0)) %>% 
   filter(pm_filter == 1) %>% 
@@
   left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
-  mutate(temp_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
+  mutate(temp_filter = if_else((sample_date >= person_time_start) & (sample_date <= person_time_end), 1, 0)) %>% 
   filter(temp_filter == 1 | child_id_num == 94) %>% 
@@
   mutate(lrti_events_di_total = if_else(lrti_events_total == 0, 0, 1),
-         filter_type2 = as.character(filter_type),
-         filter_type2 = if_else(treatment_assigned == "Education", "Ed Home", filter_type2),
-         filter_type2 = factor(filter_type2,
-                              levels = c("Honeywell", "Large Filtrete", "Small Filtrete",
-                                         "Winix", "Ed Home")),
+         treatment_filter_type = as.character(treatment_assigned),
+         filter_type = as.character(filter_type),
+         treatment_filter_type = if_else(treatment_filter_type == "Filter", 
+                                         filter_type, treatment_filter_type),
+         treatment_filter_type = factor(treatment_filter_type,
+                                        levels = c("Placebo", "Education", 
+                                                   "Honeywell", "Large Filtrete", 
+                                                   "Small Filtrete", "Winix")),
          kw_perc_exp_cat2 = as.character(kw_perc_exp_cat),
-         kw_perc_exp_cat2 = if_else(treatment_assigned == "Education", "Ed Home", kw_perc_exp_cat2),
+         kw_perc_exp_cat2 = if_else(treatment_assigned == "Filter", kw_perc_exp_cat2, treatment_assigned),
          kw_perc_exp_cat2 = factor(kw_perc_exp_cat2,
-                                  levels = c("Q1", "Q2", "Q3", "Q4", "Ed Home"))) %>% 
+                                  levels = c("Placebo", "Education", "Q1", "Q2", "Q3", "Q4"))) %>% 
   mutate(treatment = factor(treatment,
@@
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, child_age, gender,
+  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, child_age, 
+         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti, gender,
          lrti_events_winter, person_time_at_risk_winter, lrti_incidence_winter, newtx, treatment_assigned,
@@
   mutate(lrti_events_di_winter = if_else(lrti_events_winter == 0, 0, 1),
-         filter_type2 = as.character(filter_type),
-         filter_type2 = if_else(treatment_assigned == "Education", "Ed Home", filter_type2),
-         filter_type2 = factor(filter_type2,
-                              levels = c("Honeywell", "Large Filtrete", "Small Filtrete",
-                                         "Winix", "Ed Home")),
+         treatment_filter_type = as.character(treatment_assigned),
+         filter_type = as.character(filter_type),
+         treatment_filter_type = if_else(treatment_filter_type == "Filter", 
+                                         filter_type, treatment_filter_type),
+         treatment_filter_type = factor(treatment_filter_type,
+                                        levels = c("Placebo", "Education", 
+                                                   "Honeywell", "Large Filtrete", 
+                                                   "Small Filtrete", "Winix")),
          kw_perc_exp_cat2 = as.character(kw_perc_exp_cat),
-         kw_perc_exp_cat2 = if_else(treatment_assigned == "Education", "Ed Home", kw_perc_exp_cat2),
+         kw_perc_exp_cat2 = if_else(treatment_assigned == "Filter", kw_perc_exp_cat2, treatment_assigned),
          kw_perc_exp_cat2 = factor(kw_perc_exp_cat2,
-                                  levels = c("Q1", "Q2", "Q3", "Q4", "Ed Home"))) %>% 
+                                  levels = c("Placebo", "Education", "Q1", "Q2", "Q3", "Q4"))) %>% 
   mutate(treatment = factor(treatment,

### 2020-11-18 main_analysis_primary_models.Rmd
major analysis updates

@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) 
-  rownames_to_column() %>% 
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         lrti_events_di_total = as.factor(lrti_events_di_total),
+         person_time_at_risk_iqr = person_time_at_risk/20) 
+  #filter(gender == "Female")
+  #filter(home_sqft_2level == "1500+")
+
+analysis_data_outliers <- analysis_data %>% 
+  #rownames_to_column() %>% 
   #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
   mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
-  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk)) 
-  #filter(lrti_incidence < 0.7) %>% 
-  filter(filter_outliers == 0)
+  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk),
+         lrti_events_total = as.factor(lrti_events_total),
+         lrti_events_di_total = as.factor(lrti_events_di_total))
 
-health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+analysis_data_winter1 <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+  filter(person_time_at_risk_winter > 0) %>% 
+  filter(winter_id == 1)
+
+analysis_data_winter2 <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+  filter(person_time_at_risk_winter > 0) %>% 
+  filter(winter_id == 2)
 ```
@@
 ```
+
+# Final options for primary model framework
+
+## Poisson model
+```{r, echo=TRUE, eval=FALSE}
+poisson_results <- glmer(lrti_events_total ~ treatment_assigned + child_age + 
+                         (1 | home:cohort:area), 
+                         data = analysis_data, family = poisson(link = "log"), 
+                         offset = log_person_time_at_risk)
+#write_rds(poisson_results, paste0(file_path, "Output/poisson_results.rds"))
+```
+
+```{r}
+poisson_results <- read_rds(paste0(file_path, "Output/poisson_results.rds"))
+summary(poisson_results)
+#tidy(poisson_results, conf.int = TRUE)
+
+poisson_simres <- simulateResiduals(poisson_results)
+testDispersion(poisson_simres, alternative = "less")
+plot(poisson_simres, rank = FALSE, quantreg = TRUE)
+```
+
+## Negative binomial model
+```{r, echo=TRUE, eval=FALSE}
+negbin_results <- glmer.nb(lrti_events_total ~ treatment_assigned + child_age + 
+                           (1 | home:cohort:area) + 
+                           offset(log_person_time_at_risk), 
+                           data = analysis_data)
+#write_rds(negbin_results, paste0(file_path, "Output/negbin_results.rds"))
+```
+
+```{r}
+negbin_results <- read_rds(paste0(file_path, "Output/negbin_results.rds"))
+summary(negbin_results)
+#tidy(negbin_results, conf.int = TRUE)
+
+negbin_simres <- simulateResiduals(negbin_results)
+testDispersion(negbin_simres, alternative = "less")
+plot(negbin_simres, rank = FALSE, quantreg = TRUE)
+```
+
+## Logistic regression model
+```{r, echo=TRUE, eval=FALSE}
+logreg_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                        child_age + person_time_at_risk +
+                        (1 | home:cohort:area),
+                        data = analysis_data, family = binomial, nAGQ = 20)
+#write_rds(logreg_results, paste0(file_path, "Output/logreg_results.rds"))
+```
+
+```{r}
+logreg_results <- read_rds(paste0(file_path, "Output/logreg_results.rds"))
+summary(logreg_results)
+#tidy(logreg_results, conf.int = TRUE)
+
+logreg_simres <- simulateResiduals(logreg_results)
+testDispersion(logreg_simres, alternative = "two.sided")
+plot(logreg_simres, rank = FALSE, quantreg = TRUE)
+```
+
+```{r}
+# Function to clean results
+results_function <- function(model_results, model_name) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_assigned", "", term),
+         model = model_name) %>% 
+  dplyr::select(model, term, estimate, p.value, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+poisson_tidy <- results_function(poisson_results, "Poisson")
+negbin_tidy <- results_function(negbin_results, "Neg Bin")
+logreg_tidy <- results_function(logreg_results, "Log Reg")
+
+# Combine cleaned results
+itt_results_combined <- rbind(poisson_tidy, negbin_tidy, logreg_tidy)
+
+# Plot results
+itt_plot_estimates <- itt_results_combined %>%
+  mutate(model = factor(model,
+                        levels = c("Poisson", "Neg Bin", "Log Reg"))) %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=model, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=model, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Results for ITT framework, LRTI outcome") +
+  labs(y = "Estimate compared to placebo") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black")) 
+itt_plot_estimates
+
+# Save plot as jpg
+#ggsave("itt_plot_estimates.jpg", width = 12, height = 6)
+```
+
+# Interpretation:
+As Jon and I discussed 11/16/20, the Poisson and Negative Binomial models appear to be
+underdispersed (variance < mean), not overdispersed (variance > mean). These models 
+don't appear to be the best fit for our data. Additionally, results from the models 
+were essentially the same.  
+The Logistic Regression model appears to be a better fit for the data (no concerns 
+in the diagnostic plots). The good news is that the results were also very similar to 
+the Poisson/NegBin models, so we don't have to explain any big differences in results 
+if model selection concerns come up with reviewers. The bad news is that regardless 
+of model, there is no difference in LRTI between treatment arms.   
+Given these results, I'm leaning towards using the logistic regression model for the 
+primary ITT framework. There are models that can be used for underdispersed count 
+data (generalized Poisson regression?), but I can't find a good way to run those 
+as mixed models like we need to do. I think the logistic regression model will be 
+easier to explain/interpret, and we aren't really losing much information since there 
+were so few children with 2+ cases of LRTI.

### 2020-11-18 stove_grade_usage_cleaning.Rmd
major analysis updates

@@
   replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
-  mutate(kw_perc_expected = as.numeric(kw_perc_expected))
+  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) %>% 
+  filter(treatment_assigned == "Filter")
 

### 2020-12-04 main_analysis_final.Rmd
analysis updates

@@
 author: "Ethan Walker"
-date: "Updated 18 Nov 2020"
+date: "Updated 30 Nov 2020"
 output: pdf_document
@@
 plot(itt_res, rank = FALSE, quantreg = TRUE)
+```
+
+```{r, eval=FALSE, include=FALSE}
+# Effect modification/interaction analysis
+# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
+
+# Run primary model and print results
+itt_results_int <- glmer(lrti_events_di_total ~ treatment_assigned*residents_smoke + 
+                     child_age + 
+                     person_time_at_risk + (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+#write_rds(itt_results_int, paste0(file_path, "Output/itt_results_int_age.rds"))
+summary(itt_results_int)
+
+int_emmeans <- emmeans(itt_results_int, pairwise ~ treatment_assigned | residents_smoke)
+summary(int_emmeans, conf.int = TRUE)
+confint(int_emmeans)
+
+table(model.frame(itt_results_int)$treatment_assigned)
+nobs(itt_results_int)
 ```
@@
   ggtitle(label = "Results for ITT framework, LRTI outcome") +
-  labs(y = "LRTI events per child-week \n at risk compared to placebo") +
+  labs(y = "Odds of LRTI compared to placebo") +
   labs(x = "", group = "", shape = "") +
@@
 
+ITT Primary = Logistic regression mixed model, as specified above  
 ITT Winter 1 = Primary model with Winter 1 data only  
@@
 ITT outlier = Primary model with data that fixes incidence outlier
+
+## Analysis to assess effect modification
+
+There were small differences in the associations across age groups (<1 vs >=1)  
+and gender groups, but confidence intervals overlap substantially.  
+There were minimal differences in associations across groups for levels/floors  
+in the home, sqft of the home, or residents smoking.
 
@@
 # Effect modification/interaction analysis
-# Try child_age, gender, home_floors_2level, home_sqft_2level
+# Try home_floors_2level, home_sqft_2level, education_3level, child_age
+# home_year_built_2level, stove_age_3level, stove_grade_3level, chimney_clean_3level,
+# wood_collect_2level, burn_level_3level, residents_smoke
 
 # Run primary model and print results
-itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*home_sqft_2level + 
+itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*residents_smoke + 
                               child_age + 
-                       (1 | cohort:area), data = analysis_data)
+                       (1 | cohort:area), data = analysis_data_winter1)
 #write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+#summary(itt_pm_results_int)
 
-summary(itt_pm_results_int)
+int_emmeans <- emmeans(itt_pm_results_int, pairwise ~ treatment_assigned | residents_smoke)
+#summary(int_emmeans, conf.int = TRUE)
+confint(int_emmeans)
 
-int_emtrends <- emtrends(itt_pm_results_int, home_sqft_2level ~ treatment_assigned, var = "log(pm_mean_sampling_period)")
-summary(int_emtrends, conf.int = TRUE)
-# This line will plot the interaction term
-#emmip(itt_pm_results_int, child_age ~ treatment_assigned, cov.reduce = range)
-table(model.frame(itt_pm_results_int)$treatment_assigned)
-nobs(itt_pm_results_int)
+table(model.frame(itt_results_int)$treatment_assigned)
+nobs(itt_results_int)
+
+((exp(-.418)-1)*100)
 ```
@@
          model = model_name) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(conf.low)-1)*100, digits = 2),
+         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
   dplyr::select(model, term, estimate, conf.low, conf.high)
@@
   ggtitle(label = "Results for ITT framework, PM outcome") +
-  labs(y = "Difference in log(PM2.5) \n compared to placebo") +
+  labs(y = "% difference in PM2.5 \n compared to placebo") +
   labs(x = "", group = "", shape = "") +
@@
 
+ITT Primary = Linear mixed model, as specified above  
 ITT + income = Primary model plus categorical income variable  
@@
 ITT + amb temp = Primary model plus amb temp during LRTI at-risk period
+
+## Analysis to assess effect modification
+
+There were small differences in the associations across groups for levels/floors,    
+age of home, when chimney was cleaned, and resident smoking status,  
+but confidence intervals overlap substantially.  
+There were minimal differences in associations across groups for other variables.
 
@@
 er_res <- simulateResiduals(er_results)
+testDispersion(er_res, alternative = "two.sided")
 plot(er_res, rank = TRUE, quantreg = TRUE)
@@
 ```{r, eval=FALSE, include=FALSE}
+# Effect modification/interaction analysis
+# Try education_3level, child_age, chimney_clean_3level, wood_collect_2level, residents_smoke
+
+# Run primary model and print results
+er_results_int <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr*chimney_clean_3level + 
+                          child_age + 
+                          education_3level + 
+                          residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+#write_rds(er_results_int, paste0(file_path, "Output/er_results_int_age.rds"))
+#summary(er_results_int)
+
+int_emtrends <- emtrends(er_results_int, pairwise ~ chimney_clean_3level, var = "pm_mean_sampling_period_iqr")
+summary(int_emtrends, conf.int = TRUE)
+
+table(model.frame(itt_results_int)$treatment_assigned)
+nobs(itt_results_int)
+```
+
+```{r, eval=FALSE, include=FALSE}
 # Run and save model results
+
+er_results_lr <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + person_time_at_risk_winter +
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 20)
+write_rds(er_results_lr, paste0(file_path, "Output/er_results_lr.rds"))
 
@@
 er_results <- read_rds(paste0(file_path, "Output/er_results.rds"))
+er_results_lr <- read_rds(paste0(file_path, "Output/er_results_lr.rds"))
 er_income_results <- read_rds(paste0(file_path, "Output/er_income_results.rds"))
@@
 er_tidy <- results_function(er_results, "ER Primary")
+er_lr_tidy <- results_function(er_results_lr, "ER Log Reg")
 er_income_tidy <- results_function(er_income_results, "ER + income")
@@
 # Combine cleaned results
-er_results_combined <- rbind(er_tidy, er_income_tidy, er_filter_comp_tidy, er_burn_level_tidy, 
+er_results_combined <- rbind(er_tidy, er_lr_tidy, er_income_tidy, er_filter_comp_tidy, er_burn_level_tidy, 
                               er_smoking_tidy, er_windows_tidy, er_door_tidy,
@@
   mutate(model = factor(model,
-                        levels = c("ER Primary", "ER + income", "ER + filt comp", "ER + burn level",
+                        levels = c("ER Primary", "ER Log Reg", "ER + income", "ER + filt comp", "ER + burn level",
                                    "ER + smoke act", "ER + window act", "ER + door act",
@@
   ggtitle(label = "Results for ER framework, LRTI outcome") +
-  labs(y = "LRTI cases per IQR \n increase in PM2.5") +
+  labs(y = "LRTI events per IQR \n increase in PM2.5") +
   labs(x = "", group = "", shape = "") +
@@
 
+ER Primary = Negative binomial mixed model, as specified above  
+ER Log Reg = Logistic regression instead of Negative binomial model framework  
 ER + income = Primary model using income instead of education  
@@
 ER - home = Primary model without home as random term
+
+```{r, eval=FALSE, include=FALSE}
+################ Plot/check individual models and various results ##############
+
+check_model_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + moisture_closest + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 20)
+
+summary(check_model_results)
+
+# Run function to clean results
+results_function <- function(model_results) {
+
+tidy_results <- tidy(model_results, conf.int = TRUE) %>% 
+  #mutate(group_filter = if_else(grepl("pm_mean", term), 1, 0)) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  dplyr::select(term, estimate, conf.low, conf.high)
+tidy_results
+
+}
+
+# Run results through function, save, combine for plotting
+results_tidy <- results_function(check_model_results)
+
+# Plot results
+plot_estimates <- results_tidy %>%
+  #filter(term != "treatment_filter_typeHoneywell") %>% 
+  ggplot(aes(group = term, shape = term)) +
+  geom_point(aes(x=1, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=1, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  labs(y = "Results compared to placebo") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "right",
+          legend.text = element_text(size = 14, colour = "black")) 
+plot_estimates
+
+# Check summary data
+grouped_data <- analysis_data %>% 
+  group_by(treatment_assigned) %>% 
+  summarize(min(moisture_closest, na.rm = TRUE),
+            median(moisture_closest, na.rm = TRUE),
+            max(moisture_closest, na.rm = TRUE),
+            mean(moisture_closest, na.rm = TRUE),
+            sd(moisture_closest, na.rm = TRUE))
+grouped_data
+#table(analysis_data$moisture_closest)
+#summary(analysis_data$moisture_closest)
+anova_model <- lm(moisture_closest ~ treatment_assigned, data = analysis_data)
+anova(anova_model)
+
+# Check model observation numbers
+table(model.frame(itt_results_int)$treatment_assigned)
+nobs(itt_results_int)
+
+# backtransform model results for log(pm) models
+((exp(-.418)-1)*100)
+```

### 2020-12-04 stove_grade_usage_cleaning.Rmd
analysis updates

--- 2020-11-13 stove_grade_usage_cleaning.Rmd
@@
 library(zoo)
+library(gridExtra)
 ```
@@
   mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
+         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
          kw_perc_expected = abs(kw_perc_expected)) %>% 
@@
   mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
+  filter(treatment_assigned == "Filter")
 
@@
   #filter(area == "WMT") %>% 
-  group_by(filter_type) %>% 
+  group_by(treatment_assigned) %>% 
   summarize("Mean KW % Exp" = mean(kw_perc_expected),
@@
 kw_perc_exp_summary
+
+png("Kids_KW_perc_exp.png", width=960, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()
 
@@
 ```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+pm_data <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  arrange(area, home, pm_mean_sampling_period) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  select(area, home, pm_mean_sampling_period)
+kids_kw_report_clean <- 
+  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds"))
+
+kw_perc_exp_summary <- kids_kw_report_clean %>% 
+  left_join(pm_data, by = c("area", "home")) %>% 
+  filter(!is.na(kw_perc_expected)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type, treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_expected),
+            "SD KW % Exp" = sd(kw_perc_expected),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_expected),
+            "Med KW % Exp" = median(kw_perc_expected),
+            "Max KW % Exp" = max(kw_perc_expected),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+kw_perc_exp_summary
+
+png("Kids_KW_perc_exp.png", width=1200, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()
+```
+
+# Try to extract sampling dates from KW hour data
+```{r}
+wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = ymd(interventiondate),
+         sampling_date = ymd(samplingdate),
+         pickup_date = ymd(pickupdate),
+         area = "WMT") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         sampling_date = mdy(samplingdate),
+         pickup_date = mdy(pickupdate),
+         area = "NN") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         sampling_date = mdy(samplingdate),
+         pickup_date = mdy(pickupdate),
+         area = "AK") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+kids_sampling_dates <- rbind(wmt_sampling_dates, nn_sampling_dates, ak_sampling_dates)
+
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+kids_kw_clean <- 
+  read_rds(paste0(file_path, "Output/kids_kw_clean.rds"))
+filter_units <- 
+  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
+  select(home_winter_id, area, filter_type)
+
+kids_kw_dates <- kids_kw_clean %>% 
+  full_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
+  mutate(filter_var = if_else(kw_date == sampling_date | kw_date == pickup_date, 1, 0)) %>% 
+  filter(filter_var == 1) %>% 
+  arrange(area, home, kw_date) %>% 
+  mutate(kw_diff = lead(kw_hours) - kw_hours) %>% 
+  filter(!is.na(kw_diff)) %>% 
+  mutate(date_diff = as.duration(interval(sampling_date, pickup_date)),
+         date_diff = as.numeric(date_diff)/86400) %>% 
+  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
+  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
+                              2.04, 0),
+         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
+                              4.224, multiplier),
+         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
+                              0.888, multiplier),
+         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
+                              1.896, multiplier),
+         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
+                              1.896, multiplier),
+         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
+                              3.48, multiplier),
+         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
+                              0.0144, multiplier),
+         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
+                              0.48, multiplier)) %>% 
+  mutate(kw_expected = date_diff*multiplier,
+         kw_perc_exp = kw_diff/kw_expected*100)
+
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+pm_data <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  arrange(area, home, pm_mean_sampling_period) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  select(area, home, pm_mean_sampling_period)
+
+kw_perc_exp_summary <- kids_kw_dates %>% 
+  left_join(pm_data, by = c("area", "home")) %>% 
+  filter(!is.na(kw_perc_exp)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type, treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_exp),
+            "SD KW % Exp" = sd(kw_perc_exp),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_exp),
+            "Med KW % Exp" = median(kw_perc_exp),
+            "Max KW % Exp" = max(kw_perc_exp),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+kw_perc_exp_summary
+
+png("Kids_KW_perc_exp_sampling.png", width=1200, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()
+```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_kw_clean <- 
+  read_rds(paste0(file_path, "Output/kids_kw_clean.rds"))
+
+filter_units <- 
+  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
+  select(home_winter_id, area, filter_type)
+
+wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = ymd(interventiondate),
+         sampling_date = ymd(samplingdate),
+         pickup_date = ymd(pickupdate),
+         area = "WMT") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         sampling_date = mdy(samplingdate),
+         pickup_date = mdy(pickupdate),
+         area = "NN") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         sampling_date = mdy(samplingdate),
+         pickup_date = mdy(pickupdate),
+         area = "AK") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+kids_sampling_dates <- rbind(wmt_sampling_dates, nn_sampling_dates, ak_sampling_dates)
+
+kids_kw_filtered <- kids_kw_clean %>% 
+  group_by(area, home_winter_id) %>% 
+  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
+  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
+  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
+  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
+  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
+  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
+  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
+  mutate(max_kw_hours = if_else(kw_hours == max(kw_hours), 1, 0)) %>% 
+  filter(max_kw_hours == 1) %>% 
+  left_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
+  mutate(date_diff = as.duration(interval(intervention_date, kw_date)),
+         date_diff = as.numeric(date_diff)/86400) %>% 
+  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
+  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
+                              2.04, 0),
+         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
+                              4.224, multiplier),
+         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
+                              0.888, multiplier),
+         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
+                              1.896, multiplier),
+         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
+                              1.896, multiplier),
+         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
+                              3.48, multiplier),
+         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
+                              0.0144, multiplier),
+         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
+                              0.48, multiplier)) %>% 
+  mutate(kw_expected = date_diff*multiplier,
+         kw_perc_exp = kw_hours/kw_expected*100)
+
+pm_data <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  arrange(area, home, pm_mean_sampling_period) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  select(area, home, pm_mean_sampling_period)
+
+kw_perc_exp_summary <- kids_kw_filtered %>% 
+  left_join(pm_data, by = c("area", "home")) %>% 
+  filter(!is.na(kw_perc_exp)) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type, treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_exp),
+            "SD KW % Exp" = sd(kw_perc_exp),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_exp),
+            "Med KW % Exp" = median(kw_perc_exp),
+            "Max KW % Exp" = max(kw_perc_exp),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+kw_perc_exp_summary
+
+png("Kids_KW_perc_exp_new.png", width=1200, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()  
+```

### 2020-12-08 main_analysis_final.Rmd
main analysis updates

@@
 author: "Ethan Walker"
-date: "Updated 30 Nov 2020"
+date: "Updated 7 Dec 2020"
 output: pdf_document
@@
   filter(winter_id == 2)
+
+analysis_data_noak <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         lrti_events_di_total = as.factor(lrti_events_di_total),
+         person_time_at_risk_iqr = person_time_at_risk/20) %>% 
+  filter(area != "AK")
 ```
@@
 write_rds(itt_winter2_results, paste0(file_path, "Output/itt_winter2_results.rds"))
+
+itt_noak_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk +
+                     (1 | home:cohort:area),
+                     data = analysis_data_noak, family = binomial, nAGQ = 20)
+write_rds(itt_noak_results, paste0(file_path, "Output/itt_noak_results.rds"))
 
@@
 itt_winter1_results <- read_rds(paste0(file_path, "Output/itt_winter1_results.rds"))
+itt_noak_results <- read_rds(paste0(file_path, "Output/itt_noak_results.rds"))
 #itt_winter2_results <- read_rds(paste0(file_path, "Output/itt_winter2_results.rds"))
@@
 itt_winter1_tidy <- results_function(itt_winter1_results, "ITT Winter 1")
+itt_noak_tidy <- results_function(itt_noak_results, "ITT no AK")
 #itt_winter2_tidy <- results_function(itt_winter2_results, "ITT Winter 2")
@@
 # Combine cleaned results
-itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, 
+itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, itt_noak_tidy,
                               itt_income_tidy, itt_filter_comp_tidy,
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT Winter 1", 
+                        levels = c("ITT Primary", "ITT Winter 1", "ITT no AK",
                                    "ITT + income", "ITT + filt comp",
@@
 ITT Winter 1 = Primary model with Winter 1 data only  
+ITT no AK = Primary model without homes from AK (different placebo group)  
 ITT + income = Primary model plus categorical income variable  
@@
 
-```{r, eval=FALSE, include=FALSE}
+```{r, eval=FALSE, include=FALSE, fig.width=8, fig.height=5}
 # Effect modification/interaction analysis
-# Try home_floors_2level, home_sqft_2level, education_3level, child_age
-# home_year_built_2level, stove_age_3level, stove_grade_3level, chimney_clean_3level,
-# wood_collect_2level, burn_level_3level, residents_smoke
+
+int_function <- function(model_data, interaction_var) {
+
+  new_model_data <- model_data %>% 
+    rename(int_var = interaction_var)
 
 # Run primary model and print results
-itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*residents_smoke + 
+itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
                               child_age + 
-                       (1 | cohort:area), data = analysis_data_winter1)
+                       (1 | cohort:area), data = new_model_data)
 #write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
@@
 
-int_emmeans <- emmeans(itt_pm_results_int, pairwise ~ treatment_assigned | residents_smoke)
+int_emmeans <- emmeans(itt_pm_results_int, pairwise ~ treatment_assigned | int_var)
 #summary(int_emmeans, conf.int = TRUE)
-confint(int_emmeans)
+save_results <- confint(int_emmeans)
+save_results <<- confint(int_emmeans$contrasts)
+#save_results
 
-table(model.frame(itt_results_int)$treatment_assigned)
-nobs(itt_results_int)
+plot_results <- as.data.frame(save_results$contrasts) %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(lower.CL)-1)*100, digits = 2),
+         conf.high = round((exp(upper.CL)-1)*100, digits = 2)) %>% 
+  ggplot(aes(group = contrast, shape = contrast)) +
+  geom_point(aes(x=int_var, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Interaction contrasts for ITT framework, PM outcome",
+          subtitle = paste0("Interaction term = ", interaction_var)) +
+  labs(y = "% difference in PM2.5") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black"))
 
-((exp(-.418)-1)*100)
+summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+
+plot_results
+
+}
+
+int_function(analysis_data_winter1, "home_sqft_2level")
+
+# Try: home_floors_2level, home_sqft_2level, education_3level, child_age
+# home_year_built_2level, stove_age_3level, stove_grade_3level, chimney_clean_3level,
+# wood_collect_2level, wood_collect_method_2level, burn_level_3level, residents_smoke
+
+#table(model.frame(itt_results_int)$treatment_assigned)
+#((exp(-.418)-1)*100)
 ```
@@
 # Run and save model results
+
+itt_pm_noak_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
+                       (1 | cohort:area), data = analysis_data_noak)
+write_rds(itt_pm_noak_results, paste0(file_path, "Output/itt_pm_noak_results.rds"))
 
@@
                                   child_age + (1 | cohort:area), data = analysis_data_winter1)
+summary(itt_pm_filter_type_results)
+tidy_results <- tidy(itt_pm_filter_type_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_filter_type", "", term),
+         model = "Filter type") %>% 
+  mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
+         conf.low = round((exp(conf.low)-1)*100, digits = 2),
+         conf.high = round((exp(conf.high)-1)*100, digits = 2)) %>% 
+  dplyr::select(model, term, estimate, conf.low, conf.high)
+tidy_results
+table(model.frame(itt_pm_filter_type_results)$treatment_filter_type)
 write_rds(itt_pm_filter_type_results, paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
@@
 itt_pm_results <- read_rds(paste0(file_path, "Output/itt_pm_results.rds"))
+itt_pm_noak_results <- read_rds(paste0(file_path, "Output/itt_pm_noak_results.rds"))
 itt_pm_income_results <- read_rds(paste0(file_path, "Output/itt_pm_income_results.rds"))
@@
 itt_pm_tidy <- results_function(itt_pm_results, "ITT Primary")
+itt_pm_noak_tidy <- results_function(itt_pm_noak_results, "ITT no AK")
 itt_pm_income_tidy <- results_function(itt_pm_income_results, "ITT + income")
@@
 # Combine cleaned results
-itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_income_tidy, itt_pm_filter_comp_tidy, 
+itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_noak_tidy, itt_pm_income_tidy, itt_pm_filter_comp_tidy, 
                               itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT + income", "ITT + filt comp", 
+                        levels = c("ITT Primary", "ITT no AK", "ITT + income", "ITT + filt comp", 
                                    "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
@@
 ITT Primary = Linear mixed model, as specified above  
+ITT no AK = Primary model without homes from AK (different placebo group)  
 ITT + income = Primary model plus categorical income variable  
@@
 # backtransform model results for log(pm) models
-((exp(-.418)-1)*100)
+((exp(-.374)-1)*100)
 ```
+
+```{rr, eval=FALSE, include=FALSE}
+# ER framework using a Spline for PM
+
+# s() is the command for a spline
+# fx=TRUE makes a fixed df regression spline, as opposed to a penalized spline
+# bs = "cr" makes this a cubic regression spline
+## Default is "tp" for thin plate regression spline
+## Both seem to have the same effect on the model
+# k = 6 sets the degrees of freedom (df = k-1)
+
+library(splines)
+library(gamm4)
+
+model_pm_spline <- gamm4(lrti_events_di_winter ~ s(pm_mean_sampling_period_iqr, 
+                          fx = TRUE, bs = "tp", k = 6) + 
+                          child_age +  education_3level + residents_smoke +
+                          person_time_at_risk_winter, 
+                          data = analysis_data_winter1, family = binomial(link = "logit"), 
+                          random = ~(1 | home:cohort:area))
+plot(model_pm_spline$gam, seWithMean = TRUE, select = 1, shade = TRUE,
+     xlab = expression(paste("IQR increase in area PM"[2.5], 
+                             " (", mu, g/m^3, ")")),
+     ylab = expression(paste("OR, spline trend for area PM  "[2.5])))
+summary(model_pm_spline$gam)
+summary(model_pm_spline$mer)
+
+# Not sure how to add the offset in
+er_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + 
+                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
+                          data = analysis_data_winter1)
+```

### 2020-12-09 main_analysis_final.Rmd
main analysis updates

@@
 author: "Ethan Walker"
-date: "Updated 7 Dec 2020"
+date: "Updated 9 Dec 2020"
 output: pdf_document
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 #write_rds(itt_results, paste0(file_path, "Output/itt_results.rds"))
+
+#table(model.frame(itt_results)$treatment_assigned)
+#nobs(itt_results)
 ```
@@
                      (1 | home:cohort:area),
-                     data = analysis_data_winter1, family = binomial, nAGQ = 20)
+                     data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(itt_winter1_results, paste0(file_path, "Output/itt_winter1_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data_winter2, family = binomial, nAGQ = 20)
+                     data = analysis_data_winter2, family = binomial, nAGQ = 10)
 write_rds(itt_winter2_results, paste0(file_path, "Output/itt_winter2_results.rds"))
+
+itt_filter_type_results <- glmer(lrti_events_di_total ~ treatment_filter_type + 
+                     child_age + person_time_at_risk +
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 10)
+
+summary(itt_filter_type_results)
+
+tidy_results <- tidy(itt_filter_type_results, conf.int = TRUE) %>% 
+  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
+  filter(group_filter == 1) %>% 
+  mutate(term = gsub("treatment_filter_type", "", term),
+         model = "Filter type") %>% 
+  mutate(estimate = round(estimate, digits = 2),
+         conf.low = round(conf.low, digits = 2),
+         conf.high = round(conf.high, digits = 2)) %>% 
+  dplyr::select(model, term, estimate, conf.low, conf.high)
+tidy_results
+
+table(model.frame(itt_filter_type_results)$treatment_filter_type)
+
+analysis_check <- analysis_data %>% 
+  mutate(pm_na = if_else(!is.na(pm_mean_sampling_period), "PM Yes", "PM Missing"))
+
+save_table <- table(analysis_check$pm_na, analysis_check$lrti_events_di_total, analysis_check$treatment_filter_type)
+save_table
+ftable(save_table, 1)
+
+write_rds(itt_filter_type_results, paste0(file_path, "Output/itt_filter_type_results.rds"))
 
@@
                      (1 | home:cohort:area),
-                     data = analysis_data_noak, family = binomial, nAGQ = 20)
+                     data = analysis_data_noak, family = binomial, nAGQ = 10)
 write_rds(itt_noak_results, paste0(file_path, "Output/itt_noak_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_income_results, paste0(file_path, "Output/itt_income_results.rds"))
-
-itt_filter_type_results <- glmer(lrti_events_di_total ~ treatment_filter_type +  
-                     child_age + person_time_at_risk + 
-                     (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
-write_rds(itt_filter_type_results, paste0(file_path, "Output/itt_filter_type_results.rds"))
 
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_filter_comp_results, paste0(file_path, "Output/itt_filter_comp_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_smoke_results, paste0(file_path, "Output/itt_smoke_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_burn_level_results, paste0(file_path, "Output/itt_burn_level_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_smoking_results, paste0(file_path, "Output/itt_smoking_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_windows_results, paste0(file_path, "Output/itt_windows_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_door_results, paste0(file_path, "Output/itt_door_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_sweep_results, paste0(file_path, "Output/itt_sweep_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_ambpm_results, paste0(file_path, "Output/itt_ambpm_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_ambtemp_results, paste0(file_path, "Output/itt_ambtemp_results.rds"))
@@
                      (1 | cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
+                     data = analysis_data, family = binomial, nAGQ = 10)
 write_rds(itt_home_results, paste0(file_path, "Output/itt_home_results.rds"))
@@
                      (1 | home:cohort:area),
-                     data = analysis_data_outliers, family = binomial, nAGQ = 20)
+                     data = analysis_data_outliers, family = binomial, nAGQ = 10)
 write_rds(itt_outlier_results, paste0(file_path, "Output/itt_outlier_results.rds"))
@@
 #itt_filter_type_results <- read_rds(paste0(file_path, "Output/itt_filter_type_results.rds"))
-itt_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_filter_comp_results.rds"))
+#itt_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_filter_comp_results.rds"))
 itt_smoke_results <- read_rds(paste0(file_path, "Output/itt_smoke_results.rds"))
@@
 itt_ambtemp_results <- read_rds(paste0(file_path, "Output/itt_ambtemp_results.rds"))
-itt_home_results <- read_rds(paste0(file_path, "Output/itt_home_results.rds"))
+#itt_home_results <- read_rds(paste0(file_path, "Output/itt_home_results.rds"))
 itt_outlier_results <- read_rds(paste0(file_path, "Output/itt_outlier_results.rds"))
@@
 #itt_filter_type_tidy <- results_function(itt_filter_type_results, "ITT + filt type")
-itt_filter_comp_tidy <- results_function(itt_filter_comp_results, "ITT + filt comp")
+#itt_filter_comp_tidy <- results_function(itt_filter_comp_results, "ITT + filt comp")
 itt_smoke_tidy <- results_function(itt_smoke_results, "ITT + smoke hx")
@@
 itt_ambtemp_tidy <- results_function(itt_ambtemp_results, "ITT + amb temp")
-itt_home_tidy <- results_function(itt_home_results, "ITT - home")
+#itt_home_tidy <- results_function(itt_home_results, "ITT - home")
 itt_outlier_tidy <- results_function(itt_outlier_results, "ITT outlier")
@@
 itt_results_combined <- rbind(itt_tidy, itt_winter1_tidy, itt_noak_tidy,
-                              itt_income_tidy, itt_filter_comp_tidy,
+                              itt_income_tidy, 
                               itt_smoke_tidy, itt_burn_level_tidy, itt_smoking_tidy,
@@
                               itt_ambpm_tidy, itt_ambtemp_tidy,
-                              itt_home_tidy, itt_outlier_tidy)
+                              itt_outlier_tidy)
 
@@
                         levels = c("ITT Primary", "ITT Winter 1", "ITT no AK",
-                                   "ITT + income", "ITT + filt comp",
+                                   "ITT + income", 
                                    "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
@@
                                    "ITT + amb PM", "ITT + amb temp",
-                                   "ITT - home", "ITT outlier"))) %>% 
+                                   "ITT outlier"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
 ITT + income = Primary model plus categorical income variable  
-ITT + filt comp = Primary model plus variable for filter compliance  
 ITT + smoke hx = Primary model plus self-reported smoking variable  
@@
 ITT + amb temp = Primary model plus amb temp during LRTI at-risk period  
-ITT - home = Primary model without home as random term  
 ITT outlier = Primary model with data that fixes incidence outlier
@@
 # Run primary model and print results
-itt_pm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
+itt_pm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                        (1 | cohort:area), data = analysis_data_winter1)
 #write_rds(itt_pm_results, paste0(file_path, "Output/itt_pm_results.rds"))
+# Adding the randomization variable for child_age leads to the exact same model results
+# Adding a random term for home adversely impacts the model since there is only one
+# distinct observation per home
+
+#table(model.frame(itt_pm_results)$treatment_assigned)
+#nobs(itt_pm_results)
 ```
@@
 itt_pm_income_results <- read_rds(paste0(file_path, "Output/itt_pm_income_results.rds"))
-itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
-itt_pm_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
+#itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
+#itt_pm_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
 itt_pm_smoke_results <- read_rds(paste0(file_path, "Output/itt_pm_smoke_results.rds"))
@@
 itt_pm_income_tidy <- results_function(itt_pm_income_results, "ITT + income")
-itt_pm_filter_type_tidy <- results_function(itt_pm_filter_type_results, "ITT filter type")
-itt_pm_filter_comp_tidy <- results_function(itt_pm_filter_comp_results, "ITT + filt comp")
+#itt_pm_filter_type_tidy <- results_function(itt_pm_filter_type_results, "ITT filter type")
+#itt_pm_filter_comp_tidy <- results_function(itt_pm_filter_comp_results, "ITT + filt comp")
 itt_pm_smoke_tidy <- results_function(itt_pm_smoke_results, "ITT + smoke hx")
@@
 # Combine cleaned results
-itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_noak_tidy, itt_pm_income_tidy, itt_pm_filter_comp_tidy, 
+itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_noak_tidy, itt_pm_income_tidy, 
                               itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT no AK", "ITT + income", "ITT + filt comp", 
+                        levels = c("ITT Primary", "ITT no AK", "ITT + income", 
                                    "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
@@
 # Run primary model and print results
-er_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + 
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+# The Neg Bin model was overfit; see isSingular() for further explanation
+
+er_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          education_3level + residents_smoke + person_time_at_risk_winter +
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 #write_rds(er_results, paste0(file_path, "Output/er_results.rds"))
+
+#nobs(er_results)
+#isSingular(er_results)
 ```
@@
 # Run primary model and print results
-er_results_int <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr*chimney_clean_3level + 
+er_results_int <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*chimney_clean_3level + 
                           child_age + 
@@
                           residents_smoke + 
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+                          person_time_at_risk_winter +
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 #write_rds(er_results_int, paste0(file_path, "Output/er_results_int_age.rds"))
@@
 
-er_results_lr <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + person_time_at_risk_winter +
+er_income_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          residents_smoke + person_time_at_risk_winter + 
+                          income_3level + 
                           (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 20)
-write_rds(er_results_lr, paste0(file_path, "Output/er_results_lr.rds"))
-
-er_income_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + 
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_income_results, paste0(file_path, "Output/er_income_results.rds"))
 
-er_filter_comp_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + kw_perc_exp_cat2 +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_filter_comp_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter + 
+                          kw_perc_exp_cat2 + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_filter_comp_results, paste0(file_path, "Output/er_filter_comp_results.rds"))
 
-er_burn_level_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + burn_level_3level +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_burn_level_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter + 
+                          burn_level_3level + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_burn_level_results, paste0(file_path, "Output/er_burn_level_results.rds"))
 
-er_smoking_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + home_act_smoking_sum +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_smoking_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + person_time_at_risk_winter + 
+                          home_act_smoking_sum + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_smoking_results, paste0(file_path, "Output/er_smoking_results.rds"))
 
-er_windows_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + home_act_windows_sum +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_windows_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter + 
+                          home_act_windows_sum + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_windows_results, paste0(file_path, "Output/er_windows_results.rds"))
 
-er_door_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + home_act_door_sum +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_door_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter + 
+                          home_act_door_sum + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_door_results, paste0(file_path, "Output/er_door_results.rds"))
 
-er_sweep_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + home_act_sweep_sum +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_sweep_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter +
+                          home_act_sweep_sum + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_sweep_results, paste0(file_path, "Output/er_sweep_results.rds"))
 
-er_ambpm_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + amb_pm_lrti +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_ambpm_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter +
+                          amb_pm_lrti + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_ambpm_results, paste0(file_path, "Output/er_ambpm_results.rds"))
 
-er_ambtemp_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + amb_temp_lrti +
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
+er_ambtemp_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
+                          income_3level + residents_smoke + person_time_at_risk_winter +
+                          amb_temp_lrti + 
+                          (1 | home:cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
 write_rds(er_ambtemp_results, paste0(file_path, "Output/er_ambtemp_results.rds"))
-
-er_home_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + 
-                          (1 | cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
-write_rds(er_home_results, paste0(file_path, "Output/er_home_results.rds"))
 ```
@@
 er_results <- read_rds(paste0(file_path, "Output/er_results.rds"))
-er_results_lr <- read_rds(paste0(file_path, "Output/er_results_lr.rds"))
 er_income_results <- read_rds(paste0(file_path, "Output/er_income_results.rds"))
-er_filter_comp_results <- read_rds(paste0(file_path, "Output/er_filter_comp_results.rds"))
+#er_filter_comp_results <- read_rds(paste0(file_path, "Output/er_filter_comp_results.rds"))
 er_burn_level_results <- read_rds(paste0(file_path, "Output/er_burn_level_results.rds"))
@@
 er_ambtemp_results <- read_rds(paste0(file_path, "Output/er_ambtemp_results.rds"))
-er_home_results <- read_rds(paste0(file_path, "Output/er_home_results.rds"))
 
@@
 er_tidy <- results_function(er_results, "ER Primary")
-er_lr_tidy <- results_function(er_results_lr, "ER Log Reg")
 er_income_tidy <- results_function(er_income_results, "ER + income")
-er_filter_comp_tidy <- results_function(er_filter_comp_results, "ER + filt comp")
+#er_filter_comp_tidy <- results_function(er_filter_comp_results, "ER + filt comp")
 er_burn_level_tidy <- results_function(er_burn_level_results, "ER + burn level")
@@
 er_ambtemp_tidy <- results_function(er_ambtemp_results, "ER + amb temp")
-er_home_tidy <- results_function(er_home_results, "ER - home")
 
 # Combine cleaned results
-er_results_combined <- rbind(er_tidy, er_lr_tidy, er_income_tidy, er_filter_comp_tidy, er_burn_level_tidy, 
+er_results_combined <- rbind(er_tidy, er_income_tidy, er_burn_level_tidy, 
                               er_smoking_tidy, er_windows_tidy, er_door_tidy,
-                              er_sweep_tidy, er_ambpm_tidy, er_ambtemp_tidy, er_home_tidy)
+                              er_sweep_tidy, er_ambpm_tidy, er_ambtemp_tidy)
 
@@
   mutate(model = factor(model,
-                        levels = c("ER Primary", "ER Log Reg", "ER + income", "ER + filt comp", "ER + burn level",
+                        levels = c("ER Primary", "ER + income", "ER + burn level",
                                    "ER + smoke act", "ER + window act", "ER + door act",
-                                   "ER + sweep act", "ER + amb PM", "ER + amb temp", "ER - home"))) %>% 
+                                   "ER + sweep act", "ER + amb PM", "ER + amb temp"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
 
-ER Primary = Negative binomial mixed model, as specified above  
-ER Log Reg = Logistic regression instead of Negative binomial model framework  
+ER Primary = Logistic regression mixed model, as specified above  
 ER + income = Primary model using income instead of education  
-ER + filt comp = Primary model plus variable for filter compliance  
 ER + burn level = Primary model plus self-reported burn-level compared to typical  
@@
 ER + amb temp = Primary model plus amb temp during LRTI at-risk period
-ER - home = Primary model without home as random term
 

### 2020-12-09 stove_grade_usage_cleaning.Rmd
main analysis updates

@@
 ```
+
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+kids_kw_clean <- 
+  read_rds(paste0(file_path, "Output/kids_kw_clean.rds"))
+
+filter_units <- 
+  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
+  select(home_winter_id, area, filter_type)
+
+wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = ymd(interventiondate),
+         sampling_date = ymd(samplingdate),
+         pickup_date = ymd(pickupdate),
+         area = "WMT") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         sampling_date = mdy(samplingdate),
+         pickup_date = mdy(pickupdate),
+         area = "NN") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, interventiondate, 
+         samplingdate, pickupdate) %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         intervention_date = mdy(interventiondate),
+         sampling_date = mdy(samplingdate),
+         pickup_date = mdy(pickupdate),
+         area = "AK") %>% 
+  select(home_winter_id, intervention_date, 
+         sampling_date, pickup_date, area)
+
+kids_sampling_dates <- rbind(wmt_sampling_dates, nn_sampling_dates, ak_sampling_dates)
+
+kids_kw_filtered <- kids_kw_clean %>% 
+  left_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
+  mutate(date_diff = as.duration(interval(intervention_date, kw_date)),
+         date_diff = as.numeric(date_diff)/86400) %>% 
+  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
+  distinct(area, home_winter_id, .keep_all = TRUE) %>% 
+  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
+                              2.04, 0),
+         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
+                              4.224, multiplier),
+         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
+                              0.888, multiplier),
+         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
+                              1.896, multiplier),
+         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
+                              1.896, multiplier),
+         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
+                              3.48, multiplier),
+         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
+                              0.0144, multiplier),
+         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
+                              0.48, multiplier)) %>% 
+  mutate(kw_expected = date_diff*multiplier,
+         kw_perc_exp = kw_hours/kw_expected*100)
+
+pm_data <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  arrange(area, home, pm_mean_sampling_period) %>% 
+  distinct(area, home, .keep_all = TRUE) %>% 
+  select(area, home, pm_mean_sampling_period)
+
+kw_perc_exp_summary <- kids_kw_filtered %>% 
+  left_join(pm_data, by = c("area", "home")) %>% 
+  filter(!is.na(kw_perc_exp)) %>% 
+  filter(kw_perc_exp < 30000) %>% 
+  #filter(area == "WMT") %>% 
+  group_by(filter_type, treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_exp, na.rm = TRUE),
+            "SD KW % Exp" = sd(kw_perc_exp, na.rm = TRUE),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_exp, na.rm = TRUE),
+            "Med KW % Exp" = median(kw_perc_exp, na.rm = TRUE),
+            "Max KW % Exp" = max(kw_perc_exp, na.rm = TRUE),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+kw_perc_exp_summary
+
+png("Kids_KW_perc_exp_new.png", width=1200, height=480, bg = "white")
+grid.table(kw_perc_exp_summary)
+dev.off()  
+```

### 2020-12-11 main_analysis_final.Rmd
update function for interaction analysis

@@
 # Run primary model and print results
-itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
+itt_pm_results_int <<- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
                               child_age + 
@@
 #summary(int_emmeans, conf.int = TRUE)
-save_results <- confint(int_emmeans)
-save_results <<- confint(int_emmeans$contrasts)
+save_contrasts_full <<- summary(int_emmeans)
+save_contrasts <<- confint(int_emmeans$contrasts)
+obs <<- table(model.frame(itt_pm_results_int)$treatment_assigned)
+n_table <<- table(new_model_data$int_var, new_model_data$treatment_assigned)
 #save_results
@@
 int_function(analysis_data_winter1, "home_sqft_2level")
+n_table
+save_contrasts
+summary_table
+save_contrasts_full
 

### 2020-12-12 main_analysis_final.Rmd
interaction analysis updates

--- 2020-12-09 main_analysis_final.Rmd
@@
 author: "Ethan Walker"
-date: "Updated 9 Dec 2020"
+date: "Updated 11 Dec 2020"
 output: pdf_document
@@
 
-```{r, eval=FALSE, include=FALSE}
+```{r, eval=FALSE, include=FALSE, fig.width=8, fig.height=5}
 # Effect modification/interaction analysis
-# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
+
+int_function <- function(model_data, interaction_var) {
+
+  new_model_data <- model_data %>% 
+    rename(int_var = interaction_var)
 
 # Run primary model and print results
-itt_results_int <- glmer(lrti_events_di_total ~ treatment_assigned*residents_smoke + 
+itt_results_int <<- glmer(lrti_events_di_total ~ treatment_assigned*int_var + 
                           child_age + 
                           person_time_at_risk + (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 20)
-#write_rds(itt_results_int, paste0(file_path, "Output/itt_results_int_age.rds"))
-summary(itt_results_int)
-
-int_emmeans <- emmeans(itt_results_int, pairwise ~ treatment_assigned | residents_smoke)
-summary(int_emmeans, conf.int = TRUE)
-confint(int_emmeans)
-
-table(model.frame(itt_results_int)$treatment_assigned)
-nobs(itt_results_int)
+                          data = new_model_data, family = binomial, nAGQ = 20)
+#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+#summary(itt_pm_results_int)
+
+int_emmeans <- emmeans(itt_results_int, revpairwise ~ treatment_assigned | int_var)
+#summary(int_emmeans, conf.int = TRUE)
+save_contrasts_full <<- summary(int_emmeans)
+save_contrasts <<- confint(int_emmeans$contrasts) %>% 
+  mutate(estimate = round(estimate, digits = 2),
+         conf.low = round(asymp.LCL, digits = 2),
+         conf.high = round(asymp.UCL, digits = 2)) %>% 
+  dplyr::select(contrast, int_var, estimate, conf.low, conf.high)
+obs <<- table(model.frame(itt_results_int)$treatment_assigned)
+n_table <<- table(new_model_data$int_var, new_model_data$treatment_assigned)
+#save_results
+
+plot_results <- save_contrasts %>% 
+  ggplot(aes(group = contrast, shape = contrast)) +
+  geom_point(aes(x=int_var, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Interaction contrasts for ITT framework, LRTI outcome",
+          subtitle = paste0("Interaction term = ", interaction_var)) +
+  labs(y = "Odds of LRTI vs placebo") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black"))
+
+pm_summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+
+plot_results
+
+}
+
+int_function(analysis_data, "residents_smoke")
+n_table
+save_contrasts
+pm_summary_table
+save_contrasts_full
+
+# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
 ```
@@
 # Run primary model and print results
-itt_pm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+itt_pm_results <- lmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
                        (1 | cohort:area), data = analysis_data_winter1)
@@
 # Run primary model and print results
-itt_pm_results_int <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
+itt_pm_results_int <<- lmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
                               child_age + 
@@
 
-int_emmeans <- emmeans(itt_pm_results_int, pairwise ~ treatment_assigned | int_var)
+int_emmeans <- emmeans(itt_pm_results_int, revpairwise ~ treatment_assigned | int_var)
 #summary(int_emmeans, conf.int = TRUE)
-save_results <- confint(int_emmeans)
-save_results <<- confint(int_emmeans$contrasts)
-#save_results
-
-plot_results <- as.data.frame(save_results$contrasts) %>% 
+save_contrasts_full <<- summary(int_emmeans)
+save_contrasts <<- confint(int_emmeans$contrasts) %>% 
   mutate(estimate = round((exp(estimate)-1)*100, digits = 2),
@@
          conf.high = round((exp(upper.CL)-1)*100, digits = 2)) %>% 
+  dplyr::select(contrast, int_var, estimate, conf.low, conf.high)
+obs <<- table(model.frame(itt_pm_results_int)$treatment_assigned)
+n_table <<- table(new_model_data$int_var, new_model_data$treatment_assigned)
+#save_results
+
+plot_results <- save_contrasts %>% 
   ggplot(aes(group = contrast, shape = contrast)) +
@@
 
-summary_table <<- new_model_data %>% 
+pm_summary_table <<- new_model_data %>% 
   filter(!is.na(int_var)) %>% 
@@
 
-int_function(analysis_data_winter1, "home_sqft_2level")
+int_function(analysis_data_winter1, "burn_level_3level")
+n_table
+save_contrasts
+pm_summary_table
+save_contrasts_full
 
@@
 # Run primary model and print results
-er_results_int <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*chimney_clean_3level + 
-                          child_age + 
+er_results_int <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*child_age + 
+                          #child_age + 
                           education_3level + 
@@
                           (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
+                          data = analysis_data_winter1, family = binomial, nAGQ = 20)
 #write_rds(er_results_int, paste0(file_path, "Output/er_results_int_age.rds"))
@@
 
-int_emtrends <- emtrends(er_results_int, pairwise ~ chimney_clean_3level, var = "pm_mean_sampling_period_iqr")
+int_emtrends <- emtrends(er_results_int, pairwise ~ child_age, var = "pm_mean_sampling_period_iqr")
 summary(int_emtrends, conf.int = TRUE)
 
-table(model.frame(itt_results_int)$treatment_assigned)
 nobs(itt_results_int)
+```
+
+```{r, eval=FALSE, include=FALSE, fig.width=8, fig.height=5}
+# Effect modification/interaction analysis
+
+int_function <- function(model_data, interaction_var) {
+
+  new_model_data <- model_data %>% 
+    rename(int_var = interaction_var)
+
+# Run primary model and print results
+er_results_int <<- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*int_var + 
+                          #child_age + 
+                          education_3level + 
+                          residents_smoke + 
+                          person_time_at_risk_winter +
+                          (1 | home:cohort:area), 
+                          data = new_model_data, family = binomial, nAGQ = 20)
+#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+#summary(itt_pm_results_int)
+
+int_emtrends <- emtrends(er_results_int, pairwise ~ int_var, var = "pm_mean_sampling_period_iqr")
+
+save_contrasts <<- confint(int_emtrends$emtrends) %>% 
+  mutate(estimate = round(pm_mean_sampling_period_iqr.trend, digits = 2),
+         conf.low = round(asymp.LCL, digits = 2),
+         conf.high = round(asymp.UCL, digits = 2)) %>% 
+  dplyr::select(int_var, estimate, conf.low, conf.high)
+
+plot_results <- save_contrasts %>% 
+  ggplot() +
+  geom_point(aes(x=int_var, y=estimate), 
+             position = position_dodge(width = 0.5), size = 4) +
+  geom_errorbar(aes(x=int_var, ymin=conf.low, ymax=conf.high), 
+             position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +  
+  ggtitle(label = "Interaction contrasts for ER framework, LRTI outcome",
+          subtitle = paste0("Interaction term = ", interaction_var)) +
+  labs(y = "Odds of LRTI IQR increase in PM2.5") +
+  labs(x = "", group = "", shape = "") +
+  theme(title = element_text(size = 16), 
+          axis.text.x = element_text(size = 16, colour = "black", angle = 45,
+                                     hjust = 1, vjust = .8),
+          axis.text.y = element_text(size = 16, colour = "black"),
+          axis.title.y = element_text(size = 16,
+                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
+          axis.line.x = element_line(colour = "black", size = 1.2), 
+          axis.line.y = element_line(colour = "black", size = 1.2), 
+          axis.ticks = element_blank(), 
+          panel.border = element_blank(), 
+          panel.grid = element_blank(),
+          legend.position = "top",
+          legend.text = element_text(size = 14, colour = "black"))
+
+pm_summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var) %>% 
+  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+
+plot_results
+
+}
+
+int_function(analysis_data_winter1, "child_age")
+save_contrasts
+pm_summary_table
+
+# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
 ```

### 2020-12-15 main_analysis_final.Rmd
main analysis updates

@@
 author: "Ethan Walker"
-date: "Updated 11 Dec 2020"
+date: "Updated 14 Dec 2020"
 output: pdf_document
@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25) %>% 
+         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         lrti_events_di_winter = as.factor(lrti_events_di_winter)) %>% 
   filter(person_time_at_risk_winter > 0) %>% 
@@
                      (1 | home:cohort:area),
-                     data = analysis_data, family = binomial, nAGQ = 10)
+                     data = analysis_data, family = binomial, nAGQ = 10,
+                     control = glmerControl(optimizer = "bobyqa"))
 #write_rds(itt_results, paste0(file_path, "Output/itt_results.rds"))
@@
 summary(itt_results)
+#tidy(itt_results, conf.int = T)
 itt_res <- simulateResiduals(itt_results)
@@
                           child_age + 
-                          person_time_at_risk + (1 | home:cohort:area),
-                          data = new_model_data, family = binomial, nAGQ = 20)
+                          person_time_at_risk + (1 | cohort:area),
+                          data = new_model_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
 #write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
@@
 save_contrasts <<- confint(int_emmeans$contrasts) %>% 
-  mutate(estimate = round(estimate, digits = 2),
-         conf.low = round(asymp.LCL, digits = 2),
-         conf.high = round(asymp.UCL, digits = 2)) %>% 
+  mutate(estimate = round((exp(estimate)), digits = 2),
+         conf.low = round((exp(asymp.LCL)), digits = 2),
+         conf.high = round((exp(asymp.UCL)), digits = 2)) %>% 
   dplyr::select(contrast, int_var, estimate, conf.low, conf.high)
@@
              position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 1) +
   theme_bw() +  
@@
          model = "Filter type") %>% 
-  mutate(estimate = round(estimate, digits = 2),
-         conf.low = round(conf.low, digits = 2),
-         conf.high = round(conf.high, digits = 2)) %>% 
+  mutate(estimate = round(exp(estimate), digits = 2),
+         conf.low = round(exp(conf.low), digits = 2),
+         conf.high = round(exp(conf.high), digits = 2)) %>% 
   dplyr::select(model, term, estimate, conf.low, conf.high)
@@
          model = model_name) %>% 
+  mutate(estimate = round((exp(estimate)), digits = 2),
+         conf.low = round((exp(conf.low)), digits = 2),
+         conf.high = round((exp(conf.high)), digits = 2)) %>% 
   dplyr::select(model, term, estimate, p.value, conf.low, conf.high)
@@
              position = position_dodge(width = 0.5), size = 1.2, width = 0.5) +
-  geom_hline(yintercept = 0) +
   theme_bw() +  
+  geom_hline(yintercept =  1) +
   ggtitle(label = "Results for ITT framework, LRTI outcome") +
@@
                           (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
 #write_rds(er_results, paste0(file_path, "Output/er_results.rds"))
@@
 
-```{r, eval=FALSE, include=FALSE}
+```{r, eval=FALSE, include=FALSE, fig.width=8, fig.height=5}
 # Effect modification/interaction analysis
-# Try education_3level, child_age, chimney_clean_3level, wood_collect_2level, residents_smoke
+
+int_function <- function(model_data, interaction_var) {
+
+  new_model_data <- model_data %>% 
+    rename(int_var = interaction_var)
 
 # Run primary model and print results
-er_results_int <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*child_age + 
-                          #child_age + 
+er_results_int <<- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*int_var + 
+                          child_age + 
                           education_3level + 
@@
                           person_time_at_risk_winter +
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 20)
-#write_rds(er_results_int, paste0(file_path, "Output/er_results_int_age.rds"))
-#summary(er_results_int)
-
-int_emtrends <- emtrends(er_results_int, pairwise ~ child_age, var = "pm_mean_sampling_period_iqr")
-summary(int_emtrends, conf.int = TRUE)
-
-nobs(itt_results_int)
-```
-
-```{r, eval=FALSE, include=FALSE, fig.width=8, fig.height=5}
-# Effect modification/interaction analysis
-
-int_function <- function(model_data, interaction_var) {
-
-  new_model_data <- model_data %>% 
-    rename(int_var = interaction_var)
-
-# Run primary model and print results
-er_results_int <<- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*int_var + 
-                          #child_age + 
-                          education_3level + 
-                          residents_smoke + 
-                          person_time_at_risk_winter +
-                          (1 | home:cohort:area), 
-                          data = new_model_data, family = binomial, nAGQ = 20)
+                          (1 | cohort:area), 
+                          data = new_model_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
 #write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
@@
 save_contrasts <<- confint(int_emtrends$emtrends) %>% 
-  mutate(estimate = round(pm_mean_sampling_period_iqr.trend, digits = 2),
-         conf.low = round(asymp.LCL, digits = 2),
-         conf.high = round(asymp.UCL, digits = 2)) %>% 
+  mutate(estimate = round(exp(pm_mean_sampling_period_iqr.trend), digits = 2),
+         conf.low = round(exp(asymp.LCL), digits = 2),
+         conf.high = round(exp(asymp.UCL), digits = 2)) %>% 
   dplyr::select(int_var, estimate, conf.low, conf.high)
@@
 
-int_function(analysis_data_winter1, "child_age")
+int_function(analysis_data_winter1, "gender")
 save_contrasts
@@
          model = model_name) %>% 
+  mutate(estimate = round((exp(estimate)), digits = 2),
+         conf.low = round((exp(conf.low)), digits = 2),
+         conf.high = round((exp(conf.high)), digits = 2)) %>% 
   dplyr::select(model, term, estimate, conf.low, conf.high)
@@
 ((exp(-.374)-1)*100)
+
+# Check PM concentrations for subgroups
+pm_summary_table <- analysis_data_winter1 %>% 
+  filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
+           !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
+  #group_by(treatment_assigned) %>% 
+  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+pm_summary_table
 ```

### 2020-12-17 main_analysis_final.Rmd
analysis updates - splines

@@
 author: "Ethan Walker"
-date: "Updated 14 Dec 2020"
+date: "Updated 16 Dec 2020"
 output: pdf_document
@@
          lrti_events_di_total = as.factor(lrti_events_di_total),
-         person_time_at_risk_iqr = person_time_at_risk/20) 
+         person_time_at_risk_iqr = person_time_at_risk/20,
+         child_age = factor(child_age, levels = c("< 1", "1+"))) 
   #filter(gender == "Female")
@@
          lrti_events_total = as.factor(lrti_events_total),
-         lrti_events_di_total = as.factor(lrti_events_di_total))
+         lrti_events_di_total = as.factor(lrti_events_di_total),
+         child_age = factor(child_age, levels = c("< 1", "1+")))
 
@@
          pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
-         lrti_events_di_winter = as.factor(lrti_events_di_winter)) %>% 
+         lrti_events_di_winter = as.factor(lrti_events_di_winter),
+         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
   filter(person_time_at_risk_winter > 0) %>% 
@@
          lrti_events_di_total = as.factor(lrti_events_di_total),
-         person_time_at_risk_iqr = person_time_at_risk/20) %>% 
+         person_time_at_risk_iqr = person_time_at_risk/20,
+         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
   filter(area != "AK")
@@
                           control = glmerControl(optimizer = "bobyqa"))
-#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
-#summary(itt_pm_results_int)
+#write_rds(itt_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
+save_summary <<- summary(itt_results_int)
 
@@
 
+lrti_summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  mutate(lrti_events_num = as.numeric(lrti_events_di_total),
+         lrti_events_num = lrti_events_num - 1) %>%  
+  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
+            "Group %" = `Total LRTI`/n()*100)
+
 plot_results
@@
 pm_summary_table
+lrti_summary_table
+save_summary
 save_contrasts_full
@@
 
-tidy_results <- tidy(itt_filter_type_results, conf.int = TRUE) %>% 
+tidy_results <- tidy(itt_winter1_results, conf.int = TRUE) %>% 
   mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
@@
 
-table(model.frame(itt_filter_type_results)$treatment_filter_type)
+table(model.frame(itt_winter1_results)$treatment_assigned)
 
@@
 #write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
-#summary(itt_pm_results_int)
+save_summary <<- summary(itt_pm_results_int)
 
@@
 
+lrti_summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
+         lrti_events_num = lrti_events_num - 1) %>%  
+  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
+            "Group %" = `Total LRTI`/n()*100)
+
 plot_results
@@
                           education_3level + 
-                          residents_smoke + 
+                          #residents_smoke + 
                           person_time_at_risk_winter +
@@
 #write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
-#summary(itt_pm_results_int)
+save_summary <<- summary(er_results_int)
 
@@
 
+lrti_summary_table <<- new_model_data %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var) %>% 
+  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
+         lrti_events_num = lrti_events_num - 1) %>%  
+  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
+            "Group %" = `Total LRTI`/n()*100)
+
 plot_results
@@
 
-int_function(analysis_data_winter1, "gender")
+int_function(analysis_data_winter1, "residents_smoke")
 save_contrasts
 pm_summary_table
+lrti_summary_table
+save_summary
 
@@
 pm_summary_table
-```
-
-```{rr, eval=FALSE, include=FALSE}
+
+pm_summary_table <- analysis_data_winter1 %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  group_by(treatment_filter_type) %>% 
+  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+pm_summary_table
+
+lrti_summary_table <- analysis_data_winter1 %>% 
+  filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
+           !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
+  #group_by(treatment_assigned) %>% 
+  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
+         lrti_events_num = lrti_events_num - 1) %>%  
+  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
+            "Group %" = `Total LRTI`/n()*100)
+lrti_summary_table
+
+lrti_summary_table <- analysis_data_winter1 %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
+         lrti_events_num = lrti_events_num - 1) %>% 
+  group_by(treatment_filter_type) %>% 
+  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
+            "Group %" = `Total LRTI`/n()*100)
+lrti_summary_table
+```
+
+```{r, eval=FALSE, include=FALSE}
 # ER framework using a Spline for PM
+
+spline_data <- analysis_data_winter1 %>% 
+  filter(pm_mean_sampling_period < 200) %>% 
+  mutate(pm_cat = cut(pm_mean_sampling_period,
+                      breaks = c(0, 10, 20, 35, 300),
+                      labels = c("0 to 10", "10 to 20",
+                                 "20 to 35", "35+")))
 
@@
 
-model_pm_spline <- gamm4(lrti_events_di_winter ~ s(pm_mean_sampling_period_iqr, 
-                          fx = TRUE, bs = "tp", k = 6) + 
+model_pm_spline <- gamm4(lrti_events_di_winter ~ s(pm_mean_sampling_period, 
+                          fx = TRUE, bs = "tp", k = 5) + 
                           child_age +  education_3level + residents_smoke +
                           person_time_at_risk_winter, 
-                          data = analysis_data_winter1, family = binomial(link = "logit"), 
-                          random = ~(1 | home:cohort:area))
+                          data = spline_data, family = binomial(link = "logit"), 
+                          random = ~(1 | cohort:area))
 plot(model_pm_spline$gam, seWithMean = TRUE, select = 1, shade = TRUE,
-     xlab = expression(paste("IQR increase in area PM"[2.5], 
+     xlab = expression(paste("Area PM"[2.5], 
                              " (", mu, g/m^3, ")")),
@@
 
-# Not sure how to add the offset in
-er_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + 
-                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
-                          data = analysis_data_winter1)
-```
+# Try using PM grouped by quartiles
+
+er_results <- glmer(lrti_events_di_winter ~ pm_cat + child_age + 
+                          education_3level + residents_smoke + person_time_at_risk_winter +
+                          (1 | cohort:area), 
+                          data = spline_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
+summary(er_results)
+```

### 2020-12-24 main_analysis_final.Rmd
filter compliance data updates

@@
 author: "Ethan Walker"
-date: "Updated 16 Dec 2020"
+date: "Updated 23 Dec 2020"
 output: pdf_document
@@
          child_age = factor(child_age, levels = c("< 1", "1+"))) 
+  #filter(area == "WMT")
   #filter(gender == "Female")
@@
   filter(person_time_at_risk_winter > 0) %>% 
+  #filter(area == "WMT") %>% 
   filter(winter_id == 1)
@@
 
-itt_filter_comp_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
-                     child_age + person_time_at_risk + kw_perc_exp_cat2 + 
+itt_filter_comp_results <- glmer(lrti_events_di_total ~ filter_compliance + 
+                     child_age + person_time_at_risk + 
                      (1 | home:cohort:area),
                      data = analysis_data, family = binomial, nAGQ = 10)
+summary(itt_filter_comp_results)
 write_rds(itt_filter_comp_results, paste0(file_path, "Output/itt_filter_comp_results.rds"))
@@
 
-itt_pm_filter_comp_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                  child_age + (1 | cohort:area) + 
-                                  kw_perc_exp_cat2, data = analysis_data_winter1)
+itt_pm_filter_comp_results <- glmer(log(pm_mean_sampling_period) ~ filter_compliance + 
+                                  child_age + (1 | cohort:area), 
+                                  data = analysis_data_winter1)
 write_rds(itt_pm_filter_comp_results, paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))

### 2020-12-24 main_analysis_initial.Rmd
filter compliance data updates

@@
 
-kids_kw_report_1obs <- read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
-  group_by(home) %>% 
-  mutate(mean_kw_per_exp = mean(kw_perc_expected, na.rm = TRUE)) %>% 
-  ungroup() %>% 
-  distinct(area, home, .keep_all = TRUE) %>% 
-  mutate(kw_perc_exp_cat = cut(kw_perc_expected, 
-                               breaks = c(0, 30, 56, 106, 1000),
-                               labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
-  dplyr::select(-home_winter_id, -treatment_assigned, -treatment_actual, -home_id)
+kids_filter_compliance_total <- read_rds(paste0(file_path, "Output/kids_filter_compliance_total.rds")) %>% 
+  dplyr::select(-home_winter_id, -winter_id, -treatment_assigned)
 
-kids_kw_report_2obs <- read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
-  group_by(area, home_winter_id) %>% 
-  mutate(mean_kw_per_exp = mean(kw_perc_expected, na.rm = TRUE)) %>% 
-  ungroup() %>% 
-  distinct(area, home_winter_id, .keep_all = TRUE) %>% 
-  mutate(kw_perc_exp_cat = cut(kw_perc_expected, 
-                               breaks = c(0, 29, 53, 96, 1000),
-                               labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
-  dplyr::select(-home_winter_id, -treatment_assigned, -treatment_actual, -home_id)
+kids_filter_compliance_winter1 <- read_rds(paste0(file_path, "Output/kids_filter_compliance_winter1.rds")) %>% 
+  dplyr::select(-home, -treatment_assigned, -winter_id)
 
@@
   right_join(lrti_1obs_total, by = c("area", "home", "child_id_num")) %>% 
-  left_join(kids_kw_report_1obs, by = c("area", "home")) %>% 
+  left_join(kids_filter_compliance_total, by = c("area", "home")) %>% 
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
@@
          home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum,
-         filter_type, mean_kw_per_exp, kw_perc_exp_cat) %>% 
+         filter_type, filter_compliance) %>% 
   rename(treatment = newtx,
@@
                                                    "Small Filtrete", "Winix")),
-         kw_perc_exp_cat2 = as.character(kw_perc_exp_cat),
-         kw_perc_exp_cat2 = if_else(treatment_assigned == "Filter", kw_perc_exp_cat2, treatment_assigned),
-         kw_perc_exp_cat2 = factor(kw_perc_exp_cat2,
-                                  levels = c("Placebo", "Education", "Q1", "Q2", "Q3", "Q4"))) %>% 
+         filter_compliance2 = as.character(filter_compliance),
+         filter_compliance2 = if_else(treatment_assigned == "Education", treatment_assigned, filter_compliance2),
+         filter_compliance = factor(filter_compliance2,
+                                  levels = c("Placebo", "Education", 
+                                             "0-25 compliance", "25-50 compliance",
+                                             "50-75 compliance", "75-100 compliance"))) %>% 
   mutate(treatment = factor(treatment,
@@
   right_join(lrti_1obs_per_winter, by = c("area", "home", "child_id_num", "winter_id")) %>%  
-  left_join(kids_kw_report_1obs, by = c("area", "home", "winter_id")) %>% 
+  left_join(kids_filter_compliance_winter1, by = c("area", "home_winter_id")) %>% 
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
@@
          home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum,
-         filter_type, mean_kw_per_exp, kw_perc_exp_cat) %>% 
+         filter_type, filter_compliance) %>% 
   rename(treatment = newtx) %>% 

### 2020-12-24 manuscript_tables_figures.Rmd
filter compliance data updates

@@
   filter(winter_id == 1) %>% 
-  filter(treatment_assigned == "Placebo") %>% 
+  filter(treatment_assigned == "Filter") %>% 
   arrange(new_var) %>% 
@@
   filter(!is.na(new_var)) %>% 
-  select(area, home, new_var) %>% 
+  dplyr::select(area, home, new_var) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
   filter(!is.na(new_var)) %>% 
-  select(area, home, new_var) %>% 
+  dplyr::select(area, home, new_var) %>% 
   group_by(area) %>% 

### 2020-12-24 stove_grade_usage_cleaning.Rmd
filter compliance data updates

@@
   left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
+  #filter(winter_id == 1) %>% 
   arrange(area, home_winter_id) %>% 
@@
 
+# WMT filter unit compliance phone calls conducted by Carolyn
+```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+wmt_filter_checks <- read_xlsx(paste0(file_path, "Input/WMT/EdChecks-WithComments.xlsx"), na = "NULL") %>% 
+  rename_all(tolower) %>% 
+  select(homewinterid, date, 
+         filterrunning, 'filter details') %>% 
+  mutate(home_winter_id = as.character(homewinterid),
+         date = ymd(date),
+         filter_level = as.factor(filterrunning),
+         filter_comment = as.character(`filter details`),
+         area = "WMT") %>% 
+  select(home_winter_id, date, 
+         filter_level, filter_comment, area) %>% 
+  arrange(home_winter_id, date)
+
+write_rds(wmt_filter_checks, paste0(file_path, "Output/wmt_filter_checks.rds"))
+```
+
+# Summarize KWhour and PM data; print table
 ```{r}
@@
   distinct(area, home, .keep_all = TRUE) %>% 
-  select(area, home, pm_mean_sampling_period)
+  dplyr::select(area, home, pm_mean_sampling_period, treatment_assigned, home_sqft_2level)
 kids_kw_report_clean <- 
-  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds"))
+  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
+  dplyr::select(-treatment_assigned)
+wmt_filter_checks <- 
+  read_rds(paste0(file_path, "Output/wmt_filter_checks.rds"))
 
@@
   left_join(pm_data, by = c("area", "home")) %>% 
-  filter(!is.na(kw_perc_expected)) %>% 
-  #filter(area == "WMT") %>% 
-  group_by(filter_type, treatment_assigned) %>% 
-  summarize("Mean KW % Exp" = mean(kw_perc_expected),
-            "SD KW % Exp" = sd(kw_perc_expected),
-            "N KW % Exp" = n(),
-            "Min KW % Exp" = min(kw_perc_expected),
-            "Med KW % Exp" = median(kw_perc_expected),
-            "Max KW % Exp" = max(kw_perc_expected),
+  right_join(wmt_filter_checks, by = c("area", "home_winter_id")) %>% 
+  #filter(!is.na(kw_perc_expected)) %>% 
+  #filter(treatment_assigned == "Filter") %>% 
+  #distinct(area, home, .keep_all = TRUE) %>% 
+  #filter(!is.na(filter_type)) %>% 
+  #group_by(area, treatment_assigned, filter_type, home_sqft_2level) %>% 
+  filter(area == "WMT") %>% 
+  group_by(filter_type, treatment_assigned, filter_level) %>% 
+  summarize("N KW % Exp" = n(),
+            "Mean KW % Exp" = mean(kw_perc_expected, na.rm = TRUE),
+            #"SD KW % Exp" = sd(kw_perc_expected, na.rm = TRUE),
+            #"Min KW % Exp" = min(kw_perc_expected, na.rm = TRUE),
+            "Med KW % Exp" = median(kw_perc_expected, na.rm = TRUE),
+            #"Max KW % Exp" = max(kw_perc_expected),
             "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
-            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
-            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
-            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE))
 kw_perc_exp_summary
+
+# wmt_filte_checks data seems to validate the KW % expected as compliance data,
+# at least for WMT. KW % Exp is lower in homes that reported filter as too low or off.
 
@@
 
-# Try to extract sampling dates from KW hour data
+# Extract sampling dates and use KWhour data from PM sampling period
 ```{r}
@@
 
+# Use KWhour data only if KWhours increased in subsequent dates
 ```{r}
@@
   read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
-  select(home_winter_id, area, filter_type)
+  dplyr::select(home_winter_id, area, filter_type)
 
@@
   rename_all(tolower) %>% 
-  select(homewinterid, interventiondate, 
+  dplyr::select(homewinterid, interventiondate, 
          samplingdate, pickupdate) %>% 
@@
          area = "WMT") %>% 
-  select(home_winter_id, intervention_date, 
+  dplyr::select(home_winter_id, intervention_date, 
          sampling_date, pickup_date, area)
@@
          area = "NN") %>% 
-  select(home_winter_id, intervention_date, 
+  dplyr::select(home_winter_id, intervention_date, 
          sampling_date, pickup_date, area)
@@
          area = "AK") %>% 
-  select(home_winter_id, intervention_date, 
+  dplyr::select(home_winter_id, intervention_date, 
          sampling_date, pickup_date, area)
@@
 
-kids_kw_filtered <- kids_kw_clean %>% 
+kids_filter_compliance <- kids_kw_clean %>% 
   group_by(area, home_winter_id) %>% 
@@
   mutate(kw_expected = date_diff*multiplier,
-         kw_perc_exp = kw_hours/kw_expected*100)
+         kw_perc_exp = kw_hours/kw_expected*100) %>% 
+  dplyr::select(area, home_winter_id, winter_id, home, treatment_assigned, 
+                kw_date, kw_hours, date_diff, filter_type, multiplier,
+                kw_expected, kw_perc_exp)
+
+compliance_summary1 <- kids_filter_compliance %>% 
+  group_by(home) %>% 
+  mutate(kw_perc_exp_total = mean(kw_perc_exp, na.rm = TRUE)) %>%  
+  filter(treatment_assigned == "Filter") %>% 
+  filter(!is.na(kw_perc_exp_total)) %>% 
+  group_by(treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_exp_total),
+            "SD KW % Exp" = sd(kw_perc_exp_total),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_exp_total),
+            "25Q KW % Exp" = quantile(kw_perc_exp_total, 0.25),
+            "Med KW % Exp" = median(kw_perc_exp_total),
+            "75Q KW % Exp" = quantile(kw_perc_exp_total, 0.75),
+            "Max KW % Exp" = max(kw_perc_exp_total))
+compliance_summary1
+
+kids_filter_compliance_total <- kids_filter_compliance %>% 
+  group_by(home) %>% 
+  mutate(kw_perc_exp_total = mean(kw_perc_exp), 
+         treatment_assigned = as.character(treatment_assigned),
+         perc_exp_quart = cut(kw_perc_exp_total,
+                                  breaks = c(0, 30, 64, 100, 1000),
+                                  labels = c("0-25 compliance", "25-50 compliance",
+                                             "50-75 compliance", "75-100 compliance")),
+         perc_exp_quart = as.character(perc_exp_quart),
+         filter_compliance = if_else(treatment_assigned == "Filter",
+                                     perc_exp_quart, 
+                                     treatment_assigned)) %>% 
+  distinct(area, home, .keep_all = TRUE)
+
+compliance_summary2 <- kids_filter_compliance %>% 
+  group_by(area, home_winter_id) %>% 
+  mutate(kw_perc_exp_winter = mean(kw_perc_exp, na.rm = TRUE)) %>%  
+  filter(treatment_assigned == "Filter") %>% 
+  filter(!is.na(kw_perc_exp_winter)) %>% 
+  group_by(winter_id, treatment_assigned) %>% 
+  summarize("Mean KW % Exp" = mean(kw_perc_exp_winter),
+            "SD KW % Exp" = sd(kw_perc_exp_winter),
+            "N KW % Exp" = n(),
+            "Min KW % Exp" = min(kw_perc_exp_winter),
+            "25Q KW % Exp" = quantile(kw_perc_exp_winter, 0.25),
+            "Med KW % Exp" = median(kw_perc_exp_winter),
+            "75Q KW % Exp" = quantile(kw_perc_exp_winter, 0.75),
+            "Max KW % Exp" = max(kw_perc_exp_winter))
+compliance_summary2
+
+kids_filter_compliance_winter1 <- kids_filter_compliance %>% 
+  filter(winter_id == 1) %>% 
+  mutate(treatment_assigned = as.character(treatment_assigned),
+         perc_exp_quart = cut(kw_perc_exp,
+                                  breaks = c(0, 32, 60, 116, 1000),
+                                  labels = c("0-25 compliance", "25-50 compliance",
+                                             "50-75 compliance", "75-100 compliance")),
+         perc_exp_quart = as.character(perc_exp_quart),
+         filter_compliance = if_else(treatment_assigned == "Filter",
+                                     perc_exp_quart, 
+                                     treatment_assigned))
+
+write_rds(kids_filter_compliance_total, paste0(file_path, "Output/kids_filter_compliance_total.rds"))
+write_rds(kids_filter_compliance_winter1, paste0(file_path, "Output/kids_filter_compliance_winter1.rds"))
 
@@
 
+# Changing the placebo multipliers above to the same as the filter multiplier
+# changes the picture quite a bit. It seems likely that participants ran the 
+# placebo filters on high more often than expected, resulting in high % Exp KWhours
+
 png("Kids_KW_perc_exp_new.png", width=1200, height=480, bg = "white")

### 2021-01-21 intervention_questionnaire_cleaning.Rmd
New RMDs for intervention questionnaires

---
title: "KidsAIR intervention questionnaires cleaning"
author: "Ethan Walker"
date: "Started 20 Jan 2021, Updated 20 Jan 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) 
```

# Load, format, and save Post Winter Questionnaire - filter homes
```{r}
##### WMT #####
# Load data
pwq_wmt <- read_csv(paste0(file_path, "Input/WMT/eow.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:helpful) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### NN #####
# Load data
pwq_nn <- read_csv(paste0(file_path, "Input/NN/eow.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:helpful) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### AK #####
# Load data
pwq_ak <- read_csv(paste0(file_path, "Input/AK/eow.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:helpful) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

kids_pwq_filter <- rbind(pwq_wmt, pwq_nn, pwq_ak)

write_rds(kids_pwq_filter, paste0(file_path, "Output/kids_pwq_filter.rds"))
```

# Clean up all Post Winter Questionnaire - filter homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_pwq_filter <- read_rds(paste0(file_path, "Output/kids_pwq_filter.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

kids_pwq_filter_clean <- kids_pwq_filter %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate(enteredby = as.factor(enteredby),
         fuse = as.factor(fuse),
         continuously = as.factor(continuously),
         reasonnoise = as.factor(reasonnoise),
         reasoncost = as.factor(reasoncost),
         reasonappearance = as.factor(reasonappearance),
         reasonspace = as.factor(reasonspace),
         reasoninconvenience = as.factor(reasoninconvenience),
         mostimp = as.factor(mostimp),
         helpful = as.factor(helpful)) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_pwq_filter_clean)

write_rds(kids_pwq_filter_clean, paste0(file_path, "Output/kids_pwq_filter_clean.rds"))
```

###############################################

# Load, format, and save Post Winter Questionnaire - education homes
```{r}
##### WMT #####
# Load data
pwq_wmt <- read_csv(paste0(file_path, "Input/WMT/eowe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### NN #####
# Load data
pwq_nn <- read_csv(paste0(file_path, "Input/NN/eowe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### AK #####
# Load data
pwq_ak <- read_csv(paste0(file_path, "Input/AK/eowe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

kids_pwq_education <- rbind(pwq_wmt, pwq_nn, pwq_ak)

write_rds(kids_pwq_education, paste0(file_path, "Output/kids_pwq_education.rds"))
```

# Clean up all Post Winter Questionnaire - education homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_pwq_education <- read_rds(paste0(file_path, "Output/kids_pwq_education.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

kids_pwq_education_clean <- kids_pwq_education %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate_if(is.character, as.factor) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_pwq_education_clean)

write_rds(kids_pwq_education_clean, paste0(file_path, "Output/kids_pwq_education_clean.rds"))
```

###############################################

# Load, format, and save KAB Questionnaire - filter homes
```{r}
##### WMT #####
# Load data
kab_wmt <- read_csv(paste0(file_path, "Input/WMT/kab.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### NN #####
# Load data
kab_nn <- read_csv(paste0(file_path, "Input/NN/kab.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### AK #####
# Load data
kab_ak <- read_csv(paste0(file_path, "Input/AK/kab.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

kids_kab_filter <- rbind(kab_wmt, kab_nn, kab_ak)

write_rds(kids_kab_filter, paste0(file_path, "Output/kids_kab_filter.rds"))
```

# Clean up all KAB Questionnaire - filter homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kab_filter <- read_rds(paste0(file_path, "Output/kids_kab_filter.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

kids_kab_filter_clean <- kids_kab_filter %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate_all(as.factor) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_kab_filter_clean)

write_rds(kids_kab_filter_clean, paste0(file_path, "Output/kids_kab_filter_clean.rds"))
```

###############################################

# Load, format, and save KAB Questionnaire - education homes
```{r}
##### WMT #####
# Load data
kab_wmt <- read_csv(paste0(file_path, "Input/WMT/kabe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### NN #####
# Load data
kab_nn <- read_csv(paste0(file_path, "Input/NN/kabe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

##### AK #####
# Load data
kab_ak <- read_csv(paste0(file_path, "Input/AK/kabe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)

kids_kab_education <- rbind(kab_wmt, kab_nn, kab_ak)

write_rds(kids_kab_education, paste0(file_path, "Output/kids_kab_education.rds"))
```

# Clean up all KAB Questionnaire - education homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kab_education <- read_rds(paste0(file_path, "Output/kids_kab_education.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))

kids_kab_education_clean <- kids_kab_education %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate_all(as.factor) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_kab_education_clean)

write_rds(kids_kab_education_clean, paste0(file_path, "Output/kids_kab_education_clean.rds"))
```

### 2021-01-21 intervention_questionnaire_summary.Rmd
New RMDs for intervention questionnaires

---
title: "KidsAIR intervention questionnaires summary"
author: "Ethan Walker"
date: "Started 20 Jan 2021, Updated 20 Jan 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_pwq_filter_clean <- read_rds(paste0(file_path, "Output/kids_pwq_filter_clean.rds")) 
kids_pwq_education_clean <- read_rds(paste0(file_path, "Output/kids_pwq_education_clean.rds")) 
kids_kab_filter_clean <- read_rds(paste0(file_path, "Output/kids_kab_filter_clean.rds")) 
kids_kab_education_clean <- read_rds(paste0(file_path, "Output/kids_kab_education_clean.rds")) 
```

# Summarize questionnaires for K99/R00 proposal
```{r}
kids_pwq_filter_summary <- kids_pwq_filter_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_pwq_filter_summary)

kids_pwq_education_summary <- kids_pwq_education_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_pwq_education_summary)

kids_kab_filter_summary <- kids_kab_filter_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_kab_filter_summary)

kids_kab_education_summary <- kids_kab_education_clean %>% 
  filter(area == "WMT") %>% 
  filter(winter_id == 2) 
summary(kids_kab_education_summary)
```

### 2021-01-21 manuscript_tables_figures.Rmd
New RMDs for intervention questionnaires

@@
 
-demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))
+demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
+  filter(area == "WMT") %>% 
+  filter(winter_id == 1) %>% 
+  distinct(home_winter_id, .keep_all = TRUE)
 
@@
   filter(winter_id == 1) %>% 
-  filter(treatment_assigned == "Placebo") %>% 
+  #filter(treatment_assigned == "Placebo") %>% 
   arrange(new_var) %>% 

### 2021-01-21 pm_summary.Rmd
New RMDs for intervention questionnaires

@@
 boxplots_kids_pm <- health_exposure_data_sampling_day %>% 
-  filter(winter_id == 1 & treatment_assigned == "Placebo") %>% 
+  filter(winter_id == 1) %>% 
+  #filter(treatment_assigned == "Placebo") %>% 
   arrange(area, home, sampling_day) %>% 
@@
   ungroup() %>% 
-  ggplot(aes(area, pm_mean_sampling_period)) +
+  ggplot(aes(treatment_assigned, pm_mean_sampling_period)) +
     geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
@@
 
-ggsave("boxplots_kids_pm.jpg", height = 7, width = 4)
+#ggsave("boxplots_kids_pm.jpg", height = 7, width = 4)
 ```

### 2021-02-26 main_analysis_final.Rmd
Various analysis updates to help with primary manuscript

@@
 author: "Ethan Walker"
-date: "Updated 23 Dec 2020"
+date: "Updated 26 Feb 2021"
 output: pdf_document
@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         pm_mean_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         pm_mean_at_home_sampling_period = pm_at_home_sampling_period,
          lrti_events_di_total = as.factor(lrti_events_di_total),
          person_time_at_risk_iqr = person_time_at_risk/20,
-         child_age = factor(child_age, levels = c("< 1", "1+"))) 
+         child_age = factor(child_age, levels = c("< 1", "1+")),
+         moisture_di = cut(moisture_closest, breaks = c(0, 11, 100),
+                           labels = c("<11%", ">=11%")),
+         stove_age_di = fct_collapse(stove_age_3level, 
+                                     "0-5 years old" = "0-5 years old",
+                                     "6+ years old" = c("6-15 years old", "16 + years old"),
+                                     "Do not know" = "Do not know")) 
   #filter(area == "WMT")
@@
          lrti_events_di_total = as.factor(lrti_events_di_total),
-         child_age = factor(child_age, levels = c("< 1", "1+")))
+         child_age = factor(child_age, levels = c("< 1", "1+")),
+         moisture_di = cut(moisture_closest, breaks = c(0, 11, 100),
+                           labels = c("<11%", ">=11%")))
 
@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         pm_mean_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         pm_mean_at_home_sampling_period = pm_at_home_sampling_period,
          lrti_events_di_winter = as.factor(lrti_events_di_winter),
-         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
+         child_age = factor(child_age, levels = c("< 1", "1+")),
+         moisture_di = cut(moisture_closest, breaks = c(0, 11, 100),
+                           labels = c("<11%", ">=11%")),
+         stove_age_di = fct_collapse(stove_age_3level, 
+                                     "0-5 years old" = "0-5 years old",
+                                     "6+ years old" = c("6-15 years old", "16 + years old"),
+                                     "Do not know" = "Do not know")) %>% 
   filter(person_time_at_risk_winter > 0) %>% 
@@
   mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
-         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         pm_mean_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         pm_mean_at_home_sampling_period = pm_at_home_sampling_period,
          lrti_events_di_total = as.factor(lrti_events_di_total),
          person_time_at_risk_iqr = person_time_at_risk/20,
-         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
+         child_age = factor(child_age, levels = c("< 1", "1+")),
+         moisture_di = cut(moisture_closest, breaks = c(0, 11, 100),
+                           labels = c("<11%", ">=11%")),
+         stove_age_di = fct_collapse(stove_age_3level, 
+                                     "0-5 years old" = "0-5 years old",
+                                     "6+ years old" = c("6-15 years old", "16 + years old"),
+                                     "Do not know" = "Do not know")) %>% 
   filter(area != "AK")
@@
 
-int_function(analysis_data, "residents_smoke")
+int_function(analysis_data, "home_sqft_2level")
 n_table
@@
 
-int_function(analysis_data_winter1, "burn_level_3level")
+int_function(analysis_data_winter1, "stove_grade_3level")
 n_table
@@
 # Try: home_floors_2level, home_sqft_2level, education_3level, child_age
-# home_year_built_2level, stove_age_3level, stove_grade_3level, chimney_clean_3level,
+# home_year_built_2level, stove_age_3level, stove_age_di, stove_grade_3level, chimney_clean_3level,
 # wood_collect_2level, wood_collect_method_2level, burn_level_3level, residents_smoke
@@
 write_rds(itt_pm_noak_results, paste0(file_path, "Output/itt_pm_noak_results.rds"))
+
+itt_pm_at_home_results <- glmer(log(pm_mean_at_home_sampling_period) ~ treatment_assigned + 
+                               child_age + (1 | cohort:area), 
+                               data = analysis_data_winter1)
+write_rds(itt_pm_at_home_results, paste0(file_path, "Output/itt_pm_at_home_results.rds"))
 
@@
                                   data = analysis_data_winter1)
+summary(itt_pm_filter_comp_results)
 write_rds(itt_pm_filter_comp_results, paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
@@
 itt_pm_results <- read_rds(paste0(file_path, "Output/itt_pm_results.rds"))
+itt_pm_at_home_results <- read_rds(paste0(file_path, "Output/itt_pm_at_home_results.rds"))
 itt_pm_noak_results <- read_rds(paste0(file_path, "Output/itt_pm_noak_results.rds"))
 itt_pm_income_results <- read_rds(paste0(file_path, "Output/itt_pm_income_results.rds"))
-#itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
-#itt_pm_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
+itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
+itt_pm_filter_comp_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_comp_results.rds"))
 itt_pm_smoke_results <- read_rds(paste0(file_path, "Output/itt_pm_smoke_results.rds"))
@@
 itt_pm_tidy <- results_function(itt_pm_results, "ITT Primary")
+itt_pm_at_home_tidy <- results_function(itt_pm_at_home_results, "ITT PM at-home")
 itt_pm_noak_tidy <- results_function(itt_pm_noak_results, "ITT no AK")
@@
 # Combine cleaned results
-itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_noak_tidy, itt_pm_income_tidy, 
+itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_at_home_tidy, itt_pm_noak_tidy, itt_pm_income_tidy, 
                               itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT no AK", "ITT + income", 
+                        levels = c("ITT Primary", "ITT PM at-home", "ITT no AK", "ITT + income", 
                                    "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
@@
 ITT Primary = Linear mixed model, as specified above  
+ITT PM at-home = Primary model using PM when child was at home  
 ITT no AK = Primary model without homes from AK (different placebo group)  
@@
 
-er_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          education_3level + residents_smoke + person_time_at_risk_winter +
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10,
+er_results <- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr + 
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
+                          data = analysis_data, family = binomial, nAGQ = 10,
                           control = glmerControl(optimizer = "bobyqa"))
@@
 # Run primary model and print results
-er_results_int <<- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*int_var + 
-                          child_age + 
-                          education_3level + 
-                          #residents_smoke + 
-                          person_time_at_risk_winter +
-                          (1 | cohort:area), 
+er_results_int <<- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr*int_var + 
+                          #child_age + 
+                          gender + 
+                          residents_smoke + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + 
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
                           data = new_model_data, family = binomial, nAGQ = 10,
@@
   group_by(int_var) %>% 
-  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
+  mutate(lrti_events_num = as.numeric(lrti_events_di_total),
          lrti_events_num = lrti_events_num - 1) %>%  
@@
 
-int_function(analysis_data_winter1, "residents_smoke")
+int_function(analysis_data, "child_age")
 save_contrasts
@@
 
-# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
+# Try child_age, gender, residents_smoke
 ```
@@
 
-er_income_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          residents_smoke + person_time_at_risk_winter + 
-                          income_3level + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_income_results, paste0(file_path, "Output/er_income_results.rds"))
-
-er_filter_comp_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter + 
-                          kw_perc_exp_cat2 + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_filter_comp_results, paste0(file_path, "Output/er_filter_comp_results.rds"))
-
-er_burn_level_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter + 
+er_results_at_home <- er_results <- glmer(lrti_events_di_total ~ pm_mean_at_home_sampling_period_iqr + 
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
+                          data = analysis_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
+write_rds(er_results_at_home, paste0(file_path, "Output/er_results_at_home.rds"))
+
+er_winter1_results <- er_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + 
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk_winter +
+                          (1 | cohort:area), 
+                          data = analysis_data_winter1, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
+write_rds(er_winter1_results, paste0(file_path, "Output/er_winter1_results.rds"))
+
+er_burn_level_results <- er_results <- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr + 
                           burn_level_3level + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
+                          data = analysis_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
 write_rds(er_burn_level_results, paste0(file_path, "Output/er_burn_level_results.rds"))
 
-er_smoking_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + person_time_at_risk_winter + 
-                          home_act_smoking_sum + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_smoking_results, paste0(file_path, "Output/er_smoking_results.rds"))
-
-er_windows_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter + 
-                          home_act_windows_sum + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_windows_results, paste0(file_path, "Output/er_windows_results.rds"))
-
-er_door_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter + 
-                          home_act_door_sum + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_door_results, paste0(file_path, "Output/er_door_results.rds"))
-
-er_sweep_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter +
-                          home_act_sweep_sum + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_sweep_results, paste0(file_path, "Output/er_sweep_results.rds"))
-
-er_ambpm_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter +
-                          amb_pm_lrti + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_ambpm_results, paste0(file_path, "Output/er_ambpm_results.rds"))
-
-er_ambtemp_results <- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr + child_age + 
-                          income_3level + residents_smoke + person_time_at_risk_winter +
-                          amb_temp_lrti + 
-                          (1 | home:cohort:area), 
-                          data = analysis_data_winter1, family = binomial, nAGQ = 10)
-write_rds(er_ambtemp_results, paste0(file_path, "Output/er_ambtemp_results.rds"))
+er_home_size_results <- er_results <- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr + 
+                          home_sqft_2level + 
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
+                          data = analysis_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
+write_rds(er_home_size_results, paste0(file_path, "Output/er_home_size_results.rds"))
+
+er_stove_grade_results <- er_results <- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr + 
+                          stove_grade_3level + 
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
+                          data = analysis_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
+write_rds(er_stove_grade_results, paste0(file_path, "Output/er_stove_grade_results.rds"))
+
+er_stove_age_results <- er_results <- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr + 
+                          stove_age_3level + 
+                          child_age + gender + 
+                          race_parent + gender_parent +
+                          income_3level + education_3level + residents_smoke +
+                          amb_temp_lrti + amb_pm_lrti + 
+                          person_time_at_risk +
+                          (1 | home:cohort:area), 
+                          data = analysis_data, family = binomial, nAGQ = 10,
+                          control = glmerControl(optimizer = "bobyqa"))
+write_rds(er_stove_age_results, paste0(file_path, "Output/er_stove_age_results.rds"))
 ```
@@
 er_results <- read_rds(paste0(file_path, "Output/er_results.rds"))
-er_income_results <- read_rds(paste0(file_path, "Output/er_income_results.rds"))
-#er_filter_comp_results <- read_rds(paste0(file_path, "Output/er_filter_comp_results.rds"))
+er_results_at_home <- read_rds(paste0(file_path, "Output/er_results_at_home.rds"))
+er_winter1_results <- read_rds(paste0(file_path, "Output/er_winter1_results.rds"))
 er_burn_level_results <- read_rds(paste0(file_path, "Output/er_burn_level_results.rds"))
-er_smoking_results <- read_rds(paste0(file_path, "Output/er_smoking_results.rds"))
-er_windows_results <- read_rds(paste0(file_path, "Output/er_windows_results.rds"))
-er_door_results <- read_rds(paste0(file_path, "Output/er_door_results.rds"))
-er_sweep_results <- read_rds(paste0(file_path, "Output/er_sweep_results.rds"))
-er_ambpm_results <- read_rds(paste0(file_path, "Output/er_ambpm_results.rds"))
-er_ambtemp_results <- read_rds(paste0(file_path, "Output/er_ambtemp_results.rds"))
+er_home_size_results <- read_rds(paste0(file_path, "Output/er_home_size_results.rds"))
+er_stove_grade_results <- read_rds(paste0(file_path, "Output/er_stove_grade_results.rds"))
+er_stove_age_results <- read_rds(paste0(file_path, "Output/er_stove_age_results.rds"))
 
@@
 er_tidy <- results_function(er_results, "ER Primary")
-er_income_tidy <- results_function(er_income_results, "ER + income")
-#er_filter_comp_tidy <- results_function(er_filter_comp_results, "ER + filt comp")
+er_at_home_tidy <- results_function(er_results_at_home, "ER PM at-home")
+er_winter1_tidy <- results_function(er_winter1_results, "ER Winter 1 only")
 er_burn_level_tidy <- results_function(er_burn_level_results, "ER + burn level")
-er_smoking_tidy <- results_function(er_smoking_results, "ER + smoke act")
-er_windows_tidy <- results_function(er_windows_results, "ER + window act")
-er_door_tidy <- results_function(er_door_results, "ER + door act")
-er_sweep_tidy <- results_function(er_sweep_results, "ER + sweep act")
-er_ambpm_tidy <- results_function(er_ambpm_results, "ER + amb PM")
-er_ambtemp_tidy <- results_function(er_ambtemp_results, "ER + amb temp")
+er_home_size_tidy <- results_function(er_home_size_results, "ER + home size")
+er_stove_grade_tidy <- results_function(er_stove_grade_results, "ER + stove grade")
+er_stove_age_tidy <- results_function(er_stove_age_results, "ER + stove age")
 
 # Combine cleaned results
-er_results_combined <- rbind(er_tidy, er_income_tidy, er_burn_level_tidy, 
-                              er_smoking_tidy, er_windows_tidy, er_door_tidy,
-                              er_sweep_tidy, er_ambpm_tidy, er_ambtemp_tidy)
+er_results_combined <- rbind(er_tidy, er_at_home_tidy, er_winter1_tidy, er_burn_level_tidy, 
+                              er_home_size_tidy, er_stove_grade_tidy, er_stove_age_tidy)
 
@@
   mutate(model = factor(model,
-                        levels = c("ER Primary", "ER + income", "ER + burn level",
-                                   "ER + smoke act", "ER + window act", "ER + door act",
-                                   "ER + sweep act", "ER + amb PM", "ER + amb temp"))) %>% 
+                        levels = c("ER Primary", "ER PM at-home", "ER Winter 1 only", 
+                                   "ER + burn level", "ER + home size", 
+                                   "ER + stove grade", "ER + stove age"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
 ER Primary = Logistic regression mixed model, as specified above  
-ER + income = Primary model using income instead of education  
+ER PM at-home = Primary model using PM from when the child was at home  
+ER Winter 1 only = Primary model using data only from Winter 1  
 ER + burn level = Primary model plus self-reported burn-level compared to typical  
-ER + smoke act = Primary model plus # days during sampling with smoking  
-ER + window act = Primary model plus # days during sampling with windows open  
-ER + door act = Primary model plus # days during sampling with doors open  
-ER + sweep act = Primary model plus # days during sampling with sweeping/cleaning  
-ER + amb PM = Primary model plus amb PM during LRTI at-risk period  
-ER + amb temp = Primary model plus amb temp during LRTI at-risk period
+ER + home size = Primary model plus home size (sqaure meters)  
+ER + stove grade = Primary model plus assigned stove grade  
+ER + stove age = Primary model plus caregiver reported stove age
 
@@
 
+pm_summary_table <- analysis_data_winter1 %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  group_by(filter_compliance) %>% 
+  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "n" = n(),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+pm_summary_table
+
 lrti_summary_table <- analysis_data_winter1 %>% 

### 2021-02-26 main_analysis_initial.Rmd
Various analysis updates to help with primary manuscript

@@
                                   "< 1" = "Less than 1",
-                                  "1+" = c(1, 2, 3, 4))) %>% 
+                                  "1+" = c(1, 2, 3, 4)),
+         home_sqft_2level = cut(home_sqft, breaks = c(0, 1400, 100000),
+                                labels = c("<130 sq m", "130+ sq m"))) %>% 
   left_join(treatments_masked, by = "home") %>% 
@@
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, child_age, lrti_events_total, 
+  dplyr::select(area, home, cohort, child_id_num, child_age, race_child, hispanic_child, lrti_events_total, 
          person_time_at_risk_total, lrti_incidence_total, newtx, treatment_assigned,
@@
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, child_age, 
+  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, child_age, race_child, hispanic_child,
          amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti, gender,

### 2021-02-26 manuscript_tables_figures.Rmd
Various analysis updates to help with primary manuscript

@@
 health_exposure_data_sampling_day <- 
-  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
+  mutate(home_sqft_2level = cut(home_sqft, breaks = c(0, 1400, 100000),
+                                labels = c("<130 sq m", "130+ sq m")))
 
@@
 
+# Demographics summary
+```{r}
+table(demographics$hispanic_parent, demographics$race_parent, demographics$gender_parent)
+```
+
 # Count number of households and kids that participated by area and winter
@@
   filter(winter_id == 1) %>% 
-  filter(treatment_assigned == "Placebo") %>%
+  #filter(treatment_assigned == "Placebo") %>%
   arrange(new_var) %>% 
@@
   distinct(new_var, .keep_all = TRUE) %>% 
-  select(new_var, n, percent)
+  dplyr::select(new_var, n, percent)
 demographics_summary
@@
   mutate(total_n = n()) %>% 
-  group_by(area) %>% 
+  group_by(treatment_assigned) %>% 
   mutate(area_n = n()) %>% 
-  group_by(area, new_var) %>% 
+  group_by(treatment_assigned, new_var) %>% 
   mutate(n = n(),
@@
   distinct(new_var, .keep_all = TRUE) %>% 
-  select(new_var, n, percent) %>% 
-  arrange(area, new_var)
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(treatment_assigned, new_var)
 demographics_summary
@@
 char_funct_area("gender_parent")
+char_funct_total("gender")
+char_funct_area("gender")
 char_funct_total("hispanic_parent")
@@
 char_funct_area("home_floors_2level")
+char_funct_total("home_sqft_2level")
+char_funct_area("home_sqft_2level")
 char_funct_total("home_year_built_2level")
@@
 other_vars_area
+
+other_vars_total <- health_exposure_data_sampling_day %>% 
+  dplyr::select(area, child_id_num, race_child, treatment_assigned) %>% 
+  arrange(race_child) %>% 
+  distinct(area, child_id_num, .keep_all = TRUE) 
 ```
@@
 ```
+
+# Table S1 summary data
+```{r}
+analysis_data <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_mean_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         lrti_events_di_total = as.factor(lrti_events_di_total),
+         person_time_at_risk_iqr = person_time_at_risk/20,
+         child_age = factor(child_age, levels = c("< 1", "1+")),
+         moisture_di = cut(moisture_closest, breaks = c(0, 11, 100),
+                           labels = c("<11%", ">=11%")),
+         stove_age_di = fct_collapse(stove_age_3level, 
+                                     "0-5 years old" = "0-5 years old",
+                                     "6+ years old" = c("6-15 years old", "16 + years old"),
+                                     "Do not know" = "Do not know")) 
+  #filter(area == "WMT")
+  #filter(gender == "Female")
+  #filter(home_sqft_2level == "1500+")
+
+analysis_data_winter1 <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
+  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
+         pm_mean_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
+         lrti_events_di_winter = as.factor(lrti_events_di_winter),
+         child_age = factor(child_age, levels = c("< 1", "1+")),
+         moisture_di = cut(moisture_closest, breaks = c(0, 11, 100),
+                           labels = c("<11%", ">=11%")),
+         stove_age_di = fct_collapse(stove_age_3level, 
+                                     "0-5 years old" = "0-5 years old",
+                                     "6+ years old" = c("6-15 years old", "16 + years old"),
+                                     "Do not know" = "Do not know")) %>% 
+  filter(person_time_at_risk_winter > 0) %>% 
+  #filter(area == "WMT") %>% 
+  filter(winter_id == 1)
+
+n_summary_table <- analysis_data_winter1 %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  mutate(int_var = stove_grade_3level) %>% 
+  filter(!is.na(int_var)) %>% 
+  dplyr::select(area, home, int_var, treatment_assigned) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  summarize("n" = n())
+n_summary_table
+
+pm_summary_table <- analysis_data_winter1 %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  mutate(int_var = stove_grade_3level) %>% 
+  filter(!is.na(int_var)) %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  summarize("n" = n(),
+            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
+            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
+            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
+            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
+            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
+pm_summary_table
+
+n_summary_table <- analysis_data %>% 
+  distinct(area, child_id_num, .keep_all = T) %>% 
+  mutate(int_var = stove_grade_3level) %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  summarize("n" = n())
+n_summary_table
+
+lrti_summary_table <- analysis_data %>% 
+  distinct(area, child_id_num, .keep_all = T) %>% 
+  mutate(int_var = stove_grade_3level) %>% 
+  filter(!is.na(int_var)) %>% 
+  group_by(int_var, treatment_assigned) %>% 
+  mutate(lrti_events_num = as.numeric(lrti_events_di_total),
+         lrti_events_num = lrti_events_num - 1) %>%  
+  summarize("n" = n(),
+            "Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
+            "Group %" = `Total LRTI`/n()*100)
+lrti_summary_table
+```

### 2021-02-26 pm_summary.Rmd
Various analysis updates to help with primary manuscript

@@
     theme_minimal() +
-    labs(title = "Red triangle = mean",
+    #labs(title = "Red triangle = mean",
          y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +

### 2021-03-24 lrti_tracking.Rmd
small analysis updates for paper tables

@@
   read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
-  select(area, home_winter_id, treatment_assigned) %>% 
+  dplyr::select(area, home_winter_id, treatment_assigned) %>% 
   distinct(area, home_winter_id, .keep_all = TRUE)
@@
   filter(winter_id != 0) %>% 
-  group_by(area, winter_id, treatment_assigned) %>% 
+  group_by(treatment_assigned) %>% 
   filter(!is.na(person_time_possible)) %>% 
@@
          child_weeks_obs = sum(child_weeks_obs)/7) %>% 
-  distinct(area, winter_id, treatment_assigned, .keep_all = TRUE) %>% 
-  select(area, winter_id, treatment_assigned, child_weeks_obs)
+  distinct(treatment_assigned, .keep_all = TRUE) %>% 
+  dplyr::select(treatment_assigned, child_weeks_obs)
 

### 2021-03-24 main_analysis_final.Rmd
small analysis updates for paper tables

@@
 author: "Ethan Walker"
-date: "Updated 26 Feb 2021"
+date: "Updated 24 March 2021"
 output: pdf_document
@@
 
+obs <- analysis_data %>% 
+  filter(!is.na(lrti_events_di_total)) %>% 
+  filter(!is.na(child_age)) %>%
+  filter(!is.na(area)) %>% 
+  filter(!is.na(cohort)) %>% 
+  filter(!is.na(treatment_assigned)) %>% 
+  filter(!is.na(residents_smoke)) %>%
+  group_by(residents_smoke, treatment_assigned) %>% 
+  count()
+obs
+
 # Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
@@
                               itt_windows_tidy, itt_door_tidy, itt_sweep_tidy,
-                              itt_ambpm_tidy, itt_ambtemp_tidy,
-                              itt_outlier_tidy)
+                              itt_ambpm_tidy, itt_ambtemp_tidy)
 
@@
                                    "ITT + window act", "ITT + door act", "ITT + sweep act",
-                                   "ITT + amb PM", "ITT + amb temp",
-                                   "ITT outlier"))) %>% 
+                                   "ITT + amb PM", "ITT + amb temp"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
   ggtitle(label = "Results for ITT framework, LRTI outcome") +
-  labs(y = "Odds of LRTI compared to placebo") +
+  labs(y = "Odds of LRTI compared to Control") +
   labs(x = "", group = "", shape = "") +
@@
   new_model_data <- model_data %>% 
+    #distinct(area, child_id_num, .keep_all = T) %>% 
     rename(int_var = interaction_var)
@@
 
+obs <- analysis_data_winter1 %>% 
+  distinct(area, child_id_num, .keep_all = T) %>% 
+  filter(!is.na(pm_mean_sampling_period)) %>% 
+  filter(!is.na(child_age)) %>%
+  filter(!is.na(area)) %>% 
+  filter(!is.na(cohort)) %>% 
+  filter(!is.na(treatment_assigned)) %>% 
+  filter(!is.na(home_sqft_2level)) %>%
+  group_by(home_sqft_2level, treatment_assigned) %>% 
+  count()
+obs
+
 # Try: home_floors_2level, home_sqft_2level, education_3level, child_age
@@
 
-#table(model.frame(itt_results_int)$treatment_assigned)
 #((exp(-.418)-1)*100)
@@
   theme_bw() +  
-  ggtitle(label = "Results for ITT framework, PM outcome") +
-  labs(y = "% difference in PM2.5 \n compared to placebo") +
+  ggtitle(label = "Results for ITT framework, PM2.5 outcome") +
+  labs(y = "% difference in PM2.5 \n compared to Control") +
   labs(x = "", group = "", shape = "") +
@@
                           race_parent + gender_parent +
-                          income_3level + education_3level + residents_smoke +
-                          amb_temp_lrti + amb_pm_lrti + 
+                          income_3level + education_3level + 
+                          amb_temp_lrti + amb_pm_lrti + residents_smoke +
                           person_time_at_risk +
@@
 #nobs(er_results)
+#table(model.frame(er_results)$residents_smoke)
+
 #isSingular(er_results)

### 2021-06-09 intervention_questionnaire_summary.Rmd
PM analysis updates. Manuscript revisions from co-author comments.

@@
 kids_pwq_filter_summary <- kids_pwq_filter_clean %>% 
-  filter(area == "WMT") %>% 
-  filter(winter_id == 2) 
-summary(kids_pwq_filter_summary)
+  filter(treatment_assigned == "Filter") %>% 
+  select(area, winter_id, fuse, helpful, continuously) %>% 
+  group_by(winter_id) %>% 
+  mutate(winter_n = n()) %>% 
+  group_by(winter_id, winter_n, helpful) %>% 
+  count() %>% 
+  mutate(var_perc = round(n/winter_n*100, digits = 1)) %>% 
+  ungroup() %>% 
+  select(winter_id, helpful, n, var_perc)
+kids_pwq_filter_summary
 
 kids_pwq_education_summary <- kids_pwq_education_clean %>% 
-  filter(area == "WMT") %>% 
-  filter(winter_id == 2) 
-summary(kids_pwq_education_summary)
+  select(area, winter_id, mmuse, mmhelpfulness,
+         thuse, thhelpfulness, fsuse, fshelpfulness) %>% 
+  group_by(winter_id) %>% 
+  mutate(winter_n = n()) %>% 
+  group_by(winter_id, winter_n, fshelpfulness) %>% 
+  count() %>% 
+  mutate(var_perc = round(n/winter_n*100, digits = 1)) %>% 
+  ungroup() %>% 
+  select(winter_id, fshelpfulness, n, var_perc)
+kids_pwq_education_summary
 

### 2021-06-09 main_analysis_final.Rmd
PM analysis updates. Manuscript revisions from co-author comments.

@@
 author: "Ethan Walker"
-date: "Updated 24 March 2021"
+date: "Updated 9 June 2021"
 output: pdf_document
@@
 library(broom.mixed)
+library(car)
 library(zoo)
@@
 library(faraway)
-library(DHARMa)
+#library(DHARMa)
 ```
@@
                                      "Do not know" = "Do not know")) %>% 
-  filter(person_time_at_risk_winter > 0) %>% 
+  #filter(person_time_at_risk_winter > 0) %>% 
   #filter(area == "WMT") %>% 
@@
                           child_age + 
-                          person_time_at_risk + (1 | cohort:area),
+                          person_time_at_risk + (1 | home:cohort:area),
                           data = new_model_data, family = binomial, nAGQ = 10,
@@
 plot_results 
+Anova(itt_results_int)
 
@@
 
-int_function(analysis_data, "home_sqft_2level")
+int_function(analysis_data, "residents_smoke")
 n_table
@@
 # Intent-to-treat framework, PM2.5 outcome
+```{r}
+# select distinct homes within datasets
+winter_1_data_pm <- analysis_data_winter1 %>% 
+  distinct(area, home, .keep_all = T)
+
+winter_1_data_pm_noak <- winter_1_data_pm %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  filter(area != "AK")
+```
 
@@
 itt_pm_results <- lmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                       (1 | cohort:area), data = analysis_data_winter1)
+                       (1 | cohort:area), data = winter_1_data_pm)
 #write_rds(itt_pm_results, paste0(file_path, "Output/itt_pm_results.rds"))
@@
 itt_pm_results_int <<- lmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
-                              child_age + 
                        (1 | cohort:area), data = new_model_data)
@@
 plot_results 
+Anova(itt_pm_results_int)
 
@@
 
-int_function(analysis_data_winter1, "stove_grade_3level")
+int_function(winter_1_data_pm, "stove_grade_3level")
+
 n_table
@@
 
-obs <- analysis_data_winter1 %>% 
+obs <- winter_1_data_pm %>% 
   distinct(area, child_id_num, .keep_all = T) %>% 
@@
   filter(!is.na(treatment_assigned)) %>% 
-  filter(!is.na(home_sqft_2level)) %>%
-  group_by(home_sqft_2level, treatment_assigned) %>% 
+  filter(!is.na(stove_grade_3level)) %>%
+  group_by(stove_grade_3level, treatment_assigned) %>% 
   count()
@@
 
-#((exp(-.418)-1)*100)
+((exp(-0.07214)-1)*100)
 ```
@@
 
-itt_pm_noak_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + child_age + 
-                       (1 | cohort:area), data = analysis_data_noak)
+itt_pm_noak_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                       (1 | cohort:area), data = winter_1_data_pm_noak)
+summary(itt_pm_noak_results)
 write_rds(itt_pm_noak_results, paste0(file_path, "Output/itt_pm_noak_results.rds"))
@@
 itt_pm_at_home_results <- glmer(log(pm_mean_at_home_sampling_period) ~ treatment_assigned + 
-                               child_age + (1 | cohort:area), 
-                               data = analysis_data_winter1)
+                               (1 | cohort:area), 
+                               data = winter_1_data_pm)
 write_rds(itt_pm_at_home_results, paste0(file_path, "Output/itt_pm_at_home_results.rds"))
@@
 itt_pm_income_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                               child_age + (1 | cohort:area) + 
-                               income_3level, data = analysis_data_winter1)
+                               (1 | cohort:area) + 
+                               income_3level, data = winter_1_data_pm)
 write_rds(itt_pm_income_results, paste0(file_path, "Output/itt_pm_income_results.rds"))
 
+itt_pm_sqft_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
+                               (1 | cohort:area) + 
+                               home_sqft_2level, data = winter_1_data_pm)
+write_rds(itt_pm_sqft_results, paste0(file_path, "Output/itt_pm_sqft_results.rds"))
+
 itt_pm_filter_type_results <- glmer(log(pm_mean_sampling_period) ~ treatment_filter_type +
-                                  child_age + (1 | cohort:area), data = analysis_data_winter1)
+                                  (1 | cohort:area), data = winter_1_data_pm)
 summary(itt_pm_filter_type_results)
@@
 itt_pm_filter_comp_results <- glmer(log(pm_mean_sampling_period) ~ filter_compliance + 
-                                  child_age + (1 | cohort:area), 
-                                  data = analysis_data_winter1)
+                                  (1 | cohort:area), 
+                                  data = winter_1_data_pm)
 summary(itt_pm_filter_comp_results)
@@
 itt_pm_smoke_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                              child_age + (1 | cohort:area) + 
-                              residents_smoke, data = analysis_data_winter1)
+                              (1 | cohort:area) + 
+                              residents_smoke, data = winter_1_data_pm)
 write_rds(itt_pm_smoke_results, paste0(file_path, "Output/itt_pm_smoke_results.rds"))
@@
 itt_pm_burn_level_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   burn_level_3level, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   burn_level_3level, data = winter_1_data_pm)
 write_rds(itt_pm_burn_level_results, paste0(file_path, "Output/itt_pm_burn_level_results.rds"))
@@
 itt_pm_smoking_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   home_act_smoking_sum, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   home_act_smoking_sum, data = winter_1_data_pm)
 write_rds(itt_pm_smoking_results, paste0(file_path, "Output/itt_pm_smoking_results.rds"))
@@
 itt_pm_windows_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   home_act_windows_sum, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   home_act_windows_sum, data = winter_1_data_pm)
 write_rds(itt_pm_windows_results, paste0(file_path, "Output/itt_pm_windows_results.rds"))
@@
 itt_pm_door_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   home_act_door_sum, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   home_act_door_sum, data = winter_1_data_pm)
 write_rds(itt_pm_door_results, paste0(file_path, "Output/itt_pm_door_results.rds"))
@@
 itt_pm_sweep_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   home_act_sweep_sum, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   home_act_sweep_sum, data = winter_1_data_pm)
 write_rds(itt_pm_sweep_results, paste0(file_path, "Output/itt_pm_sweep_results.rds"))
@@
 itt_pm_ambpm_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   amb_pm_lrti, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   amb_pm_lrti, data = winter_1_data_pm)
 write_rds(itt_pm_ambpm_results, paste0(file_path, "Output/itt_pm_ambpm_results.rds"))
@@
 itt_pm_ambtemp_results <- glmer(log(pm_mean_sampling_period) ~ treatment_assigned + 
-                                   child_age + (1 | cohort:area) + 
-                                   amb_temp_lrti, data = analysis_data_winter1)
+                                   (1 | cohort:area) + 
+                                   amb_temp_lrti, data = winter_1_data_pm)
 write_rds(itt_pm_ambtemp_results, paste0(file_path, "Output/itt_pm_ambtemp_results.rds"))
@@
 itt_pm_income_results <- read_rds(paste0(file_path, "Output/itt_pm_income_results.rds"))
+itt_pm_sqft_results <- read_rds(paste0(file_path, "Output/itt_pm_sqft_results.rds"))
 itt_pm_filter_type_results <- read_rds(paste0(file_path, "Output/itt_pm_filter_type_results.rds"))
@@
 itt_pm_income_tidy <- results_function(itt_pm_income_results, "ITT + income")
+itt_pm_sqft_tidy <- results_function(itt_pm_sqft_results, "ITT + home size")
 #itt_pm_filter_type_tidy <- results_function(itt_pm_filter_type_results, "ITT filter type")
@@
 # Combine cleaned results
-itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_at_home_tidy, itt_pm_noak_tidy, itt_pm_income_tidy, 
+itt_pm_results_combined <- rbind(itt_pm_tidy, itt_pm_at_home_tidy, itt_pm_noak_tidy, 
+                                 itt_pm_income_tidy, itt_pm_sqft_tidy, 
                               itt_pm_smoke_tidy, itt_pm_burn_level_tidy, itt_pm_smoking_tidy,
@@
   mutate(model = factor(model,
-                        levels = c("ITT Primary", "ITT PM at-home", "ITT no AK", "ITT + income", 
+                        levels = c("ITT Primary", "ITT PM at-home", "ITT no AK", 
+                                   "ITT + income", "ITT + home size",
                                    "ITT + smoke hx", "ITT + burn level", "ITT + smoke act",
@@
 er_results_int <<- glmer(lrti_events_di_total ~ pm_mean_sampling_period_iqr*int_var + 
-                          #child_age + 
+                          child_age + 
                           gender + 
-                          residents_smoke + 
+                          #residents_smoke + 
                           race_parent + gender_parent +
@@
 plot_results 
+Anova(er_results_int)
 
@@
 
-int_function(analysis_data, "child_age")
+int_function(analysis_data, "residents_smoke")
 save_contrasts
@@
   theme_bw() +  
-  ggtitle(label = "Results for ER framework, LRTI outcome") +
-  labs(y = "LRTI events per IQR \n increase in PM2.5") +
+  #ggtitle(label = "Results for ER framework, LRTI outcome") +
+  labs(y = "Odds of LRTI per IQR \n increase in PM2.5") +
   labs(x = "", group = "", shape = "") +

### 2021-06-09 main_analysis_initial.Rmd
PM analysis updates. Manuscript revisions from co-author comments.

--- 2020-12-24 main_analysis_initial.Rmd
@@
   distinct(area, child_id_num, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, child_age, lrti_events_total, 
+  dplyr::select(area, home, cohort, child_id_char, child_id_num, child_age, race_child, hispanic_child, lrti_events_total, 
          person_time_at_risk_total, lrti_incidence_total, newtx, treatment_assigned,
@@
   distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
-  dplyr::select(area, home, cohort, child_id_num, home_winter_id, winter_id, child_age, 
+  dplyr::select(area, home, cohort, child_id_char, child_id_num, home_winter_id, winter_id, child_age, race_child, hispanic_child,
          amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti, gender,
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+analysis_data <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_1obs_per_child.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+
 # Summary analysis for categorical variables
 
-# clean up data for table 1 summary
-table_1_data <- health_exposure_data_sampling_day %>% 
-  mutate(treatment_assigned = factor(treatment_assigned,
-                                     levels = c("Placebo", "Filter", "Education"))) %>% 
-    #select(-treatment_assigned, -treatment_actual) %>% 
-  #left_join(treatments_masked, by = "home") %>%
-  #rename(treatment = newtx) %>% 
-  arrange(area, home, winter_id, sampling_day) %>% 
-  filter(winter_id == 1) %>% 
-  filter(area == "AK") %>% # Change area filter as necessary
-  distinct(area, home, .keep_all = TRUE)
-
 # summary data for manuscript tables
-char_funct_total <- function(var, data = table_1_data){
+char_funct_total <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  filter(winter_id == 1) %>% 
   rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   mutate(total_n = n()) %>% 
@@
 
-char_funct_treatment <- function(var, data = table_1_data){
+char_funct_treatment <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  filter(winter_id == 1) %>% 
   rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   group_by(treatment_assigned) %>% 
@@
 char_funct_treatment("home_act_door_sum")
+
+# Summary analysis for categorical variables - by distinct child
+
+# summary data for manuscript tables
+char_funct_total <- function(var, data = health_exposure_data_sampling_day){
+
+demographics_summary <- data %>% 
+  mutate(child_age = fct_collapse(age_child_nov,
+                                  "< 1" = "Less than 1",
+                                  "1+" = c(1, 2, 3, 4))) %>% 
+  filter(winter_id == 1) %>% 
+  filter(!is.na(child_id_char)) %>% 
+  rename(new_var = var) %>% 
+  arrange(area, child_id_char, new_var) %>% 
+  distinct(area, child_id_char, .keep_all = T) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = round(n()/total_n*100)) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
+demographics_summary
+}
+
+char_funct_treatment <- function(var, data = health_exposure_data_sampling_day){
+
+demographics_summary <- data %>% 
+  mutate(child_age = fct_collapse(age_child_nov,
+                                  "< 1" = "Less than 1",
+                                  "1+" = c(1, 2, 3, 4))) %>% 
+  filter(winter_id == 1) %>% 
+  rename(new_var = var) %>% 
+  arrange(area, child_id_char, new_var) %>% 
+  distinct(area, child_id_char, .keep_all = T) %>% 
+  group_by(treatment_assigned) %>% 
+  mutate(treatment_n = n()) %>% 
+  group_by(treatment_assigned, new_var) %>% 
+    mutate(n = n(),
+         percent = round(n()/treatment_n*100)) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(treatment_assigned, new_var)
+demographics_summary
+}
+
+char_funct_total("gender")
+char_funct_treatment("gender")
+char_funct_total("child_age")
+char_funct_treatment("child_age")
+char_funct_total("race_child")
+char_funct_treatment("race_child")
+char_funct_total("hispanic_child")
+char_funct_treatment("hispanic_child")
 ```
@@
 
-# clean up data for table 1 summary
-table_1_data <- health_exposure_data_sampling_day %>% 
-  mutate(treatment_assigned = factor(treatment_assigned,
-                                     levels = c("Placebo", "Filter", "Education"))) %>% 
-  #select(-treatment_assigned, -treatment_actual) %>% 
-  #left_join(treatments_masked, by = "home") %>%
-  #rename(treatment = newtx) %>% 
-  arrange(area, home, winter_id, sampling_day) %>% 
-  filter(winter_id == 1) %>% 
-  #filter(area == "AK") %>% # Change area filter as necessary
-  distinct(area, home, .keep_all = TRUE)
-
 # summary data for manuscript tables
-num_funct_total <- function(var, data = table_1_data){
+num_funct_total <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  filter(winter_id == 1) %>% 
   rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
@@
   filter(!is.na(new_var)) %>% 
-  select(treatment_assigned, home, new_var) %>% 
+  dplyr::select(treatment_assigned, home, new_var) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
@@
 
-num_funct_treatment <- function(var, data = table_1_data){
+num_funct_treatment <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  filter(winter_id == 1) %>% 
   rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
@@
   filter(!is.na(new_var)) %>% 
-  select(treatment_assigned, home, new_var) %>% 
+  dplyr::select(treatment_assigned, home, new_var) %>% 
   group_by(treatment_assigned) %>% 
@@
 incidence_summary <- analysis_data %>% 
-  filter(area == "NN") %>% # Change area filter as necessary  
+  #filter(!is.na(child_id_char)) %>% 
+  filter(area == "WMT") %>% # Change area filter as necessary  
   #group_by(treatment) %>% # Change treatment grouping as necessary  
   #filter(lrti_incidence < 0.7) %>% 
-  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
-            sd_inc = sd(lrti_incidence, na.rm = TRUE),
+  summarize(mean_inc = mean(lrti_incidence_total, na.rm = TRUE),
+            sd_inc = sd(lrti_incidence_total, na.rm = TRUE),
             n_inc = n(),
-            min_inc = min(lrti_incidence, na.rm = TRUE),
-            median_inc = median(lrti_incidence, na.rm = TRUE),
-            max_inc = max(lrti_incidence, na.rm = TRUE))
+            min_inc = min(lrti_incidence_total, na.rm = TRUE),
+            median_inc = median(lrti_incidence_total, na.rm = TRUE),
+            max_inc = max(lrti_incidence_total, na.rm = TRUE))
 incidence_summary

### 2021-06-09 manuscript_tables_figures.Rmd
PM analysis updates. Manuscript revisions from co-author comments.

@@
   filter(winter_id != 2)
+kids_demographics
 ```
@@
   filter(original == 1)
+
+kids_ids
 ```
@@
   distinct(home, winter_id, .keep_all = TRUE) %>%
-  count(area, winter_id, treatment_assigned)
+  count(winter_id, treatment_assigned)
+study_homes
 
 study_kids <- health_exposure_data_sampling_day %>% 
-  select(area, home, child_id_char, child_id_num, winter_id, treatment_assigned) %>% 
+  dplyr::select(area, home, child_id_char, child_id_num, winter_id, treatment_assigned) %>% 
   filter(!is.na(child_id_char)) %>% 
-  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
-  count(area, winter_id, treatment_assigned)
+  distinct(area, child_id_char, winter_id, .keep_all = TRUE) %>% 
+  count(winter_id, treatment_assigned)
+study_kids
 ```
@@
 ```{r}
+demographics_summary <- health_exposure_data_sampling_day %>% 
+  filter(winter_id == 1 | winter_id == 0) %>% 
+  arrange(winter_id) %>% 
+  #filter(treatment_assigned == "Placebo") %>%
+  distinct(child_id_char, .keep_all = TRUE)
+
 # summary data for manuscript tables
@@
   rename(new_var = var) %>% 
-  filter(winter_id == 1) %>% 
+  filter(winter_id != 2) %>% 
   #filter(treatment_assigned == "Placebo") %>%
-  arrange(new_var) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
+  arrange(area, child_id_char, new_var) %>% 
+  distinct(area, child_id_char, .keep_all = TRUE) %>% # change to child_id_char or home as necessary
   mutate(total_n = n()) %>% 
@@
   rename(new_var = var) %>% 
-  filter(winter_id == 1) %>% 
+  filter(winter_id != 2) %>% 
   #filter(treatment_assigned == "Placebo") %>% 
-  arrange(new_var) %>% 
-  distinct(home, .keep_all = TRUE) %>% 
+  arrange(area, child_id_char, new_var) %>% 
+  distinct(area, child_id_char, .keep_all = TRUE) %>% # change to child_id_char or home as necessary
   mutate(total_n = n()) %>% 
@@
 
+# distinct by child - change var in functions above to child_id_char
+char_funct_total("gender")
+char_funct_area("gender")
+char_funct_total("hispanic_parent")
+char_funct_area("hispanic_parent")
+char_funct_total("race_child")
+char_funct_area("race_child")
+
+# distinct by home - change var in functions above to home
 char_funct_total("gender_parent")
 char_funct_area("gender_parent")
-char_funct_total("gender")
-char_funct_area("gender")
 char_funct_total("hispanic_parent")
@@
 ```
+
+# Check ambient temp within groups of self-reported relative stove use
+```{r}
+analysis_data <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_2obs_per_child.rds"))
+
+summary_results <- analysis_data %>% 
+  select(home, area, winter_id, burn_level_3level, amb_temp_lrti) %>% 
+  group_by(burn_level_3level) %>% 
+  summarize(mean(amb_temp_lrti, na.rm = T),
+            median(amb_temp_lrti, na.rm = T))
+
+summary_results
+```

### 2021-09-07 main_analysis_final.Rmd
Annotate/organize code for main outcomes paper

@@
 author: "Ethan Walker"
-date: "Updated 9 June 2021"
+date: "Updated 7 September 2021"
 output: pdf_document
@@
 
+# Primary analysis dataset - load and format a few variables
 analysis_data <- read_rds(paste0(file_path, 
@@
                                      "Do not know" = "Do not know"))
+  #filter(!is.na(home_sqft_2level))
   #filter(area == "WMT")
@@
 
+# Dataset for sensitivity analysis - replace "outlier" for person time with mean value
 analysis_data_outliers <- analysis_data %>% 
@@
 
+# Dataset for sensitivity analysis - filter for Winter 1 data only
 analysis_data_winter1 <- read_rds(paste0(file_path, 
@@
 
+# Dataset for sensitivity analysis - filter for Winter 2 data only
 analysis_data_winter2 <- read_rds(paste0(file_path, 
@@
 
+# Dataset for sensitivity analysis - filter out AK data
 analysis_data_noak <- read_rds(paste0(file_path, 
@@
 
+# Run function with selected interaction term
+# Use next few lines to help fill in Tables for publication
 int_function(analysis_data, "residents_smoke")
+
 n_table
@@
 ```{r, eval=FALSE, include=FALSE} 
-# Run and save model results
+# Run and save model results for various sensitivity analyses
 
@@
 
+# Run function with selected interaction term
+# Use next few lines to help fill in Tables for publication
 int_function(winter_1_data_pm, "stove_grade_3level")
@@
 
-((exp(-0.07214)-1)*100)
+((exp(-0.07214)-1)*100) # formula for exponentiating results (percent difference)
 ```
@@
 summary(er_results)
+#summary(er_results2)
 er_res <- simulateResiduals(er_results)
@@
 
+# Run function with selected interaction term
+# Use next few lines to help fill in Tables for publication
 int_function(analysis_data, "residents_smoke")
@@
 ```{r, eval=FALSE, include=FALSE}
-# Run and save model results
+# Run and save model results for various sensitivity analyses
 
@@
 er_stove_age_results <- read_rds(paste0(file_path, "Output/er_stove_age_results.rds"))
+
+nobs(er_stove_age_results)
 
@@
 ```{r, eval=FALSE, include=FALSE}
-################ Plot/check individual models and various results ##############
+################ Code to plot/check individual models and various results ##############
 

### 2021-09-07 manuscript_tables_figures.Rmd
Annotate/organize code for main outcomes paper

@@
 author: "Ethan Walker"
-date: "Started 13 May 2020, Updated 30 June 2020"
+date: "Started 13 May 2020, Updated 7 September 2021"
 output:
@@
 
+# Character/factor data summary stats
 ```{r}
@@
 
+# Numeric data summary stats
 ```{r}
@@
 
+# Summarize home activity variables
 ```{r}

### 2021-11-19 child_health_data_summary.Rmd
Minor analysis updates for main paper revisions

--- 2020-05-14 child_health_data_summary.Rmd
@@
 ```
+
+\pagebreak
+
+```{r}
+kids_at_home_log <- read_rds(paste0(file_path, "Output/kids_at_home_log.rds"))
+
+at_home_summary <- kids_at_home_log %>% 
+  mutate(treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education"))) %>% 
+  filter(!is.na(at_home)) %>% 
+  arrange(area, home, child_id_num, sampling_day, sample_time) %>% 
+  group_by(area, home, child_id_num) %>% 
+  mutate(total_n = n(),
+         time_home = sum(at_home, na.rm = TRUE),
+         at_home_percent = time_home/total_n*100) %>% 
+  ungroup() %>% 
+  select(area, home, child_id_num, sampling_day, sample_time, treatment_assigned,
+         total_n, time_home, at_home, at_home_percent) %>% 
+  filter(!is.na(treatment_assigned)) %>% 
+  distinct(area, home, child_id_num, .keep_all = T) %>% 
+  #group_by(treatment_assigned) %>% 
+  summarize(mean_at_home = mean(at_home_percent, na.rm = TRUE),
+            sd_at_home = sd(at_home_percent, na.rm = TRUE),
+            min_at_home = min(at_home_percent, na.rm = TRUE),
+            p25 = quantile(at_home_percent, probs = 0.25),
+            median_at_home = median(at_home_percent, na.rm = TRUE),
+            p75 = quantile(at_home_percent, probs = 0.75),
+            max_at_home = max(at_home_percent, na.rm = TRUE))
+at_home_summary
+
+at_home_n <- kids_at_home_log %>% 
+  mutate(treatment_assigned = factor(treatment_assigned,
+                                     levels = c("Placebo", "Filter", "Education"))) %>% 
+  filter(!is.na(treatment_assigned)) %>% 
+  filter(!is.na(at_home)) %>% 
+  distinct(area, home, child_id_num, .keep_all = TRUE) %>% 
+  count(treatment_assigned)
+at_home_n
+```

### 2021-11-19 main_analysis_final.Rmd
Minor analysis updates for main paper revisions

@@
 author: "Ethan Walker"
-date: "Updated 7 September 2021"
+date: "Updated 19 Oct 2021"
 output: pdf_document
@@
                      control = glmerControl(optimizer = "bobyqa"))
+
 #write_rds(itt_results, paste0(file_path, "Output/itt_results.rds"))
@@
 # Use next few lines to help fill in Tables for publication
-int_function(analysis_data, "residents_smoke")
+int_function(analysis_data, "flu_shot")
 
@@
 
-# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke
+# Try child_age, gender, home_floors_2level, home_sqft_2level, residents_smoke, flu_shot
 ```
@@
 
+itt_fluvax_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + person_time_at_risk + flu_shot + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 10)
+write_rds(itt_fluvax_results, paste0(file_path, "Output/itt_fluvax_results.rds"))
+
 itt_home_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
@@
 write_rds(itt_home_results, paste0(file_path, "Output/itt_home_results.rds"))
+
+itt_area_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + area + person_time_at_risk + 
+                     (1 | home:cohort),
+                     data = analysis_data, family = binomial, nAGQ = 10)
+write_rds(itt_area_results, paste0(file_path, "Output/itt_area_results.rds"))
+
+itt_offset_results <- glmer(lrti_events_di_total ~ treatment_assigned + 
+                     child_age + 
+                     (1 | home:cohort:area),
+                     data = analysis_data, family = binomial, nAGQ = 10)
+write_rds(itt_offset_results, paste0(file_path, "Output/itt_offset_results.rds"))
 
@@
 itt_ambtemp_results <- read_rds(paste0(file_path, "Output/itt_ambtemp_results.rds"))
-#itt_home_results <- read_rds(paste0(file_path, "Output/itt_home_results.rds"))
+itt_fluvax_results <- read_rds(paste0(file_path, "Output/itt_fluvax_results.rds"))
+itt_home_results <- read_rds(paste0(file_path, "Output/itt_home_results.rds"))
+itt_area_results <- read_rds(paste0(file_path, "Output/itt_area_results.rds"))
+itt_offset_results <- read_rds(paste0(file_path, "Output/itt_offset_results.rds"))
 itt_outlier_results <- read_rds(paste0(file_path, "Output/itt_outlier_results.rds"))
@@
 itt_ambtemp_tidy <- results_function(itt_ambtemp_results, "ITT + amb temp")
-#itt_home_tidy <- results_function(itt_home_results, "ITT - home")
+itt_fluvax_tidy <- results_function(itt_fluvax_results, "ITT + flu vax")
+itt_home_tidy <- results_function(itt_home_results, "ITT no Home")
+itt_area_tidy <- results_function(itt_area_results, "ITT area fixed")
+itt_offset_tidy <- results_function(itt_offset_results, "ITT no offset")
 itt_outlier_tidy <- results_function(itt_outlier_results, "ITT outlier")
@@
                               itt_windows_tidy, itt_door_tidy, itt_sweep_tidy,
-                              itt_ambpm_tidy, itt_ambtemp_tidy)
+                              itt_ambpm_tidy, itt_ambtemp_tidy, itt_fluvax_tidy, 
+                              itt_area_tidy, itt_offset_tidy, itt_home_tidy)
 
@@
                                    "ITT + window act", "ITT + door act", "ITT + sweep act",
-                                   "ITT + amb PM", "ITT + amb temp"))) %>% 
+                                   "ITT + amb PM", "ITT + amb temp", "ITT + flu vax", 
+                                   "ITT area fixed", "ITT no offset", "ITT no Home"))) %>% 
   ggplot(aes(group = term, shape = term)) +
@@
 
-ITT Primary = Logistic regression mixed model, as specified above  
+ITT Primary = Logistic regression mixed model, as specified in main results  
 ITT Winter 1 = Primary model with Winter 1 data only  
-ITT no AK = Primary model without homes from AK (different placebo group)  
+ITT no AK = Primary model without homes from Alaska (different placebo group)  
 ITT + income = Primary model plus categorical income variable  
@@
 ITT + amb temp = Primary model plus amb temp during LRTI at-risk period  
-ITT outlier = Primary model with data that fixes incidence outlier
+ITT + flu vax = Primary model plus caregiver-reported influenza vaccination status of child  
+ITT area fixed = Primary model with area as a fixed term instead of a random term
+ITT no offset = Primary model without fixed offset term for person-time of follow-up
+ITT no Home = Primary model without the random nested term for household/home
 
@@
                           income_3level + education_3level + 
-                          amb_temp_lrti + amb_pm_lrti + residents_smoke +
+                          amb_temp_lrti + 
+                          #amb_pm_lrti + 
+                          residents_smoke +
                           person_time_at_risk +
@@
 
-lrti_summary_table <- analysis_data_winter1 %>% 
-  filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
-           !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
+lrti_summary_table <- analysis_data_winter2 %>% 
+  #filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
+   #        !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
   #group_by(treatment_assigned) %>% 
@@
          lrti_events_num = lrti_events_num - 1) %>%  
+  filter(area == "AK") %>% 
+  group_by(treatment_assigned) %>% 
   summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
-            "Group %" = `Total LRTI`/n()*100)
+            "Group %" = `Total LRTI`/n()*100,
+            n())
 lrti_summary_table
@@
 lrti_summary_table
+
+person_time_summary <- analysis_data %>% 
+  group_by(treatment_assigned) %>% 
+  summarize(mean_var = mean(person_time_at_risk), 
+            sd_var = sd(person_time_at_risk), 
+            n_var = n(), 
+            min_var = min(person_time_at_risk), 
+            p25 = quantile(person_time_at_risk, probs = 0.25),
+            med_var = median(person_time_at_risk), 
+            p75 = quantile(person_time_at_risk, probs = 0.75), 
+            max_var = max(person_time_at_risk))
+person_time_summary
 ```

### 2021-11-19 main_analysis_initial.Rmd
Minor analysis updates for main paper revisions

--- 2020-12-24 main_analysis_initial.Rmd
@@
          home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum,
-         filter_type, filter_compliance) %>% 
+         filter_type, filter_compliance, flu_shot, flu_shot_planned) %>% 
   rename(treatment = newtx,
@@
          home_act_smoking_sum, home_act_windows_sum, home_act_sweep_sum, home_act_door_sum,
-         filter_type, filter_compliance) %>% 
+         filter_type, filter_compliance, flu_shot, flu_shot_planned) %>% 
   rename(treatment = newtx) %>% 
@@
 ```{r}
+file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
+
+analysis_data <- read_rds(paste0(file_path, 
+                                 "Output/kids_analysis_data_1obs_per_child.rds"))
+
+health_exposure_data_sampling_day <- 
+  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
+
 # Summary analysis for categorical variables
 
-# clean up data for table 1 summary
-table_1_data <- health_exposure_data_sampling_day %>% 
-  mutate(treatment_assigned = factor(treatment_assigned,
-                                     levels = c("Placebo", "Filter", "Education"))) %>% 
-    #select(-treatment_assigned, -treatment_actual) %>% 
-  #left_join(treatments_masked, by = "home") %>%
-  #rename(treatment = newtx) %>% 
-  arrange(area, home, winter_id, sampling_day) %>% 
-  filter(winter_id == 1) %>% 
-  filter(area == "AK") %>% # Change area filter as necessary
-  distinct(area, home, .keep_all = TRUE)
-
 # summary data for manuscript tables
-char_funct_total <- function(var, data = table_1_data){
+char_funct_total <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  filter(winter_id == 1) %>% 
+  filter(area == "WMT") %>% 
   rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   mutate(total_n = n()) %>% 
@@
 
-char_funct_treatment <- function(var, data = table_1_data){
+char_funct_treatment <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  filter(winter_id == 1) %>% 
+  filter(area == "WMT") %>% 
   rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   group_by(treatment_assigned) %>% 
@@
 ```{r}
-# Summary analysis for numeric variables
-
-# clean up data for table 1 summary
-table_1_data <- health_exposure_data_sampling_day %>% 
-  mutate(treatment_assigned = factor(treatment_assigned,
-                                     levels = c("Placebo", "Filter", "Education"))) %>% 
-  #select(-treatment_assigned, -treatment_actual) %>% 
-  #left_join(treatments_masked, by = "home") %>%
-  #rename(treatment = newtx) %>% 
-  arrange(area, home, winter_id, sampling_day) %>% 
-  filter(winter_id == 1) %>% 
-  #filter(area == "AK") %>% # Change area filter as necessary
-  distinct(area, home, .keep_all = TRUE)
+# Summary analysis for categorical variables - by distinct child
 
 # summary data for manuscript tables
-num_funct_total <- function(var, data = table_1_data){
+char_funct_total <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  mutate(child_age = fct_collapse(age_child_nov,
+                                  "< 1" = "Less than 1",
+                                  "1+" = c(1, 2, 3, 4))) %>% 
+  filter(area == "WMT") %>% 
+  filter(winter_id == 1) %>% 
+  filter(!is.na(child_id_char)) %>% 
   rename(new_var = var) %>% 
-  mutate(new_var = as.numeric(new_var)) %>% 
-  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
-  filter(!is.na(new_var)) %>% 
-  select(treatment_assigned, home, new_var) %>% 
-  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
-            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+  arrange(area, child_id_char, new_var) %>% 
+  distinct(area, child_id_char, .keep_all = T) %>% 
+  mutate(total_n = n()) %>% 
+  group_by(new_var) %>% 
+  mutate(n = n(),
+         percent = round(n()/total_n*100)) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(new_var)
 demographics_summary
@@
 
-num_funct_treatment <- function(var, data = table_1_data){
+char_funct_treatment <- function(var, data = health_exposure_data_sampling_day){
 
 demographics_summary <- data %>% 
+  mutate(child_age = fct_collapse(age_child_nov,
+                                  "< 1" = "Less than 1",
+                                  "1+" = c(1, 2, 3, 4))) %>% 
+  filter(area == "WMT") %>% 
+  filter(winter_id == 1) %>% 
+  filter(!is.na(child_id_char)) %>% 
   rename(new_var = var) %>% 
+  arrange(area, child_id_char, new_var) %>% 
+  distinct(area, child_id_char, .keep_all = T) %>% 
+  group_by(treatment_assigned) %>% 
+  mutate(treatment_n = n()) %>% 
+  group_by(treatment_assigned, new_var) %>% 
+    mutate(n = n(),
+         percent = round(n()/treatment_n*100)) %>% 
+  distinct(new_var, .keep_all = TRUE) %>% 
+  dplyr::select(new_var, n, percent) %>% 
+  arrange(treatment_assigned, new_var)
+demographics_summary
+}
+
+char_funct_total("gender")
+char_funct_treatment("gender")
+char_funct_total("child_age")
+char_funct_treatment("child_age")
+char_funct_total("race_child")
+char_funct_treatment("race_child")
+char_funct_total("hispanic_child")
+char_funct_treatment("hispanic_child")
+```
+
+```{r}
+# Summary analysis for numeric variables
+
+# summary data for manuscript tables
+num_funct_total <- function(var, data = health_exposure_data_sampling_day){
+
+demographics_summary <- data %>% 
+  filter(area == "WMT") %>% 
+  filter(winter_id == 1) %>% 
+  rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
   mutate(new_var = as.numeric(new_var)) %>% 
-  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
+  mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
   filter(!is.na(new_var)) %>% 
-  select(treatment_assigned, home, new_var) %>% 
+  dplyr::select(treatment_assigned, home, new_var) %>% 
+  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
+            min_var = min(new_var), p25 = quantile(new_var, probs = 0.25), 
+            med_var = median(new_var), p75 = quantile(new_var, probs = 0.75), 
+            max_var = max(new_var))
+demographics_summary
+}
+
+num_funct_treatment <- function(var, data = health_exposure_data_sampling_day){
+
+demographics_summary <- data %>% 
+  filter(area == "WMT") %>% 
+  filter(winter_id == 1) %>% 
+  rename(new_var = var) %>% 
+  arrange(area, home, new_var) %>% 
+  distinct(area, home, .keep_all = T) %>% 
+  mutate(new_var = as.numeric(new_var)) %>% 
+  mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
+  filter(!is.na(new_var)) %>% 
+  dplyr::select(treatment_assigned, home, new_var) %>% 
   group_by(treatment_assigned) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
-            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+            min_var = min(new_var), p25 = quantile(new_var, probs = 0.25), 
+            med_var = median(new_var), p75 = quantile(new_var, probs = 0.75), 
+            max_var = max(new_var))
 demographics_summary
@@
 at_home_pm_summary <- health_exposure_data_sampling_day %>% 
-  select(-treatment_assigned, -treatment_actual) %>% 
-  left_join(treatments_masked, by = "home") %>%
-  rename(treatment = newtx) %>% 
+  #select(-treatment_assigned, -treatment_actual) %>% 
+  #left_join(treatments_masked, by = "home") %>%
+  rename(treatment = treatment_assigned) %>% 
   arrange(area, home, winter_id, sampling_day) %>% 
@@
   filter(!is.na(new_var)) %>%
-  filter(area == "NN") %>% # Change area filter as necessary 
+  #filter(area == "NN") %>% # Change area filter as necessary 
   group_by(treatment) %>% # Change treatment grouping as necessary 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
-            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+            min_var = min(new_var), p25 = quantile(new_var, probs = 0.25), 
+            med_var = median(new_var), p75 = quantile(new_var, probs = 0.75), 
+            max_var = max(new_var))
 at_home_pm_summary

### 2021-11-19 manuscript_tables_figures.Rmd
Minor analysis updates for main paper revisions

@@
   mutate(home_sqft_2level = cut(home_sqft, breaks = c(0, 1400, 100000),
-                                labels = c("<130 sq m", "130+ sq m")))
+                                labels = c("<130 sq m", "130+ sq m")),
+         home_bedrooms = as.factor(home_bedrooms),
+         home_bedrooms_cat = fct_collapse(home_bedrooms,
+                                          "0" = "0",
+                                          "1" = "1",
+                                          "2" = "2",
+                                          "3+" = c("3", "4", "5", "6", "11")))
 
@@
   #filter(treatment_assigned == "Placebo") %>%
-  distinct(child_id_char, .keep_all = TRUE)
+  distinct(area, child_id_char, .keep_all = TRUE)
 
@@
 demographics_summary <- data %>% 
-  rename(new_var = var) %>% 
+  rename(new_var = as.factor(var)) %>% 
   filter(winter_id != 2) %>% 
+  filter(area == "AK") %>% 
   #filter(treatment_assigned == "Placebo") %>%
@@
   distinct(area, child_id_char, .keep_all = TRUE) %>% # change to child_id_char or home as necessary
+  filter(!is.na(child_id_char)) %>% 
   mutate(total_n = n()) %>% 
@@
   distinct(new_var, .keep_all = TRUE) %>% 
-  dplyr::select(new_var, n, percent)
+  dplyr::select(new_var, n, percent, total_n)
 demographics_summary
@@
 demographics_summary <- data %>% 
-  rename(new_var = var) %>% 
+  rename(new_var = as.factor(var)) %>% 
   filter(winter_id != 2) %>% 
+  filter(area == "AK") %>% 
   #filter(treatment_assigned == "Placebo") %>% 
@@
   distinct(new_var, .keep_all = TRUE) %>% 
-  dplyr::select(new_var, n, percent) %>% 
+  dplyr::select(new_var, n, percent, area_n) %>% 
   arrange(treatment_assigned, new_var)
@@
 char_funct_area("race_child")
+char_funct_total("flu_shot")
+char_funct_area("flu_shot")
+char_funct_total("flu_shot_planned")
+char_funct_area("flu_shot_planned")
 
@@
 char_funct_area("ruca_code_3")
+char_funct_total("residents_under_five")
+char_funct_area("residents_under_five")
+char_funct_total("home_bedrooms_cat")
+char_funct_area("home_bedrooms_cat")
 ```
@@
   filter(winter_id == 1) %>% 
-  filter(treatment_assigned == "Filter") %>% 
+  #filter(treatment_assigned == "Filter") %>% 
   arrange(new_var) %>% 
@@
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
-            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+            min_var = min(new_var), p25 = quantile(new_var, probs = 0.25),
+            med_var = median(new_var), 
+            p75 = quantile(new_var, probs = 0.75),max_var = max(new_var))
 demographics_summary
@@
   filter(!is.na(new_var)) %>% 
-  dplyr::select(area, home, new_var) %>% 
-  group_by(area) %>% 
+  dplyr::select(treatment_assigned, home, new_var) %>% 
+  group_by(treatment_assigned) %>% 
   summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
-            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
+            min_var = min(new_var), p25 = quantile(new_var, probs = 0.25),
+            med_var = median(new_var), 
+            p75 = quantile(new_var, probs = 0.75),max_var = max(new_var))
 demographics_summary
@@
 
-# Check ambient temp within groups of self-reported relative stove use
+# Check ambient and sums temp within groups of self-reported relative stove use
 ```{r}
@@
 summary_results
+
+summary_results <- health_exposure_data_sampling_day %>% 
+  select(home, area, winter_id, burn_level_3level, sums_mean_winter) %>% 
+  group_by(burn_level_3level) %>% 
+  summarize(mean(sums_mean_winter, na.rm = T),
+            median(sums_mean_winter, na.rm = T))
+
+summary_results
 ```

### 2021-11-19 stove_grade_usage_cleaning.Rmd
Minor analysis updates for main paper revisions

@@
   replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
-  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
-  filter(treatment_assigned == "Filter")
+  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) %>% 
+  filter(kw_perc_expected < 1000)
 
@@
   filter(!is.na(kw_perc_expected)) %>% 
-  #filter(area == "WMT") %>% 
-  group_by(treatment_assigned) %>% 
+  filter(winter_id == 1) %>% 
+  #group_by(treatment_assigned) %>% 
   summarize("Mean KW % Exp" = mean(kw_perc_expected),
